### 2018-12-10 Papers/SHADE/elsarticle-template/paperSHADE.tex
Started paper redaction

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{mathtools}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\modulolinenumbers[5]

\journal{Applied Soft Computing}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

\section{Constrained Clustering}

Clustering is the task of grouping instances of a dataset into subsets. The criterion used to assign an instance to a given cluster is the similarity with the rest of elements in that cluster and the dissimilarity with the rest of instances of the dataset. In this way, the clustering process obtains subsets of instances of the initial data set that present high intra-similarity and low inter-similarity.

In most clustering applications it is common to have some kind of information about the data set to be analyzed. In constrained clustering this information is given in the form of pairs of instances. A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster. Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms.

Before formalizing the definition of constraints we must give a notation for the data set. We will note our dataset as $X$, which is a set of $N$ instances with $D$ features. We well refer to the instances of $X$ as $x$. Now, given the definition of dataset, we can define two types of constraints:

\begin{itemize}

	\item Must-link constraints $c_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.

	\item Cannot-link constraints $c_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.

\end{itemize}

The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraints set. As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $N = |X| = \sum_{i = 1}^{K} |c_i| $.

Knowing how a constraint is defined, Must-link constraints---from now on ML---are an example of an equivalence relation; therefore, CL constraints are reflexive, transitive and symmetric. This way, given the constraints $c_=(x_a,x_b)$ and $c_=(x_b,x_c)$ then $c_=(x_a,x_c)$ is verified. In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $c_=(x_a,x_b)$, then $c_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.

On the other hand, Cannot-link restrictions---from now on CL---do not constitute an equivalence relation, but they can be entailed. To illustrate that, we consider the instances $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then is also true that $c_{\neq}(x_c,x_d)$ for all $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.

\textcolor{red}{Comentar la complejidad del problema?}

\section{Background}

\section{Brief Review of BRKGA Algorithm}

The biased random-key genetic algorithm---from now on BRKGA---was first proposed by Gonçalves and Resende \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.

In BRKGA each solution---also called chromosome---is represented as a vector of values in the interval $[0,1]$. However, this vector is not a solution to the problem, it is only a representation of it. This allows us to abstract the details of the problem to apply a genetic algorithm that operates with real coding. Therefore, a deterministic decoder is necessary to obtain the actual solution to the problem based on the chromosome that encodes it.

BRKGA also requires a fitness function to evaluate each chromosome. This function takes the decoded solution as input and provides its fitness value as output. It is designed specifically for each problem.

The BRKGA optimization process is completely independent of the problem. First, a population $P$ of $p_{size}$ vectors of random-keys is initialized, this will be the initial generation. Each random-key vector has $N$ random-keys in it. The population is sorted by the fitness value $f_i$ of each chromosome to select the first $p_e$ individuals, this is, the elite of the population. The elite will be preserved in the next generation without modification. To introduce diversity, a number $p_m$ of new random-keys are also included in the next generation. The remaining chromosome of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents \cite{de2017comparison}.

\subsection{Adaptation of BRKGA for Constrained Clustering}

Oliviera, Chaves and Lorena (2017) proposed an adaptation of BRKGA to apply it to constrained clustering \cite{de2017comparison}.

Their method randomly initializes a population of random-key vectors, each one with $N$ random-keys. The decoder divides the interval $[0,1]$ in $K$ intervals, so there exist a correspondence between each random-key and the integer corresponding to the interval in which it is. Table \ref{tab:table1} shows an example of random-key decodification for $K = 3$. Note that extreme values 0 and 1 can also appear in a random-key vector.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}

	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
		\hline
		Index & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\ 
		\hline
		Random-key & 0.12 & 0.37 & 0.66 & 0.56 & 0.00 & 0.97 & 0.23 & 0.25 & 0.15 & 1.00 \\
		\hline
		Clusters & 1 & 2 & 2 & 2 & 1 & 3 & 1 & 3 & 1 & 3 \\ 
		\hline

	\end{tabular}
	\caption{Random-key decodification example \cite{de2017comparison}}
	\label{tab:table1}
\end{table}

Once the decoded solution has been obtained, the fitness value of the solution is computed as in Equation \eqref{eq1}.

\begin{equation}
f_i = z_i + \overbracket{(\mu * N * {infeasibility}_i)}^{penalty}
\label{eq1}
\end{equation}

Where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.

\begin{equation}
z_i = \frac{\sum_{x_a, x_b \in c_i} d_D^2(x_a,x_b)}{|c_i|} \;\;\; \forall c_i \in C
\label{eq2}
\end{equation}

Where $x_a$ and $x_b$ are instances in the cluster $c_i$ such that $a \neq b$ and the distance between each pair of instances in $c_i$ is included in the sum only once.

Oliviera, Chaves and Lorena also added a new element to BRKGA, which is a local search. This local optimization method is applied to each decoded chromosome in each generation, but its results are not transferred to the chromosome, in order to maintain the diversity. So the results of the local search are only used to update the best solution found so far if needed.

\section{The SHADE Optimization Method}

Success history based adaptive differential evolution---de ahora en adelante SHADE---is an optimization method based on differential evolution. Proposed by Tanabe and Fukunaga (2013), it was an improvement for the JADE model, created by Zhang and Sanderson \cite{tanabe2013success}\cite{zhang2009jade}.

SHADE uses a historical record of successful parameters as a mechanism for the adaptation of parameters that are involved in the process of creating new generations. Unlike JADE, which, in order to create a new generation, only considers the parameters used in the previous generation.

\subsection{SHADE Operators}

As in differential evolution, SHADE uses muntant generation, crossover and replacement operators to explore the space of solutions and bring in new generations of individuals.

The mutation strategy used by SHADE is known as current-to-pbest/1, and the expression that generates new individuals is described in the Equation \eqref{eq3}.

\begin{equation}
m_{i,G} = p_{i,G} + F_i * (p_{pbest, G} - p_{i,G}) + F_i * (p_{r1, G} - p_{r2,G})
\label{eq3}
\end{equation}

Where $m_{i,G}$ is the mutant vector, which is generated based on an individual from the population $p_{i,G}$. The indices $r1$ and $r2$ are random values in the range $[0,N]$, different from $i$ and from each other. The individual $p_{pbest, G}$ is randomly selected from among the best $N \times p\;\;|\;\;p\in [0,1]$ individuals in the population. In this way the $p$ parameter controls the greediness of the mutation strategy. The parameter $F_i$ defines the magnitude of the operator.

After generating the mutant vector $m_{i,G}$, a crossover operator is applied to combine it with the parent $p_{i,G}$ vector, the result is the trial vector. SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.

\begin{equation}
t_{j,i,G} = \left\{ \begin{array}{lc}
m_{j,i,G} &   if \;\; rand[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
p_{j,i,G} &  otherwise
\end{array}
\right.
\label{eq4}
\end{equation}

Where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.

The SHADE replacement operator determines the survivors individuals for the next generation. It compares each parent $x_{i,G}$ with the trial vector $_{j,i,G}$, generated based on it. The best one of these two individuals will survive for the next generation. Equation \eqref{eq5} shows this concept.

\begin{equation}
p_{i,G + 1} = \left\{ \begin{array}{lc}
t_{i,G} &   if \;\; f(t_{i,G}) \le f(p_{i,G}) \\
p_{i,G} &  otherwise
\end{array}
\right.
\label{eq5}
\end{equation}

To maintain diversity, SHADE can make use of an external $A$ archive, in which parents $p_{i,G}$ who were replaced by their associated trial vector $t_{i,G}$ are saved. To make use of the archive, we will consider that the individual $p_{r2,G}$ from Equation \eqref{eq3} is selected from $P \cup A$, which is the union of the population and the file $A$. The file size is the same as the population size. When the size of $A$ exceeds the size $P$, a random individual is selected to remove it and make room for the new one.

The parameters $p$, $F_i$ y $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process. These are the parameters that SHADE adaptively optimizes using the mentioned history record.

\subsection{SHADE Parameter Adaptation Method}

The SHADE method saves in memory a structure with $H$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:table2}. This structure is initialized with the value $0.5$ for all entries.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{10pt}
	\renewcommand{\arraystretch}{0.9}

	\begin{tabular}{|>{\centering\arraybackslash}m{1cm}  |>{\centering\arraybackslash}m{1cm}|>{\centering\arraybackslash}m{1cm}|>{\centering\arraybackslash}m{1cm}|>{\centering\arraybackslash}m{1.4cm}|>{\centering\arraybackslash}m{1.2cm}|}
		\hline
		 Index & 1 & 2 & $\cdots$ & $H - 1$ & $H$ \\ 
		 \hline
		 \hline
		 $M_{CR}$ & $M_{CR,1}$ & $M_{CR,1}$ & $\cdots$ & $M_{CR,H-1}$ & $M_{CR,H}$ \\ 
		 \hline
		 $M_{F}$ & $M_{F,1}$ & $M_{F,1}$ & $\cdots$ & $M_{F,H-1}$ & $M_{F,H}$ \\ 
		\hline

	\end{tabular}
	\caption{The historical memory $M_{CR}$, $M_{F}$  \cite{tanabe2013success}}
	\label{tab:table2}
\end{table}

In each generation, parameters $CR_i$ and $F_i$ are calculated on the basis of the history by applying the Equations \eqref{eq6} and \eqref{eq7}:

\begin{equation}
CR_i = randn_i(M_{CR,r_i}, 0.1)
\label{eq6}
\end{equation}

\begin{equation}
F_i = randc_i(M_{F,r_i}, 0.1)
\label{eq7}
\end{equation}

Where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$. When a value for $CR_i$ outside of the range $[0,1]$ is generated, its replaced by the allowed value. If $F_i > 1$, then it is truncated to $1$, and if $F_i < 0$, then Equation \eqref{eq7} is applied as many times as needed to obtain a legal value.

SHADE uses two auxiliar sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent. At the end of each generation, the content of the memory $M$ is updated following Equations \eqref{eq8} and \eqref{eq9}.

\begin{equation}
M_{CR,h,G+1} = \left\{ \begin{array}{lc}
mean_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
M_{CR,h,G+1} &  otherwise
\end{array}
\right.
\label{eq8}
\end{equation}

\begin{equation}
M_{F,h,G+1} = \left\{ \begin{array}{lc}
mean_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
M_{F,h,G+1} &  otherwise
\end{array}
\right.
\label{eq9}
\end{equation}

The $h \;\; (0 \le h \le H)$ index specifies the position of the memory $M$ to be updated. This index is initialized to $0$ at the beginning of the optimization process and increased by one after each generation. When $h \ge H$ then $h$ is restarted to 1. It is worth noting that, when in a generation there is no trial vector better than the parent which generated it, no update of $M$ is done.

In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the wighted mean,  which is computed following Equations \eqref{eq10} and \eqref{eq11}, which were proposed by Peng et al (2009) \cite{peng2009multi} to prevent $CR$ to converge to small value.

\begin{equation}
mean_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{RC,i}
\label{eq10}
\end{equation}

\begin{equation}
\omega_i = \frac{\Delta f_i}{\sum_{i = 1}^{|S_{RC}|} \Delta f_i}
\label{eq11}
\end{equation}

where $\Delta f_i = |f(t_i,G) - f(p_i, G)|$, which is the amount of improvement obtained from the trial vector respect to to the parent.

In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12}, proposed by Tanabe and Fukunaga (2013) \cite{tanabe2013success}.

\begin{equation}
mean_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{F,i}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{F,i}}
\label{eq12}
\end{equation}

Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the greedyness of the mutation strategy, is not included in the optimization process. However it is not static, it is calculated for each individual of the population $p_i$ following Equation \eqref{eq13}.

\begin{equation}
pbest_i = randn[p_{min}, 0.2]
\label{eq13}
\end{equation}

where $p_{min}$ is set such that there is always al least two individuals to choose between, so $p_{min} = 2/N$.

\subsection{Adaptation of SHADE for constrained clustering}

\newpage

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2018-12-10 Software/BRKGAMain.py

@@ -13,9 +13,9 @@ def main():
 
     names_array, datasets_array, labels_array = load_all_datasets()
 
-    names_array = names_array[0:4]
-    datasets_array = datasets_array[0:4]
-    labels_array = labels_array[0:4]
+    names_array = names_array[0:5]
+    datasets_array = datasets_array[0:5]
+    labels_array = labels_array[0:5]
 
     const_percent_vector = [0.05, 0.1, 0.15, 0.2]
     #const_percent_vector = [0.1, 0.2]

### 2018-12-10 Software/BRKGAMainParallel.py

@@ -44,9 +44,9 @@ def main():
 
     names_array, datasets_array, labels_array = load_all_datasets()
 
-    names_array = names_array[:4]
-    datasets_array = datasets_array[:4]
-    labels_array = labels_array[:4]
+    names_array = names_array[:5]
+    datasets_array = datasets_array[:5]
+    labels_array = labels_array[:5]
 
     const_percent_vector = [0.05, 0.1, 0.15, 0.2]
     # const_percent_vector = [0.05]

### 2018-12-10 Software/SHADE/SHADE.py

@@ -4,7 +4,7 @@ import scipy.stats
 
 class SHADE:
 
-    def __init__(self, data, ml_const, cl_const, population_size, nb_clust, ls = False):
+    def __init__(self, data, ml_const, cl_const, population_size, nb_clust, prt_elite, ls = False):
 
         self._data = data
         self._ml = ml_const
@@ -16,6 +16,7 @@ class SHADE:
         self._h_record_cr = np.empty((0, 0))
         self._h_record_f = np.empty((0, 0))
         self._ls = ls
+        self._prt_elite = prt_elite
 
 
     def init_population(self):
@@ -149,12 +150,23 @@ class SHADE:
                 f_i = scipy.stats.cauchy.rvs(loc=self._h_record_f[r_i], scale=0.1)
                 p_i = np.random.uniform(2.0/self._population_size, 0.2)
 
+                if cr_i < 0: cr_i = 0
+                if cr_i > 1: cr_i = 1
+                while f_i <= 0: f_i = scipy.stats.cauchy.rvs(loc=self._h_record_f[r_i], scale=0.1)
+                if f_i > 1: f_i = 1
+
                 #Obtenemos los individuos que intervienen en la generacion
                 x_i = self._population[j]
                 x_pbest = self._population[np.random.randint(0, int(self._population_size * p_i))]
-                x_r1 = self._population[np.random.randint(0, self._population_size)]
+
+                r1_index = np.random.randint(0, self._population_size)
+                while r1_index == j: r1_index = np.random.randint(0, self._population_size)
+
+                x_r1 = self._population[r1_index]
 
                 r2_index = np.random.randint(0, self._population_size + external_archive_size)
+                while r2_index == j or r2_index == r1_index:
+                    r2_index = np.random.randint(0, self._population_size + external_archive_size)
 
                 if r2_index < self._population_size:
 
@@ -175,7 +187,9 @@ class SHADE:
                 #         mutant[i] = np.random.rand()
 
                 #Obtenemos los puntos de cruce segun cr_i
-                cross_points = np.random.rand(self._dim) <= cr_i
+                cross_points1 = np.random.rand(self._dim) <= cr_i
+                cross_points2 = np.array(range(self._dim)) == np.random.randint(self._dim, size=self._dim)
+                cross_points = np.logical_or(cross_points1, cross_points2)
 
                 #Obtenemos el vector hijo de x_i
                 trial = np.where(cross_points, mutant, x_i)
@@ -240,7 +254,9 @@ class SHADE:
[…]

### 2018-12-10 Software/SHADEMain.py

@@ -13,12 +13,12 @@ def main():
 
     names_array, datasets_array, labels_array = load_all_datasets()
 
-    names_array = names_array[0:4]
-    datasets_array = datasets_array[0:4]
-    labels_array = labels_array[0:4]
+    names_array = names_array[0:2]
+    datasets_array = datasets_array[0:2]
+    labels_array = labels_array[0:2]
 
     const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    # const_percent_vector = [0.1, 0.2]
+    const_percent_vector = [0.1, 0.2]
     const_array = load_constraints(names_array, const_percent_vector)
 
     general_table_file = open("Results/SHADE_general_table_file.txt", "w+")
@@ -30,75 +30,67 @@ def main():
     max_eval = 300000
     population_size = 100
 
-    for label_percent in range(len(const_percent_vector)):
+    for i in range(1, 11):
 
-        print("------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[label_percent]) + " ------------")
-        general_table_file.write("------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[label_percent]) + " ------------\n")
-        general_table_file.write("Dataset RandIndex   Time(s)   Unsat(%)   TotalRestr   ML(%)   CL(%)\n")
-        results_file.write("------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[label_percent]) + " ------------\n")
-        results_file.write("Dataset RandIndex   Time(s)   Unsat(%)\n")
+        prt_elite = i/10
 
-        for i in range(len(names_array)):
+        print("------------ Probando con elite: " + str(prt_elite) + " ------------\n")
+        general_table_file.write("------------ Probando con elite: " + str(prt_elite) + " ------------\n")
+        results_file.write("------------ Probando con elite: " + str(prt_elite) + " ------------\n")
 
-            name = names_array[i]
-            data_set = datasets_array[i]
-            labels = labels_array[i]
-            nb_clust = len(set(labels))
+        for label_percent in range(len(const_percent_vector)):
 
-            print("------------ Procesando " + name + " Dataset ------------ (" + str(const_percent_vector[label_percent]) + ")")
-            const = const_array[label_percent][i]
-            ml_const, cl_const = get_const_list(const)
[…]

### 2018-12-10 Software/SHADEMainParallel.py

@@ -22,7 +22,7 @@ def apply(data_set, p_size, nb_clust, max_eval, nb_runs, name, label_percent, la
     mean_unsat_percent = 0
 
     for j in range(nb_runs):
-        shade = SHADE(data_set, ml_const, cl_const, p_size, nb_clust, True)
+        shade = SHADE(data_set, ml_const, cl_const, p_size, nb_clust, 0.25, True)
         start = time.clock()
         de_assignment = shade.run(max_eval)[1]
         end = time.clock()
@@ -34,7 +34,7 @@ def apply(data_set, p_size, nb_clust, max_eval, nb_runs, name, label_percent, la
     mean_execution_time /= nb_runs
     mean_unsat_percent /= nb_runs
 
-    return tuple((name, nb_const, ml_const_percent, cl_const_percent, mean_ars, mean_execution_time, mean_unsat_percent))
+    return tuple((name, mean_ars, mean_execution_time, mean_unsat_percent))
 
 
 def main():
@@ -49,55 +49,46 @@ def main():
     datasets_array = datasets_array[:4]
     labels_array = labels_array[:4]
 
-    # names_array = [names_array[0]]
-    # datasets_array = [datasets_array[0]]
-    # labels_array = [labels_array[0]]
-
-    general_table_file = open("Results/SHADE_general_table_file.txt", "w+")
-    results_file = open("Results/SHADE_results_file.txt", "w+")
-
     #BUCLE DE OBTENCION DE DATOS
 
     const_percent_vector = np.array([0.05, 0.1, 0.15, 0.2])
-    #const_percent_vector = np.array([0.05])
+    #const_percent_vector = np.array([0.1, 0.2])
     nb_runs = 1
     max_eval = 300000
     population_size = 100
 
     general_start = time.time()
+    results_matrix = np.zeros((len(names_array), len(const_percent_vector)))
 
     with Parallel(n_jobs=multiprocessing.cpu_count() - 1) as parallel:
 
-        for label_percent in const_percent_vector:
+        for label_percent in range(len(const_percent_vector)):
 
-            general_table_file.write(
-                "------------ Procesando en porcentaje de restricciones: " + str(label_percent) + " ------------\n")
-            general_table_file.write("Dataset RandIndex   Time(s)   Unsat(%)   TotalRestr   ML(%)   CL(%)\n")
-            results_file.write(
[…]

### 2018-12-10 Software/tablemaker.py

@@ -1,0 +1,42 @@
+from functions import *
+import numpy as np
+import collections
+
+def main():
+
+	names_array, datasets_array, labels_array = load_all_datasets()
+	#names_array = names_array[0:4]
+
+	for i in range(len(names_array)):
+
+		n = np.shape(datasets_array[i])[0]
+		nb_class = len(set(labels_array[i]))
+
+		print(str(n) + " " + names_array[i].title() + " & " + str(n) + " & " + str(nb_class) + " &  \\\\")
+
+
+	const_percent_vector = [0.05, 0.1, 0.15, 0.2]
+	const_array = load_constraints(names_array, const_percent_vector)
+
+	for i in range(len(names_array)):
+
+		print(names_array[i].title(), end='')
+
+		for l in range(len(const_percent_vector)):
+
+			c = const_array[l][i]
+			total = np.count_nonzero(c)
+			total = (total - np.shape(c)[0]) / 2
+			unique, counts = np.unique(c, return_counts=True)
+			d = dict(zip(unique, counts))
+			ml = (d[1] - np.shape(c)[0]) / 2
+			cl = d[-1] / 2
+
+			print(" & %d & %d & %d" % (total, ml, cl), end='')
+
+		print(" \\\\")
+
+
+
+
+if __name__ == "__main__": main()

### 2018-12-11 Papers/SHADE/elsarticle-template/paperSHADE.tex
SHADE added to paper

@@
 \usepackage{mathtools}
+\usepackage{multirow}
+\usepackage{adjustbox}
+\usepackage{chngpage}
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
@@
 
+\section{Background}
+
 \section{Constrained Clustering}
@@
 Clustering is the task of grouping instances of a dataset into subsets.
-The criterion used to assign an instance to a given cluster is the similarity with the rest of elements in that cluster and the dissimilarity with the rest of instances of the dataset.
+The criterion used to assign an instance to a given cluster is the similarity with the rest of elements in that cluster, and the dissimilarity with the rest of instances of the dataset.
 In this way, the clustering process obtains subsets of instances of the initial data set that present high intra-similarity and low inter-similarity.
@@
 On the other hand, Cannot-link restrictions---from now on CL---do not constitute an equivalence relation, but they can be entailed.
-To illustrate that, we consider the instances $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then is also true that $c_{\neq}(x_c,x_d)$ for all $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
+To illustrate that, we consider the instances $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then it is also true that $c_{\neq}(x_c,x_d)$ for all $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
 
@@
 
-The BRKGA optimization process is completely independent of the problem.
+The BRKGA optimization process is completely independent of the problem to solve.
 First, a population $P$ of $p_{size}$ vectors of random-keys is initialized, this will be the initial generation.
@@
 
-\subsection{Adaptation of BRKGA for Constrained Clustering}
+\subsection{Adaptation of BRKGA for Constrained Clustering} \label{sec:AdaptationofBRKGA}
 
@@
 	\renewcommand{\arraystretch}{0.9}
-
+	\resizebox{\textwidth}{!}{
 	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
@@
 
-	\end{tabular}
+	\end{tabular}}
 	\caption{Random-key decodification example \cite{de2017comparison}}
@@
 
-Success history based adaptive differential evolution---de ahora en adelante SHADE---is an optimization method based on differential evolution.
+Success history based adaptive differential evolution---from now on SHADE---is an optimization method based on differential evolution.
 Proposed by Tanabe and Fukunaga (2013), it was an improvement for the JADE model, created by Zhang and Sanderson \cite{tanabe2013success}\cite{zhang2009jade}.
@@
 
-To maintain diversity, SHADE can make use of an external $A$ archive, in which parents $p_{i,G}$ who were replaced by their associated trial vector $t_{i,G}$ are saved.
+To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{i,G}$ who were replaced by their associated trial vector $t_{i,G}$ are saved.
 To make use of the archive, we will consider that the individual $p_{r2,G}$ from Equation \eqref{eq3} is selected from $P \cup A$, which is the union of the population and the file $A$.
@@
 	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{10pt}
-	\renewcommand{\arraystretch}{0.9}
-
-	\begin{tabular}{|>{\centering\arraybackslash}m{1cm}  |>{\centering\arraybackslash}m{1cm}|>{\centering\arraybackslash}m{1cm}|>{\centering\arraybackslash}m{1cm}|>{\centering\arraybackslash}m{1.4cm}|>{\centering\arraybackslash}m{1.2cm}|}
+	\setlength{\tabcolsep}{13pt}
+	%\renewcommand{\arraystretch}{0.9}
+	\resizebox{\textwidth}{!}{
+	\begin{tabular}{|c|c|c|c|c|c|}
 		\hline
@@
 
-	\end{tabular}
+	\end{tabular}}
 	\caption{The historical memory $M_{CR}$, $M_{F}$  \cite{tanabe2013success}}
@@
 
-In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the wighted mean,  which is computed following Equations \eqref{eq10} and \eqref{eq11}, which were proposed by Peng et al (2009) \cite{peng2009multi} to prevent $CR$ to converge to small value.
+In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the wighted mean,  which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed by Peng et al (2009) \cite{peng2009multi} to prevent $CR$ to converge to small values.
 
@@
 
-\newpage
+Para adaptar SHADE para resolver el porblema del clustering con restricciones tomaremos algunos de los elementos que permitian adaptarlo a BRKGA.
+Haremos uso del concepto de random-key para crear los individuos de la población, siendo cada individuo de la misma un vector de random-keys.
+Utilizaremos el mismo método de inicialización y decodificación que el empleado en BRKGA (consultar Sección \ref{sec:AdaptationofBRKGA}).
+Proponemos una nueva función fitness para evaluar los inidividuos de la población, definida en la Ecuación \eqref{eq14}, donde $z_i$ se obtiene como en la Ecuacion \eqref{eq2}.
+
+To adapt SHADE to solve the constrained clustering problem we will take some of the elements that allowed us to adapt it to BRKGA.
+We will use random-keys to create the individuals of the population, this way each individual is defined as a vector of random-keys.
+We will also use the same initialization method and the same decoder used in BRKGA (see Section \ref{sec:AdaptationofBRKGA}).
+
+For SHADE, we propose a new fitness function to evaluate the individuals of the population.
+Shade fitness function is defined in Equation \eqref{eq14}, where $z_i$ is obtained as in Equation \eqref{eq2}.
+
+\begin{equation}
+f_i = z_i * (infeasibility + 1)
+\label{eq14}
+\end{equation}
+
+Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated based on the dataset or constraints, whereas in \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.
+
+We found that with the fitness function defined in \eqref{eq1} there is no competition between the penalty term and the within-cluster-sum-of-squares term, because penalty is always way larger than it or zero.
+This can bias the  exploration of the solutions space to those that satisfies all the constraints, even if the within-cluster-sum-of-squares is still improvable by moving two instances implied in a constraint to different clusters without violating that constraint.
+
+With Equation \eqref{eq14} we try to find a trade-off between the within-cluster-sum-of-squares and the penalty term, allowing solutions that violate a certain number of constraints to compete with those who violate a lesser number of them but find a better within-cluster-sum-of-squares.
+
+\textcolor{red}{Poner una pintura que ejemplifique lo anterior.}\\
+\textcolor{red}{Explicar la busqueda local. Con pseudocódigo?}
+
+\clearpage
+
+\section{Experimental Setup}
+
+\subsection{Dataset}
+
+\begin{table}[!h]
+	\centering
+	%\setlength{\arrayrulewidth}{1mm}
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{0.9}
+	\resizebox{\textwidth}{!}{
+	\begin{tabular}{c c c c}
+		\hline
+		Name & Number of Instances & Number of Classes & Features \\
+		\hline
+		& & & \\
+		& & & \\
+		& & & \\
+		\hline
+
+	\end{tabular}}
+	\caption{Summary of datasets used for the experiments \textcolor{red}{citar keel}}
+	\label{undefined}
+\end{table}
+
+\subsection{Constraints Generation}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{0.9}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+	\begin{tabular}{lcc c cc c cc c cc}
+		\hline
+		\multirow{2}{*}{Dataset} &
+		\multicolumn{2}{c}{Percent 1\%} &&  \multicolumn{2}{c}{Percent 5\%} && \multicolumn{2}{c}{Percent 15\%} && \multicolumn{2}{c}{Percent 20\%} \\
+		\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
+		& ML & CL && ML & CL && ML & CL && ML & CL \\
+		\hline
+		& & && & && & && & \\
+		& & && & && & && & \\
+		& & && & && & && & \\
+		\hline
+
+	\end{tabular}}
+	%\end{adjustwidth}
+
+	\caption{Number of constraints used in experiments}
+	\label{undefined}
+\end{table}
+
+\subsection{Evaluation Method}
+
+We use the evaluation method proposed by Wagstaff et al. (2001) \cite{wagstaff2001constrained}, the authors of the Constrained K-Means algorithm, proposed in their work to evaluate the results.
+
+Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
+We will use Rand Index \cite{rand1971objective} to measure the accuracy of the predictions resulting from each method we test.
+Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
+$C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions.
+
+For each pair of $x_i$ and $x_j$ instances in $X$, $C_i$ assigns them to the same cluster or to different clusters.
+We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in $C_1$ and $C_2$ , and $b$ as the opposite event.
+Then, the degree of similarity between $C_1$ and $C_2$ is calculated as in Equation \eqref{eq15}.
+
+\begin{equation}
+Rand(C_1, C_2) = \frac{a + b}{N(N - 1)/2}
+\label{eq15}
+\end{equation}
+
+We will use this measure co compute the accuracy for all of our experiments.
+
+\section{Experimental Results}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{0.9}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{l ccc c ccc}
+			\hline
+			\multirow{2}{*}{Dataset} &
+			\multicolumn{3}{c}{BRKGA} &&  \multicolumn{3}{c}{SHADE} \\
+			\cline{2-4} \cline{6-8}
+			& RandIndex & Unsat(\%) & Time(s) && RandIndex & Unsat(\%) & Time(s) \\
+			\hline
+			&  &  &  &&  &  &  \\
+			&  &  &  &&  &  &  \\
+			&  &  &  &&  &  &  \\
+			\hline
+			Means &  &  &  &&  &  &  \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained with X\% of labeled data}
+	\label{undefined}
+\end{table}
+
+\section{Acknowledgement}
+
+\section{Conclusions and Future Work}
+
+\clearpage
 

### 2018-12-17 Papers/SHADE/paperSHADE.tex
Correcciones en memoria

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{chngpage}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\modulolinenumbers[5]

\journal{Applied Soft Computing}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
\newtheorem{definition}{Definition}
\begin{document}

\begin{frontmatter}

\title{Enhancing instance-level constrained clustering through differential evolution}

\author[mymainaddress]{Germ\'an Gonz\'alez Almagro\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{germangalmagro@ugr.es}

\author[mymainaddress]{Salvador Garc\'ia}

\address[mymainaddress]{Department of Computer Science and Artificial Intelligence, University of Granada, 18071 Granada, Spain}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

\section{Background}

\subsection{Constrained Clustering}

Clustering is the task of grouping instances of a dataset into subsets. The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset. In this way, the clustering process obtains subsets of instances of the initial data set that present high intra-similarity and low inter-similarity.

In most clustering applications it is common to have some kind of information about the data set to be analyzed. In constrained clustering this information is given in the form of pairs of instances. A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster. Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms.

Before formalizing the definition of constraints we must give a notation for the data set. We will note our dataset as $X$, which is a set of $N$ instances with $D$ features. We well refer to the instances of $X$ as $x$. Now, given the definition of dataset, we can distinguish between two types of constraints:

\begin{itemize}

	\item Must-link constraints $c_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.

	\item Cannot-link constraints $c_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.

\end{itemize}

The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set. As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $N = |X| = \sum_{i = 1}^{K} |c_i|$.

Knowing how a constraint is defined, Must-link constraints---from now on ML---are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric. This way, given constraints $c_=(x_a,x_b)$ and $c_=(x_b,x_c)$ then $c_=(x_a,x_c)$ is verified. In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $c_=(x_a,x_b)$, then $c_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.

On the other hand, Cannot-link constraints---from now on CL---do not constitute an equivalence relation***, but they also can be entailed***. However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then it is also true that $c_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.

\subsection{The Feasibility Problem}

Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the problem complexity. The feasibility problem for constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.

\begin{definition}

	\textbf{Feasibility Problem}: given a dataset $X$, a set of constraints $R$, and the bounds on the number of clusters $K_l \leq k \leq K_u$. Does there exists a partition $C$ of $X$ with $k$ clusters such that all constraints in $R$ are satisfied? \cite{davidson2005clustering}
	\label{def1}

\end{definition}

In \cite{davidson2005clustering} it is proven that, when $K_l = 1$ and $K_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem. (It is also proven that it is not harder than it, so both have the same complexity.) Table \ref{tab:feasibility} shows feasibility problem complexity for different types of constraints.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	%\resizebox{\textwidth}{!}{
		\begin{tabular}{c c}
			\hline
			Constraints & Complexity \\
			\hline
			Must-Link & $\mathbf{P}$\\
			Cannot-Link & $\mathbf{NP}$-complete\\
			ML and CL & $\mathbf{NP}$-complete\\
			\hline

		\end{tabular}%}
	\caption{Feasibility problem complexity \cite{davidson2005clustering}}
	\label{tab:feasibility}
\end{table}

These complexity results show that feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too. For more details on constrained clustering problem complexity see \cite{davidson2005clustering}.

Intractable problems are hard to solve with deterministic and exact methods. That is the reason why applying heuristics-guided and population-based algorithms is a good approach to find quality solutions to the constrained clustering problem.

\section{Brief Review of the BRKGA Algorithm}

The biased random-key genetic algorithm---from now on BRKGA---was first proposed by \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.

In BRKGA, each solution---also called individual---is represented as a vector of values within the interval $[0,1]$. However, this vector is not a solution to the problem but a representation of it. This allows us to abstract the details of the problem in order to apply a genetic algorithm that operates with real coding. Therefore, a deterministic decoder is necessary to obtain the actual solution to the problem from the vector that encodes it.

BRKGA also requires a fitness function to evaluate each individual. This function takes the decoded solution as input and provides its fitness value as output. It is designed specifically for each problem.

The BRKGA optimization process is completely independent of the problem to solve. First, a population $P$ of $p_{size}$ vectors of random-keys is initialized, and this will be the initial generation. Each random-key vector $p_i$ has $N$ random-keys in it. The population is sorted by the fitness value $f_i$ of each $p_i$ to select the first $p_e$ individuals, this is, the elite of the population. The elite will be preserved in the next generation without modification. To introduce diversity, a number $p_m$ of new random-keys vectors are also included in the next generation. The remaining individuals of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents \cite{de2017comparison}.

\subsection{Adaptation of BRKGA for Constrained Clustering} \label{sec:AdaptationofBRKGA}

An adaptation of BRKGA for constrained clustering was proposed in \cite{de2017comparison}.

Their method randomly initializes a population of random-key vectors, each one with $N$ random-keys. The decoder divides the interval $[0,1]$ in $K$ intervals, so there exists a correspondence between each random-key and the integer corresponding to the interval which it lies in. Table \ref{tab:decodingrk} shows an example of random-key decoding for $K = 3$. Note that extreme values 0 and 1 can also appear in a random-key vector.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
		\hline
		Index & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
		\hline
		Random-key & 0.12 & 0.37 & 0.66 & 0.56 & 0.00 & 0.97 & 0.23 & 0.25 & 0.15 & 1.00 \\
		\hline
		Clusters & 1 & 2 & 2 & 2 & 1 & 3 & 1 & 3 & 1 & 3 \\
		\hline

	\end{tabular}}
	\caption{Random-key decodification example \cite{de2017comparison}}
	\label{tab:decodingrk}
\end{table}

Once the decoded solution has been obtained, the fitness value of the solution is computed as in Equation \eqref{eq1}.

\begin{equation}
f_i = z_i + \overbrace{(\mu * N * {infeasibility}_i)}^{penalty}
\label{eq1}
\end{equation}

Where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.

\begin{equation}
z_i = \sum_{c_j \in C_i} \left[ \frac{\sum_{x_a, x_b \in c_j} d_D^2(x_a,x_b)}{|c_j|}\right]
\label{eq2}
\end{equation}

Where $C_i$ is the partition of dataset $X$ defined by solution $p_i$, and $(x_a, x_b)$ are instances in the cluster $c_j$ such that $a \neq b$ and the distance between each pair of instances in $c_i$ is included in the sum only once.

In \cite{de2017comparison} authors also added a new element to BRKGA, a local search procedure. This local optimization method is applied to each decoded individual in each generation, but its results are not transferred to the individual, so that diversity is maintained. Thus, the individuals produced by the local search are only used to update the best solution found so far if needed. For more details on BRKGA see \cite{de2017comparison}.

\section{The SHADE Optimization Method}

Success history based adaptive differential evolution---from now on SHADE---is an optimization method based on differential evolution. Proposed by \cite{tanabe2013success}, it was an improvement for the JADE model, created by \cite{zhang2009jade}.

Unlike JADE, which only considers the parameters used in the previous generation, SHADE uses a historical record of successful parameters as a mechanism for adapting parameters involved in the process of creating new generations.

\subsection{SHADE Operators}

As in differential evolution, SHADE employs mutant generation, crossover and replacement operators to explore the space of solutions and bring in new generations of individuals.

The mutation strategy used by SHADE is known as current-to-pbest/1, and the expression that generates new individuals is described in Equation \eqref{eq3}.

\begin{equation}
m_{i,G} = p_{i,G} + F_i * (p_{pbest, G} - p_{i,G}) + F_i * (p_{r1, G} - p_{r2,G})
\label{eq3}
\end{equation}

Where $m_{i,G}$ is the mutant vector, which is generated with an individual $p_{i,G}$ from the population [serving as a starting point]. Indices $r1$ and $r2$ are random values in the range $[0,N]$, different from $i$ and from each other. The individual $p_{pbest, G}$ is randomly selected from among the best $N \times pbest\;|\;pbest\in [0,1]$ individuals in the population. This way, the parameter $p$ controls the greediness of the mutation strategy. The parameter $F_i$ defines the magnitude of the operator.

After generating the mutant vector $m_{i,G}$, it is combined with the parent $p_{i,G}$ vector by means of a crossover operator; the result is the trial vector. SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.

\begin{equation}
t_{j,i,G} = \left\{ \begin{array}{lc}
m_{j,i,G} &   if \;\; rand[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
p_{j,i,G} &  otherwise
\end{array}
\right.
\label{eq4}
\end{equation}

Where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.

The SHADE replacement operator determines the survivors individuals for the next generation. It compares each parent $x_{i,G}$ with the trial vector $_{j,i,G}$, generated based on it. The best one of these two individuals will survive for the next generation. Equation \eqref{eq5} shows this concept.

\begin{equation}
p_{i,G + 1} = \left\{ \begin{array}{lc}
t_{i,G} &   if \;\; f(t_{i,G}) \le f(p_{i,G}) \\
p_{i,G} &  otherwise
\end{array}
\right.
\label{eq5}
\end{equation}

To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{i,G}$ who were replaced by their associated trial vector $t_{i,G}$ are saved. To make use of the archive, we will consider that the individual $p_{r2,G}$ from Equation \eqref{eq3} is selected from $P \cup A$. The archive size is the same as the population size. When the size of $A$ exceeds the size of $P$, random individuals are selected for removal to make room for new ones.

The parameters $p$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process. These are the parameters that SHADE adaptively optimizes using the mentioned history record.

\subsection{SHADE Parameter Adaptation Method}

The SHADE method stores in memory a structure with $H$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:SHADEmemory}. This structure is initialized with the value $0.5$ for all entries.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{13pt}
	%\renewcommand{\arraystretch}{0.9}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		 Index & 1 & 2 & $\cdots$ & $H - 1$ & $H$ \\
		 \hline
		 \hline
		 $M_{CR}$ & $M_{CR,1}$ & $M_{CR,1}$ & $\cdots$ & $M_{CR,H-1}$ & $M_{CR,H}$ \\
		 \hline
		 $M_{F}$ & $M_{F,1}$ & $M_{F,1}$ & $\cdots$ & $M_{F,H-1}$ & $M_{F,H}$ \\
		\hline

	\end{tabular}}
	\caption{Historical memory $M_{CR}$, $M_{F}$ used by SHADE \cite{tanabe2013success}}
	\label{tab:SHADEmemory}
\end{table}

In each generation, parameters $CR_i$ and $F_i$ are calculated on the basis of the existing history by applying Equations \eqref{eq6} and \eqref{eq7}:

\begin{equation}
CR_i = randn_i(M_{CR,r_i}, 0.1)
\label{eq6}
\end{equation}

\begin{equation}
F_i = randc_i(M_{F,r_i}, 0.1)
\label{eq7}
\end{equation}

where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$. When a value for $CR_i$ outside of the range $[0,1]$ is generated, it is replaced by the corresponding limit value. If $F_i > 1$, then it is truncated to $1$; conversely, if $F_i < 0$, then Equation \eqref{eq7} is applied as many times as needed to obtain a legal value.

SHADE keeps two auxiliar sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent. At the end of each generation, the content of the memory $M$ is updated following Equations \eqref{eq8} and \eqref{eq9}.

\begin{equation}
M_{CR,h,G+1} = \left\{ \begin{array}{lc}
mean_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
M_{CR,h,G+1} &  otherwise
\end{array}
\right.
\label{eq8}
\end{equation}

\begin{equation}
M_{F,h,G+1} = \left\{ \begin{array}{lc}
mean_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
M_{F,h,G+1} &  otherwise
\end{array}
\right.
\label{eq9}
\end{equation}

The $h \;\; (0 \le h \le H)$ index specifies the position of the memory $M$ to be updated. This index is initialized to $0$ at the beginning of the optimization process and increased by one after each generation. When $h \ge H$ then $h$ is reset to 1. It is worth noting that, when there is a generation with no trial vectors successfully replacing their parents, no update of $M$ is done.

In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the weighted mean, which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed by \cite{peng2009multi} to prevent $CR$ from converging to small values.

\begin{equation}
mean_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{RC,i}
\label{eq10}
\end{equation}

\begin{equation}
\omega_i = \frac{\Delta f_i}{\sum_{i = 1}^{|S_{RC}|} \Delta f_i}
\label{eq11}
\end{equation}

where $\Delta f_i = |f(t_i,G) - f(p_i, G)|$, which is the amount of improvement obtained from the trial vector with respect to the parent.

In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12} (proposed by \cite{tanabe2013success}).

\begin{equation}
mean_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{F,i}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{F,i}}
\label{eq12}
\end{equation}

Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the qreediness of the mutation strategy, is not included in the optimization process. However, it is not static: it is calculated for each individual $p_i$ of the population following Equation \eqref{eq13}.

\begin{equation}
pbest_i = randn[2/N, 0.2]
\label{eq13}
\end{equation}

such that there is always at least two individuals to choose between.

\subsection{Adaptation of SHADE for constrained clustering}

For the adaptation of SHADE to the problem of constrained clustering, we will take some of the elements that allowed us to adapt it to BRKGA. To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of random-keys. Additionally, we will also reuse the same initialization method and the same decoder used in BRKGA (see Section \ref{sec:AdaptationofBRKGA}).

However, we propose a new fitness function to evaluate the quality of the individuals of the population, Equation \eqref{eq14}, whose element $z_i$ is obtained as in Equation \eqref{eq2}.

\begin{equation}
f_i = z_i * (infeasibility + 1)
\label{eq14}
\end{equation}

Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated from the dataset or constraints, whereas in \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.

We found that with the fitness function defined in \eqref{eq1} there is no competition between the penalty term and the within-cluster-sum-of-squares term, because the penalty is always significantly larger than it or zero. This can bias the exploration of the solution space, restricting it in practice to those that satisfy all the constraints, even if the within-cluster-sum-of-squares is still improvable by moving two instances involved in a constraint to different clusters without violating that constraint.

With Equation \eqref{eq14} we try to find a trade-off between the within-cluster-sum-of-squares and the penalty term, allowing solutions that violate a certain number of constraints to compete with those who violate a smaller number of them but score a better within-cluster-sum-of-squares.

\textcolor{red}{Poner una pintura que ejemplifique lo anterior.}\\
\textcolor{red}{Explicar la busqueda local. Con pseudocódigo?}

\clearpage

\section{Experimental Setup}

\subsection{Dataset}

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{5pt}
	\renewcommand{\arraystretch}{0.8}
	%\resizebox{\textwidth}{!}{
	\small
	\begin{tabular}{l c c c}
		\hline
		Name & No. Instances & No. Classes & Features \\
		\hline
		Appendicitis & 106 & 2 &  \\
		Balance & 625 & 3 &  \\
		Boston & 506 & 229 &  \\
		Breast Cancer & 569 & 2 &  \\
		Bupa & 345 & 2 &  \\
		Circles & 300 & 2 &  \\
		Contraceptive & 1473 & 3 &  \\
		Diabetes & 442 & 214 &  \\
		Ecoli & 336 & 8 &  \\
		Glass & 214 & 6 &  \\
		Haberman & 306 & 2 &  \\
		Hayesroth & 160 & 3 &  \\
		Heart & 270 & 2 &  \\
		Ionosphere & 351 & 2 &  \\
		Iris & 150 & 3 &  \\
		Led7Digit & 500 & 10 &  \\
		Monk2 & 432 & 2 &  \\
		Moons & 300 & 2 &  \\
		Movement Libras & 360 & 15 &  \\
		Newthyroid & 215 & 3 &  \\
		Pima & 768 & 2 &  \\
		Rand & 150 & 3 &  \\
		Saheart & 462 & 2 &  \\
		Sonar & 208 & 2 &  \\
		Soybean & 47 & 4 &  \\
		Spectfheart & 267 & 2 &  \\
		Spiral & 300 & 2 &  \\
		Tae & 151 & 3 &  \\
		Vehicle & 846 & 4 &  \\
		Vowel & 990 & 11 &  \\
		Wdbc & 569 & 2 &  \\
		Wine & 178 & 3 &  \\
		Zoo & 101 & 7 &  \\
		\hline

	\end{tabular}%}
	\caption{Summary of datasets used for the experiments \textcolor{red}{Citar keel}}
	\label{tab:datasets}
\end{table}

\subsection{Constraints Generation}

Since we have the true labels associated with each data set, we will use the method proposed by Wagstaff and Cardie to generate artificial constraint sets \cite{wagstaff2001constrained}. This method consists in randomly selecting two instances of a data set, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.

We will generate four different sets of constraints for each data set. The number of constraints is given by the number of edges of the complete constraints graph that we can build with a percentage of labeled data. Each constraint set is associated with a different percentage of labeled data, namely: 5\%, 10\%, 15\% and 20\%. It is worth noting that constraints are not the complete constraint graph made up of the tagged instances, this graph only sets the number of constraints to be created, this way there is a lower probability of biasing the set of constraints so that there are classes with poor representation in it.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.8}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{lcc c cc c cc c cc}
		\hline
		\multirow{2}{*}{Dataset} &
		\multicolumn{2}{c}{5\%} &&  \multicolumn{2}{c}{10\%} && \multicolumn{2}{c}{15\%} && \multicolumn{2}{c}{20\%} \\
		\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
		& ML & CL && ML & CL && ML & CL && ML & CL \\
		\hline
		Appendicitis & 13 & 2 && 39 & 16 && 71 & 49 && 164 & 67 \\
		Balance & 198 & 298 && 838 & 1115 && 1865 & 2506 && 3336 & 4414 \\
		Boston & 0 & 325 && 3 & 1272 && 13 & 2837 && 26 & 5125 \\
		Breast Cancer & 216 & 190 && 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
		Bupa & 79 & 74 && 323 & 272 && 699 & 627 && 1201 & 1145 \\
		Circles & 50 & 55 && 208 & 227 && 502 & 488 && 853 & 917 \\
		Contraceptive & 994 & 1707 && 3765 & 7113 && 8501 & 15809 && 15399 & 27966 \\
		Diabetes & 2 & 251 && 5 & 985 && 7 & 2204 && 23 & 3893 \\
		Ecoli & 30 & 106 && 163 & 398 && 357 & 918 && 609 & 1669 \\
		Glass & 11 & 44 && 52 & 179 && 139 & 389 && 259 & 644 \\
		Haberman & 76 & 44 && 304 & 161 && 634 & 401 && 1135 & 756 \\
		Hayesroth & 12 & 16 && 39 & 81 && 102 & 174 && 177 & 319 \\
		Heart & 41 & 50 && 178 & 173 && 396 & 424 && 744 & 687 \\
		Ionosphere & 92 & 61 && 330 & 300 && 732 & 646 && 1299 & 1186 \\
		Iris & 9 & 19 && 26 & 79 && 82 & 171 && 136 & 299 \\
		Led7Digit & 25 & 275 && 126 & 1099 && 267 & 2508 && 460 & 4490 \\
		Monk2 & 101 & 130 && 473 & 473 && 979 & 1101 && 1917 & 1824 \\
		Moons & 55 & 50 && 200 & 235 && 494 & 496 && 900 & 870 \\
		Movement Libras & 6 & 147 && 27 & 603 && 112 & 1319 && 158 & 2398 \\
		Newthyroid & 25 & 30 && 108 & 123 && 270 & 258 && 449 & 454 \\
		Pima & 412 & 329 && 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
		Rand & 8 & 20 && 25 & 80 && 76 & 177 && 151 & 284 \\
		Saheart & 152 & 124 && 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
		Sonar & 29 & 26 && 100 & 110 && 245 & 251 && 436 & 425 \\
		Soybean & 0 & 3 && 4 & 6 && 6 & 22 && 12 & 33 \\
		Spectfheart & 56 & 35 && 233 & 118 && 543 & 277 && 965 & 466 \\
		Spiral & 52 & 53 && 224 & 211 && 487 & 503 && 918 & 852 \\
		Tae & 8 & 20 && 40 & 80 && 82 & 171 && 151 & 314 \\
		Vehicle & 221 & 682 && 874 & 2696 && 1955 & 6046 && 3589 & 10776 \\
		Vowel & 107 & 1118 && 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
		Wdbc & 209 & 197 && 840 & 756 && 1925 & 1730 && 3472 & 2969 \\
		Wine & 14 & 22 && 49 & 104 && 121 & 230 && 217 & 413 \\
		Zoo & 7 & 8 && 21 & 34 && 29 & 91 && 41 & 169 \\
		\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Number of constraints used in experiments}
	\label{tab:constraints}
\end{table}

\subsection{Evaluation Method}

We use the evaluation method proposed by Wagstaff et al. (2001) \cite{wagstaff2001constrained}, the authors of the Constrained K-Means algorithm, proposed in their work to evaluate the results.

Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method. We will use Rand Index \cite{rand1971objective} to measure the accuracy of the predictions resulting from each method we test. Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$. $C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions.

For each pair of $x_i$ and $x_j$ instances in $X$, $C_i$ assigns them to the same cluster or to different clusters. We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in $C_1$ and $C_2$ , and $b$ as the opposite event. Then, the degree of similarity between $C_1$ and $C_2$ is calculated as in Equation \eqref{eq15}.

\begin{equation}
Rand(C_1, C_2) = \frac{a + b}{N(N - 1)/2}
\label{eq15}
\end{equation}

We will use this measure co compute the accuracy for all of our experiments.

\subsection{Calibration}

\section{Experimental Results}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA} &&  \multicolumn{3}{c}{SHADE} \\
			\cline{2-4} \cline{6-8}
			& RandIndex & Unsat(\%) & Time(s) && RandIndex & Unsat(\%) & Time(s) \\
			\hline
			&  &  &  &&  &  &  \\
			&  &  &  &&  &  &  \\
			&  &  &  &&  &  &  \\
			\hline
			Means &  &  &  &&  &  &  \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Experimental results obtained with X\% of labeled data}
	\label{undefined}
\end{table}

\section{Conclusions and Future Work}

\section{Acknowledgements}

\clearpage

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2018-12-17 Software/BRKGA/BRKGA.py

@@ -1,6 +1,7 @@
 import numpy as np
 from sklearn.neighbors.nearest_centroid import NearestCentroid
 import copy as cp
+import time
 
 class BRKGA:
 
@@ -77,11 +78,6 @@ class BRKGA:
                 tot = 0.
                 for k in range(clust.shape[0] - 1):
                     tot += ((((clust[k + 1:] - clust[k]) ** 2).sum(1)) ** .5).sum()
-
-                if ((clust.shape[0] - 1) * (clust.shape[0]) / 2.) == 0.0:
-                    print(tot)
-                    print(cromosome)
-                    print(clust)
 
                 avg = tot / ((clust.shape[0] - 1) * (clust.shape[0]) / 2.)
 
@@ -182,10 +178,12 @@ class BRKGA:
         num_mutants = int(self._population_size * self._pbt_mutation)
         offspring_size = self._population_size - num_elite - num_mutants
         generations = 0
+        gen_times = []
 
         #Mientras no se haya alcanzado el numero maximo de evaluaciones
         while self._evals_done < max_eval:
 
+            start_gen = time.time()
             #Guardar la elite de la generacion actual
             elite = self._population[:num_elite, :]
             non_elite = self._population[num_elite:, :]
@@ -217,10 +215,9 @@ class BRKGA:
                 self.local_search()
 
             generations += 1
+            gen_times.append(time.time() - start_gen)
 
-            #print(str(generations) + " " + str(fitness[sorted_fitness[0]]))
-
-        return self.decode_single_random_key(self._population[sorted_fitness[0]]), self._best
+        return self.decode_single_random_key(self._population[sorted_fitness[0]]), self._best, np.median(gen_times) * generations
 
     #Busqueda local por trayectorias simples
     def local_search(self):

### 2018-12-17 Software/Paper1Main.py

@@ -1,0 +1,91 @@
+import numpy as np
+from SHADE.SHADE import SHADE
+from BRKGA.BRKGA import BRKGA
+from functions import *
+from sklearn.metrics import adjusted_rand_score
+import time
+import sys
+
+
+def main():
+
+    if len(sys.argv) != 2:
+        print("Numero de argumentos incorrecto")
+        return -1
+
+    np.random.seed(11)
+
+    #CARGA DE DATOS
+
+    names_array, datasets_array, labels_array = load_all_datasets()
+
+    const_percent_vector = [float(sys.argv[1])]
+    const_array = load_constraints(names_array, const_percent_vector)
+    results_matrix = np.zeros((len(names_array), 6))
+
+    #BUCLE DE OBTENCION DE DATOS
+
+    nb_runs = 2
+    max_eval = 300000
+    population_size = 100
+
+    for i in range(len(names_array)):
+
+        print("Procesando " + names_array[i] + " dataset")
+
+        data_set = datasets_array[i]
+        labels = labels_array[i]
+        nb_clust = len(set(labels))
+
+        const = const_array[0][i]
+        ml_const, cl_const = get_const_list(const)
+
+        mean_ars = 0
+        mean_execution_time = 0
+        mean_unsat_percent = 0
+
+        for j in range(nb_runs):
+
+            shade = SHADE(data_set, ml_const, cl_const, population_size, nb_clust, 0.25, True)
+            de_assignment, mean_execution_time = shade.run(max_eval)[1:3]
+            mean_ars += adjusted_rand_score(labels, de_assignment)
+            mean_unsat_percent += get_usat_percent(ml_const, cl_const, de_assignment)
+
+        mean_ars /= nb_runs
+        mean_execution_time /= nb_runs
+        mean_unsat_percent /= nb_runs
+
+        results_matrix[i, 0:3] = [mean_ars, mean_unsat_percent, mean_execution_time]
+
+        mean_ars = 0
+        mean_execution_time = 0
+        mean_unsat_percent = 0
+
+        for j in range(nb_runs):
+
+            brkga = BRKGA(data_set, ml_const, cl_const, population_size, 0.2, 0.2, 0.5, nb_clust, 10)
+            brkga_assignment, mean_execution_time = brkga.run(max_eval, True)[1:3]
+            mean_ars += adjusted_rand_score(labels, brkga_assignment)
[…]

### 2018-12-17 Software/SHADE/SHADE.py

@@ -1,6 +1,8 @@
 import numpy as np
 import copy as cp
 import scipy.stats
+import random
+import time
 
 class SHADE:
 
@@ -17,7 +19,7 @@ class SHADE:
         self._h_record_f = np.empty((0, 0))
         self._ls = ls
         self._prt_elite = prt_elite
-
+        #self._ls_max_neighbors =
 
     def init_population(self):
 
@@ -35,6 +37,7 @@ class SHADE:
 
         decoded = np.ceil(cromosome * self._result_nb_clust)
         decoded[decoded == 0] = 1
+
         return decoded - 1
 
 
@@ -115,9 +118,12 @@ class SHADE:
         generations = 0
         h_index = 0
         external_archive_size = 0
+        gen_times = []
 
         # Mientras no se haya alcanzado el numero maximo de evaluaciones
         while self._evals_done < max_eval:
+
+            start_gen = time.time()
 
             #Ordenamos la poblacion segun el valor fitness de cada individuo
             self._population_fitness = self.get_fitness()[0]
@@ -133,8 +139,6 @@ class SHADE:
             if self._ls:
 
                 self.local_search()
-
-            #print(str(generations) + " " + str(self._best_fitness))
 
             #Inicializamos estructuras de almacenamiento auxiliares
             s_cr = np.empty((0, 0))
@@ -160,12 +164,15 @@ class SHADE:
                 x_pbest = self._population[np.random.randint(0, int(self._population_size * p_i))]
 
                 r1_index = np.random.randint(0, self._population_size)
+
                 while r1_index == j: r1_index = np.random.randint(0, self._population_size)
 
                 x_r1 = self._population[r1_index]
 
                 r2_index = np.random.randint(0, self._population_size + external_archive_size)
+
                 while r2_index == j or r2_index == r1_index:
+
                     r2_index = np.random.randint(0, self._population_size + external_archive_size)
 
                 if r2_index < self._population_size:
@@ -176,15 +183,9 @@ class SHADE:
 
                     x_r2 = self._external_archive[r2_index - self._population_size]
 
-                #Generamos el mutante segun el operador current-to-best/1
+                #Generamos el mutante segun el operador current-to-pbest/1
                 mutant = x_i + f_i * (x_pbest - x_i) + f_i * (x_r1 - x_r2)
                 mutant = np.clip(mutant, 0, 1)
-
-                #Corregimos los valores fuera el rango [0,1] del mutante
[…]

### 2018-12-17 Software/SHADEMainParallel.py

@@ -39,15 +39,15 @@ def apply(data_set, p_size, nb_clust, max_eval, nb_runs, name, label_percent, la
 
 def main():
 
-    np.random.seed(43)
+    np.random.seed(11)
 
     #CARGA DE DATOS
 
     names_array, datasets_array, labels_array = load_all_datasets()
 
-    names_array = names_array[:4]
-    datasets_array = datasets_array[:4]
-    labels_array = labels_array[:4]
+    names_array = names_array[0:5]
+    datasets_array = datasets_array[0:5]
+    labels_array = labels_array[0:5]
 
     #BUCLE DE OBTENCION DE DATOS
 
@@ -60,7 +60,7 @@ def main():
     general_start = time.time()
     results_matrix = np.zeros((len(names_array), len(const_percent_vector)))
 
-    with Parallel(n_jobs=multiprocessing.cpu_count() - 1) as parallel:
+    with Parallel(n_jobs=4) as parallel:
 
         for label_percent in range(len(const_percent_vector)):
 

### 2018-12-17 Software/functions.py

@@ -96,7 +96,7 @@ def load_all_datasets():
     # CARGA DE DATOS
 
     iris = datasets.load_iris()
-    iris_set = iris.data[:, :2]
+    iris_set = iris.data#[:, :2]
     iris_labels = iris.target
     datasets_array.append(iris_set)
     labels_array.append(iris_labels)
@@ -116,12 +116,6 @@ def load_all_datasets():
     labels_array.append(soybean_labels)
     names_array.append("soybean")
 
-    # breastA_set = np.transpose(np.loadtxt("Datasets/Breast_A.gct.txt", skiprows=3, usecols=np.array(range(2, 100))))
-    # breastA_labels = np.asarray(np.loadtxt("Datasets/Breast_A.cls", skiprows=2))
-    # datasets_array.append(breastA_set)
-    # labels_array.append(breastA_labels)
-    # names_array.append("breastA")
-
     pima = np.loadtxt("Datasets/pima.dat", skiprows=13, delimiter=',', dtype=str)
     pima_labels = pima[:, 8]
     pima_set = np.asarray(pima[:, :8].astype(float))
@@ -136,6 +130,27 @@ def load_all_datasets():
     labels_array.append(balance_labels)
     names_array.append("balance")
 
+    boston = datasets.load_boston()
+    boston_labels = boston.target
+    boston_set = boston.data
+    datasets_array.append(boston_set)
+    labels_array.append(boston_labels)
+    names_array.append("boston")
+
+    diabetes = datasets.load_diabetes()
+    diabetes_labels = diabetes.target
+    diabetes_set = diabetes.data
+    datasets_array.append(diabetes_set)
+    labels_array.append(diabetes_labels)
+    names_array.append("diabetes")
+
+    breast_cancer = datasets.load_breast_cancer()
+    breast_cancer_labels = breast_cancer.target
+    breast_cancer_set = breast_cancer.data
+    datasets_array.append(breast_cancer_set)
+    labels_array.append(breast_cancer_labels)
+    names_array.append("breast_cancer")
+
     bupa = np.loadtxt("Datasets/bupa.dat", skiprows=11, delimiter=',', dtype=str)
     bupa_labels = bupa[:, 6]
     bupa_set = np.asarray(bupa[:, :6].astype(float))
@@ -143,12 +158,12 @@ def load_all_datasets():
     labels_array.append(bupa_labels)
     names_array.append("bupa")
 
[…]

### 2018-12-17 Software/prueba0.py

@@ -1,89 +1,0 @@
-import numpy as np
-from BRKGA.BRKGA import BRKGA
-from SHADE.SHADE import SHADE
-from functions import *
-from sklearn.metrics import adjusted_rand_score
-import time
-
-
-def main():
-
-    np.random.seed(43)
-
-    #CARGA DE DATOS
-
-    names_array, datasets_array, labels_array = load_all_datasets()
-
-    names_array = names_array[0:1]
-    datasets_array = datasets_array[0:1]
-    labels_array = labels_array[0:1]
-
-    const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    const_percent_vector = [0.1, 0.2]
-
-    const_array = load_constraints(names_array, const_percent_vector)
-
-    file = open("max_eval_comp0.txt", "w+")
-
-    #BUCLE DE OBTENCION DE DATOS
-
-    population_size = 100
-    general_start = time.time()
-
-    name = names_array[0]
-    data_set = datasets_array[0]
-    labels = labels_array[0]
-
-    nb_clust = len(set(labels))
-
-    for i in range(2, 11):
-
-        max_eval = i * 100000
-
-        print("------------ Probando con " + str(max_eval) + " evaluaciones ------------")
-        file.write("------------ Probando con " + str(max_eval) + " evaluaciones ------------\n")
-
-        for j in range(len(const_array)):
-
-            for k in range(len(const_array[j])):
-
-                print("------------ Procesando " + name + " Dataset ------------ (" + str(const_percent_vector[j]) + ")")
-                const = const_array[j][k]
-                ml_const, cl_const = get_const_list(const)
-
-                print("Procesando en porcentaje de restricciones " + str(const_percent_vector[j]))
-
-                file.write(
-                    "------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[j]) + " ------------\n")
-                file.write("Dataset RandIndex   Time(s)   Unsat(%)\n")
-
-                print("Procesando con BRKGA")
-                brkga = BRKGA(data_set, ml_const, cl_const, population_size, 0.2, 0.2, 0.5, nb_clust, 10)
-                start = time.time()
-                brkga_assignment = brkga.run(max_eval, True)[1]
-                end = time.time()
-                brkga_ars = adjusted_rand_score(labels, brkga_assignment)
-                brkga_execution_time = end - start
-                brkga_unsat_percent = get_usat_percent(ml_const, cl_const, brkga_assignment)
-
-                print("Procesando con SHADE")
[…]

### 2018-12-17 Software/prueba1.py

@@ -1,89 +1,0 @@
-import numpy as np
-from BRKGA.BRKGA import BRKGA
-from SHADE.SHADE import SHADE
-from functions import *
-from sklearn.metrics import adjusted_rand_score
-import time
-
-
-def main():
-
-    np.random.seed(43)
-
-    #CARGA DE DATOS
-
-    names_array, datasets_array, labels_array = load_all_datasets()
-
-    names_array = names_array[1:2]
-    datasets_array = datasets_array[1:2]
-    labels_array = labels_array[1:2]
-
-    const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    const_percent_vector = [0.1, 0.2]
-
-    const_array = load_constraints(names_array, const_percent_vector)
-
-    file = open("max_eval_comp1.txt", "w+")
-
-    #BUCLE DE OBTENCION DE DATOS
-
-    population_size = 100
-    general_start = time.time()
-
-    name = names_array[0]
-    data_set = datasets_array[0]
-    labels = labels_array[0]
-
-    nb_clust = len(set(labels))
-
-    for i in range(2, 11):
-
-        max_eval = i * 100000
-
-        print("------------ Probando con " + str(max_eval) + " evaluaciones ------------")
-        file.write("------------ Probando con " + str(max_eval) + " evaluaciones ------------\n")
-
-        for j in range(len(const_array)):
-
-            for k in range(len(const_array[j])):
-
-                print("------------ Procesando " + name + " Dataset ------------ (" + str(const_percent_vector[j]) + ")")
-                const = const_array[j][k]
-                ml_const, cl_const = get_const_list(const)
-
-                print("Procesando en porcentaje de restricciones " + str(const_percent_vector[j]))
-
-                file.write(
-                    "------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[j]) + " ------------\n")
-                file.write("Dataset RandIndex   Time(s)   Unsat(%)\n")
-
-                print("Procesando con BRKGA")
-                brkga = BRKGA(data_set, ml_const, cl_const, population_size, 0.2, 0.2, 0.5, nb_clust, 10)
-                start = time.time()
-                brkga_assignment = brkga.run(max_eval, True)[1]
-                end = time.time()
-                brkga_ars = adjusted_rand_score(labels, brkga_assignment)
-                brkga_execution_time = end - start
-                brkga_unsat_percent = get_usat_percent(ml_const, cl_const, brkga_assignment)
-
-                print("Procesando con SHADE")
[…]

### 2018-12-17 Software/prueba2.py

@@ -1,89 +1,0 @@
-import numpy as np
-from BRKGA.BRKGA import BRKGA
-from SHADE.SHADE import SHADE
-from functions import *
-from sklearn.metrics import adjusted_rand_score
-import time
-
-
-def main():
-
-    np.random.seed(43)
-
-    #CARGA DE DATOS
-
-    names_array, datasets_array, labels_array = load_all_datasets()
-
-    names_array = names_array[2:3]
-    datasets_array = datasets_array[2:3]
-    labels_array = labels_array[2:3]
-
-    const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    const_percent_vector = [0.1, 0.2]
-
-    const_array = load_constraints(names_array, const_percent_vector)
-
-    file = open("max_eval_comp2.txt", "w+")
-
-    #BUCLE DE OBTENCION DE DATOS
-
-    population_size = 100
-    general_start = time.time()
-
-    name = names_array[0]
-    data_set = datasets_array[0]
-    labels = labels_array[0]
-
-    nb_clust = len(set(labels))
-
-    for i in range(2, 11):
-
-        max_eval = i * 100000
-
-        print("------------ Probando con " + str(max_eval) + " evaluaciones ------------")
-        file.write("------------ Probando con " + str(max_eval) + " evaluaciones ------------\n")
-
-        for j in range(len(const_array)):
-
-            for k in range(len(const_array[j])):
-
-                print("------------ Procesando " + name + " Dataset ------------ (" + str(const_percent_vector[j]) + ")")
-                const = const_array[j][k]
-                ml_const, cl_const = get_const_list(const)
-
-                print("Procesando en porcentaje de restricciones " + str(const_percent_vector[j]))
-
-                file.write(
-                    "------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[j]) + " ------------\n")
-                file.write("Dataset RandIndex   Time(s)   Unsat(%)\n")
-
-                print("Procesando con BRKGA")
-                brkga = BRKGA(data_set, ml_const, cl_const, population_size, 0.2, 0.2, 0.5, nb_clust, 10)
-                start = time.time()
-                brkga_assignment = brkga.run(max_eval, True)[1]
-                end = time.time()
-                brkga_ars = adjusted_rand_score(labels, brkga_assignment)
-                brkga_execution_time = end - start
-                brkga_unsat_percent = get_usat_percent(ml_const, cl_const, brkga_assignment)
-
-                print("Procesando con SHADE")
[…]

### 2018-12-17 Software/prueba3.py

@@ -1,89 +1,0 @@
-import numpy as np
-from BRKGA.BRKGA import BRKGA
-from SHADE.SHADE import SHADE
-from functions import *
-from sklearn.metrics import adjusted_rand_score
-import time
-
-
-def main():
-
-    np.random.seed(43)
-
-    #CARGA DE DATOS
-
-    names_array, datasets_array, labels_array = load_all_datasets()
-
-    names_array = names_array[3:4]
-    datasets_array = datasets_array[3:4]
-    labels_array = labels_array[3:4]
-
-    const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    const_percent_vector = [0.1, 0.2]
-
-    const_array = load_constraints(names_array, const_percent_vector)
-
-    file = open("max_eval_comp3.txt", "w+")
-
-    #BUCLE DE OBTENCION DE DATOS
-
-    population_size = 100
-    general_start = time.time()
-
-    name = names_array[0]
-    data_set = datasets_array[0]
-    labels = labels_array[0]
-
-    nb_clust = len(set(labels))
-
-    for i in range(2, 11):
-
-        max_eval = i * 100000
-
-        print("------------ Probando con " + str(max_eval) + " evaluaciones ------------")
-        file.write("------------ Probando con " + str(max_eval) + " evaluaciones ------------\n")
-
-        for j in range(len(const_array)):
-
-            for k in range(len(const_array[j])):
-
-                print("------------ Procesando " + name + " Dataset ------------ (" + str(const_percent_vector[j]) + ")")
-                const = const_array[j][k]
-                ml_const, cl_const = get_const_list(const)
-
-                print("Procesando en porcentaje de restricciones " + str(const_percent_vector[j]))
-
-                file.write(
-                    "------------ Procesando en porcentaje de restricciones: " + str(const_percent_vector[j]) + " ------------\n")
-                file.write("Dataset RandIndex   Time(s)   Unsat(%)\n")
-
-                print("Procesando con BRKGA")
-                brkga = BRKGA(data_set, ml_const, cl_const, population_size, 0.2, 0.2, 0.5, nb_clust, 10)
-                start = time.time()
-                brkga_assignment = brkga.run(max_eval, True)[1]
-                end = time.time()
-                brkga_ars = adjusted_rand_score(labels, brkga_assignment)
-                brkga_execution_time = end - start
-                brkga_unsat_percent = get_usat_percent(ml_const, cl_const, brkga_assignment)
-
-                print("Procesando con SHADE")
[…]

### 2018-12-17 Software/tablemaker.py

@@ -6,13 +6,14 @@ def main():
 
 	names_array, datasets_array, labels_array = load_all_datasets()
 	#names_array = names_array[0:4]
+	print(len(names_array))
 
 	for i in range(len(names_array)):
 
 		n = np.shape(datasets_array[i])[0]
 		nb_class = len(set(labels_array[i]))
 
-		print(str(n) + " " + names_array[i].title() + " & " + str(n) + " & " + str(nb_class) + " &  \\\\")
+		print(names_array[i].title() + " & " + str(n) + " & " + str(nb_class) + " &  \\\\")
 
 
 	const_percent_vector = [0.05, 0.1, 0.15, 0.2]
@@ -20,7 +21,7 @@ def main():
 
 	for i in range(len(names_array)):
 
-		print(names_array[i].title(), end='')
+		print(names_array[i].title() + " & ", end='')
 
 		for l in range(len(const_percent_vector)):
 
@@ -32,7 +33,10 @@ def main():
 			ml = (d[1] - np.shape(c)[0]) / 2
 			cl = d[-1] / 2
 
-			print(" & %d & %d & %d" % (total, ml, cl), end='')
+			if l < len(const_percent_vector) - 1:
+				print("%d & %d && " % (ml, cl), end='')
+			else:
+				print("%d & %d" % (ml, cl), end='')
 
 		print(" \\\\")
 

### 2018-12-19 Papers/SHADE/paperSHADE.tex
Actualizada la memoria

@@
 \usepackage{chngpage}
+\usepackage{setspace}
+\usepackage[ruled, lined, onelanguage, linesnumbered]{algorithm2e}
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\newcommand{\myfloatalign}{\centering}
 \modulolinenumbers[5]
@@
 
+\author[mymainaddress]{Another Juan}
+
 \address[mymainaddress]{Department of Computer Science and Artificial Intelligence, University of Granada, 18071 Granada, Spain}
@@
 
+\textcolor{red}{Citar a los de las hormigas}\\
+\textcolor{red}{Poner un parrafo de aplicaciones del clustering con restricciones}
+
 \section{Background}
 
-\subsection{Constrained Clustering}
+%\subsection{Constrained Clustering}
 
@@
 
-On the other hand, Cannot-link constraints---from now on CL---do not constitute an equivalence relation***, but they also can be entailed***.
+On the other hand, Cannot-link constraints---from now on CL---do not constitute an equivalence relation.
 However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then it is also true that $c_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
@@
 Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the problem complexity.
-The feasibility problem for constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.
+The feasibility problem for non hierarchical constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.
 
@@
 Does there exists a partition $C$ of $X$ with $k$ clusters such that all constraints in $R$ are satisfied?
-\cite{davidson2005clustering}
+\cite{davidson2007survey}\cite{davidson2005clustering}
 	\label{def1}
@@
 	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{0.9}
+	\renewcommand{\arraystretch}{1.2}
 	%\resizebox{\textwidth}{!}{
@@
 These complexity results show that feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too.
-For more details on constrained clustering problem complexity see \cite{davidson2005clustering}.
+For more details on constrained clustering complexity see \cite{davidson2005clustering}.
+
+\textcolor{red}{El parrafo de abajo esta bien?}
 
 Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why applying heuristics-guided and population-based algorithms is a good approach to find quality solutions to the constrained clustering problem.
+That is the reason why applying heuristics-guided and population-based algorithms constitute a good approach to find quality solutions to the constrained clustering problem.
 
@@
 	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{0.9}
+	\renewcommand{\arraystretch}{1.2}
 	\resizebox{\textwidth}{!}{
@@
 
-The mutation strategy used by SHADE is known as current-to-pbest/1, and the expression that generates new individuals is described in Equation \eqref{eq3}.
+The mutation strategy used by SHADE is known as current-to-$p$best/1, and the expression that generates new individuals is described in Equation \eqref{eq3}.
 
 \begin{equation}
-m_{i,G} = p_{i,G} + F_i * (p_{pbest, G} - p_{i,G}) + F_i * (p_{r1, G} - p_{r2,G})
+m_{[i,G]} = p_{[i,G]} + F_i * (p_{[pbest, G]} - p_{[i,G]}) + F_i * (p_{[r1, G]} - p_{[r2,G]})
 \label{eq3}
@@
 
-Where $m_{i,G}$ is the mutant vector, which is generated with an individual $p_{i,G}$ from the population [serving as a starting point].
+Where $m_{[i,G]}$ is the mutant vector, which is generated with an individual $p_{[i,G]}$ from the population [serving as a starting point].
 Indices $r1$ and $r2$ are random values in the range $[0,N]$, different from $i$ and from each other.
-The individual $p_{pbest, G}$ is randomly selected from among the best $N \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
+The individual $p_{[pbest, G]}$ is randomly selected from among the best $N \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
 This way, the parameter $p$ controls the greediness of the mutation strategy.
@@
 
-After generating the mutant vector $m_{i,G}$, it is combined with the parent $p_{i,G}$ vector by means of a crossover operator; the result is the trial vector.
+After generating the mutant vector $m_{[i,G]}$, it is combined with the parent $p_{[i,G]}$ vector by means of a crossover operator; the result is the trial vector.
 SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.
@@
 \begin{equation}
-t_{j,i,G} = \left\{ \begin{array}{lc}
-m_{j,i,G} &   if \;\; rand[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
-p_{j,i,G} &  otherwise
+t_{[j,i,G]} = \left\{ \begin{array}{lc}
+m_{[j,i,G]} &   if \;\; \text{rand}[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
+p_{[j,i,G]} &  \text{otherwise}
 \end{array}
@@
 The SHADE replacement operator determines the survivors individuals for the next generation.
-It compares each parent $x_{i,G}$ with the trial vector $_{j,i,G}$, generated based on it.
+It compares each parent $x_{[i,G]}$ with the trial vector $_{[j,i,G]}$, generated based on it.
 The best one of these two individuals will survive for the next generation.
@@
 \begin{equation}
-p_{i,G + 1} = \left\{ \begin{array}{lc}
-t_{i,G} &   if \;\; f(t_{i,G}) \le f(p_{i,G}) \\
-p_{i,G} &  otherwise
+p_{[i,G + 1]} = \left\{ \begin{array}{lc}
+t_{[i,G]} &   if \;\; f(t_{[i,G]}) \le f(p_{[i,G]}) \\
+p_{[i,G]} &  \text{otherwise}
 \end{array}
@@
 
-To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{i,G}$ who were replaced by their associated trial vector $t_{i,G}$ are saved.
-To make use of the archive, we will consider that the individual $p_{r2,G}$ from Equation \eqref{eq3} is selected from $P \cup A$.
+To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{[i,G]}$ who were replaced by their associated trial vector $t_{[i,G]}$ are saved.
+To make use of the archive, we will consider that the individual $p_{[r2,G]}$ from Equation \eqref{eq3} is selected from $P \cup A$.
 The archive size is the same as the population size.
@@
 		 \hline
-		 $M_{CR}$ & $M_{CR,1}$ & $M_{CR,1}$ & $\cdots$ & $M_{CR,H-1}$ & $M_{CR,H}$ \\
+		 $M_{CR}$ & $M_{[CR,1]}$ & $M_{[CR,1]}$ & $\cdots$ & $M_{[CR,H-1]}$ & $M_{[CR,H]}$ \\
 		 \hline
-		 $M_{F}$ & $M_{F,1}$ & $M_{F,1}$ & $\cdots$ & $M_{F,H-1}$ & $M_{F,H}$ \\
+		 $M_{F}$ & $M_{[F,1]}$ & $M_{[F,1]}$ & $\cdots$ & $M_{[F,H-1]}$ & $M_{[F,H]}$ \\
 		\hline
@@
 \begin{equation}
-CR_i = randn_i(M_{CR,r_i}, 0.1)
+CR_i = \text{randn}_i(M_{[CR,r_i]}, 0.1)
 \label{eq6}
@@
 \begin{equation}
-F_i = randc_i(M_{F,r_i}, 0.1)
+F_i = \text{randc}_i(M_{[F,r_i]}, 0.1)
 \label{eq7}
@@
 \begin{equation}
-M_{CR,h,G+1} = \left\{ \begin{array}{lc}
-mean_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
-M_{CR,h,G+1} &  otherwise
+M_{[CR,h,G+1]} = \left\{ \begin{array}{lc}
+\text{mean}_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
+M_{[CR,h,G+1]} &  otherwise
 \end{array}
@@
 \begin{equation}
-M_{F,h,G+1} = \left\{ \begin{array}{lc}
-mean_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
-M_{F,h,G+1} &  otherwise
+M_{[F,h,G+1]} = \left\{ \begin{array}{lc}
+\text{mean}_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
+M_{[F,h,G+1]} &  otherwise
 \end{array}
@@
 \begin{equation}
-mean_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{RC,i}
+\text{mean}_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{[RC,i]}
 \label{eq10}
@@
 
-where $\Delta f_i = |f(t_i,G) - f(p_i, G)|$, which is the amount of improvement obtained from the trial vector with respect to the parent.
+where $\Delta f_i = |f(t_{[i,G]}) - f(p_{[i, G]})|$, which is the amount of improvement obtained from the trial vector with respect to the parent.
 
@@
 \begin{equation}
-mean_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{F,i}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{F,i}}
+\text{mean}_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{[F,i]}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{[F,i]}}
 \label{eq12}
@@
 
-Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the qreediness of the mutation strategy, is not included in the optimization process.
+Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the greediness of the mutation strategy, is not included in the adaptive optimization process.
 However, it is not static: it is calculated for each individual $p_i$ of the population following Equation \eqref{eq13}.
@@
 \begin{equation}
-pbest_i = randn[2/N, 0.2]
+pbest_i = \text{randn}[2/N, 0.2]
 \label{eq13}
@@
 
-Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated from the dataset or constraints, whereas in \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.
+Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated from the dataset or constraints, whereas in Equation \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.
 
@@
 
-\textcolor{red}{Poner una pintura que ejemplifique lo anterior.}\\
-\textcolor{red}{Explicar la busqueda local. Con pseudocódigo?}
+To put those concepts clear we consider a toy dataset and three partitions over it.
+Dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
+Partitions $C_0$ and $C_1$ do not violate the constraint ($infeasibility = 0$), while $C_2$ does violate it ($infeasibility = 1$).
+
+\begin{figure}[bth]
+	\myfloatalign
+	{\includegraphics[width=.45\linewidth]{Figures/Dataset}} \quad
+	%\subfloat[Imagen original]
+	{\includegraphics[width=.45\linewidth]{Figures/C0}} \quad
+	%\subfloat[Clustering sin restricciones]
+	{\includegraphics[width=.45\linewidth]{Figures/C1}} \quad
+	%\subfloat[Clustering con restricciones]
+	{\includegraphics[width=.45\linewidth]{Figures/C2}} \quad
+	\caption{Three partitions over a toy dataset.}
+	\label{img:toydatasets}
+\end{figure}
+
+Table\ref{tab:fitnessfunctions} shows fitness functions values for each partition from Figure \ref{img:toydatasets}.
+We call $f_1$ to the function presented in Equation \eqref{eq1} and $f_2$ to the function presented in Equation \eqref{eq14}.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.3}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{c c c c c c}
+			\hline
+			\multirow{2}{*}{Partition} &
+			\multicolumn{2}{c}{Expression} &&
+			\multicolumn{2}{c}{Value} \\
+			\cline{2-3} \cline{5-6}
+			& $f_1$ & $f_2$ && $f_1$ & $f_2$ \\
+			\hline
+			$C_0$ & $z_0$ & $z_0$ && $0.457$  & $0.457$  \\
+			$C_1$ & $z_1$ & $z_1$  && $0.540$  & $0.540$  \\
+			$C_2$ & $z_2 + (\mu * N * infs)$ & $z_2 * (infs + 1)$  && $0.503 + 1000 = 1000.503$ &  $0.503 * 2 = 1.005$ \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Expression and value of fitness functions over three partitions. ($\mu = 10$)}
+	\label{tab:fitnessfunctions}
+\end{table}
+
+Results in Table \ref{tab:fitnessfunctions} are a clear example of the above.
+In the general case, with $f_2$ the penalty for violating constraints is proportional to the within-cluster-sum-of-squares, the lower it is the lower the penalty.
+This is not the case with $f_1$, whose penalty term is independent of the within-cluster-sum-of-squares, which can result in a difference of several orders of magnitude between partitions satisfying different number of constraints.
+
+As in \cite{de2017comparison} we apply a local optimization procedure to the individuals of the population, without transferring its results to the individual to maintain diversity.
+The difference is that we do not apply it to the whole population but only to a number $p_e$ of individuals considered as the elite of the population.
+The percentage of individuals considered as elite must be determined for each case.
+
+Algorithms \ref{alg:SHADE} shows overall SHADE optimization process and is an adaptation of the one that can be found on \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{Sort}{Sort}
+	\SetKwFunction{LocalSearch}{LocalSearch}
+	\KwIn{Dataset $X$, constraints sets $c_=$ and $c_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $K$.}
+	\tcp{Initialization phase}
+	$G \leftarrow 0$\\
+	Initialize population $P_0 \leftarrow \{p_{[1,0]}, \cdots, p_{[p_{size},0]}\}$\\
+	Initialize all values in $M_{CR}$, $M_F$ to 0.5\\
+	$A \leftarrow \emptyset$; $h \leftarrow 1$\\
+	\tcp{Main loop}
+	\While{Termination criteria are not met}{
+
+		$S_{CR} \leftarrow S_F \leftarrow \emptyset$\\
+		$P_G \leftarrow$ \Sort{$P_G$} \tcp{Sort the population}
+		\tcp{Apply a LS procedure to the elite of the population}
+		\LocalSearch{$\{p_{[1,G]}, \cdots, p_{[p_{e},G]}\}$}\\
+		\For{$i \in [1,p_{size}]$}{
+			$r_i \leftarrow$ randInt$[1,H]$\\
+			$CR_{[i,G]} \leftarrow$ randn$_i(M_{[CR,r_i]}, 0.1)$\\
+			$F_{[i,G]} \leftarrow$ randc$_i(M_{[F,r_i]}, 0.1)$\\
+			$p_{[i,G]} \leftarrow$ rand$[N/2, 0.2]$\\
+			Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin\\
+
+			\eIf{$f(t_{[i,G]}) \le f(p_{[i,G]})$}{
+				$p_{[i,G + 1]} \leftarrow t_{[i,G]}$
+			}{
+				$p_{[i,G + 1]} \leftarrow p_{[i,G]}$
+			}
+
+			\If{$f(t_{[i,G]}) < f(p_{[i,G]})$}{
+				$p_{[i,G]} \rightarrow A$;
+				$CR_{[i,G]} \rightarrow S_{CR}$;
+				$F_{[i,G]} \rightarrow S_{F}$
+			}
+
+		}
+		Whenever $|A| > |P|$, randomly select an individual from $A$ to be deleted so that $|A| \le |P|$\\
+		\If{$S_{CR} \neq \emptyset$ \textbf{and} $S_{F} \neq \emptyset$}{
+			Update $M_{[CR,h]}$ and $M_{[F,h]}$ based on $S_{CR}$ and $S_{F}$\\
+			$h \leftarrow (h + 1) \mod H$
+		}
+		$G++$
+	}
+	\caption{Modified SHADE}\label{alg:SHADE}
+\end{algorithm}
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{RandomShuffle}{RandomShuffle}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Dataset $X$, constraints sets $c_=$ and $c_{\neq}$, decoded random-key vector (solution) $S$, number of clusters $K$.}
+	\BlankLine
+	\While{$improvement$}{
+		$improvement \leftarrow$ \texttt{false} \\
+
+		$s_i \leftarrow $ Select random object from $S$\\
+		\tcp{Random shuffle labels set}
+		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
+		\For{$c \in RSL$}{
+			$S^\prime \leftarrow S$\\
+			$S^\prime$\texttt{[}$s_i$\texttt{]} $\leftarrow c$ \tcp{Move object $s_i$ to cluster $c$}
+
+			\If{$f(S^\prime) < f(S)$}{
+				$S \leftarrow S^\prime$\\
+				$improvement \leftarrow$ \texttt{true} \\
+			}
+		}	
+	}
+	\BlankLine
+	\KwRet ($S$)
+
+\caption{Local Search}\label{alg:LS}
+\end{algorithm}
 
@@
 
-\subsection{Dataset}
+For our experiments we will compare the results obtained by BRKGA and SHADE over XX datasets.
+Most of the datasets used can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
+\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
+Table \ref{tab:datasets} present a summary the main features of all datasets used.
+\textcolor{red}{Esta bien citar tambien al sklearn?
+Que paper de keel cito, 2017 o 2011?}
 
@@
 	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{5pt}
-	\renewcommand{\arraystretch}{0.8}
+	%\setlength{\tabcolsep}{5pt}
+	%\renewcommand{\arraystretch}{1.2}
 	%\resizebox{\textwidth}{!}{
@@
 	\end{tabular}%}
-	\caption{Summary of datasets used for the experiments \textcolor{red}{Citar keel}}
+	\caption{Summary of datasets used for the experiments \cite{triguero2017keel}\cite{scikit-learn}}
 	\label{tab:datasets}
@@
 
+\clearpage
+
 \subsection{Constraints Generation}
 
-Since we have the true labels associated with each data set, we will use the method proposed by Wagstaff and Cardie to generate artificial constraint sets \cite{wagstaff2001constrained}.
+Since we have the true labels associated with each data set, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraint sets.
 This method consists in randomly selecting two instances of a data set, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
@@
 Each constraint set is associated with a different percentage of labeled data, namely: 5\%, 10\%, 15\% and 20\%.
-It is worth noting that constraints are not the complete constraint graph made up of the tagged instances, this graph only sets the number of constraints to be created, this way there is a lower probability of biasing the set of constraints so that there are classes with poor representation in it.
+It is worth noting that constraints are not the complete constraint graph made up of the labeled instances, this graph only sets the number of constraints to be created, this way there is a lower probability of biasing the set of constraints so that there are classes with poor representation in it.
+Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.
 
@@
 	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{0.8}
+	\renewcommand{\arraystretch}{1.2}
 	%\begin{adjustwidth}{-1in}{-1in}
@@
 
+Note that the greater the number of classes present in the data set, the fewer ML restrictions obtained with the method proposed by \cite{wagstaff2001constrained}.
+This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the data set increases.
+
+\clearpage
+
 \subsection{Evaluation Method}
 
-We use the evaluation method proposed by Wagstaff et al. (2001) \cite{wagstaff2001constrained}, the authors of the Constrained K-Means algorithm, proposed in their work to evaluate the results.
-
 Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
-We will use Rand Index \cite{rand1971objective} to measure the accuracy of the predictions resulting from each method we test.
-Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
-$C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions.
+We will use Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.Basic Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
+$C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions \cite{rand1971objective}.
 
@@
 
-We will use this measure co compute the accuracy for all of our experiments.
+The Adjusted Rand Index is a corrected-for-chance version of Rand Index.
+This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline.
+The Adjusted Rand Index is computed as in Equation \eqref{eq16}.
+
+\begin{equation}
+ARI(C_1, C_2) = \frac{Rand(C_1, C_2) - ExpectedIndex}{MaximumIndex - ExpectedIndex}
+\label{eq16}
+\end{equation}
+
+where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
+It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that a $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing the obtained partitions with random partitions.
+To summarize, the higher the $ARI$, the greater the degree of similarity between $C_1$ and $C_2$.
+For more details on Adjusted Rand Index see \cite{hubert1985comparing}.
 
 \subsection{Calibration}
+
+\clearpage
 
@@
 	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{0.9}
+	\renewcommand{\arraystretch}{1.4}
 	%\begin{adjustwidth}{-1in}{-1in}

### 2018-12-21 Papers/SHADE/paperSHADE.tex
Nuevas secciones en background

@@
 \usepackage{setspace}
+\usepackage{amsfonts}
 \usepackage[ruled, lined, onelanguage, linesnumbered]{algorithm2e}
@@
 
+\subsection{Short insertion to Differential Evolution}
+
+To introduce Differential Evolution--from now on DE--we take as basis the review of the state-of-the-art performed by \cite{das2011differential}, where you can find a very detailed analysis on DE an its variants.
+
+The first article written on differential evolution emerged in 1995 as a technical report written by \cite{noman2008differential}.
+From that moment on, DE's reputation was manifested in congresses and competitions where it was able to obtain competitive results that rivaled the state of the art of that moment.
+Specifically DE excelled in real optimization problems.
+
+DE arises as a variant of population-based evolutionary algorithms, its goal is to perform an intelligent search in the problem's solutions space in the most optimized and effective possible way.
+
+Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{i,1},\cdots,p_{i,D}\}$ that serve as parameters for the function to be optimized.
+These individuals are considered solutions to the problem.
+
+To guide the optimization process, we use an objective function.
+DE's task is to seek the parameter vector $p_i$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \Omega$ with omega a nonempty finite set that is the search domain.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.4]{Figures/DEop.png}
+	\caption{DE optimization process \cite{das2011differential}.}\label{img:DE}
+\end{figure}
+
+DE optimization process works through a simple cycle of stages, presented in Figure \ref{img:DE} that we can summarize stages in as follow:
+
+\begin{itemize}
+
+	\item \textbf{Initialization of Vectors}: DE begins with a randomly initiated populations of $N$ $D$-dimensionals arrays of parameters.
+These vectors constitute a candidate solution for the problem to solve.
+
+	\item \textbf{Difference-vector based mutation}: DE introduces diversity on the population via mutation operator.
+This operator is applied to each individual of the population, known ins this context as parent vector.
+It combines the parent vector with two randomly selected vectors from the population by applying an arithmetic operator to them.
+The result of this process is a mutant vector.
+
+	\item \textbf{Crossover/Recombination}: The mutant vector exanges some components with its associated parent to from a trial vector, which will be the candidate to replace the parent vector.
+There exist several strategies to generate this trial vector, we will not delve into them but they can be found in \cite{das2011differential}.
+
+	\item \textbf{Selection}: The newly generated trial vector is compared with its associated parent using the objective function.
+If the trial vector is better than the parent, then it will replace the parent in the next generation.
+It is worth noting that, following this strategy, the population never degenerates, only get better or remains the same.
+
+\end{itemize}
+
+Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
+We will focus on the SHADE variant, which adaptively optimizes some of the DE parameters (see Section \ref{sec:SHADE}).
+
 \section{Brief Review of the BRKGA Algorithm}
@@
 
-\section{The SHADE Optimization Method}
+\section{The SHADE Optimization Method} \label{sec:SHADE}
 

### 2018-12-23 Papers/SHADE/paperSHADE.tex
Terminado cuerpo paper SHADE

@@
 
+\textcolor{red}{Revisar que todos los subindices en los que aparecan comas lleven corchetes. Revisar que las asignaciones en los algoritmo sean flechas y no =}
+
 \section{Introduction}
 
-\textcolor{red}{Citar a los de las hormigas}\\
-\textcolor{red}{Poner un parrafo de aplicaciones del clustering con restricciones}
+\textcolor{red}{Citar a los de las hormigas. Poner un parrafo de aplicaciones del clustering con restricciones}
 
@@
 
-\subsection{Short insertion to Differential Evolution}
+\subsection{Short Introduction to Differential Evolution}
 
-To introduce Differential Evolution--from now on DE--we take as basis the review of the state-of-the-art performed by \cite{das2011differential}, where you can find a very detailed analysis on DE an its variants.
+\textcolor{red}{Eliminar toda la notacion matematica de esta seccion?}
 
-The first article written on differential evolution emerged in 1995 as a technical report written by \cite{noman2008differential}.
-From that moment on, DE's reputation was manifested in congresses and competitions where it was able to obtain competitive results that rivaled the state of the art of that moment.
+To introduce Differential Evolution--from now on DE--we take as basis the review of the state-of-the-art performed by \cite{das2011differential}.
+You can also find in \cite{das2011differential} a very detailed analysis on DE an its variants.
+
+The first article written on DE emerged in 1995 as a technical report written by \cite{noman2008differential}.
+From that moment on, DE's reputation was manifested in congresses and competitions where it was able to obtain competitive results that rivaled the state-of-the-art of that moment.
 Specifically DE excelled in real optimization problems.
@@
 We will focus on the SHADE variant, which adaptively optimizes some of the DE parameters (see Section \ref{sec:SHADE}).
+
+\subsection{(Classic) Resolution Methods}
+
+\textcolor{red}{Clasicos clasicos no son}
+
+Constrained clustering has many applications and has been widely studied in the literature.
+The first adaptation of a classic clustering method for constrained clustering was proposed by \cite{wagstaff2001constrained}.
+It involved modifying the widely studied K-medias algorithm to take into account instance level constraints, the already known ML and CL.
+This method was named COP-kmeans.
+
+In \cite{antoine2012cecm} authors propose an adaptation of the evidential C-means algorithm, within the fuzzy clustering framework, for constrained clustering.
+The particularity of this resolution method is that the membership of instances to cluster is defined by a probabilistic belief function.
+This method redefines constraints from the point of view of belief functions and includes them in the cost function.
+
+In \cite{pelleg2007k} LCVQE, a modification of CVQE (Constrained Vector Cuantization Error), is proposed.
+The adaptation of CVQE for constrained clustering was proven to produce high quality results, at the cost of a very high computational complexity.
+LCVQE (Linear CVQE) introduces a modification on CVQE objective function to make it more intuitive and less computationally complex.
+The experimentation resulted in a dramatic improvement on clustering quality over both noisy and clean constraints sets.
+
+TVClust (Two Views Clustering) and RDP-means (Relation Dirichlet Process - Means) were prosed by \cite{khashabi2015clustering}.
+TVClust is able to incorporate the constraints to the clustering problem making a relaxed interpretation of them.
+The authors model the data set and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
+Using this model as basis authors derive the deterministic algorithm RDP-means.
+This method can ve viewed as an extension of K-means that include side information (constraints) and has the property that the number of clusters ($K$) does not need to be specified.
+
+\textcolor{red}{Decir que voy a comparar con ellos? Incluir mas métodos ``clasicos``?}
 
@@
 
+\textcolor{red}{En esta seccion podria simplemente decir el numero de restricciones que genero como un numero magico sin meterme en nada de grafos, pero igual es un poco burro.}
+
 Since we have the true labels associated with each data set, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraint sets.
@@
 Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
-We will use Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.Basic Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
+We will use Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
+Basic Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
 $C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions \cite{rand1971objective}.

### 2018-12-23 Software/EstimandoXi.py

@@ -1,0 +1,62 @@
+import numpy as np
+from ILSNueva.ILSNueva import ILSNueva
+from functions import *
+from sklearn.metrics import adjusted_rand_score
+import sys
+
+
+def main():
+
+    np.random.seed(11)
+
+    #CARGA DE DATOS
+
+    names_array, datasets_array, labels_array = load_all_datasets()
+
+    names_array = names_array[0:5]
+    datasets_array = datasets_array[0:5]
+    labels_array = labels_array[0:5]
+
+    const_percent_vector = [0.05, 0.1, 0.15, 0.2]
+    const_array = load_constraints(names_array, const_percent_vector)
+
+
+    #BUCLE DE OBTENCION DE DATOS
+
+    max_eval = 300000
+
+    for p in range(0, 6):
+
+        xi = p/10.0
+
+        results_matrix = np.zeros((len(names_array), 4))
+
+        for i in range(len(names_array)):
+
+            data_set = datasets_array[i]
+            labels = labels_array[i]
+            nb_clust = len(set(labels))
+
+            for j in range(len(const_array)):
+
+                const = const_array[j][i]
+                ml_const, cl_const = get_const_list(const)
+
+                ils = ILSNueva(data_set, ml_const, cl_const, nb_clust, 0.3, 300, 0.3, xi)
+                ils_assignment = ils.run(max_eval)[1]
+                ars = adjusted_rand_score(labels, ils_assignment)
+
+                results_matrix[i, j] = ars
+
+        print("---------------- Procesando con xi = " + str(xi) + " ----------------")
+        for i in range(results_matrix.shape[0]):
+
+            print(names_array[i].title(), end='')
+
+            for j in range(results_matrix.shape[1]):
+
+                print(" & %.3f" % (results_matrix[i, j]), end='')
+
+            print(" \\\\")
+
+if __name__ == "__main__": main()

### 2018-12-23 Software/ILSNueva/ILSNueva.py

@@ -4,7 +4,7 @@ import copy as cp
 
 class ILSNueva:
 
-    def __init__(self, data, ml_const, cl_const, nb_clust, segment_percent, ls_max_neighbors, pbt_inherit, restart_percent):
+    def __init__(self, data, ml_const, cl_const, nb_clust, segment_percent, ls_max_neighbors, pbt_inherit, xi):
 
         self._data = data
         self._ml = ml_const
@@ -15,7 +15,7 @@ class ILSNueva:
         self._segment_size = int(np.ceil(segment_percent * self._dim))
         self._max_neighbors = ls_max_neighbors
         self._pbt_inherit = pbt_inherit
-        self._restart_percent = restart_percent
+        self._xi = xi
 
     def init_ils(self):
 
@@ -179,8 +179,8 @@ class ILSNueva:
 
             worst = np.argmax(self._best_fitness)
             best = (worst + 1) % 2
-            print("Evaluaciones: " + str(self._evals_done) + " best: " + str(self._best_fitness[best])
-                  + " worst: " + str(self._best_fitness[worst]))
+            #print("Evaluaciones: " + str(self._evals_done) + " best: " + str(self._best_fitness[best])
+             #     + " worst: " + str(self._best_fitness[worst]))
 
             new_chromosome = self.uniform_crossover_operator(self._best_solution[best], self._best_solution[worst])
             mutant = self.segment_mutation_operator(new_chromosome)
@@ -191,10 +191,12 @@ class ILSNueva:
                 self._best_solution[worst] = mutant
                 self._best_fitness[worst] = improved_mutant_fitness
 
-            if np.abs(self._best_fitness[0] - self._best_fitness[1]) < self._best_fitness[best] * self._restart_percent:
-
-                self._best_solution[worst, :] = np.random.randint(0, self._result_nb_clust, self._dim)
-                self._best_fitness[worst] = self.get_single_fitness(self._best_solution[worst, :])[0]
+            #if np.abs(self._best_fitness[0] - self._best_fitness[1]) < self._best_fitness[best] * self._xi:
+            if self._best_fitness[best] - self._best_fitness[worst] > self._best_fitness[best] * self._xi:
+                self._best_solution[best, :] = np.random.randint(0, self._result_nb_clust, self._dim)
+                self._best_fitness[best] = self.get_single_fitness(self._best_solution[best, :])[0]
[…]

### 2019-01-09 Papers/SHADE/paperSHADE.tex
Correcciones de javi

--- 2018-12-19 Papers/SHADE/paperSHADE.tex
@@
 
+\textcolor{red}{Revisar que todos los subindices en los que aparecan comas lleven corchetes.
+Revisar que las asignaciones en los algoritmo sean flechas y no =.
+Revisar que restrictions nunca aparece, siempre constraints.
+Revisar que es siempre instance-level y no instance level.}
+
 \section{Introduction}
 
-\textcolor{red}{Citar a los de las hormigas}\\
-\textcolor{red}{Poner un parrafo de aplicaciones del clustering con restricciones}
+\textcolor{red}{Citar a los de las hormigas. Poner un parrafo de aplicaciones del clustering con restricciones}
 
@@
 A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster.
-Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms.
+**Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms**.
+[It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms]
 
@@
 
-Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the problem complexity.
-The feasibility problem for non hierarchical constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.
+Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the problem's complexity.
+The feasibility problem for non-hierarchical constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.
 
@@
 
-	\textbf{Feasibility Problem}: given a dataset $X$, a set of constraints $R$, and the bounds on the number of clusters $K_l \leq k \leq K_u$.
-Does there exists a partition $C$ of $X$ with $k$ clusters such that all constraints in $R$ are satisfied?
+	\textbf{Feasibility Problem}: given a dataset $X$, a set of constraints $R$, and the bounds on the number of clusters $K_l \leq k \leq K_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $R$ are satisfied?
 \cite{davidson2007survey}\cite{davidson2005clustering}
@@
 
-In \cite{davidson2005clustering} it is proven that, when $K_l = 1$ and $K_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem.
-(It is also proven that it is not harder than it, so both have the same complexity.) Table \ref{tab:feasibility} shows feasibility problem complexity for different types of constraints.
+In \cite{davidson2005clustering} it is proven that, when $K_l = 1$ and $K_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem (it is also proven that it is not harder, so both have the same complexity).
+Table \ref{tab:feasibility} shows the feasibility problem's complexity for different types of constraints.
 
@@
 
-These complexity results show that feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too.
-For more details on constrained clustering complexity see \cite{davidson2005clustering}.
+These complexity results show that the feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too.
+For more details on the complexity of constrained clustering see \cite{davidson2005clustering}.
 
@@
 Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why applying heuristics-guided and population-based algorithms constitute a good approach to find quality solutions to the constrained clustering problem.
+That is the reason why heuristics-guided and population-based algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
+
+\subsection{Short Introduction to Differential Evolution}
+
+\textcolor{red}{Eliminar toda la notacion matematica de esta seccion?}
+
+To introduce Differential Evolution---from now on DE---we take as a starting point the review of the state-of-the-art carried out by \cite{das2011differential}.
+A highly detailed analysis of DE and its variants can also be found in \cite{das2011differential}.
+
+The first article written on DE emerged in 1995 as a technical report by \cite{noman2008differential}.
+From that point on, DE's reputation was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
+In particular, DE excelled in real-coded optimization problems.
+
+DE arose as a variant of population-based evolutionary algorithms, its goal being to perform an intelligent search in the problem's solution space in the most optimized and effective way possible.
+
+Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{i,1},\cdots,p_{i,D}\}$ that serve as parameters for the function to be optimized.
+These individuals are considered solutions to the problem.
+
+To guide the optimization process, we use an objective function.
+DE's task is to seek the parameter vector $p_i$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.4]{Figures/DEop.png}
+	\caption{DE's optimization process \cite{das2011differential}.}\label{img:DE}
+\end{figure}
+
+DE's optimization process works through a simple cycle of stages, presented in Figure \ref{img:DE}, that we can summarize in as follows:
+
+\begin{itemize}
+
+	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of $N$ $D$-dimensional arrays of parameters.
+These vectors constitute candidate solutions for the problem to solve.
+
+	\item \textbf{Difference-vector-based mutation}: DE introduces diversity into the population via mutation operator.
+This operator is applied to each individual of the population, known in this context as the parent vector.
+It combines the parent vector with two randomly selected vectors from the population by applying an arithmetic operator to them.
+The result of this process is a mutant vector.
+
+	\item \textbf{Crossover/Recombination}: The mutant vector exchanges some components with its associated parent to form a trial vector, which will be the candidate to replace the parent vector.
+Several strategies exist to generate this trial vector, which we will not delve into; however, they can be found in \cite{das2011differential}.
+
+	\item \textbf{Selection}: The newly generated trial vector is compared with its associated parent using the objective function.
+If the trial vector is better than the parent, then it will replace the parent in the next generation.
+It is worth noting that, following this strategy, the population never degenerates but only gets better or remains the same.
+
+\end{itemize}
+
+Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
+We will focus on the SHADE variant, which adaptively optimizes some of DE's parameters (see Section \ref{sec:SHADE}).
+
+\subsection{(Classic) Resolution Methods}
+
+\textcolor{red}{Clasicos clasicos no son}
+
+Constrained clustering has many applications and has been widely studied in the literature.
+The first adaptation of a classic clustering method for constrained clustering was proposed by \cite{wagstaff2001constrained}.
+It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
+This method was named COP-kmeans.
+
+The authors of \cite{antoine2012cecm} propose an adaptation of the evidential C-means algorithm, within the fuzzy clustering framework, for constrained clustering.
+The particularity of this resolution method is that the membership of instances to a cluster is defined by a probabilistic belief function.
+This method redefines constraints from the point of view of belief functions and includes them in the cost function.
+
+A modification of CVQE (Constrained Vector Quantization Error), LCVQE, is proposed in \cite{pelleg2007k}.
+The adaptation of CVQE for constrained clustering was proven to produce high quality results, at the cost of a very high computational complexity.
+LCVQE (Linear CVQE) introduces a modification of CVQE's objective function to make it more intuitive and less computationally complex.
+The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraints sets.
+
+TVClust (Two Views Clustering) and RDP-means (Relation Dirichlet Process - Means) were proposed by \cite{khashabi2015clustering}.
+TVClust is able to incorporate the constraints into the clustering problem by making a relaxed interpretation of them.
+The authors model the data set and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
+Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
+This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($K$) does not need to be specified.
+
+\textcolor{red}{Decir que voy a comparar con ellos? Incluir mas métodos ``clasicos''?}
 
@@
 
-In \cite{de2017comparison} authors also added a new element to BRKGA, a local search procedure.
+In \cite{de2017comparison}, the authors also added a new element to BRKGA, a local search procedure.
 This local optimization method is applied to each decoded individual in each generation, but its results are not transferred to the individual, so that diversity is maintained.
@@
 The $h \;\; (0 \le h \le H)$ index specifies the position of the memory $M$ to be updated.
-This index is initialized to $0$ at the beginning of the optimization process and increased by one after each generation.
+This index is initialized to $0$ at the beginning of the optimization process and increased by one after every generation.
 When $h \ge H$ then $h$ is reset to 1.
@@
 Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the greediness of the mutation strategy, is not included in the adaptive optimization process.
-However, it is not static: it is calculated for each individual $p_i$ of the population following Equation \eqref{eq13}.
+However, it is not static: it is calculated for every individual $p_i$ of the population following Equation \eqref{eq13}.
 
@@
 
-To put those concepts clear we consider a toy dataset and three partitions over it.
+To make these concepts clear we consider a toy dataset and three partitions over it.
 Dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
-Partitions $C_0$ and $C_1$ do not violate the constraint ($infeasibility = 0$), while $C_2$ does violate it ($infeasibility = 1$).
+Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($infeasibility = 0$), whereas $C_2$ does violate it ($infeasibility = 1$).
+However, $C_2$ shows a possibility, for later iterations, of adding both instances to their nearest cluster while also satisfying the ML constraint.
 
@@
 
-Table\ref{tab:fitnessfunctions} shows fitness functions values for each partition from Figure \ref{img:toydatasets}.
-We call $f_1$ to the function presented in Equation \eqref{eq1} and $f_2$ to the function presented in Equation \eqref{eq14}.
+Table \ref{tab:fitnessfunctions} shows fitness values for each partition from Figure \ref{img:toydatasets}.
+We refer to the function presented in Equation \eqref{eq1} as $f_1$ and to the function presented in Equation \eqref{eq14} as $f_2$.
 
@@
 
+%---------------------------------------------------------------------%
+
 Results in Table \ref{tab:fitnessfunctions} are a clear example of the above.
-In the general case, with $f_2$ the penalty for violating constraints is proportional to the within-cluster-sum-of-squares, the lower it is the lower the penalty.
-This is not the case with $f_1$, whose penalty term is independent of the within-cluster-sum-of-squares, which can result in a difference of several orders of magnitude between partitions satisfying different number of constraints.
+In the general case, by using $f_2$ the penalty for violating constraints is proportional to the within-cluster-sum-of-squares**,** [---] the lower it is **[,] the lower the penalty.
+This is not the case with $f_1$, whose penalty term is independent of the within-cluster-sum-of-squares, which can result in a difference of several orders of magnitude between partitions satisfying different amounts of constraints.
 
-As in \cite{de2017comparison} we apply a local optimization procedure to the individuals of the population, without transferring its results to the individual to maintain diversity.
+%---------------------------------------------------------------------%
+
+As in \cite{de2017comparison} we apply a local optimization procedure to the individuals of the population, but without transferring its results to the original individual in order to maintain diversity.
 The difference is that we do not apply it to the whole population but only to a number $p_e$ of individuals considered as the elite of the population.
@@
 
-Algorithms \ref{alg:SHADE} shows overall SHADE optimization process and is an adaptation of the one that can be found on \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
+Algorithm \ref{alg:SHADE} describes the overall SHADE optimization process and is an adaptation of the one that can be found on \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
 
@@
 For our experiments we will compare the results obtained by BRKGA and SHADE over XX datasets.
-Most of the datasets used can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
+Most of these data sets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-Table \ref{tab:datasets} present a summary the main features of all datasets used.
+Table \ref{tab:datasets} displays a summary of every data set.
 \textcolor{red}{Esta bien citar tambien al sklearn?
@@
 \subsection{Constraints Generation}
+
+\textcolor{red}{En esta seccion podria simplemente decir el numero de restricciones que genero como un numero magico sin meterme en nada de grafos, pero igual es un poco burro.}
 
@@
 
-Note that the greater the number of classes present in the data set, the fewer ML restrictions obtained with the method proposed by \cite{wagstaff2001constrained}.
+Note that the greater the number of classes present in the data set, the fewer ML constraints obtained with the method proposed by \cite{wagstaff2001constrained}.
 This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the data set increases.
@@
 Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
-We will use Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.Basic Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
+We will use the Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
+The Basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
 $C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions \cite{rand1971objective}.
 
-For each pair of $x_i$ and $x_j$ instances in $X$, $C_i$ assigns them to the same cluster or to different clusters.
-We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in $C_1$ and $C_2$ , and $b$ as the opposite event.
+For each pair of instances $x_i$ and $x_j$ in $X$, $C_i$ assigns them to the same cluster or to different clusters.
+We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in both $C_1$ and $C_2$, and $b$ as the opposite event ($x_i$ and $x_j$ are in different clusters in $C_1$ and $C_2$).
 Then, the degree of similarity between $C_1$ and $C_2$ is calculated as in Equation \eqref{eq15}.
@@
 
-The Adjusted Rand Index is a corrected-for-chance version of Rand Index.
+The Adjusted Rand Index is a corrected-for-chance version of the Rand Index.
 This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline.
@@
 where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
-It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that a $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing the obtained partitions with random partitions.
+It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that an $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing the obtained partitions with random partitions.
 To summarize, the higher the $ARI$, the greater the degree of similarity between $C_1$ and $C_2$.

### 2019-01-11 Papers/DILS/paperDILS.tex
Nuevo paper añadido

--- 2019-01-09 Papers/SHADE/paperSHADE.tex
@@
 
-\title{Enhancing instance-level constrained clustering through differential evolution}
+\title{Improving instance-level constrained clustering through Dual Iterated Local Search}
 
@@
 Revisar que restrictions nunca aparece, siempre constraints.
-Revisar que es siempre instance-level y no instance level.}
+Revisar que es siempre instance-level y no instance level.
+Comprobar que es siempre dataset y no data set}
 
 \section{Introduction}
-
-\textcolor{red}{Citar a los de las hormigas. Poner un parrafo de aplicaciones del clustering con restricciones}
 
@@
 The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset.
-In this way, the clustering process obtains subsets of instances of the initial data set that present high intra-similarity and low inter-similarity.
+In this way, the clustering process obtains subsets of instances of the initial dataset that present high intra-similarity and low inter-similarity.
 
-In most clustering applications it is common to have some kind of information about the data set to be analyzed.
+In most clustering applications it is common to have some kind of information about the dataset to be analyzed.
 In constrained clustering this information is given in the form of pairs of instances.
@@
 
-Before formalizing the definition of constraints we must give a notation for the data set.
+Before formalizing the definition of constraints we must give a notation for the dataset.
 We will note our dataset as $X$, which is a set of $N$ instances with $D$ features.
@@
 
-\subsection{Short Introduction to Differential Evolution}
+\subsection{Short Introduction to Iterated Local Search}
 
-\textcolor{red}{Eliminar toda la notacion matematica de esta seccion?}
+Before defining the iterated local search---from now on ILS---,it is necessary to understand the local search and the motivation that leads to designing methods with greater exploratory capacity.
 
-To introduce Differential Evolution---from now on DE---we take as a starting point the review of the state-of-the-art carried out by \cite{das2011differential}.
-A highly detailed analysis of DE and its variants can also be found in \cite{das2011differential}.
+A local search is a procedure that consists of performing an heuristics-guided iterative improvement.
+Taking as initial state a well defined solution to the problem, the local search uses a neighbor generation operator to explore new solutions.
+When a solution better than the current one is found, it replaces the current one.
+In order to evaluate the solutions, the local search uses a function that assigns a value to each one of them, this is known as a target function.
+***This value will be better the closer the solution that evaluates the optimal solution.***
 
-The first article written on DE emerged in 1995 as a technical report by \cite{noman2008differential}.
-From that point on, DE's reputation was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
-In particular, DE excelled in real-coded optimization problems.
+Both the neighbor generation operator and the target function used by the local search must be defined for each problem.
+The target function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem.
+The neighbor generation operator takes as base a solution to the problem $s$, and generates a new solution $s^\prime$ by modifying some of its components in a local way.
 
-DE arose as a variant of population-based evolutionary algorithms, its goal being to perform an intelligent search in the problem's solution space in the most optimized and effective way possible.
+The Algorithm \ref{alg:LS} presents a basic scheme for a local search procedure.
+It is worth noting that the condition of line 4 must be adapted if the problem to be solved is one of maximization or minimization.
+In addition, the general scheme considers that the optimization process ends when the current solution is not improved, however the stop condition can be adapted to each problem.
 
-Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{i,1},\cdots,p_{i,D}\}$ that serve as parameters for the function to be optimized.
-These individuals are considered solutions to the problem.
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{GenerateNeighbor}{GenerateNeighbor}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Initial solution $s$}
+	\BlankLine
+	\While{$improvement$}{
+		$improvement \leftarrow$ \texttt{false} \\
 
-To guide the optimization process, we use an objective function.
-DE's task is to seek the parameter vector $p_i$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
+		$s^\prime \leftarrow $ \GenerateNeighbor{$s$}
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[scale=0.4]{Figures/DEop.png}
-	\caption{DE's optimization process \cite{das2011differential}.}\label{img:DE}
-\end{figure}
+		\If{$f(s^\prime) < f(s)$}{
 
-DE's optimization process works through a simple cycle of stages, presented in Figure \ref{img:DE}, that we can summarize in as follows:
+			$s \leftarrow s^\prime$\\
+			$improvement \leftarrow$ \texttt{true} \\
 
-\begin{itemize}
+			}	
+	}
+	\BlankLine
+	\KwRet ($s$)
 
-	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of $N$ $D$-dimensional arrays of parameters.
-These vectors constitute candidate solutions for the problem to solve.
+	\caption{Local Search}\label{alg:LS}
+\end{algorithm}
 
-	\item \textbf{Difference-vector-based mutation}: DE introduces diversity into the population via mutation operator.
-This operator is applied to each individual of the population, known in this context as the parent vector.
-It combines the parent vector with two randomly selected vectors from the population by applying an arithmetic operator to them.
-The result of this process is a mutant vector.
+With the local search defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$.
 
-	\item \textbf{Crossover/Recombination}: The mutant vector exchanges some components with its associated parent to form a trial vector, which will be the candidate to replace the parent vector.
-Several strategies exist to generate this trial vector, which we will not delve into; however, they can be found in \cite{das2011differential}.
+\textcolor{red}{Poner un dibujo que lo ejemplifique}
 
-	\item \textbf{Selection}: The newly generated trial vector is compared with its associated parent using the objective function.
-If the trial vector is better than the parent, then it will replace the parent in the next generation.
-It is worth noting that, following this strategy, the population never degenerates but only gets better or remains the same.
+It is therefore necessary to appeal to techniques that introduce diversity in the search to avoid the problem of local optimal. This is exactly what ILS is trying to do.
 
-\end{itemize}
+With ILS, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^prime$ that will not be a local optimum and will be far enough away from $s^*$.
+After that, we will apply the simple local search method to find a new local optimal $s^{*\prime}$.
 
-Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
-We will focus on the SHADE variant, which adaptively optimizes some of DE's parameters (see Section \ref{sec:SHADE}).
+\textcolor{red}{Poner un dibujo que lo ejemplifique, el del handbook of MH pag 368}
+
+It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^prime$ is $s*$ and therefore the solution space will not be explored effectively.
+On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.
+
+The method for generating the perturbation applied to the local optimum $s^*$ may vary depending on the problem, and may be based on previous solutions.
+In this way, it is possible to keep a history of the solutions reached so far, so they can be used when applying the perturbation.
+Algorithm \ref{alg:ILS} summarizes the ILS overall optimization process.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{LocalSearch}{LocalSearch}
+	\SetKwFunction{Perturbation}{Perturbation}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Initial solution $s$}
+	\BlankLine
+	$s^* \leftarrow$ \LocalSearch{$s$}\\
+	\While{Termination criteria are not met}{
+
+		$s^\prime \leftarrow $ \Perturbation{$s^*$, $history$}\\
+		$s^{*\prime} \leftarrow$ \LocalSearch{$s^\prime$}
+
+		\If{$f(s^{*\prime}) < f(s^*)$}{
+
+			$s^* \leftarrow s^{*\prime}$
+
+		}	
+	}
+	\BlankLine
+	\KwRet ($s*$)
+
+	\caption{Iterated Local Search}\label{alg:ILS}
+\end{algorithm}
+
+In the ILS scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only use the difference in the cost of the solutions to be compared to choose between them.
+This is the most simple scheme and does not depends on previous solutions recorded in the history.
+
+\clearpage
 
@@
 TVClust is able to incorporate the constraints into the clustering problem by making a relaxed interpretation of them.
-The authors model the data set and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
+The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
 Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
@@
 
-\section{The SHADE Optimization Method} \label{sec:SHADE}
-
-Success history based adaptive differential evolution---from now on SHADE---is an optimization method based on differential evolution.
-Proposed by \cite{tanabe2013success}, it was an improvement for the JADE model, created by \cite{zhang2009jade}.
-
-Unlike JADE, which only considers the parameters used in the previous generation, SHADE uses a historical record of successful parameters as a mechanism for adapting parameters involved in the process of creating new generations.
-
-\subsection{SHADE Operators}
-
-As in differential evolution, SHADE employs mutant generation, crossover and replacement operators to explore the space of solutions and bring in new generations of individuals.
-
-The mutation strategy used by SHADE is known as current-to-$p$best/1, and the expression that generates new individuals is described in Equation \eqref{eq3}.
-
-\begin{equation}
-m_{[i,G]} = p_{[i,G]} + F_i * (p_{[pbest, G]} - p_{[i,G]}) + F_i * (p_{[r1, G]} - p_{[r2,G]})
-\label{eq3}
-\end{equation}
-
-Where $m_{[i,G]}$ is the mutant vector, which is generated with an individual $p_{[i,G]}$ from the population [serving as a starting point].
-Indices $r1$ and $r2$ are random values in the range $[0,N]$, different from $i$ and from each other.
-The individual $p_{[pbest, G]}$ is randomly selected from among the best $N \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
-This way, the parameter $p$ controls the greediness of the mutation strategy.
-The parameter $F_i$ defines the magnitude of the operator.
-
-After generating the mutant vector $m_{[i,G]}$, it is combined with the parent $p_{[i,G]}$ vector by means of a crossover operator; the result is the trial vector.
-SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.
-
-\begin{equation}
-t_{[j,i,G]} = \left\{ \begin{array}{lc}
-m_{[j,i,G]} &   if \;\; \text{rand}[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
-p_{[j,i,G]} &  \text{otherwise}
-\end{array}
-\right.
-\label{eq4}
-\end{equation}
-
-Where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.
-
-The SHADE replacement operator determines the survivors individuals for the next generation.
-It compares each parent $x_{[i,G]}$ with the trial vector $_{[j,i,G]}$, generated based on it.
-The best one of these two individuals will survive for the next generation.
-Equation \eqref{eq5} shows this concept.
-
-\begin{equation}
-p_{[i,G + 1]} = \left\{ \begin{array}{lc}
-t_{[i,G]} &   if \;\; f(t_{[i,G]}) \le f(p_{[i,G]}) \\
-p_{[i,G]} &  \text{otherwise}
-\end{array}
-\right.
-\label{eq5}
-\end{equation}
-
-To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{[i,G]}$ who were replaced by their associated trial vector $t_{[i,G]}$ are saved.
-To make use of the archive, we will consider that the individual $p_{[r2,G]}$ from Equation \eqref{eq3} is selected from $P \cup A$.
-The archive size is the same as the population size.
-When the size of $A$ exceeds the size of $P$, random individuals are selected for removal to make room for new ones.
-
-The parameters $p$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process.
-These are the parameters that SHADE adaptively optimizes using the mentioned history record.
-
-\subsection{SHADE Parameter Adaptation Method}
-
-The SHADE method stores in memory a structure with $H$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:SHADEmemory}.
-This structure is initialized with the value $0.5$ for all entries.
-
-\begin{table}[!h]
-	\centering
-	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{13pt}
-	%\renewcommand{\arraystretch}{0.9}
-	\resizebox{\textwidth}{!}{
-	\begin{tabular}{|c|c|c|c|c|c|}
-		\hline
-		 Index & 1 & 2 & $\cdots$ & $H - 1$ & $H$ \\
-		 \hline
-		 \hline
-		 $M_{CR}$ & $M_{[CR,1]}$ & $M_{[CR,1]}$ & $\cdots$ & $M_{[CR,H-1]}$ & $M_{[CR,H]}$ \\
-		 \hline
-		 $M_{F}$ & $M_{[F,1]}$ & $M_{[F,1]}$ & $\cdots$ & $M_{[F,H-1]}$ & $M_{[F,H]}$ \\
-		\hline
-
-	\end{tabular}}
-	\caption{Historical memory $M_{CR}$, $M_{F}$ used by SHADE \cite{tanabe2013success}}
-	\label{tab:SHADEmemory}
-\end{table}
-
-In each generation, parameters $CR_i$ and $F_i$ are calculated on the basis of the existing history by applying Equations \eqref{eq6} and \eqref{eq7}:
-
-\begin{equation}
-CR_i = \text{randn}_i(M_{[CR,r_i]}, 0.1)
-\label{eq6}
-\end{equation}
-
-\begin{equation}
-F_i = \text{randc}_i(M_{[F,r_i]}, 0.1)
-\label{eq7}
-\end{equation}
-
-where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$.
-When a value for $CR_i$ outside of the range $[0,1]$ is generated, it is replaced by the corresponding limit value.
-If $F_i > 1$, then it is truncated to $1$; conversely, if $F_i < 0$, then Equation \eqref{eq7} is applied as many times as needed to obtain a legal value.
-
-SHADE keeps two auxiliar sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent.
-At the end of each generation, the content of the memory $M$ is updated following Equations \eqref{eq8} and \eqref{eq9}.
-
-\begin{equation}
-M_{[CR,h,G+1]} = \left\{ \begin{array}{lc}
-\text{mean}_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
-M_{[CR,h,G+1]} &  otherwise
-\end{array}
-\right.
-\label{eq8}
-\end{equation}
-
-\begin{equation}
-M_{[F,h,G+1]} = \left\{ \begin{array}{lc}
-\text{mean}_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
-M_{[F,h,G+1]} &  otherwise
-\end{array}
-\right.
-\label{eq9}
-\end{equation}
-
-The $h \;\; (0 \le h \le H)$ index specifies the position of the memory $M$ to be updated.
-This index is initialized to $0$ at the beginning of the optimization process and increased by one after every generation.
-When $h \ge H$ then $h$ is reset to 1.
-It is worth noting that, when there is a generation with no trial vectors successfully replacing their parents, no update of $M$ is done.
-
-In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the weighted mean, which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed by \cite{peng2009multi} to prevent $CR$ from converging to small values.
-
-\begin{equation}
-\text{mean}_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{[RC,i]}
-\label{eq10}
-\end{equation}
-
-\begin{equation}
-\omega_i = \frac{\Delta f_i}{\sum_{i = 1}^{|S_{RC}|} \Delta f_i}
-\label{eq11}
-\end{equation}
-
-where $\Delta f_i = |f(t_{[i,G]}) - f(p_{[i, G]})|$, which is the amount of improvement obtained from the trial vector with respect to the parent.
-
-In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12} (proposed by \cite{tanabe2013success}).
-
-\begin{equation}
-\text{mean}_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{[F,i]}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{[F,i]}}
-\label{eq12}
-\end{equation}
-
-Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the greediness of the mutation strategy, is not included in the adaptive optimization process.
-However, it is not static: it is calculated for every individual $p_i$ of the population following Equation \eqref{eq13}.
-
-\begin{equation}
-pbest_i = \text{randn}[2/N, 0.2]
-\label{eq13}
-\end{equation}
-
-such that there is always at least two individuals to choose between.
-
-\subsection{Adaptation of SHADE for constrained clustering}
-
-For the adaptation of SHADE to the problem of constrained clustering, we will take some of the elements that allowed us to adapt it to BRKGA.
-To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of random-keys.
-Additionally, we will also reuse the same initialization method and the same decoder used in BRKGA (see Section \ref{sec:AdaptationofBRKGA}).
-
-However, we propose a new fitness function to evaluate the quality of the individuals of the population, Equation \eqref{eq14}, whose element $z_i$ is obtained as in Equation \eqref{eq2}.
-
-\begin{equation}
-f_i = z_i * (infeasibility + 1)
-\label{eq14}
-\end{equation}
-
-Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated from the dataset or constraints, whereas in Equation \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.
-
-We found that with the fitness function defined in \eqref{eq1} there is no competition between the penalty term and the within-cluster-sum-of-squares term, because the penalty is always significantly larger than it or zero.
-This can bias the exploration of the solution space, restricting it in practice to those that satisfy all the constraints, even if the within-cluster-sum-of-squares is still improvable by moving two instances involved in a constraint to different clusters without violating that constraint.
-
-With Equation \eqref{eq14} we try to find a trade-off between the within-cluster-sum-of-squares and the penalty term, allowing solutions that violate a certain number of constraints to compete with those who violate a smaller number of them but score a better within-cluster-sum-of-squares.
-
-To make these concepts clear we consider a toy dataset and three partitions over it.
-Dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
-Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($infeasibility = 0$), whereas $C_2$ does violate it ($infeasibility = 1$).
-However, $C_2$ shows a possibility, for later iterations, of adding both instances to their nearest cluster while also satisfying the ML constraint.
-
-\begin{figure}[bth]
-	\myfloatalign
-	{\includegraphics[width=.45\linewidth]{Figures/Dataset}} \quad
-	%\subfloat[Imagen original]
-	{\includegraphics[width=.45\linewidth]{Figures/C0}} \quad
-	%\subfloat[Clustering sin restricciones]
-	{\includegraphics[width=.45\linewidth]{Figures/C1}} \quad
-	%\subfloat[Clustering con restricciones]
-	{\includegraphics[width=.45\linewidth]{Figures/C2}} \quad
-	\caption{Three partitions over a toy dataset.}
-	\label{img:toydatasets}
-\end{figure}
-
-Table \ref{tab:fitnessfunctions} shows fitness values for each partition from Figure \ref{img:toydatasets}.
-We refer to the function presented in Equation \eqref{eq1} as $f_1$ and to the function presented in Equation \eqref{eq14} as $f_2$.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.3}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{c c c c c c}
-			\hline
-			\multirow{2}{*}{Partition} &
-			\multicolumn{2}{c}{Expression} &&
-			\multicolumn{2}{c}{Value} \\
-			\cline{2-3} \cline{5-6}
-			& $f_1$ & $f_2$ && $f_1$ & $f_2$ \\
-			\hline
-			$C_0$ & $z_0$ & $z_0$ && $0.457$  & $0.457$  \\
-			$C_1$ & $z_1$ & $z_1$  && $0.540$  & $0.540$  \\
-			$C_2$ & $z_2 + (\mu * N * infs)$ & $z_2 * (infs + 1)$  && $0.503 + 1000 = 1000.503$ &  $0.503 * 2 = 1.005$ \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Expression and value of fitness functions over three partitions. ($\mu = 10$)}
-	\label{tab:fitnessfunctions}
-\end{table}
-
-%---------------------------------------------------------------------%
-
-Results in Table \ref{tab:fitnessfunctions} are a clear example of the above.
-In the general case, by using $f_2$ the penalty for violating constraints is proportional to the within-cluster-sum-of-squares**,** [---] the lower it is **[,] the lower the penalty.
-This is not the case with $f_1$, whose penalty term is independent of the within-cluster-sum-of-squares, which can result in a difference of several orders of magnitude between partitions satisfying different amounts of constraints.
-
-%---------------------------------------------------------------------%
-
-As in \cite{de2017comparison} we apply a local optimization procedure to the individuals of the population, but without transferring its results to the original individual in order to maintain diversity.
-The difference is that we do not apply it to the whole population but only to a number $p_e$ of individuals considered as the elite of the population.
-The percentage of individuals considered as elite must be determined for each case.
-
-Algorithm \ref{alg:SHADE} describes the overall SHADE optimization process and is an adaptation of the one that can be found on \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
-
-\begin{algorithm}
-	\SetNlSty{textbf}{[}{]}
-	\SetNlSkip{0.5em}
-	\setstretch{1.2}
-	\SetKwFunction{Sort}{Sort}
-	\SetKwFunction{LocalSearch}{LocalSearch}
-	\KwIn{Dataset $X$, constraints sets $c_=$ and $c_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $K$.}
-	\tcp{Initialization phase}
-	$G \leftarrow 0$\\
-	Initialize population $P_0 \leftarrow \{p_{[1,0]}, \cdots, p_{[p_{size},0]}\}$\\
-	Initialize all values in $M_{CR}$, $M_F$ to 0.5\\
-	$A \leftarrow \emptyset$; $h \leftarrow 1$\\
-	\tcp{Main loop}
-	\While{Termination criteria are not met}{
-
-		$S_{CR} \leftarrow S_F \leftarrow \emptyset$\\
-		$P_G \leftarrow$ \Sort{$P_G$} \tcp{Sort the population}
-		\tcp{Apply a LS procedure to the elite of the population}
-		\LocalSearch{$\{p_{[1,G]}, \cdots, p_{[p_{e},G]}\}$}\\
-		\For{$i \in [1,p_{size}]$}{
-			$r_i \leftarrow$ randInt$[1,H]$\\
-			$CR_{[i,G]} \leftarrow$ randn$_i(M_{[CR,r_i]}, 0.1)$\\
-			$F_{[i,G]} \leftarrow$ randc$_i(M_{[F,r_i]}, 0.1)$\\
-			$p_{[i,G]} \leftarrow$ rand$[N/2, 0.2]$\\
-			Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin\\
-
-			\eIf{$f(t_{[i,G]}) \le f(p_{[i,G]})$}{
-				$p_{[i,G + 1]} \leftarrow t_{[i,G]}$
-			}{
-				$p_{[i,G + 1]} \leftarrow p_{[i,G]}$
-			}
-
-			\If{$f(t_{[i,G]}) < f(p_{[i,G]})$}{
-				$p_{[i,G]} \rightarrow A$;
-				$CR_{[i,G]} \rightarrow S_{CR}$;
-				$F_{[i,G]} \rightarrow S_{F}$
-			}
-
-		}
-		Whenever $|A| > |P|$, randomly select an individual from $A$ to be deleted so that $|A| \le |P|$\\
-		\If{$S_{CR} \neq \emptyset$ \textbf{and} $S_{F} \neq \emptyset$}{
-			Update $M_{[CR,h]}$ and $M_{[F,h]}$ based on $S_{CR}$ and $S_{F}$\\
-			$h \leftarrow (h + 1) \mod H$
-		}
-		$G++$
-	}
-	\caption{Modified SHADE}\label{alg:SHADE}
-\end{algorithm}
-
-\begin{algorithm}
-	\SetNlSty{textbf}{[}{]}
-	\SetNlSkip{0.5em}
-	\setstretch{1.2}
-	\SetKwFunction{RandomShuffle}{RandomShuffle}
-	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Dataset $X$, constraints sets $c_=$ and $c_{\neq}$, decoded random-key vector (solution) $S$, number of clusters $K$.}
-	\BlankLine
-	\While{$improvement$}{
-		$improvement \leftarrow$ \texttt{false} \\
-
-		$s_i \leftarrow $ Select random object from $S$\\
-		\tcp{Random shuffle labels set}
-		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
-		\For{$c \in RSL$}{
-			$S^\prime \leftarrow S$\\
-			$S^\prime$\texttt{[}$s_i$\texttt{]} $\leftarrow c$ \tcp{Move object $s_i$ to cluster $c$}
-
-			\If{$f(S^\prime) < f(S)$}{
-				$S \leftarrow S^\prime$\\
-				$improvement \leftarrow$ \texttt{true} \\
-			}
-		}	
-	}
-	\BlankLine
-	\KwRet ($S$)
-
-\caption{Local Search}\label{alg:LS}
-\end{algorithm}
+\section{The DILS Optimization Method}
 
@@
 For our experiments we will compare the results obtained by BRKGA and SHADE over XX datasets.
-Most of these data sets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
+Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-Table \ref{tab:datasets} displays a summary of every data set.
+Table \ref{tab:datasets} displays a summary of every dataset.
 \textcolor{red}{Esta bien citar tambien al sklearn?
-Que paper de keel cito, 2017 o 2011?}
+Que paper de keel cito, 2017 o 2011?
+Decir que algunos datasets son artifiales si se quedan, igual para los del sklearn}
 
@@
 
-Since we have the true labels associated with each data set, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraint sets.
-This method consists in randomly selecting two instances of a data set, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
+Since we have the true labels associated with each dataset, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraint sets.
+This method consists in randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
 
-We will generate four different sets of constraints for each data set.
+We will generate four different sets of constraints for each dataset.
 The number of constraints is given by the number of edges of the complete constraints graph that we can build with a percentage of labeled data.
@@
 
-Note that the greater the number of classes present in the data set, the fewer ML constraints obtained with the method proposed by \cite{wagstaff2001constrained}.
-This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the data set increases.
+Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed by \cite{wagstaff2001constrained}.
+This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.
 
@@
 
+\textcolor{red}{Poner tabla con parametros para los dos algoritmos.
+Decir que comparamos rendimiento medio y no rendimiento pico, porque el proceso de optimización de parámetros neceseario para cada algoritmo y dataset cae fuera de este estudio.
+Comentar que el curterio de finalización es número de evaluaciones de la funcion objetivo}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
+			\hline
+			Parameter & Meaning & BRKGA & SHADE \\
+			\hline
+			$p_{size}$ & Population size & 100 & 100 \\
+			$Evals$ & Fitness function evaluations & 300000 & 300000 \\
+			$p_e$ & Size of the elite set in population & $0.2 * p_{size}$ & $0.25 * p_{size}$ \\
+			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_{size}$ & $\times$ \\
+			$p_{inherit}$ & Probability that a feature is inherit from an elite parent & $50\%$ & $\times$ \\
+			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Depends on the dataset and it has the same value for both methods} \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+		\caption{Parameters setup used for BRKGA and SHADE.}
+		\label{undefined}
+	\end{table}
+
 \begin{table}[!h]

### 2019-02-07 Papers/DILS/paperDILS.tex
Añadidos algoritmos TFG, resultados TFG y modificado paper SHADE

@@
 
+%\usepackage{pdfpages}
+\usepackage{graphics, graphicx, float, url}
 \usepackage{lineno,hyperref}
@@
 \modulolinenumbers[5]
+
+\usepackage{epstopdf}
+\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
+\AppendGraphicsExtensions{.tiff}
 
@@
 With the local search defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$.
+Figure \ref{img:LS} shows a pictorical representation of the above.
 
-\textcolor{red}{Poner un dibujo que lo ejemplifique}
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{Figures/LS.png}
+	\caption{Local optimum achieved by a local search procedure}\label{img:LS}
+\end{figure}
 
@@
 
-With ILS, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^prime$ that will not be a local optimum and will be far enough away from $s^*$.
-After that, we will apply the simple local search method to find a new local optimal $s^{*\prime}$.
+With ILS, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^\prime$ that will not be a local optimum and will be far enough away from $s^*$.
+After that, we will apply a local search procedure to find a new local optimal $s^{*\prime}$.
+Figure \ref{img:ILS} shows a representation of the above concepts.
 
-\textcolor{red}{Poner un dibujo que lo ejemplifique, el del handbook of MH pag 368}
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{Figures/ILS.png}
+	\caption{No se como llamar a esta puta imagen}\label{img:ILS}
+\end{figure}
 
-It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^prime$ is $s*$ and therefore the solution space will not be explored effectively.
+It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^\prime$ is $s^*$ and therefore the solution space will not be explored effectively.
 On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.

### 2019-02-07 Papers/SHADE/paperSHADE.tex
Añadidos algoritmos TFG, resultados TFG y modificado paper SHADE

@@
 Revisar que restrictions nunca aparece, siempre constraints.
-Revisar que es siempre instance-level y no instance level.}
+Revisar que es siempre instance-level y no instance level.
+Comprobar que es siempre dataset y no data set}
 
@@
 
-The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set.
+The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraints set.
 As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $N = |X| = \sum_{i = 1}^{K} |c_i|$.
@@
 To guide the optimization process, we use an objective function.
-DE's task is to seek the parameter vector $p_i$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
+DE's task is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
 
@@
 	\centering
-	\includegraphics[scale=0.4]{Figures/DEop.png}
+	\includegraphics[scale=0.3]{Figures/DEop.png}
 	\caption{DE's optimization process \cite{das2011differential}.}\label{img:DE}
@@
 
-SHADE keeps two auxiliar sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent.
+SHADE keeps two auxiliary sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent.
 At the end of each generation, the content of the memory $M$ is updated following Equations \eqref{eq8} and \eqref{eq9}.
@@
 To make these concepts clear we consider a toy dataset and three partitions over it.
-Dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
+The dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
 Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($infeasibility = 0$), whereas $C_2$ does violate it ($infeasibility = 1$).
@@
 
-%---------------------------------------------------------------------%
-
 Results in Table \ref{tab:fitnessfunctions} are a clear example of the above.
-In the general case, by using $f_2$ the penalty for violating constraints is proportional to the within-cluster-sum-of-squares**,** [---] the lower it is **[,] the lower the penalty.
+In the general case, by using $f_2$ the penalty for violating constraints is proportional to the within-cluster-sum-of-squares---the lower it is, the lower the penalty.
 This is not the case with $f_1$, whose penalty term is independent of the within-cluster-sum-of-squares, which can result in a difference of several orders of magnitude between partitions satisfying different amounts of constraints.
-
-%---------------------------------------------------------------------%
 
@@
 
-For our experiments we will compare the results obtained by BRKGA and SHADE over XX datasets.
-Most of these data sets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
+For our experiments we will compare the results obtained by BRKGA and SHADE over 25 datasets.
+Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-Table \ref{tab:datasets} displays a summary of every data set.
-\textcolor{red}{Esta bien citar tambien al sklearn?
-Que paper de keel cito, 2017 o 2011?}
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at GitHub.
+Table \ref{tab:datasets} displays a summary of every dataset.
 
@@
 		\hline
-		Name & No. Instances & No. Classes & Features \\
+		Name & No. Instances & No. Classes & No. Features \\
 		\hline
-		Appendicitis & 106 & 2 &  \\
-		Balance & 625 & 3 &  \\
-		Boston & 506 & 229 &  \\
-		Breast Cancer & 569 & 2 &  \\
-		Bupa & 345 & 2 &  \\
-		Circles & 300 & 2 &  \\
-		Contraceptive & 1473 & 3 &  \\
-		Diabetes & 442 & 214 &  \\
-		Ecoli & 336 & 8 &  \\
-		Glass & 214 & 6 &  \\
-		Haberman & 306 & 2 &  \\
-		Hayesroth & 160 & 3 &  \\
-		Heart & 270 & 2 &  \\
-		Ionosphere & 351 & 2 &  \\
-		Iris & 150 & 3 &  \\
-		Led7Digit & 500 & 10 &  \\
-		Monk2 & 432 & 2 &  \\
-		Moons & 300 & 2 &  \\
-		Movement Libras & 360 & 15 &  \\
-		Newthyroid & 215 & 3 &  \\
-		Pima & 768 & 2 &  \\
-		Rand & 150 & 3 &  \\
-		Saheart & 462 & 2 &  \\
-		Sonar & 208 & 2 &  \\
-		Soybean & 47 & 4 &  \\
-		Spectfheart & 267 & 2 &  \\
-		Spiral & 300 & 2 &  \\
-		Tae & 151 & 3 &  \\
-		Vehicle & 846 & 4 &  \\
-		Vowel & 990 & 11 &  \\
-		Wdbc & 569 & 2 &  \\
-		Wine & 178 & 3 &  \\
-		Zoo & 101 & 7 &  \\
+		Appendicitis & 106 & 2 & 7 \\
+		Boston & 506 & 229 & 13 \\
+		Breast Cancer & 569 & 2 & 30 \\
+		Circles & 300 & 2 & 2 \\
+		Diabetes & 442 & 214 & 10 \\
+		Ecoli & 336 & 8 & 7 \\
+		Glass & 214 & 6 & 9 \\
+		Haberman & 306 & 2 & 3 \\
+		Hayesroth & 160 & 3 & 4 \\
+		Ionosphere & 351 & 2 & 33 \\
+		Iris & 150 & 3 & 4 \\
+		Led7Digit & 500 & 10 & 7 \\
+		Monk2 & 432 & 2 & 6 \\
+		Moons & 300 & 2 & 2 \\
+		Pima & 768 & 2 & 8 \\
+		Saheart & 462 & 2 & 9 \\
+		Sonar & 208 & 2 & 60 \\
+		Soybean & 47 & 4 & 35 \\
+		Spectfheart & 267 & 2 & 44 \\
+		Spiral & 300 & 2 & 2 \\
+		Tae & 151 & 3 & 5 \\
+		Vehicle & 846 & 4 & 18 \\
+		Vowel & 990 & 11 & 13 \\
+		Wdbc & 569 & 2 & 30 \\
+		Zoo & 101 & 7 & 16 \\
 		\hline
@@
 
-Since we have the true labels associated with each data set, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraint sets.
-This method consists in randomly selecting two instances of a data set, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
+Since we have the true labels associated with each dataset, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraints sets.
+This method consists in randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
 
-We will generate four different sets of constraints for each data set.
+We will generate four different sets of constraints for each dataset.
 The number of constraints is given by the number of edges of the complete constraints graph that we can build with a percentage of labeled data.
-Each constraint set is associated with a different percentage of labeled data, namely: 5\%, 10\%, 15\% and 20\%.
+Each constraints set is associated with a different percentage of labeled data, namely: 5\%, 10\%, 15\% and 20\%.
 It is worth noting that constraints are not the complete constraint graph made up of the labeled instances, this graph only sets the number of constraints to be created, this way there is a lower probability of biasing the set of constraints so that there are classes with poor representation in it.
@@
 		Appendicitis & 13 & 2 && 39 & 16 && 71 & 49 && 164 & 67 \\
-		Balance & 198 & 298 && 838 & 1115 && 1865 & 2506 && 3336 & 4414 \\
 		Boston & 0 & 325 && 3 & 1272 && 13 & 2837 && 26 & 5125 \\
 		Breast Cancer & 216 & 190 && 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
-		Bupa & 79 & 74 && 323 & 272 && 699 & 627 && 1201 & 1145 \\
 		Circles & 50 & 55 && 208 & 227 && 502 & 488 && 853 & 917 \\
-		Contraceptive & 994 & 1707 && 3765 & 7113 && 8501 & 15809 && 15399 & 27966 \\
 		Diabetes & 2 & 251 && 5 & 985 && 7 & 2204 && 23 & 3893 \\
@@
 		Hayesroth & 12 & 16 && 39 & 81 && 102 & 174 && 177 & 319 \\
-		Heart & 41 & 50 && 178 & 173 && 396 & 424 && 744 & 687 \\
 		Ionosphere & 92 & 61 && 330 & 300 && 732 & 646 && 1299 & 1186 \\
@@
 		Moons & 55 & 50 && 200 & 235 && 494 & 496 && 900 & 870 \\
-		Movement Libras & 6 & 147 && 27 & 603 && 112 & 1319 && 158 & 2398 \\
-		Newthyroid & 25 & 30 && 108 & 123 && 270 & 258 && 449 & 454 \\
 		Pima & 412 & 329 && 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
-		Rand & 8 & 20 && 25 & 80 && 76 & 177 && 151 & 284 \\
 		Saheart & 152 & 124 && 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
@@
 		Wdbc & 209 & 197 && 840 & 756 && 1925 & 1730 && 3472 & 2969 \\
-		Wine & 14 & 22 && 49 & 104 && 121 & 230 && 217 & 413 \\
 		Zoo & 7 & 8 && 21 & 34 && 29 & 91 && 41 & 169 \\
@@
 
-\clearpage
+Table \ref{tab:params} shows a summary of the parameters setup used for both BRKGA+LS and SHADE+LS algorithms.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
+			\hline
+			Parameter & Meaning & BRKGA & SHADE \\
+			\hline
+			$p_{size}$ & Population size & 100 & 100 \\
+			$Evals$ & Fitness function evaluations & 300000 & 300000 \\
+			$p_e$ & Size of the elite set in population & $0.2 * p_{size}$ & $0.25 * p_{size}$ \\
+			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_{size}$ & - \\
+			$p_{inherit}$ & Probability that a feature is inherit from an elite parent & $50\%$ & - \\
+			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Known for each dataset, it has the same value for both methods} \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+		\caption{Parameters setup used for BRKGA and SHADE.}
+		\label{tab:params}
+	\end{table}
+
+It should be noted that the study presented in this document aims to compare average performance, not peak performance.
+The parameter optimization process required for each algorithm and data set falls outside of this study.
+
+In both cases---BRKGA+LS and SHADE+LS---the population size used will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
 
 \section{Experimental Results}
+
+In this section we present the Tables \ref{tab:results5} to \ref{tab:results20}, which record the results obtained by the methods to be compared for each dataset and constraints set.
+
+The ARI column shows the Adjusted Rand Index for each case, the Unsat column displays the percentage of violated constraints and the Time columns shows the time---measured in seconds---taken for each algorithm to deliver results.
 
@@
 			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA} &&  \multicolumn{3}{c}{SHADE} \\
+			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
 			\cline{2-4} \cline{6-8}
-			& RandIndex & Unsat(\%) & Time(s) && RandIndex & Unsat(\%) & Time(s) \\
+			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
 			\hline
-			&  &  &  &&  &  &  \\
-			&  &  &  &&  &  &  \\
-			&  &  &  &&  &  &  \\
-			\hline
-			Means &  &  &  &&  &  &  \\
+			Appendicitis & 0.001 & 0.000 & 374.388 && -0.047 & 0.000 & 360.150 \\
+			Boston & -0.002 & 0.000 & 1405.381 && 0.001 & 0.000 & 1425.668 \\
+			Breast Cancer & 0.052 & 9.704 & 3427.396 && 0.099 & 7.685 & 3453.035 \\
+			Circles & 0.020 & 1.333 & 1042.597 && 0.027 & 1.143 & 1004.287 \\
+			Diabetes & 0.001 & 0.000 & 1173.404 && 0.000 & 0.633 & 1374.711 \\
+			Ecoli & 0.005 & 6.912 & 1098.664 && 0.004 & 1.618 & 1045.699 \\
+			Glass & -0.009 & 0.000 & 703.172 && -0.007 & 0.000 & 686.843 \\
+			Haberman & 0.010 & 2.000 & 1098.043 && -0.003 & 2.000 & 1065.608 \\
+			Hayesroth & -0.009 & 0.000 & 534.294 && 0.002 & 0.000 & 507.300 \\
+			Ionosphere & 0.025 & 3.399 & 1821.049 && 0.044 & 3.530 & 1780.769 \\
+			Iris & 0.433 & 0.000 & 494.135 && 0.670 & 0.000 & 451.290 \\
+			Led7Digit & 0.006 & 4.867 & 1668.206 && 0.002 & 1.933 & 1648.570 \\
+			Monk2 & 0.043 & 3.810 & 1743.874 && 0.035 & 3.637 & 1702.858 \\
+			Moons & 0.081 & 1.905 & 1040.926 && 0.157 & 0.571 & 1021.028 \\
+			Pima & 0.011 & 15.169 & 3803.816 && 0.014 & 12.605 & 3723.091 \\
+			Saheart & 0.029 & 6.160 & 2013.580 && 0.006 & 6.739 & 1977.839 \\
+			Sonar & 0.010 & 0.364 & 1076.248 && 0.001 & 0.727 & 1061.966 \\
+			Soybean & 0.371 & 0.000 & 175.224 && 0.483 & 0.000 & 159.017 \\
+			Spectfheart & 0.104 & 1.099 & 1395.389 && 0.084 & 0.659 & 1365.181 \\
+			Spiral & 0.016 & 0.952 & 1044.697 && 0.004 & 0.953 & 1021.158 \\
+			Tae & 0.004 & 0.000 & 506.495 && 0.010 & 0.000 & 476.786 \\
+			Vehicle & 0.002 & 19.557 & 4008.912 && 0.003 & 10.255 & 4024.292 \\
+			Vowel & 0.001 & 10.465 & 3796.154 && 0.001 & 8.000 & 3763.485 \\
+			Wdbc & 0.038 & 10.246 & 3424.397 && 0.025 & 9.212 & 3400.415 \\
+			Zoo & 0.016 & 0.000 & 332.475 && 0.129 & 0.000 & 318.684 \\
 			\hline
@@
 
-	\caption{Experimental results obtained with X\% of labeled data}
-	\label{undefined}
+	\caption{Experimental results obtained with 5\% of labeled data}
+	\label{tab:results5}
 \end{table}
+
+Table \ref{tab:results5} shows the results for the constraints set obtained with 5\% of the labeled data.
+We can note that SHADE+LS represents a consistent improvement over BRKGA + LS in both Unsat and Time, while in ARI both methods seems similar.
+This is because, operating with a small set of restrictions, the penalty applied to the fitness function of both methods has little presence, and since they are population-based and heurist-guided methods, they have similar exploration and exploitation capabilities.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{l ccc c ccc}
+			\hline
+			\multirow{2}{*}{Dataset} &
+			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
+			\cline{2-4} \cline{6-8}
+			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
+			\hline
+			Appendicitis & 0.022 & 5.454 & 383.642 && 0.086 & 5.091 & 366.491 \\
+			Boston & -0.000 & 0.173 & 1502.452 && -0.001 & 0.235 & 1570.547 \\
+			Breast\_Cancer & 0.395 & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
+			Circles & 0.166 & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
+			Diabetes & 0.000 & 0.364 & 1277.982 && 0.003 & 0.465 & 1284.486 \\
+			Ecoli & 0.009 & 20.321 & 1169.798 && 0.020 & 12.763 & 1112.981 \\
+			Glass & 0.016 & 8.658 & 735.507 && 0.013 & 4.848 & 694.962 \\
+			Haberman & 0.047 & 15.699 & 1155.985 && 0.084 & 15.441 & 1146.023 \\
+			Hayesroth & 0.027 & 1.500 & 550.117 && 0.025 & 2.000 & 529.424 \\
+			Ionosphere & 0.125 & 17.587 & 1894.390 && 0.332 & 12.603 & 1872.308 \\
+			Iris & 0.238 & 0.190 & 505.112 && 0.342 & 0.000 & 483.584 \\
+			Led7Digit & 0.009 & 11.347 & 1786.943 && 0.004 & 8.866 & 1784.708 \\
+			Monk2 & 0.421 & 14.926 & 1848.543 && 0.270 & 16.385 & 1821.208 \\
+			Moons & 0.277 & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
+			Pima & 0.236 & 25.195 & 4072.087 && 0.275 & 22.468 & 3990.383 \\
+			Saheart & 0.147 & 20.555 & 2135.624 && 0.265 & 16.725 & 2095.554 \\
+			Sonar & 0.124 & 7.619 & 1106.039 && 0.116 & 8.952 & 1081.994 \\
+			Soybean & 0.496 & 0.000 & 176.333 && 0.515 & 0.000 & 151.163 \\
+			Spectfheart & 0.334 & 9.573 & 1451.235 && 0.343 & 8.604 & 1405.243 \\
+			Spiral & 0.058 & 15.402 & 1098.622 && 0.114 & 13.793 & 1099.747 \\
+			Tae & 0.018 & 0.833 & 520.855 && 0.024 & 1.333 & 508.988 \\
+			Vehicle & 0.005 & 28.515 & 4366.104 && 0.006 & 22.151 & 4352.663 \\
+			Vowel & 0.003 & 13.330 & 4245.672 && 0.002 & 11.523 & 4290.154 \\
+			Wdbc & 0.478 & 16.316 & 3596.517 && 0.357 & 17.093 & 3528.029 \\
+			Zoo & 0.105 & 1.091 & 336.345 && 0.098 & 2.182 & 317.674 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained with 10\% of labeled data}
+	\label{tab:results:10}
+\end{table}
+
+The Table \ref{tab:results:10} presents the results obtained for the constraints set based on 10\% labeled data.
+On the results for Unsat and Time, we can draw similar conclusions to those obtained with Table \ref{tab:results5}, that is, the advantage remains for SHADE + LS, even with the general increase of these values derived from a larger set of constraints.
+It is at ARI where we begin to see the difference between the two methods.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{l ccc c ccc}
+			\hline
+			\multirow{2}{*}{Dataset} &
+			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
+			\cline{2-4} \cline{6-8}
+			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
+			\hline
+			Appendicitis & 0.296 & 6.167 & 381.625 && 0.332 & 5.833 & 362.106 \\
+			Boston & 0.000 & 0.526 & 1648.008 && 0.001 & 0.547 & 1688.985 \\
+			Breast\_Cancer & 0.772 & 10.172 & 3854.911 && 0.911 & 3.310 & 3766.402 \\
+			Circles & 0.828 & 6.101 & 1136.484 && 0.982 & 0.404 & 1108.928 \\
+			Diabetes & 0.004 & 0.317 & 1364.941 && -0.000 & 0.371 & 1424.047 \\
+			Ecoli & 0.015 & 24.314 & 1200.463 && 0.025 & 19.090 & 1127.568 \\
+			Glass & 0.017 & 17.500 & 745.987 && 0.031 & 11.743 & 722.009 \\
+			Haberman & 0.866 & 4.677 & 1209.471 && 0.796 & 5.314 & 1147.479 \\
+			Hayesroth & 0.107 & 7.536 & 545.861 && 0.099 & 8.261 & 520.702 \\
+			Ionosphere & 0.803 & 7.722 & 1986.543 && 0.591 & 11.974 & 1944.085 \\
+			Iris & 0.412 & 3.873 & 506.548 && 0.260 & 6.087 & 486.723 \\
+			Led7Digit & 0.004 & 12.836 & 1902.272 && 0.007 & 10.919 & 1897.511 \\
+			Monk2 & 0.751 & 10.827 & 1949.305 && 0.945 & 2.067 & 1920.361 \\
+			Moons & 0.958 & 1.334 & 1133.858 && 0.995 & 0.020 & 1113.687 \\
+			Pima & 0.481 & 22.051 & 4451.193 && 0.762 & 9.928 & 4402.693 \\
+			Saheart & 0.749 & 10.940 & 2281.992 && 0.945 & 1.839 & 2225.210 \\
+			Sonar & 0.800 & 5.040 & 1124.700 && 0.890 & 2.177 & 1069.862 \\
+			Soybean & 0.428 & 0.000 & 170.674 && 0.422 & 0.000 & 155.948 \\
+			Spectfheart & 0.924 & 2.854 & 1561.040 && 0.778 & 5.732 & 1485.260 \\
+			Spiral & 0.847 & 5.677 & 1133.962 && 0.949 & 1.495 & 1102.920 \\
+			Tae & 0.119 & 7.115 & 522.088 && 0.066 & 8.380 & 513.289 \\
+			Vehicle & 0.089 & 28.044 & 4806.559 && 0.018 & 26.856 & 4749.812 \\
+			Vowel & 0.002 & 14.486 & 4807.165 && 0.001 & 13.341 & 4867.849 \\
+			Wdbc & 0.805 & 8.985 & 3829.407 && 0.930 & 2.801 & 3750.711 \\
+			Zoo & 0.157 & 3.500 & 333.984 && 0.128 & 2.834 & 315.131 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained with 15\% of labeled data}
+	\label{tab:results15}
+\end{table}
+
+Table \ref{tab:results15} presents the results obtained with 15\% of labeled data.
+While conclusions for Unsat and Time remain unchanged, we see a clear difference as far as ARI is concerned.
+We observe that, for constraint sets obtained with 15\% labeled data, both methods already report ARI results above 0.7, which in an ARI ranking is considered to be a good result.
+We can even observe that the SHADE + LS method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA + LS that only obtains similar results in 2 datasets.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{l ccc c ccc}
+			\hline
+			\multirow{2}{*}{Dataset} &
+			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
+			\cline{2-4} \cline{6-8}
+			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
+			\hline
+			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
+			Boston & 0.001 & 0.636 & 1820.313 && 0.001 & 0.660 & 1878.863 \\
+			Breast\_Cancer & 0.807 & 9.086 & 4061.162 && 0.977 & 0.891 & 3984.370 \\
+			Circles & 0.935 & 2.814 & 1180.268 && 1.000 & 0.000 & 1170.452 \\
+			Diabetes & 0.003 & 0.695 & 1480.035 && 0.001 & 0.710 & 1589.821 \\
+			Ecoli & 0.028 & 25.716 & 1267.419 && 0.047 & 21.466 & 1277.903 \\
+			Glass & 0.038 & 23.610 & 766.861 && 0.082 & 16.412 & 796.162 \\
+			Haberman & 0.943 & 2.073 & 1266.131 && 0.997 & 0.138 & 1235.441 \\
+			Hayesroth & 0.548 & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
+			Ionosphere & 0.902 & 4.314 & 2051.251 && 0.993 & 0.258 & 2019.558 \\
+			Iris & 0.691 & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
+			Led7Digit & 0.007 & 13.653 & 2053.264 && 0.005 & 12.170 & 2066.223 \\
+			Monk2 & 0.855 & 6.025 & 2055.425 && 0.991 & 0.379 & 2013.468 \\
+			Moons & 0.937 & 2.689 & 1177.848 && 0.995 & 0.158 & 1157.222 \\
+			Pima & 0.753 & 11.544 & 4855.446 && 0.899 & 4.419 & 4820.684 \\
+			Saheart & 0.881 & 5.250 & 2402.513 && 0.981 & 0.706 & 2353.931 \\
+			Sonar & 0.992 & 0.186 & 1138.769 && 1.000 & 0.000 & 1127.020 \\
+			Soybean & 0.416 & 0.000 & 171.281 && 0.563 & 0.000 & 153.406 \\
+			Spectfheart & 0.953 & 1.677 & 1602.229 && 1.000 & 0.000 & 1526.925 \\
+			Spiral & 0.921 & 3.604 & 1178.580 && 0.998 & 0.091 & 1155.194 \\
+			Tae & 0.351 & 9.807 & 527.292 && 0.392 & 9.419 & 508.610 \\
+			Vehicle & 0.144 & 28.781 & 5189.066 && 0.044 & 29.008 & 5284.288 \\
+			Vowel & 0.002 & 14.682 & 5450.331 && 0.003 & 13.766 & 5514.052 \\
+			Wdbc & 0.823 & 8.281 & 4019.683 && 0.972 & 1.068 & 3983.564 \\
+			Zoo & 0.176 & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained with 20\% of labeled data}
+	\label{tab:results20}
+\end{table}
+
+It is in Table \ref{tab:results20}, which gathers results obtained with constraint sets generated from 20\% of labeled data for each dataset, that we observe the major differences between the two methods.
+We found that SHADE + LS is capable of producing clusters identical to the original ones for 4 datasets, against the single dataset for which BRKGA + LS obtains a similar result.
+Let's recall that an ARI of 1 represents a total matching between the two partitions given as arguments for ARI's calculation.
+This improvement in ARI is clearly reflected in the percentage of violated constraints--Unsat column---, value for which, once again, SHADE + LS gets the advantage.
+Conclusions on execution time remain the same.
+
+\section{Results Analysis}
+
+With the results obtained by both methods for a total of 100 different datasets---the 25 datasets in combination with the 4 constraint sets for each one of them--we can perform an empirical analysis of them.
+This way we can statistically determine which of the two methods offers the best results.
+
+In order to carry out this task we will make use the widely known frequentist non parametric two-tailed Wilcoxon test, as well as the lesser-known Bayesian sign test.
+
+\subsection{Applying Wilcoxon test to our results}
+
+\textcolor{red}{Buscar a alguien que sepa explicarme que cojones es lo que hace el Keel.
+Las hipótesis no pueden ser las que he planteado y problemente el alpha no se use para calcular el estadístico T}
+
+We will apply the Wilcoxon test for each of the measures obtained as a result, namely: ARI, Unsat and Time. Remember that the Wilcoxon test is a hypothesis test, whose hypotheses are established as:
+
+\begin{itemize}
+
+	\item $H_0$: differences between pairs of results follow a symmetric distribution around zero.
+
+	\item $H_1$: differences between pairs of results do not follow a symmetric distribution around zero.
+
+\end{itemize}
+
+It is also necessary to establish a confidence level to obtain the statistical values that will allow us to make a decision about $H_0$.
+In this case we consider a confidence level of 95\%, which means that the value of $\alpha$ for the calculation of the statistic $T$ will be $\alpha = 1 - 0.95 = 0.05$.
+Table \ref{tab:freqTest} shows the statistical values obtained using Wilcoxon test for each comparison.
+
+\begin{table}[!htp]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	\begin{tabular}{cccc}
+		\hline
+		VS & $R^{+}$ & $R^{-}$ & P-value \\ \hline 
+		ARI & 3203.5 & 1746.5 & 0.006936\\
+		Time & 4008.0 & 1042.0 & 0\\  
+		Unsat & 3829.5 & 1120.5 & 0.000002\\ \hline 
+	\end{tabular}
+	\caption{Results obtained by the Wilcoxon test for algorithm SHADE (vs BRKGA)}
+	\label{tab:freqTest}
+\end{table}
+
+It is worth noting that, in the case of ARI, this is a maximization problem, while in the remaining two measures we try to minimize.
+
+The results are clear, in all cases $p-value < \alpha$, which means that we can reject the null hypothesis.
+We conclude that there are significant differences between the two methods in favour of SHADE+LS.
+
+\clearpage
 

### 2019-02-07 Resultados/ResultadosDefinitivos/Tests/vignetes.rmd
Añadidos algoritmos TFG, resultados TFG y modificado paper SHADE

---
title: "Statistical Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Statistical Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = FALSE, error= TRUE)
```

```{r load.libraries, message=FALSE, warning=FALSE}
library(rNPBST)
library(dplyr)
library(scmamp)
library(xtable)
library(ggplot2)
```

# Parametric analysis

In this section we illustrate the application of the parametric tests
to the comparison of algorithms in the study case of the competition
of CEC2017. With this purpose, we use the results obtained at the end
of the 100\% of the steps and compute the mean among different
iterations of each algorithm on each benchmark function.

## Preliminars

The traditional parametric test used in the context of a comparison of
multiple algorithms over multiple problems (benchmarks) is the ANOVA
test. This test makes some assumptions that should be checked before
it is performed:
- The distribution of the results among different benchmarks is normal.
- The standard deviation of results is equal.

### Normality

```{r shapiro, echo=TRUE, results="asis"}
shapiro.test <- split(cec17.final, cec17.final$Dimension) %>%
   lapply(function(x) split(x, x$Algorithm) %>%
   lapply(function(y) shapiro.test(y$Result)$p.value)) %>%
   unlist() %>% matrix(ncol = 4, byrow = F) %>% as.data.frame()
AdjustFormatTable(shapiro.test, 
                  colnames = paste("Dim.", levels(cec17.final$Dimension)),
                  rownames = levels(cec17.final$Algorithm), print.code = TRUE,
                  digits = -2, type = "latex")
```

We can see that all the null hypotheses are rejected because the
associated p-values are less than $0.05$. This means that the mean
results of the algorithms among the different iterations do not follow
a normal distribution, as could be expected due to the different
difficulty of the multiple benchmark functions throug the dimensions.

### Homocedasticity

```{r levene}
levene.test <- split(cec17.final, cec17.final$Dimension) %>%
  lapply(function(x) lawstat::levene.test(x$Result,  x$Algorithm)$p.value) %>% 
  unlist() %>% matrix(ncol = 4, byrow = T)
AdjustFormatTable(levene.test, type ="latex", digits = -2, 
                  colnames = paste("Dim.", levels(cec17.final$Dimension)),
                  rownames = "p.value", caption = "Levene test for homocedasticity",
                  label = "tab:levene", print.code = TRUE)
```

Although normality is not satisfied, we check the homocedasticity for
the different dimensions. As the $p$-value are greater than $\alpha =
0.05$ we cannot reject the hypothesis of the equivalence of the
variances.

```{r t.test}
split(cec17.final, cec17.final$Dimension) %>%
   lapply(function(x){
      y <- t.test(select(filter(x, Algorithm == "EBO"), Result),
                  select(filter(x, Algorithm == "jSO"), Result))$p.value
   }) %>%
    unlist() %>%
    matrix(nrow = 1, dimnames = list("p.value", paste("Dim.", levels(cec17.final$Dimension)))) %>%
    AdjustFormatTable(caption = "EBO vs. jSO t-test results", label = "tab:t-test")
```

```{r anova}
split(cec17.final, cec17.final$Dimension) %>%
   lapply(function(x){
      y <- aov(Result ~ Algorithm, x) %>% summary()
      y[[1]][1,5]
   }) %>%
    unlist() %>%
    matrix(nrow = 1, dimnames = list("p.value", paste("Dim.", levels(cec17.final$Dimension)))) %>%
    AdjustFormatTable(caption = "ANOVA results", label = "tab:ANOVA")
```

We perform the ANOVA test and see how the equivalence of the means by
algorithms can not be rejected for any dimensions, a circumstance that
is consistent with the previous statements about the power of the
ANOVA test when the assumption of the test are not fulfilled.

# Non-parametric tests

In this section we perform the most popular tests in the field of the
comparison of optimization algorithms. We continue using the
aggregated results across the different runs at the end of the
iterations, except for the Page test for the study of the convergence.

## Classic tests

```{r}
ApplyNPPairwise <- function(data, test){
split(data, data$Dimension) %>%
   lapply(function(x){
      y <- test(matrix(c(unlist(select(filter(x, Algorithm == "EBO"), Result)),
                                              unlist(select(filter(x, Algorithm == "jSO"), Result))),
                                            ncol = 2))$p.value[3]
   }) %>%
    unlist
}
results.sign <- ApplyNPPairwise(cec17.final, rNPBST::binomialSign.test)
results.wilcoxon <- ApplyNPPairwise(cec17.final, rNPBST::wilcoxon.test)
results.wilcoxonRS <- ApplyNPPairwise(cec17.final, rNPBST::wilcoxonRankSum.test)
results.nppairwise <- matrix(c(results.sign, results.wilcoxon,results.wilcoxonRS),
                             nrow = 3, byrow = T,
                             dimnames = list(c("Sign","Wilcoxon", "Wilcoxon Rank-Sum"), 
                                             paste("Dim.", levels(cec17.final$Dimension)))) %>%
    AdjustFormatTable(caption = "Non-parametric pairwise comparison between EBO and jSO", label = "tab:np-pairwise", print.code = T)
```

```{r apply.np.test}
ApplyNPTest <- function(data, test){
    split(data, data$Dimension) %>%
    lapply(function(x)
        select(x, -Dimension) %>% 
        tidyr::spread(Algorithm, Result) %>%
        select(-Benchmark) %>%
        test %>% get("p.value", .)) %>%
    unlist
}
```

```{r np.test}
friedman.results <- ApplyNPTest(cec17.final, friedman.test)
friedman.ar.results <- ApplyNPTest(cec17.final, friedmanAR.test)
iman.davenport.results <- ApplyNPTest(cec17.final, imanDavenport.test)
quade.results <- split(cec17.final, cec17.final$Dimension) %>%
  lapply(function(x){
    x <- select(x, -Dimension)
    quade.test(x$Result, x$Algorithm, x$Benchmark) %>% get("p.value", .)
  }) %>%
  unlist(use.names = T)
```

```{r np.results}
non.parametric <- matrix(c(friedman.results, friedman.ar.results,
                           iman.davenport.results, quade.results),
                         byrow = T, ncol = 4)
AdjustFormatTable(non.parametric,
                  colnames = paste("Dim.", levels(cec17.final$Dimension)),
                  rownames = c("Friedman" ,"Friedman AR", "Iman-Davenport", "Quade"),
                  caption = "Non-Parametric tests",
                  label = "tab:np-tests")
```
We can see how the four tests reject the null hypotheses, it is, the
equivalence of the medians of the results of the different
benchmarks. We must have in mind that a comparison between thirteen
algorithms is not the recommended procedure if we want to compare our
proposal, but we should include in the comparison the state-of-the-art
algorithms, because the inclusion of an algorithm with lower
performance will be reflected in the rejection of the null hypothesis
while that would be not rejected if this bogus algorithm would not
have been included.

## Post-hoc tests

Then, we proceed to perform the post-hoc tests in order to determinate
the location of the differences between these algorithms. We use the
modification of the classic non-parametric tests to obtain the
$p$-value associated with each hypothesis, although we should adjust
these $p$-values with a post-hoc procedure.

### Control algorithm

To illustrate this, we first suppose that we are in an One versus all
scenario where we are presenting our algorithm (we are going to use
EBOwithCMAR, the winner of the CEC'2017 competition). The possible
approach here, as in the rest of the analysis are:

- Considering all the results in the different dimensions as if they
  were different function or benchmarks, we would only obtain a single
  $p$-value for the comparison between EBO-CMAR with each contestant
  algorithm. The adjusted $p$-values are shown in the following table
  for the Friedman, Friedman Aligned-Rank and Quade test.

```{r post.hoc}
cec17.final.10 <- filter(cec17.final, Dimension == 10)
ApplyNPTestPostHoc <- function(data, test, adjust, control = NULL){
    data %>%
        tidyr::spread(Algorithm, Result) %>%
        select(-Benchmark, -Dimension) %>%
        test(control = control) %>%
        adjust
}
results.post.friedman <- ApplyNPTestPostHoc(cec17.final, scmamp::friedmanPost,
                                           scmamp::adjustHolland, control = "EBO")
results.post.friedmanAR <- ApplyNPTestPostHoc(cec17.final, scmamp::friedmanPost,
                                             scmamp::adjustHolland, control = "EBO")
results.post.quade <- ApplyNPTestPostHoc(cec17.final, scmamp::quadePost, scmamp::adjustHolland,
                                        control = "EBO")
results.post.global.holland <- rbind(results.post.friedman,
                                    results.post.friedmanAR,
                                    results.post.quade)
AdjustFormatTable(as.data.frame(t(results.post.global.holland)),
                  colnames = c("Friedman","FriedmanAR","Quade"),
                  caption = "Post-Hoc Non-Parametric tests with control algorithm",
                  label = "tab:np-tests-ph-control", print.code = T)
```

- If we wanted to show that the differences between the algorithms
  also persist in each group of results obtained across the different
  dimensions, we should perform these tests repeatedly and apply later
  the appropiate post-hoc procedure. In the following table we show
  the adjusted $p$-value obtained.

```{r post.hoc.grouped}
ApplyNPTestPostHoc.Grouped <- function(data, test, adjust, control = NULL){
  results <- split(data, data$Dimension) %>%
    sapply(function(x)
      select(x, -Dimension) %>% 
        tidyr::spread(Algorithm, Result) %>%
        select(-Benchmark) %>%
        test(control = control)) %>%
    adjust
  return(results)
}
post.hoc.grouped <- ApplyNPTestPostHoc.Grouped(cec17.final, 
                                               scmamp::friedmanPost, 
                                               scmamp::adjustHolland,
                                              control = "EBO")
AdjustFormatTable(post.hoc.grouped,
                  colnames = paste("Dim", colnames(post.hoc.grouped)),
                  rownames = levels(cec17.final$Algorithm),
                  caption = "Results grouped by dimension. Friedman test + Holland adjust.",
                  label = "tab:grp-friedman-holland")
```

### $n$ versus $n$ scenario

In the scenario of the statistical analysis of the results obtained
during a competition, we would not focus in the comparison between the
results of a single algorithm but we would make all the posible pairs,
so we would not use the control algorithm.

As in the scenario with a control algorithm, we could be interested in
the  global comparison  between  all the  posible  pairs of  algorithm
considering the  results in the  different dimensions as if  they came
from the same distribution or pay attention to the differences between
the algorithms in the different contexts according to the dimension.

```{r all.vs.all}
ApplyNPTestPostHoc(cec17.final.10,
                   scmamp::friedmanPost,
                   scmamp::adjustHolland)  %>%
  AdjustFormatTable(caption = "Results $n$ vs $n$. Friedman test + Holland adjust.",
                    label = "tab:nvsn-friedman-holland", print.code = F)
```

```{r all.vs.all.group}
all.vs.all.group <- scmamp::postHocTest(select(tidyr::spread(cec17.final, Algorithm, Result),
                                             -Benchmark),
                                      group.by = "Dimension",
                                      test = "friedman", adjust = "holland")
all.vs.all.group$corrected.pval[, ,1] %>%
  AdjustFormatTable(caption = "Results $n$ vs $n$, dimension 10. Friedman test + Holland adjust.",
                    label = "tab:nvsn-dim10-friedman-holland")
# all.vs.all.group$corrected.pval[, ,3] %>%
#   AdjustFormatTable(caption = "Results $n$ vs $n$, dimension 30. Friedman test + Holland adjust.",
#                     label = "tab:nvsn-dim30-friedman-holland")
```

## Convergence test

For the comparison of the convergence of two algorithms, we use the
mean result acrossthe different runs for each benchmark function of
two algorithms. This results could be equally extended using the
convenient adjustments. It is relevant to note that the algorithms
EBOwithCMAR (first classified) and LSHADE-cnEpSin (third classified)
do not provide their partial results, so we compare here the
convergence of the jSO and LSHADE\_SPACMA algorithms in the
competition for the 10 dimension scenario, where the null hypothesis
of LSHADE\_SPACMA cannot be rejected and 50 dimension scenario.

```{r convergence}
convergence.comparison.lshade <- filter(cec17.mean,
                                Algorithm == "LSSPA",
                                Dimension == 10) %>%
    select(-Dimension)
convergence.comparison.jso <- filter(cec17.mean,
                                Algorithm == "jSO",
                                Dimension == 10) %>%
    select(-Dimension)
convergence.comparison <- rbind(convergence.comparison.jso,
                                convergence.comparison.lshade) %>%
  as.data.frame() %>% reshape2::melt(id.vars = c("Algorithm", "Benchmark"), value.name = "Result") %>%
  transform(Perc.Iterations = as.numeric(gsub("perc.completion.", "", variable))) %>% select(-variable)
convergence.comparison %>%
  filter(Benchmark %in% c(3, 9, 11, 15)) %>%
  ggplot(aes(x = Perc.Iterations, y = Result, col = Benchmark, linetype = Algorithm)) + 
  geom_line()
```

```{r}
plot(as.numeric((convergence.comparison.lshade - convergence.comparison.jso)[2, ]))
```

```{r}
convergence.comparison.jso <- select(convergence.comparison.jso, -Benchmark, -Algorithm)
convergence.comparison.lshade <- select(convergence.comparison.lshade, -Benchmark, -Algorithm)
cat(rNPBST::htest2Tex(rNPBST::page.test(-convergence.comparison.lshade+convergence.comparison.jso)))
```

```{r}
convergence.comparison.jso <- filter(cec17.mean,
                                Algorithm == "jSO",
                                Dimension == 100) %>%
    select(-Algorithm, -Benchmark, -Dimension)
convergence.comparison.dyypo <- filter(cec17.mean,
                                Algorithm == "DYYPO",
                                Dimension == 100) %>%
    select(-Algorithm, -Benchmark, -Dimension)
rNPBST::page.test(convergence.comparison.jso - convergence.comparison.dyypo)
rNPBST::page.test(-convergence.comparison.jso + convergence.comparison.dyypo)
```

## Confidence Intervals and Confidence Curves

In this Section we show the use of confidence intervals and confidence
curves in the comparison of optimization results. First, we have to
warn that these comparisons only take care of two algorithm at each
time, and a post-hoc correction would be needed if the comparison
involved a greater number of algorithms as we are going to see in
the following examples.

- We suppose that we want to compare the final and mean results of the
  two first classified algorithms in the benchmark function for the 10
  dimension. Then, we select in the dataset the corresponding data and
  obtain the following result.

```{r confidence.intervals}
conf.interval.PPSO <- filter(cec17.final,Algorithm == "PPSO",
                           Dimension == 10) %>%
    select(Result)
conf.interval.jSO <- filter(cec17.final,Algorithm == "jSO",
                           Dimension == 10) %>%
    select(Result)
```

As the 0 effect is not included in the Non-parametric confidence
interval, the null hypothesis is then rejected, although the interval
is very wide, so we have not much certainty about that. If we only had
done the Wilcoxon test, we would have obtained the associated
$p$-value, and the null hypothesis would be also rejected, so the
difference between the medians are detected with both methods. If we
look at the confidence curve, we can reject the classic null
hypothesis if the intersection of the vertical line at the 0 level and
the horizontal line at the $\alpha$ value remains under the curve.

```{r confidence.curve}
conf.interval.matrix <- cbind(conf.interval.PPSO, conf.interval.jSO)
colnames(conf.interval.matrix) <- c("PPSO", "jSO")
cat(rNPBST::htest2Tex(rNPBST::wilcoxon.test(conf.interval.matrix)))
```

```{r}
ApproximateNPConfidenceInterval(x = conf.interval.matrix[ ,1],
                                                           y = conf.interval.matrix[ ,2], paired = T)
PlotConfidenceCurve(x = conf.interval.matrix[ ,1],
                                               y = conf.interval.matrix[ ,2], m = 50, paired = T)
```

## Multiple measures test - GLRT

```{r multiple.measures.nonparametric}
matrix.EBO <-  filter(cec17.final, Algorithm == "EBO") %>%
    select(-Algorithm) %>% tidyr::spread(Dimension, Result) %>%
    select(-Benchmark)
matrix.jSO <-  filter(cec17.final, Algorithm == "jSO") %>%
    select(-Algorithm) %>% tidyr::spread(Dimension, Result) %>%
    select(-Benchmark)
rNPBST::multipleMeasuresGLRT(matrix.EBO, matrix.jSO) 
```

In order to perform the Non-Parametric Multiple Measures test, we
consider that the results of the different dimensions as if they were
different measures in the same benchmark function. We select the means
of the executions of the two best algorithms and reshape them into a
matrix with the results of each benchmark in the rows and the
different dimensions in the columns. Then we use the test to see which
hypothesis of dominance is the most probable and if we can state that
the probability of this dominance statement is significant. According
to the results, we obtain that the most observed dominance statement
in the third position (which corresponds to the number 2, 0010 in
binary) and the configuration $[<,<,>,<]$, it is, EBO-CMAR obtains a
better result for the dimensions 10, 30 and 100 while jSO obtains
better results in the 50 dimension scenario. However, the associated
null hypothesis which states that the mentioned configuration is no
more probable than the following one, obtain a $p$-value of 0.25, so
this hypothesis cannot be rejected. The second most probable
configuration is in the 15th position (or number 14) and the
configuration $[>,>,>,<]$, which means that EBO-CMAR obtains worse
results in dimensions 10,30 and 50, so we cannot assure which is the
most probable situation in the lower dimension scenarios.

# Bayesian tests

In this section we illustrate the use of the proposed Bayesian
tests. The considerations are analogous to the ones made in the
frequentist case, as the described tests use the aggregations of the
runs to compare the results of the different benchmark functions, or
use these runs with the drawback of obtaining a restrained statement
about the results in one single problem.

## Bayesian Friedman test

We start with the Bayesian version of the Friedman test. In this test
we do not obtain a single $p$-value, but the accepted hypothesis. Due
to the high number of contestant algorithms and the memory needed to
allocate the covariance matrix of all the possible permutations, we
perform here the imprecise version of the test. The null hypothesis of
the equivalence of the mean ranks for the 10 dimension scenario is
rejected.

```{r bayesian.friedman.test}
dim.10 <- filter(cec17.final, Dimension == 10) %>%
    tidyr::spread(Algorithm, Result) %>% select(-Benchmark, -Dimension)
bft.results <- bayesianFriedman.test(dim.10, imprecise = TRUE, n.samples = 200)
bft.results$h
meanranks <- bft.results$meanranks
names(meanranks) <- colnames(bft.results$covariance)
xtable::xtable(as.matrix(meanranks))
```

## Bayesian Sign and Signed-Rank test

The originial proposal of the use of the Bayesian Sign and Signed-Rank
tests are destinated to the comparison of classification algorithms
and the proposed region of practical equivalence is $[-0.01,0.01]$ for
a measure in the range $[0,1]$. In the scenario of optimization
problems, we should be concern that the possible outcomes are
lower-bounded by 0 but in many functions there is not an upper bound
or the maximum is very high so we must follow another approach. As the
difference in the 30 dimension comparison is between 0 and 1527, we
state that the region of practical equivalence is $[-10,10]$.

```{r bayesian.pair.test}
jso <- filter(cec17.final, Algorithm == "jSO", Dimension == 30) %>%
    select(Result) %>% unlist()
ebo.cmar <- filter(cec17.final, Algorithm == "EBO-CMAR", Dimension == 30) %>%
    select(Result) %>% unlist()
bst.results <- rNPBST::bayesianSign.test(ebo.cmar, jso,
                                       rope.min = -10, rope.max = 10)
bsrt.results <- rNPBST::bayesianSignedRank.test(ebo.cmar, jso,
                                              rope.min = -10, rope.max = 10)
```
The test compute the probability of the true location of
$\textrm{EBO-CMAR} - \textrm{jSO}$ with respect to $0$, so both tests'
results gives that the left region (jSO obtain worse results) a
greater probability, although this is not significative. Rope
probability is low in comparison with the right and left region, and
this is because, although the number of wins and losses is similar for
both algorithms, the number of benchmark functions where the results
are quite similar is lower.

We can see the posterior probability density of the parameter in the
following plots, where each point represents an estimation of the
probability of the parameter to belong to each region of
interest. This means that we have repeatedly obtained the triplets of
the probability of each region to be the true location of the
difference between the two samples and then we have plotted these
triplets to obtain the posterior distribution of the parameter. If we
compare these results with a paired Wilcoxon test we see that the null
hypothesis of the equivalence of the means cannot be
rejected. However, using the Bayesian paradigm we cannot establish the
dominance of one algorithm over the other either, but while in the
frequentist paradigm we could be tempted to (erroneously) establish
that there is no difference between these algorithms with these plots
we obtain valuable information and we can see that here is not the same
situation, as the probability of the rope is very low.

```{r plot.pair.bayesian.tests}
wilcox.test(jso, ebo.cmar, paired = T)
plotSimplex(bst.results$sample, num.points = 10000) 
plotSimplex(bsrt.results$sample, num.points = 10000)
```

## Imprecise Dirichlet Process Test

The Imprecise Dirichlet Process consists in a more complex idea of the
previous tests although the implications of the use of the Bayesian
Tests could be clearer. In this test we try to not to introduce any
prior distribution, not even the prior distribution where both
algorithms have the same performance, but all the possible probability
measures $\alpha$, and then obtain an upper and a lower bounds for the
probability in which we are interested. The input consits in the
aggregated results among the different runs for all the benchmark
functions for a single dimension. The other parameters of the function
are the strong parameter $s$ and the pseudo-observation $c$. With
these data, we obtain two bounds for the posterior probability of the
first algorithm outperforming the second one, i.e. the probability of
$P(X \leq Y) \geq 0.5$. These are the possible scenarios:

- Both bounds are greater than 0.95: Then we can say that the first
  algorithm outperforms the second algorithm with a 95\% of
  probability.
- Both bounds are lower than 0.05: This is the inverse case. In this
  situation, the second algorithm outperforms the first algorithm with
  a 95\% of probability.
- Both bounds are between 0.05 and 0.95: Then we can say that the
  probability of one algorithm outperforming the other is lower than
  the desired probability of 0.95.
- Finally, if only one of the bounds is greater than 0.95 or lower
  than 0.05, the situation is indetermined and we cannot decide.

```{r bayesian.imprecise}
library(dplyr)
library(rNPBST)
jso <- filter(cec17.final, Algorithm == "jSO", Dimension == 50) %>%
    select(Result) %>% unlist()
ebo.cmar <- filter(cec17.final, Algorithm == "EBO", Dimension == 50) %>%
    select(Result) %>% unlist()
bayesian.imprecise.results <- bayesian.imprecise(jso, ebo.cmar)
plot(bayesian.imprecise.results)
```

According to the results of the Bayesian Imprecise Dirichlet Process,
the probability $\mathcal{P}[P(EBO-CMAR \leq jSO) \geq 0.5]$, that is
the probability of EBO-CMAR outperforming jSO, is between 0.92 and
0.96, so there is an indetermination and we cannot say that there is 
a probability greater than 0.95 of EBO-CMAR outperforming jSO, 
nor there is not such probability. These numbers represent
the area under the curve of the upper and lower distributions when
$P(X \leq Y) \geq 0.5$

## Bayesian Multiple Measures Test

In the Bayesian version of the Multiple Measures test the results are
analogous to the frequentist version. The most probable configuration
is also the dominance of the EBO-CMAR in the 10,30 and 100 dimensions
while the algorithm jSO obtains better results in the 50 dimension
according to this test, but the posterior probability is 0.841, so we
cannot say that the difference with respect the remaining
configurations is determinative.

```{r bayesian.multiple}
bayesianmultipleconditions.results <- rNPBST::bayesianMultipleConditions.test(a,b)
bayesianmultipleconditions.results$probabilities
```

# Appendix

```{r}
base.data <- cec17.final %>%
  filter(Dimension == 10) %>%
  select(-Dimension) %>%
  reshape(idvar = "Benchmark", timevar = "Algorithm", direction = "wide") %>%
  transform(Benchmark = as.integer(levels(Benchmark))[Benchmark]) %>% arrange(Benchmark)
colnames(base.data)[-1] <- levels(cec17.final$Algorithm)
knitr::kable(base.data)
```

### 2019-02-07 Resultados/ResultadosDefinitivos/Tests/combine.r

@@ -1,0 +1,12 @@
+
+ARI <- read.csv("ARI.csv", dec = ",", row.names = 1)
+Unsat <- read.csv("Unsat.csv", dec = ",", row.names = 1)
+Time <- read.csv("Time.csv", dec = ",", row.names = 1)
+
+colnames(ARI) <- c("BRKGA", "SHADE")
+colnames(Unsat) <- c("BRKGA", "SHADE")
+colnames(Time) <- c("BRKGA", "SHADE")
+
+write.csv(ARI, "ARI.csv", row.names = T, col.names = T, sep = ",")
+write.csv(Unsat, "Unsat.csv", row.names = T, col.names = T, sep = ",")
+write.csv(Time, "Time.csv", row.names = T, col.names = T, sep = ",")

### 2019-02-07 Resultados/ResultadosDefinitivos/Tests/testBayesianos.r

@@ -1,11 +1,43 @@
 library(rNPBST)
 
-installed <- rownames(installed.packages())
-pkgs <- dir("~/R/x86_64-pc-linux-gnu-library/3.4")
-new <- setdiff(pkgs, installed)
-new
-install.packages(new)
+ARI <- read.csv("ARI.csv", dec = ".", row.names = 1)
+Unsat <- read.csv("Unsat.csv", dec = ".", row.names = 1)
+Time <- read.csv("Time.csv", dec = ".", row.names = 1)
 
-if (!requireNamespace("BiocManager", quietly = TRUE))
-  install.packages("BiocManager")
-BiocManager::install("graph", version = "3.8")
+#n.samples = paramatero para el numero de puntitos
+
+ari.test = rNPBST::bayesianSign.test(ARI$BRKGA, ARI$SHADE)
+unsat.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHADE)
+time.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHADE)
+
+ari.test$probabilities
+unsat.test$probabilities
+time.test$probabilities
+
+pdf(file = "BayesSignoARI.pdf", width= 6, height = 4, useDingbats=F)
+plot(ari.test)
+dev.off()
+
+pdf(file = "BayesSignoUnsat.pdf", width= 6, height = 4, useDingbats=F)
+plot(unsat.test)
+dev.off()
+
+pdf(file = "BayesSignoTime.pdf", width= 6, height = 4, useDingbats=F)
+plot(time.test)
+dev.off()
+
+ari.test = rNPBST::bayesianSignedRank.test(ARI$BRKGA, ARI$SHADE)
+unsat.test = rNPBST::bayesianSignedRank.test(Unsat$SHADE, Unsat$BRKGA)
+time.test = rNPBST::bayesianSignedRank.test(Unsat$SHADE, Unsat$BRKGA)
+
+plot(ari.test)
+plot(unsat.test)
+plot(time.test)
+
+ari.test = rNPBST::bayesian.imprecise(ARI$BRKGA, ARI$SHADE)
+unsat.test = rNPBST::bayesian.imprecise(Unsat$SHADE, Unsat$BRKGA)
+time.test = rNPBST::bayesian.imprecise(Unsat$SHADE, Unsat$BRKGA)
+
+plot(ari.test)
+plot(unsat.test)
+plot(time.test)

### 2019-02-07 Software/TFG_COPKM.py

@@ -24,7 +24,7 @@ def main():
 
     #BUCLE DE OBTENCION DE DATOS
 
-    nb_runs = 1
+    nb_runs = 5
 
     for i in range(len(names_array)):
 
@@ -43,7 +43,7 @@ def main():
             copkm_assignment, wine_centroid = COPKM(data_set, nb_clust, const, max_iter = 300)
 
             if copkm_assignment == None:
-            	mean_ars = -1
+            	mean_ars += -1
             else:
             	mean_ars += adjusted_rand_score(labels, copkm_assignment)
 

### 2019-02-07 Software/TFG_LCVQE.py

@@ -37,7 +37,7 @@ def main():
 
     #BUCLE DE OBTENCION DE DATOS
 
-    nb_runs = 1
+    nb_runs = 5
 
     for i in range(len(names_array)):
 

### 2019-02-07 Software/TFG_RDPM.py

@@ -23,7 +23,7 @@ def main():
 
 	#BUCLE DE OBTENCION DE DATOS
 
-	nb_runs = 1
+	nb_runs = 5
 
 	for i in range(len(names_array)):
 
@@ -36,11 +36,11 @@ def main():
 		const = const_array[0][i]
 		T = np.mean(data_set, 0)
 		lamb_arr = np.array(np.sqrt(np.sum((np.matlib.repmat(T, np.shape(data_set)[0], 1) - data_set) ** 2, 1)))
+		lamb_arr[::-1].sort()
 		mean_ars = 0
 
 		for j in range(nb_runs):
-			#Mirar como se hacia la eleccion del lambda
-			rdpm_assignment, rdpm_nbc = RDPM(data_set, lamb_arr[2], const, 20000, 0.1, 1)
+			rdpm_assignment, rdpm_nbc = RDPM(data_set, lamb_arr[nb_clust - 1], const, 20000, 0.1, 1)
 			mean_ars += adjusted_rand_score(labels, rdpm_assignment)
 
 		mean_ars /= nb_runs

### 2019-02-07 Software/TFG_TVClust.py

@@ -33,7 +33,7 @@ def main():
 
     #BUCLE DE OBTENCION DE DATOS
 
-    nb_runs = 5
+    nb_runs = 1
 
     for i in range(len(names_array)):
 

### 2019-02-11 Papers/SHADE/paperSHADE.tex
Nuevos resultados y secciones en paper SHADE

@@
 
+\usepackage{graphics, float, url}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{lineno,hyperref}
@@
 \usepackage{amsfonts}
+\usepackage{subcaption}
+\usepackage{float}
 \usepackage[ruled, lined, onelanguage, linesnumbered]{algorithm2e}
@@
 
-In order to carry out this task we will make use the widely known frequentist non parametric two-tailed Wilcoxon test, as well as the lesser-known Bayesian sign test.
-
-\subsection{Applying Wilcoxon test to our results}
-
-\textcolor{red}{Buscar a alguien que sepa explicarme que cojones es lo que hace el Keel.
-Las hipótesis no pueden ser las que he planteado y problemente el alpha no se use para calcular el estadístico T}
-
-We will apply the Wilcoxon test for each of the measures obtained as a result, namely: ARI, Unsat and Time. Remember that the Wilcoxon test is a hypothesis test, whose hypotheses are established as:
+In order to carry out this task we will make use the widely known frequentist non-parametric Wilcoxon's test, as well as the lesser-known Bayesian sign test.
+
+\subsection{Applying Wilcoxon's test to our results}
+
+We will apply the Wilcoxon's test for each of the measures obtained, namely: ARI, Unsat and Time. Remember that the Wilcoxon's test is a hypothesis test, whose hypotheses are established as:
 
@@
 
-	\item $H_0$: differences between pairs of results follow a symmetric distribution around zero.
-
-	\item $H_1$: differences between pairs of results do not follow a symmetric distribution around zero.
+	\item $H_0$: differences between pairs of results follow a symmetric distribution around $\mu$.
+
+	\item $H_1$: differences between pairs of results do not follow a symmetric distribution around $\mu$.
 
@@
 
-It is also necessary to establish a confidence level to obtain the statistical values that will allow us to make a decision about $H_0$.
-In this case we consider a confidence level of 95\%, which means that the value of $\alpha$ for the calculation of the statistic $T$ will be $\alpha = 1 - 0.95 = 0.05$.
-Table \ref{tab:freqTest} shows the statistical values obtained using Wilcoxon test for each comparison.
+In this case we set $\mu = 0$, and work with a confidence level of 95\%, which means that the significance level is 0.05 ($\alpha = 0.05$).
+Table \ref{tab:freqTest} shows the statistical values obtained using Wilcoxon's test for each comparison.
 
@@
 	\renewcommand{\arraystretch}{1.4}
-	\begin{tabular}{cccc}
-		\hline
-		VS & $R^{+}$ & $R^{-}$ & P-value \\ \hline 
-		ARI & 3203.5 & 1746.5 & 0.006936\\
-		Time & 4008.0 & 1042.0 & 0\\  
-		Unsat & 3829.5 & 1120.5 & 0.000002\\ \hline 
+	\begin{tabular}{cccccc}
+		\hline
+		\multirow{2}{*}{Measure} & \multirow{2}{*}{$R^{+}$} & \multirow{2}{*}{$R^{-}$} & \multirow{2}{*}{P-value}
+		&  \multicolumn{2}{c}{Median} \\
+		\cline{5-6}
+		& & & & BRKGA + LS & SHADE + LS \\
+		\hline
+		ARI & 3088 & 1665 & 0.01051 & 0.30356 & 0.32573\\
+		Time & 4008 & 1042 & 0.000000344 & 1204.967 & 1202.947 \\  
+		Unsat & 3082.5 & 745.5 & 0.000000767 & 5.851 & 3.072 \\ 
+		\hline 
 	\end{tabular}
-	\caption{Results obtained by the Wilcoxon test for algorithm SHADE (vs BRKGA)}
+	\caption{Results obtained by the Wilcoxon's test for algorithm SHADE (vs BRKGA)}
 	\label{tab:freqTest}
@@
 
-It is worth noting that, in the case of ARI, this is a maximization problem, while in the remaining two measures we try to minimize.
-
-The results are clear, in all cases $p-value < \alpha$, which means that we can reject the null hypothesis.
-We conclude that there are significant differences between the two methods in favour of SHADE+LS.
+For all three considered measures, $\alpha < P-value$, therefore in all cases we can reject $H_0$ and find that there are significant differences between the two methods.
+To choose the method that gets the lead we observe the medians, we choose the one with greater or lesser median depending on whether it is a matter of maximization or minimization.
+In conclusion, we can safely admit that Wilcoxon's test provides empirical evidence in favour of SHADE + LS.
+
+\subsection{Applying Bayesian sign test to our results}
+
+Te Bayesian sign test is the Bayesian version of the frequentist non-parametric tests.
+It allows us to conclude whether there are significant differences between two sets of results.\cite{carrasco2017rnpbst}
+
+It is based on deriving the distribution of a parameter $\rho$ according to the difference between the sets of results.
+We will refer to the set of results obtained by BRKGA - LS as $A$, and to the set of results obtained with SHADE + LS as $B$.
+It is also necessary to define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, for those cases in which the difference between $A$ and $B$ is not significant.
+
+To get the distribution of $\rho$ we count the number of times the advantage is for $A$, the number of times where there are no significant differences, and the number of times the advantage is for $B$.
+We then calculate the weights of a Dirichlet distribution and sample it.
+The result of this sampling will be a set of triplets in the form of $[(\rho < r_{min}), (\rho \in rope), (\rho > r_{max}]$.
+
+\textcolor{red}{Traducir el párrafo de abajo al inglés}
+
+Podemos obtener una representación del conjunto de tripletas en forma de un diagrama de calor.
+Para ello, hacemos corresponder los vértices de un triángulo equilátero con los tres valores de una tripleta.
+Para representar un valor, trazamos una linea paralela al lado opuesto del vértice asociado al valor que queremos representar, a una distancia relativa de el calculada en base al valor cuestión.
+El punto en el que intersecan las líneas asociadas a los tres valores es el punto asociado a la tripleta que los contiene.
+Cabe destacar que, dado que los valores de cada tripleta representan una distribución de probabilidad, su suma siempre vale 1, y por tanto siempre existe un punto de intersección entre las líneas que los representan.
+
+\begin{figure}[ht!]
+	\centering
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/BayesSignoARI.pdf}
+		\caption{ARI}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/BayesSignoTime.pdf}
+		\caption{Time}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/BayesSignoUnsat.pdf}
+		\caption{Usat}
+	\end{subfigure}
+	\caption{Heat diagrams for each measure}
+\end{figure}
 
@@
 \end{document}
+
+\begin{figure}[bth]
+	\myfloatalign
+	%\subfloat[ARI]
+	{\includegraphics[width=.45\linewidth]{Figures/BayesSignoARI.pdf}} \quad
+	%\subfloat[Time]
+	{\includegraphics[width=.45\linewidth]{Figures/BayesSignoTime.pdf}} \quad
+	%\subfloat[Unsat]
+	{\includegraphics[width=.45\linewidth]{Figures/BayesSignoUnsat.pdf}} \quad
+	\caption{Heat diagrams for each measure obtained.}
+	\label{img:heatDiag}
+\end{figure}

### 2019-02-11 Resultados/ResultadosDefinitivos/Tests/combine.r

@@ -1,12 +1,19 @@
+library(dplyr)
 
-ARI <- read.csv("ARI.csv", dec = ",", row.names = 1)
-Unsat <- read.csv("Unsat.csv", dec = ",", row.names = 1)
-Time <- read.csv("Time.csv", dec = ",", row.names = 1)
+ARI <- read.csv("ARI.csv", dec = ".", row.names = 1)
+Unsat <- read.csv("Unsat.csv", dec = ".", row.names = 1)
+Time <- read.csv("Time.csv", dec = ".", row.names = 1)
 
 colnames(ARI) <- c("BRKGA", "SHADE")
 colnames(Unsat) <- c("BRKGA", "SHADE")
 colnames(Time) <- c("BRKGA", "SHADE")
 
-write.csv(ARI, "ARI.csv", row.names = T, col.names = T, sep = ",")
-write.csv(Unsat, "Unsat.csv", row.names = T, col.names = T, sep = ",")
-write.csv(Time, "Time.csv", row.names = T, col.names = T, sep = ",")
+test.ari = wilcox.test(ARI$SHADE, ARI$BRKGA, paired = T)
+test.unsat = wilcox.test(Unsat$SHADE, Unsat$BRKGA, paired = T)
+test.time = wilcox.test(Time$SHADE, Time$BRKGA, paired = T)
+
+median(ARI$BRKGA)
+
+#write.csv(ARI, "ARI.csv", row.names = T, col.names = T, sep = ",")
+#write.csv(Unsat, "Unsat.csv", row.names = T, col.names = T, sep = ",")
+#write.csv(Time, "Time.csv", row.names = T, col.names = T, sep = ",")

### 2019-02-11 Resultados/ResultadosDefinitivos/Tests/testBayesianos.r

@@ -6,9 +6,9 @@ Time <- read.csv("Time.csv", dec = ".", row.names = 1)
 
 #n.samples = paramatero para el numero de puntitos
 
-ari.test = rNPBST::bayesianSign.test(ARI$BRKGA, ARI$SHADE)
-unsat.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHADE)
-time.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHADE)
+ari.test = rNPBST::bayesianSign.test(ARI$BRKGA, ARI$SHADE, n.samples = 2000)
+unsat.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHAD, n.samples = 2000)
+time.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHADE, n.samples = 2000)
 
 ari.test$probabilities
 unsat.test$probabilities

### 2019-02-12 Papers/SHADE/paperSHADE.tex
Completadas secciones de analisis de resultados

@@
 Revisar que es siempre instance-level y no instance level.
-Comprobar que es siempre dataset y no data set}
+Comprobar que es siempre dataset y no data set.
+Puedo decir en alguna parte que se hace una iterpretación relajada de las restricciones}
 
@@
 where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
-It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that an $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing the obtained partitions with random partitions.
+It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that an $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing with random partitions.
 To summarize, the higher the $ARI$, the greater the degree of similarity between $C_1$ and $C_2$.
@@
 			$p_{inherit}$ & Probability that a feature is inherit from an elite parent & $50\%$ & - \\
-			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Known for each dataset, it has the same value for both methods} \\
+			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Known for each dataset (Table \ref{tab:datasets}. It has the same value for both methods} \\
 			\hline
@@
 
-The ARI column shows the Adjusted Rand Index for each case, the Unsat column displays the percentage of violated constraints and the Time columns shows the time---measured in seconds---taken for each algorithm to deliver results.
+Since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
+To lessen the effect this may have on the results, we will apply each method 5 times to each dataset and set of constraints, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+
+The ARI column shows the Adjusted Rand Index for each case, the Unsat column displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
 
@@
 While conclusions for Unsat and Time remain unchanged, we see a clear difference as far as ARI is concerned.
-We observe that, for constraint sets obtained with 15\% labeled data, both methods already report ARI results above 0.7, which in an ARI ranking is considered to be a good result.
+We observe that, for constraints sets obtained with 15\% labeled data, both methods already report ARI results above 0.7, which in an ARI ranking is considered to be a good result.
 We can even observe that the SHADE + LS method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA + LS that only obtains similar results in 2 datasets.
@@
 
-It is in Table \ref{tab:results20}, which gathers results obtained with constraint sets generated from 20\% of labeled data for each dataset, that we observe the major differences between the two methods.
+It is in Table \ref{tab:results20}, which gathers results obtained with constraints sets generated from 20\% of labeled data for each dataset, that we observe the major differences between the two methods.
 We found that SHADE + LS is capable of producing clusters identical to the original ones for 4 datasets, against the single dataset for which BRKGA + LS obtains a similar result.
@@
 
-With the results obtained by both methods for a total of 100 different datasets---the 25 datasets in combination with the 4 constraint sets for each one of them--we can perform an empirical analysis of them.
+With the results obtained by both methods for a total of 100 different datasets---the 25 datasets in combination with the 4 constraints sets for each one of them--we can perform an empirical analysis of them.
 This way we can statistically determine which of the two methods offers the best results.
 
-In order to carry out this task we will make use the widely known frequentist non-parametric Wilcoxon's test, as well as the lesser-known Bayesian sign test.
-
-\subsection{Applying Wilcoxon's test to our results}
-
-We will apply the Wilcoxon's test for each of the measures obtained, namely: ARI, Unsat and Time. Remember that the Wilcoxon's test is a hypothesis test, whose hypotheses are established as:
-
-\begin{itemize}
-
-	\item $H_0$: differences between pairs of results follow a symmetric distribution around $\mu$.
-
-	\item $H_1$: differences between pairs of results do not follow a symmetric distribution around $\mu$.
-
-\end{itemize}
+In order to carry out this task we will make use the widely known frequentist non-parametric Wilcoxon's signed rank test, as well as the lesser-known Bayesian sign test.
+
+From now on, and as far as this section is concerned, we will refer to the results obtained by BRKGA - LS as sample $A$, and to the results obtained with SHADE + LS as sample $B$.
+
+\subsection{Applying Wilcoxon's signed rank test to our results}
+
+We will apply the Wilcoxon's signed rank test for each of the measures obtained, namely: ARI, Unsat and Time.
+We set as null hypothesis $H_0 = $ the distribution of $A - B$ is symmetric about $\mu$.
+However, given that in the case of ARI it is a problem of maximisation, and in the case of Time and Unsat it is a problem of minimisation, we will consider a different alternative hypothesis for each case.
+In the case of maximization we propose $H_{max} =$ the distribution of $A - B$ is shifted to the left with respect to $\mu$.
+On the other hand, in the case of minimization we propose $H_{min} =$ the distribution of $A - B$ is shifted to the right with respect to $\mu$.
+In this way we assume that SHADE + LS is better than BRKGA + LS and we try to prove it.
 
@@
 	\renewcommand{\arraystretch}{1.4}
-	\begin{tabular}{cccccc}
-		\hline
-		\multirow{2}{*}{Measure} & \multirow{2}{*}{$R^{+}$} & \multirow{2}{*}{$R^{-}$} & \multirow{2}{*}{P-value}
-		&  \multicolumn{2}{c}{Median} \\
-		\cline{5-6}
-		& & & & BRKGA + LS & SHADE + LS \\
-		\hline
-		ARI & 3088 & 1665 & 0.01051 & 0.30356 & 0.32573\\
-		Time & 4008 & 1042 & 0.000000344 & 1204.967 & 1202.947 \\  
-		Unsat & 3082.5 & 745.5 & 0.000000767 & 5.851 & 3.072 \\ 
+	\begin{tabular}{ccccc}
+		\hline
+		Measure & $R^{+}$ & $R^{-}$ & P-value & Alternative hypothesis  \\
+		\hline
+		ARI & 3088 & 1665 & 0.005254 & $H_{max}$ \\
+		Time & 4008 & 1042 & 0.0000001723 & $H_{min}$ \\  
+		Unsat & 3082.5 & 745.5 & 0.0000003838 & $H_{min}$ \\ 
 		\hline 
 	\end{tabular}
-	\caption{Results obtained by the Wilcoxon's test for algorithm SHADE (vs BRKGA)}
+	\caption{Results obtained by the Wilcoxon's signed rank test for the comparison SHADE vs BRKGA}
 	\label{tab:freqTest}
@@
 
-For all three considered measures, $\alpha < P-value$, therefore in all cases we can reject $H_0$ and find that there are significant differences between the two methods.
-To choose the method that gets the lead we observe the medians, we choose the one with greater or lesser median depending on whether it is a matter of maximization or minimization.
-In conclusion, we can safely admit that Wilcoxon's test provides empirical evidence in favour of SHADE + LS.
+For all three considered measures, $P-value < \alpha$, therefore in all cases we can reject $H_0$ and find that there are significant differences between the two methods, giving the adavantage to SHADE + LS with basis to the proposed alternative hypothesis.
+In conclusion, we can safely admit that Wilcoxon's signed rank test provides empirical evidence in favour of SHADE + LS.
 
@@
 
-Te Bayesian sign test is the Bayesian version of the frequentist non-parametric tests.
+Te Bayesian sign test is the Bayesian version of the frequentist non-parametric sign test.
 It allows us to conclude whether there are significant differences between two sets of results.\cite{carrasco2017rnpbst}
 
-It is based on deriving the distribution of a parameter $\rho$ according to the difference between the sets of results.
-We will refer to the set of results obtained by BRKGA - LS as $A$, and to the set of results obtained with SHADE + LS as $B$.
-It is also necessary to define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, for those cases in which the difference between $A$ and $B$ is not significant.
-
-To get the distribution of $\rho$ we count the number of times the advantage is for $A$, the number of times where there are no significant differences, and the number of times the advantage is for $B$.
-We then calculate the weights of a Dirichlet distribution and sample it.
+It is based on obtaining the $\rho$ parameter of a statistical distribution according to the difference between the results.
+To get the distribution of $\rho$ we count the number of times that $A - B<0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
+In order to identify cases where there are no significant differences we define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, so that $P(A = B) = P(\rho \in rope)$.
+Based on these results we calculate the weights of a Dirichlet distribution and sample it.
 The result of this sampling will be a set of triplets in the form of $[(\rho < r_{min}), (\rho \in rope), (\rho > r_{max}]$.
 
-\textcolor{red}{Traducir el párrafo de abajo al inglés}
-
-Podemos obtener una representación del conjunto de tripletas en forma de un diagrama de calor.
-Para ello, hacemos corresponder los vértices de un triángulo equilátero con los tres valores de una tripleta.
-Para representar un valor, trazamos una linea paralela al lado opuesto del vértice asociado al valor que queremos representar, a una distancia relativa de el calculada en base al valor cuestión.
-El punto en el que intersecan las líneas asociadas a los tres valores es el punto asociado a la tripleta que los contiene.
-Cabe destacar que, dado que los valores de cada tripleta representan una distribución de probabilidad, su suma siempre vale 1, y por tanto siempre existe un punto de intersección entre las líneas que los representan.
+We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point.
+With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle.
+In order to find out where a certain triplet will be placed within the triangle, we will take each of its three values and draw a parallel line to the opposing side of the corresponding vertex; the separation between a triangle side and its parallel line will be proportional to the associated triplet value, so that the higher the value, the closer the line will be to the vertex.
+The location where the three lines intersect is where we draw a point.
+Since the values of every triplet describe a probability distribution and therefore they must add up to one, we can be sure that all triplets will lie in some point within the triangle.
+The color indicates the density of points in a given region, with yellow representing a high density and red a low density.
 
@@
 		\caption{ARI}
+		\label{fig:bayesARI}
 	\end{subfigure}
@@
 		\caption{Time}
+		\label{fig:bayesTime}
 	\end{subfigure}
@@
 		\includegraphics[width=1\linewidth]{Figures/BayesSignoUnsat.pdf}
-		\caption{Usat}
+		\caption{Unsat}
+		\label{fig:bayesUnsat}
 	\end{subfigure}
@@
 \end{figure}
+
+Figure \ref{fig:bayesARI} shows the heatmap associated with the ARI measurement.
+The diagram suggest that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle.
+It also indicates that in some cases there are no significant differences between $A$ and $B$, since we also find triplets in the rope area.
+Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE + LS performs better than BRKGA + LS.
+
+On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} are very similar, they show the heatmaps obtained for Time and Unsat measurements.
+In both, all triplets are represented in the lower right third of the triangle, indicating that the Bayesian sign test assigns a high probability to $A - B > 0$.
+With this, and bearing in mind that both measures must be minimized, we conclude that SHADE + LS represents a solid improvement over BRKGA + LS as far a this measures are concerned.
 

### 2019-02-13 Resultados/ResultadosDefinitivos/Tests/freqTest.r
Salvando nuevos resultados

@@ -8,9 +8,10 @@ colnames(ARI) <- c("BRKGA", "SHADE")
 colnames(Unsat) <- c("BRKGA", "SHADE")
 colnames(Time) <- c("BRKGA", "SHADE")
 
-test.ari = wilcox.test(ARI$SHADE, ARI$BRKGA, paired = T)
-test.unsat = wilcox.test(Unsat$SHADE, Unsat$BRKGA, paired = T)
-test.time = wilcox.test(Time$SHADE, Time$BRKGA, paired = T)
+#https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/wilcox.test
+test.ari = wilcox.test(x = ARI$BRKGA, y = ARI$SHADE, paired = T, alternative = "l")
+test.unsat = wilcox.test(Unsat$BRKGA, Unsat$SHADE, paired = T, alternative = "g")
+test.time = wilcox.test(Time$BRKGA, Time$SHADE, paired = T, alternative = "g")
 
 median(ARI$BRKGA)
 

### 2019-02-14 Papers/SHADE/paperSHADE.tex
Escrita introducción para SHADE

@@
 Comprobar que es siempre dataset y no data set.
-Puedo decir en alguna parte que se hace una iterpretación relajada de las restricciones}
+Puedo decir en alguna parte que se hace una iterpretación relajada de las restricciones.
+Tengo que decidir como llamar a la funcion de coste, "cost", "fitness" o "heuristic"}
 
@@
 
-\textcolor{red}{Citar a los de las hormigas. Poner un parrafo de aplicaciones del clustering con restricciones}
-
-\section{Background}
-
-%\subsection{Constrained Clustering}
-
-Clustering is the task of grouping instances of a dataset into subsets.
+One of the most widely known and studied data analysis tools is clustering.
+It is one of the most successful methods that has been able to provide solutions in a large number of knowledge fields, such as medicine, marketing, banking or psychology, among others.
+
+Clustering is the task of grouping instances of a dataset into subsets, so that new information can be extracted from them.
 The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset.
 In this way, the clustering process obtains subsets of instances of the initial dataset that present high intra-similarity and low inter-similarity.
+
+It is well known that the clustering problem is $\mathbf{NP}$-hard, therefore finding the optimal solution is not feasible with the technology we have today.
+However, there are many approximated resolution methods, including constrained clustering.
+
+Clustering with constraints is a semi-supervised learning method.
+Its goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraints set.
+In \cite{davidson2007survey} we find two constraint models, as well as an extensive review of some resolution methods.
+This study focuses specifically on instance-based constraints, which will be defined in Section \ref{sec:background}.
+
+Some of the constrained clustering applications mentioned in \cite{davidson2007survey} include the application of this technique to data obtained from image, video, audio, text, etc.
+
+It should be noted that, as we will study in more depth in Section \ref{sec:ugly}, the incorporation of constraints to clustering modifies its complexity, making it harder.
+The SHADE algorithm has proven to be one of the most successful methods to solve real-domain problems in an approximate way.
+The aim of this paper is to propose an adaptation of it to solve clustering with restrictions.
+
+There have been previous attempts to adapt population-based methods to constrained clustering, in particular in \cite{de2017comparison} an adaptation of the BRKGA algorithm is proposed.
+There are also some adaptations of swarm intelligence based methods, such as the one presented in \cite{xu2013improving}.
+As far as this paper is concerned, we will use some of the techniques presented in \cite{de2017comparison} to adapt SHADE to constrained clustering, as well as we propose a new cost function to improve the results.
+We will properly compare the results obtained by both methods.
+On the other hand we will check whether the population-based methods represent an improvement over some of the non ehuristic-based methods, presented in \cite{davidson2007survey} and briefly discussed in this paper in Section \ref{sec:NHres}.
+
+\textcolor{red}{Incluir un párrafo en el que se comete la organización del paper?}
+
+\section{Background} \label{sec:background}
+
+%\subsection{Constrained Clustering}
 
@@
 A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster.
-**Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms**.
-[It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms]
+It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms
 
@@
 
-\subsection{The Feasibility Problem}
+\subsection{The Feasibility Problem} \label{sec:feas}
 
@@
 
-\subsection{(Classic) Resolution Methods}
-
-\textcolor{red}{Clasicos clasicos no son}
+\subsection{(Classic) Resolution Methods} \label{sec:NHres}
+
+\textcolor{red}{Clasicos clasicos no son, puedo cambiar classic por non-heuristic?}
 
@@
 
-Where $C_i$ is the partition of dataset $X$ defined by solution $p_i$, and $(x_a, x_b)$ are instances in the cluster $c_j$ such that $a \neq b$ and the distance between each pair of instances in $c_i$ is included in the sum only once.
+Where $C_i$ is the partition of dataset $X$ defined by solution $p_i$, and $(x_a, x_b)$ are instances in the cluster $c_j$ such that $a \neq b$ and the Euclidean distance ($d_D^2(\cdot, \cdot)$) between each pair of instances in $c_i$ is included in the sum only once.
 

### 2019-02-22 Papers/SHADE/paperSHADE2.tex
Terminando PaperShade

--- 2019-02-14 Papers/SHADE/paperSHADE.tex
@@
 Puedo decir en alguna parte que se hace una iterpretación relajada de las restricciones.
-Tengo que decidir como llamar a la funcion de coste, "cost", "fitness" o "heuristic"}
+Tengo que decidir como llamar a la funcion de coste, "cost", "fitness" o "heuristic".
+Cambiar los = de las hipotesis por :.
+Cambiar constraints set por constraint set}
 
 \section{Introduction}
+
+%%CORREGIR ESTA SECCION
 
@@
 
-Clustering with constraints is a semi-supervised learning method.
+Constrained clustering is a semi-supervised learning method.
 Its goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraints set.
@@
 
-Some of the constrained clustering applications mentioned in \cite{davidson2007survey} include the application of this technique to data obtained from image, video, audio, text, etc.
-
-It should be noted that, as we will study in more depth in Section \ref{sec:ugly}, the incorporation of constraints to clustering modifies its complexity, making it harder.
+[Some of the constrained clustering applications mentioned in \cite{davidson2007survey} include: object recognition and tracking in video data, grouping of face images according to their orientation, detection of simultaneous appearance of genes and proteins in biological data, grouping of text documents according to their content, grouping of URLs obtained as a result of ambiguous searches according to their content, identification of speakers in audio data, and the widely known grouping of vehicles and lanes based on GPS data.]
+
+It should be noted that, as we will study in more depth in Section \ref{sec:feas}, the incorporation of constraints to clustering modifies its complexity, making it harder.
 The SHADE algorithm has proven to be one of the most successful methods to solve real-domain problems in an approximate way.
-The aim of this paper is to propose an adaptation of it to solve clustering with restrictions.
+The aim of this paper is to propose an adaptation of it to solve constrained clustering.
 
@@
 We will properly compare the results obtained by both methods.
-On the other hand we will check whether the population-based methods represent an improvement over some of the non ehuristic-based methods, presented in \cite{davidson2007survey} and briefly discussed in this paper in Section \ref{sec:NHres}.
-
-\textcolor{red}{Incluir un párrafo en el que se comete la organización del paper?}
+On the other hand we will check whether the population-based methods represent an improvement over some of the non heuristic-based methods, presented in \cite{davidson2007survey} and briefly discussed in this paper in Section \ref{sec:NHres}.
+
+Regarding the organization of this paper, Section \ref{sec:background} reviews the general knowledge concerning constrained clustering.
+In the Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
+On the other hand the SHADE algorithm will be presented in the section \ref{sec:SHADE}, as well as in it we will present its adaptation for constrained clustering and a new cost function.
+Sections \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis.
+Finally, in Section \ref{sec:conclusiones} we discuss conclusions and thoughts about future work.
 
@@
 
-\section{Brief Review of the BRKGA Algorithm}
+\section{Brief Review of the BRKGA Algorithm} \label{sec:brkga}
 
@@
 
-\section{Experimental Setup}
+\section{Experimental Setup} \label{sec:expSetup}
 
@@
 
+%% CORREGIR DESDE AQUI HASTA EL FINAL
+
 \subsection{Calibration}
@@
 
+In both cases---BRKGA+LS and SHADE+LS---the population size used will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+
+To compare with the state of the art methods mentioned in section \ref{sec:NHres} we will use the parameters recommended by the authors, setting the maximum limit of iterations as 300 in all cases.
+
 It should be noted that the study presented in this document aims to compare average performance, not peak performance.
@@
 
-In both cases---BRKGA+LS and SHADE+LS---the population size used will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
-
-\section{Experimental Results}
+\clearpage
+
+\section{Experimental Results} \label{sec:results}
 
@@
 To lessen the effect this may have on the results, we will apply each method 5 times to each dataset and set of constraints, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+
+\subsection{Comparing BRKGA + LS with SHADE + LS}
 
@@
 We can note that SHADE+LS represents a consistent improvement over BRKGA + LS in both Unsat and Time, while in ARI both methods seems similar.
-This is because, operating with a small set of restrictions, the penalty applied to the fitness function of both methods has little presence, and since they are population-based and heurist-guided methods, they have similar exploration and exploitation capabilities.
+This is because, operating with a small set of constraints, the penalty applied to the fitness function of both methods has little presence, and since they are population-based and heurist-guided methods, they have similar exploration and exploitation capabilities.
 
@@
 	\caption{Experimental results obtained with 10\% of labeled data}
-	\label{tab:results:10}
+	\label{tab:results10}
 \end{table}
 
-The Table \ref{tab:results:10} presents the results obtained for the constraints set based on 10\% labeled data.
+The Table \ref{tab:results10} presents the results obtained for the constraints set based on 10\% labeled data.
 On the results for Unsat and Time, we can draw similar conclusions to those obtained with Table \ref{tab:results5}, that is, the advantage remains for SHADE + LS, even with the general increase of these values derived from a larger set of constraints.
@@
 
-\section{Results Analysis}
+\subsection{Comparing SHADE + LS with the state-of-the-art}
+
+\clearpage
+
+\section{Results Analysis} \label{sec:analisis}
 
@@
 In order to identify cases where there are no significant differences we define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, so that $P(A = B) = P(\rho \in rope)$.
-Based on these results we calculate the weights of a Dirichlet distribution and sample it.
-The result of this sampling will be a set of triplets in the form of $[(\rho < r_{min}), (\rho \in rope), (\rho > r_{max}]$.
+Based on these results we calculate the weights of a Dirichlet distribution and sample it to get a set of triplets in the form of $[(\rho < r_{min}), (\rho \in rope), (\rho > r_{max})]$.
 
@@
 
-\clearpage
-
-\section{Conclusions and Future Work}
+\section{Conclusions} \label{sec:conclusiones}
+
+In this paper we proposed an adaptation of SHADE to solve the constrained clustering problem.
+We took concepts such as the random-key and the encoder/decoder from BRKGA to transform constrained clustering into a real-domain optimization problem.
+We also proposed a new cost function that has proven to be more adaptable and less restrictive.
+
+The SHADE algorithm is one of the most successful methods to approximately solve difficult problems, and constrained clustering does not seem to be an exception.
+We obtained results on 25 different data sets and 4 different constraints sets for each, applying all the resolution methods named in this paper.
+Thanks to statistical tests, like Wilcoxon's signed rank test, Bayesian signed rank test and Friedman's test, we were able to objectively prove that the SHADE + LS approach is significantly better than previous proposals.
+This last statement also applies to the comparison with some of the state-of-the-art methods.
 
@@
 \end{figure}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{c >{\centering\arraybackslash}c m{5cm} c}
+			\hline
+			Algorithm & Parameter & Meaning & Value \\
+			\hline
+			All & $MaxIter$ & Iterations limit for the algorithm & 300 \\
+			\multirow{2}{*}{COPKM}
+			& $Tolerance$ & Used to determine the lower threshold limit of centroid movement & $1 \times 10^{-4}$ \\
+			& $InitMethod$ & Method for centroid initialization & Random \\
+			\multirow{5}{*}{CEKM}
+			& $\alpha$ & exponente que controla la penalización
+			por asignar instancias a conjuntos con alta cardinalidad &
+			\hline \\
+			\hline
+
+		\end{tabular}}		
+		\caption{Parameters setup used for the state-of-the-art methods.}
+		\label{tab:paramsSOTA}
+	\end{table}

### 2019-03-09 Papers/SHADE/descartados.tex
Correcciones en memoria SHADE y nuevos resultados

%%%%%%%%%%%%%%%%% Test frecuentiastas %%%%%%%%%%%%%%%%%%%

\subsection{Applying the Wilcoxon's signed-rank test to our results}

We will apply the Wilcoxon's signed-rank test for each of the measures obtained, namely: ARI, Unsat and Time. We set as null hypothesis $H_0: $ the distribution of $A - B$ is symmetric about $\mu$. However, given that in the case of ARI it is a problem of maximisation, and in the case of Time and Unsat it is a problem of minimisation, we will consider two different alternative hypotheses. In the case of maximization we propose $H_{max}:$ the distribution of $A - B$ is shifted to the left with respect to $\mu$. On the other hand, in the case of minimization we propose $H_{min}:$ the distribution of $A - B$ is shifted to the right with respect to $\mu$. In this way we assume that SHADE+LS is better than BRKGA+LS and we try to prove it.

In this case we set $\mu = 0$, and work with a confidence level of 95\%, which means that the significance level is 0.05 ($\alpha = 0.05$). Table \ref{tab:freqTest} shows the statistical values obtained using Wilcoxon's signed-rank test for each comparison.

\begin{table}[!htp]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	\begin{tabular}{ccccc}
		\hline
		Measure & $R^{+}$ & $R^{-}$ & P-value & Alternative hypothesis  \\
		\hline
		ARI & 3088 & 1665 & 0.005254 & $H_{max}$ \\
		Time & 4008 & 1042 & 0.0000001723 & $H_{min}$ \\
		Unsat & 3082.5 & 745.5 & 0.0000003838 & $H_{min}$ \\
		\hline
	\end{tabular}
	\caption{Results obtained by the Wilcoxon's signed-rank test for the comparison SHADE vs BRKGA}
	\label{tab:freqTest}
\end{table}

For all three considered measures, $P-value < \alpha$, therefore in all cases we can reject $H_0$ and infer that there are significant differences between the two methods, giving the advantage to SHADE+LS with basis to the proposed alternative hypothesis. In conclusion, we can safely admit that [the] Wilcoxon's signed-rank test provides empirical evidence in favour of SHADE+LS.

### 2019-03-09 Papers/SHADE/paperSHADE.tex
Correcciones en memoria SHADE y nuevos resultados

--- 2019-02-12 Papers/SHADE/paperSHADE.tex
@@
 
-\author[mymainaddress]{Germ\'an Gonz\'alez Almagro\corref{mycorrespondingauthor}}
+\author[mymainaddress]{Germ\'an Gonz\'alez-Almagro\corref{mycorrespondingauthor}}
 \cortext[mycorrespondingauthor]{Corresponding author}
@@
 
+\author[mymainaddress]{Juli\'an Luengo}
+
+\author[mysecondaddress]{Jos\'e-Ram\'on Cano}
+
 \author[mymainaddress]{Salvador Garc\'ia}
 
-\author[mymainaddress]{Another Juan}
-
-\address[mymainaddress]{Department of Computer Science and Artificial Intelligence, University of Granada, 18071 Granada, Spain}
+\address[mymainaddress]{DaSCI Andalusian Institute of Data Science and Computational Intelligence, University of Granada, Spain}
+
+\address[mysecondaddress]{Dept. of Computer Science, EPS of Linares, University of Ja\'en, Campus Cient\'ifico Tecnol\'ogico de Linares, Cintur\'on Sur S/N, Linares 23700, Ja\'en, Spain}
 
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+This paper proposes an adaptation of the successful SHADE method to solve the constrained clustering problem.
+This problem consists of a modification of the traditional clustering to consider a new type of information, which is given in the form of must-link and cannot-link constraints.
+We will compare the results obtained on 25 datasets by this new SHADE adaptation with a previous heuristic-based approach---BRKGA---as well as with some of the state-of-the-art-methods, supporting our conclusions with the aid of Bayesian tests.
 \end{abstract}
@@
 \begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
+constrained clustering, must-link, cannot-link, genetic algorithm, differential evolution.
 \end{keyword}
@@
 
-\textcolor{red}{Revisar que todos los subindices en los que aparecan comas lleven corchetes.
-Revisar que las asignaciones en los algoritmo sean flechas y no =.
-Revisar que restrictions nunca aparece, siempre constraints.
-Revisar que es siempre instance-level y no instance level.
-Comprobar que es siempre dataset y no data set.
-Puedo decir en alguna parte que se hace una iterpretación relajada de las restricciones}
-
-\section{Introduction}
-
-\textcolor{red}{Citar a los de las hormigas. Poner un parrafo de aplicaciones del clustering con restricciones}
-
-\section{Background}
-
-%\subsection{Constrained Clustering}
-
-Clustering is the task of grouping instances of a dataset into subsets.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset.
-In this way, the clustering process obtains subsets of instances of the initial dataset that present high intra-similarity and low inter-similarity.
+\section{Introduction} \label{sec:Intro}
+
+One of the most widely known and studied data analysis problems is clustering.
+It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
+Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology and genetics among others \cite{ClusterAnalysis}.
+
+In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
+The main difference is that in hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{ClusterAnalysis}.
+In this paper we will focus on partitional clustering.
+\textcolor{red}{Poner ejemplos de cada uno (con referencias)?}
+
+We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them.
+A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features.
+More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,d]})$.
+A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$.
+As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$.
+The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$.
+An important detail is that each instance must belong to a single cluster.
+\cite{jain1999data}
+
+\textcolor{red}{Bucar una referencia fiable de ssl}
+
+Constrained clustering is a semi-supervised learning method.
+Its goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
+It has been succesfully applied in many knowledge fields.
+It has been used to guide the movement of walking robots, as well as in other advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}.
+In \cite{seret2014new} constrained clustering is presented as a useful tool in the context of applied marketing.
+Biology has also made use of constrained clustering, being able to detect simultaneous appearances of genes and proteins in biological data bases \cite{segal2003discovering}.
+In \cite{levy2008structural} a method of segmenting musical audio into separated sections using constrained clustering is described.
+Constrained clustering has also been useful to classify Internet traffic, which is considered to be one of the most fundamental functionalities in the network management field \cite{wang2014internet}.
+Electoral district design problems can also be approached using constrained clustering, as shown in \cite{brieden2017constrained}.
+Lastly, it is mandatory to name the widely known application to GPS data of constrained clustering, presented in \cite{wagstaff2001constrained}.
+
+Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level, instance-level and feature-level constrained clustering \cite{bradley2000constrained,davidson2007survey,schmidt2011clustering}.
+Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.
+
+In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints.
+Pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
+On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure.\cite{davidson2007survey} This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \textcolor{red}{X}.
+
+Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard and soft constraints.
+Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new, wagstaff2001constrained, law2004clustering}.
+For the purposes of this paper, we will employ the latter.
+
+Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem.
+Therefore, the incorporation of constraints may modify the complexity of the clustering problem, depending on the type of constraints used.
+As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete.
+\cite{davidson2005clustering}
+
+The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it.
+As mentioned earlier, it is a difficult problem to solve, so nature-inspired techniques are presented as a promising option to find quality approximate solutions.
+Nature is the best example of adaptive problem solving since it can apply an optimal strategy suited for each natural phenomenon \cite{fausto2019ants}.
+Nature-inspired algorithms are designed to emulate natural optimization phenomena, such as evolution, collective behavior of animals, physics laws or even human being-related processes.
+There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the Biased Random-key Genetic Algorithm (BRKGA), which can be found in \cite{de2017comparison}.
+Swarm-based methods have also been applied to constrained clustering, such as the one presented in \cite{xu2013improving}.
+
+Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}; in this paper we propose a new DE variant to find quality solutions for the constrained clustering problem (see Section \ref{sec:SHADE}).
+We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant.
+We will properly compare the results obtained by our new method with previous nature-inspired algorithms, as well as with the constrained clustering state-of-the-art.
+
+\textcolor{red}{Cambiar el parrafo de abajo en funcion de como quede la organizacion de secciones final}
+
+Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
+In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
+Afterward, a variant of DE algorithm will be presented in Section \ref{sec:SHADE}, followed by its adaptation for constrained clustering and a new fitness function.
+Sections \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis.
+Finally, in Section \ref{sec:conclusiones} we discuss conclusions.
+
+\section{Background} \label{sec:background}
+
+In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering \ref{sec:BackSOTA}.
+We will also present the basis of the DE optimization method (Section \ref{sec:BackDE}).
+
+\subsection{Constrained Clustering} \label{sec:BackCC}
 
 In most clustering applications it is common to have some kind of information about the dataset to be analyzed.
-In constrained clustering this information is given in the form of pairs of instances.
+In pairwise instance-level constrained clustering this information is given in the form of pairs of instances.
 A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster.
-**Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms**.
-[It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms]
-
-Before formalizing the definition of constraints we must give a notation for the dataset.
-We will note our dataset as $X$, which is a set of $N$ instances with $D$ features.
-We well refer to the instances of $X$ as $x$.
-Now, given the definition of dataset, we can distinguish between two types of constraints:
+It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms.
+
+Given the notation for the clustering problem introduced in Section \ref{sec:Intro}, we can now formalize the two type of constraints mentioned:
 
@@
 
-	\item Must-link constraints $c_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
-
-	\item Cannot-link constraints $c_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.
+	\item Must-link constraints $C_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
+
+	\item Cannot-link constraints $C_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.
 
@@
 
-The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraints set.
-As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $N = |X| = \sum_{i = 1}^{K} |c_i|$.
-
-Knowing how a constraint is defined, Must-link constraints---from now on ML---are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric.
-This way, given constraints $c_=(x_a,x_b)$ and $c_=(x_b,x_c)$ then $c_=(x_a,x_c)$ is verified.
-In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $c_=(x_a,x_b)$, then $c_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.
-
-On the other hand, Cannot-link constraints---from now on CL---do not constitute an equivalence relation.
-However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then it is also true that $c_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
-
-\subsection{The Feasibility Problem}
-
-Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the problem's complexity.
-The feasibility problem for non-hierarchical constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.
+The goal of constrained clustering is to find a partition (or clustering) of $k$ clusters $C = \{c_1, \cdots, c_k\}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set.
+As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $n = |X| = \sum_{i = 1}^{k} |c_i|$.
+
+Knowing how a constraint is defined, ML constraints are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric.
+This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified.
+In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.
+
+On the other hand, CL constraints do not constitute an equivalence relation.
+However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $C_{\neq}(x_a,x_b)$, then it is also true that $C_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
+
+\textcolor{red}{Opciones para cambiar el nombre de los conjuntos de restricciones: $CS_=$ y $CS_{\neq}$, con $CS$ para referirse a las restricciones en general.
+O simplemente $ML$ y $CL$, con $CS$ para referirse a las restricciones en general}
+
+\subsection{The Feasibility Problem} \label{sec:BackFeas}
+
+Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem.
+The feasibility problem for non-hierarchical constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
 
@@
 
-	\textbf{Feasibility Problem}: given a dataset $X$, a set of constraints $R$, and the bounds on the number of clusters $K_l \leq k \leq K_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $R$ are satisfied?
+	\textbf{Feasibility Problem}: given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
 \cite{davidson2007survey}\cite{davidson2005clustering}
@@
 
-In \cite{davidson2005clustering} it is proven that, when $K_l = 1$ and $K_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem (it is also proven that it is not harder, so both have the same complexity).
-Table \ref{tab:feasibility} shows the feasibility problem's complexity for different types of constraints.
+In \cite{davidson2005clustering} it is proven that, when $k_l = 1$ and $k_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem (it is also proven that it is not harder, so both have the same complexity).
+Table \ref{tab:feasibility} shows the complexity of the feasibility for different types of constraints.
 
@@
 
-\textcolor{red}{El parrafo de abajo esta bien?}
-
 Intractable problems are hard to solve with deterministic and exact methods.
@@
 
-\subsection{Short Introduction to Differential Evolution}
-
-\textcolor{red}{Eliminar toda la notacion matematica de esta seccion?}
-
-To introduce Differential Evolution---from now on DE---we take as a starting point the review of the state-of-the-art carried out by \cite{das2011differential}.
-A highly detailed analysis of DE and its variants can also be found in \cite{das2011differential}.
-
-The first article written on DE emerged in 1995 as a technical report by \cite{noman2008differential}.
-From that point on, DE's reputation was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
-In particular, DE excelled in real-coded optimization problems.
-
-DE arose as a variant of population-based evolutionary algorithms, its goal being to perform an intelligent search in the problem's solution space in the most optimized and effective way possible.
-
-Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{i,1},\cdots,p_{i,D}\}$ that serve as parameters for the function to be optimized.
-These individuals are considered solutions to the problem.
-
-To guide the optimization process, we use an objective function.
-DE's task is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
-
-\begin{figure}[!h]
-	\centering
-	\includegraphics[scale=0.3]{Figures/DEop.png}
-	\caption{DE's optimization process \cite{das2011differential}.}\label{img:DE}
-\end{figure}
-
-DE's optimization process works through a simple cycle of stages, presented in Figure \ref{img:DE}, that we can summarize in as follows:
-
-\begin{itemize}
-
-	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of $N$ $D$-dimensional arrays of parameters.
-These vectors constitute candidate solutions for the problem to solve.
-
-	\item \textbf{Difference-vector-based mutation}: DE introduces diversity into the population via mutation operator.
-This operator is applied to each individual of the population, known in this context as the parent vector.
-It combines the parent vector with two randomly selected vectors from the population by applying an arithmetic operator to them.
-The result of this process is a mutant vector.
-
-	\item \textbf{Crossover/Recombination}: The mutant vector exchanges some components with its associated parent to form a trial vector, which will be the candidate to replace the parent vector.
-Several strategies exist to generate this trial vector, which we will not delve into; however, they can be found in \cite{das2011differential}.
-
-	\item \textbf{Selection}: The newly generated trial vector is compared with its associated parent using the objective function.
-If the trial vector is better than the parent, then it will replace the parent in the next generation.
-It is worth noting that, following this strategy, the population never degenerates but only gets better or remains the same.
-
-\end{itemize}
-
-Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
-We will focus on the SHADE variant, which adaptively optimizes some of DE's parameters (see Section \ref{sec:SHADE}).
-
-\subsection{(Classic) Resolution Methods}
-
-\textcolor{red}{Clasicos clasicos no son}
+\subsection{State-of-the-art Methods} \label{sec:BackSOTA}
 
 Constrained clustering has many applications and has been widely studied in the literature.
-The first adaptation of a classic clustering method for constrained clustering was proposed by \cite{wagstaff2001constrained}.
+The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}.
 It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
-This method was named COP-kmeans.
-
-The authors of \cite{antoine2012cecm} propose an adaptation of the evidential C-means algorithm, within the fuzzy clustering framework, for constrained clustering.
-The particularity of this resolution method is that the membership of instances to a cluster is defined by a probabilistic belief function.
-This method redefines constraints from the point of view of belief functions and includes them in the cost function.
-
-A modification of CVQE (Constrained Vector Quantization Error), LCVQE, is proposed in \cite{pelleg2007k}.
-The adaptation of CVQE for constrained clustering was proven to produce high quality results, at the cost of a very high computational complexity.
-LCVQE (Linear CVQE) introduces a modification of CVQE's objective function to make it more intuitive and less computationally complex.
-The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraints sets.
-
-TVClust (Two Views Clustering) and RDP-means (Relation Dirichlet Process - Means) were proposed by \cite{khashabi2015clustering}.
+This method was named COP-kmeans, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
+
+In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), an \textcolor{red}{adaptation} of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods, for constrained clustering.
+The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
+This method redefines constraints from the point of view of belief functions and includes them in the fitness function.
+
+A modification of the Constrained Vector Quantization Error algorithm (CVQE \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}.
+The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity.
+Linear CVQE (LCVQE) introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex.
+The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.
+
+Two Views Clustering (TVClust) and Relation Dirichlet Process - Means (RDPM) were proposed in \cite{khashabi2015clustering}.
 TVClust is able to incorporate the constraints into the clustering problem by making a relaxed interpretation of them.
@@
 
-\textcolor{red}{Decir que voy a comparar con ellos? Incluir mas métodos ``clasicos''?}
-
-\section{Brief Review of the BRKGA Algorithm}
-
-The biased random-key genetic algorithm---from now on BRKGA---was first proposed by \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.
+\subsection{Short Introduction to Differential Evolution} \label{sec:BackDE}
+
+Evolution-based methods comprise a group of algorithms developed on the basis of the natural laws of evolution.
+In this type of techniques a population of individuals---representing potential solutions to the problem---is usually employed, which compete and combine through a certain number of generations so that the population evolves in such a way that only the best individuals, and thus the best solutions, remain at the end.
+This type of techniques involves the development of a series of operators that allow us to simulate the evolution process, such as crossover, mutation and selection operators.
+\cite{fausto2019ants}
+
+DE, which emerged in 1995 as a technical report presented in \cite{storn1997differential}, is a good example of evolution-based methods.
+From that point on, the reputation of DE was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
+In particular, DE excelled in real-coded optimization problems.
+A highly detailed analysis of DE and its variants can be found in \cite{das2011differential}.
+\textcolor{red}{No he podido encontrar las referencias a los resultados de las compteciones en las que ganaba DE (ni SHADE)}
+
+DE arose as a variant of population-based evolutionary algorithms, its goal being to perform an intelligent search in the solution space of the problem in the most optimized and effective way possible.
+
+Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,D]}\}$ that serve as parameters for the function to be optimized.
+These individuals are considered solutions to the problem.
+
+To guide the optimization process, we use a fitness function.
+The task of DE is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.3]{Figures/DEop.png}
+	\caption{The optimization process of DE \cite{das2011differential}.}\label{img:DE}
+\end{figure}
+
+The optimization process of DE the works through a simple cycle of stages, presented in Figure \ref{img:DE}, that we can summarize as follows:
+
+\begin{itemize}
+
+	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of a given number of $D$-dimensional arrays of parameters.
+These vectors constitute candidate solutions for the problem to be solved.
+
+	\item \textbf{Difference-vector-based mutation}: DE introduces diversity into the population via the mutation operator.
+This operator is applied to each individual of the population, known in this context as the parent vector.
+It combines the parent vector with two randomly selected vectors from the population by applying an arithmetic operator to them.
+The result of this process is a mutant vector.
+
+	\item \textbf{Crossover/Recombination}: The mutant vector exchanges some components with its associated parent to form a trial vector, which will be the candidate to replace the parent vector.
+Several strategies exist to generate this trial vector, which we will not delve into; however, they can be found in \cite{das2011differential}.
+
+	\item \textbf{Selection}: The newly generated trial vector is compared with its associated parent using the fitness function.
+If the trial vector is better than the parent, then it will replace the parent in the next generation.
+It is worth noting that, following this strategy, the population never degenerates but only gets better or remains the same.
+
+\end{itemize}
+
+Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
+We will focus on the SHADE variant, which adaptively optimizes some of the parameters of DE (see Section \ref{sec:SHADE}).
+
+\section{Brief Review of the BRKGA Algorithm} \label{sec:brkga}
+
+The biased random-key genetic algorithm---from now on BRKGA---was first proposed in \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.
 
@@
 
-Their method randomly initializes a population of random-key vectors, each one with $N$ random-keys.
-The decoder divides the interval $[0,1]$ in $K$ intervals, so there exists a correspondence between each random-key and the integer corresponding to the interval which it lies in.
-Table \ref{tab:decodingrk} shows an example of random-key decoding for $K = 3$.
+This method randomly initializes a population of random-key vectors, each one with $n$ random-keys.
+The decoder divides the interval $[0,1]$ in $k$ intervals, so there exists a correspondence between each random-key (instance $x_i$) and the integer (label $l_i$) corresponding to the interval which it lies in.
+Table \ref{tab:decodingrk} shows an example of random-key decoding for a dataset with 10 instances ($n$ = 10) and $k = 3$.
 Note that extreme values 0 and 1 can also appear in a random-key vector.
@@
 		\hline
-		Index & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
+		Index (instance) & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
 		\hline
@@
 		\hline
-		Clusters & 1 & 2 & 2 & 2 & 1 & 3 & 1 & 3 & 1 & 3 \\
+		Cluster (label) & 1 & 2 & 2 & 2 & 1 & 3 & 1 & 3 & 1 & 3 \\
 		\hline
@@
 \begin{equation}
-f_i = z_i + \overbrace{(\mu * N * {infeasibility}_i)}^{penalty}
+f_i = z_i + \overbrace{(\mu * n * {infeasibility}_i)}^{penalty},
 \label{eq1}
@@
 
-Where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.
-
-\begin{equation}
-z_i = \sum_{c_j \in C_i} \left[ \frac{\sum_{x_a, x_b \in c_j} d_D^2(x_a,x_b)}{|c_j|}\right]
+\noindent where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.
+
+\begin{equation}
+z_i = \sum_{c_j \in C_i} \left[ \frac{\sum_{x_a, x_b \in c_j} d^2(x_a,x_b)}{|c_j|}\right],
 \label{eq2}
@@
 
-Where $C_i$ is the partition of dataset $X$ defined by solution $p_i$, and $(x_a, x_b)$ are instances in the cluster $c_j$ such that $a \neq b$ and the distance between each pair of instances in $c_i$ is included in the sum only once.
+\noindent where $C_i$ is the partition of dataset $X$ defined by solution (individual) $p_i$, and $(x_a, x_b)$ are instances in the cluster $c_j$ such that $a \neq b$ and the Euclidean distance ($d^2(\cdot, \cdot)$) between each pair of instances in $C_i$ is included in the sum only once.
 
@@
 
-Success history based adaptive differential evolution---from now on SHADE---is an optimization method based on differential evolution.
-Proposed by \cite{tanabe2013success}, it was an improvement for the JADE model, created by \cite{zhang2009jade}.
+Success History Based Adaptive Differential Evolution (SHADE) is an optimization method based on differential evolution.
+Proposed in \cite{tanabe2013success}, it was an improvement for the JADE model, presented in \cite{zhang2009jade}.
 
@@
 \begin{equation}
-m_{[i,G]} = p_{[i,G]} + F_i * (p_{[pbest, G]} - p_{[i,G]}) + F_i * (p_{[r1, G]} - p_{[r2,G]})
+m_{[i,G]} = p_{[i,G]} + F_i * (p_{[pbest, G]} - p_{[i,G]}) + F_i * (p_{[r1, G]} - p_{[r2,G]}),
 \label{eq3}
@@
 
-Where $m_{[i,G]}$ is the mutant vector, which is generated with an individual $p_{[i,G]}$ from the population [serving as a starting point].
-Indices $r1$ and $r2$ are random values in the range $[0,N]$, different from $i$ and from each other.
+\noindent where $m_{[i,G]}$ is the mutant vector, which is generated with an individual $p_{[i,G]}$ from the population serving as a starting point.
+Indices $r1$ and $r2$ are random values in the range $[0,p_{size}]$, different from $i$ and from each other.
 The individual $p_{[pbest, G]}$ is randomly selected from among the best $N \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
-This way, the parameter $p$ controls the greediness of the mutation strategy.
+This way, the parameter $pbest$ controls the greediness of the mutation strategy.
 The parameter $F_i$ defines the magnitude of the operator.
@@
 \end{array}
-\right.
+\right.,
 \label{eq4}
@@
 
-Where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.
+\noindent where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.
 
@@
 
-The parameters $p$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process.
+The parameters $pbest$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process.
 These are the parameters that SHADE adaptively optimizes using the mentioned history record.
@@
 \begin{equation}
-F_i = \text{randc}_i(M_{[F,r_i]}, 0.1)
+F_i = \text{randc}_i(M_{[F,r_i]}, 0.1),
 \label{eq7}
@@
 
-where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$.
+\noindent where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$.
 When a value for $CR_i$ outside of the range $[0,1]$ is generated, it is replaced by the corresponding limit value.
@@
 
-In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the weighted mean, which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed by \cite{peng2009multi} to prevent $CR$ from converging to small values.
+In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the weighted mean, which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed in \cite{peng2009multi} to prevent $CR$ from converging to small values.
 
@@
 \begin{equation}
-\omega_i = \frac{\Delta f_i}{\sum_{i = 1}^{|S_{RC}|} \Delta f_i}
+\omega_i = \frac{\Delta f_i}{\sum_{i = 1}^{|S_{RC}|} \Delta f_i},
 \label{eq11}
@@
 
-where $\Delta f_i = |f(t_{[i,G]}) - f(p_{[i, G]})|$, which is the amount of improvement obtained from the trial vector with respect to the parent.
-
-In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12} (proposed by \cite{tanabe2013success}).
+\noindent where $\Delta f_i = |f(t_{[i,G]}) - f(p_{[i, G]})|$, which is the amount of improvement obtained from the trial vector with respect to the parent.
+
+In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12} (proposed in \cite{tanabe2013success}).
 
@@
 \begin{equation}
-pbest_i = \text{randn}[2/N, 0.2]
+pbest_i = \text{randn}[2/N, 0.2],
 \label{eq13}
@@
 
-such that there is always at least two individuals to choose between.
-
-\subsection{Adaptation of SHADE for constrained clustering}
+\noindent such that there is always at least two individuals to choose between.
+
+\subsection{Adaptation of SHADE for constrained clustering} \label{sec:SHADEadapt}
 
 For the adaptation of SHADE to the problem of constrained clustering, we will take some of the elements that allowed us to adapt it to BRKGA.
-To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of random-keys.
+To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of $n$ random-keys.
 Additionally, we will also reuse the same initialization method and the same decoder used in BRKGA (see Section \ref{sec:AdaptationofBRKGA}).
@@
 
-Algorithm \ref{alg:SHADE} describes the overall SHADE optimization process and is an adaptation of the one that can be found on \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
+Algorithm \ref{alg:SHADE} describes the overall SHADE optimization process and it is an adaptation of the one that can be found in \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
 
@@
 	\SetKwFunction{LocalSearch}{LocalSearch}
-	\KwIn{Dataset $X$, constraints sets $c_=$ and $c_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $K$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $K$.}
 	\tcp{Initialization phase}
@@
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Dataset $X$, constraints sets $c_=$ and $c_{\neq}$, decoded random-key vector (solution) $S$, number of clusters $K$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, decoded random-key vector (solution) $S$, number of clusters $K$.}
 	\BlankLine
@@
 
-\subsection{Constraints Generation}
-
-\textcolor{red}{En esta seccion podria simplemente decir el numero de restricciones que genero como un numero magico sin meterme en nada de grafos, pero igual es un poco burro.}
-
-Since we have the true labels associated with each dataset, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraints sets.
+\subsection{Constraint Generation}
+
+Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets.
 This method consists in randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
 
-We will generate four different sets of constraints for each dataset.
-The number of constraints is given by the number of edges of the complete constraints graph that we can build with a percentage of labeled data.
-Each constraints set is associated with a different percentage of labeled data, namely: 5\%, 10\%, 15\% and 20\%.
-It is worth noting that constraints are not the complete constraint graph made up of the labeled instances, this graph only sets the number of constraints to be created, this way there is a lower probability of biasing the set of constraints so that there are classes with poor representation in it.
+We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with four small percentages of the size of the dataset: 10\%, 15\% and 20\%.
+With $n$ being the fraction of the size of the dataset associated to each of these percentages, the formula $\frac{n(n-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n$ vertices would have.
+
+The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
 Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.
@@
 	\resizebox{\textwidth}{!}{
-	\begin{tabular}{lcc c cc c cc c cc}
+	\begin{tabular}{lcc c cc c cc}
 		\hline
 		\multirow{2}{*}{Dataset} &
-		\multicolumn{2}{c}{5\%} &&  \multicolumn{2}{c}{10\%} && \multicolumn{2}{c}{15\%} && \multicolumn{2}{c}{20\%} \\
-		\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
-		& ML & CL && ML & CL && ML & CL && ML & CL \\
-		\hline
-		Appendicitis & 13 & 2 && 39 & 16 && 71 & 49 && 164 & 67 \\
-		Boston & 0 & 325 && 3 & 1272 && 13 & 2837 && 26 & 5125 \\
-		Breast Cancer & 216 & 190 && 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
-		Circles & 50 & 55 && 208 & 227 && 502 & 488 && 853 & 917 \\
-		Diabetes & 2 & 251 && 5 & 985 && 7 & 2204 && 23 & 3893 \\
-		Ecoli & 30 & 106 && 163 & 398 && 357 & 918 && 609 & 1669 \\
-		Glass & 11 & 44 && 52 & 179 && 139 & 389 && 259 & 644 \\
-		Haberman & 76 & 44 && 304 & 161 && 634 & 401 && 1135 & 756 \\
-		Hayesroth & 12 & 16 && 39 & 81 && 102 & 174 && 177 & 319 \\
-		Ionosphere & 92 & 61 && 330 & 300 && 732 & 646 && 1299 & 1186 \\
-		Iris & 9 & 19 && 26 & 79 && 82 & 171 && 136 & 299 \\
-		Led7Digit & 25 & 275 && 126 & 1099 && 267 & 2508 && 460 & 4490 \\
-		Monk2 & 101 & 130 && 473 & 473 && 979 & 1101 && 1917 & 1824 \\
-		Moons & 55 & 50 && 200 & 235 && 494 & 496 && 900 & 870 \\
-		Pima & 412 & 329 && 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
-		Saheart & 152 & 124 && 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
-		Sonar & 29 & 26 && 100 & 110 && 245 & 251 && 436 & 425 \\
-		Soybean & 0 & 3 && 4 & 6 && 6 & 22 && 12 & 33 \\
-		Spectfheart & 56 & 35 && 233 & 118 && 543 & 277 && 965 & 466 \\
-		Spiral & 52 & 53 && 224 & 211 && 487 & 503 && 918 & 852 \\
-		Tae & 8 & 20 && 40 & 80 && 82 & 171 && 151 & 314 \\
-		Vehicle & 221 & 682 && 874 & 2696 && 1955 & 6046 && 3589 & 10776 \\
-		Vowel & 107 & 1118 && 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
-		Wdbc & 209 & 197 && 840 & 756 && 1925 & 1730 && 3472 & 2969 \\
-		Zoo & 7 & 8 && 21 & 34 && 29 & 91 && 41 & 169 \\
+		\multicolumn{2}{c}{$CS_{10}$} && \multicolumn{2}{c}{$CS_{15}$} && \multicolumn{2}{c}{$CS_{20}$} \\
+		\cline{2-3} \cline{5-6} \cline{8-9}
+		& ML & CL && ML & CL && ML & CL \\
+		\hline
+		Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
+		Boston & 3 & 1272 && 13 & 2837 && 26 & 5125 \\
+		Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
+		Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
+		Diabetes & 5 & 985 && 7 & 2204 && 23 & 3893 \\
+		Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
+		Glass & 52 & 179 && 139 & 389 && 259 & 644 \\
+		Haberman & 304 & 161 && 634 & 401 && 1135 & 756 \\
+		Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
+		Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
+		Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
+		Led7Digit & 126 & 1099 && 267 & 2508 && 460 & 4490 \\
+		Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
+		Moons & 200 & 235 && 494 & 496 && 900 & 870 \\
+		Pima & 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
+		Saheart & 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
+		Sonar & 100 & 110 && 245 & 251 && 436 & 425 \\
+		Soybean & 4 & 6 && 6 & 22 && 12 & 33 \\
+		Spectfheart & 233 & 118 && 543 & 277 && 965 & 466 \\
+		Spiral & 224 & 211 && 487 & 503 && 918 & 852 \\
+		Tae & 40 & 80 && 82 & 171 && 151 & 314 \\
+		Vehicle & 874 & 2696 && 1955 & 6046 && 3589 & 10776 \\
+		Vowel & 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
+		Wdbc & 840 & 756 && 1925 & 1730 && 3472 & 2969 \\
+		Zoo & 21 & 34 && 29 & 91 && 41 & 169 \\
 		\hline
@@
 
-Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed by \cite{wagstaff2001constrained}.
+Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
 This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.
@@
 \begin{equation}
-ARI(C_1, C_2) = \frac{Rand(C_1, C_2) - ExpectedIndex}{MaximumIndex - ExpectedIndex}
+ARI(C_1, C_2) = \frac{Rand(C_1, C_2) - ExpectedIndex}{MaximumIndex - ExpectedIndex},
 \label{eq16}
@@
 
-where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
+\noindent where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
 It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that an $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing with random partitions.
@@
 
+%% CORREGIR DESDE AQUI HASTA EL FINAL
+
 \subsection{Calibration}
 
-Table \ref{tab:params} shows a summary of the parameters setup used for both BRKGA+LS and SHADE+LS algorithms.
+Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA+LS and SHADE+LS algorithms.
 
@@
 			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_{size}$ & - \\
-			$p_{inherit}$ & Probability that a feature is inherit from an elite parent & $50\%$ & - \\
-			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Known for each dataset (Table \ref{tab:datasets}. It has the same value for both methods} \\
+			$p_{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
+			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Known for each dataset (Table \ref{tab:datasets}). It has the same value for both methods} \\
 			\hline
@@
 
+In both cases---BRKGA+LS and SHADE+LS---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+
+To compare with the state-of-the-art methods mentioned in section \ref{sec:BackSOTA} we will use the parameters recommended by the authors, setting the maximum limit of iterations as 300 in all cases.
+
 It should be noted that the study presented in this document aims to compare average performance, not peak performance.
-The parameter optimization process required for each algorithm and data set falls outside of this study.
-
-In both cases---BRKGA+LS and SHADE+LS---the population size used will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
-
-\section{Experimental Results}
-
-In this section we present the Tables \ref{tab:results5} to \ref{tab:results20}, which record the results obtained by the methods to be compared for each dataset and constraints set.
+The parameter optimization process required for each algorithm and dataset falls outside the scope of this study.
+
+\section{Experimental Results} \label{sec:results}
+
+In this section we present Tables \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.
 
 Since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
-To lessen the effect this may have on the results, we will apply each method 5 times to each dataset and set of constraints, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
-
-The ARI column shows the Adjusted Rand Index for each case, the Unsat column displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
+To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+
+\subsection{Comparing BRKGA+LS with SHADE+LS}
+
+In Tables \ref{tab:results10} to \ref{tab:results20} the ARI columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
 
@@
 			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
-			\cline{2-4} \cline{6-8}
-			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
-			\hline
-			Appendicitis & 0.001 & 0.000 & 374.388 && -0.047 & 0.000 & 360.150 \\
-			Boston & -0.002 & 0.000 & 1405.381 && 0.001 & 0.000 & 1425.668 \\
-			Breast Cancer & 0.052 & 9.704 & 3427.396 && 0.099 & 7.685 & 3453.035 \\
-			Circles & 0.020 & 1.333 & 1042.597 && 0.027 & 1.143 & 1004.287 \\
-			Diabetes & 0.001 & 0.000 & 1173.404 && 0.000 & 0.633 & 1374.711 \\
-			Ecoli & 0.005 & 6.912 & 1098.664 && 0.004 & 1.618 & 1045.699 \\
-			Glass & -0.009 & 0.000 & 703.172 && -0.007 & 0.000 & 686.843 \\
-			Haberman & 0.010 & 2.000 & 1098.043 && -0.003 & 2.000 & 1065.608 \\
-			Hayesroth & -0.009 & 0.000 & 534.294 && 0.002 & 0.000 & 507.300 \\
-			Ionosphere & 0.025 & 3.399 & 1821.049 && 0.044 & 3.530 & 1780.769 \\
-			Iris & 0.433 & 0.000 & 494.135 && 0.670 & 0.000 & 451.290 \\
-			Led7Digit & 0.006 & 4.867 & 1668.206 && 0.002 & 1.933 & 1648.570 \\
-			Monk2 & 0.043 & 3.810 & 1743.874 && 0.035 & 3.637 & 1702.858 \\
-			Moons & 0.081 & 1.905 & 1040.926 && 0.157 & 0.571 & 1021.028 \\
-			Pima & 0.011 & 15.169 & 3803.816 && 0.014 & 12.605 & 3723.091 \\
-			Saheart & 0.029 & 6.160 & 2013.580 && 0.006 & 6.739 & 1977.839 \\
-			Sonar & 0.010 & 0.364 & 1076.248 && 0.001 & 0.727 & 1061.966 \\
-			Soybean & 0.371 & 0.000 & 175.224 && 0.483 & 0.000 & 159.017 \\
-			Spectfheart & 0.104 & 1.099 & 1395.389 && 0.084 & 0.659 & 1365.181 \\
-			Spiral & 0.016 & 0.952 & 1044.697 && 0.004 & 0.953 & 1021.158 \\
-			Tae & 0.004 & 0.000 & 506.495 && 0.010 & 0.000 & 476.786 \\
-			Vehicle & 0.002 & 19.557 & 4008.912 && 0.003 & 10.255 & 4024.292 \\
-			Vowel & 0.001 & 10.465 & 3796.154 && 0.001 & 8.000 & 3763.485 \\
-			Wdbc & 0.038 & 10.246 & 3424.397 && 0.025 & 9.212 & 3400.415 \\
-			Zoo & 0.016 & 0.000 & 332.475 && 0.129 & 0.000 & 318.684 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained with 5\% of labeled data}
-	\label{tab:results5}
-\end{table}
-
-Table \ref{tab:results5} shows the results for the constraints set obtained with 5\% of the labeled data.
-We can note that SHADE+LS represents a consistent improvement over BRKGA + LS in both Unsat and Time, while in ARI both methods seems similar.
-This is because, operating with a small set of restrictions, the penalty applied to the fitness function of both methods has little presence, and since they are population-based and heurist-guided methods, they have similar exploration and exploitation capabilities.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{l ccc c ccc}
-			\hline
-			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
+			\multicolumn{3}{c}{BRKGA+LS} &&  \multicolumn{3}{c}{SHADE+LS} \\
 			\cline{2-4} \cline{6-8}
@@
 
-	\caption{Experimental results obtained with 10\% of labeled data}
-	\label{tab:results:10}
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE+LS and BRKGA+LS}
+	\label{tab:results10}
 \end{table}
 
-The Table \ref{tab:results:10} presents the results obtained for the constraints set based on 10\% labeled data.
-On the results for Unsat and Time, we can draw similar conclusions to those obtained with Table \ref{tab:results5}, that is, the advantage remains for SHADE + LS, even with the general increase of these values derived from a larger set of constraints.
-It is at ARI where we begin to see the difference between the two methods.
+Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
+We can note that SHADE+LS represents a consistent improvement over BRKGA+LS in both Unsat and Time, while in terms of ARI both methods seem similar.
+This is because the penalty applied to the fitness function of both methods has little presence when operating with a small constraint set, and since they are population-based and heuristic-guided methods, they have similar exploration and exploitation capabilities.
+
+ it is at ARI where we begin to see a meaningful difference between the two methods.
 
@@
 			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA + LS} &&  \multicolumn{3}{c}{SHADE + LS} \\
+			\multicolumn{3}{c}{BRKGA+LS} &&  \multicolumn{3}{c}{SHADE+LS} \\
 			\cline{2-4} \cline{6-8}
@@
 
-	\caption{Experimental results obtained with 15\% of labeled data}
+	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE+LS and BRKGA+LS}
 	\label{tab:results15}
@@
 
-Table \ref{tab:results15} presents the results obtained with 15\% of labeled data.
+Table \ref{tab:results15} presents the results obtained with the $CS_{15}$ constraint set.
 While conclusions for Unsat and Time remain unchanged, we see a clear difference as far as ARI is concerned.
-We observe that, for constraints sets obtained with 15\% labeled data, both methods already report ARI results above 0.7, which in an ARI ranking is considered to be a good result.
-We can even observe that the SHADE + LS method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA + LS that only obtains similar results in 2 datasets.
+We observe that both methods already report ARI results above 0.7, which in an ARI ranking is considered to be an acceptable result.
+We can even observe that the SHADE+LS method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA+LS that only obtains similar results in 2 datasets.
 
@@
 
-	\caption{Experimental results obtained with 20\% of labeled data}
+	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE+LS and BRKGA+LS}
 	\label{tab:results20}
@@
 
-It is in Table \ref{tab:results20}, which gathers results obtained with constraints sets generated from 20\% of labeled data for each dataset, that we observe the major differences between the two methods.
-We found that SHADE + LS is capable of producing clusters identical to the original ones for 4 datasets, against the single dataset for which BRKGA + LS obtains a similar result.
-Let's recall that an ARI of 1 represents a total matching between the two partitions given as arguments for ARI's calculation.
-This improvement in ARI is clearly reflected in the percentage of violated constraints--Unsat column---, value for which, once again, SHADE + LS gets the advantage.
+It is in Table \ref{tab:results20}, which gathers results obtained for the constraint set $CS_{20}$, that we observe major differences between the two methods.
+We find that SHADE+LS is capable of producing clusters identical to the original ones for 4 datasets, against the single dataset for which BRKGA+LS achieves it.
+Let us remember that an ARI of 1 represents a perfect match between the two partitions given as arguments.
+This improvement in ARI is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, SHADE+LS gets the advantage.
 Conclusions on execution time remain the same.
 
-\section{Results Analysis}
-
-With the results obtained by both methods for a total of 100 different datasets---the 25 datasets in combination with the 4 constraints sets for each one of them--we can perform an empirical analysis of them.
-This way we can statistically determine which of the two methods offers the best results.
-
-In order to carry out this task we will make use the widely known frequentist non-parametric Wilcoxon's signed rank test, as well as the lesser-known Bayesian sign test.
-
-From now on, and as far as this section is concerned, we will refer to the results obtained by BRKGA - LS as sample $A$, and to the results obtained with SHADE + LS as sample $B$.
-
-\subsection{Applying Wilcoxon's signed rank test to our results}
-
-We will apply the Wilcoxon's signed rank test for each of the measures obtained, namely: ARI, Unsat and Time.
-We set as null hypothesis $H_0 = $ the distribution of $A - B$ is symmetric about $\mu$.
-However, given that in the case of ARI it is a problem of maximisation, and in the case of Time and Unsat it is a problem of minimisation, we will consider a different alternative hypothesis for each case.
-In the case of maximization we propose $H_{max} =$ the distribution of $A - B$ is shifted to the left with respect to $\mu$.
-On the other hand, in the case of minimization we propose $H_{min} =$ the distribution of $A - B$ is shifted to the right with respect to $\mu$.
-In this way we assume that SHADE + LS is better than BRKGA + LS and we try to prove it.
-
-In this case we set $\mu = 0$, and work with a confidence level of 95\%, which means that the significance level is 0.05 ($\alpha = 0.05$).
-Table \ref{tab:freqTest} shows the statistical values obtained using Wilcoxon's test for each comparison.
-
-\begin{table}[!htp]
+\clearpage
+
+\subsection{Comparing SHADE+LS with the state-of-the-art} \label{sec:SHADEvsSOTA}
+
+In this section we present a brief comparison between the proposed SHADE+LS method and some of the non-heuristic state-of-the-art methods.
+Tables \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
+This time we will only compare the results obtained for ARI.
+
+It should be noted that there are some missing results in these tables.
+In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
+It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
+In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
+
+\begin{table}[!h]
 	\centering
@@
 	\renewcommand{\arraystretch}{1.4}
-	\begin{tabular}{ccccc}
-		\hline
-		Measure & $R^{+}$ & $R^{-}$ & P-value & Alternative hypothesis  \\
-		\hline
-		ARI & 3088 & 1665 & 0.005254 & $H_{max}$ \\
-		Time & 4008 & 1042 & 0.0000001723 & $H_{min}$ \\  
-		Unsat & 3082.5 & 745.5 & 0.0000003838 & $H_{min}$ \\ 
-		\hline 
-	\end{tabular}
-	\caption{Results obtained by the Wilcoxon's signed rank test for the comparison SHADE vs BRKGA}
-	\label{tab:freqTest}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lcccccc}
+			\hline
+			Dataset & SHADE+LS & COPKM & LCVQE & RDPM & TVClust & CECM \\
+			\hline
+			Appendicitis & 0.086 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
+			Boston & -0.001 & 0.010 & 0.010 & 0.003 & 0.007 & - \\
+			Breast Cancer & 0.159 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Circles & 0.063 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
+			Diabetes & 0.003 & 0.022 & 0.005 & 0.001 & 0.002 & - \\
+			Ecoli & 0.020 & - & 0.387 & 0.417 & 0.265 & - \\
+			Glass & 0.013 & 0.184 & 0.268 & 0.197 & 0.211 & - \\
+			Haberman & 0.084 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
+			Hayesroth & 0.025 & - & 0.106 & 0.097 & 0.054 & 0.139 \\
+			Ionosphere & 0.332 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
+			Iris & 0.342 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
+			Led7Digit & 0.004 & 0.497 & 0.425 & 0.369 & 0.316 & - \\
+			Monk2 & 0.270 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
+			Moons & 0.272 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
+			Pima & 0.275 & - & 0.076 & 0.075 & 0.145 & -0.008 \\
+			Saheart & 0.265 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
+			Sonar & 0.116 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
+			Soybean & 0.515 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
+			Spectfheart & 0.343 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
+			Spiral & 0.114 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
+			Tae & 0.024 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
+			Vehicle & 0.006 & - & 0.121 & 0.081 & 0.280 & - \\
+			Vowel & 0.002 & - & 0.063 & -0.003 & 0.067 & - \\
+			Wdbc & 0.357 & - & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Zoo & 0.098 & 0.715 & 0.666 & 0.412 & 0.335 & - \\
+			\hline
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE+LS and the state-of-the-art}
+	\label{tab:results10SOTA}
 \end{table}
 
-For all three considered measures, $P-value < \alpha$, therefore in all cases we can reject $H_0$ and find that there are significant differences between the two methods, giving the adavantage to SHADE + LS with basis to the proposed alternative hypothesis.
-In conclusion, we can safely admit that Wilcoxon's signed rank test provides empirical evidence in favour of SHADE + LS.
-
-\subsection{Applying Bayesian sign test to our results}
-
-Te Bayesian sign test is the Bayesian version of the frequentist non-parametric sign test.
-It allows us to conclude whether there are significant differences between two sets of results.\cite{carrasco2017rnpbst}
-
-It is based on obtaining the $\rho$ parameter of a statistical distribution according to the difference between the results.
-To get the distribution of $\rho$ we count the number of times that $A - B<0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
-In order to identify cases where there are no significant differences we define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, so that $P(A = B) = P(\rho \in rope)$.
-Based on these results we calculate the weights of a Dirichlet distribution and sample it.
-The result of this sampling will be a set of triplets in the form of $[(\rho < r_{min}), (\rho \in rope), (\rho > r_{max}]$.
-
-We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point.
+Table \ref{tab:results10SOTA} presents the results obtained for the $CS_{10}$ constraint set.
+LCVQE, RDPM and TVClust methods stand out, achieving results similar or better than SHADE+LS with a low number of constraints.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lcccccc}
+			\hline
+			Dataset & SHADE+LS & COPKM & LCVQE & RDPM & TVClust & CECM \\
+			\hline
+			Appendicitis & 0.332 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
+			Boston & 0.001 & 0.033 & 0.010 & 0.004 & 0.007 & - \\
+			Breast Cancer & 0.911 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Circles & 0.982 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
+			Diabetes & -0.000 & 0.026 & 0.005 & 0.003 & 0.001 & - \\
+			Ecoli & 0.025 & - & 0.387 & 0.372 & 0.719 & - \\
+			Glass & 0.031 & - & 0.280 & 0.253 & 0.229 & - \\
+			Haberman & 0.796 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
+			Hayesroth & 0.099 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
+			Ionosphere & 0.591 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
+			Iris & 0.260 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
+			Led7Digit & 0.007 & - & 0.425 & 0.490 & 0.324 & - \\
+			Monk2 & 0.945 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
+			Moons & 0.995 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Pima & 0.762 & - & 0.076 & 0.075 & 1.000 & 0.000 \\
+			Saheart & 0.945 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
+			Sonar & 0.890 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
+			Soybean & 0.422 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
+			Spectfheart & 0.778 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
+			Spiral & 0.949 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
+			Tae & 0.066 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
+			Vehicle & 0.018 & 1.000 & 0.121 & 0.081 & 0.576 & - \\
+			Vowel & 0.001 & - & 0.063 & -0.003 & 0.073 & - \\
+			Wdbc & 0.930 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Zoo & 0.128 & 0.435 & 0.642 & 0.450 & 0.353 & - \\
+			\hline
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE+LS and the state-of-the-art}
+	\label{tab:results15SOTA}
+\end{table}
+
+In Table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE+LS.
+On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by SHADE+LS, with the exception of COPKM, which is able to reach the optimum for some of the cases in which it finds a solution.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lcccccc}
+			\hline
+			Dataset & SHADE+LS & COPKM & LCVQE & RDPM & TVClust & CECM \\
+			\hline
+			Appendicitis & 1.000 & - & 0.305 & 0.331 & 0.012 & -0.006 \\
+			Boston & 0.001 & 0.052 & 0.010 & 0.003 & 0.008 & - \\
+			Breast Cancer & 0.977 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Circles & 1.000 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
+			Diabetes & 0.001 & 0.065 & 0.005 & 0.007 & 0.002 & - \\
+			Ecoli & 0.047 & - & 0.387 & 0.459 & 0.763 & - \\
+			Glass & 0.082 & - & 0.271 & 0.287 & 0.218 & - \\
+			Haberman & 0.997 & 1.000 & -0.002 & 0.106 & 0.737 & 0.000 \\
+			Hayesroth & 0.531 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
+			Ionosphere & 0.993 & 1.000 & 0.173 & 0.309 & 0.000 & 0.115 \\
+			Iris & 0.553 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
+			Led7Digit & 0.005 & - & 0.425 & 0.571 & 0.327 & - \\
+			Monk2 & 0.991 & 1.000 & 0.072 & 0.253 & -0.002 & 0.160 \\
+			Moons & 0.995 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
+			Pima & 0.899 & 1.000 & 0.076 & 0.076 & 1.000 & -0.005 \\
+			Saheart & 0.981 & 1.000 & 0.018 & 0.026 & 1.000 & -0.006 \\
+			Sonar & 1.000 & 1.000 & 0.004 & 0.127 & 0.000 & 0.003 \\
+			Soybean & 0.563 & -0.218 & 0.545 & 0.631 & 0.000 & -0.016 \\
+			Spectfheart & 1.000 & 1.000 & -0.107 & -0.112 & 0.000 & -0.054 \\
+			Spiral & 0.998 & 1.000 & -0.003 & 0.011 & 0.006 & 0.045 \\
+			Tae & 0.392 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
+			Vehicle & 0.044 & 1.000 & 0.122 & 0.081 & 0.689 & - \\
+			Vowel & 0.003 & - & 0.063 & -0.003 & 0.071 & - \\
+			Wdbc & 0.972 & 1.000 & 0.486 & 0.522 & 0.000 & 0.000 \\
+			Zoo & 0.153 & 0.821 & 0.642 & 0.439 & 0.335 & - \\
+			\hline
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE+LS and the state-of-the-art}
+	\label{tab:results20SOTA}
+\end{table}
+
+Table \ref{tab:results20SOTA} presents the results obtained for $CS_{20}$.
+For SHADE+LS we continue to observe its progression in accuracy.
+As far as the state-of-the-art methods are concerned, we find a general improvement in ARI, except for CECM.
+Even with this, it is SHADE+LS and COPKM that seem to obtain the highest quality results, with the difference that SHADE+LS is able to produce results in all cases.
+
+\clearpage
+
+\section{Analysis of Results} \label{sec:analisis}
+
+With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
+This way we can statistically determine whether SHADE+LS represents a significant improvement over previous proposals.
+
+In order to carry out this study we will use Bayesian statistical tests, instead of the classic null hypothesis statistical tests (NHST).
+In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
+\textit{"In a nutshell: NHST do not answer the question we ask"}.
+To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
+
+To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
+On the other hand, the hypothesis that the differences between results obtained with different methods are exactly zero is almost always false!
+In the cases where NHST allows us to reject the null hypothesis we conclude that the occurrence of this is unlikely, but this is known even before carrying out experiments.
+
+Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
+This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.
+
+In most cases the researcher would also like to have information about the magnitude of the effects and the uncertainty of its estimate, but NHST do not provide them.
+As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
+
+Furthermore, and this is a situation that all the researchers have faced, NHST do not provide any information about the null hypothesis!
+That is: What can we conclude when NHST do not reject $H_0$?
+We cannot infer anything since NHST cannot provide evidence in its favor.
+
+Finally, there are two other problems that researchers face when performing NHST.
+The choice of the significance level $\alpha$ is critical to the test results, and there is no principle to establish it objectively.
+Additionally, it is worth remembering that, in order to carry out a NHST, there is a previous need to fix and formalize the intentions of the sampling of the results; however, these intentions are commonly fixed a posteriori, which can lead to a misreading of the results.
+
+As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST.
+In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test.
+To make use of it we will employ the R package \texttt{rNPBST}, whose documentation and guide can be found in \cite{carrasco2017rnpbst}.
+
+From now on, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA+LS as sample $A$, and to the results obtained with SHADE+LS as sample $B$.
+
+The Bayesian sign test is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
+To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
+In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, so that $P(A \approx B) = P(\rho \in rope)$.
+Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
+$$[P(\rho < r_{min}) = P(A - B < 0),\;\; P(\rho \in rope),\;\; P(\rho > r_{max}) = P(A - B > 0)]$$
+
+One of the major advantages of the Bayesian sign test over the null hypothesis tests such as the (Wilcoxon's signed-rank test) is that we can obtain a visual representation of its results.
+We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
 With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle.
@@
 	\end{subfigure}
-	\caption{Heat diagrams for each measure}
+	\caption{Heat diagrams for the comparison with BRKGA+LS}
 \end{figure}
@@
 Figure \ref{fig:bayesARI} shows the heatmap associated with the ARI measurement.
-The diagram suggest that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle.
+The diagram suggests that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle.
 It also indicates that in some cases there are no significant differences between $A$ and $B$, since we also find triplets in the rope area.
-Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE + LS performs better than BRKGA + LS.
-
-On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} are very similar, they show the heatmaps obtained for Time and Unsat measurements.
+Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE+LS performs better than BRKGA+LS.
+
+On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} which show the heatmaps obtained for Time and Unsat measurements, are very similar.
 In both, all triplets are represented in the lower right third of the triangle, indicating that the Bayesian sign test assigns a high probability to $A - B > 0$.
-With this, and bearing in mind that both measures must be minimized, we conclude that SHADE + LS represents a solid improvement over BRKGA + LS as far a this measures are concerned.
+With this, and bearing in mind that both measures must be minimized, we conclude that SHADE+LS represents a solid improvement over BRKGA+LS as far as this measures are concerned.
+
+Regarding the comparison of SHADE+LS with the state-of-the-art, we will keep the results obtained by SHADE+LS as $B$, and note the results obtained by the state-of-the-art methods as $A$.
+
+This way, we apply the process described earlier in this section to obtain the heatmaps corresponding to the comparison of SHADE+LS with each of the 5 state-of-the-art methods.
+This time the only measure we compare is ARI.
+Figures \ref{fig:bayesCOPKM} to \ref{fig:bayesCECM} show said comparison.
+
+\begin{figure}[ht!]
+	\centering
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/AriCOPKM.pdf}
+		\caption{SHADE+LS vs COPKM}
+		\label{fig:bayesCOPKM}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/AriLCVQE.pdf}
+		\caption{SHADE+LS vs LCVQE}
+		\label{fig:bayesLCVQE}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/AriRDPM.pdf}
+		\caption{SHADE+LS vs RDPM}
+		\label{fig:bayesRDPM}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/AriTVClust.pdf}
+		\caption{SHADE+LS vs TVClust}
+		\label{fig:bayesTVClust}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/AriCECM.pdf}
+		\caption{SHADE+LS vs CECM}
+		\label{fig:bayesCECM}
+	\end{subfigure}
+	\caption{Heat diagrams for the comparison with the state-of-the-art}
+\end{figure}
+
+We see very similar diagrams for COPKM (\ref{fig:bayesCOPKM}), LVCQE (\ref{fig:bayesLCVQE}), RDPM (\ref{fig:bayesRDPM}) and TVClust (\ref{fig:bayesTVClust}).
+In all of them most of the triplets are represented in the lower middle region of the diagram, with a slight shift of the point cloud to the left.
+That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---, but the advantage is for one or the other depending on the case, being SHADE+LS the one that gets it in most of them.
+However, this does not apply in the case of CECM (\ref{fig:bayesCECM}), where the advantage is indisputably for SHADE+LS.
+
+\section{Conclusions} \label{sec:conclusiones}
+
+In this paper we proposed an adaptation of SHADE to solve the constrained clustering problem.
+We took concepts such as the random-key and the encoder/decoder from BRKGA to transform constrained clustering into a real-domain optimization problem.
+We also proposed a new fitness function that has proven to be more adaptable and less restrictive.
+
+The SHADE algorithm is one of the most successful methods to approximately solve difficult problems, and constrained clustering does not appear to be an exception.
+We obtained results on 25 different datasets and 3 different constraint sets for each one, applying all the resolution methods named in this paper.
+Thanks to the Bayesian statistical tests, which offer us the advantageous option of producing a visualization of differences between results obtained by different methods, we were able to objectively prove that the SHADE+LS approach is significantly better than previous population-based proposals.
+As far as the state-of-the-art is concerned, the SHADE+LS approach has proven to be better or equivalent when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
+
+\section{Acknowledgements}
 
@@
 
-\section{Conclusions and Future Work}
-
-\section{Acknowledgements}
-
-\clearpage
-
 \section*{References}
@@
 \end{document}
-
-\begin{figure}[bth]
-	\myfloatalign
-	%\subfloat[ARI]
-	{\includegraphics[width=.45\linewidth]{Figures/BayesSignoARI.pdf}} \quad
-	%\subfloat[Time]
-	{\includegraphics[width=.45\linewidth]{Figures/BayesSignoTime.pdf}} \quad
-	%\subfloat[Unsat]
-	{\includegraphics[width=.45\linewidth]{Figures/BayesSignoUnsat.pdf}} \quad
-	\caption{Heat diagrams for each measure obtained.}
-	\label{img:heatDiag}
-\end{figure}

### 2019-03-09 Software/CEKM/CEKM.py

@@ -1,9 +1,8 @@
 import numpy as np
 from numpy import matlib
-import skfuzzy as fuzz
-from set_centers_ecm import set_centers_ecm
-from set_distances import set_distances
-from solqp import solqp
+from CEKM.set_centers import set_centers_ecm
+from CEKM.set_distances import set_distances
+from CEKM.solqp import solqp
 
 
 def get_bit(numb, K, ind):
@@ -45,10 +44,8 @@ def CEKM(X, K, constraints, max_iter=300, rho=100, xi=0.5, stop_thr=1e-3, init='
     beq = np.ones((rows, 1))
 
     # centroids inicialization
-    if (init == 'rand'):
-        g = np.random.rand(K, cols) * np.matlib.repmat(np.max(X) - np.min(X), K, 1) + np.matlib.repmat(np.min(X), K, 1)
-    else:
-        g = fuzz.cluster.cmeans(X.T, K, 2, 1e-5, 100)[0]
+    g = np.random.rand(K, cols) * np.matlib.repmat(np.max(X) - np.min(X), K, 1) + np.matlib.repmat(np.min(X), K, 1)
+
 
     # centers calculus for all the subsets
     gplus = np.zeros((nb_foc-1,cols))

### 2019-03-09 Software/EstimandoXi.py

@@ -7,56 +7,58 @@ import sys
 
 def main():
 
-    np.random.seed(11)
+	np.random.seed(11)
 
-    #CARGA DE DATOS
+	#CARGA DE DATOS
 
-    names_array, datasets_array, labels_array = load_all_datasets()
+	names_array, datasets_array, labels_array = load_all_datasets()
 
-    names_array = names_array[0:5]
-    datasets_array = datasets_array[0:5]
-    labels_array = labels_array[0:5]
+	names_array = names_array[0:5]
+	datasets_array = datasets_array[0:5]
+	labels_array = labels_array[0:5]
 
-    const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    const_array = load_constraints(names_array, const_percent_vector)
+	const_percent_vector = [0.05, 0.1, 0.15, 0.2]
+	const_array = load_constraints(names_array, const_percent_vector)
 
 
-    #BUCLE DE OBTENCION DE DATOS
+	#BUCLE DE OBTENCION DE DATOS
 
-    max_eval = 100
+	max_eval = 300000
 
-    for p in range(0, 6):
+	for p in range(4, 6):
 
-        xi = p/10.0
+		xi = -p/10.0
 
-        results_matrix = np.zeros((len(names_array), 4))
+		results_matrix = np.zeros((len(names_array), 4))
 
-        for i in range(len(names_array)):
+		for i in range(len(names_array)):
 
-            data_set = datasets_array[i]
-            labels = labels_array[i]
-            nb_clust = len(set(labels))
+			print("Procesando " + names_array[i] + " dataset")
 
-            for j in range(len(const_array)):
+			data_set = datasets_array[i]
+			labels = labels_array[i]
+			nb_clust = len(set(labels))
 
-                const = const_array[j][i]
-                ml_const, cl_const = get_const_list(const)
+			for j in range(len(const_array)):
 
-                ils = ILSNueva(data_set, ml_const, cl_const, nb_clust, 0.3, 300, 0.3, xi)
-                ils_assignment = ils.run(max_eval)[0]
-                ars = adjusted_rand_score(labels, ils_assignment)
+				const = const_array[j][i]
+				ml_const, cl_const = get_const_list(const)
 
-                results_matrix[i, j] = ars
+				ils = ILSNueva(data_set, ml_const, cl_const, nb_clust, 0.3, 300, 0.3, xi)
+				ils_assignment = ils.run(max_eval)[0]
+				ars = adjusted_rand_score(labels, ils_assignment)
 
[…]

### 2019-03-09 Software/ILSNueva/ILSNueva.py

@@ -180,8 +180,8 @@ class ILSNueva:
         restarts_array = []
         restarts = 0
 
-        #while self._evals_done < max_evals:
-        while len(gen_times) < max_evals:
+        while self._evals_done < max_evals:
+        #while len(gen_times) < max_evals:
             start_gen = time.time()
             worst = np.argmax(self._best_fitness)
             best = (worst + 1) % 2
@@ -208,7 +208,7 @@ class ILSNueva:
 
             gen_times.append(time.time() - start_gen)
             restarts_array.append(restarts)
-            print(str(len(gen_times)) + " " + str(restarts))
+            #print(str(len(gen_times)) + " " + str(restarts))
 
         best = np.argmin(self._best_fitness)
         return self._best_solution[best, :], np.sum(gen_times), restarts_array

### 2019-03-09 Software/TFG_CEKM.py

@@ -1,0 +1,68 @@
+import numpy as np
+from CEKM.CEKM import CEKM
+from functions import *
+from sklearn.metrics import adjusted_rand_score
+import time
+import sys
+
+def main():
+
+	if len(sys.argv) != 2:
+		print("Numero de argumentos incorrecto")
+		return -1
+
+	np.random.seed(11)
+
+	#CARGA DE DATOS
+
+	names_array, datasets_array, labels_array = load_all_datasets()
+
+	# datasets_array = datasets_array[5:7]
+	# names_array = names_array[5:7]
+	# labels_array = labels_array[5:7]
+	
+	const_percent_vector = [float(sys.argv[1])]
+	const_array = load_constraints(names_array, const_percent_vector)
+	results_array = []
+
+	#BUCLE DE OBTENCION DE DATOS
+
+	nb_runs = 1
+
+	for i in range(len(names_array)):
+
+		print("Procesando " + names_array[i] + " dataset")
+
+		data_set = datasets_array[i]
+		labels = labels_array[i]
+		nb_clust = len(set(labels))
+
+		const = const_array[0][i]
+		mean_ars = 0
+		broke = False
+		for j in range(nb_runs):
+			try:
+				cekm_assignment = CEKM(data_set, nb_clust, np.asmatrix(const))[0]
+				mean_ars += adjusted_rand_score(labels, cekm_assignment)
+			except:
+				print("Excepcion manejada")
+				print(sys.exc_info()[0])
+				mean_ars += -1
+				broke = True
+				break
+
+
+		if not broke:
+			mean_ars /= nb_runs
+
+		results_array.append(mean_ars)
+		print(names_array[i].title() + " & %.3f" % results_array[i] + " \\\\")
+
+	print("-------------------- Resultados --------------------")
+
+	for i in range(len(results_array)):
+
+		print(names_array[i].title() + " & %.3f" % results_array[i] + " \\\\")
+
+
+if __name__ == "__main__": main()

### 2019-03-14 Papers/SHADE/paperSHADE.tex
Correcciones en paper SHADE

@@
 \begin{abstract}
-This paper proposes an adaptation of the successful SHADE method to solve the constrained clustering problem.
-This problem consists of a modification of the traditional clustering to consider a new type of information, which is given in the form of must-link and cannot-link constraints.
-We will compare the results obtained on 25 datasets by this new SHADE adaptation with a previous heuristic-based approach---BRKGA---as well as with some of the state-of-the-art-methods, supporting our conclusions with the aid of Bayesian tests.
+Clustering has always been a powerful tool in knowledge discovery.
+Traditionally unsupervised, it has received attention when adding new types of information to it has proven to provide better results than completely unsupervised procedures, leading to a new kind of semi-supervised learning.
+This paper proposes a variant of the successful SHADE method to solve the constrained clustering problem.
+This problem consists of a generalization of the traditional clustering to consider a new type of information, which is given in the form of constraints.
+This paper focuses on instance-level must-link and cannot-link constraints.
+We will compare the results obtained on 25 datasets, in conjunction with various constraint sets, by this new proposal with a previous nature-inspired approach---BRKGA---as well as with some of the state-of-the-art algorithms, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 \begin{keyword}
-constrained clustering, must-link, cannot-link, genetic algorithm, differential evolution.
+constrained clustering, instance-level, must-link, cannot-link, genetic algorithm, differential evolution, Bayesian statistical tests.
 \end{keyword}
@@
 It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
-Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology and genetics among others \cite{ClusterAnalysis}.
+Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology and genetics among others \cite{Everitt:2009:CA:1538772}.
 
 In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
-The main difference is that in hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{ClusterAnalysis}.
+The main difference is that in hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
+A representative example of partitional clustering is the widely studied and well-known algorithm K-means, while for hierarchical clustering the CURE method should be highlighted \cite{wu2009top, guha1998cure}.
 In this paper we will focus on partitional clustering.
-\textcolor{red}{Poner ejemplos de cada uno (con referencias)?}
 
@@
 
-\textcolor{red}{Bucar una referencia fiable de ssl}
+Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
+We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering.
+The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
+The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper.
+\cite{triguero2015self}
 
@@
 Pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
-On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure.\cite{davidson2007survey} This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \textcolor{red}{X}.
+On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure.\cite{davidson2007survey} This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
 
@@
 
-\textcolor{red}{Cambiar el parrafo de abajo en funcion de como quede la organizacion de secciones final}
-
 Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
 In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
-Afterward, a variant of DE algorithm will be presented in Section \ref{sec:SHADE}, followed by its adaptation for constrained clustering and a new fitness function.
+Afterward, a variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
+In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
 Sections \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis.
@@
 
-In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering \ref{sec:BackSOTA}.
+In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering (Section \ref{sec:BackSOTA}).
 We will also present the basis of the DE optimization method (Section \ref{sec:BackDE}).
@@
 This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified.
-In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.
+In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$.
+\cite{davidson2007survey}
 
@@
 
-\textcolor{red}{Opciones para cambiar el nombre de los conjuntos de restricciones: $CS_=$ y $CS_{\neq}$, con $CS$ para referirse a las restricciones en general.
-O simplemente $ML$ y $CL$, con $CS$ para referirse a las restricciones en general}
-
 \subsection{The Feasibility Problem} \label{sec:BackFeas}
@@
 	\textbf{Feasibility Problem}: given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
-\cite{davidson2007survey}\cite{davidson2005clustering}
+\cite{davidson2005clustering}
 	\label{def1}
@@
 
-Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,D]}\}$ that serve as parameters for the function to be optimized.
+Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,v]}\}$ that serve as parameters for the function to be optimized.
 These individuals are considered solutions to the problem.
@@
 To guide the optimization process, we use a fitness function.
-The task of DE is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^D \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
+The task of DE is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^{v} \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
 
@@
 
-	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of a given number of $D$-dimensional arrays of parameters.
+	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of a given number of $v$-dimensional arrays of parameters.
 These vectors constitute candidate solutions for the problem to be solved.
@@
 
-The biased random-key genetic algorithm---from now on BRKGA---was first proposed in \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.
-
-In BRKGA, each solution---also called individual---is represented as a vector of values within the interval $[0,1]$.
-However, this vector is not a solution to the problem but a representation of it.
-This allows us to abstract the details of the problem in order to apply a genetic algorithm that operates with real coding.
-Therefore, a deterministic decoder is necessary to obtain the actual solution to the problem from the vector that encodes it.
-
-BRKGA also requires a fitness function to evaluate each individual.
-This function takes the decoded solution as input and provides its fitness value as output.
-It is designed specifically for each problem.
-
-The BRKGA optimization process is completely independent of the problem to solve.
-First, a population $P$ of $p_{size}$ vectors of random-keys is initialized, and this will be the initial generation.
-Each random-key vector $p_i$ has $N$ random-keys in it.
+Genetic Algorithms (GA) emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
+It is one of the most studied and successful evolutionary algorithms thanks in part to its simplicity and easy implementation.
+\cite{fausto2019ants}
+
+In GA a Population $P$ of a given number of solutions $p_i = \{p_{[i,1]}, \cdots, p_{[i,v]}\}$ is initialized, these solutions are also called chromosomes or individuals.
+Each element of a chromosome is called gene and the value it takes is known as allele.
+To simulate the natural processes of evolution and natural selections the GA uses a set of evolutionary operator similar to the one used by DE, this is: crossover, mutation and selection.
+The crossover operator consist of combining two different chromosomes (parents) from the population and obtaining a new chromosome (offspring) which inherit alleles for its genes from both parents.
+The mutation operator is used by GA to avoid local optimum, it consist of introducing random changes in the population to favor diversity.
+The selection operator is uses to determine if a chromosome is suitable to survive for the next generation, to accomplish this task it uses a fitness function, which assign a value to each chromosome in order to determine if it is a good solution to the problem or not.
+GA applies these operator through a series of generations to hopefully obtain a solution for the problem to solve close to the optimal.
+\cite{fausto2019ants}
+
+GAs have been applied to all kinds of problems, including problems whose solutions are subject to strong restrictions such as machine scheduling, resource location and the quadratic assignment problem.
+These problems present the peculiarity that, by applying generalized crossover operators to two individuals that represent a feasible solution to the problem, we can obtain one that is not feasible.
+The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to this kind of problems, it uses random numbers to represent solutions to the problem (individuals), which are used as keys to obtain an actual solutions.
+Random-key vectors are sampled from the space $[0,1]^{v}$, which is used as a surrogate space for the literal solutions space.
+Points in the random-key space are transfered to the literal solutions space via a decoder for fitness evaluation.
+A decoder is a deterministic procedure that allow us to map random-key vectors to the literal solutions space.
+We can build a decoder that always provides as a result a feasible solution to the problem we want to solve, thus avoiding the feasibility problem that we found in the classic GA.
+
+The Biased Random-key Genetic Algorithm (BRKGA) was first proposed in \cite{gonccalves2011biased} as a variant of the RKGA.
+In the BRKGA optimization process, first, the population $P$ of $p_{size}$ vectors of random-keys is initialized, this will be the initial generation.
+Each random-key vector $p_i$ has $v$ random-keys in it.
 The population is sorted by the fitness value $f_i$ of each $p_i$ to select the first $p_e$ individuals, this is, the elite of the population.
@@
 To introduce diversity, a number $p_m$ of new random-keys vectors are also included in the next generation.
-The remaining individuals of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents \cite{de2017comparison}.
-
-\subsection{Adaptation of BRKGA for Constrained Clustering} \label{sec:AdaptationofBRKGA}
+The remaining individuals of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents.
+
+It should be noted that in BRKGA there is a clear differentiation between problem-dependent and problem-independent elements.
+The independent part of the algorithm has no knowledge of the problem it solves, it is limited to performing the operations related to the population optimization process.
+The dependent part of the problem is the decoder and the fitness function, which allow us to obtain real solutions for the problem and its evaluation.
+Therefore, to specify a BRKGA optimization scheme it is only necessary to define the decoder and the fitness function.
+\cite{gonccalves2011biased}
+
+\subsection{BRKGA optimization scheme for Constrained Clustering} \label{sec:AdaptationofBRKGA}
 
@@
 
-This method randomly initializes a population of random-key vectors, each one with $n$ random-keys.
+In this scheme each random-key vector has $n$ random-keys in it.
 The decoder divides the interval $[0,1]$ in $k$ intervals, so there exists a correspondence between each random-key (instance $x_i$) and the integer (label $l_i$) corresponding to the interval which it lies in.
@@
 
+Since SHADE is an evolution-based algorithm, it shares elements with BRKGA (Section \ref{sec:brkga}), so we will use a similar notation to define its elements.
+We will note the population as $P$, formed by $p_{size}$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
+
 \subsection{SHADE Operators}
 
-As in differential evolution, SHADE employs mutant generation, crossover and replacement operators to explore the space of solutions and bring in new generations of individuals.
+As in differential evolution, SHADE employs mutant generation, crossover and selection operators to explore the space of solutions and bring in new generations of individuals.
 
@@
 Indices $r1$ and $r2$ are random values in the range $[0,p_{size}]$, different from $i$ and from each other.
-The individual $p_{[pbest, G]}$ is randomly selected from among the best $N \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
+The individual $p_{[pbest, G]}$ is randomly selected from among the best $p_{size} \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
 This way, the parameter $pbest$ controls the greediness of the mutation strategy.
@@
 
-\noindent where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.
-
-The SHADE replacement operator determines the survivors individuals for the next generation.
+\noindent where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,v]$, and $CR \in [0,1]$ is the crossover ratio.
+
+The SHADE selection operator determines the survivors individuals for the next generation.
 It compares each parent $x_{[i,G]}$ with the trial vector $_{[j,i,G]}$, generated based on it.
@@
 
-\subsection{Adaptation of SHADE for constrained clustering} \label{sec:SHADEadapt}
+\section{Adaptation of SHADE for constrained clustering} \label{sec:SHADEadapt}
 
@@
 		$P_G \leftarrow$ \Sort{$P_G$} \tcp{Sort the population}
-		\tcp{Apply a LS procedure to the elite of the population}
+		\tcp{Apply an LS procedure to the elite of the population}
 		\LocalSearch{$\{p_{[1,G]}, \cdots, p_{[p_{e},G]}\}$}\\
@@
 			$p_{[i,G]} \leftarrow$ rand$[N/2, 0.2]$\\
-			Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin\\
+			\tcp{Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin}
 
@@
 		}
-		Whenever $|A| > |P|$, randomly select an individual from $A$ to be deleted so that $|A| \le |P|$\\
+		\tcp{Whenever $|A| > |P|$, randomly select an individual from $A$ to be deleted so that $|A| \le |P|$}
 		\If{$S_{CR} \neq \emptyset$ \textbf{and} $S_{F} \neq \emptyset$}{
@@
 		Appendicitis & 106 & 2 & 7 \\
-		Boston & 506 & 229 & 13 \\
 		Breast Cancer & 569 & 2 & 30 \\
+		Bupa & 345 & 2 & 6 \\
 		Circles & 300 & 2 & 2 \\
-		Diabetes & 442 & 214 & 10 \\
 		Ecoli & 336 & 8 & 7 \\
@@
 		Moons & 300 & 2 & 2 \\
+		Movement Libras & 360 & 15 & 90 \\
 		Pima & 768 & 2 & 8 \\
@@
 		Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
-		Boston & 3 & 1272 && 13 & 2837 && 26 & 5125 \\
 		Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
+		Bupa & 323 & 272 && 699 & 627 && 1201 & 1145 \\
 		Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
-		Diabetes & 5 & 985 && 7 & 2204 && 23 & 3893 \\
 		Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
@@
 		Moons & 200 & 235 && 494 & 496 && 900 & 870 \\
+		Movement Libras & 27 & 603 && 112 & 1319 && 158 & 2398 \\
 		Pima & 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
@@
 			Appendicitis & 0.022 & 5.454 & 383.642 && 0.086 & 5.091 & 366.491 \\
-			Boston & -0.000 & 0.173 & 1502.452 && -0.001 & 0.235 & 1570.547 \\
-			Breast\_Cancer & 0.395 & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
+			Breast Cancer & 0.395 & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
+			Bupa & 0.281 & 14.151 & 1420.117 && 0.104 & 16.672 & 1409.487 \\
 			Circles & 0.166 & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
-			Diabetes & 0.000 & 0.364 & 1277.982 && 0.003 & 0.465 & 1284.486 \\
 			Ecoli & 0.009 & 20.321 & 1169.798 && 0.020 & 12.763 & 1112.981 \\
@@
 			Moons & 0.277 & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
+			Movement Libras & 0.004 & 4.159 & 1508.279 && 0.003 & 3.396 & 1519.185 \\
 			Pima & 0.236 & 25.195 & 4072.087 && 0.275 & 22.468 & 3990.383 \\
@@
 			Appendicitis & 0.296 & 6.167 & 381.625 && 0.332 & 5.833 & 362.106 \\
-			Boston & 0.000 & 0.526 & 1648.008 && 0.001 & 0.547 & 1688.985 \\
-			Breast\_Cancer & 0.772 & 10.172 & 3854.911 && 0.911 & 3.310 & 3766.402 \\
+			Breast Cancer & 0.772 & 10.172 & 3854.911 && 0.911 & 3.310 & 3766.402 \\
+			Bupa & 0.840 & 5.973 & 1463.152 && 0.801 & 5.641 & 1437.639 \\
 			Circles & 0.828 & 6.101 & 1136.484 && 0.982 & 0.404 & 1108.928 \\
-			Diabetes & 0.004 & 0.317 & 1364.941 && -0.000 & 0.371 & 1424.047 \\
 			Ecoli & 0.015 & 24.314 & 1200.463 && 0.025 & 19.090 & 1127.568 \\
@@
 			Moons & 0.958 & 1.334 & 1133.858 && 0.995 & 0.020 & 1113.687 \\
+			Movement Libras & 0.006 & 8.679 & 1584.028 && 0.001 & 8.889 & 1590.321 \\
 			Pima & 0.481 & 22.051 & 4451.193 && 0.762 & 9.928 & 4402.693 \\
@@
 			\hline
-			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
-			Boston & 0.001 & 0.636 & 1820.313 && 0.001 & 0.660 & 1878.863 \\
-			Breast\_Cancer & 0.807 & 9.086 & 4061.162 && 0.977 & 0.891 & 3984.370 \\
+			Breast Cancer & 0.807 & 9.086 & 4061.162 && 0.977 & 0.891 & 3984.370 \\
+			Bupa & 0.936 & 2.975 & 1529.151 && 0.993 & 0.281 & 1508.772 \\
 			Circles & 0.935 & 2.814 & 1180.268 && 1.000 & 0.000 & 1170.452 \\
-			Diabetes & 0.003 & 0.695 & 1480.035 && 0.001 & 0.710 & 1589.821 \\
 			Ecoli & 0.028 & 25.716 & 1267.419 && 0.047 & 21.466 & 1277.903 \\
@@
 			Moons & 0.937 & 2.689 & 1177.848 && 0.995 & 0.158 & 1157.222 \\
+			Movement Libras & 0.006 & 8.381 & 1654.159 && 0.004 & 7.676 & 1645.174 \\
 			Pima & 0.753 & 11.544 & 4855.446 && 0.899 & 4.419 & 4820.684 \\
@@
 			Appendicitis & 0.086 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
-			Boston & -0.001 & 0.010 & 0.010 & 0.003 & 0.007 & - \\
 			Breast Cancer & 0.159 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.104 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
 			Circles & 0.063 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
-			Diabetes & 0.003 & 0.022 & 0.005 & 0.001 & 0.002 & - \\
 			Ecoli & 0.020 & - & 0.387 & 0.417 & 0.265 & - \\
@@
 			Moons & 0.272 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
+			Movement Libras & 0.003 & 0.285 & 0.293 & 0.256 & 0.000 & - \\
 			Pima & 0.275 & - & 0.076 & 0.075 & 0.145 & -0.008 \\
@@
 			Appendicitis & 0.332 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Boston & 0.001 & 0.033 & 0.010 & 0.004 & 0.007 & - \\
 			Breast Cancer & 0.911 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Bupa & 0.801 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
 			Circles & 0.982 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Diabetes & -0.000 & 0.026 & 0.005 & 0.003 & 0.001 & - \\
 			Ecoli & 0.025 & - & 0.387 & 0.372 & 0.719 & - \\
@@
 			Moons & 0.995 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Movement Libras & 0.001 & -0.742 & 0.293 & 0.255 & 0.000 & - \\
 			Pima & 0.762 & - & 0.076 & 0.075 & 1.000 & 0.000 \\
@@
 			Appendicitis & 1.000 & - & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Boston & 0.001 & 0.052 & 0.010 & 0.003 & 0.008 & - \\
 			Breast Cancer & 0.977 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.993 & 1.000 & -0.005 & -0.007 & -0.006 & 0.000 \\
 			Circles & 1.000 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
-			Diabetes & 0.001 & 0.065 & 0.005 & 0.007 & 0.002 & - \\
 			Ecoli & 0.047 & - & 0.387 & 0.459 & 0.763 & - \\
@@
 			Moons & 0.995 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
+			Movement Libras & 0.004 & - & 0.293 & 0.293 & 0.000 & - \\
 			Pima & 0.899 & 1.000 & 0.076 & 0.076 & 1.000 & -0.005 \\

### 2019-03-15 Papers/SHADE/paperSHADE.tex
Añadida fila de medias a las tablas

@@
 
-Genetic Algorithms (GA) emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
-It is one of the most studied and successful evolutionary algorithms thanks in part to its simplicity and easy implementation.
+Genetic Algorithms (GAs) emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
+It is one of the most studied and successful evolutionary algorithms, thanks in part to its simplicity and ease of implementation.
 \cite{fausto2019ants}
 
-In GA a Population $P$ of a given number of solutions $p_i = \{p_{[i,1]}, \cdots, p_{[i,v]}\}$ is initialized, these solutions are also called chromosomes or individuals.
+In a GA, a population $P$ of a given number of solutions $p_i = \{p_{[i,1]}, \cdots, p_{[i,v]}\}$ is first initialized; these solutions are also called chromosomes or individuals.
 Each element of a chromosome is called gene and the value it takes is known as allele.
-To simulate the natural processes of evolution and natural selections the GA uses a set of evolutionary operator similar to the one used by DE, this is: crossover, mutation and selection.
-The crossover operator consist of combining two different chromosomes (parents) from the population and obtaining a new chromosome (offspring) which inherit alleles for its genes from both parents.
-The mutation operator is used by GA to avoid local optimum, it consist of introducing random changes in the population to favor diversity.
-The selection operator is uses to determine if a chromosome is suitable to survive for the next generation, to accomplish this task it uses a fitness function, which assign a value to each chromosome in order to determine if it is a good solution to the problem or not.
-GA applies these operator through a series of generations to hopefully obtain a solution for the problem to solve close to the optimal.
+To simulate the natural processes of evolution and natural selection, the set of evolutionary operators that GAs make use of is similar to that used by DE: crossover, mutation and selection.
+The crossover operator consists in combining two different chromosomes (parents) from the population and obtaining a new chromosome (offspring) which inherits alleles for its genes from both parents.
+The mutation operator is used by GAs to avoid local optima, and it consists in introducing random changes into the population to favor diversity.
+The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function whose task is to assign a value to each chromosome to assess its quality as a solution for the problem.
+GAs apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal.
 \cite{fausto2019ants}
@@
 GAs have been applied to all kinds of problems, including problems whose solutions are subject to strong restrictions such as machine scheduling, resource location and the quadratic assignment problem.
-These problems present the peculiarity that, by applying generalized crossover operators to two individuals that represent a feasible solution to the problem, we can obtain one that is not feasible.
-The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to this kind of problems, it uses random numbers to represent solutions to the problem (individuals), which are used as keys to obtain an actual solutions.
-Random-key vectors are sampled from the space $[0,1]^{v}$, which is used as a surrogate space for the literal solutions space.
-Points in the random-key space are transfered to the literal solutions space via a decoder for fitness evaluation.
-A decoder is a deterministic procedure that allow us to map random-key vectors to the literal solutions space.
-We can build a decoder that always provides as a result a feasible solution to the problem we want to solve, thus avoiding the feasibility problem that we found in the classic GA.
+These problems present the peculiarity that, by applying generalized crossover operators to two individuals that represent a feasible solution to the problem, we can obtain non-feasible ones.
+The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to this kind of problems.
+It uses random numbers to represent solutions to the problem (individuals), which are used as keys to obtain the actual solutions.
+Random-key vectors are sampled from the space $[0,1]^{v}$, which is used as a surrogate space for the literal solution space.
+Points in the random-key space are transferred to the literal solution space via a decoder for fitness evaluation.
+A decoder is a deterministic procedure that allows us to map random-key vectors to the literal solution space.
+We can build a decoder that always provides as a result a feasible solution to the problem we want to solve, thus avoiding the feasibility problem that we found in classic GAs.
 
 The Biased Random-key Genetic Algorithm (BRKGA) was first proposed in \cite{gonccalves2011biased} as a variant of the RKGA.
-In the BRKGA optimization process, first, the population $P$ of $p_{size}$ vectors of random-keys is initialized, this will be the initial generation.
+In the BRKGA optimization process, first, the population $P$ of $p_{size}$ vectors of random-keys is initialized; this will be the initial generation.
 Each random-key vector $p_i$ has $v$ random-keys in it.
@@
 The independent part of the algorithm has no knowledge of the problem it solves, it is limited to performing the operations related to the population optimization process.
-The dependent part of the problem is the decoder and the fitness function, which allow us to obtain real solutions for the problem and its evaluation.
+The dependent part of the problem is composed of the decoder and the fitness function, which allow us to obtain actual solutions for the problem and its evaluation.
 Therefore, to specify a BRKGA optimization scheme it is only necessary to define the decoder and the fitness function.
@@
 
-An adaptation of BRKGA for constrained clustering was proposed in \cite{de2017comparison}.
-
-In this scheme each random-key vector has $n$ random-keys in it.
+An adaptation of BRKGA for constrained clustering was proposed in \cite{de2017comparison}, we will refer to it as BRKGA$_{CC}$.
+
+In this scheme each random-key vector has $n$ random-keys in it ($v = n$).
 The decoder divides the interval $[0,1]$ in $k$ intervals, so there exists a correspondence between each random-key (instance $x_i$) and the integer (label $l_i$) corresponding to the interval which it lies in.
@@
 
-The SHADE method stores in memory a structure with $H$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:SHADEmemory}.
+The SHADE method stores in memory a structure $M$ with $|M|$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:SHADEmemory}.
 This structure is initialized with the value $0.5$ for all entries.
@@
 		\hline
-		 Index & 1 & 2 & $\cdots$ & $H - 1$ & $H$ \\
-		 \hline
-		 \hline
-		 $M_{CR}$ & $M_{[CR,1]}$ & $M_{[CR,1]}$ & $\cdots$ & $M_{[CR,H-1]}$ & $M_{[CR,H]}$ \\
-		 \hline
-		 $M_{F}$ & $M_{[F,1]}$ & $M_{[F,1]}$ & $\cdots$ & $M_{[F,H-1]}$ & $M_{[F,H]}$ \\
+		 Index & 1 & 2 & $\cdots$ & $|M| - 1$ & $|M|$ \\
+		 \hline
+		 \hline
+		 $M_{CR}$ & $M_{[CR,1]}$ & $M_{[CR,1]}$ & $\cdots$ & $M_{[CR,|M|-1]}$ & $M_{[CR,|M|]}$ \\
+		 \hline
+		 $M_{F}$ & $M_{[F,1]}$ & $M_{[F,1]}$ & $\cdots$ & $M_{[F,|M|-1]}$ & $M_{[F,|M|]}$ \\
 		\hline
@@
 \text{mean}_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
-M_{[CR,h,G+1]} &  otherwise
+M_{[CR,h,G]} &  otherwise
 \end{array}
@@
 \text{mean}_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
-M_{[F,h,G+1]} &  otherwise
+M_{[F,h,G]} &  otherwise
 \end{array}
@@
 
-The $h \;\; (0 \le h \le H)$ index specifies the position of the memory $M$ to be updated.
+The $h \;\; (0 \le h \le |M|)$ index specifies the position of the memory $M$ to be updated.
 This index is initialized to $0$ at the beginning of the optimization process and increased by one after every generation.
-When $h \ge H$ then $h$ is reset to 1.
+When $h \ge |M|$ then $h$ is reset to 1.
 It is worth noting that, when there is a generation with no trial vectors successfully replacing their parents, no update of $M$ is done.
@@
 
-\section{Adaptation of SHADE for constrained clustering} \label{sec:SHADEadapt}
-
-For the adaptation of SHADE to the problem of constrained clustering, we will take some of the elements that allowed us to adapt it to BRKGA.
+\section{SHADE optimization scheme for constrained clustering} \label{sec:SHADEadapt}
+
+In this section, we present an optimization scheme for the SHADE algorithm when applied to constrained clustering.
+We will call this new approach $SHADE_{CC}$.
 To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of $n$ random-keys.
-Additionally, we will also reuse the same initialization method and the same decoder used in BRKGA (see Section \ref{sec:AdaptationofBRKGA}).
+Additionally, we will also make use of the initialization method and the decoder found in Section \ref{sec:AdaptationofBRKGA}.
+However, we propose a new fitness function to evaluate the quality of the individuals of the population, Equation \eqref{eq14}, whose element $z_i$ is obtained as in Equation \eqref{eq2}.
 
@@
 	\SetKwFunction{LocalSearch}{LocalSearch}
-	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $K$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $k$.}
 	\tcp{Initialization phase}
@@
 		\For{$i \in [1,p_{size}]$}{
-			$r_i \leftarrow$ randInt$[1,H]$\\
+			$r_i \leftarrow$ randInt$[1,|M|]$\\
 			$CR_{[i,G]} \leftarrow$ randn$_i(M_{[CR,r_i]}, 0.1)$\\
@@
 		}
-		\tcp{Whenever $|A| > |P|$, randomly select an individual from $A$ to be deleted so that $|A| \le |P|$}
+		\tcc{Whenever $|A| > |P|$, randomly select an individual from $A$ to be deleted so that $|A| \le |P|$}
 		\If{$S_{CR} \neq \emptyset$ \textbf{and} $S_{F} \neq \emptyset$}{
 			Update $M_{[CR,h]}$ and $M_{[F,h]}$ based on $S_{CR}$ and $S_{F}$\\
-			$h \leftarrow (h + 1) \mod H$
+			$h \leftarrow (h + 1) \mod |M|$
 		}
@@
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, decoded random-key vector (solution) $S$, number of clusters $K$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, decoded random-key vector (solution) $S$, number of clusters $k$.}
 	\BlankLine
@@
 		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
-		\For{$c \in RSL$}{
+		\For{$l \in RSL$}{
 			$S^\prime \leftarrow S$\\
-			$S^\prime$\texttt{[}$s_i$\texttt{]} $\leftarrow c$ \tcp{Move object $s_i$ to cluster $c$}
-
+			\tcp{Move object $s_i$ to the cluster associated with label $l$}
+			$S^\prime$\texttt{[}$s_i$\texttt{]} $\leftarrow l$\\
 			\If{$f(S^\prime) < f(S)$}{
@@
 
-For our experiments we will compare the results obtained by BRKGA and SHADE over 25 datasets.
+For our experiments we will compare the results obtained by BRKGA and SHADE over 25 datasets and 3 constraints set for each one of them.
 Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
@@
 	\end{tabular}%}
-	\caption{Summary of datasets used for the experiments \cite{triguero2017keel}\cite{scikit-learn}}
+	\caption{Summary of datasets used for the experiments}
 	\label{tab:datasets}
@@
 
+Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
+They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Rand Index (see Section \ref{sec:EvalMet}).
+
 \clearpage
 
-\subsection{Constraint Generation}
+\subsection{Constraint Generation} \label{sec:ConstGent}
 
 Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets.
-This method consists in randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
+This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
 
 We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with four small percentages of the size of the dataset: 10\%, 15\% and 20\%.
-With $n$ being the fraction of the size of the dataset associated to each of these percentages, the formula $\frac{n(n-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n$ vertices would have.
+With $n_f$ being the fraction of the size of the dataset associated to each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
 
@@
 
-\subsection{Evaluation Method}
+\subsection{Evaluation Method} \label{sec:EvalMet}
 
 Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
-We will use the Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
-The Basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
-$C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions \cite{rand1971objective}.
-
-For each pair of instances $x_i$ and $x_j$ in $X$, $C_i$ assigns them to the same cluster or to different clusters.
+We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
+The basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
+$C_1$ and $C_2$ are viewed as collections of $n(n - 1)/2$ pairwise decisions \cite{rand1971objective}.
+
+For each pair of instances $x_i$ and $x_j$ in $X$, a partition assigns them to the same cluster or to different clusters.
 We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in both $C_1$ and $C_2$, and $b$ as the opposite event ($x_i$ and $x_j$ are in different clusters in $C_1$ and $C_2$).
@@
 \begin{equation}
-Rand(C_1, C_2) = \frac{a + b}{N(N - 1)/2}
+Rand(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
 \label{eq15}
@@
 
+Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper.
+To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth.
+
 %% CORREGIR DESDE AQUI HASTA EL FINAL
@@
 
-Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA+LS and SHADE+LS algorithms.
+Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA$_{CC}$ and SHADE$_{CC}$ algorithms.
 
@@
 
-In both cases---BRKGA+LS and SHADE+LS---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
 
@@
 
-In this section we present Tables \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.
+In this section we present Tables from \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.
 
@@
 
-\subsection{Comparing BRKGA+LS with SHADE+LS}
-
-In Tables \ref{tab:results10} to \ref{tab:results20} the ARI columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
+\subsection{Comparing BRKGA$_{CC}$ with SHADE$_{CC}$}
+
+In Tables from \ref{tab:results10} to \ref{tab:results20} the ARI columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
 
@@
 			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA+LS} &&  \multicolumn{3}{c}{SHADE+LS} \\
+			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
 			\cline{2-4} \cline{6-8}
@@
 			\hline
+			Mean & 0.161 & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
+			\hline
 
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE+LS and BRKGA+LS}
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$}
 	\label{tab:results10}
@@
 Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-We can note that SHADE+LS represents a consistent improvement over BRKGA+LS in both Unsat and Time, while in terms of ARI both methods seem similar.
+We can note that SHADE$_{CC}$ represents a consistent improvement over BRKGA$_{CC}$ in both Unsat and Time, while in terms of ARI both methods are similar.
 This is because the penalty applied to the fitness function of both methods has little presence when operating with a small constraint set, and since they are population-based and heuristic-guided methods, they have similar exploration and exploitation capabilities.
-
- it is at ARI where we begin to see a meaningful difference between the two methods.
 
@@
 			\hline
+			Mean & 0.483 & 9.456 & 1784.931 && 0.507 & 6.997 & 1751.368 \\
+			\hline
 
@@
 
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE+LS and BRKGA+LS}
+	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$}
 	\label{tab:results15}
@@
 We observe that both methods already report ARI results above 0.7, which in an ARI ranking is considered to be an acceptable result.
-We can even observe that the SHADE+LS method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA+LS that only obtains similar results in 2 datasets.
+We can even observe that the SHADE$_{CC}$ method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA$_{CC}$ that only obtains similar results in 2 datasets.
 
@@
 			\hline
+			Mean & 0.602 & 8.086 & 1895.671 && 0.647 & 5.443 & 1880.921 \\
+			\hline
 
@@
 
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE+LS and BRKGA+LS}
+	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$}
 	\label{tab:results20}
@@
 It is in Table \ref{tab:results20}, which gathers results obtained for the constraint set $CS_{20}$, that we observe major differences between the two methods.
-We find that SHADE+LS is capable of producing clusters identical to the original ones for 4 datasets, against the single dataset for which BRKGA+LS achieves it.
+We find that SHADE$_{CC}$ is capable of producing clusters identical to the original ones for 4 datasets, regardless the single dataset for which BRKGA$_{CC}$ achieves it.
 Let us remember that an ARI of 1 represents a perfect match between the two partitions given as arguments.
-This improvement in ARI is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, SHADE+LS gets the advantage.
+This improvement in ARI is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, SHADE$_{CC}$ outperforms BRKGA$_{CC}$.
 Conclusions on execution time remain the same.
@@
 
-\subsection{Comparing SHADE+LS with the state-of-the-art} \label{sec:SHADEvsSOTA}
-
-In this section we present a brief comparison between the proposed SHADE+LS method and some of the non-heuristic state-of-the-art methods.
-Tables \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
-This time we will only compare the results obtained for ARI.
+\subsection{Comparing SHADE$_{CC}$ with the state-of-the-art} \label{sec:SHADEvsSOTA}
+
+In this section we present a brief comparison between the proposed SHADE$_{CC}$ method and some of the state-of-the-art methods.
+Tables from \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
+To shorten the presented results this time we will only compare the ARI
 
@@
 			\hline
-			Dataset & SHADE+LS & COPKM & LCVQE & RDPM & TVClust & CECM \\
+			Dataset & SHADE$_{CC}$ & COPKM & LCVQE & RDPM & TVClust & CECM \\
 			\hline
@@
 			\hline
+			Mean  & 0.156 & -0.51 & 0.215 & 0.211 & 0.13 & -0.226 \\
+			\hline
 		\end{tabular}}
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE+LS and the state-of-the-art}
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art}
 	\label{tab:results10SOTA}
@@
 Table \ref{tab:results10SOTA} presents the results obtained for the $CS_{10}$ constraint set.
-LCVQE, RDPM and TVClust methods stand out, achieving results similar or better than SHADE+LS with a low number of constraints.
+LCVQE, RDPM and TVClust methods stand out, achieving results as competitive as SHADE$_{CC}$ with a low number of constraints.
 
@@
 			\hline
+			Mean  & 0.507 & 0.01 & 0.213 & 0.229 & 0.253 & -0.238 \\
+			\hline
 		\end{tabular}}
@@
 
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE+LS and the state-of-the-art}
+	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and the state-of-the-art}
 	\label{tab:results15SOTA}
@@
 
-In Table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE+LS.
-On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by SHADE+LS, with the exception of COPKM, which is able to reach the optimum for some of the cases in which it finds a solution.
+In Table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE$_{CC}$.
+On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by SHADE$_{CC}$, with the exception of COPKM, which is able to reach the optimum for some of the cases in which it finds a solution.
 
@@
 			\hline
+			Mean  & 0.647 & 0.224 & 0.212 & 0.28 & 0.319 & -0.228 \\
+			\hline
 		\end{tabular}}
@@
 
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE+LS and the state-of-the-art}
+	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and the state-of-the-art}
 	\label{tab:results20SOTA}
@@
 Table \ref{tab:results20SOTA} presents the results obtained for $CS_{20}$.
-For SHADE+LS we continue to observe its progression in accuracy.
+For SHADE$_{CC}$ we continue to observe its progression in accuracy.
 As far as the state-of-the-art methods are concerned, we find a general improvement in ARI, except for CECM.
-Even with this, it is SHADE+LS and COPKM that seem to obtain the highest quality results, with the difference that SHADE+LS is able to produce results in all cases.
+Even with this, it is SHADE$_{CC}$ and COPKM that seem to obtain the highest quality results, with the difference that SHADE$_{CC}$ is able to produce results in all cases.
 
@@
 
-\section{Analysis of Results} \label{sec:analisis}
+\section{Statistical Analysis of Results} \label{sec:analisis}
 
 With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
-This way we can statistically determine whether SHADE+LS represents a significant improvement over previous proposals.
-
-In order to carry out this study we will use Bayesian statistical tests, instead of the classic null hypothesis statistical tests (NHST).
+This way we can statistically determine whether SHADE$_{CC}$ represents a significant improvement over previous proposals.
+
+In order to carry out this study we will use Bayesian statistical tests, instead of the classic Hull Hypothesis Statistical Tests (NHST).
 In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
@@
 To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
-On the other hand, the hypothesis that the differences between results obtained with different methods are exactly zero is almost always false!
-In the cases where NHST allows us to reject the null hypothesis we conclude that the occurrence of this is unlikely, but this is known even before carrying out experiments.
-
 Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
@@
 
-In most cases the researcher would also like to have information about the magnitude of the effects and the uncertainty of its estimate, but NHST do not provide them.
+Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate.
 As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
 
-Furthermore, and this is a situation that all the researchers have faced, NHST do not provide any information about the null hypothesis!
+Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
 That is: What can we conclude when NHST do not reject $H_0$?
-We cannot infer anything since NHST cannot provide evidence in its favor.
+We can not infer anything since NHST can not provide evidence in its favor.
 
 Finally, there are two other problems that researchers face when performing NHST.
-The choice of the significance level $\alpha$ is critical to the test results, and there is no principle to establish it objectively.
-Additionally, it is worth remembering that, in order to carry out a NHST, there is a previous need to fix and formalize the intentions of the sampling of the results; however, these intentions are commonly fixed a posteriori, which can lead to a misreading of the results.
+The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results.
+The second one is the need to previously formalize the intentions of the sampling of the results, which are usually fixed a posteriori; this could lead to a misreading of said results.
 
@@
 
-From now on, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA+LS as sample $A$, and to the results obtained with SHADE+LS as sample $B$.
+From now on, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA$_{CC}$ as sample $A$, and to the results obtained with SHADE$_{CC}$ as sample $B$.
 
@@
 	\end{subfigure}
-	\caption{Heat diagrams for the comparison with BRKGA+LS}
+	\caption{Heat diagrams for the comparison with BRKGA$_{CC}$}
 \end{figure}
@@
 It also indicates that in some cases there are no significant differences between $A$ and $B$, since we also find triplets in the rope area.
-Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE+LS performs better than BRKGA+LS.
+Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE$_{CC}$ performs better than BRKGA$_{CC}$.
 
@@
 In both, all triplets are represented in the lower right third of the triangle, indicating that the Bayesian sign test assigns a high probability to $A - B > 0$.
-With this, and bearing in mind that both measures must be minimized, we conclude that SHADE+LS represents a solid improvement over BRKGA+LS as far as this measures are concerned.
-
-Regarding the comparison of SHADE+LS with the state-of-the-art, we will keep the results obtained by SHADE+LS as $B$, and note the results obtained by the state-of-the-art methods as $A$.
-
-This way, we apply the process described earlier in this section to obtain the heatmaps corresponding to the comparison of SHADE+LS with each of the 5 state-of-the-art methods.
+With this, and bearing in mind that both measures must be minimized, we conclude that SHADE$_{CC}$ represents a solid improvement over BRKGA$_{CC}$ as far as this measures are concerned.
+
+Regarding the comparison of SHADE$_{CC}$ with the state-of-the-art, we will keep the results obtained by SHADE$_{CC}$ as $B$, and note the results obtained by the state-of-the-art methods as $A$.
+
+This way, we apply the process described earlier in this section to obtain the heatmaps corresponding to the comparison of SHADE$_{CC}$ with each of the 5 state-of-the-art methods.
 This time the only measure we compare is ARI.
@@
 		\includegraphics[width=1\linewidth]{Figures/AriCOPKM.pdf}
-		\caption{SHADE+LS vs COPKM}
+		\caption{SHADE$_{CC}$ vs COPKM}
 		\label{fig:bayesCOPKM}
@@
 		\includegraphics[width=1\linewidth]{Figures/AriLCVQE.pdf}
-		\caption{SHADE+LS vs LCVQE}
+		\caption{SHADE$_{CC}$ vs LCVQE}
 		\label{fig:bayesLCVQE}
@@
 		\includegraphics[width=1\linewidth]{Figures/AriRDPM.pdf}
-		\caption{SHADE+LS vs RDPM}
+		\caption{SHADE$_{CC}$ vs RDPM}
 		\label{fig:bayesRDPM}
@@
 		\includegraphics[width=1\linewidth]{Figures/AriTVClust.pdf}
-		\caption{SHADE+LS vs TVClust}
+		\caption{SHADE$_{CC}$ vs TVClust}
 		\label{fig:bayesTVClust}
@@
 		\includegraphics[width=1\linewidth]{Figures/AriCECM.pdf}
-		\caption{SHADE+LS vs CECM}
+		\caption{SHADE$_{CC}$ vs CECM}
 		\label{fig:bayesCECM}
@@
 In all of them most of the triplets are represented in the lower middle region of the diagram, with a slight shift of the point cloud to the left.
-That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---, but the advantage is for one or the other depending on the case, being SHADE+LS the one that gets it in most of them.
-However, this does not apply in the case of CECM (\ref{fig:bayesCECM}), where the advantage is indisputably for SHADE+LS.
+That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---, but the advantage is for one or the other depending on the case, being SHADE$_{CC}$ the one that gets it in most of them.
+However, this does not apply in the case of CECM (\ref{fig:bayesCECM}), where the advantage is indisputably for SHADE$_{CC}$.
 
@@
 We obtained results on 25 different datasets and 3 different constraint sets for each one, applying all the resolution methods named in this paper.
-Thanks to the Bayesian statistical tests, which offer us the advantageous option of producing a visualization of differences between results obtained by different methods, we were able to objectively prove that the SHADE+LS approach is significantly better than previous population-based proposals.
-As far as the state-of-the-art is concerned, the SHADE+LS approach has proven to be better or equivalent when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
+Thanks to the Bayesian statistical tests, which offer us the advantageous option of producing a visualization of differences between results obtained by different methods, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous population-based proposals.
+As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be better or equivalent when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
 

### 2019-03-17 Papers/SHADE/paperSHADE.tex
Nuevas correcciones en SHADE

@@
 Clustering has always been a powerful tool in knowledge discovery.
-Traditionally unsupervised, it has received attention when adding new types of information to it has proven to provide better results than completely unsupervised procedures, leading to a new kind of semi-supervised learning.
-This paper proposes a variant of the successful SHADE method to solve the constrained clustering problem.
-This problem consists of a generalization of the traditional clustering to consider a new type of information, which is given in the form of constraints.
-This paper focuses on instance-level must-link and cannot-link constraints.
-We will compare the results obtained on 25 datasets, in conjunction with various constraint sets, by this new proposal with a previous nature-inspired approach---BRKGA---as well as with some of the state-of-the-art algorithms, supporting our conclusions with the aid of Bayesian statistical tests.
+Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning.
+This paper proposes a variant of SHADE, a successful approach to the constrained clustering problem, which is a generalization of traditional clustering that considers additional information in the form of constraints.
+In particular, this work focuses on instance-level must-link and cannot-link constraints.
+We will compare the results obtained by this proposal to those obtained by a previous nature-inspired technique---BRKGA---and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 
-Constrained clustering is a semi-supervised learning method.
+Constrained clustering is an SSL learning method.
 Its goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
@@
 In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
-Sections \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis.
+Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis.
 Finally, in Section \ref{sec:conclusiones} we discuss conclusions.
@@
 
-	\item Must-link constraints $C_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
+	\item Must-link constraints $C_=(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
 
@@
 Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem.
-The feasibility problem for non-hierarchical constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
+The feasibility problem for non-hierarchical instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
 
@@
 		\end{tabular}%}
-	\caption{Feasibility problem complexity \cite{davidson2005clustering}}
+	\caption{Feasibility problem complexity \cite{davidson2005clustering}.}
 	\label{tab:feasibility}
@@
 Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why heuristics-guided and population-based algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
+That is the reason why nature-inspired algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
 
@@
 
-In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), an \textcolor{red}{adaptation} of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods, for constrained clustering.
+In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods, for constrained clustering.
 The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
@@
 Two Views Clustering (TVClust) and Relation Dirichlet Process - Means (RDPM) were proposed in \cite{khashabi2015clustering}.
-TVClust is able to incorporate the constraints into the clustering problem by making a relaxed interpretation of them.
+TVClust is able to incorporate the constraints into the clustering problem by making a soft interpretation of them.
 The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
 Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
-This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($K$) does not need to be specified.
+This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($k$) does not need to be specified.
 
@@
 The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to this kind of problems.
-It uses random numbers to represent solutions to the problem (individuals), which are used as keys to obtain the actual solutions.
+It uses vectors of random numbers to represent solutions to the problem (individuals), which are used as keys to obtain the actual solutions.
 Random-key vectors are sampled from the space $[0,1]^{v}$, which is used as a surrogate space for the literal solution space.
@@
 	\end{tabular}}
-	\caption{Random-key decodification example \cite{de2017comparison}}
+	\caption{Random-key decodification example \cite{de2017comparison}.}
 	\label{tab:decodingrk}
@@
 Since SHADE is an evolution-based algorithm, it shares elements with BRKGA (Section \ref{sec:brkga}), so we will use a similar notation to define its elements.
-We will note the population as $P$, formed by $p_{size}$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
+We will note the population as $P_G$, formed by $p_{size}$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
 
@@
 
-After generating the mutant vector $m_{[i,G]}$, it is combined with the parent $p_{[i,G]}$ vector by means of a crossover operator; the result is the trial vector.
+After generating the mutant vector $m_{[i,G]}$, it is combined with the parent $p_{[i,G]}$ vector by means of a crossover operator; the result is the trial vector $t_{[i,G]} = \{t_{[i,G,1]}, \cdots, t_{[i,G,v]}\}$.
 SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.
@@
 \begin{equation}
-t_{[j,i,G]} = \left\{ \begin{array}{lc}
-m_{[j,i,G]} &   if \;\; \text{rand}[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
-p_{[j,i,G]} &  \text{otherwise}
+t_{[i,G,j]} = \left\{ \begin{array}{lc}
+m_{[i,G,j]} &   if \;\; \text{rand}[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
+p_{[i,G,j]} &  \text{otherwise}
 \end{array}
@@
 The SHADE selection operator determines the survivors individuals for the next generation.
-It compares each parent $x_{[i,G]}$ with the trial vector $_{[j,i,G]}$, generated based on it.
+It compares each parent $x_{[i,G]}$ with the trial vector $_{[i,G,j]}$, generated based on it.
 The best one of these two individuals will survive for the next generation.
@@
 	\end{tabular}}
-	\caption{Historical memory $M_{CR}$, $M_{F}$ used by SHADE \cite{tanabe2013success}}
+	\caption{Historical memory $M_{CR}$, $M_{F}$ used by SHADE \cite{tanabe2013success}.}
 	\label{tab:SHADEmemory}
@@
 We will call this new approach $SHADE_{CC}$.
-To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of $n$ random-keys.
+To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of $n$ random-keys ($v = n$).
 Additionally, we will also make use of the initialization method and the decoder found in Section \ref{sec:AdaptationofBRKGA}.
-However, we propose a new fitness function to evaluate the quality of the individuals of the population, Equation \eqref{eq14}, whose element $z_i$ is obtained as in Equation \eqref{eq2}.
-
 However, we propose a new fitness function to evaluate the quality of the individuals of the population, Equation \eqref{eq14}, whose element $z_i$ is obtained as in Equation \eqref{eq2}.
@@
 
-	\caption{Expression and value of fitness functions over three partitions. ($\mu = 10$)}
+	\caption{Expression and value of fitness functions over three partitions ($\mu = 10$).}
 	\label{tab:fitnessfunctions}
@@
 	}
-	\caption{Modified SHADE}\label{alg:SHADE}
+	\caption{SHADE$_{CC}$}\label{alg:SHADE}
 \end{algorithm}
@@
 	\end{tabular}%}
-	\caption{Summary of datasets used for the experiments}
+	\caption{Summary of datasets used for the experiments.}
 	\label{tab:datasets}
@@
 Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
-They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Rand Index (see Section \ref{sec:EvalMet}).
+They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).
 
@@
 
-The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
+The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
 Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.
@@
 
-	\caption{Number of constraints used in experiments}
+	\caption{Number of constraints used in experiments.}
 	\label{tab:constraints}
@@
 
-The Adjusted Rand Index is a corrected-for-chance version of the Rand Index.
+The ARI is a corrected-for-chance version of the Rand Index.
 This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline.
-The Adjusted Rand Index is computed as in Equation \eqref{eq16}.
+The ARI is computed as in Equation \eqref{eq16}.
 
@@
 To summarize, the higher the $ARI$, the greater the degree of similarity between $C_1$ and $C_2$.
-For more details on Adjusted Rand Index see \cite{hubert1985comparing}.
+For more details on ARI see \cite{hubert1985comparing}.
 
 Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper.
-To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth.
+To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth labels.
 
@@
 			$p_{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
-			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Known for each dataset (Table \ref{tab:datasets}). It has the same value for both methods} \\
+			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets}).} \\
 			\hline
@@
 
-To compare with the state-of-the-art methods mentioned in section \ref{sec:BackSOTA} we will use the parameters recommended by the authors, setting the maximum limit of iterations as 300 in all cases.
+To compare with the state-of-the-art methods mentioned in section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} c}
+			\hline
+			Common Parameters & Meaning & Value \\
+			\hline
+			\texttt{max\_iter} & Maximum number of iterations & 300 \\
+			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}). \\
+			\hline
+			\hline
+			Specific Parameters & \multicolumn{2}{l}{Name and Value} \\
+			\hline
+			COPKM & \multicolumn{2}{l}{\texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''}} \\
+			CECM & \multicolumn{2}{l}{$\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''}} \\
+			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
+			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
+			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Parameters setup used for the state-of-the-art algorithms.}
+	\label{tab:paramsSOTA}
+\end{table}
 
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$}
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
 	\label{tab:results10}
@@
 
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$}
+	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
 	\label{tab:results15}
@@
 
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$}
+	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
 	\label{tab:results20}
@@
 We find that SHADE$_{CC}$ is capable of producing clusters identical to the original ones for 4 datasets, regardless the single dataset for which BRKGA$_{CC}$ achieves it.
-Let us remember that an ARI of 1 represents a perfect match between the two partitions given as arguments.
+Let us remember that an ARI of 1 represents a perfect match between the two partitions given to compute it.
 This improvement in ARI is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, SHADE$_{CC}$ outperforms BRKGA$_{CC}$.
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art}
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
 	\label{tab:results10SOTA}
@@
 
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and the state-of-the-art}
+	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
 	\label{tab:results15SOTA}
@@
 
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and the state-of-the-art}
+	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
 	\label{tab:results20SOTA}
@@
 	\end{subfigure}
-	\caption{Heat diagrams for the comparison with BRKGA$_{CC}$}
+	\caption{Heat diagrams for the comparison with BRKGA$_{CC}$.}
 \end{figure}
@@
 	\end{subfigure}
-	\caption{Heat diagrams for the comparison with the state-of-the-art}
+	\caption{Heat diagrams for the comparison with the state-of-the-art.}
 \end{figure}
@@
 
+\clearpage
+
 \section{Conclusions} \label{sec:conclusiones}
 
-In this paper we proposed an adaptation of SHADE to solve the constrained clustering problem.
-We took concepts such as the random-key and the encoder/decoder from BRKGA to transform constrained clustering into a real-domain optimization problem.
+In this paper we proposed SHADE$_{CC}$, a new variant of the SHADE algorithm to solve the constrained clustering problem.
+We took concepts such as the random-key and the encoder/decoder from RKGA to transform constrained clustering into a real-domain optimization problem.
 We also proposed a new fitness function that has proven to be more adaptable and less restrictive.
 
-The SHADE algorithm is one of the most successful methods to approximately solve difficult problems, and constrained clustering does not appear to be an exception.
-We obtained results on 25 different datasets and 3 different constraint sets for each one, applying all the resolution methods named in this paper.
-Thanks to the Bayesian statistical tests, which offer us the advantageous option of producing a visualization of differences between results obtained by different methods, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous population-based proposals.
-As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be better or equivalent when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
+The SHADE algorithm is one of the most successful methods to approximately solve difficult problems, and constrained clustering is not an exception.
+We obtained results on 25 different datasets and 3 different constraint sets for each one, applying all the algorithms named in this paper.
+Thanks to the Bayesian statistical tests, which offer us the advantageous option of producing a visualization of differences between results obtained by different methods, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous evolution-based proposals.
+As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be equivalent or better when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
 

### 2019-03-23 Papers/SHADE/paperSHADE.tex
Resultados de DILS completos

@@
 The Biased Random-key Genetic Algorithm (BRKGA) was first proposed in \cite{gonccalves2011biased} as a variant of the RKGA.
-In the BRKGA optimization process, first, the population $P$ of $p_{size}$ vectors of random-keys is initialized; this will be the initial generation.
+In the BRKGA optimization process, first, the population $P$ of $p_\text{size}$ vectors of random-keys is initialized; this will be the initial generation.
 Each random-key vector $p_i$ has $v$ random-keys in it.
@@
 To introduce diversity, a number $p_m$ of new random-keys vectors are also included in the next generation.
-The remaining individuals of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents.
+The remaining individuals of the new generation ($p_\text{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents.
 
@@
 \begin{equation}
-f_i = z_i + \overbrace{(\mu * n * {infeasibility}_i)}^{penalty},
+f_i = z_i + \overbrace{(\mu * n * \text{infeasibility}_i)}^\text{penalty},
 \label{eq1}
@@
 
-\noindent where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.
+\noindent where $\mu$ is a high value, $\text{infeasibility}_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.
 
@@
 Since SHADE is an evolution-based algorithm, it shares elements with BRKGA (Section \ref{sec:brkga}), so we will use a similar notation to define its elements.
-We will note the population as $P_G$, formed by $p_{size}$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
+We will note the population as $P_G$, formed by $p_\text{size}$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
 
@@
 \begin{equation}
-m_{[i,G]} = p_{[i,G]} + F_i * (p_{[pbest, G]} - p_{[i,G]}) + F_i * (p_{[r1, G]} - p_{[r2,G]}),
+m_{[i,G]} = p_{[i,G]} + F_i * (p_{[p\text{best}, G]} - p_{[i,G]}) + F_i * (p_{[r1, G]} - p_{[r2,G]}),
 \label{eq3}
@@
 \noindent where $m_{[i,G]}$ is the mutant vector, which is generated with an individual $p_{[i,G]}$ from the population serving as a starting point.
-Indices $r1$ and $r2$ are random values in the range $[0,p_{size}]$, different from $i$ and from each other.
-The individual $p_{[pbest, G]}$ is randomly selected from among the best $p_{size} \times pbest\;|\;pbest\in [0,1]$ individuals in the population.
-This way, the parameter $pbest$ controls the greediness of the mutation strategy.
+Indices $r1$ and $r2$ are random values in the range $[0,p_\text{size}]$, different from $i$ and from each other.
+The individual $p_{[p\text{best}, G]}$ is randomly selected from among the best $p_\text{size} \times p\text{best}\;|\;p\text{best}\in [0,1]$ individuals in the population.
+This way, the parameter $p\text{best}$ controls the greediness of the mutation strategy.
 The parameter $F_i$ defines the magnitude of the operator.
@@
 t_{[i,G,j]} = \left\{ \begin{array}{lc}
-m_{[i,G,j]} &   if \;\; \text{rand}[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
+m_{[i,G,j]} &   \text{if} \;\; \text{rand}[0,1) \le CR_i \;\; \text{or} \;\;j = j_{rand} \\
 p_{[i,G,j]} &  \text{otherwise}
@@
 
-\noindent where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,v]$, and $CR \in [0,1]$ is the crossover ratio.
+\noindent where $\text{rand}[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_\text{rand}$ is a random integer selected from the range $[1,v]$, and $CR \in [0,1]$ is the crossover ratio.
 
 The SHADE selection operator determines the survivors individuals for the next generation.
-It compares each parent $x_{[i,G]}$ with the trial vector $_{[i,G,j]}$, generated based on it.
+It compares each parent $x_{[i,G]}$ with the trial vector $t_{[i,G,j]}$, generated based on it.
 The best one of these two individuals will survive for the next generation.
@@
 p_{[i,G + 1]} = \left\{ \begin{array}{lc}
-t_{[i,G]} &   if \;\; f(t_{[i,G]}) \le f(p_{[i,G]}) \\
+t_{[i,G]} &   \text{if} \;\; f(t_{[i,G]}) \le f(p_{[i,G]}) \\
 p_{[i,G]} &  \text{otherwise}
@@
 
-The parameters $pbest$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process.
+The parameters $p\text{best}$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process.
 These are the parameters that SHADE adaptively optimizes using the mentioned history record.
@@
 
-\noindent where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$.
+\noindent where $\text{randn}(\mu, \sigma^2)$ and $\text{randc}(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$.
 When a value for $CR_i$ outside of the range $[0,1]$ is generated, it is replaced by the corresponding limit value.
@@
 M_{[CR,h,G+1]} = \left\{ \begin{array}{lc}
-\text{mean}_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
-M_{[CR,h,G]} &  otherwise
+\text{mean}_{WA} (S_{CR}) & \text{if} \;\; S_{CR} \neq \emptyset \\
+M_{[CR,h,G]} &  \text{otherwise}
 \end{array}
@@
 M_{[F,h,G+1]} = \left\{ \begin{array}{lc}
-\text{mean}_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
-M_{[F,h,G]} &  otherwise
+\text{mean}_{WL} (S_{F}) & \text{if} \;\; S_{F} \neq \emptyset \\
+M_{[F,h,G]} &  \text{otherwise}
 \end{array}
@@
 
-In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the weighted mean, which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed in \cite{peng2009multi} to prevent $CR$ from converging to small values.
+In Equation \ref{eq8}, the term $\text{mean}_{WA} (S_{CR})$ is the weighted mean, which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed in \cite{peng2009multi} to prevent $CR$ from converging to small values.
 
@@
 
-In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12} (proposed in \cite{tanabe2013success}).
+In Equation \ref{eq9} the term $\text{mean}_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12} (proposed in \cite{tanabe2013success}).
 
@@
 
-Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the greediness of the mutation strategy, is not included in the adaptive optimization process.
+Unlike parameters $S_{CR}$ and $S_F$, the parameter $p\text{best}$ from Equation \eqref{eq3}, used to set the greediness of the mutation strategy, is not included in the adaptive optimization process.
 However, it is not static: it is calculated for every individual $p_i$ of the population following Equation \eqref{eq13}.
@@
 \begin{equation}
-pbest_i = \text{randn}[2/N, 0.2],
+p\text{best}_i = \text{randn}[2/N, 0.2],
 \label{eq13}
@@
 \begin{equation}
-f_i = z_i * (infeasibility + 1)
+f_i = z_i * (\text{infeasibility}_i + 1)
 \label{eq14}
@@
 The dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
-Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($infeasibility = 0$), whereas $C_2$ does violate it ($infeasibility = 1$).
+Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($\text{infeasibility} = 0$), whereas $C_2$ does violate it ($\text{infeasibility} = 1$).
 However, $C_2$ shows a possibility, for later iterations, of adding both instances to their nearest cluster while also satisfying the ML constraint.
@@
 			$C_1$ & $z_1$ & $z_1$ && $0.540$  & $0.540$  \\
-			$C_2$ & $z_2 + (\mu * N * infs)$ & $z_2 * (infs + 1)$  && $0.503 + 1000 = 1000.503$ &  $0.503 * 2 = 1.005$ \\
+			$C_2$ & $z_2 + (\mu * N * \text{infsblt})$ & $z_2 * (\text{infsblt} + 1)$  && $0.503 + 1000 = 1000.503$ &  $0.503 * 2 = 1.005$ \\
 			\hline
@@
 \begin{equation}
-Rand(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
+\text{Rand}(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
 \label{eq15}
@@
 \begin{equation}
-ARI(C_1, C_2) = \frac{Rand(C_1, C_2) - ExpectedIndex}{MaximumIndex - ExpectedIndex},
+\text{ARI}(C_1, C_2) = \frac{\text{Rand}(C_1, C_2) - \text{Expected Index}}{\text{Maximum Index} - \text{Expected Index}},
 \label{eq16}
@@
 
-\noindent where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
-It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that an $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing with random partitions.
-To summarize, the higher the $ARI$, the greater the degree of similarity between $C_1$ and $C_2$.
+\noindent where Maximum Index is expected to be 1 and Expected Index is the already mentioned expected degree of similarity with a random model.
+It is easy to see that $\text{ARI}(C_1, C_2) \in [-1,1]$, such that an ARI value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $\text{Rand}(C_1, C_2)$ is less than expected when comparing with random partitions.
+To summarize, the higher the ARI, the greater the degree of similarity between $C_1$ and $C_2$.
 For more details on ARI see \cite{hubert1985comparing}.
@@
 			\hline
-			$p_{size}$ & Population size & 100 & 100 \\
-			$Evals$ & Fitness function evaluations & 300000 & 300000 \\
-			$p_e$ & Size of the elite set in population & $0.2 * p_{size}$ & $0.25 * p_{size}$ \\
-			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_{size}$ & - \\
-			$p_{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
+			$p_\text{size}$ & Population size & 100 & 100 \\
+			Evals & Fitness function evaluations & 300000 & 300000 \\
+			$p_e$ & Size of the elite set in population & $0.2 * p_\text{size}$ & $0.25 * p_\text{size}$ \\
+			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_\text{size}$ & - \\
+			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
 			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets}).} \\
@@
 			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
-			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
+			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
 			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
@@
 
-It should be noted that the study presented in this document aims to compare average performance, not peak performance.
-The parameter optimization process required for each algorithm and dataset falls outside the scope of this study.
+Parameter values have been assigned following the guidelines of the original creators of the different proposals.
+Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible.
+Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible.
+Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
 
@@
 Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
-That is: What can we conclude when NHST do not reject $H_0$?
+That is: What can we conclude when NHST do not reject the null hypothesis?
 We can not infer anything since NHST can not provide evidence in its favor.
@@
 To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
-In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_{min}, r_{max}]$, so that $P(A \approx B) = P(\rho \in rope)$.
+In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$.
 Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
-$$[P(\rho < r_{min}) = P(A - B < 0),\;\; P(\rho \in rope),\;\; P(\rho > r_{max}) = P(A - B > 0)]$$
+$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$
 

### 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman10/Friedman10.tex
Resultados de DILS completos

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage{lscape}
\title{Output tables for 1xN statistical comparisons.}
\author{}
\date{\today}
\begin{document}
\pagestyle{empty}
\maketitle
\thispagestyle{empty}

\section{Average rankings of Friedman test}

Average ranks obtained by each method in the Friedman test.

\begin{table}[!htp]
\centering
\begin{tabular}{|c|c|}\hline
Algorithm&Ranking\\\hline
"SHADE"&3.04\\"COPKM"&4.6\\"LCVQE"&2.7\\"RDPM"&2.78\\"TVClust"&3.22\\"CEKM"&4.66\\\hline\end{tabular}
\caption{Average Rankings of the algorithms (Friedman)}
\end{table}

Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 28.6. \newline P-value computed by Friedman Test: 0.000028.\newline

\newpage

\section{Post hoc comparison (Friedman)}

P-values obtained in by applying post hoc methods over the results of Friedman procedure.

\begin{table}[!htp]
\centering\footnotesize
\begin{tabular}{cccccccc}
$i$&algorithm&$z=(R_0 - R_i)/SE$&$p$&Holm &Holland&Finner&Li\\
\hline5&"CEKM"&3.704052&0.000212&0.01&0.010206&0.010206&0.006325\\4&"COPKM"&3.590662&0.00033&0.0125&0.012741&0.020308&0.006325\\3&"TVClust"&0.982708&0.325751&0.016667&0.016952&0.030307&0.006325\\2&"SHADE"&0.64254&0.520523&0.025&0.025321&0.040204&0.006325\\1&"RDPM"&0.151186&0.879829&0.05&0.05&0.05&0.05\\\hline
\end{tabular}
\caption{Post Hoc comparison Table for $\alpha=0.05$ (FRIEDMAN)}
\end{table}Holm's procedure rejects those hypotheses that have an unadjusted p-value $\le0.016667$.

Holland's procedure rejects those hypotheses that have an unadjusted p-value $\le0.016952$.

Finner's procedure rejects those hypotheses that have an unadjusted p-value $\le0.030307$.

Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.006325$.

\newpage

\section{Adjusted P-Values (Friedman)}

Adjusted P-values obtained through the application of the post hoc methods (Friedman).

\begin{table}[!htp]
\centering\small
\begin{tabular}{cccc}
i&algorithm&unadjusted $p$&$p_{Holm}$\\
\hline1&"CEKM"&0.000212&0.001061\\2&"COPKM"&0.00033&0.001319\\3&"TVClust"&0.325751&0.977254\\4&"SHADE"&0.520523&1.041046\\5&"RDPM"&0.879829&1.041046\\\hline
\end{tabular}
\caption{Adjusted $p$-values (FRIEDMAN) (I)}
\end{table}
\begin{table}[!htp]
\centering\small
\begin{tabular}{cccccc}
i&algorithm&unadjusted $p$&$p_{Holland}$&$p_{Finner}$&$p_{Li}$\\
\hline1&"CEKM"&0.000212&0.00106&0.00106&0.001763\\2&"COPKM"&0.00033&0.001319&0.00106&0.002737\\3&"TVClust"&0.325751&0.693479&0.481558&0.730512\\4&"SHADE"&0.520523&0.770102&0.601012&0.812436\\5&"RDPM"&0.879829&0.879829&0.879829&0.879829\\\hline
\end{tabular}
\caption{Adjusted $p$-values (FRIEDMAN) (II)}
\end{table}

\newpage
\end{document}

### 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman15/Friedman15.tex
Resultados de DILS completos

--- 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman10/Friedman10.tex
@@
 Algorithm&Ranking\\\hline
-"SHADE"&3.04\\"COPKM"&4.6\\"LCVQE"&2.7\\"RDPM"&2.78\\"TVClust"&3.22\\"CEKM"&4.66\\\hline\end{tabular}
+"SHADE"&2.92\\"COPKM"&3.24\\"LCVQE"&3.24\\"RDPM"&3.24\\"TVClust"&3.42\\"CEKM"&4.94\\\hline\end{tabular}
 \caption{Average Rankings of the algorithms (Friedman)}
@@
 
-Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 28.6. \newline P-value computed by Friedman Test: 0.000028.\newline
+Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 18.708571. \newline P-value computed by Friedman Test: 0.002178.\newline
 
@@
 $i$&algorithm&$z=(R_0 - R_i)/SE$&$p$&Holm &Holland&Finner&Li\\
-\hline5&"CEKM"&3.704052&0.000212&0.01&0.010206&0.010206&0.006325\\4&"COPKM"&3.590662&0.00033&0.0125&0.012741&0.020308&0.006325\\3&"TVClust"&0.982708&0.325751&0.016667&0.016952&0.030307&0.006325\\2&"SHADE"&0.64254&0.520523&0.025&0.025321&0.040204&0.006325\\1&"RDPM"&0.151186&0.879829&0.05&0.05&0.05&0.05\\\hline
+\hline5&"CEKM"&3.817441&0.000135&0.01&0.010206&0.010206&0.023929\\4&"TVClust"&0.944911&0.344704&0.0125&0.012741&0.020308&0.023929\\3&"LCVQE"&0.604743&0.54535&0.016667&0.016952&0.030307&0.023929\\2&"RDPM"&0.604743&0.54535&0.025&0.025321&0.040204&0.023929\\1&"COPKM"&0.604743&0.54535&0.05&0.05&0.05&0.05\\\hline
 \end{tabular}
 \caption{Post Hoc comparison Table for $\alpha=0.05$ (FRIEDMAN)}
-\end{table}Holm's procedure rejects those hypotheses that have an unadjusted p-value $\le0.016667$.
+\end{table}Holm's procedure rejects those hypotheses that have an unadjusted p-value $\le0.0125$.
 
-Holland's procedure rejects those hypotheses that have an unadjusted p-value $\le0.016952$.
+Holland's procedure rejects those hypotheses that have an unadjusted p-value $\le0.012741$.
 
-Finner's procedure rejects those hypotheses that have an unadjusted p-value $\le0.030307$.
+Finner's procedure rejects those hypotheses that have an unadjusted p-value $\le0.020308$.
 
-Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.006325$.
+Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.023929$.
 
@@
 i&algorithm&unadjusted $p$&$p_{Holm}$\\
-\hline1&"CEKM"&0.000212&0.001061\\2&"COPKM"&0.00033&0.001319\\3&"TVClust"&0.325751&0.977254\\4&"SHADE"&0.520523&1.041046\\5&"RDPM"&0.879829&1.041046\\\hline
+\hline1&"CEKM"&0.000135&0.000674\\2&"TVClust"&0.344704&1.378817\\3&"LCVQE"&0.54535&1.636049\\4&"RDPM"&0.54535&1.636049\\5&"COPKM"&0.54535&1.636049\\\hline
 \end{tabular}
@@
 i&algorithm&unadjusted $p$&$p_{Holland}$&$p_{Finner}$&$p_{Li}$\\
-\hline1&"CEKM"&0.000212&0.00106&0.00106&0.001763\\2&"COPKM"&0.00033&0.001319&0.00106&0.002737\\3&"TVClust"&0.325751&0.693479&0.481558&0.730512\\4&"SHADE"&0.520523&0.770102&0.601012&0.812436\\5&"RDPM"&0.879829&0.879829&0.879829&0.879829\\\hline
+\hline1&"CEKM"&0.000135&0.000674&0.000674&0.000296\\2&"TVClust"&0.344704&0.815605&0.652389&0.431228\\3&"LCVQE"&0.54535&0.906021&0.731179&0.54535\\4&"RDPM"&0.54535&0.906021&0.731179&0.54535\\5&"COPKM"&0.54535&0.906021&0.731179&0.54535\\\hline
 \end{tabular}

### 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman20/Friedman20.tex
Resultados de DILS completos

--- 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman15/Friedman15.tex
@@
 Algorithm&Ranking\\\hline
-"SHADE"&2.92\\"COPKM"&3.24\\"LCVQE"&3.24\\"RDPM"&3.24\\"TVClust"&3.42\\"CEKM"&4.94\\\hline\end{tabular}
+"SHADE"&2.8\\"COPKM"&2.86\\"LCVQE"&3.66\\"RDPM"&3.28\\"TVClust"&3.3\\"CEKM"&5.1\\\hline\end{tabular}
 \caption{Average Rankings of the algorithms (Friedman)}
@@
 
-Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 18.708571. \newline P-value computed by Friedman Test: 0.002178.\newline
+Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 25.525714. \newline P-value computed by Friedman Test: 0.00011.\newline
 
@@
 $i$&algorithm&$z=(R_0 - R_i)/SE$&$p$&Holm &Holland&Finner&Li\\
-\hline5&"CEKM"&3.817441&0.000135&0.01&0.010206&0.010206&0.023929\\4&"TVClust"&0.944911&0.344704&0.0125&0.012741&0.020308&0.023929\\3&"LCVQE"&0.604743&0.54535&0.016667&0.016952&0.030307&0.023929\\2&"RDPM"&0.604743&0.54535&0.025&0.025321&0.040204&0.023929\\1&"COPKM"&0.604743&0.54535&0.05&0.05&0.05&0.05\\\hline
+\hline5&"CEKM"&4.346591&0.000014&0.01&0.010206&0.010206&0.004751\\4&"LCVQE"&1.625247&0.10411&0.0125&0.012741&0.020308&0.004751\\3&"TVClust"&0.944911&0.344704&0.016667&0.016952&0.030307&0.004751\\2&"RDPM"&0.907115&0.364346&0.025&0.025321&0.040204&0.004751\\1&"COPKM"&0.113389&0.909722&0.05&0.05&0.05&0.05\\\hline
 \end{tabular}
@@
 
-Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.023929$.
+Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.004751$.
 
@@
 i&algorithm&unadjusted $p$&$p_{Holm}$\\
-\hline1&"CEKM"&0.000135&0.000674\\2&"TVClust"&0.344704&1.378817\\3&"LCVQE"&0.54535&1.636049\\4&"RDPM"&0.54535&1.636049\\5&"COPKM"&0.54535&1.636049\\\hline
+\hline1&"CEKM"&0.000014&0.000069\\2&"LCVQE"&0.10411&0.41644\\3&"TVClust"&0.344704&1.034113\\4&"RDPM"&0.364346&1.034113\\5&"COPKM"&0.909722&1.034113\\\hline
 \end{tabular}
@@
 i&algorithm&unadjusted $p$&$p_{Holland}$&$p_{Finner}$&$p_{Li}$\\
-\hline1&"CEKM"&0.000135&0.000674&0.000674&0.000296\\2&"TVClust"&0.344704&0.815605&0.652389&0.431228\\3&"LCVQE"&0.54535&0.906021&0.731179&0.54535\\4&"RDPM"&0.54535&0.906021&0.731179&0.54535\\5&"COPKM"&0.54535&0.906021&0.731179&0.54535\\\hline
+\hline1&"CEKM"&0.000014&0.000069&0.000069&0.000153\\2&"LCVQE"&0.10411&0.355803&0.240309&0.535578\\3&"TVClust"&0.344704&0.718608&0.505618&0.792456\\4&"RDPM"&0.364346&0.718608&0.505618&0.801423\\5&"COPKM"&0.909722&0.909722&0.909722&0.909722\\\hline
 \end{tabular}

### 2019-03-23 Resultados/ResultadosDefinitivos/Tests/testBayesianos.r

@@ -1,14 +1,14 @@
 library(rNPBST)
 
-ARI <- read.csv("ARI.csv", dec = ".", row.names = 1)
-Unsat <- read.csv("Unsat.csv", dec = ".", row.names = 1)
-Time <- read.csv("Time.csv", dec = ".", row.names = 1)
+ARI <- read.csv("ARI.csv", dec = ",", row.names = 1)
+Unsat <- read.csv("Unsat.csv", dec = ",", row.names = 1)
+Time <- read.csv("Time.csv", dec = ",", row.names = 1)
 
 #n.samples = paramatero para el numero de puntitos
 
-ari.test = rNPBST::bayesianSign.test(ARI$BRKGA, ARI$SHADE, n.samples = 2000)
-unsat.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHAD, n.samples = 2000)
-time.test = rNPBST::bayesianSign.test(Unsat$BRKGA, Unsat$SHADE, n.samples = 2000)
+ari.test = rNPBST::bayesianSign.test(ARI$BRKGA[25:100], ARI$SHADE[25:100], n.samples = 2000)
+unsat.test = rNPBST::bayesianSign.test(Unsat$BRKGA[25:100], Unsat$SHAD[25:100], n.samples = 2000)
+time.test = rNPBST::bayesianSign.test(Unsat$BRKGA[25:100], Unsat$SHADE[25:100], n.samples = 2000)
 
 ari.test$probabilities
 unsat.test$probabilities
@@ -41,3 +41,30 @@ time.test = rNPBST::bayesian.imprecise(Unsat$SHADE, Unsat$BRKGA)
 plot(ari.test)
 plot(unsat.test)
 plot(time.test)
+
+
+ari.copkm = rNPBST::bayesianSign.test(ARI$COPKM[76:100], ARI$SHADE[76:100], n.samples = 2000)
+ari.lcvqe = rNPBST::bayesianSign.test(ARI$LCVQE[76:100], ARI$SHADE[76:100], n.samples = 2000)
+ari.rdpm = rNPBST::bayesianSign.test(ARI$RDPM[76:100], ARI$SHADE[76:100], n.samples = 2000)
+ari.tvclust = rNPBST::bayesianSign.test(ARI$TVClust[76:100], ARI$SHADE[76:100], n.samples = 2000)
[…]

### 2019-03-23 Resultados/ResultadosTFG/combine.r

@@ -1,0 +1,6 @@
+
+ARI <- read.csv("ARI_TFG.csv", dec = ".", row.names = 1)
+write.csv(ARI[1:25, ], "ARI_TFG05.csv")
+write.csv(ARI[26:50, ], "ARI_TFG10.csv")
+write.csv(ARI[51:75, ], "ARI_TFG15.csv")
+write.csv(ARI[76:100, ], "ARI_TFG20.csv")

### 2019-03-23 Resultados/ResultadosTFG/testBayesianos.r

@@ -1,0 +1,43 @@
+library(rNPBST)
+
+ARI <- read.csv("ARI_TFG.csv", dec = ",", row.names = 1)
+
+#n.samples = paramatero para el numero de puntitos
+
+ari.copkm = rNPBST::bayesianSign.test(ARI$COPKM[26:100], ARI$SHADE[26:100], n.samples = 2000)
+ari.lcvqe = rNPBST::bayesianSign.test(ARI$LCVQE[26:100], ARI$SHADE[26:100], n.samples = 2000)
+ari.rdpm = rNPBST::bayesianSign.test(ARI$RDPM[26:100], ARI$SHADE[26:100], n.samples = 2000)
+ari.tvclust = rNPBST::bayesianSign.test(ARI$TVClust[26:100], ARI$SHADE[26:100], n.samples = 2000)
+ari.cekm = rNPBST::bayesianSign.test(ARI$CEKM[26:100], ARI$SHADE[26:100], n.samples = 2000)
+
+ari.copkm$probabilities
+ari.lcvqe$probabilities
+ari.rdpm$probabilities
+ari.tvclust$probabilities
+ari.cekm$probabilities
+
+plot(ari.copkm)
+plot(ari.lcvqe)
+plot(ari.rdpm)
+plot(ari.tvclust)
+plot(ari.cekm)
+
+pdf(file = "AriCOPKM.pdf", width= 6, height = 4, useDingbats=F)
+plot(ari.copkm)
+dev.off()
+
+pdf(file = "AriLCVQE.pdf", width= 6, height = 4, useDingbats=F)
+plot(ari.lcvqe)
+dev.off()
+
+pdf(file = "AriRDPM.pdf", width= 6, height = 4, useDingbats=F)
+plot(ari.rdpm)
+dev.off()
+
+pdf(file = "AriTVClust.pdf", width= 6, height = 4, useDingbats=F)
+plot(ari.tvclust)
+dev.off()
+
+pdf(file = "AriCEKM.pdf", width= 6, height = 4, useDingbats=F)
+plot(ari.cekm)
+dev.off()

### 2019-03-23 Resultados/combine.r

@@ -17,16 +17,16 @@ tabla15 = tabla15.1 * 0.4 + tabla15.2 * 0.6
 tabla20 = tabla20.1 * 0.4 + tabla20.2 * 0.6
 
 tabla5 = tabla5[!(row.names(tabla5) %in% 
-                    c("Wine ", "Balance ", "Bupa ", "Newthyroid ", "Movement_Libras ", "Heart ", "Rand ")), ]
+                     c("Wine ", "Balance ", "Boston ", "Diabetes ", "Newthyroid ", "Heart ", "Rand ")), ]
 
 tabla10 = tabla10[!(row.names(tabla10) %in% 
-                    c("Wine ", "Balance ", "Bupa ", "Newthyroid ", "Movement_Libras ", "Heart ", "Rand ")), ]
+                      c("Wine ", "Balance ", "Boston ", "Diabetes ", "Newthyroid ", "Heart ", "Rand ")), ]
 
 tabla15 = tabla15[!(row.names(tabla15) %in% 
-                    c("Wine ", "Balance ", "Bupa ", "Newthyroid ", "Movement_Libras ", "Heart ", "Rand ")), ]
+                      c("Wine ", "Balance ", "Boston ", "Diabetes ", "Newthyroid ", "Heart ", "Rand ")), ]
 
 tabla20 = tabla20[!(row.names(tabla20) %in% 
-                    c("Wine ", "Balance ", "Bupa ", "Newthyroid ", "Movement_Libras ", "Heart ", "Rand ")), ]
+                      c("Wine ", "Balance ", "Boston ", "Diabetes ", "Newthyroid ", "Heart ", "Rand ")), ]
 
 write.table(tabla5, "tabla5.dat", row.names = F, col.names = F)
 write.table(tabla10, "tabla10.dat", row.names = F, col.names = F)

### 2019-03-23 Resultados/printTables.py

@@ -3,11 +3,11 @@ import csv
 
 
 def main():
-
-	names_array = ['iris', 'soybean', 'pima', 'boston', 'diabetes', 
-	'breast_cancer', 'ecoli', 'haberman', 'led7digit', 'monk2', 
-	'vehicle', 'zoo', 'sonar', 'ionosphere', 'wdbc', 'vowel', 
-	'appendicitis', 'saheart', 'spectfheart', 'hayesroth', 'tae', 'glass', 'spiral', 'moons', 'circles']
+	#"wine ", "balance ", "boston ", "diabetes ", "newthyroid ", "heart ", "rand "
+	names_array = ['iris', 'soybean', 'pima', 
+	'breast_cancer', 'bupa', 'ecoli', 'haberman', 'led7digit', 'monk2',
+	'vehicle', 'zoo', 'sonar', 'ionosphere', 'wdbc', 'vowel', 'movement_libras', 
+	'appendicitis', 'saheart', 'spectfheart', 'hayesroth', 'tae', 'glass', 'spiral', 'moons', 'circles', "means"]
 
 	tabla5 = np.loadtxt("tabla5.dat")
 	tabla10 = np.loadtxt("tabla10.dat")

### 2019-03-23 Resultados/select.py

@@ -9,10 +9,10 @@ def main():
 	'vehicle', 'zoo', 'sonar', 'heart', 'ionosphere', 'wdbc', 'vowel', 'movement_libras', 
 	'appendicitis', 'saheart', 'spectfheart', 'hayesroth', 'tae', 'glass', 'rand', 'spiral', 'moons', 'circles', "means"]
 
-	tabla5 = np.loadtxt("tabla5.dat")
-	tabla10 = np.loadtxt("tabla10.dat")
-	tabla15 = np.loadtxt("tabla15.dat")
-	tabla20 = np.loadtxt("tabla20.dat")
+	tabla5 = np.loadtxt("ResultadosCompletos/tabla5.dat")
+	tabla10 = np.loadtxt("ResultadosCompletos/tabla10.dat")
+	tabla15 = np.loadtxt("ResultadosCompletos/tabla15.dat")
+	tabla20 = np.loadtxt("ResultadosCompletos/tabla20.dat")
 
 	comp = np.empty((33,4), dtype = np.bool)
 	comp2 = np.empty((33,4), dtype = np.float)
@@ -26,6 +26,8 @@ def main():
 	comp2[:, 1] = tabla10[:, 0] - tabla10[:, 3]
 	comp2[:, 2] = tabla15[:, 0] - tabla15[:, 3]
 	comp2[:, 3] = tabla20[:, 0] - tabla20[:, 3]
+
+	#"Wine ", "Balance ", "Bupa ", "Newthyroid ", "Movement_Libras ", "Heart ", "Rand "
 
 	for i in range(comp.shape[0]):
 

### 2019-03-23 Software/EstimandoXi.py

@@ -25,7 +25,7 @@ def main():
 
 	max_eval = 300000
 
-	for p in range(4, 6):
+	for p in range(5, 6):
 
 		xi = -p/10.0
 

### 2019-03-23 Software/ILSNueva/ILSNueva.py

@@ -6,212 +6,203 @@ import time
 
 class ILSNueva:
 
-    def __init__(self, data, ml_const, cl_const, nb_clust, segment_percent, ls_max_neighbors, pbt_inherit, xi):
+	def __init__(self, data, ml_const, cl_const, nb_clust, segment_percent, ls_max_neighbors, pbt_inherit, xi):
 
-        self._data = data
-        self._ml = ml_const
-        self._cl = cl_const
-        self._dim = data.shape[0]
-        self._result_nb_clust = nb_clust
-        self._evals_done = 0
-        self._segment_size = int(np.ceil(segment_percent * self._dim))
-        self._max_neighbors = ls_max_neighbors
-        self._pbt_inherit = pbt_inherit
-        self._xi = xi
+		self._data = data
+		self._ml = ml_const
+		self._cl = cl_const
+		self._dim = data.shape[0]
+		self._result_nb_clust = nb_clust
+		self._evals_done = 0
+		self._segment_size = int(np.ceil(segment_percent * self._dim))
+		self._max_neighbors = ls_max_neighbors
+		self._pbt_inherit = pbt_inherit
+		self._xi = xi
 
-    def init_ils(self):
+	def init_ils(self):
 
-        self._best_solution = np.random.randint(0, self._result_nb_clust, (2, self._dim))
-        self._best_fitness = np.empty(2)
-        self._best_fitness[0] = self.get_single_fitness(self._best_solution[0, :])[0]
-        self._best_fitness[1] = self.get_single_fitness(self._best_solution[1, :])[0]
+		self._best_solution = np.random.randint(0, self._result_nb_clust, (2, self._dim))
+		self._best_fitness = np.empty(2)
+		self._best_fitness[0] = self.get_single_fitness(self._best_solution[0, :])[0]
+		self._best_fitness[1] = self.get_single_fitness(self._best_solution[1, :])[0]
 
 
-    # Funcionalidad para evaluar un unico individuo decodificado
-    def get_single_fitness(self, cromosome):
+	# Funcionalidad para evaluar un unico individuo decodificado
+	def get_single_fitness(self, cromosome):
 
-        current_clustering = cromosome
-        # Inicializamos la distancia media de las instancias de los clusters
-        total_mean_distance = 0
-        # Obtenemos el numero de clusters del clustering actual
-        nb_clusters = len(set(current_clustering))
+		current_clustering = cromosome
[…]

### 2019-04-04 Papers/SHADE/descartados.tex
Segunda iteracion de correcciones paper SHADE (1)

@@
 In conclusion, we can safely admit that [the] Wilcoxon's signed-rank test provides empirical evidence in favour of SHADE+LS.
+
+-----------------------------------------------
+
+Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
+It has been succesfully applied in many knowledge fields.
+It has been used to guide the movement of walking robots, as well as in other advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}.
+In \cite{seret2014new} constrained clustering is presented as a useful tool in the context of applied marketing.
+Biology has also made use of constrained clustering, being able to detect simultaneous appearances of genes and proteins in biological data bases \cite{segal2003discovering}.
+In \cite{levy2008structural} a method of segmenting musical audio into separated sections using constrained clustering is described.
+Constrained clustering has also been useful to classify Internet traffic, which is considered to be one of the most fundamental functionalities in the network management field \cite{wang2014internet}.
+Electoral district design problems can also be approached using constrained clustering, as shown in \cite{brieden2017constrained}.
+Lastly, it is mandatory to name the widely known application to GPS data of constrained clustering, presented in \cite{wagstaff2001constrained}.

### 2019-04-04 Papers/SHADE/paperSHADE.tex
Segunda iteracion de correcciones paper SHADE (1)

@@
 Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning.
-This paper proposes a variant of SHADE, a successful approach to the constrained clustering problem, which is a generalization of traditional clustering that considers additional information in the form of constraints.
-In particular, this work focuses on instance-level must-link and cannot-link constraints.
-We will compare the results obtained by this proposal to those obtained by a previous nature-inspired technique---BRKGA---and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
+This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
+We propose the first application of Differential Evolution to the constrained clustering problem, which is a generalization of traditional clustering that considers additional information in the form of constraints.
+We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
-Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology and genetics among others \cite{Everitt:2009:CA:1538772}.
+Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics, cartographic labeling and image segmentation among others \cite{Everitt:2009:CA:1538772, araujo2019improving, verma2016improved, aparajeeta2016modified, wang2018non}.
 
@@
 The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$.
-An important detail is that each instance must belong to a single cluster.
-\cite{jain1999data}
+An important detail is that each instance must belong to a single cluster \cite{jain1999data}.
 
@@
 The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
-The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper.
-\cite{triguero2015self}
-
-Constrained clustering is an SSL learning method.
-Its goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
-It has been succesfully applied in many knowledge fields.
-It has been used to guide the movement of walking robots, as well as in other advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}.
-In \cite{seret2014new} constrained clustering is presented as a useful tool in the context of applied marketing.
-Biology has also made use of constrained clustering, being able to detect simultaneous appearances of genes and proteins in biological data bases \cite{segal2003discovering}.
-In \cite{levy2008structural} a method of segmenting musical audio into separated sections using constrained clustering is described.
-Constrained clustering has also been useful to classify Internet traffic, which is considered to be one of the most fundamental functionalities in the network management field \cite{wang2014internet}.
-Electoral district design problems can also be approached using constrained clustering, as shown in \cite{brieden2017constrained}.
-Lastly, it is mandatory to name the widely known application to GPS data of constrained clustering, presented in \cite{wagstaff2001constrained}.
+The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{chapelle2009semi, triguero2015self}.
+
+Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
+It has been successfully applied in many knowledge fields, among which it is worth mentioning: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, and lane finding in GPS data \cite{wagstaff2001constrained} among others.
 
@@
 In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints.
-Pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
-On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure.\cite{davidson2007survey} This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
+In one hand, pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
+On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure \cite{davidson2007survey}.
+This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
 
@@
 Therefore, the incorporation of constraints may modify the complexity of the clustering problem, depending on the type of constraints used.
-As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete.
-\cite{davidson2005clustering}
+As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete \cite{davidson2005clustering}.
 
@@
 Nature-inspired algorithms are designed to emulate natural optimization phenomena, such as evolution, collective behavior of animals, physics laws or even human being-related processes.
-There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the Biased Random-key Genetic Algorithm (BRKGA), which can be found in \cite{de2017comparison}.
+There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the Biased Random-key Genetic Algorithm (BRKGA) presented in \cite{de2017comparison}.
 Swarm-based methods have also been applied to constrained clustering, such as the one presented in \cite{xu2013improving}.
 
-Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}; in this paper we propose a new DE variant to find quality solutions for the constrained clustering problem (see Section \ref{sec:SHADE}).
+Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}; in this paper we propose the first DE application to find quality solutions for the constrained clustering problem (see Section \ref{sec:SHADE}).
+In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
 We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant.
@@
 In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
-Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis.
+Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
 Finally, in Section \ref{sec:conclusiones} we discuss conclusions.
@@
 It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms.
-
-Given the notation for the clustering problem introduced in Section \ref{sec:Intro}, we can now formalize the two type of constraints mentioned:
+We can now formalize the two type of constraints mentioned:
 
@@
 This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified.
-In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$.
-\cite{davidson2007survey}
-
-On the other hand, CL constraints do not constitute an equivalence relation.
+In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
+
+It can also be proven that CL constraints do not constitute an equivalence relation.
 However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $C_{\neq}(x_a,x_b)$, then it is also true that $C_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
@@
 
-In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods, for constrained clustering.
+In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
 The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
@@
 
-\subsection{Short Introduction to Differential Evolution} \label{sec:BackDE}
+\subsection{Differential Evolution} \label{sec:BackDE}
 
@@
 In this type of techniques a population of individuals---representing potential solutions to the problem---is usually employed, which compete and combine through a certain number of generations so that the population evolves in such a way that only the best individuals, and thus the best solutions, remain at the end.
-This type of techniques involves the development of a series of operators that allow us to simulate the evolution process, such as crossover, mutation and selection operators.
-\cite{fausto2019ants}
-
-DE, which emerged in 1995 as a technical report presented in \cite{storn1997differential}, is a good example of evolution-based methods.
+This type of techniques involves the development of a series of operators that allow us to simulate the evolution process, such as crossover, mutation and selection operators \cite{fausto2019ants}.
+
+DE, which emerged in \cite{storn1997differential}, is a good example of evolution-based methods.
 From that point on, the reputation of DE was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
@@
 A highly detailed analysis of DE and its variants can be found in \cite{das2011differential}.
-\textcolor{red}{No he podido encontrar las referencias a los resultados de las compteciones en las que ganaba DE (ni SHADE)}
 
@@
 Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
-We will focus on the SHADE variant, which adaptively optimizes some of the parameters of DE (see Section \ref{sec:SHADE}).
+We will focus on the SHADE variant, which adaptively optimizes some of the parameters of DE.
+To carry out this task SHADE uses a historic record of values for these parameters that allows it to set them in favor of the optimization process, based on how good the algorithm behaved with these parameter values in the past (see Section \ref{sec:SHADE}).
 
@@
 Genetic Algorithms (GAs) emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
-It is one of the most studied and successful evolutionary algorithms, thanks in part to its simplicity and ease of implementation.
-\cite{fausto2019ants}
+It is one of the most studied and successful evolutionary algorithms, thanks in part to its simplicity and ease of implementation \cite{goldberg1989genetic, fausto2019ants}.
 
@@
 The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function whose task is to assign a value to each chromosome to assess its quality as a solution for the problem.
-GAs apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal.
-\cite{fausto2019ants}
-
-GAs have been applied to all kinds of problems, including problems whose solutions are subject to strong restrictions such as machine scheduling, resource location and the quadratic assignment problem.
+GAs apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal \cite{fausto2019ants}.
+
+%GAs have been applied to all kinds of problems, including problems whose solutions are subject to strong restrictions such as machine scheduling, resource location and the quadratic assignment problem.
 These problems present the peculiarity that, by applying generalized crossover operators to two individuals that represent a feasible solution to the problem, we can obtain non-feasible ones.
-The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to this kind of problems.
+
+The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to high constrained problems.
 It uses vectors of random numbers to represent solutions to the problem (individuals), which are used as keys to obtain the actual solutions.
@@
 The Biased Random-key Genetic Algorithm (BRKGA) was first proposed in \cite{gonccalves2011biased} as a variant of the RKGA.
-In the BRKGA optimization process, first, the population $P$ of $p_\text{size}$ vectors of random-keys is initialized; this will be the initial generation.
+In the BRKGA optimization process, first, the population $P$ of $|P|$ vectors of random-keys is initialized; this will be the initial generation.
 Each random-key vector $p_i$ has $v$ random-keys in it.
-The population is sorted by the fitness value $f_i$ of each $p_i$ to select the first $p_e$ individuals, this is, the elite of the population.
+The population is sorted by the fitness value $f_i$ of each $p_i$ to select the first $P_e$ individuals, this is, the elite of the population.
 The elite will be preserved in the next generation without modification.
-To introduce diversity, a number $p_m$ of new random-keys vectors are also included in the next generation.
-The remaining individuals of the new generation ($p_\text{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents.
+To introduce diversity, a number $P_m$ of new random-keys vectors are also included in the next generation.
+The remaining individuals of the new generation ($|P| - P_e - P_m$) are obtained through crossovers between elite and non-elite parents.
 
@@
 The dependent part of the problem is composed of the decoder and the fitness function, which allow us to obtain actual solutions for the problem and its evaluation.
-Therefore, to specify a BRKGA optimization scheme it is only necessary to define the decoder and the fitness function.
-\cite{gonccalves2011biased}
+Therefore, to specify a BRKGA optimization scheme it is only necessary to define the decoder and the fitness function \cite{gonccalves2011biased}.
 
@@
 Thus, the individuals produced by the local search are only used to update the best solution found so far if needed.
-For more details on BRKGA see \cite{de2017comparison}.
+For more details on BRKGA$_{CC}$ see \cite{de2017comparison}.
 
@@
 Since SHADE is an evolution-based algorithm, it shares elements with BRKGA (Section \ref{sec:brkga}), so we will use a similar notation to define its elements.
-We will note the population as $P_G$, formed by $p_\text{size}$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
+We will note the population as $P_G$, formed by $|P|$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
 
@@
 \noindent where $m_{[i,G]}$ is the mutant vector, which is generated with an individual $p_{[i,G]}$ from the population serving as a starting point.
-Indices $r1$ and $r2$ are random values in the range $[0,p_\text{size}]$, different from $i$ and from each other.
-The individual $p_{[p\text{best}, G]}$ is randomly selected from among the best $p_\text{size} \times p\text{best}\;|\;p\text{best}\in [0,1]$ individuals in the population.
+Indices $r1$ and $r2$ are random values in the range $[0,|P|]$, different from $i$ and from each other.
+The individual $p_{[p\text{best}, G]}$ is randomly selected from among the best $|P| \times p\text{best}\;|\;p\text{best}\in [0,1]$ individuals in the population.
 This way, the parameter $p\text{best}$ controls the greediness of the mutation strategy.
@@
 \end{array}
-\right.
+\right..
 \label{eq5}
@@
 \begin{equation}
-CR_i = \text{randn}_i(M_{[CR,r_i]}, 0.1)
+CR_i = \text{randn}_i(M_{[CR,r_i]}, 0.1),
 \label{eq6}
@@
 \end{array}
-\right.
+\right.,
 \label{eq8}
@@
 \end{array}
-\right.
+\right..
 \label{eq9}
@@
 \begin{equation}
-\text{mean}_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{[RC,i]}
+\text{mean}_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{[RC,i]},
 \label{eq10}
@@
 \begin{equation}
-\text{mean}_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{[F,i]}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{[F,i]}}
+\text{mean}_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{[F,i]}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{[F,i]}}.
 \label{eq12}
@@
 \begin{equation}
-f_i = z_i * (\text{infeasibility}_i + 1)
+f_i = z_i * (\text{infeasibility}_i + 1).
 \label{eq14}
@@
 
-Table \ref{tab:fitnessfunctions} shows fitness values for each partition from Figure \ref{img:toydatasets}.
+Table \ref{tab:fitnessfunctions} shows the fitness values for each partition from Figure \ref{img:toydatasets}.
 We refer to the function presented in Equation \eqref{eq1} as $f_1$ and to the function presented in Equation \eqref{eq14} as $f_2$.
@@
 As in \cite{de2017comparison} we apply a local optimization procedure to the individuals of the population, but without transferring its results to the original individual in order to maintain diversity.
-The difference is that we do not apply it to the whole population but only to a number $p_e$ of individuals considered as the elite of the population.
+The difference is that we do not apply it to the whole population but only to a number $P_e$ of individuals considered as the elite of the population.
 The percentage of individuals considered as elite must be determined for each case.
 
-Algorithm \ref{alg:SHADE} describes the overall SHADE optimization process and it is an adaptation of the one that can be found in \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
+Algorithm \ref{alg:SHADE} describes the overall SHADE$_{CC}$ optimization process, on the basis of the one that can be found in \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
+The goal of this procedure is to locally improve solutions (individuals from $P$) in a non-exhaustive way.
+To do so, it randomly choses an instance from the dataset and iteratively assigns it to different clusters.
+When improvement in the fitness function is detected, the change is transferred to the solution, when there is no possible improvement the local optimization process stops.
 
@@
 	\SetKwFunction{LocalSearch}{LocalSearch}
-	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, population size $p_{size}$, elite size $p_e$, number of clusters $k$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, population size $p_{size}$, elite size $P_e$, number of clusters $k$.}
 	\tcp{Initialization phase}
@@
 
-For our experiments we will compare the results obtained by BRKGA and SHADE over 25 datasets and 3 constraints set for each one of them.
-Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
+For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraints set for each one of them.
+Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
@@
 
-We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with four small percentages of the size of the dataset: 10\%, 15\% and 20\%.
-With $n_f$ being the fraction of the size of the dataset associated to each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
+We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with three small percentages of the size of the dataset: 10\%, 15\% and 20\%.
+With $n_f$ being the fraction of the size of the dataset associated with each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
 
@@
 			\hline
-			$p_\text{size}$ & Population size & 100 & 100 \\
+			$|P|$ & Population size & 100 & 100 \\
 			Evals & Fitness function evaluations & 300000 & 300000 \\
-			$p_e$ & Size of the elite set in population & $0.2 * p_\text{size}$ & $0.25 * p_\text{size}$ \\
-			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_\text{size}$ & - \\
+			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
+			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
 			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
@@
 
-To compare with the state-of-the-art methods mentioned in section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}
 
@@
 
-\clearpage
-
 \subsection{Comparing SHADE$_{CC}$ with the state-of-the-art} \label{sec:SHADEvsSOTA}
@@
 Tables from \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
-To shorten the presented results this time we will only compare the ARI
+To shorten the presented results this time we will only compare the ARI.
 
@@
 
-In Table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE$_{CC}$.
+In one hand, table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE$_{CC}$.
 On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by SHADE$_{CC}$, with the exception of COPKM, which is able to reach the optimum for some of the cases in which it finds a solution.
@@
 
-Figure \ref{fig:bayesARI} shows the heatmap associated with the ARI measurement.
+In one hand, figure \ref{fig:bayesARI} shows the heatmap associated with the ARI measurement.
 The diagram suggests that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle.
@@
 Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE$_{CC}$ performs better than BRKGA$_{CC}$.
-
 On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} which show the heatmaps obtained for Time and Unsat measurements, are very similar.
@@
 
+Our work has been supported by the research project TIN2017-89517-P and PP2016.PRI.I.02.
+
 \clearpage

### 2019-04-10 Papers/DILS/paperDILS.tex
Nuevos resultados y secciones en paper SHADE
Tablas y graficas añadidas a DILS

@@
 
-%\usepackage{pdfpages}
-\usepackage{graphics, graphicx, float, url}
+\usepackage{graphics, float, url}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{lineno,hyperref}
@@
 
-\title{Improving instance-level constrained clustering through Dual Iterated Local Search}
+\title{Enhancing instance-level constrained clustering through differential evolution}
 
-\author[mymainaddress]{Germ\'an Gonz\'alez Almagro\corref{mycorrespondingauthor}}
+\author[mymainaddress]{Germ\'an Gonz\'alez-Almagro\corref{mycorrespondingauthor}}
 \cortext[mycorrespondingauthor]{Corresponding author}
@@
 
+\author[mymainaddress]{Juli\'an Luengo}
+
+\author[mysecondaddress]{Jos\'e-Ram\'on Cano}
+
 \author[mymainaddress]{Salvador Garc\'ia}
 
-\author[mymainaddress]{Another Juan}
+\address[mymainaddress]{DaSCI Andalusian Institute of Data Science and Computational Intelligence, University of Granada, Spain}
 
-\address[mymainaddress]{Department of Computer Science and Artificial Intelligence, University of Granada, 18071 Granada, Spain}
+\address[mysecondaddress]{Dept. of Computer Science, EPS of Linares, University of Ja\'en, Campus Cient\'ifico Tecnol\'ogico de Linares, Cintur\'on Sur S/N, Linares 23700, Ja\'en, Spain}
 
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+Clustering has always been a powerful tool in knowledge discovery.
+Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning.
+This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
+We propose the first application of Differential Evolution to the constrained clustering problem, which is a generalization of traditional clustering that considers additional information in the form of constraints.
+We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 \begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
+constrained clustering, instance-level, must-link, cannot-link, genetic algorithm, differential evolution, Bayesian statistical tests.
 \end{keyword}
@@
 
-\textcolor{red}{Revisar que todos los subindices en los que aparecan comas lleven corchetes.
-Revisar que las asignaciones en los algoritmo sean flechas y no =.
-Revisar que restrictions nunca aparece, siempre constraints.
-Revisar que es siempre instance-level y no instance level.
-Comprobar que es siempre dataset y no data set}
+\section{Introduction} \label{sec:Intro}
 
-\section{Introduction}
+One of the most widely known and studied data analysis problems is clustering.
+It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
+Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics, cartographic labeling and image segmentation among others \cite{Everitt:2009:CA:1538772, araujo2019improving, verma2016improved, aparajeeta2016modified, wang2018non}.
 
-\section{Background}
+In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
+The main difference is that in hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
+A representative example of partitional clustering is the widely studied and well-known algorithm K-means, while for hierarchical clustering the CURE method should be highlighted \cite{wu2009top, guha1998cure}.
+In this paper we will focus on partitional clustering.
 
-%\subsection{Constrained Clustering}
+We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them.
+A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features.
+More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,d]})$.
+A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$.
+As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$.
+The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$.
+An important detail is that each instance must belong to a single cluster \cite{jain1999data}.
 
-Clustering is the task of grouping instances of a dataset into subsets.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset.
-In this way, the clustering process obtains subsets of instances of the initial dataset that present high intra-similarity and low inter-similarity.
+Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
+We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering.
+The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
+The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{chapelle2009semi, triguero2015self}.
+
+Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
+It has been successfully applied in many knowledge fields, among which it is worth mentioning: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, and lane finding in GPS data \cite{wagstaff2001constrained} among others.
+
+Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level, instance-level and feature-level constrained clustering \cite{bradley2000constrained,davidson2007survey,schmidt2011clustering}.
+Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.
+
+In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints.
+In one hand, pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
+On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure \cite{davidson2007survey}.
+This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
+
+Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard and soft constraints.
+Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new, wagstaff2001constrained, law2004clustering}.
+For the purposes of this paper, we will employ the latter.
+
+Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem.
+Therefore, the incorporation of constraints may modify the complexity of the clustering problem, depending on the type of constraints used.
+As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete \cite{davidson2005clustering}.
+
+The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it.
+As mentioned earlier, it is a difficult problem to solve, so nature-inspired techniques are presented as a promising option to find quality approximate solutions.
+Nature is the best example of adaptive problem solving since it can apply an optimal strategy suited for each natural phenomenon \cite{fausto2019ants}.
+Nature-inspired algorithms are designed to emulate natural optimization phenomena, such as evolution, collective behavior of animals, physics laws or even human being-related processes.
+There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the Biased Random-key Genetic Algorithm (BRKGA) presented in \cite{de2017comparison}.
+Swarm-based methods have also been applied to constrained clustering, such as the one presented in \cite{xu2013improving}.
+
+Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}; in this paper we propose the first DE application to find quality solutions for the constrained clustering problem (see Section \ref{sec:SHADE}).
+In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
+We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant.
+We will properly compare the results obtained by our new method with previous nature-inspired algorithms, as well as with the constrained clustering state-of-the-art.
+
+Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
+In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
+Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
+In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
+Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
+Finally, in Section \ref{sec:conclusiones} we discuss conclusions.
+
+\section{Background} \label{sec:background}
+
+In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering (Section \ref{sec:BackSOTA}).
+We will also present the basis of the DE optimization method (Section \ref{sec:BackDE}).
+
+\subsection{Constrained Clustering} \label{sec:BackCC}
 
 In most clustering applications it is common to have some kind of information about the dataset to be analyzed.
-In constrained clustering this information is given in the form of pairs of instances.
+In pairwise instance-level constrained clustering this information is given in the form of pairs of instances.
 A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster.
-**Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms**.
-[It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms]
-
-Before formalizing the definition of constraints we must give a notation for the dataset.
-We will note our dataset as $X$, which is a set of $N$ instances with $D$ features.
-We well refer to the instances of $X$ as $x$.
-Now, given the definition of dataset, we can distinguish between two types of constraints:
+It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms.
+We can now formalize the two type of constraints mentioned:
 
@@
 
-	\item Must-link constraints $c_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
+	\item Must-link constraints $C_=(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
 
-	\item Cannot-link constraints $c_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.
+	\item Cannot-link constraints $C_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.
 
@@
 
-The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set.
-As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $N = |X| = \sum_{i = 1}^{K} |c_i|$.
+The goal of constrained clustering is to find a partition (or clustering) of $k$ clusters $C = \{c_1, \cdots, c_k\}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set.
+As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $n = |X| = \sum_{i = 1}^{k} |c_i|$.
 
-Knowing how a constraint is defined, Must-link constraints---from now on ML---are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric.
-This way, given constraints $c_=(x_a,x_b)$ and $c_=(x_b,x_c)$ then $c_=(x_a,x_c)$ is verified.
-In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $c_=(x_a,x_b)$, then $c_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.
+Knowing how a constraint is defined, ML constraints are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric.
+This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified.
+In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
 
-On the other hand, Cannot-link constraints---from now on CL---do not constitute an equivalence relation.
-However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then it is also true that $c_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
+It can also be proven that CL constraints do not constitute an equivalence relation.
+However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $C_{\neq}(x_a,x_b)$, then it is also true that $C_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
 
-\subsection{The Feasibility Problem}
+\subsection{The Feasibility Problem} \label{sec:BackFeas}
 
-Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the problem's complexity.
-The feasibility problem for non-hierarchical constrained clustering was defined by \cite{davidson2005clustering} as in Definition \ref{def1}.
+Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem.
+The feasibility problem for non-hierarchical instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
 
@@
 
-	\textbf{Feasibility Problem}: given a dataset $X$, a set of constraints $R$, and the bounds on the number of clusters $K_l \leq k \leq K_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $R$ are satisfied?
-\cite{davidson2007survey}\cite{davidson2005clustering}
+	\textbf{Feasibility Problem}: given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
+\cite{davidson2005clustering}
 	\label{def1}
@@
 
-\textcolor{red}{El parrafo de abajo esta bien?}
+Intractable problems are hard to solve with deterministic and exact methods.
+That is the reason why nature-inspired algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
 
-Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why heuristics-guided and population-based algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
+\subsection{State-of-the-art Methods} \label{sec:BackSOTA}
+
+Constrained clustering has many applications and has been widely studied in the literature.
+The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}.
+It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
+This method was named COP-kmeans, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
+
+In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
+The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
+This method redefines constraints from the point of view of belief functions and includes them in the fitness function.
+
+A modification of the Constrained Vector Quantization Error algorithm (CVQE \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}.
+The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity.
+Linear CVQE (LCVQE) introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex.
+The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.
+
+Two Views Clustering (TVClust) and Relation Dirichlet Process - Means (RDPM) were proposed in \cite{khashabi2015clustering}.
+TVClust is able to incorporate the constraints into the clustering problem by making a soft interpretation of them.
+The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
+Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
+This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($k$) does not need to be specified.
 
@@
 
-\subsection{(Classic) Resolution Methods}
-
-\textcolor{red}{Clasicos clasicos no son}
-
-Constrained clustering has many applications and has been widely studied in the literature.
-The first adaptation of a classic clustering method for constrained clustering was proposed by \cite{wagstaff2001constrained}.
-It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
-This method was named COP-kmeans.
-
-The authors of \cite{antoine2012cecm} propose an adaptation of the evidential C-means algorithm, within the fuzzy clustering framework, for constrained clustering.
-The particularity of this resolution method is that the membership of instances to a cluster is defined by a probabilistic belief function.
-This method redefines constraints from the point of view of belief functions and includes them in the cost function.
-
-A modification of CVQE (Constrained Vector Quantization Error), LCVQE, is proposed in \cite{pelleg2007k}.
-The adaptation of CVQE for constrained clustering was proven to produce high quality results, at the cost of a very high computational complexity.
-LCVQE (Linear CVQE) introduces a modification of CVQE's objective function to make it more intuitive and less computationally complex.
-The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraints sets.
-
-TVClust (Two Views Clustering) and RDP-means (Relation Dirichlet Process - Means) were proposed by \cite{khashabi2015clustering}.
-TVClust is able to incorporate the constraints into the clustering problem by making a relaxed interpretation of them.
-The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
-Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
-This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($K$) does not need to be specified.
-
-\textcolor{red}{Decir que voy a comparar con ellos? Incluir mas métodos ``clasicos''?}
-
-\section{Brief Review of the BRKGA Algorithm}
-
-The biased random-key genetic algorithm---from now on BRKGA---was first proposed by \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.
-
-In BRKGA, each solution---also called individual---is represented as a vector of values within the interval $[0,1]$.
-However, this vector is not a solution to the problem but a representation of it.
-This allows us to abstract the details of the problem in order to apply a genetic algorithm that operates with real coding.
-Therefore, a deterministic decoder is necessary to obtain the actual solution to the problem from the vector that encodes it.
-
-BRKGA also requires a fitness function to evaluate each individual.
-This function takes the decoded solution as input and provides its fitness value as output.
-It is designed specifically for each problem.
-
-The BRKGA optimization process is completely independent of the problem to solve.
-First, a population $P$ of $p_{size}$ vectors of random-keys is initialized, and this will be the initial generation.
-Each random-key vector $p_i$ has $N$ random-keys in it.
-The population is sorted by the fitness value $f_i$ of each $p_i$ to select the first $p_e$ individuals, this is, the elite of the population.
-The elite will be preserved in the next generation without modification.
-To introduce diversity, a number $p_m$ of new random-keys vectors are also included in the next generation.
-The remaining individuals of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents \cite{de2017comparison}.
-
-\subsection{Adaptation of BRKGA for Constrained Clustering} \label{sec:AdaptationofBRKGA}
-
-An adaptation of BRKGA for constrained clustering was proposed in \cite{de2017comparison}.
-
-Their method randomly initializes a population of random-key vectors, each one with $N$ random-keys.
-The decoder divides the interval $[0,1]$ in $K$ intervals, so there exists a correspondence between each random-key and the integer corresponding to the interval which it lies in.
-Table \ref{tab:decodingrk} shows an example of random-key decoding for $K = 3$.
-Note that extreme values 0 and 1 can also appear in a random-key vector.
-
-\begin{table}[!h]
-	\centering
-	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	\resizebox{\textwidth}{!}{
-	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
-		\hline
-		Index & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
-		\hline
-		Random-key & 0.12 & 0.37 & 0.66 & 0.56 & 0.00 & 0.97 & 0.23 & 0.25 & 0.15 & 1.00 \\
-		\hline
-		Clusters & 1 & 2 & 2 & 2 & 1 & 3 & 1 & 3 & 1 & 3 \\
-		\hline
-
-	\end{tabular}}
-	\caption{Random-key decodification example \cite{de2017comparison}}
-	\label{tab:decodingrk}
-\end{table}
-
-Once the decoded solution has been obtained, the fitness value of the solution is computed as in Equation \eqref{eq1}.
-
-\begin{equation}
-f_i = z_i + \overbrace{(\mu * N * {infeasibility}_i)}^{penalty}
-\label{eq1}
-\end{equation}
-
-Where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.
-
-\begin{equation}
-z_i = \sum_{c_j \in C_i} \left[ \frac{\sum_{x_a, x_b \in c_j} d_D^2(x_a,x_b)}{|c_j|}\right]
-\label{eq2}
-\end{equation}
-
-Where $C_i$ is the partition of dataset $X$ defined by solution $p_i$, and $(x_a, x_b)$ are instances in the cluster $c_j$ such that $a \neq b$ and the distance between each pair of instances in $c_i$ is included in the sum only once.
-
-In \cite{de2017comparison}, the authors also added a new element to BRKGA, a local search procedure.
-This local optimization method is applied to each decoded individual in each generation, but its results are not transferred to the individual, so that diversity is maintained.
-Thus, the individuals produced by the local search are only used to update the best solution found so far if needed.
-For more details on BRKGA see \cite{de2017comparison}.
-
-\section{The DILS Optimization Method}
+\section{The Dual Iterative Local Search Method}
 
@@
 
-\section{Experimental Setup}
+\section{Experimental Setup} \label{sec:expSetup}
 
-For our experiments we will compare the results obtained by BRKGA and SHADE over XX datasets.
-Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\cite{triguero2017keel}, though some of them have been obtained via
+For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraints set for each one of them.
+Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at GitHub.
 Table \ref{tab:datasets} displays a summary of every dataset.
-\textcolor{red}{Esta bien citar tambien al sklearn?
-Que paper de keel cito, 2017 o 2011?
-Decir que algunos datasets son artifiales si se quedan, igual para los del sklearn}
 
@@
 		\hline
-		Name & No. Instances & No. Classes & Features \\
+		Name & No. Instances & No. Classes & No. Features \\
 		\hline
-		Appendicitis & 106 & 2 &  \\
-		Balance & 625 & 3 &  \\
-		Boston & 506 & 229 &  \\
-		Breast Cancer & 569 & 2 &  \\
-		Bupa & 345 & 2 &  \\
-		Circles & 300 & 2 &  \\
-		Contraceptive & 1473 & 3 &  \\
-		Diabetes & 442 & 214 &  \\
-		Ecoli & 336 & 8 &  \\
-		Glass & 214 & 6 &  \\
-		Haberman & 306 & 2 &  \\
-		Hayesroth & 160 & 3 &  \\
-		Heart & 270 & 2 &  \\
-		Ionosphere & 351 & 2 &  \\
-		Iris & 150 & 3 &  \\
-		Led7Digit & 500 & 10 &  \\
-		Monk2 & 432 & 2 &  \\
-		Moons & 300 & 2 &  \\
-		Movement Libras & 360 & 15 &  \\
-		Newthyroid & 215 & 3 &  \\
-		Pima & 768 & 2 &  \\
-		Rand & 150 & 3 &  \\
-		Saheart & 462 & 2 &  \\
-		Sonar & 208 & 2 &  \\
-		Soybean & 47 & 4 &  \\
-		Spectfheart & 267 & 2 &  \\
-		Spiral & 300 & 2 &  \\
-		Tae & 151 & 3 &  \\
-		Vehicle & 846 & 4 &  \\
-		Vowel & 990 & 11 &  \\
-		Wdbc & 569 & 2 &  \\
-		Wine & 178 & 3 &  \\
-		Zoo & 101 & 7 &  \\
+		Appendicitis & 106 & 2 & 7 \\
+		Breast Cancer & 569 & 2 & 30 \\
+		Bupa & 345 & 2 & 6 \\
+		Circles & 300 & 2 & 2 \\
+		Ecoli & 336 & 8 & 7 \\
+		Glass & 214 & 6 & 9 \\
+		Haberman & 306 & 2 & 3 \\
+		Hayesroth & 160 & 3 & 4 \\
+		Heart & 270 & 2 & 13 \\
+		Ionosphere & 351 & 2 & 33 \\
+		Iris & 150 & 3 & 4 \\
+		Led7Digit & 500 & 10 & 7 \\
+		Monk2 & 432 & 2 & 6 \\
+		Moons & 300 & 2 & 2 \\
+		Movement Libras & 360 & 15 & 90 \\
+		Newthyroid & 215 & 3 & 5 \\
+		Saheart & 462 & 2 & 9 \\
+		Sonar & 208 & 2 & 60 \\
+		Soybean & 47 & 4 & 35 \\
+		Spectfheart & 267 & 2 & 44 \\
+		Spiral & 300 & 2 & 2 \\
+		Tae & 151 & 3 & 5 \\
+		Vowel & 990 & 11 & 13 \\
+		Wine & 178 & 3 & 13 \\
+		Zoo & 101 & 7 & 16 \\
 		\hline
@@
 	\end{tabular}%}
-	\caption{Summary of datasets used for the experiments \cite{triguero2017keel}\cite{scikit-learn}}
+	\caption{Summary of datasets used for the experiments.}
 	\label{tab:datasets}
@@
 
+Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
+They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).
+
 \clearpage
 
-\subsection{Constraints Generation}
+\subsection{Constraint Generation} \label{sec:ConstGent}
 
-\textcolor{red}{En esta seccion podria simplemente decir el numero de restricciones que genero como un numero magico sin meterme en nada de grafos, pero igual es un poco burro.}
+Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets.
+This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
 
-Since we have the true labels associated with each dataset, we will use the method proposed by \cite{wagstaff2001constrained} to generate artificial constraint sets.
-This method consists in randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
+We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with three small percentages of the size of the dataset: 10\%, 15\% and 20\%.
+With $n_f$ being the fraction of the size of the dataset associated with each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
 
-We will generate four different sets of constraints for each dataset.
-The number of constraints is given by the number of edges of the complete constraints graph that we can build with a percentage of labeled data.
-Each constraint set is associated with a different percentage of labeled data, namely: 5\%, 10\%, 15\% and 20\%.
-It is worth noting that constraints are not the complete constraint graph made up of the labeled instances, this graph only sets the number of constraints to be created, this way there is a lower probability of biasing the set of constraints so that there are classes with poor representation in it.
+The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
 Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.
@@
 	\resizebox{\textwidth}{!}{
-	\begin{tabular}{lcc c cc c cc c cc}
+	\begin{tabular}{lcc c cc c cc}
 		\hline
 		\multirow{2}{*}{Dataset} &
-		\multicolumn{2}{c}{5\%} &&  \multicolumn{2}{c}{10\%} && \multicolumn{2}{c}{15\%} && \multicolumn{2}{c}{20\%} \\
-		\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
-		& ML & CL && ML & CL && ML & CL && ML & CL \\
+		\multicolumn{2}{c}{$CS_{10}$} && \multicolumn{2}{c}{$CS_{15}$} && \multicolumn{2}{c}{$CS_{20}$} \\
+		\cline{2-3} \cline{5-6} \cline{8-9}
+		& ML & CL && ML & CL && ML & CL \\
 		\hline
-		Appendicitis & 13 & 2 && 39 & 16 && 71 & 49 && 164 & 67 \\
-		Balance & 198 & 298 && 838 & 1115 && 1865 & 2506 && 3336 & 4414 \\
-		Boston & 0 & 325 && 3 & 1272 && 13 & 2837 && 26 & 5125 \\
-		Breast Cancer & 216 & 190 && 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
-		Bupa & 79 & 74 && 323 & 272 && 699 & 627 && 1201 & 1145 \\
-		Circles & 50 & 55 && 208 & 227 && 502 & 488 && 853 & 917 \\
-		Contraceptive & 994 & 1707 && 3765 & 7113 && 8501 & 15809 && 15399 & 27966 \\
-		Diabetes & 2 & 251 && 5 & 985 && 7 & 2204 && 23 & 3893 \\
-		Ecoli & 30 & 106 && 163 & 398 && 357 & 918 && 609 & 1669 \\
-		Glass & 11 & 44 && 52 & 179 && 139 & 389 && 259 & 644 \\
-		Haberman & 76 & 44 && 304 & 161 && 634 & 401 && 1135 & 756 \\
-		Hayesroth & 12 & 16 && 39 & 81 && 102 & 174 && 177 & 319 \\
-		Heart & 41 & 50 && 178 & 173 && 396 & 424 && 744 & 687 \\
-		Ionosphere & 92 & 61 && 330 & 300 && 732 & 646 && 1299 & 1186 \\
-		Iris & 9 & 19 && 26 & 79 && 82 & 171 && 136 & 299 \\
-		Led7Digit & 25 & 275 && 126 & 1099 && 267 & 2508 && 460 & 4490 \\
-		Monk2 & 101 & 130 && 473 & 473 && 979 & 1101 && 1917 & 1824 \\
-		Moons & 55 & 50 && 200 & 235 && 494 & 496 && 900 & 870 \\
-		Movement Libras & 6 & 147 && 27 & 603 && 112 & 1319 && 158 & 2398 \\
-		Newthyroid & 25 & 30 && 108 & 123 && 270 & 258 && 449 & 454 \\
-		Pima & 412 & 329 && 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
-		Rand & 8 & 20 && 25 & 80 && 76 & 177 && 151 & 284 \\
-		Saheart & 152 & 124 && 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
-		Sonar & 29 & 26 && 100 & 110 && 245 & 251 && 436 & 425 \\
-		Soybean & 0 & 3 && 4 & 6 && 6 & 22 && 12 & 33 \\
-		Spectfheart & 56 & 35 && 233 & 118 && 543 & 277 && 965 & 466 \\
-		Spiral & 52 & 53 && 224 & 211 && 487 & 503 && 918 & 852 \\
-		Tae & 8 & 20 && 40 & 80 && 82 & 171 && 151 & 314 \\
-		Vehicle & 221 & 682 && 874 & 2696 && 1955 & 6046 && 3589 & 10776 \\
-		Vowel & 107 & 1118 && 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
-		Wdbc & 209 & 197 && 840 & 756 && 1925 & 1730 && 3472 & 2969 \\
-		Wine & 14 & 22 && 49 & 104 && 121 & 230 && 217 & 413 \\
-		Zoo & 7 & 8 && 21 & 34 && 29 & 91 && 41 & 169 \\
+		Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
+		Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
+		Bupa & 323 & 272 && 699 & 627 && 1201 & 1145 \\
+		Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
+		Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
+		Glass & 52 & 179 && 139 & 389 && 259 & 644 \\
+		Haberman & 304 & 161 && 634 & 401 && 1135 & 756 \\
+		Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
+		Heart & 178 & 173 && 396 & 424 && 744 & 687 \\
+		Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
+		Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
+		Led7Digit & 126 & 1099 && 267 & 2508 && 460 & 4490 \\
+		Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
+		Moons & 200 & 235 && 494 & 496 && 900 & 870 \\
+		Movement Libras & 27 & 603 && 112 & 1319 && 158 & 2398 \\
+		Newthyroid & 108 & 123 && 270 & 258 && 449 & 454 \\
+		Saheart & 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
+		Sonar & 100 & 110 && 245 & 251 && 436 & 425 \\
+		Soybean & 4 & 6 && 6 & 22 && 12 & 33 \\
+		Spectfheart & 233 & 118 && 543 & 277 && 965 & 466 \\
+		Spiral & 224 & 211 && 487 & 503 && 918 & 852 \\
+		Tae & 40 & 80 && 82 & 171 && 151 & 314 \\
+		Vowel & 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
+		Wine & 49 & 104 && 121 & 230 && 217 & 413 \\
+		Zoo & 21 & 34 && 29 & 91 && 41 & 169 \\
 		\hline
@@
 
-\subsection{Evaluation Method}
+\subsection{Evaluation Method} \label{sec:EvalMet}
 
 Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
-We will use the Adjusted Rand Index to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
-The Basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
-$C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions \cite{rand1971objective}.
+We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
+The basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
+$C_1$ and $C_2$ are viewed as collections of $n(n - 1)/2$ pairwise decisions \cite{rand1971objective}.
 
-For each pair of instances $x_i$ and $x_j$ in $X$, $C_i$ assigns them to the same cluster or to different clusters.
+For each pair of instances $x_i$ and $x_j$ in $X$, a partition assigns them to the same cluster or to different clusters.
 We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in both $C_1$ and $C_2$, and $b$ as the opposite event ($x_i$ and $x_j$ are in different clusters in $C_1$ and $C_2$).
@@
 \begin{equation}
-Rand(C_1, C_2) = \frac{a + b}{N(N - 1)/2}
+\text{Rand}(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
 \label{eq15}
@@
 
-The Adjusted Rand Index is a corrected-for-chance version of the Rand Index.
+The ARI is a corrected-for-chance version of the Rand Index.
 This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline.
-The Adjusted Rand Index is computed as in Equation \eqref{eq16}.
+The ARI is computed as in Equation \eqref{eq16}.
 
 \begin{equation}
-ARI(C_1, C_2) = \frac{Rand(C_1, C_2) - ExpectedIndex}{MaximumIndex - ExpectedIndex}
+\text{ARI}(C_1, C_2) = \frac{\text{Rand}(C_1, C_2) - \text{Expected Index}}{\text{Maximum Index} - \text{Expected Index}},
 \label{eq16}
@@
 
-where $MaximumIndex$ is expected to be 1 and $ExpectedIndex$ is the already mentioned expected degree of similarity with a random model.
-It is easy to see that $ARI(C_1, C_2) \in [-1,1]$, such that an $ARI$ value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $Rand(C_1, C_2)$ is less than expected when comparing the obtained partitions with random partitions.
-To summarize, the higher the $ARI$, the greater the degree of similarity between $C_1$ and $C_2$.
-For more details on Adjusted Rand Index see \cite{hubert1985comparing}.
+\noindent where Maximum Index is expected to be 1 and Expected Index is the already mentioned expected degree of similarity with a random model.
+It is easy to see that $\text{ARI}(C_1, C_2) \in [-1,1]$, such that an ARI value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $\text{Rand}(C_1, C_2)$ is less than expected when comparing with random partitions.
+To summarize, the higher the ARI, the greater the degree of similarity between $C_1$ and $C_2$.
+For more details on ARI see \cite{hubert1985comparing}.
+
+Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper.
+To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth labels.
+
+%% CORREGIR DESDE AQUI HASTA EL FINAL
 
@@
 
-\clearpage
-
-\section{Experimental Results}
-
-\textcolor{red}{Poner tabla con parametros para los dos algoritmos.
-Decir que comparamos rendimiento medio y no rendimiento pico, porque el proceso de optimización de parámetros neceseario para cada algoritmo y dataset cae fuera de este estudio.
-Comentar que el curterio de finalización es número de evaluaciones de la funcion objetivo}
+Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA$_{CC}$ and SHADE$_{CC}$ algorithms.
 
@@
 			\hline
-			$p_{size}$ & Population size & 100 & 100 \\
-			$Evals$ & Fitness function evaluations & 300000 & 300000 \\
-			$p_e$ & Size of the elite set in population & $0.2 * p_{size}$ & $0.25 * p_{size}$ \\
-			$p_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * p_{size}$ & $\times$ \\
-			$p_{inherit}$ & Probability that a feature is inherit from an elite parent & $50\%$ & $\times$ \\
-			$K$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{Depends on the dataset and it has the same value for both methods} \\
+			$|P|$ & Population size & 100 & 100 \\
+			Evals & Fitness function evaluations & 300000 & 300000 \\
+			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
+			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
+			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
+			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets}).} \\
 			\hline
@@
 	\caption{Parameters setup used for BRKGA and SHADE.}
-		\label{undefined}
+	\label{tab:params}
+\end{table}
+
+In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} c}
+			\hline
+			Common Parameters & Meaning & Value \\
+			\hline
+			\texttt{max\_iter} & Maximum number of iterations & 300 \\
+			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}). \\
+			\hline
+			\hline
+			Specific Parameters & \multicolumn{2}{l}{Name and Value} \\
+			\hline
+			COPKM & \multicolumn{2}{l}{\texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''}} \\
+			CECM & \multicolumn{2}{l}{$\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''}} \\
+			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
+			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
+			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Parameters setup used for the state-of-the-art algorithms.}
+	\label{tab:paramsSOTA}
+\end{table}
+
+Parameter values have been assigned following the guidelines of the original creators of the different proposals.
+Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible.
+Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible.
+Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
+
+\section{Experimental Results} \label{sec:results}
+
+In this section we present Tables from \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.
+
+Since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
+To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+
+In this section we present a brief comparison between the proposed SHADE$_{CC}$ method and some of the state-of-the-art methods.
+Tables from \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
+To shorten the presented results this time we will only compare the ARI.
+
+It should be noted that there are some missing results in these tables.
+In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
+It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
+In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lccccccc}
+			\hline
+			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			\hline
+			Appendicitis & 0.611 & 0.022 & -1.000 & 0.335 & 0.316 & 0.025 & -0.005 \\
+			Breast Cancer & 0.755 & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.870 & 0.281 & -1.000 & -0.005 & -0.005 & -0.004 & -0.011 \\
+			Circles & 0.798 & 0.166 & -1.000 & -0.003 & 0.162 & 0.137 & 0.133 \\
+			Ecoli & 0.069 & 0.009 & -1.000 & 0.387 & 0.417 & 0.265 & -1.000 \\
+			Glass & 0.009 & 0.016 & 0.184 & 0.268 & 0.197 & 0.211 & -1.000 \\
+			Haberman & 0.638 & 0.047 & -1.000 & -0.002 & 0.127 & 0.218 & -0.004 \\
+			Hayesroth & 0.031 & 0.027 & -1.000 & 0.106 & 0.097 & 0.054 & 0.139 \\
+			Heart & 0.793 & 0.483 & -1.000 & 0.027 & 0.036 & 0.423 & -0.003 \\
+			Ionosphere & 0.792 & 0.125 & -1.000 & 0.168 & 0.197 & 0.000 & 0.030 \\
+			Iris & 0.621 & 0.238 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
+			Led7Digit & 0.016 & 0.009 & 0.497 & 0.425 & 0.369 & 0.316 & -1.000 \\
+			Monk2 & 0.815 & 0.421 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
+			Moons & 0.968 & 0.277 & -1.000 & 0.241 & 0.319 & 0.785 & 0.092 \\
+			Movement Libras & 0.022 & 0.004 & 0.285 & 0.293 & 0.256 & 0.000 & -1.000 \\
+			Newthyroid & 0.028 & 0.017 & -1.000 & 0.579 & 0.289 & 0.846 & 0.002 \\
+			Saheart & 0.800 & 0.147 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
+			Sonar & 0.649 & 0.124 & -1.000 & 0.004 & 0.013 & 0.000 & 0.000 \\
+			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
+			Spectfheart & 0.896 & 0.334 & -1.000 & -0.107 & -0.114 & 0.000 & 0.050 \\
+			Spiral & 0.856 & 0.058 & -1.000 & -0.003 & 0.012 & 0.034 & -0.002 \\
+			Tae & 0.031 & 0.018 & -1.000 & 0.009 & -0.000 & 0.067 & 0.000 \\
+			Vowel & 0.002 & 0.003 & -1.000 & 0.063 & -0.003 & 0.067 & -1.000 \\
+			Wine & 0.334 & 0.120 & -1.000 & 0.360 & 0.368 & 0.281 & 0.004 \\
+			Zoo & 0.151 & 0.105 & 0.715 & 0.666 & 0.412 & 0.335 & -1.000 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
+	\label{tab:results10SOTA}
 \end{table}
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabular}{l ccc c ccc}
+		\begin{tabular}{lccccccc}
 			\hline
-			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA} &&  \multicolumn{3}{c}{SHADE} \\
-			\cline{2-4} \cline{6-8}
-			& RandIndex & Unsat(\%) & Time(s) && RandIndex & Unsat(\%) & Time(s) \\
+			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
 			\hline
-			&  &  &  &&  &  &  \\
-			&  &  &  &&  &  &  \\
-			&  &  &  &&  &  &  \\
-			\hline
-			Means &  &  &  &&  &  &  \\
+			Appendicitis & 0.957 & 0.296 & -1.000 & 0.305 & 0.284 & 0.025 & -0.006 \\
+			Breast Cancer & 0.788 & 0.772 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Bupa & 0.984 & 0.840 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
+			Circles & 1.000 & 0.828 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
+			Ecoli & 0.134 & 0.015 & -1.000 & 0.387 & 0.372 & 0.719 & -1.000 \\
+			Glass & 0.057 & 0.017 & -1.000 & 0.280 & 0.253 & 0.229 & -1.000 \\
+			Haberman & 1.000 & 0.866 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
+			Hayesroth & 0.386 & 0.107 & -1.000 & 0.106 & 0.107 & 0.150 & 0.135 \\
+			Heart & 1.000 & 0.692 & 1.000 & 0.025 & 0.033 & 0.442 & 0.000 \\
+			Ionosphere & 0.970 & 0.803 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
+			Iris & 0.850 & 0.412 & -1.000 & 0.730 & 0.547 & 0.421 & 0.684 \\
+			Led7Digit & 0.020 & 0.004 & -1.000 & 0.425 & 0.490 & 0.324 & -1.000 \\
+			Monk2 & 0.902 & 0.751 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
+			Moons & 1.000 & 0.958 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Movement Libras & 0.025 & 0.006 & -0.742 & 0.293 & 0.255 & 0.000 & -1.000 \\
+			Newthyroid & 0.440 & 0.320 & -1.000 & 0.568 & 0.356 & 0.923 & 0.052 \\
+			Saheart & 0.862 & 0.749 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
+			Sonar & 0.981 & 0.800 & -1.000 & 0.004 & 0.019 & 0.000 & 0.000 \\
+			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
+			Spectfheart & 1.000 & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
+			Spiral & 1.000 & 0.847 & -1.000 & -0.003 & 0.014 & 0.006 & 0.051 \\
+			Tae & 0.299 & 0.119 & -1.000 & 0.008 & -0.004 & 0.062 & 0.000 \\
+			Vowel & 0.003 & 0.002 & -1.000 & 0.063 & -0.003 & 0.073 & -1.000 \\
+			Wine & 0.725 & 0.528 & -1.000 & 0.360 & 0.367 & 0.348 & -0.001 \\
+			Zoo & 0.185 & 0.157 & 0.435 & 0.642 & 0.450 & 0.353 & -1.000 \\
 			\hline
@@
 
-	\caption{Experimental results obtained with X\% of labeled data}
-	\label{undefined}
+		\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
+		\label{tab:results15SOTA}
 	\end{table}
 
-\section{Conclusions and Future Work}
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lccccccc}
+			\hline
+			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			\hline
+			Appendicitis & 1.000 & 1.000 & -1.000 & 0.305 & 0.331 & 0.012 & -0.006 \\
+			Breast Cancer & 0.807 & 0.807 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.991 & 0.936 & 1.000 & -0.005 & -0.007 & -0.006 & 0.000 \\
+			Circles & 1.000 & 0.935 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
+			Ecoli & 0.201 & 0.028 & -1.000 & 0.387 & 0.459 & 0.763 & -1.000 \\
+			Glass & 0.322 & 0.038 & -1.000 & 0.271 & 0.287 & 0.218 & -1.000 \\
+			Haberman & 1.000 & 0.943 & 1.000 & -0.002 & 0.106 & 0.737 & 0.000 \\
+			Hayesroth & 0.812 & 0.548 & -1.000 & 0.106 & 0.107 & 0.208 & 0.056 \\
+			Heart & 1.000 & 0.942 & 1.000 & 0.025 & 0.033 & 1.000 & 0.000 \\
+			Ionosphere & 0.982 & 0.902 & 1.000 & 0.173 & 0.309 & 0.000 & 0.115 \\
+			Iris & 0.968 & 0.691 & -1.000 & 0.708 & 0.540 & 0.585 & 0.684 \\
+			Led7Digit & 0.017 & 0.007 & -1.000 & 0.425 & 0.571 & 0.327 & -1.000 \\
+			Monk2 & 0.904 & 0.855 & 1.000 & 0.072 & 0.253 & -0.002 & 0.160 \\
+			Moons & 1.000 & 0.937 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
+			Movement Libras & 0.018 & 0.006 & -1.000 & 0.293 & 0.293 & 0.000 & -1.000 \\
+			Newthyroid & 0.842 & 0.848 & -1.000 & 0.568 & 0.393 & 0.933 & 0.004 \\
+			Saheart & 0.881 & 0.881 & 1.000 & 0.018 & 0.026 & 1.000 & -0.006 \\
+			Sonar & 1.000 & 0.992 & 1.000 & 0.004 & 0.127 & 0.000 & 0.003 \\
+			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & 0.631 & 0.000 & -0.016 \\
+			Spectfheart & 1.000 & 0.953 & 1.000 & -0.107 & -0.112 & 0.000 & -0.054 \\
+			Spiral & 1.000 & 0.921 & 1.000 & -0.003 & 0.011 & 0.006 & 0.045 \\
+			Tae & 0.892 & 0.351 & -1.000 & 0.008 & 0.000 & 0.035 & 0.000 \\
+			Vowel & 0.003 & 0.002 & -1.000 & 0.063 & -0.003 & 0.071 & -1.000 \\
+			Wine & 0.916 & 0.937 & -1.000 & 0.360 & 0.363 & 0.380 & 0.015 \\
+			Zoo & 0.203 & 0.176 & 0.821 & 0.642 & 0.439 & 0.335 & -1.000 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+		\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
+		\label{tab:results25SOTA}
+	\end{table}
+
+\clearpage
+
+\section{Statistical Analysis of Results} \label{sec:analisis}
+
+With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
+This way we can statistically determine whether SHADE$_{CC}$ represents a significant improvement over previous proposals.
+
+In order to carry out this study we will use Bayesian statistical tests, instead of the classic Hull Hypothesis Statistical Tests (NHST).
+In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
+\textit{"In a nutshell: NHST do not answer the question we ask"}.
+To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
+
+To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
+Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
+This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.
+
+Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate.
+As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
+
+Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
+That is: What can we conclude when NHST do not reject the null hypothesis?
+We can not infer anything since NHST can not provide evidence in its favor.
+
+Finally, there are two other problems that researchers face when performing NHST.
+The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results.
+The second one is the need to previously formalize the intentions of the sampling of the results, which are usually fixed a posteriori; this could lead to a misreading of said results.
+
+As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST.
+In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test.
+To make use of it we will employ the R package \texttt{rNPBST}, whose documentation and guide can be found in \cite{carrasco2017rnpbst}.
+
+From now on, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA$_{CC}$ as sample $A$, and to the results obtained with SHADE$_{CC}$ as sample $B$.
+
+The Bayesian sign test is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
+To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
+In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$.
+Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
+$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$
+
+One of the major advantages of the Bayesian sign test over the null hypothesis tests such as the (Wilcoxon's signed-rank test) is that we can obtain a visual representation of its results.
+We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
+With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle.
+In order to find out where a certain triplet will be placed within the triangle, we will take each of its three values and draw a parallel line to the opposing side of the corresponding vertex; the separation between a triangle side and its parallel line will be proportional to the associated triplet value, so that the higher the value, the closer the line will be to the vertex.
+The location where the three lines intersect is where we draw a point.
+Since the values of every triplet describe a probability distribution and therefore they must add up to one, we can be sure that all triplets will lie in some point within the triangle.
+The color indicates the density of points in a given region, with yellow representing a high density and red a low density.
+
+\begin{figure}[ht!]
+	\centering
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/DILSvsBRKGA.pdf}
+		\caption{ARI}
+		\label{fig:DILSvsBRKGA}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/DILSvsCOPKM.pdf}
+		\caption{Time}
+		\label{fig:DILSvsCOPKM}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/DILSvsLCVQE.pdf}
+		\caption{Unsat}
+		\label{fig:DILSvsLCVQE}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/DILSvsTVClust.pdf}
+		\caption{Unsat}
+		\label{fig:DILSvsTVClust}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/DILSvsRDPM.pdf}
+		\caption{Unsat}
+		\label{fig:DILSvsRDPM}
+	\end{subfigure}
+	\begin{subfigure}{.45\textwidth}
+		\includegraphics[width=1\linewidth]{Figures/DILSvsCEKM.pdf}
+		\caption{Unsat}
+		\label{fig:DILSvsCEKM}
+	\end{subfigure}
+	\caption{Heat diagrams.}
+\end{figure}
+
+\clearpage
+
+\section{Conclusions} \label{sec:conclusiones}
+
+In this paper we proposed SHADE$_{CC}$, the first application of a DE-based algorithm to the SSL constrained clustering problem.
+Focusing on the instance-level ML and CL constraints, SHADE$_{CC}$ has proven that nature-inspired evolution-based techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
+
+Supported by the Bayesian statistical tests, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous evolution-based proposals.
+As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be equivalent or better when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
 
 \section{Acknowledgements}
+
+Our work has been supported by the research project TIN2017-89517-P and PP2016.PRI.I.02.
 

### 2019-04-10 Papers/SHADE/paperSHADE.tex
Tablas y graficas añadidas a DILS

@@
 In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
-Afterward, a variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
+Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
 In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
@@
 We will focus on the SHADE variant, which adaptively optimizes some of the parameters of DE.
-To carry out this task SHADE uses a historic record of values for these parameters that allows it to set them in favor of the optimization process, based on how good the algorithm behaved with these parameter values in the past (see Section \ref{sec:SHADE}).
+To carry out this task SHADE uses a historic record of values for these parameters that allows it to set them in favor of the optimization process, based on how good the algorithm behaved with certain parameter values in the past (see Section \ref{sec:SHADE}).
 
@@
 
-In this paper we proposed SHADE$_{CC}$, a new variant of the SHADE algorithm to solve the constrained clustering problem.
-We took concepts such as the random-key and the encoder/decoder from RKGA to transform constrained clustering into a real-domain optimization problem.
-We also proposed a new fitness function that has proven to be more adaptable and less restrictive.
-
-The SHADE algorithm is one of the most successful methods to approximately solve difficult problems, and constrained clustering is not an exception.
-We obtained results on 25 different datasets and 3 different constraint sets for each one, applying all the algorithms named in this paper.
-Thanks to the Bayesian statistical tests, which offer us the advantageous option of producing a visualization of differences between results obtained by different methods, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous evolution-based proposals.
+In this paper we proposed SHADE$_{CC}$, the first application of a DE-based algorithm to the SSL constrained clustering problem.
+Focusing on the instance-level ML and CL constraints, SHADE$_{CC}$ has proven that nature-inspired evolution-based techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
+
+Supported by the Bayesian statistical tests, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous evolution-based proposals.
 As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be equivalent or better when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.

### 2019-06-11 Papers/DILS/paperDILS.tex
SHADE terminado y DILS avanzado

@@
 It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
-Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics, cartographic labeling and image segmentation among others \cite{Everitt:2009:CA:1538772, araujo2019improving, verma2016improved, aparajeeta2016modified, wang2018non}.
+Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics,\cite{Everitt:2009:CA:1538772} cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
 
 In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
-The main difference is that in hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
-A representative example of partitional clustering is the widely studied and well-known algorithm K-means, while for hierarchical clustering the CURE method should be highlighted \cite{wu2009top, guha1998cure}.
+In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
+A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
 In this paper we will focus on partitional clustering.
@@
 A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features.
-More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,d]})$.
+More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$.
 A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$.
 As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$.
-An important detail is that each instance must belong to a single cluster \cite{jain1999data}.
+The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$ \cite{jain1999data}.
 
 Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
-We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering.
+We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}.
 The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
-The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{chapelle2009semi, triguero2015self}.
+The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{triguero2015self}.
 
@@
 
-Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level, instance-level and feature-level constrained clustering \cite{bradley2000constrained,davidson2007survey,schmidt2011clustering}.
+Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level \cite{bradley2000constrained}, instance-level \cite{davidson2007survey} and feature-level constrained clustering \cite{schmidt2011clustering}.
 Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.
@@
 
-Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard and soft constraints.
-Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new, wagstaff2001constrained, law2004clustering}.
+Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained} and soft \cite{law2004clustering} constraints.
+Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}.
 For the purposes of this paper, we will employ the latter.
@@
 
-Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}; in this paper we propose the first DE application to find quality solutions for the constrained clustering problem (see Section \ref{sec:SHADE}).
+Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}.
+In this paper we propose the first DE application to find quality solutions for the constrained clustering problem.
 In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
@@
 Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
-In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
+In Section \ref{sec:brkga} the BRKGA algorithm and its adaptation for constrained clustering will be briefly reviewed.
 Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
-In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
+In Section \ref{sec:SHADEadapt} the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function, will be presented.
 Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
-Finally, in Section \ref{sec:conclusiones} we discuss conclusions.
+Finally, in Section \ref{sec:conclusiones} conclusions are discussed.
+
+\textcolor{red}{Decir que mientras que en los algoritmos memeticos el host es el GA en DILS el host es ILS, conservando todas sus caracteristicas y añadiendo algunas tecnicas de los GA}
 
@@
 
-\subsection{Short Introduction to Iterated Local Search}
+\textcolor{red}{Poner aqui una descripción del BRKGA?}
 
-Before defining the iterated local search---from now on ILS---,it is necessary to understand the local search and the motivation that leads to designing methods with greater exploratory capacity.
+\clearpage
 
-A local search is a procedure that consists of performing an heuristics-guided iterative improvement.
-Taking as initial state a well defined solution to the problem, the local search uses a neighbor generation operator to explore new solutions.
-When a solution better than the current one is found, it replaces the current one.
-In order to evaluate the solutions, the local search uses a function that assigns a value to each one of them, this is known as a target function.
-***This value will be better the closer the solution that evaluates the optimal solution.***
+\section{The Dual Iterative Local Search Method}
 
-Both the neighbor generation operator and the target function used by the local search must be defined for each problem.
-The target function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem.
-The neighbor generation operator takes as base a solution to the problem $s$, and generates a new solution $s^\prime$ by modifying some of its components in a local way.
+The Dual Iterative Local Search (DILS) is a new variant of the classic Iterated Local Search (ILS) method.
+Its goal is to perform a search in the solutions space to find the solution with the best fitness function $f$ value by introducing diversity in a guided way.
+While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-introducing methods and to avoid local optima.
+The individual $m_b$ is the one providing the best fitness value, so $m_w$ provides the worst fitness value.
 
-The Algorithm \ref{alg:LS} presents a basic scheme for a local search procedure.
-It is worth noting that the condition of line 4 must be adapted if the problem to be solved is one of maximization or minimization.
-In addition, the general scheme considers that the optimization process ends when the current solution is not improved, however the stop condition can be adapted to each problem.
+DILS mimics some of the genetic algorithms techniques to introduce diversity into the optimization process, while also manteining the esentials of ILS.
+In each generation $G$ of the optimization process DILS builds a new individual by applying a crossover operator to $m_b$ and $m_w$.
+After that, it applies a strong mutation operator to the newly generated individual.
+This is the invidual DILS applies the local search procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
+The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the replacement operator for DILS.
+Equation \ref{eq1} displays the expression for this operator in the case of a minimization problem.
+
+\begin{equation}
+m_{w,G} = \left\{ \begin{array}{lc}
+m_t &   \text{if} \;\; f(m_t) < f(m_w)\\
+m_{w,G} &  \text{otherwise}
+\end{array}
+\right..
+\label{eq1}
+\end{equation}
+
+Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solutions space.
+To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory $\{m_b, m_w\}$, Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.
+
+\begin{equation}
+m_{w,G+1} = \left\{ \begin{array}{lc}
+\text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
+m_{w,G} &  \text{otherwise}
+\end{array}
+\right.,
+\label{eq2}
+\end{equation}
+
+\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
+This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via crossover operator) as the best parent.
+Figure \ref{img:DILS_diag} summarizes the DILS optimization process.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{Figures/DILSDiagram.png}
+	\caption{Diagram summarizing DILS optimization process. \textcolor{red}{La figura es horrible (llamar a javi)}}\label{img:DILS_diag}
+\end{figure}
+
+It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reasigned to the true best and worst individual respectively.
+This way, at the point where the reinitialization criterion is tested, it is possible that $f(m_w)$ provides a better value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
+
+This leads us to note the influence of the parameter $\xi$.
+In one hand, if we are dealing with a minimization problem and we set $\xi > 0$, then the true worst individual is only reinitializatized when the trial individual $m_t$ replaced $m_w$ and is also better than $m_b$ by a margin.
+On the other hand and also for the minimization case but this time with $\xi < 0$ we allow DILS to reinitialize the true worst individual even $m_t$ is not better than $m_b$ but they are close.
+It comes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
+Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$.
+To summarize, the closer $\xi$ is to 1 the more restrictive the reinitialization criterion is---in the minimization case.
+It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages, resembling some fatures of the simmulated annealing algorithm \cite{van1987simulated}.
+Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.
 
@@
 	\setstretch{1.2}
-	\SetKwFunction{GenerateNeighbor}{GenerateNeighbor}
-	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Initial solution $s$}
-	\BlankLine
-	\While{$improvement$}{
-		$improvement \leftarrow$ \texttt{false} \\
+	\SetKwFunction{RandInit}{RandInit}
+	\SetKwFunction{Best}{Best}
+	\SetKwFunction{Worst}{Worst}
+	\SetKwFunction{Crossover}{Crossover}
+	\SetKwFunction{Mutation}{Mutation}
+	\SetKwFunction{LocalSearch}{LocalSearch}
+	\KwIn{}
+	\tcp{Initialization phase}
+	$G \leftarrow 0$\\
+	$m_b \leftarrow$ \RandInit{};
+	$m_w \leftarrow$ \RandInit{}\\
+	\tcp{Main loop}
+	\While{Termination criteria are not met}{
+		\tcp{Find best and worst individual}
+		$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $};
+		$m_w \leftarrow $ \Worst{$ \{m_b, m_w\} $}\\
+		\tcp{Generate new individual}
+		$m_t \leftarrow $ \Crossover{$m_b, \; m_w$}\\
+		$m_t \leftarrow $ \Mutation{$m_t$}\\
+		\tcp{Improve new individual}
+		$m_t \leftarrow $ \LocalSearch{$m_t$}\\
+		\tcp{Apply replacement operator}
+		\If{$f(m_t) < f(m_w)$}{
 
-		$s^\prime \leftarrow $ \GenerateNeighbor{$s$}
-
-		\If{$f(s^\prime) < f(s)$}{
-
-			$s \leftarrow s^\prime$\\
-			$improvement \leftarrow$ \texttt{true} \\
+			$m_w \leftarrow m_t$\\
 
 		}
+		\tcp{Check restart criterion}
+		\If{$f(m_b) - f(m_w) > f(m_b) \times \xi$}{
+
+			$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
+			$m_w \leftarrow$ \RandInit{}\\
+
 		}
-	\BlankLine
-	\KwRet ($s$)
 
-	\caption{Local Search}\label{alg:LS}
+	$G \leftarrow G + 1$
+}
+\caption{DILS}\label{alg:DILS}
 \end{algorithm}
 
-With the local search defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$.
-Figure \ref{img:LS} shows a pictorical representation of the above.
+For the mutation and crossover operators that appear in the Algorithm \ref{alg:DILS}, we use the uniform crossover operator \cite{spears1995virtues} and the segment mutation operator \textcolor{red}{REFERENCIA?} respectively.
+This type of mutation consists of selecting a fixed size segment from the features of the individual and replacing it with a random segment of the same size.
+The size of the segment must be fixed at the beginning of the optimization process.
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[scale=0.25]{Figures/LS.png}
-	\caption{Local optimum achieved by a local search procedure}\label{img:LS}
-\end{figure}
+Cabe destacar que el proceso de optimización descrito hasta el momento es aplicable a cualquier problema de minimización, aunque es claro que axiste una adaptación directa de DILS para problemas de maximización.
+Además el proceso de búsqueda local empleado para obtener el individuo trial debe ser especificado para cada problema, así como cuestiones de representación numérica y las criterios de parada.
 
-It is therefore necessary to appeal to techniques that introduce diversity in the search to avoid the problem of local optimal. This is exactly what ILS is trying to do.
+It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems.
+In addition, the local search procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation matters and termination criteria.
 
-With ILS, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^\prime$ that will not be a local optimum and will be far enough away from $s^*$.
-After that, we will apply a local search procedure to find a new local optimal $s^{*\prime}$.
-Figure \ref{img:ILS} shows a representation of the above concepts.
+\section{DILS optimization scheme for constrained clustering}
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[scale=0.25]{Figures/ILS.png}
-	\caption{No se como llamar a esta puta imagen}\label{img:ILS}
-\end{figure}
+In this section, we present an optimization scheme for the DILS algorithm when applied to constrained clustering.
+We will call this new approach DILS$_{CC}$.
+To begin with, we will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---being $n$ the number of instances of the dataset.
+Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, si each individual $m$ is a solution to the clustering problem.
+Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to intially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.
 
-It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^\prime$ is $s^*$ and therefore the solution space will not be explored effectively.
-On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.
+Regarding the fitness function $f$, for each individual $m$ its fitness value $f_m$ can be calculated as in Equation \ref{eq3}.
 
-The method for generating the perturbation applied to the local optimum $s^*$ may vary depending on the problem, and may be based on previous solutions.
-In this way, it is possible to keep a history of the solutions reached so far, so they can be used when applying the perturbation.
-Algorithm \ref{alg:ILS} summarizes the ILS overall optimization process.
+\begin{equation}
+f_m = z_m * (\text{infeasibility}_m + 1),
+\label{eq3}
+\end{equation}
+
+\noindent where $\text{infeasibility}_m$ is the number of non-satisfied constraints and $z_m$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq4}.
+
+\begin{equation}
+z_m = \sum_{c_i \in C_m} \left[ \frac{\sum_{x_a, x_b \in c_i} d^2(x_a,x_b)}{|c_i|}\right].
+\label{eq4}
+\end{equation}
+
+We also need to specify a local search procedure for DILS to be able to improve the mutant individual to obtain the trial vector $m_t$.
+Algorithm \ref{alg:LS} shows the local search precedure used in DILS$_{CC}$.
+Note that, in order to not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
+We also stablish a maximum number of neighbors $max\_neighbors$ that can be generated, this is done to find a better exploration-exploitation balance.
 
@@
 	\setstretch{1.2}
-	\SetKwFunction{LocalSearch}{LocalSearch}
-	\SetKwFunction{Perturbation}{Perturbation}
+	\SetKwFunction{RandomShuffle}{RandomShuffle}
+	\SetKwFunction{Rand}{Rand}
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Initial solution $s$}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $max\_neighbors$.}
 	\BlankLine
-	$s^* \leftarrow$ \LocalSearch{$s$}\\
-	\While{Termination criteria are not met}{
-
-		$s^\prime \leftarrow $ \Perturbation{$s^*$, $history$}\\
-		$s^{*\prime} \leftarrow$ \LocalSearch{$s^\prime$}
-
-		\If{$f(s^{*\prime}) < f(s^*)$}{
-
-			$s^* \leftarrow s^{*\prime}$
-
+	\While{$improvement$ {\bf and} $generated < max\_neighbors$}{
+		$improvement \leftarrow$ \texttt{false} \\
+		\tcp{Select random index (object) from $m$}
+		$i \leftarrow $ \Rand{$1$,$n$} \\
+		\tcp{Random shuffle labels set}
+		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
+		\For{$l \in RSL$}{
+			$m^\prime \leftarrow m$\\
+			\tcc{Move object $i$ from $m$ to the cluster associated with label $l$}
+			$m^\prime_i \leftarrow l$\\
+			\If{$f(m^\prime) < f(m)$}{
+				$m \leftarrow m^\prime$\\
+				$improvement \leftarrow$ \texttt{true} \\
+			}
+			$generated \leftarrow generated + 1$\\
 		}
@@
 	\BlankLine
-	\KwRet ($s*$)
+	\KwRet ($m$)
 
-	\caption{Iterated Local Search}\label{alg:ILS}
+	\caption{Local Search}\label{alg:LS}
 \end{algorithm}
-
-In the ILS scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only use the difference in the cost of the solutions to be compared to choose between them.
-This is the most simple scheme and does not depends on previous solutions recorded in the history.
-
-\clearpage
-
-\section{The Dual Iterative Local Search Method}
 
@@
 
-%% CORREGIR DESDE AQUI HASTA EL FINAL
+\subsection{Validation of results} \label{sec:ValidtnMethod}
 
-\subsection{Calibration}
-
-Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA$_{CC}$ and SHADE$_{CC}$ algorithms.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
-			\hline
-			Parameter & Meaning & BRKGA & SHADE \\
-			\hline
-			$|P|$ & Population size & 100 & 100 \\
-			Evals & Fitness function evaluations & 300000 & 300000 \\
-			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
-			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
-			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
-			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets}).} \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Parameters setup used for BRKGA and SHADE.}
-	\label{tab:params}
-\end{table}
-
-In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
-
-To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} c}
-			\hline
-			Common Parameters & Meaning & Value \\
-			\hline
-			\texttt{max\_iter} & Maximum number of iterations & 300 \\
-			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}). \\
-			\hline
-			\hline
-			Specific Parameters & \multicolumn{2}{l}{Name and Value} \\
-			\hline
-			COPKM & \multicolumn{2}{l}{\texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''}} \\
-			CECM & \multicolumn{2}{l}{$\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''}} \\
-			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
-			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
-			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Parameters setup used for the state-of-the-art algorithms.}
-	\label{tab:paramsSOTA}
-\end{table}
-
-Parameter values have been assigned following the guidelines of the original creators of the different proposals.
-Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible.
-Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible.
-Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
-
-\section{Experimental Results} \label{sec:results}
-
-In this section we present Tables from \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.
-
-Since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
-To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
-
-In this section we present a brief comparison between the proposed SHADE$_{CC}$ method and some of the state-of-the-art methods.
-Tables from \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
-To shorten the presented results this time we will only compare the ARI.
-
-It should be noted that there are some missing results in these tables.
-In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
-It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
-In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & 0.611 & 0.022 & -1.000 & 0.335 & 0.316 & 0.025 & -0.005 \\
-			Breast Cancer & 0.755 & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.870 & 0.281 & -1.000 & -0.005 & -0.005 & -0.004 & -0.011 \\
-			Circles & 0.798 & 0.166 & -1.000 & -0.003 & 0.162 & 0.137 & 0.133 \\
-			Ecoli & 0.069 & 0.009 & -1.000 & 0.387 & 0.417 & 0.265 & -1.000 \\
-			Glass & 0.009 & 0.016 & 0.184 & 0.268 & 0.197 & 0.211 & -1.000 \\
-			Haberman & 0.638 & 0.047 & -1.000 & -0.002 & 0.127 & 0.218 & -0.004 \\
-			Hayesroth & 0.031 & 0.027 & -1.000 & 0.106 & 0.097 & 0.054 & 0.139 \\
-			Heart & 0.793 & 0.483 & -1.000 & 0.027 & 0.036 & 0.423 & -0.003 \\
-			Ionosphere & 0.792 & 0.125 & -1.000 & 0.168 & 0.197 & 0.000 & 0.030 \\
-			Iris & 0.621 & 0.238 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
-			Led7Digit & 0.016 & 0.009 & 0.497 & 0.425 & 0.369 & 0.316 & -1.000 \\
-			Monk2 & 0.815 & 0.421 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
-			Moons & 0.968 & 0.277 & -1.000 & 0.241 & 0.319 & 0.785 & 0.092 \\
-			Movement Libras & 0.022 & 0.004 & 0.285 & 0.293 & 0.256 & 0.000 & -1.000 \\
-			Newthyroid & 0.028 & 0.017 & -1.000 & 0.579 & 0.289 & 0.846 & 0.002 \\
-			Saheart & 0.800 & 0.147 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
-			Sonar & 0.649 & 0.124 & -1.000 & 0.004 & 0.013 & 0.000 & 0.000 \\
-			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
-			Spectfheart & 0.896 & 0.334 & -1.000 & -0.107 & -0.114 & 0.000 & 0.050 \\
-			Spiral & 0.856 & 0.058 & -1.000 & -0.003 & 0.012 & 0.034 & -0.002 \\
-			Tae & 0.031 & 0.018 & -1.000 & 0.009 & -0.000 & 0.067 & 0.000 \\
-			Vowel & 0.002 & 0.003 & -1.000 & 0.063 & -0.003 & 0.067 & -1.000 \\
-			Wine & 0.334 & 0.120 & -1.000 & 0.360 & 0.368 & 0.281 & 0.004 \\
-			Zoo & 0.151 & 0.105 & 0.715 & 0.666 & 0.412 & 0.335 & -1.000 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
-	\label{tab:results10SOTA}
-\end{table}
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & 0.957 & 0.296 & -1.000 & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Breast Cancer & 0.788 & 0.772 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
-			Bupa & 0.984 & 0.840 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
-			Circles & 1.000 & 0.828 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Ecoli & 0.134 & 0.015 & -1.000 & 0.387 & 0.372 & 0.719 & -1.000 \\
-			Glass & 0.057 & 0.017 & -1.000 & 0.280 & 0.253 & 0.229 & -1.000 \\
-			Haberman & 1.000 & 0.866 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
-			Hayesroth & 0.386 & 0.107 & -1.000 & 0.106 & 0.107 & 0.150 & 0.135 \\
-			Heart & 1.000 & 0.692 & 1.000 & 0.025 & 0.033 & 0.442 & 0.000 \\
-			Ionosphere & 0.970 & 0.803 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
-			Iris & 0.850 & 0.412 & -1.000 & 0.730 & 0.547 & 0.421 & 0.684 \\
-			Led7Digit & 0.020 & 0.004 & -1.000 & 0.425 & 0.490 & 0.324 & -1.000 \\
-			Monk2 & 0.902 & 0.751 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
-			Moons & 1.000 & 0.958 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
-			Movement Libras & 0.025 & 0.006 & -0.742 & 0.293 & 0.255 & 0.000 & -1.000 \\
-			Newthyroid & 0.440 & 0.320 & -1.000 & 0.568 & 0.356 & 0.923 & 0.052 \\
-			Saheart & 0.862 & 0.749 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
-			Sonar & 0.981 & 0.800 & -1.000 & 0.004 & 0.019 & 0.000 & 0.000 \\
-			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
-			Spectfheart & 1.000 & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
-			Spiral & 1.000 & 0.847 & -1.000 & -0.003 & 0.014 & 0.006 & 0.051 \\
-			Tae & 0.299 & 0.119 & -1.000 & 0.008 & -0.004 & 0.062 & 0.000 \\
-			Vowel & 0.003 & 0.002 & -1.000 & 0.063 & -0.003 & 0.073 & -1.000 \\
-			Wine & 0.725 & 0.528 & -1.000 & 0.360 & 0.367 & 0.348 & -0.001 \\
-			Zoo & 0.185 & 0.157 & 0.435 & 0.642 & 0.450 & 0.353 & -1.000 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-		\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
-		\label{tab:results15SOTA}
-	\end{table}
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & 1.000 & 1.000 & -1.000 & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Breast Cancer & 0.807 & 0.807 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.991 & 0.936 & 1.000 & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & 1.000 & 0.935 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
-			Ecoli & 0.201 & 0.028 & -1.000 & 0.387 & 0.459 & 0.763 & -1.000 \\
-			Glass & 0.322 & 0.038 & -1.000 & 0.271 & 0.287 & 0.218 & -1.000 \\
-			Haberman & 1.000 & 0.943 & 1.000 & -0.002 & 0.106 & 0.737 & 0.000 \\
-			Hayesroth & 0.812 & 0.548 & -1.000 & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Heart & 1.000 & 0.942 & 1.000 & 0.025 & 0.033 & 1.000 & 0.000 \\
-			Ionosphere & 0.982 & 0.902 & 1.000 & 0.173 & 0.309 & 0.000 & 0.115 \\
-			Iris & 0.968 & 0.691 & -1.000 & 0.708 & 0.540 & 0.585 & 0.684 \\
-			Led7Digit & 0.017 & 0.007 & -1.000 & 0.425 & 0.571 & 0.327 & -1.000 \\
-			Monk2 & 0.904 & 0.855 & 1.000 & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & 1.000 & 0.937 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
-			Movement Libras & 0.018 & 0.006 & -1.000 & 0.293 & 0.293 & 0.000 & -1.000 \\
-			Newthyroid & 0.842 & 0.848 & -1.000 & 0.568 & 0.393 & 0.933 & 0.004 \\
-			Saheart & 0.881 & 0.881 & 1.000 & 0.018 & 0.026 & 1.000 & -0.006 \\
-			Sonar & 1.000 & 0.992 & 1.000 & 0.004 & 0.127 & 0.000 & 0.003 \\
-			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & 0.631 & 0.000 & -0.016 \\
-			Spectfheart & 1.000 & 0.953 & 1.000 & -0.107 & -0.112 & 0.000 & -0.054 \\
-			Spiral & 1.000 & 0.921 & 1.000 & -0.003 & 0.011 & 0.006 & 0.045 \\
-			Tae & 0.892 & 0.351 & -1.000 & 0.008 & 0.000 & 0.035 & 0.000 \\
-			Vowel & 0.003 & 0.002 & -1.000 & 0.063 & -0.003 & 0.071 & -1.000 \\
-			Wine & 0.916 & 0.937 & -1.000 & 0.360 & 0.363 & 0.380 & 0.015 \\
-			Zoo & 0.203 & 0.176 & 0.821 & 0.642 & 0.439 & 0.335 & -1.000 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-		\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
-		\label{tab:results25SOTA}
-	\end{table}
-
-\clearpage
-
-\section{Statistical Analysis of Results} \label{sec:analisis}
-
-With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
-This way we can statistically determine whether SHADE$_{CC}$ represents a significant improvement over previous proposals.
-
-In order to carry out this study we will use Bayesian statistical tests, instead of the classic Hull Hypothesis Statistical Tests (NHST).
+In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic Null Hypothesis Statistical Tests (NHST).
 In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
@@
 
-From now on, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA$_{CC}$ as sample $A$, and to the results obtained with SHADE$_{CC}$ as sample $B$.
-
-The Bayesian sign test is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
+We will use the Bayesian sign test to validate our results, which is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
 To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
@@
 
-One of the major advantages of the Bayesian sign test over the null hypothesis tests such as the (Wilcoxon's signed-rank test) is that we can obtain a visual representation of its results.
+\subsection{Calibration}
+
+Table \ref{tab:params} shows a summary of the parameter setup used for DILS$_{CC}$.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c m{6cm} c}
+			\hline
+			Parameter & Meaning & Value \\
+			\hline
+			Evals & Fitness function evaluations & 300000 \\
+			$segment\_size$ & Segment size for the mutation operator & $0.3 \times n$ \\
+			$max\_neighbors$ & Maximum number of neighbors generated in the local search procedure & Evals $\times 0.01$ \\
+			$p_\text{inherit}$ & Probability that a feature is inherited from $m_b$ & $0.3$ \\
+			$\xi$ & Reinitialization method tolerance & $0.2$ \\
+			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Parameters setup used for DILS$_{CC}$.}
+	\label{tab:params}
+\end{table}
+
+The stop criterion is given by the number of evaluations of the fitness function $f$, which at most will be 300000.
+
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
+In all cases the value for $k$ is equal to the number of classes displayed in Table \ref{tab:datasets}.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{cl}
+			\hline
+			Method name & Parameters name and values \\
+			\hline
+			BRKGA & Evals = $300000$; $|P| = 100$; $P_e = P_m = 0.2 * |P|$; $p_{\text{inherit}} = 50\%$ \\
+			COPKM & \texttt{max\_iter} = $300$; \texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''} \\
+			CECM & \texttt{max\_iter} = $300$; $\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''} \\
+			LCVQE & \texttt{max\_iter} = $300$; \texttt{initial\_centroids} = $\emptyset$ \\
+			RDPM & \texttt{max\_iter} = $300$; $\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset. \\
+			TVClust & \texttt{max\_iter} = $300$; $\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$ \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Parameters setup used for the state-of-the-art algorithms.}
+	\label{tab:paramsSOTA}
+\end{table}
+
+Parameter values have been assigned following the guidelines of the original creators of the different proposals.
+Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible.
+Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible.
+Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
+
+\section{Experimental Results} \label{sec:results}
+
+In this section we present Tables from \ref{tab:results10} to \ref{tab:results20}, which display the results obtained by DILS$_{CC}$ and six state-of-the-art methods to be compared for each dataset and constraint set.
+
+Since some of the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
+To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+
+It should be noted that there are some missing results in these tables.
+In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
+It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
+In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
+
+Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
+We can note that DILS$_{CC}$ represents a consistent improvement over BRKGA$_{CC}$ in both Unsat and Time, while in terms of ARI both methods are similar.
+This is because the penalty applied to the fitness function of both methods has little presence when operating with a small constraint set, and since they are population-based and heuristic-guided methods, they have similar exploration and exploitation capabilities.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lccccccc}
+			\hline
+			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			\hline
+			Appendicitis & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
+			Breast Cancer & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
+			Circles & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
+			Ecoli & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
+			Glass & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
+			Haberman & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
+			Hayesroth & 0.031 & 0.027 & - & \textbf{0.106} & 0.097 & 0.054 & 0.139 \\
+			Heart & \textbf{0.793} & 0.483 & - & 0.027 & 0.036 & 0.423 & -0.003 \\
+			Ionosphere & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
+			Iris & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
+			Led7Digit & 0.016 & 0.009 & \textbf{0.497} & 0.425 & 0.369 & 0.316 & - \\
+			Monk2 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
+			Moons & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
+			Movement Libras & 0.022 & 0.004 & 0.285 & \textbf{0.293} & 0.256 & 0.000 & - \\
+			Newthyroid & 0.028 & 0.017 & - & 0.579 & 0.289 & \textbf{0.846} & 0.002 \\
+			Saheart & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
+			Sonar & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
+			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
+			Spectfheart & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
+			Spiral & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
+			Tae & \textbf{0.031} & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
+			Vowel & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
+			Wine & 0.334 & 0.120 & - & 0.360 & \textbf{0.368} & 0.281 & 0.004 \\
+			Zoo & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
+			\hline
+			Mean & \textbf{0.474} & 0.158 & -0.510 & 0.226 & 0.212 & 0.175 & -0.186 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+	\caption{Experimental results obtained for $CS_{10}$ comparing DILS$_{CC}$ and the state-of-the-art.}
+	\label{tab:results10}
+\end{table}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lccccccc}
+			\hline
+			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			\hline
+			Appendicitis & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
+			Breast Cancer & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Bupa & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
+			Circles & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
+			Ecoli & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
+			Glass & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
+			Haberman & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
+			Hayesroth & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
+			Heart & \textbf{1.000} & 0.692 & \textbf{1.000} & 0.025 & 0.033 & 0.442 & 0.000 \\
+			Ionosphere & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
+			Iris & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
+			Led7Digit & 0.020 & 0.004 & - & 0.425 & \textbf{0.490} & 0.324 & - \\
+			Monk2 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
+			Moons & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Movement Libras & 0.025 & 0.006 & -0.742 & \textbf{0.293} & 0.255 & 0.000 & - \\
+			Newthyroid & 0.440 & 0.320 & - & 0.568 & 0.356 & \textbf{0.923} & 0.052 \\
+			Saheart & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
+			Sonar & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
+			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
+			Spectfheart & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
+			Spiral & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
+			Tae & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
+			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
+			Wine & \textbf{0.725} & 0.528 & - & 0.360 & 0.367 & 0.348 & -0.001 \\
+			Zoo & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
+			\hline
+			Mean & \textbf{0.641} & 0.490 & -0.070 & 0.225 & 0.233 & 0.258 & -0.196 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+		\caption{Experimental results obtained for $CS_{15}$ comparing DILS$_{CC}$ and the state-of-the-art.}
+		\label{tab:results15}
+	\end{table}
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lccccccc}
+			\hline
+			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			\hline
+			Appendicitis & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
+			Breast Cancer & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
+			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
+			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
+			Glass & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
+			Haberman & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
+			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
+			Heart & \textbf{1.000} & 0.942 & \textbf{1.000} & 0.025 & 0.033 & 1.000 & 0.000 \\
+			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
+			Iris & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
+			Led7Digit & 0.017 & 0.007 & - & 0.425 & \textbf{0.571} & 0.327 & - \\
+			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
+			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
+			Movement Libras & 0.018 & 0.006 & - & \textbf{0.293} & \textbf{0.293} & 0.000 & - \\
+			Newthyroid & 0.842 & 0.848 & - & 0.568 & 0.393 & \textbf{0.933} & 0.004 \\
+			Saheart & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
+			Sonar & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
+			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
+			Spectfheart & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
+			Spiral & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
+			Tae & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
+			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
+			Wine & 0.916 & \textbf{0.937} & - & 0.360 & 0.363 & 0.380 & 0.015 \\
+			Zoo & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
+			\hline
+			Mean & \textbf{0.733} & 0.642 & 0.064 & 0.223 & 0.285 & 0.344 & -0.187 \\
+			\hline
+
+		\end{tabular}}
+		%\end{adjustwidth}
+
+		\caption{Experimental results obtained for $CS_{20}$ comparing DILS$_{CC}$ and the state-of-the-art.}
+		\label{tab:results20}
+	\end{table}
+
+\clearpage
+
+\section{Statistical Analysis of Results} \label{sec:analisis}
+
+With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
+This way we can statistically determine whether DILS$_{CC}$ represents a significant improvement over previous proposals.
+
+Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with DILS$_{CC}$ as sample $B$.
+
+One of the major advantages of the Bayesian sign test is that we can obtain a very illustrative visual representation of its results.
 We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsBRKGA.pdf}
-		\caption{ARI}
+		\caption{BRKGA}
 		\label{fig:DILSvsBRKGA}
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsCOPKM.pdf}
-		\caption{Time}
+		\caption{COPKM}
 		\label{fig:DILSvsCOPKM}
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsLCVQE.pdf}
-		\caption{Unsat}
+		\caption{LCVQE}
 		\label{fig:DILSvsLCVQE}
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsTVClust.pdf}
-		\caption{Unsat}
+		\caption{TVClust}
 		\label{fig:DILSvsTVClust}
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsRDPM.pdf}
-		\caption{Unsat}
+		\caption{RDPM}
 		\label{fig:DILSvsRDPM}
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsCEKM.pdf}
-		\caption{Unsat}
+		\caption{CEKM}
 		\label{fig:DILSvsCEKM}

### 2019-06-11 Papers/SHADE/Sometido/paperSHADE_CC.tex
SHADE terminado y DILS avanzado

--- 2019-04-10 Papers/SHADE/paperSHADE.tex
@@
 Clustering has always been a powerful tool in knowledge discovery.
-Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning.
-This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose the first application of Differential Evolution to the constrained clustering problem, which is a generalization of traditional clustering that considers additional information in the form of constraints.
+Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering.
+This technique is a generalization of traditional clustering that considers additional information encoded by constraints.
+Constraints can be given in the form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
+We propose the first application of Differential Evolution to the constrained clustering problem, which has proven to produce a better exploration-exploitation trade-off when comparing with previous approaches.
 We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
@@
 Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
-In Section \ref{sec:brkga} we will briefly review the BRKGA algorithm and its adaptation for constrained clustering.
+In Section \ref{sec:brkga} the BRKGA algorithm and its adaptation for constrained clustering will be briefly reviewed.
 Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
-In Section \ref{sec:SHADEadapt} we present the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function.
+In Section \ref{sec:SHADEadapt} the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function, will be presented.
 Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
-Finally, in Section \ref{sec:conclusiones} we discuss conclusions.
+Finally, in Section \ref{sec:conclusiones} conclusions are discussed.
 
@@
 
-Constrained clustering has many applications and has been widely studied in the literature.
 The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}.
@@
 
-Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,v]}\}$ that serve as parameters for the function to be optimized.
+Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,v]}\}$, with $v$ being the number of features of each individual.
+Each $p_i$ serves as parameter for the function to be optimized.
 These individuals are considered solutions to the problem.
@@
 	\centering
-	\includegraphics[scale=0.3]{Figures/DEop.png}
+	\includegraphics[scale=0.3]{DEop.png}
 	\caption{The optimization process of DE \cite{das2011differential}.}\label{img:DE}
@@
 Genetic Algorithms (GAs) emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
-It is one of the most studied and successful evolutionary algorithms, thanks in part to its simplicity and ease of implementation \cite{goldberg1989genetic, fausto2019ants}.
+It is one of the most studied and successful evolutionary algorithms, thanks in part to its simplicity and ease of implementation \cite{fausto2019ants, goldberg1989genetic}.
 
@@
 The mutation operator is used by GAs to avoid local optima, and it consists in introducing random changes into the population to favor diversity.
-The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function whose task is to assign a value to each chromosome to assess its quality as a solution for the problem.
+The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function that assigns a value to each chromosome to assess its quality as a solution for the problem.
 GAs apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal \cite{fausto2019ants}.
-
-%GAs have been applied to all kinds of problems, including problems whose solutions are subject to strong restrictions such as machine scheduling, resource location and the quadratic assignment problem.
-These problems present the peculiarity that, by applying generalized crossover operators to two individuals that represent a feasible solution to the problem, we can obtain non-feasible ones.
 
@@
 Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated from the dataset or constraints, whereas in Equation \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.
+Therefore with Equation \eqref{eq14} a better generalization capability is obtained.
 
@@
 	\myfloatalign
-	{\includegraphics[width=.45\linewidth]{Figures/Dataset}} \quad
+	{\includegraphics[width=.45\linewidth]{Dataset}} \quad
 	%\subfloat[Imagen original]
-	{\includegraphics[width=.45\linewidth]{Figures/C0}} \quad
+	{\includegraphics[width=.45\linewidth]{C0}} \quad
 	%\subfloat[Clustering sin restricciones]
-	{\includegraphics[width=.45\linewidth]{Figures/C1}} \quad
+	{\includegraphics[width=.45\linewidth]{C1}} \quad
 	%\subfloat[Clustering con restricciones]
-	{\includegraphics[width=.45\linewidth]{Figures/C2}} \quad
+	{\includegraphics[width=.45\linewidth]{C2}} \quad
 	\caption{Three partitions over a toy dataset.}
@@
 As in \cite{de2017comparison} we apply a local optimization procedure to the individuals of the population, but without transferring its results to the original individual in order to maintain diversity.
-The difference is that we do not apply it to the whole population but only to a number $P_e$ of individuals considered as the elite of the population.
+The difference is that we do not apply it to the whole population but only to a number $P_e$ of individuals considered as the elite of the population, with the aim of preserving a good exploration-exploitation trade-off.
 The percentage of individuals considered as elite must be determined for each case.
 
-Algorithm \ref{alg:SHADE} describes the overall SHADE$_{CC}$ optimization process, on the basis of the one that can be found in \cite{tanabe2013success} to include the local search procedure presented in Algorithm \ref{alg:LS}.
+Algorithm \ref{alg:SHADE} describes the overall SHADE$_{CC}$ optimization process, including the local search procedure presented in Algorithm \ref{alg:LS}.
 The goal of this procedure is to locally improve solutions (individuals from $P$) in a non-exhaustive way.
@@
 		}
-		$G++$
+		$G \leftarrow G + 1$
 	}
@@
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at GitHub.
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/SHADE_CC}{GitHub}.
 Table \ref{tab:datasets} displays a summary of every dataset.
@@
 
-Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
+Classification datasets are commonly used in the literature to test constrained clustering algorithms; they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
 They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).
-
-\clearpage
 
@@
 
-Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
+Note that the greater the number of classes in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
 This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.
@@
 
-Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method.
+Since we have the true labels associated to each of the datasets, we can use them to evaluate the results provided by each method.
 We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
@@
 
-%% CORREGIR DESDE AQUI HASTA EL FINAL
+\subsection{Validation of results} \label{sec:ValidtnMethod}
+
+In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic Null Hypothesis Statistical Tests (NHST).
+In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
+\textit{"In a nutshell: NHST do not answer the question we ask"}.
+To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
+
+To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
+Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
+This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.
+
+Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate.
+As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
+
+Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
+That is: What can we conclude when NHST do not reject the null hypothesis?
+We can not infer anything since NHST can not provide evidence in its favor.
+
+Finally, there are two other problems that researchers face when performing NHST.
+The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results.
+The second one is the need to previously formalize the intentions of the sampling of the results, which are usually fixed a posteriori; this could lead to a misreading of said results.
+
+As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST.
+In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test.
+To make use of it we will employ the R package \texttt{rNPBST}, whose documentation and guide can be found in \cite{carrasco2017rnpbst}.
+
+We will use the Bayesian sign test to validate our results, which is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
+To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
+In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$.
+Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
+$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$
 
@@
 			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
-			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets}).} \\
+			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
 			\hline
@@
 In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+A Python implementation for both algorithms can be found at \href{https://github.com/GermangUgr/SHADE\_CC}{GitHub}
 
@@
 			\texttt{max\_iter} & Maximum number of iterations & 300 \\
-			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}). \\
+			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
 			\hline
@@
 			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
-			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset.} \\
+			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset} \\
 			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
@@
 	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
+	\renewcommand{\arraystretch}{1.2}
 	%\begin{adjustwidth}{-1in}{-1in}
@@
 			\hline
-			Appendicitis & 0.022 & 5.454 & 383.642 && 0.086 & 5.091 & 366.491 \\
-			Breast Cancer & 0.395 & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
-			Bupa & 0.281 & 14.151 & 1420.117 && 0.104 & 16.672 & 1409.487 \\
-			Circles & 0.166 & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
-			Ecoli & 0.009 & 20.321 & 1169.798 && 0.020 & 12.763 & 1112.981 \\
-			Glass & 0.016 & 8.658 & 735.507 && 0.013 & 4.848 & 694.962 \\
-			Haberman & 0.047 & 15.699 & 1155.985 && 0.084 & 15.441 & 1146.023 \\
-			Hayesroth & 0.027 & 1.500 & 550.117 && 0.025 & 2.000 & 529.424 \\
-			Ionosphere & 0.125 & 17.587 & 1894.390 && 0.332 & 12.603 & 1872.308 \\
-			Iris & 0.238 & 0.190 & 505.112 && 0.342 & 0.000 & 483.584 \\
-			Led7Digit & 0.009 & 11.347 & 1786.943 && 0.004 & 8.866 & 1784.708 \\
-			Monk2 & 0.421 & 14.926 & 1848.543 && 0.270 & 16.385 & 1821.208 \\
-			Moons & 0.277 & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
-			Movement Libras & 0.004 & 4.159 & 1508.279 && 0.003 & 3.396 & 1519.185 \\
-			Pima & 0.236 & 25.195 & 4072.087 && 0.275 & 22.468 & 3990.383 \\
-			Saheart & 0.147 & 20.555 & 2135.624 && 0.265 & 16.725 & 2095.554 \\
-			Sonar & 0.124 & 7.619 & 1106.039 && 0.116 & 8.952 & 1081.994 \\
-			Soybean & 0.496 & 0.000 & 176.333 && 0.515 & 0.000 & 151.163 \\
-			Spectfheart & 0.334 & 9.573 & 1451.235 && 0.343 & 8.604 & 1405.243 \\
-			Spiral & 0.058 & 15.402 & 1098.622 && 0.114 & 13.793 & 1099.747 \\
-			Tae & 0.018 & 0.833 & 520.855 && 0.024 & 1.333 & 508.988 \\
-			Vehicle & 0.005 & 28.515 & 4366.104 && 0.006 & 22.151 & 4352.663 \\
-			Vowel & 0.003 & 13.330 & 4245.672 && 0.002 & 11.523 & 4290.154 \\
-			Wdbc & 0.478 & 16.316 & 3596.517 && 0.357 & 17.093 & 3528.029 \\
-			Zoo & 0.105 & 1.091 & 336.345 && 0.098 & 2.182 & 317.674 \\
-			\hline
-			Mean & 0.161 & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
+			Appendicitis & 0.022 & 5.454 & 383.642 && \textbf{0.086} & 5.091 & 366.491 \\
+			Breast Cancer & \textbf{0.395} & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
+			Bupa & \textbf{0.281} & 14.151 & 1420.117 && 0.104 & 16.672 & 1409.487 \\
+			Circles & \textbf{0.166} & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
+			Ecoli & 0.009 & 20.321 & 1169.798 && \textbf{0.020} & 12.763 & 1112.981 \\
+			Glass & \textbf{0.016} & 8.658 & 735.507 && 0.013 & 4.848 & 694.962 \\
+			Haberman & 0.047 & 15.699 & 1155.985 && \textbf{0.084} & 15.441 & 1146.023 \\
+			Hayesroth & \textbf{0.027} & 1.500 & 550.117 && 0.025 & 2.000 & 529.424 \\
+			Ionosphere & 0.125 & 17.587 & 1894.390 && \textbf{0.332} & 12.603 & 1872.308 \\
+			Iris & 0.238 & 0.190 & 505.112 && \textbf{0.342} & 0.000 & 483.584 \\
+			Led7Digit & \textbf{0.009} & 11.347 & 1786.943 && 0.004 & 8.866 & 1784.708 \\
+			Monk2 & \textbf{0.421} & 14.926 & 1848.543 && 0.270 & 16.385 & 1821.208 \\
+			Moons & \textbf{0.277} & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
+			Movement Libras & \textbf{0.004} & 4.159 & 1508.279 && 0.003 & 3.396 & 1519.185 \\
+			Pima & 0.236 & 25.195 & 4072.087 && \textbf{0.275} & 22.468 & 3990.383 \\
+			Saheart & 0.147 & 20.555 & 2135.624 && \textbf{0.265} & 16.725 & 2095.554 \\
+			Sonar & \textbf{0.124} & 7.619 & 1106.039 && 0.116 & 8.952 & 1081.994 \\
+			Soybean & 0.496 & 0.000 & 176.333 && \textbf{0.515} & 0.000 & 151.163 \\
+			Spectfheart & 0.334 & 9.573 & 1451.235 && \textbf{0.343} & 8.604 & 1405.243 \\
+			Spiral & 0.058 & 15.402 & 1098.622 && \textbf{0.114} & 13.793 & 1099.747 \\
+			Tae & 0.018 & 0.833 & 520.855 && \textbf{0.024} & 1.333 & 508.988 \\
+			Vehicle & 0.005 & 28.515 & 4366.104 && \textbf{0.006} & 22.151 & 4352.663 \\
+			Vowel & \textbf{0.003} & 13.330 & 4245.672 && 0.002 & 11.523 & 4290.154 \\
+			Wdbc & \textbf{0.478} & 16.316 & 3596.517 && 0.357 & 17.093 & 3528.029 \\
+			Zoo & \textbf{0.105} & 1.091 & 336.345 && 0.098 & 2.182 & 317.674 \\
+			\hline
+			Mean & \textbf{0.161} & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
 			\hline
@@
 			\hline
-			Appendicitis & 0.296 & 6.167 & 381.625 && 0.332 & 5.833 & 362.106 \\
-			Breast Cancer & 0.772 & 10.172 & 3854.911 && 0.911 & 3.310 & 3766.402 \\
-			Bupa & 0.840 & 5.973 & 1463.152 && 0.801 & 5.641 & 1437.639 \\
-			Circles & 0.828 & 6.101 & 1136.484 && 0.982 & 0.404 & 1108.928 \\
-			Ecoli & 0.015 & 24.314 & 1200.463 && 0.025 & 19.090 & 1127.568 \\
-			Glass & 0.017 & 17.500 & 745.987 && 0.031 & 11.743 & 722.009 \\
-			Haberman & 0.866 & 4.677 & 1209.471 && 0.796 & 5.314 & 1147.479 \\
-			Hayesroth & 0.107 & 7.536 & 545.861 && 0.099 & 8.261 & 520.702 \\
-			Ionosphere & 0.803 & 7.722 & 1986.543 && 0.591 & 11.974 & 1944.085 \\
-			Iris & 0.412 & 3.873 & 506.548 && 0.260 & 6.087 & 486.723 \\
-			Led7Digit & 0.004 & 12.836 & 1902.272 && 0.007 & 10.919 & 1897.511 \\
-			Monk2 & 0.751 & 10.827 & 1949.305 && 0.945 & 2.067 & 1920.361 \\
-			Moons & 0.958 & 1.334 & 1133.858 && 0.995 & 0.020 & 1113.687 \\
-			Movement Libras & 0.006 & 8.679 & 1584.028 && 0.001 & 8.889 & 1590.321 \\
-			Pima & 0.481 & 22.051 & 4451.193 && 0.762 & 9.928 & 4402.693 \\
-			Saheart & 0.749 & 10.940 & 2281.992 && 0.945 & 1.839 & 2225.210 \\
-			Sonar & 0.800 & 5.040 & 1124.700 && 0.890 & 2.177 & 1069.862 \\
-			Soybean & 0.428 & 0.000 & 170.674 && 0.422 & 0.000 & 155.948 \\
-			Spectfheart & 0.924 & 2.854 & 1561.040 && 0.778 & 5.732 & 1485.260 \\
-			Spiral & 0.847 & 5.677 & 1133.962 && 0.949 & 1.495 & 1102.920 \\
-			Tae & 0.119 & 7.115 & 522.088 && 0.066 & 8.380 & 513.289 \\
-			Vehicle & 0.089 & 28.044 & 4806.559 && 0.018 & 26.856 & 4749.812 \\
-			Vowel & 0.002 & 14.486 & 4807.165 && 0.001 & 13.341 & 4867.849 \\
-			Wdbc & 0.805 & 8.985 & 3829.407 && 0.930 & 2.801 & 3750.711 \\
-			Zoo & 0.157 & 3.500 & 333.984 && 0.128 & 2.834 & 315.131 \\
-			\hline
-			Mean & 0.483 & 9.456 & 1784.931 && 0.507 & 6.997 & 1751.368 \\
+			Appendicitis & 0.296 & 6.167 & 381.625 && \textbf{0.332} & 5.833 & 362.106 \\
+			Breast Cancer & 0.772 & 10.172 & 3854.911 && \textbf{0.911} & 3.310 & 3766.402 \\
+			Bupa & \textbf{0.840} & 5.973 & 1463.152 && 0.801 & 5.641 & 1437.639 \\
+			Circles & 0.828 & 6.101 & 1136.484 && \textbf{0.982} & 0.404 & 1108.928 \\
+			Ecoli & 0.015 & 24.314 & 1200.463 && \textbf{0.025} & 19.090 & 1127.568 \\
+			Glass & 0.017 & 17.500 & 745.987 && \textbf{0.031} & 11.743 & 722.009 \\
+			Haberman & \textbf{0.866} & 4.677 & 1209.471 && 0.796 & 5.314 & 1147.479 \\
+			Hayesroth & \textbf{0.107} & 7.536 & 545.861 && 0.099 & 8.261 & 520.702 \\
+			Ionosphere & \textbf{0.803} & 7.722 & 1986.543 && 0.591 & 11.974 & 1944.085 \\
+			Iris & \textbf{0.412} & 3.873 & 506.548 && 0.260 & 6.087 & 486.723 \\
+			Led7Digit & 0.004 & 12.836 & 1902.272 && \textbf{0.007} & 10.919 & 1897.511 \\
+			Monk2 & 0.751 & 10.827 & 1949.305 && \textbf{0.945} & 2.067 & 1920.361 \\
+			Moons & 0.958 & 1.334 & 1133.858 && \textbf{0.995} & 0.020 & 1113.687 \\
+			Movement Libras & \textbf{0.006} & 8.679 & 1584.028 && 0.001 & 8.889 & 1590.321 \\
+			Pima & 0.481 & 22.051 & 4451.193 && \textbf{0.762} & 9.928 & 4402.693 \\
+			Saheart & 0.749 & 10.940 & 2281.992 && \textbf{0.945} & 1.839 & 2225.210 \\
+			Sonar & 0.800 & 5.040 & 1124.700 && \textbf{0.890} & 2.177 & 1069.862 \\
+			Soybean & \textbf{0.428} & 0.000 & 170.674 && 0.422 & 0.000 & 155.948 \\
+			Spectfheart & \textbf{0.924} & 2.854 & 1561.040 && 0.778 & 5.732 & 1485.260 \\
+			Spiral & 0.847 & 5.677 & 1133.962 && \textbf{0.949} & 1.495 & 1102.920 \\
+			Tae & \textbf{0.119} & 7.115 & 522.088 && 0.066 & 8.380 & 513.289 \\
+			Vehicle & \textbf{0.089} & 28.044 & 4806.559 && 0.018 & 26.856 & 4749.812 \\
+			Vowel & \textbf{0.002} & 14.486 & 4807.165 && 0.001 & 13.341 & 4867.849 \\
+			Wdbc & 0.805 & 8.985 & 3829.407 && \textbf{0.930} & 2.801 & 3750.711 \\
+			Zoo & \textbf{0.157} & 3.500 & 333.984 && 0.128 & 2.834 & 315.131 \\
+			\hline
+			Mean & 0.483 & 9.456 & 1784.931 && \textbf{0.507} & 6.997 & 1751.368 \\
 			\hline
@@
 			\hline
-			Breast Cancer & 0.807 & 9.086 & 4061.162 && 0.977 & 0.891 & 3984.370 \\
-			Bupa & 0.936 & 2.975 & 1529.151 && 0.993 & 0.281 & 1508.772 \\
-			Circles & 0.935 & 2.814 & 1180.268 && 1.000 & 0.000 & 1170.452 \\
-			Ecoli & 0.028 & 25.716 & 1267.419 && 0.047 & 21.466 & 1277.903 \\
-			Glass & 0.038 & 23.610 & 766.861 && 0.082 & 16.412 & 796.162 \\
-			Haberman & 0.943 & 2.073 & 1266.131 && 0.997 & 0.138 & 1235.441 \\
-			Hayesroth & 0.548 & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
-			Ionosphere & 0.902 & 4.314 & 2051.251 && 0.993 & 0.258 & 2019.558 \\
-			Iris & 0.691 & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
-			Led7Digit & 0.007 & 13.653 & 2053.264 && 0.005 & 12.170 & 2066.223 \\
-			Monk2 & 0.855 & 6.025 & 2055.425 && 0.991 & 0.379 & 2013.468 \\
-			Moons & 0.937 & 2.689 & 1177.848 && 0.995 & 0.158 & 1157.222 \\
-			Movement Libras & 0.006 & 8.381 & 1654.159 && 0.004 & 7.676 & 1645.174 \\
-			Pima & 0.753 & 11.544 & 4855.446 && 0.899 & 4.419 & 4820.684 \\
-			Saheart & 0.881 & 5.250 & 2402.513 && 0.981 & 0.706 & 2353.931 \\
-			Sonar & 0.992 & 0.186 & 1138.769 && 1.000 & 0.000 & 1127.020 \\
-			Soybean & 0.416 & 0.000 & 171.281 && 0.563 & 0.000 & 153.406 \\
-			Spectfheart & 0.953 & 1.677 & 1602.229 && 1.000 & 0.000 & 1526.925 \\
-			Spiral & 0.921 & 3.604 & 1178.580 && 0.998 & 0.091 & 1155.194 \\
-			Tae & 0.351 & 9.807 & 527.292 && 0.392 & 9.419 & 508.610 \\
-			Vehicle & 0.144 & 28.781 & 5189.066 && 0.044 & 29.008 & 5284.288 \\
-			Vowel & 0.002 & 14.682 & 5450.331 && 0.003 & 13.766 & 5514.052 \\
-			Wdbc & 0.823 & 8.281 & 4019.683 && 0.972 & 1.068 & 3983.564 \\
-			Zoo & 0.176 & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
-			\hline
-			Mean & 0.602 & 8.086 & 1895.671 && 0.647 & 5.443 & 1880.921 \\
+			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
+			Bupa & 0.936 & 2.975 & 1529.151 && \textbf{0.993} & 0.281 & 1508.772 \\
+			Circles & 0.935 & 2.814 & 1180.268 && \textbf{1.000} & 0.000 & 1170.452 \\
+			Ecoli & 0.028 & 25.716 & 1267.419 && \textbf{0.047} & 21.466 & 1277.903 \\
+			Glass & 0.038 & 23.610 & 766.861 && \textbf{0.082} & 16.412 & 796.162 \\
+			Haberman & 0.943 & 2.073 & 1266.131 && \textbf{0.997} & 0.138 & 1235.441 \\
+			Hayesroth & \textbf{0.548} & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
+			Ionosphere & 0.902 & 4.314 & 2051.251 && \textbf{0.993} & 0.258 & 2019.558 \\
+			Iris & \textbf{0.691} & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
+			Led7Digit & \textbf{0.007} & 13.653 & 2053.264 && 0.005 & 12.170 & 2066.223 \\
+			Monk2 & 0.855 & 6.025 & 2055.425 && \textbf{0.991} & 0.379 & 2013.468 \\
+			Moons & 0.937 & 2.689 & 1177.848 && \textbf{0.995} & 0.158 & 1157.222 \\
+			Movement Libras & \textbf{0.006} & 8.381 & 1654.159 && 0.004 & 7.676 & 1645.174 \\
+			Pima & 0.753 & 11.544 & 4855.446 && \textbf{0.899} & 4.419 & 4820.684 \\
+			Saheart & 0.881 & 5.250 & 2402.513 && \textbf{0.981} & 0.706 & 2353.931 \\
+			Sonar & 0.992 & 0.186 & 1138.769 && \textbf{1.000} & 0.000 & 1127.020 \\
+			Soybean & 0.416 & 0.000 & 171.281 && \textbf{0.563} & 0.000 & 153.406 \\
+			Spectfheart & 0.953 & 1.677 & 1602.229 && \textbf{1.000} & 0.000 & 1526.925 \\
+			Spiral & 0.921 & 3.604 & 1178.580 && \textbf{0.998} & 0.091 & 1155.194 \\
+			Tae & 0.351 & 9.807 & 527.292 && \textbf{0.392} & 9.419 & 508.610 \\
+			Vehicle & \textbf{0.144} & 28.781 & 5189.066 && 0.044 & 29.008 & 5284.288 \\
+			Vowel & 0.002 & 14.682 & 5450.331 && \textbf{0.003} & 13.766 & 5514.052 \\
+			Wdbc & 0.823 & 8.281 & 4019.683 && \textbf{0.972} & 1.068 & 3983.564 \\
+			Zoo & \textbf{0.176} & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
+			\hline
+			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
 			\hline
@@
 			\hline
-			Appendicitis & 0.086 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
-			Breast Cancer & 0.159 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.104 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
-			Circles & 0.063 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
-			Ecoli & 0.020 & - & 0.387 & 0.417 & 0.265 & - \\
-			Glass & 0.013 & 0.184 & 0.268 & 0.197 & 0.211 & - \\
-			Haberman & 0.084 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
-			Hayesroth & 0.025 & - & 0.106 & 0.097 & 0.054 & 0.139 \\
-			Ionosphere & 0.332 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
-			Iris & 0.342 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
-			Led7Digit & 0.004 & 0.497 & 0.425 & 0.369 & 0.316 & - \\
-			Monk2 & 0.270 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
-			Moons & 0.272 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
-			Movement Libras & 0.003 & 0.285 & 0.293 & 0.256 & 0.000 & - \\
-			Pima & 0.275 & - & 0.076 & 0.075 & 0.145 & -0.008 \\
-			Saheart & 0.265 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
-			Sonar & 0.116 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
-			Soybean & 0.515 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
-			Spectfheart & 0.343 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
-			Spiral & 0.114 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
-			Tae & 0.024 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
-			Vehicle & 0.006 & - & 0.121 & 0.081 & 0.280 & - \\
-			Vowel & 0.002 & - & 0.063 & -0.003 & 0.067 & - \\
-			Wdbc & 0.357 & - & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Zoo & 0.098 & 0.715 & 0.666 & 0.412 & 0.335 & - \\
-			\hline
-			Mean  & 0.156 & -0.51 & 0.215 & 0.211 & 0.13 & -0.226 \\
+			Appendicitis & 0.086 & - & \textbf{0.335} & 0.316 & 0.025 & -0.005 \\
+			Breast Cancer & 0.159 & -0.604 & 0.486 & \textbf{0.502} & 0.000 & 0.000 \\
+			Bupa & \textbf{0.104} & - & -0.005 & -0.005 & -0.004 & -0.011 \\
+			Circles & 0.063 & - & -0.003 & \textbf{0.162} & 0.137 & 0.133 \\
+			Ecoli & 0.020 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
+			Glass & 0.013 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
+			Haberman & 0.084 & - & -0.002 & 0.127 & \textbf{0.218} & -0.004 \\
+			Hayesroth & 0.025 & - & 0.106 & 0.097 & 0.054 & \textbf{0.139} \\
+			Ionosphere & \textbf{0.332} & - & 0.168 & 0.197 & 0.000 & 0.030 \\
+			Iris & 0.342 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
+			Led7Digit & 0.004 & \textbf{0.497} & 0.425 & 0.369 & 0.316 & - \\
+			Monk2 & 0.270 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
+			Moons & 0.272 & - & 0.241 & 0.319 & \textbf{0.785} & 0.092 \\
+			Movement Libras & 0.003 & 0.285 & \textbf{0.293} & 0.256 & 0.000 & - \\
+			Pima & \textbf{0.275} & - & 0.076 & 0.075 & 0.145 & -0.008 \\
+			Saheart & 0.265 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
+			Sonar & \textbf{0.116} & - & 0.004 & 0.013 & 0.000 & 0.000 \\
+			Soybean & 0.515 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
+			Spectfheart & \textbf{0.343} & - & -0.107 & -0.114 & 0.000 & 0.050 \\
+			Spiral & \textbf{0.114} & - & -0.003 & 0.012 & 0.034 & -0.002 \\
+			Tae & 0.024 & - & 0.009 & -0.000 & \textbf{0.067} & 0.000 \\
+			Vehicle & 0.006 & - & 0.121 & 0.081 & \textbf{0.280} & - \\
+			Vowel & 0.002 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
+			Wdbc & 0.357 & - & 0.486 & \textbf{0.502} & 0.000 & 0.000 \\
+			Zoo & 0.098 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
+			\hline
+			Mean  & 0.156 & -0.51 & \textbf{0.215} & 0.211 & 0.13 & -0.226 \\
 			\hline
@@
 			\hline
-			Appendicitis & 0.332 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Breast Cancer & 0.911 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
-			Bupa & 0.801 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
-			Circles & 0.982 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Ecoli & 0.025 & - & 0.387 & 0.372 & 0.719 & - \\
+			Appendicitis & \textbf{0.332} & - & 0.305 & 0.284 & 0.025 & -0.006 \\
+			Breast Cancer & 0.911 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Bupa & 0.801 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
+			Circles & 0.982 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
+			Ecoli & 0.025 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
 			Glass & 0.031 & - & 0.280 & 0.253 & 0.229 & - \\
-			Haberman & 0.796 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
-			Hayesroth & 0.099 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
-			Ionosphere & 0.591 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
-			Iris & 0.260 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
-			Led7Digit & 0.007 & - & 0.425 & 0.490 & 0.324 & - \\
-			Monk2 & 0.945 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
-			Moons & 0.995 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
-			Movement Libras & 0.001 & -0.742 & 0.293 & 0.255 & 0.000 & - \\
-			Pima & 0.762 & - & 0.076 & 0.075 & 1.000 & 0.000 \\
-			Saheart & 0.945 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
-			Sonar & 0.890 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
-			Soybean & 0.422 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
-			Spectfheart & 0.778 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
-			Spiral & 0.949 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
-			Tae & 0.066 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
-			Vehicle & 0.018 & 1.000 & 0.121 & 0.081 & 0.576 & - \\
-			Vowel & 0.001 & - & 0.063 & -0.003 & 0.073 & - \\
-			Wdbc & 0.930 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Zoo & 0.128 & 0.435 & 0.642 & 0.450 & 0.353 & - \\
-			\hline
-			Mean  & 0.507 & 0.01 & 0.213 & 0.229 & 0.253 & -0.238 \\
+			Haberman & 0.796 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
+			Hayesroth & 0.099 & - & 0.106 & 0.107 & \textbf{0.150} & 0.135 \\
+			Ionosphere & 0.591 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
+			Iris & 0.260 & - & \textbf{0.730} & 0.547 & 0.421 & 0.684 \\
+			Led7Digit & 0.007 & - & 0.425 & \textbf{0.490} & 0.324 & - \\
+			Monk2 & 0.945 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
+			Moons & 0.995 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Movement Libras & 0.001 & -0.742 & \textbf{0.293} & 0.255 & 0.000 & - \\
+			Pima & \textbf{0.762} & - & 0.076 & 0.075 & 1.000 & 0.000 \\
+			Saheart & 0.945 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
+			Sonar & \textbf{0.890} & - & 0.004 & 0.019 & 0.000 & 0.000 \\
+			Soybean & 0.422 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
+			Spectfheart & 0.778 & \textbf{0.983} & -0.107 & -0.117 & 0.000 & -0.070 \\
+			Spiral & \textbf{0.949} & - & -0.003 & 0.014 & 0.006 & 0.051 \\
+			Tae & \textbf{0.066} & - & 0.008 & -0.004 & 0.062 & 0.000 \\
+			Vehicle & 0.018 & \textbf{1.000} & 0.121 & 0.081 & 0.576 & - \\
+			Vowel & 0.001 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
+			Wdbc & 0.930 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Zoo & 0.128 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
+			\hline
+			Mean  & \textbf{0.507} & 0.01 & 0.213 & 0.229 & 0.253 & -0.238 \\
 			\hline
@@
 
-In one hand, table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE$_{CC}$.
+In one hand, Table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE$_{CC}$.
 On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by SHADE$_{CC}$, with the exception of COPKM, which is able to reach the optimum for some of the cases in which it finds a solution.
@@
 			\hline
-			Appendicitis & 1.000 & - & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Breast Cancer & 0.977 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.993 & 1.000 & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & 1.000 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
-			Ecoli & 0.047 & - & 0.387 & 0.459 & 0.763 & - \\
-			Glass & 0.082 & - & 0.271 & 0.287 & 0.218 & - \\
-			Haberman & 0.997 & 1.000 & -0.002 & 0.106 & 0.737 & 0.000 \\
-			Hayesroth & 0.531 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Ionosphere & 0.993 & 1.000 & 0.173 & 0.309 & 0.000 & 0.115 \\
-			Iris & 0.553 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
-			Led7Digit & 0.005 & - & 0.425 & 0.571 & 0.327 & - \\
-			Monk2 & 0.991 & 1.000 & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & 0.995 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
-			Movement Libras & 0.004 & - & 0.293 & 0.293 & 0.000 & - \\
-			Pima & 0.899 & 1.000 & 0.076 & 0.076 & 1.000 & -0.005 \\
-			Saheart & 0.981 & 1.000 & 0.018 & 0.026 & 1.000 & -0.006 \\
-			Sonar & 1.000 & 1.000 & 0.004 & 0.127 & 0.000 & 0.003 \\
-			Soybean & 0.563 & -0.218 & 0.545 & 0.631 & 0.000 & -0.016 \\
-			Spectfheart & 1.000 & 1.000 & -0.107 & -0.112 & 0.000 & -0.054 \\
-			Spiral & 0.998 & 1.000 & -0.003 & 0.011 & 0.006 & 0.045 \\
-			Tae & 0.392 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
-			Vehicle & 0.044 & 1.000 & 0.122 & 0.081 & 0.689 & - \\
-			Vowel & 0.003 & - & 0.063 & -0.003 & 0.071 & - \\
-			Wdbc & 0.972 & 1.000 & 0.486 & 0.522 & 0.000 & 0.000 \\
-			Zoo & 0.153 & 0.821 & 0.642 & 0.439 & 0.335 & - \\
-			\hline
-			Mean  & 0.647 & 0.224 & 0.212 & 0.28 & 0.319 & -0.228 \\
+			Appendicitis & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
+			Breast Cancer & 0.977 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.993 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
+			Circles & \textbf{1.000} & \textbf{1.000} & -0.003 & 0.629 & \textbf{1.000} & 0.138 \\
+			Ecoli & 0.047 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
+			Glass & 0.082 & - & 0.271 & \textbf{0.287} & 0.218 & - \\
+			Haberman & 0.997 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
+			Hayesroth & \textbf{0.531} & - & 0.106 & 0.107 & 0.208 & 0.056 \\
+			Ionosphere & 0.993 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
+			Iris & 0.553 & - & \textbf{0.708} & 0.540 & 0.585 & 0.684 \\
+			Led7Digit & 0.005 & - & 0.425 & \textbf{0.571} & 0.327 & - \\
+			Monk2 & 0.991 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
+			Moons & 0.995 & \textbf{1.000} & 0.241 & 0.831 & \textbf{1.000} & 0.180 \\
+			Movement Libras & 0.004 & - & \textbf{0.293} & \textbf{0.293} & 0.000 & - \\
+			Pima & 0.899 & \textbf{1.000} & 0.076 & 0.076 & \textbf{1.000} & -0.005 \\
+			Saheart & 0.981 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
+			Sonar & \textbf{1.000} & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
+			Soybean & 0.563 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
+			Spectfheart & \textbf{1.000} & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
+			Spiral & 0.998 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
+			Tae & \textbf{0.392} & - & 0.008 & 0.000 & 0.035 & 0.000 \\
+			Vehicle & 0.044 & \textbf{1.000} & 0.122 & 0.081 & 0.689 & - \\
+			Vowel & 0.003 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
+			Wdbc & 0.972 & \textbf{1.000} & 0.486 & 0.522 & 0.000 & 0.000 \\
+			Zoo & 0.153 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
+			\hline
+			Mean  & \textbf{0.647} & 0.224 & 0.212 & 0.28 & 0.319 & -0.228 \\
 			\hline
@@
 
-In order to carry out this study we will use Bayesian statistical tests, instead of the classic Hull Hypothesis Statistical Tests (NHST).
-In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
-\textit{"In a nutshell: NHST do not answer the question we ask"}.
-To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
-
-To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
-Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
-This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.
-
-Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate.
-As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
-
-Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
-That is: What can we conclude when NHST do not reject the null hypothesis?
-We can not infer anything since NHST can not provide evidence in its favor.
-
-Finally, there are two other problems that researchers face when performing NHST.
-The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results.
-The second one is the need to previously formalize the intentions of the sampling of the results, which are usually fixed a posteriori; this could lead to a misreading of said results.
-
-As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST.
-In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test.
-To make use of it we will employ the R package \texttt{rNPBST}, whose documentation and guide can be found in \cite{carrasco2017rnpbst}.
-
-From now on, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA$_{CC}$ as sample $A$, and to the results obtained with SHADE$_{CC}$ as sample $B$.
-
-The Bayesian sign test is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
-To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
-In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$.
-Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
-$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$
-
-One of the major advantages of the Bayesian sign test over the null hypothesis tests such as the (Wilcoxon's signed-rank test) is that we can obtain a visual representation of its results.
+Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA$_{CC}$ as sample $A$, and to the results obtained with SHADE$_{CC}$ as sample $B$.
+
+One of the major advantages of the Bayesian sign test is that we can obtain a very illustrative visual representation of its results.
 We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/BayesSignoARI.pdf}
+		\includegraphics[width=1\linewidth]{BayesSignoARI.pdf}
 		\caption{ARI}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/BayesSignoTime.pdf}
+		\includegraphics[width=1\linewidth]{BayesSignoTime.pdf}
 		\caption{Time}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/BayesSignoUnsat.pdf}
+		\includegraphics[width=1\linewidth]{BayesSignoUnsat.pdf}
 		\caption{Unsat}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriCOPKM.pdf}
+		\includegraphics[width=1\linewidth]{AriCOPKM.pdf}
 		\caption{SHADE$_{CC}$ vs COPKM}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriLCVQE.pdf}
+		\includegraphics[width=1\linewidth]{AriLCVQE.pdf}
 		\caption{SHADE$_{CC}$ vs LCVQE}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriRDPM.pdf}
+		\includegraphics[width=1\linewidth]{AriRDPM.pdf}
 		\caption{SHADE$_{CC}$ vs RDPM}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriTVClust.pdf}
+		\includegraphics[width=1\linewidth]{AriTVClust.pdf}
 		\caption{SHADE$_{CC}$ vs TVClust}
@@
 	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriCECM.pdf}
+		\includegraphics[width=1\linewidth]{AriCECM.pdf}
 		\caption{SHADE$_{CC}$ vs CECM}
@@
 
-\clearpage
-
 \section{Conclusions} \label{sec:conclusiones}
@@
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 
@@
 
+\section*{Compliance with ethical standards}
+
+\section*{Declaration of competing interest}
+
+The authors declare that there is no conflict of interest.
+
 \clearpage

### 2019-06-11 Papers/descartados.tex
SHADE terminado y DILS avanzado

--- 2019-04-04 Papers/SHADE/descartados.tex
@@
 Lastly, it is mandatory to name the widely known application to GPS data of constrained clustering, presented in \cite{wagstaff2001constrained}.
+
+\subsection{Short Introduction to Iterated Local Search}
+
+Before defining the iterated local search---from now on ILS---,it is necessary to understand the local search and the motivation that leads to designing methods with greater exploratory capacity.
+
+A local search is a procedure that consists of performing an heuristics-guided iterative improvement.
+Taking as initial state a well defined solution to the problem, the local search uses a neighbor generation operator to explore new solutions.
+When a solution better than the current one is found, it replaces the current one.
+In order to evaluate the solutions, the local search uses a function that assigns a value to each one of them, this is known as a target function.
+***This value will be better the closer the solution that evaluates the optimal solution.***
+
+Both the neighbor generation operator and the target function used by the local search must be defined for each problem.
+The target function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem.
+The neighbor generation operator takes as base a solution to the problem $s$, and generates a new solution $s^\prime$ by modifying some of its components in a local way.
+
+The Algorithm \ref{alg:LS} presents a basic scheme for a local search procedure.
+It is worth noting that the condition of line 4 must be adapted if the problem to be solved is one of maximization or minimization.
+In addition, the general scheme considers that the optimization process ends when the current solution is not improved, however the stop condition can be adapted to each problem.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{GenerateNeighbor}{GenerateNeighbor}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Initial solution $s$}
+	\BlankLine
+	\While{$improvement$}{
+		$improvement \leftarrow$ \texttt{false} \\
+
+		$s^\prime \leftarrow $ \GenerateNeighbor{$s$}
+
+		\If{$f(s^\prime) < f(s)$}{
+
+			$s \leftarrow s^\prime$\\
+			$improvement \leftarrow$ \texttt{true} \\
+
+		}	
+	}
+	\BlankLine
+	\KwRet ($s$)
+
+	\caption{Local Search}\label{alg:LS}
+\end{algorithm}
+
+With the local search defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$.
+Figure \ref{img:LS} shows a pictorical representation of the above.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{Figures/LS.png}
+	\caption{Local optimum achieved by a local search procedure}\label{img:LS}
+\end{figure}
+
+It is therefore necessary to appeal to techniques that introduce diversity in the search to avoid the problem of local optimal. This is exactly what ILS is trying to do.
+
+With ILS, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^\prime$ that will not be a local optimum and will be far enough away from $s^*$.
+After that, we will apply a local search procedure to find a new local optimal $s^{*\prime}$.
+Figure \ref{img:ILS} shows a representation of the above concepts.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{Figures/ILS.png}
+	\caption{No se como llamar a esta puta imagen}\label{img:ILS}
+\end{figure}
+
+It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^\prime$ is $s^*$ and therefore the solution space will not be explored effectively.
+On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.
+
+The method for generating the perturbation applied to the local optimum $s^*$ may vary depending on the problem, and may be based on previous solutions.
+In this way, it is possible to keep a history of the solutions reached so far, so they can be used when applying the perturbation.
+Algorithm \ref{alg:ILS} summarizes the ILS overall optimization process.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{LocalSearch}{LocalSearch}
+	\SetKwFunction{Perturbation}{Perturbation}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Initial solution $s$}
+	\BlankLine
+	$s^* \leftarrow$ \LocalSearch{$s$}\\
+	\While{Termination criteria are not met}{
+
+		$s^\prime \leftarrow $ \Perturbation{$s^*$, $history$}\\
+		$s^{*\prime} \leftarrow$ \LocalSearch{$s^\prime$}
+
+		\If{$f(s^{*\prime}) < f(s^*)$}{
+
+			$s^* \leftarrow s^{*\prime}$
+
+		}	
+	}
+	\BlankLine
+	\KwRet ($s*$)
+
+	\caption{Iterated Local Search}\label{alg:ILS}
+\end{algorithm}
+
+In the ILS scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only use the difference in the cost of the solutions to be compared to choose between them.
+This is the most simple scheme and does not depends on previous solutions recorded in the history.

### 2019-06-12 Papers/DILS/paperDILS.tex
Paper DILS completado texto

@@
 
-\journal{Applied Soft Computing}
+\journal{Computers and Operations Research}
 
@@
 This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose the first application of Differential Evolution to the constrained clustering problem, which is a generalization of traditional clustering that considers additional information in the form of constraints.
-We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
+We propose a new heuristic algorithm, the Dual Iterative Local Search, and prrove its capability to provide quiality results for the constrained clustering problem.
+We will compare the results obtained by this proposal to those obtained by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 \begin{keyword}
-constrained clustering, instance-level, must-link, cannot-link, genetic algorithm, differential evolution, Bayesian statistical tests.
+constrained clustering, instance-level, must-link, cannot-link, dual iterative local search.
 \end{keyword}
@@
 The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it.
-As mentioned earlier, it is a difficult problem to solve, so nature-inspired techniques are presented as a promising option to find quality approximate solutions.
-Nature is the best example of adaptive problem solving since it can apply an optimal strategy suited for each natural phenomenon \cite{fausto2019ants}.
-Nature-inspired algorithms are designed to emulate natural optimization phenomena, such as evolution, collective behavior of animals, physics laws or even human being-related processes.
-There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the Biased Random-key Genetic Algorithm (BRKGA) presented in \cite{de2017comparison}.
-Swarm-based methods have also been applied to constrained clustering, such as the one presented in \cite{xu2013improving}.
+As mentioned earlier, it is a difficult problem to solve, so heuristic techniques are presented as a promising option to find quality approximate solutions.
+Heuristic algorithms are designed to explore the solution space associated with a problem guided by a fitness (heuristic) function \textcolor{red}{CITA}.
+In this type of approaches the key is to find a good exploration-exploitation tradeoff.
+This is the reason why the classic Local Search (LS) algorithm derived in the Iterative Local Search (ILS) algorithm, which introduces periodics perturbations in the exploration of the solutions space to avoid local optima \textcolor{red}{CITA}.
+Even in ILS the exloitation-exploration tradeoff is unbalanced towards exploitationm, lacking from exploration.
 
-Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}.
-In this paper we propose the first DE application to find quality solutions for the constrained clustering problem.
-In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
-We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant.
-We will properly compare the results obtained by our new method with previous nature-inspired algorithms, as well as with the constrained clustering state-of-the-art.
+In this paper we propose the Dual Iterative Local Search (DILS) algorithm, which aims to address the exploration lacking problem by using techniques typical of nature-inspired algorithms such as genetic algorithms \textcolor{red}{CITA}.
+To put it simple, we use ILS as the host algorithm and introduce genetic algorithm operators and an adaptative reinitialization method to encourage exploration of the solution space.
+This is contrary to the well-known tendency to introduce exploitation techniques in over-explorative algorithm \textcolor{red}{CITA}.
+In them the host algorithm is usually evolutionary and methods such as LS or simulated annealing \textcolor{red}{CITA} are used to find a better balance between exploration and exploitation \textcolor{red}{CITA}.
+An example of this are memetic algorithms \textcolor{red}{CITA}.
 
-Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
-In Section \ref{sec:brkga} the BRKGA algorithm and its adaptation for constrained clustering will be briefly reviewed.
-Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
-In Section \ref{sec:SHADEadapt} the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function, will be presented.
+Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and the state-of-the-art.
+In Sections \ref{sec:DILS} and \ref{sec:DILS_CC} the DILS algorithm and its constrained clustering formulation will be reviewed.
 Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
 Finally, in Section \ref{sec:conclusiones} conclusions are discussed.
-
-\textcolor{red}{Decir que mientras que en los algoritmos memeticos el host es el GA en DILS el host es ILS, conservando todas sus caracteristicas y añadiendo algunas tecnicas de los GA}
 
@@
 In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering (Section \ref{sec:BackSOTA}).
-We will also present the basis of the DE optimization method (Section \ref{sec:BackDE}).
 
@@
 The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
-This method redefines constraints from the point of view of belief functions and includes them in the fitness function.
+This method redefines constraints from the point of view of belief functions and includes them in the cost function.
 
@@
 
-\textcolor{red}{Poner aqui una descripción del BRKGA?}
+An adaptation of the Biased Random-Key Genetic Algorithm (BRKGA) \cite{gonccalves2011biased}---a variant of the Random-Key genetic algorithm \cite{bean1994genetic}---for the constrained clustering problem is proposed in \cite{de2017comparison}.
+It uses a genetic algorithm, combined with a local search procedure and guided by a penalty-style fitness function, to obtain a partition of the input dataset.
+It constitutes one of the few approaches to constrained clustering from the point of view of the evolutionary computing.
 
@@
 
-\section{The Dual Iterative Local Search Method}
+\section{The Dual Iterative Local Search Method} \label{sec:DILS}
 
@@
 
+\textcolor{red}{Poner los inputs}
+
 \begin{algorithm}
@@
 
-Cabe destacar que el proceso de optimización descrito hasta el momento es aplicable a cualquier problema de minimización, aunque es claro que axiste una adaptación directa de DILS para problemas de maximización.
-Además el proceso de búsqueda local empleado para obtener el individuo trial debe ser especificado para cada problema, así como cuestiones de representación numérica y las criterios de parada.
-
 It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems.
@@
 
-\section{DILS optimization scheme for constrained clustering}
+\section{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}
 
@@
 
-For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraints set for each one of them.
+For our experiments we will compare the results obtained by DILS$_{CC}$ and state-of-the-art methods over 25 datasets and 3 constraint sets for each one of them.
 Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at GitHub.
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
 Table \ref{tab:datasets} displays a summary of every dataset.
@@
 They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).
-
-\clearpage
 
@@
 
-Table \ref{tab:params} shows a summary of the parameter setup used for DILS$_{CC}$.
+Table \ref{tab:params} shows a summary of the parameter setup used for the DILS$_{CC}$.
 
@@
 The stop criterion is given by the number of evaluations of the fitness function $f$, which at most will be 300000.
+The implementation for DILS$_{CC}$ can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
 
 To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
+The implementation for the BRKGA algorith can be found in the same link as DILS$_{CC}$.
 In all cases the value for $k$ is equal to the number of classes displayed in Table \ref{tab:datasets}.
+
+\textcolor{red}{Meter todos los algoritmos en el mismo repo?}
 
@@
 
-Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-We can note that DILS$_{CC}$ represents a consistent improvement over BRKGA$_{CC}$ in both Unsat and Time, while in terms of ARI both methods are similar.
-This is because the penalty applied to the fitness function of both methods has little presence when operating with a small constraint set, and since they are population-based and heuristic-guided methods, they have similar exploration and exploitation capabilities.
-
 \begin{table}[!h]
@@
 			\hline
-			Dataset & DILS & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
 			\hline
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ comparing DILS$_{CC}$ and the state-of-the-art.}
+	\caption{Experimental results obtained for $CS_{10}$ by DILS$_{CC}$ and the state-of-the-art methods.}
 	\label{tab:results10}
 \end{table}
+
+Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
+We can note that DILS$_{CC}$ represents a consistent improvement over all other 5 methods, specially in the case of BRKGA.
+This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods.
 
@@
 
-		\caption{Experimental results obtained for $CS_{15}$ comparing DILS$_{CC}$ and the state-of-the-art.}
+	\caption{Experimental results obtained for $CS_{15}$ by DILS$_{CC}$ and the state-of-the-art methods.}
 	\label{tab:results15}
 \end{table}
+
+Results obtained with the $SC_{15}$ constraint set are presented in Table \ref{tab:results15}.
+We observe how DILS$_{CC}$ and COPKM are able to obtain partitions of the datasets perfectly coincidents with the true partitions.
+The fact that DILS$_{CC}$ is always able to output a partition of the datasets is worth the advantage over COPKM.
+It is even the case that COPKM is not able to provide an output partition while DILS finds the optimal partition or very close to the optimal partition for the same datasets.
 
@@
 
-		\caption{Experimental results obtained for $CS_{20}$ comparing DILS$_{CC}$ and the state-of-the-art.}
+	\caption{Experimental results obtained for $CS_{20}$ by DILS$_{CC}$ and the state-of-the-art methods.}
 	\label{tab:results20}
 \end{table}
+
+In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how DILS$_{CC}$ scales the quality of its results with the number of restrictions, as well as methods such as BRKGA or COPKM, the latter being more limited in other aspects.
 
@@
 	\end{subfigure}
-	\caption{Heat diagrams.}
+	\caption{Heat diagrams comparing DILS with the state-of-the-art methods.}
+	\label{fig:heatDiag}
 \end{figure}
+
+Figure \ref{fig:heatDiag} shows all six heatmaps obtained when applying the Bayesian sign test to DILS$_{CC}$ and the state-of-the-art methods.
+All heatmaps suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle.
+Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strog evidence that DILS$_{CC}$ does represent a significan improvement over all other six methods, being te results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.
+
+\textcolor{red}{No estan citadas las subimagenes una por una}
 
@@
 
-In this paper we proposed SHADE$_{CC}$, the first application of a DE-based algorithm to the SSL constrained clustering problem.
-Focusing on the instance-level ML and CL constraints, SHADE$_{CC}$ has proven that nature-inspired evolution-based techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
+In this paper we proposed the DILS heuristic method, along with its application to the SSL constrained clustering problem, which we called DILS$_{CC}$.
+Focusing on the instance-level ML and CL constraints, DILS$_{CC}$ has proven that heuristic techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
 
-Supported by the Bayesian statistical tests, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous evolution-based proposals.
-As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be equivalent or better when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
+Supported by the Bayesian statistical tests, we were able to objectively prove that the DILS$_{CC}$ approach is significantly better than the state-of-the-art methods.
+DILS$_{CC}$ has proven to be better when considering the quality and availability of the solutions, especially in cases where large sets of constraints are analyzed.
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 
 Our work has been supported by the research project TIN2017-89517-P and PP2016.PRI.I.02.
+
+\section*{Compliance with ethical standards}
+
+\section*{Declaration of competing interest}
+
+The authors declare that there is no conflict of interest.
 

### 2019-06-12 Papers/SHADE/paperSHADE.tex
Paper DILS completado texto

@@
 \begin{keyword}
-constrained clustering, instance-level, must-link, cannot-link, genetic algorithm, differential evolution, Bayesian statistical tests.
+constrained clustering, instance-level, must-link, cannot-link, differential evolution.
 \end{keyword}
@@
 In this section, we present an optimization scheme for the SHADE algorithm when applied to constrained clustering.
-We will call this new approach $SHADE_{CC}$.
+We will call this new approach SHADE$_{CC}$.
 To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of $n$ random-keys ($v = n$).
@@
 
-For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraints set for each one of them.
+For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraint sets for each one of them.
 Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at GitHub.
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/SHADE_CC}{GitHub}.
 Table \ref{tab:datasets} displays a summary of every dataset.
@@
 			\hline
-			Parameter & Meaning & BRKGA & SHADE \\
+			Parameter & Meaning & BRKGA$_{CC}$ & SHADE$_{CC}$ \\
 			\hline
@@
 
-	\caption{Parameters setup used for BRKGA and SHADE.}
+	\caption{Parameters setup used for BRKGA$_{CC}$ and SHADE$_{CC}$.}
 	\label{tab:params}
@@
 In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
-
-To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}
+A Python implementation for both algorithms can be found at \href{https://github.com/GermangUgr/SHADE\_CC}{GitHub}
+
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
 

### 2019-06-14 Papers/DILS/paperDILS.tex
Correcciones en paper DILS y escript para estudio de xi

@@
 
-\title{Enhancing instance-level constrained clustering through differential evolution}
+\title{Improving instance-level constrained clustering through Dual Iterative Local Search}
 
@@
 This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose a new heuristic algorithm, the Dual Iterative Local Search, and prrove its capability to provide quiality results for the constrained clustering problem.
+We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to provide quiality results for the constrained clustering problem.
 We will compare the results obtained by this proposal to those obtained by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
@@
 As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$ \cite{jain1999data}.
+The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement \cite{jain1999data}.
 
@@
 The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it.
-As mentioned earlier, it is a difficult problem to solve, so heuristic techniques are presented as a promising option to find quality approximate solutions.
-Heuristic algorithms are designed to explore the solution space associated with a problem guided by a fitness (heuristic) function \textcolor{red}{CITA}.
-In this type of approaches the key is to find a good exploration-exploitation tradeoff.
-This is the reason why the classic Local Search (LS) algorithm derived in the Iterative Local Search (ILS) algorithm, which introduces periodics perturbations in the exploration of the solutions space to avoid local optima \textcolor{red}{CITA}.
-Even in ILS the exloitation-exploration tradeoff is unbalanced towards exploitationm, lacking from exploration.
+As mentioned earlier, it is a difficult problem to solve, so metaheuristic are presented as a promising option to find quality approximate solutions.
+Metaheuristic algorithms are designed to explore the solution space associated with a problem guided by a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
+In this type of approaches the key is to find a good diversification-intensification tradeoff.
+This is the reason why the classic Local Search (LS) heuristic algorithm derived in the Iterative Local Search (ILS) metaheuristic algorithm, which introduces periodics perturbations in the exploration of the solution space to avoid local optima \cite{lourencco2010iterated}.
+ILS---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
+ILS is also very commonly used to design hybrid methods with good diversification-intensification tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
 
-In this paper we propose the Dual Iterative Local Search (DILS) algorithm, which aims to address the exploration lacking problem by using techniques typical of nature-inspired algorithms such as genetic algorithms \textcolor{red}{CITA}.
-To put it simple, we use ILS as the host algorithm and introduce genetic algorithm operators and an adaptative reinitialization method to encourage exploration of the solution space.
-This is contrary to the well-known tendency to introduce exploitation techniques in over-explorative algorithm \textcolor{red}{CITA}.
-In them the host algorithm is usually evolutionary and methods such as LS or simulated annealing \textcolor{red}{CITA} are used to find a better balance between exploration and exploitation \textcolor{red}{CITA}.
-An example of this are memetic algorithms \textcolor{red}{CITA}.
+In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptative and guided way.
+We apply it to the constrained clustering problem, prooving that it constitute a good approach to obtain quality results.
 
@@
 Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why nature-inspired algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
+That is the reason why metaheuristic algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
 
@@
 An adaptation of the Biased Random-Key Genetic Algorithm (BRKGA) \cite{gonccalves2011biased}---a variant of the Random-Key genetic algorithm \cite{bean1994genetic}---for the constrained clustering problem is proposed in \cite{de2017comparison}.
-It uses a genetic algorithm, combined with a local search procedure and guided by a penalty-style fitness function, to obtain a partition of the input dataset.
+It uses a genetic algorithm, combined with an LS procedure and guided by a penalty-style fitness function, to obtain a partition of the input dataset.
 It constitutes one of the few approaches to constrained clustering from the point of view of the evolutionary computing.
@@
 
-The Dual Iterative Local Search (DILS) is a new variant of the classic Iterated Local Search (ILS) method.
-Its goal is to perform a search in the solutions space to find the solution with the best fitness function $f$ value by introducing diversity in a guided way.
+The Dual Iterative Local Search (DILS) is a new variant of the classic ILS method.
+Its goal is to perform a search in the solution space to find the solution with the best fitness function $f$ value by introducing diversity in an adaptative and guided way.
 While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-introducing methods and to avoid local optima.
@@
 
-DILS mimics some of the genetic algorithms techniques to introduce diversity into the optimization process, while also manteining the esentials of ILS.
-In each generation $G$ of the optimization process DILS builds a new individual by applying a crossover operator to $m_b$ and $m_w$.
+In each generation $G$ of the optimization process DILS builds a new individual by applying a recombination operator to $m_b$ and $m_w$.
 After that, it applies a strong mutation operator to the newly generated individual.
-This is the invidual DILS applies the local search procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
-The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the replacement operator for DILS.
-Equation \ref{eq1} displays the expression for this operator in the case of a minimization problem.
+This is the invidual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
+The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the acceptance criterion for DILS.
+Equation \ref{eq1} displays the expression for this criterion in the case of a minimization problem.
 
@@
 
-Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solutions space.
+Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solution space.
 To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory $\{m_b, m_w\}$, Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.
@@
 \noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
-This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via crossover operator) as the best parent.
+This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via recombination operator) as the best parent.
 Figure \ref{img:DILS_diag} summarizes the DILS optimization process.
@@
 	\includegraphics[scale=0.25]{Figures/DILSDiagram.png}
-	\caption{Diagram summarizing DILS optimization process. \textcolor{red}{La figura es horrible (llamar a javi)}}\label{img:DILS_diag}
+	\caption{Diagram summarizing DILS optimization process.}\label{img:DILS_diag}
 \end{figure}
@@
 In one hand, if we are dealing with a minimization problem and we set $\xi > 0$, then the true worst individual is only reinitializatized when the trial individual $m_t$ replaced $m_w$ and is also better than $m_b$ by a margin.
-On the other hand and also for the minimization case but this time with $\xi < 0$ we allow DILS to reinitialize the true worst individual even $m_t$ is not better than $m_b$ but they are close.
+On the other hand, and also for the minimization case, but this time with $\xi < 0$, we allow DILS to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
 It comes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
-Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$.
+Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$, and we recommend $[-0.5, 0.5]$.
 To summarize, the closer $\xi$ is to 1 the more restrictive the reinitialization criterion is---in the minimization case.
-It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages, resembling some fatures of the simmulated annealing algorithm \cite{van1987simulated}.
+It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages.
+The reason being is that the fitness value of the best individual $f(m_b)$ never deteriorate ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it become harder for $m_b$ to get better o close to it.
 Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.
-
-\textcolor{red}{Poner los inputs}
 
@@
 	\SetKwFunction{Worst}{Worst}
-	\SetKwFunction{Crossover}{Crossover}
+	\SetKwFunction{Recombination}{Recombination}
 	\SetKwFunction{Mutation}{Mutation}
 	\SetKwFunction{LocalSearch}{LocalSearch}
-	\KwIn{}
+	\KwIn{Probability for recombination operator $p_r$, segment size for mutation operator $p_s$, reinitialization method tolerance $\xi$.}
 	\tcp{Initialization phase}
@@
 		\tcp{Generate new individual}
-		$m_t \leftarrow $ \Crossover{$m_b, \; m_w$}\\
+		$m_t \leftarrow $ \Recombination{$m_b, \; m_w$}\\
 		$m_t \leftarrow $ \Mutation{$m_t$}\\
@@
 
-For the mutation and crossover operators that appear in the Algorithm \ref{alg:DILS}, we use the uniform crossover operator \cite{spears1995virtues} and the segment mutation operator \textcolor{red}{REFERENCIA?} respectively.
-This type of mutation consists of selecting a fixed size segment from the features of the individual and replacing it with a random segment of the same size.
-The size of the segment must be fixed at the beginning of the optimization process.
+For the recombination and mutation operators that appear in Algorithm \ref{alg:DILS}, we use the uniform recombination and the segment mutation operator respectively.
+The uniform recombination consists of selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}.
+The segment mutation operator consists of replacing a fixed size segment from the features of the individual for randomly generated features CITA.
+The size of the segment $p_s$ must be fixed at the beginning of the optimization process.
 
 It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems.
-In addition, the local search procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation matters and termination criteria.
+In addition, the LS procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation matters and termination criteria.
 
@@
 To begin with, we will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---being $n$ the number of instances of the dataset.
-Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, si each individual $m$ is a solution to the clustering problem.
+Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, so each individual $m$ is a solution to the clustering problem.
 Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to intially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.
@@
 
-We also need to specify a local search procedure for DILS to be able to improve the mutant individual to obtain the trial vector $m_t$.
-Algorithm \ref{alg:LS} shows the local search precedure used in DILS$_{CC}$.
+We also need to specify an LS procedure for DILS to be able to improve the mutant individual to obtain the trial vector $m_t$.
+Algorithm \ref{alg:LS} shows the LS used in DILS$_{CC}$.
 Note that, in order to not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
-We also stablish a maximum number of neighbors $max\_neighbors$ that can be generated, this is done to find a better exploration-exploitation balance.
+We also stablish the maximum number of neighbors $p_m$ that can be generated in each call, this is done to find a better diversification-intensification tradeoff.
 
@@
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $max\_neighbors$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $p_m$.}
 	\BlankLine
-	\While{$improvement$ {\bf and} $generated < max\_neighbors$}{
+	\While{$improvement$ {\bf and} $generated < p_m$}{
 		$improvement \leftarrow$ \texttt{false} \\
@@
 			Evals & Fitness function evaluations & 300000 \\
-			$segment\_size$ & Segment size for the mutation operator & $0.3 \times n$ \\
-			$max\_neighbors$ & Maximum number of neighbors generated in the local search procedure & Evals $\times 0.01$ \\
-			$p_\text{inherit}$ & Probability that a feature is inherited from $m_b$ & $0.3$ \\
+			$p_s$ & Segment size for the mutation operator & $0.3 \times n$ \\
+			$p_m$ & Maximum number of neighbors generated in each call to the LS procedure & Evals $\times 0.01$ \\
+			$p_r$ & Probability that a feature is selected from $m_b$ for recombination & $0.3$ \\
 			$\xi$ & Reinitialization method tolerance & $0.2$ \\
@@
 
-\textcolor{red}{Meter todos los algoritmos en el mismo repo?}
+\textcolor{red}{Meter todos los algoritmos en el mismo repo para no tener que poner tantos enlaces a github? Poner los enlaces como notas a pie de pagina tambien?}
 
@@
 Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-We can note that DILS$_{CC}$ represents a consistent improvement over all other 5 methods, specially in the case of BRKGA.
-This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods.
+We can note that DILS$_{CC}$ represents a consistent improvement over all other 5 methods.
+This is due to the fact that the intensification capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lower level of constraint-based information we work with.
 
@@
 Results obtained with the $SC_{15}$ constraint set are presented in Table \ref{tab:results15}.
-We observe how DILS$_{CC}$ and COPKM are able to obtain partitions of the datasets perfectly coincidents with the true partitions.
+We observe how DILS$_{CC}$ and COPKM are able to obtain ARI values equal to 1, meaning that the resulting partitions of the datasets perfectly match the true partitions.
 The fact that DILS$_{CC}$ is always able to output a partition of the datasets is worth the advantage over COPKM.
@@
 In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how DILS$_{CC}$ scales the quality of its results with the number of restrictions, as well as methods such as BRKGA or COPKM, the latter being more limited in other aspects.
+It is also woth noting how DILS$_{CC}$ provides the better average result for each of the three constraint set analyzed in this paper.
 
@@
 
-\textcolor{red}{No estan citadas las subimagenes una por una}
-
 \clearpage

### 2019-06-14 Software/EstudioXi.py

@@ -1,0 +1,80 @@
+import numpy as np
+from ILSNueva.ILSNueva import ILSNueva
+from functions import *
+from sklearn.metrics import adjusted_rand_score
+import time
+
+
+def main():
+
+    np.random.seed(43)
+
+    #CARGA DE DATOS (SOLO SE CARGA IRIS)
+
+    names_array, datasets_array, labels_array = load_all_datasets()
+
+    name = names_array[0]
+    dataset = datasets_array[0]
+    labels = labels_array[0]
+
+    const_array = load_constraints([name], [0.1])
+    const = const_array[0][0]
+
+    #BUCLE DE OBTENCION DE DATOS
+
+    nb_runs = 10
+    max_gen = 300
+    nb_clust = len(set(labels))
+    ml_const, cl_const = get_const_list(const)
+    xi_array = [-1, -0.5, -0.4, -0.3, -0.2, -0.1, 0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 1]
+    xi_array = [-1, -0.5]
+
+    total_restarts_matrix = np.zeros((nb_runs, len(xi_array)))
+    mean_restarts_matrix = np.zeros((max_gen, len(xi_array)))
+
+    for xi_pos in range(len(xi_array)):
+
+        print("Procesando con xi = " + str(xi_array[xi_pos]))
+
+        mean_restarts = np.zeros(max_gen)
+        total_restarts = []
+
+        for j in range(nb_runs):
+
+            dils = ILSNueva(dataset, ml_const, cl_const, nb_clust, 0.3, max_gen, 0.3, xi_array[xi_pos])
+            dils_retarts = dils.run(1000)
+            total_restarts.append(dils_retarts[np.size(dils_retarts) -1])
+            mean_restarts += dils_retarts
+
+        mean_restarts /= nb_runs
+        total_restarts_matrix[:, xi_pos] = total_restarts
+        mean_restarts_matrix[:, xi_pos] = mean_restarts
+
+
+    print("------------------------- total_restarts_matrix -------------------------")
+    
+    for i in range(np.shape(total_restarts_matrix)[0]):
+
+        print("RowName", end = '')
+
+        for j in range(np.shape(total_restarts_matrix)[1]):
+
+            print(" & %d" % (total_restarts_matrix[i,j]), end = '')
+
+        print(" \\\\")
+
+    print("------------------------- mean_restarts_matrix -------------------------")
+
[…]

### 2019-06-14 Software/ILSNueva/ILSNueva.py

@@ -177,8 +177,12 @@ class ILSNueva:
 
 		self.init_ils()
 		gen_times = []
+		nb_restarts = 0
+		nb_restarts_array = []
 
-		while self._evals_done < max_evals:
+		#while self._evals_done < max_evals:
+		while len(gen_times) < self._max_neighbors:
+
 			start_gen = time.time()
 			worst = np.argmax(self._best_fitness)
 			best = (worst + 1) % 2
@@ -199,10 +203,16 @@ class ILSNueva:
 				self._best_solution[worst, :] = np.random.randint(0, self._result_nb_clust, self._dim)
 				self._best_fitness[worst] = self.get_single_fitness(self._best_solution[worst, :])[0]
 
+				nb_restarts += 1
+
 			gen_times.append(time.time() - start_gen)
+			nb_restarts_array.append(nb_restarts)
+
+			#print(str(np.size(gen_times)) + " " + str(nb_restarts) + " (" + str(self._evals_done) + ")")
 
 		best = np.argmin(self._best_fitness)
-		return self._best_solution[best, :], np.sum(gen_times)
+		return nb_restarts_array
+		#return self._best_solution[best, :], np.sum(gen_times)
 
 
 

### 2019-06-14 Software/ILSNuevaMain.py

@@ -18,7 +18,7 @@ def main():
     labels_array = labels_array[0:1]
 
     const_percent_vector = [0.05, 0.1, 0.15, 0.2]
-    const_percent_vector = [0.05]
+    const_percent_vector = [0.1]
     const_array = load_constraints(names_array, const_percent_vector)
 
     general_table_file = open("Results/ILSNueva_general_table_file.txt", "w+")
@@ -26,8 +26,8 @@ def main():
 
     #BUCLE DE OBTENCION DE DATOS
 
-    nb_runs = 1
-    max_eval = 300000
+    nb_runs = 5
+    max_eval = 30000
     population_size = 100
     run_ls = True
 
@@ -60,9 +60,9 @@ def main():
 
             for j in range(nb_runs):
 
-                ilsn = ILSNueva(data_set, ml_const, cl_const, nb_clust, 0.3, 300, 0.3, 0.003)
+                ilsn = ILSNueva(data_set, ml_const, cl_const, nb_clust, 0.3, 300, 0.3, xi)
                 start = time.time()
-                ilsn_assignment = ilsn.run(max_eval)[1]
+                ilsn_assignment = ilsn.run(max_eval)[0]
                 end = time.time()
                 mean_ars += adjusted_rand_score(labels, ilsn_assignment)
                 mean_execution_time += end - start
@@ -85,19 +85,5 @@ def main():
 
     general_table_file.close()
     results_file.close()
-
-    #REPRESENTANDO ALGUNOS RESULTADOS
-    # iris_plot1 = draw_data_2DNC(iris_set, np.asarray(iris_labels, np.uint8), 3, alg + "Iris Dataset ML")
-    # iris_plot2 = draw_data_2DNC(iris_set, np.asarray(iris_labels, np.uint8), 3, alg + "Iris Dataset CL")
-    #
-    # iris_plot3 = draw_data_2DNC(iris_set, np.asarray(iris_ilsn_assignment, np.float), 3, alg + "Iris Dataset Results ML")
-    # iris_plot4 = draw_data_2DNC(iris_set, np.asarray(iris_ilsn_assignment, np.float), 3, alg + "Iris Dataset Results CL")
-    #
-    # ax1, ax2 = draw_const(iris_set, iris_const, iris_plot1, iris_plot2, "Ml Original", "Cl Original")
-    #
[…]

### 2019-06-18 Memoria/Chapters/Clustering.tex
Empezada memoria TFM

\chapter{Short introduction to Clustering}\label{ch:Breve introducción al Clustering}

Firstly, it will be necessary to provide an introduction, albeit a brief one, to clustering and its applications, in order to ease the understanding of the following chapters. For this purpose we will follow the study carried out in \cite{Everitt:2009:CA:1538772}.

\section{Reasons for classifying}

Classification can be understood as a way of simplifying the information contained in large datasets in a way that is easily understood by humans. In this way, the processes of extracting useful information and applying it are simplified. Thus, if we are able to meaningfully divide a large dataset into subsets or groups, we can extract information common to all elements of the subset and provide a precise description that encompasses them.

The need to analyze information in this way grows with the emergence and availability of large datasets in science. The analysis of this type of information by classification, or clustering, is known today as \textit{Data Science}. In the 21st century a particular interest in data science arises from the appearance of the \textit{World Wide Web}, known as the Internet, where the aim has become to extract relevant information from the Web pages that make up this vast network.

Es importante destacar que, en la mayoría de las ocasiones, no hay un sólo criterio de clasificación para un mismo conjunto de datos, sino que existe una amplia variedad de los mismos. En el caso de las personas, podrían ser clasificadas, por ejemplo, en base a sus ingresos económicos, o según la cantidad de calorías que consumen a lo largo de un periodo de tiempo definido. Así, distintos criterios de clasificación no tienen por qué dar como resultado la misma división en grupos del conjunto a clasificar; de esta manera, diferentes criterios servirán a diferentes propósitos.

\section{Métodos numéricos para el Clustering}

Los métodos numéricos para el clustering surgen en ramas de las ciencias naturales, como la biología o la zoología, en un intento de eliminar la subjetividad implícita en el proceso de clasificación que se desencadena al descubrir una nueva especie. El objetivo es proporcionar un método no subjetivo y estable para clasificar y agrupar elementos.

Estos métodos adoptan diversos nombres que varían según el campo en el que se apliquen: taxonomía numérica (\textit{numerical taxonomy}) en biología, $Q$ análisis (\textit{$Q$ analysis}) en psicología, o reconocimiento de patrones no supervisado (\textit{unsupervised pattern recognition}) en el campo de la inteligencia artificial. No obstante, hoy en día, análisis de clusters (\textit{Clusters analysis}) o simplemente clustering son los términos más ampliamente aceptados y extendidos para referirse a tareas que involucran el descubrimiento de subgrupos dentro de un conjunto de elementos.

En la mayoría de las aplicaciones del clustering, el objetivo es obtener una partición de los datos, en la que cada instancia u objeto pertenezca a un único cluster, y la unión de todos ellos contenga a todos los objetos individuales. Dicho esto, debe destacarse que, en algunas circunstancias, son aceptables soluciones en las que existe solapamiento entre clusters, así como el hecho de que puede no existir una partición aceptable de los datos.

La manera más ampliamente extendida de representar la información sobre la que se debe aplicar clustering es una matriz $X$ de dimensión $n\times p$, en la que cada fila corresponde a una instancia u objeto a procesar, y cada columna corresponde a una de las variables que caracterizan dichas instancias u objetos. El término comúnmente aceptado para referirse a cada fila es el de \textit{vector de características}.

$$ X = \left( \begin{array}{cccc}

x_{1,1} & x_{1,2} & \cdots & x_{1,p} \\

x_{2,1} & \cdots & \cdots & \cdots  \\

\vdots & \vdots & \vdots & \vdots \\

x_{n,1} & \cdots & \cdots & x_{n,p} \\

\end{array} \right) $$

La entrada $x_{i,j}$ en $X$ se corresponde con el valor de la variable $j$-esima en la instancia $i$.

Las variables en $X$ pueden ser una mezcla de atributos en un dominio continuo, discreto o categórico. Además, es posible que, en problemas reales, algunas entradas no estén disponibles. Esta mezcla de tipos de variables y los valores perdidos pueden complicar la tarea del clustering. Sin embargo, existen métodos para tratar estos casos, como la inferencia de valores perdidos o las transformaciones de dominio.

En algunas aplicaciones, las filas de la matriz pueden contener medidas repetidas de la misma variable, aunque bajo diferentes condiciones, o en diferentes momentos, incluso en diferentes localizaciones espaciales. Un ejemplo de ello pueden ser las medidas de altura de un grupo de niños en un mismo mes a lo largo de diferentes años. Este tipo de datos posee una estructura que, de nuevo, puede complicar la tarea del clustering.

Algunos métodos de clustering conllevan realizar transformaciones sobre la matriz $X$ para transformarla en una matriz de $n \times n$ en la que se almacenan medidas extraídas de la matriz $X$ que relacionen una instancia con todas las demás, como pueden ser la similitud, distancia o disimilitud.

El clustering es, dicho de manera simple, descubrimiento de grupos en datos, y no debe ser en ningún caso confundido con los métodos de discriminación o asignación, conocidos en el ámbito de la inteligencia artificial como aprendizaje supervisado. En ellos los grupos son conocidos a priori y el objetivo del análisis es obtener una regla de clasificación o clasificador que permita asignar nuevas instancias o individuos a uno de los grupos ya conocidos.

Una vez definida la estructura general de los métodos de clustering, estaría justificado preguntar, ¿qué es un cluster? La Sección \ref{QueEsCluster} intentará dar respuesta a esta pregunta.

\subsection{¿Qué es un cluster?} \label{QueEsCluster}

Hasta este momento, los términos cluster, grupo y clase han sido empleados de una manera completamente intuitiva, sin necesidad alguna de definición formal, una prueba más de lo innato de estos conceptos en el ser humano. De hecho, dar una definición formal de cluster resulta una tarea, no sólo complicada, sino en muchas ocasiones poco útil. Bonner, por ejemplo, propuso en 1964 una definición de cluster completamente dependiente de la interpretación del usuario. En lo que a este autor respecta, un cluster es aquello que el usuario entiende como cluster sin haberle propuesto una definición formal del mismo.

Aunque la definición de Bonner es acertada en un amplio rango de situaciones, autores como Cormack, en 1971, o Gordon en 1999, proponen una definición algo más analítica desde el punto de vista matemático, definiendo un cluster en términos de cohesión interna (homogeneidad), y aislamiento externo (separación). La Figura \ref{fig:figure1} muestra de manera informal las propiedades descritas anteriormente, de forma que a cualquier observador le resultarán aparentes los clusters presentes en ella sin necesidad de una definición formal de cluster. Este hecho puede explicar por qué alcanzar una definición matemáticamente precisa de homogeneidad y separación puede llegar a ser innecesario.

No queda completamente clara la manera en que las personas reconocen diferentes clusters cuando éstos son representados en un plano, pero una de las variables que con certeza influye es la distribución de distancias relativas entre los objetos o puntos.

\clearpage

\begin{figure}[bth]
	\myfloatalign
	{\includegraphics[width=.3\linewidth]{imagenes/c2/TwoBasicsClusters}}
	{\includegraphics[width=.3\linewidth]{imagenes/c2/MoonsBasics}}
	{\includegraphics[width=.3\linewidth]{imagenes/c2/ThreeBasicClusters}}
	\caption[Clusters con cohesión interna y/o aislamiento externo]{Clusters con cohesión interna y/o aislamiento externo.}\label{fig:figure1}
\end{figure}

Por otra parte, como ya mencionamos anteriormente en esta sección, puede darse el caso de que en un conjunto de datos no exista una partición justificada. En la Figura \ref{fig:figure2} se muestra un conjunto de datos para el que la mayoría de observadores llegaría a la conclusión de que no existen grupos diferenciados, simplemente una nube de puntos uniformemente distribuidos. Idealmente, es de esperar que un método de clustering aplicado a este mismo conjunto de datos llegue a la misma conclusión.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.2]{imagenes/c2/rand.png} 
	\caption{Distribución uniforme de puntos.}\label{fig:figure2}
\end{figure}

Sin embargo, la mayoría de métodos de aprendizaje no supervisado darán como resultado un particionamiento uniforme como el que se muestra en la Figura \ref{fig:figure3}. El número de particiones encontradas dependerá del método aplicado, si bien en cualquier caso obtendremos un particionamiento uniforme.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.2]{imagenes/c2/randClasif.png} 
	\caption{Distribución uniforme de puntos clasificados.}\label{fig:figure3}
\end{figure}

El proceso de dividir una distribución homogénea de datos en diferentes grupos se conoce como disección, y tal proceso puede ser útil en ciertas circunstancias. Sin embargo, dado que en la mayoría de las ocasiones de aplicación real de métodos de clusters, no se conoce a priori la estructura de los datos, existe el riesgo de interpretar todas las soluciones en términos de existencia de subgrupos, lo que conllevaría la imposición de una estructura ficticia en datos en los que no hay estructura presente.

\section{Aplicaciones del clustering}

Como ya se ha indicado, el problema general al que intenta dar solución el clustering está presente en muchas disciplinas: biología, botánica, medicina, psicología, geografía, marketing, procesamiento de imágenes, psiquiatría, arqueología, etc. En esta sección se presentan algunas de las aplicaciones del clustering relacionadas con los citados campos.

\subsection{Aplicaciones en marketing}

Dividir los clientes en grupos homogéneos es una de las tareas más frecuentes en marketing. Un director de marketing podría preguntarse cómo agrupar los posibles clientes según los beneficios potenciales del producto que intenta introducir en el mercado. Por otra parte, un analista de marketing podría estar interesado en agrupar las empresas según sus características financieras, para poder analizarlas y predecir sus estrategias de mercado.

Un ejemplo de aplicación del clustering en este ámbito fue publicado por Green et al. (1967). Así, con un gran número de ciudades disponibles para el análisis, debieron restringir los lugares en los que llevar a cabo sus estudios debido a motivos económicos. Para ello hicieron uso del análisis de clusters, clasificando las ciudades en pequeños grupos basándose en 14 características de las mismas, entre ellas el tamaño e ingresos medios \textit{per capita}. Dado que se esperaba que las ciudades incluidas en un mismo grupo fueran muy similares, escogieron una ciudad de cada uno de ellos para realizar sus estudios.

Otra aplicación del análisis de clusters en el marketing fue descrita por Chakrapani (2004). En este caso, un fabricante de coches cree que comprar un coche deportivo no es una decisión basada sólo en capacidades económicas o edad, sino que es una decisión relacionada con el estilo de vida que llevan aquellos que deciden comprar un coche de estas características, frente a aquellos que no lo hacen. En consecuencia, el fabricante decide realizar un estudio, empleando análisis de clusters, que le permita identificar todas las características relacionadas con las personas que comprarían un coche deportivo, para así enfocar sus campañas de marketing a este sector específicamente.

\subsection{Aplicaciones en astronomía}

Dado un conjunto de datos astronómicos, los investigadores quieren saber, por ejemplo, cuantas clases de estrellas hay presentes en ellos, basándose en algun criterio estadístico. Las preguntas más frecuentes dentro de este ámbito son: ¿Cuantos objetos estadísticamente diferentes están presentes en los datos y a qué clase debe ser asignado cada objeto? ¿Aparecen clases de objetos previamente desconocidas?. El análisis de clusters puede ser aplicado para dar respuesta a estas cuestiones, ayudando a detectar objetos estadísticamente anómalos, así como a guiar el proceso de clasificación de los mismos. Algunos ejemplos incluyen el descubrimiento de quasars con alto corrimiento al rojo, quasars de tipo 2 (altamente luminosos, núcleos galácticos activos a menudo oscurecidos por polvo y gas), y enanas marrones.

Un ejemplo específico viene dado por el estudio de Faúndez-Abans et al. (1996), quienes aplicaron técnicas de clustering a datos sobre la composición química de 192 nebulosas planetarias. Se identificaron 6 grupos diferentes que eran similares en muchos aspectos a una clasificación previa de dichos objetos, pero que también mostraban diferencias interesantes que hasta ese momento los investigadores habían pasado por alto.

Un segundo ejemplo lo encontramos en el estudio de Celeux y Govaert (1992), quienes aplicaron clustering basado en distribuciones normales a un conjunto de 2370 estrellas, descritas por su velocidad relativa al núcleo galáctico y a la rotación galáctica. Usando un modelo de tres clusters, encontraron un cluster de gran tamaño y pequeño volumen, y dos de pequeño tamaño y gran volumen.

\subsection{Aplicaciones en psiquiatría}

Las enfermedades de la mente son a menudo más difíciles de diagnosticar que las enfermedades del cuerpo; es por ello que en el campo de la psiquiatría ha crecido el interés por las técnicas de análisis de clusters que permitan refinar, o incluso redefinir, las técnicas de diagnosis para este tipo de enfermedades. Gran parte de este trabajo involucra pacientes deprimidos, casos en los que el interés reside en distinguir entre dos tipos de depresión, la endógena (congénita), y la neurótica.

Pilowsky et al. (1968), por ejemplo, usando métodos desarrollados por otros autores, aplicaron técnicas de clustering a 200 pacientes en base a sus respuestas a un cuestionario sobre la depresión, junto a información sobre su sexo, edad, estado mental y enfermedad padecida. Éste es un claro ejemplo de variables de diferentes tipos incluidas en el mismo conjunto de datos. Uno de los grupos obtenidos como resultado de este estudio fue identificado como marcador de la depresión endógena.

El análisis de clusters también ha sido empleado para encontrar una clasificación de individuos que intentaron cometer suicidio, que podría sentar las bases para estudios posteriores sobre las causas y tratamientos del problema. Paykey y Rassaby (1978), por ejemplo, estudiaron 236 casos de suicidas fallidos registrados por el servicio de emergencias de una ciudad de los Estados Unidos de América. Del conjunto de posibles variables, 14 fueron seleccionadas como particularmente relevantes para la clasificación y, por tanto, fueron usadas en el análisis. Entre ellas se encontraban: edad, número de intentos de suicidio, gravedad de la depresión y grado de hostilidad, además de una serie de características demográficas. Al conjunto de datos resultante se le aplicó métodos de clustering; el resultado más significativo obtenido corresponde a una división en tres clusters bien definidos.

\subsection{Aplicaciones en meteorología y climatología}

Diariamente se recogen enormes cantidades de datos sobre la meteorología mundial. Explorar estos datos mediante técnicas de clustering puede aportar nuevos enfoques para la climatología y el medio ambiente.

Littmann (2000), por ejemplo, aplicó clustering a los datos recogidos sobre los cambios diarios en la presión superficial en la cuenca Mediterránea, y encontró 20 grupos que explicaban la varianza de las lluvias en las regiones centrales del Mediterráneo. Otro ejemplo viene de la mano de Liu y George (2005), quienes usaron el algoritmo K-medias difuso (\textit{Fuzzy K-means}, \acs{FKM}) con datos espaciotemporales de la climatología de las regiones del sur central de EEUU.

\subsection{Aplicaciones en arqueología}

La arqueología es otra de las disciplinas en la que resulta útil la aplicación del clustering. La clasificación de los diferentes objetos encontrados en los yacimientos puede ayudar a descubrir su uso, los periodos a los que pertenecen, así como la población que los utilizó. De forma similar, el estudio de materiales fosilizados puede ayudar a revelar como vivieron las sociedades prehistóricas.

Un ejemplo temprano de la aplicación de clustering a objetos arqueológicos viene dado por Hodson et al. (1966), quienes aplicaron técnicas de clustering a un grupo de broches que datan de la Edad de Hierro, encontrando una clasificación para los mismos de demostrada relevancia arqueológica. Otro ejemplo de la mano de Hodson (1971) es la aplicación del algoritmo K-medias (\acs{KM}, Apéndice \ref{ap:kmeans}) para construir una taxonomía de hachas de mano encontradas en las Islas Británicas. Las variables tenidas en cuenta para describir cada hacha incluyen longitud, anchura y valores en una escala que describen cómo de puntiaguda era la herramienta. El clustering dio como resultado dos grupos de hachas, uno formado por las pequeñas y delgadas, y otro formado por las grandes y gruesas.

Respecto a materiales fosilizados, Sutton y Reinhard (1995) realizaron un estudio sobre 155 coprolitos encontrados en \textit{Antelope House}, un yacimiento prehistórico en el Cañón de Chelly en Arizona. El estudio arrojó como resultado una interpretación de las diferencias entre coprolitos basada en la alimentación.

\subsection{Aplicaciones en bioinformática y genética}

Tiempos recientes están siendo testigo de un tremendo auge en el interés por la Bioinformática, acompañada por la biología molecular, ciencias de la computación, matemáticas y estadística. Tal aumento ha sido acelerado por la siempre creciente base de datos genómica y proteica, que son por sí mismas resultado de un grandísimo avance en las técnicas de secuenciación del ADN, medidas de expresión de los genes y compresión de las estructuras macromoleculares. La estadística ha resultado relevante en el estudio de la expresión de los genes. Los genes contenidos en el ADN de cada célula proporcionan las plantillas necesarias para la generación de las proteínas implicadas en la mayoría de los procesos estructurales y biomecánicos que tienen lugar en cada uno de nosotros. Sin embargo, aunque la mayoría de las células en los seres humanos contienen todos los complementos genéticos que componen el genoma humano, los genes se expresan de manera selectiva en cada célula dependiendo del tipo de la misma, del tejido y de las condiciones generales tanto dentro como fuera de la célula. La biología molecular ha puesto de manifiesto que la mayoría de los procesos en la vida de una célula están regulados por factores que afectan a la expresión de sus genes.

Como hemos visto, uno de los campos de investigación más activos hoy en día es el que estudia los procesos que regulan la expresión de los genes. Con el fin de almacenar la información relativa a esta área de estudio surgen los microarrays, (Cortesse, 2000). Desde el punto de vista del análisis de datos, una de las características relevantes en este tipo de información es que el número de atributos de cada instancia ($p$), supera con creces al número de instancias disponibles ($n$); conjuntos de datos como este son calificados como \textit{datos de alta dimensionalidad}.

La mayoría de métodos estadísticos clásicos no pueden ser aplicados a este tipo de conjuntos de datos sin ser modificados de forma sustancial. Sin embargo, el análisis de clusters acepta bien tales conjuntos de datos y puede ser empleado para identificar grupos de genes con patrones de expresión similares, y dar respuesta a preguntas como por qué un gen se ve afectado por cierta enfermedad, o qué genes son responsables de enfermedades genéticas hereditarias.

Un ejemplo de aplicación lo encontramos en el trabajo de Selinski e Ickstadt (2008), quienes usaron clustering sobre polimorfismos de nucleótidos simples para detectar diferencias entre enfermedades a nivel genético.

\section{Resumen}

El clustering consiste en la exploración de conjuntos de datos. Su objetivo es discernir si pueden o no ser resumidos de manera significativa, en términos de un número relativamente pequeño de grupos o clusters de objetos o individuos, que se parecen unos a otros y que se diferencian de los que se encuentran en otros clusters.

Muchas ramas de la ciencia han hecho uso de las técnicas de clustering de manera exitosa, para avanzar en sus respectivos campos, y procesar grandes cantidades de datos, cuyo análisis sería impensable afrontar con técnicas tradicionales.

%Para referir un acronimo completo \acf
%Para referir un acronimo solo con las siglas \acs

### 2019-06-18 Memoria/Chapters/Introduction.tex
Empezada memoria TFM

\chapter{Introduction}\label{ch:introduction}

\begin{quotation}{\slshape
		An intelligent being cannot treat every object it sees as unique entity unlike anything else in the universe. It has to put objects in categories so that it may apply its hard-won knowledge about similar objects encountered in the past, to he object at hand.}
		\begin{flushright}
			\textbf{Steven Pinker, How the Mind Works, 1997} 
		\end{flushright}
\end{quotation}

One of the most basic and primitive abilities that human beings are endowed with is that of grouping similar objects to produce a classification that is useful to them. For example, they should have been able to realize which objects were edible, which were poisonous, and which would try to kill them.

Classification skills, in their broadest sense, are necessary for language development, which is made up of words that help us recognize different types of events, actions and entities. In essence, each noun is a label that we use to group a collective of beings or objects with similar features, so that we can refer to them all using the word that unites them.

Just as classification is a basic skill for people in their daily lives, it is also essential in most branches of science. In biology, for example, the classification of different types of organisms has been the subject of study since the beginning of their existence. Aristotle built an elaborated animal classification system that divided all creatures into two groups, those with red blood and those without. Later he proposed a subdivision that classified them according to the way in which new individuals came into the world, either alive, by means of eggs, chrysalis, etc.

Following Aristotle, Theophrastus wrote the first document compiling guidelines for the classification of plants. The resulting works were so extensive and detailed that they laid the foundation for research in biology over the following centuries. This work was not replaced until 1737, when Carl von Linné published his work \textit{Genera Plantarum}, from which we extract the following fragment:

\begin{quotation}{\slshape
		All the real knowledge which we possess, depends on methods by which we distinguish the similar from the dissimilar. The greater the number of natural distinctions this method comprehends the clearer becomes our idea of things. The
		more numerous the objects which employ our attention the more difficult it becomes to
		form such a method and the more necessary.
		For we must not join in the same genus the horse and the swine, though both species
		had been one hoof’d nor separate in different genera the goat, the reindeer and the elk,
		tho’ they differ in the form of their horns. We ought therefore by attentive and diligent
		observation to determine the limits of the genera, since they cannot be determined a
		priori. This is the great work, the important labour, for should the genera be confused,
		all would be confusion.} 
		\begin{flushright}
			\textbf{Carl von Linné, Genera Plantarum, 1737}
		\end{flushright}
\end{quotation}

Classification of animals and plants has been of major importance in fields such as biology and zoology. In particular, this classification laid the foundation for the development of Darwin's theory of evolution. But it has also been of high relevance in knowledge areas such as chemistry and physics, with the classification of the elements in the periodic table, proposed by Mendeleyev in the 1860s; or in astronomy, with the classification of dwarf or giant stars using the Hertzsprung-Russell guidelines, shown in Figure \ref{fig:intro_HRdiagram}.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.35]{gfx/Intro/HR_diagram} 
	\caption[Hertzsprung-Russell starts classification diagram]{Hertzsprung-Russell starts classification diagram}\label{fig:intro_HRdiagram}
\end{figure}

\section{The learning task}

We humans, like many other beings, are capable of learning a multitude of concepts throughout our lives, although generally the learning process is extremely slow. It takes us about six years to be able to begin dedicated and specific learning in primary education, and another eighteen years---generally speaking---to become computer scientists or experts in any other branch of knowledge.

Thus, the human learning process is a slow and tedious process. So, do we want to teach machines to learn the way humans do? Machines have proven to be outstanding at repetitive tasks, as well as they reach calculation speeds that humans can only dream of. It is worth asking if it is possible to develop a learning method for machines that is orders of magnitude faster than the human and equal or more effective. Although there is a major impediment to the development of such a method: we do not know if it exists! The search for a method to avoid the disadvantages of the human learning process is what drives research in the science we know today as \acf{ML}.

One of the major differences between human learning and machine learning, and we must admit that in this the machines again lead us, is the way in which knowledge is transferred. In machines it is enough to copy and paste, that is to say, once a program performs correctly the task for which it has been designed in a computer, it performs it in all of them. Just copy and paste. In humans this is not the case, we cannot extract knowledge from one brain and insert it into another, we don't even know how it is encoded! We humans must use our inefficient input-output devices to transmit knowledge. The following quote, which well reflects the reality of automatic learning, is more than thirty years old, yet still valid:

\begin{quotation}{\slshape
		...we do have reason to search for machine learning programs that will avoid the inefficiencies of human learning, although we must be alert to the possibility that such programs cannot, in principle, be constructed. The difficulty may be intrinsic in the task; human learning, though slow, may be close to optimally efficient.}
	\begin{flushright}
		\textbf{Herbert A. Simon, Machine Learning: An Artificial Intelligence Approach, 1983} 
	\end{flushright}
\end{quotation}

\section{Machine Learning}

So far, references have been made to learning without formally defining it. The concept of learning is deeply anchored in human beings. However, when trying to apply learning to computers, which are still no more than glorified calculators, we must give some kind of definition: \textit{Learning denotes changes in the system that are adaptive in the sense that they enable the system to do the same task or tasks drawn from the same population more efficiently and more effectively the next time} \cite{Michalski:2013:MLA:2588013}.

\acs{ML} is a field derived from computer science, and a branch of \acf{AI}, whose ultimate goal is to develop methods that make it possible for machines to learn. In other words, the techniques embedded in \acs{ML} aim to develop programs that guide computers to learn from a set of examples. \acs{ML} is part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work and help them.

\subsection{Types of Machine Learning}

There are several approaches within the field of \acf{ML}, these are classified according to the information that is provided to the machine for it to learn. Thus, following the classification proposed in \cite{ayodele2010types}, we find four different types of learning:

\begin{itemize}

	\item \textbf{Supervised Learning}: complete information is provided for learning. The machine is provided with a set of examples and the class (label) associated with each of them. The goal is to generate a function that approximates the behavior of the function that maps the examples in their respective classes.

	\item \textbf{Unsupervised Learning}: in this case the class to which the examples belong to is not available. The goal is to extract the class from the information contained in the set of examples.

	\item \textbf{Semi-supervised Learning}: it arises from adding incomplete information to unsupervised learning. This information is usually given in the form of a subset of labeled data, but there is no restrictions to this concern.

	\item \textbf{Reinforcement Learning}: the algorithm learns how to act under the influence of its environment, which provides feedback on the actions taken by the algorithm.

\end{itemize}

Among the four main categories of machine learning methods we find the \acf{SSL}, which is the focus of this work. As we have already mentioned, it is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.

\section{Objetivos}

\section{Estructura}

### 2019-06-18 Memoria/FrontBackmatter/Abstract.tex
Empezada memoria TFM

\pdfbookmark[1]{Abstract}{Abstract}
\begingroup

\chapter*{Abstract}

\endgroup

### 2019-06-18 Memoria/FrontBackmatter/Declaration.tex
Empezada memoria TFM

%*******************************************************
% Declaration
%*******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
\textbf{JAVIER LEÓN PALOMARES} con DNI $77138857$P, actuando en su propio nombre y derecho, \\

\textbf{DECLARA, BAJO SU RESPONSABILIDAD:} \\

Que el Trabajo Fin de Máster presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con fecha \_\_ de julio de 2019, titulado <<\textit{\_\_}>> es original, no es copia ni adaptación de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona. \\

Y para que así conste a los efectos oportunos, firmo la presente en Granada, a \_\_ de julio de 2019.

\bigskip

%\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

Asimismo, yo, Javier León Palomares, alumno del Máster en Ciencia de Datos e Ingeniería de Computadores de la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con DNI $77138857$P, autorizo la ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del centro para que pueda ser consultada por las personas que lo deseen.

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

### 2019-06-18 Memoria/FrontBackmatter/Resumen.tex
Empezada memoria TFM

\pdfbookmark[1]{Resumen}{Resumen}
\begingroup

\chapter*{Resumen}

\endgroup

### 2019-06-18 Papers/DILS/paperDILS.tex
Empezada memoria TFM

@@
 This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to provide quiality results for the constrained clustering problem.
+We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to provide quality results for the constrained clustering problem.
 We will compare the results obtained by this proposal to those obtained by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
@@
 Metaheuristic algorithms are designed to explore the solution space associated with a problem guided by a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
-In this type of approaches the key is to find a good diversification-intensification tradeoff.
+In this type of approaches the key is to find a good exploration-exploitation tradeoff.
 This is the reason why the classic Local Search (LS) heuristic algorithm derived in the Iterative Local Search (ILS) metaheuristic algorithm, which introduces periodics perturbations in the exploration of the solution space to avoid local optima \cite{lourencco2010iterated}.
 ILS---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
-ILS is also very commonly used to design hybrid methods with good diversification-intensification tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
+ILS is also very commonly used to design hybrid methods with good exploration-exploitation tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
 
-In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptative and guided way.
+In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptive and guided way.
 We apply it to the constrained clustering problem, prooving that it constitute a good approach to obtain quality results.
@@
 The Dual Iterative Local Search (DILS) is a new variant of the classic ILS method.
-Its goal is to perform a search in the solution space to find the solution with the best fitness function $f$ value by introducing diversity in an adaptative and guided way.
+Its goal is to perform a search in the solution space to find the solution with the best fitness function $f$ value by introducing diversity in an adaptive and guided way.
 While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-introducing methods and to avoid local optima.
@@
 After that, it applies a strong mutation operator to the newly generated individual.
-This is the invidual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
+This is the individual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
 The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the acceptance criterion for DILS.
@@
 	\centering
-	\includegraphics[scale=0.25]{Figures/DILSDiagram.png}
+	\includegraphics[scale=0.25]{Figures/DILSDiagram.jpg}
 	\caption{Diagram summarizing DILS optimization process.}\label{img:DILS_diag}
@@
 	}
+
+	$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
+	\KwRet{$m_b$}
 \caption{DILS}\label{alg:DILS}
@@
 The uniform recombination consists of selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}.
-The segment mutation operator consists of replacing a fixed size segment from the features of the individual for randomly generated features CITA.
+The segment mutation operator consists of replacing a fixed size segment from the features of the individual for randomly generated features.
 The size of the segment $p_s$ must be fixed at the beginning of the optimization process.
@@
 Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, so each individual $m$ is a solution to the clustering problem.
-Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to intially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.
+Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to initially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.
 
@@
 Note that, in order to not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
-We also stablish the maximum number of neighbors $p_m$ that can be generated in each call, this is done to find a better diversification-intensification tradeoff.
+We also stablish the maximum number of neighbors $p_m$ that can be generated in each call, this is done to find a better exploration-exploitation tradeoff.
 
@@
 	\BlankLine
-	\KwRet ($m$)
+	\KwRet{$m$}
 
@@
 To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
-The implementation for the BRKGA algorith can be found in the same link as DILS$_{CC}$.
+The implementation for the BRKGA algorithm can be found in the same link as DILS$_{CC}$.
 In all cases the value for $k$ is equal to the number of classes displayed in Table \ref{tab:datasets}.
@@
 We can note that DILS$_{CC}$ represents a consistent improvement over all other 5 methods.
-This is due to the fact that the intensification capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lower level of constraint-based information we work with.
+This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lower level of constraint-based information we work with.
 
@@
 In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how DILS$_{CC}$ scales the quality of its results with the number of restrictions, as well as methods such as BRKGA or COPKM, the latter being more limited in other aspects.
-It is also woth noting how DILS$_{CC}$ provides the better average result for each of the three constraint set analyzed in this paper.
+It is also worth noting how DILS$_{CC}$ provides the better average result for each of the three constraint set analyzed in this paper.
 
@@
 All heatmaps suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle.
-Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strog evidence that DILS$_{CC}$ does represent a significan improvement over all other six methods, being te results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.
+Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that DILS$_{CC}$ does represent a significant improvement over all other six methods, being te results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.
 

### 2019-06-19 Memoria/Chapters/Clustering.tex
Escrito capitulo 2

@@
 
-Es importante destacar que, en la mayoría de las ocasiones, no hay un sólo criterio de clasificación para un mismo conjunto de datos, sino que existe una amplia variedad de los mismos.
-En el caso de las personas, podrían ser clasificadas, por ejemplo, en base a sus ingresos económicos, o según la cantidad de calorías que consumen a lo largo de un periodo de tiempo definido.
-Así, distintos criterios de clasificación no tienen por qué dar como resultado la misma división en grupos del conjunto a clasificar; de esta manera, diferentes criterios servirán a diferentes propósitos.
+It is important to note that, in most cases, there is no single classification criterion for the same data set, but rather a wide variety of data.
+In the case of individuals, they could be classified, for example, on the basis of their income, or according to the amount of calories they consume over a defined period of time.
+Thus, different classification criteria do not necessarily result in the same division into groups of the set to be classified; in that way, different criteria will serve different purposes.
 
-\section{Métodos numéricos para el Clustering}
+\section{Numerical Methods of Classification}
 
-Los métodos numéricos para el clustering surgen en ramas de las ciencias naturales, como la biología o la zoología, en un intento de eliminar la subjetividad implícita en el proceso de clasificación que se desencadena al descubrir una nueva especie.
-El objetivo es proporcionar un método no subjetivo y estable para clasificar y agrupar elementos.
+Numerical methods for clustering arise in branches of the natural sciences, such as biology or zoology, in an attempt to eliminate the subjectivity implicit in the classification process that is triggered by the discovery of a new species.
+The aim is to provide a non-subjective and stable method for classifying and grouping elements.
 
-Estos métodos adoptan diversos nombres que varían según el campo en el que se apliquen: taxonomía numérica (\textit{numerical taxonomy}) en biología, $Q$ análisis (\textit{$Q$ analysis}) en psicología, o reconocimiento de patrones no supervisado (\textit{unsupervised pattern recognition}) en el campo de la inteligencia artificial.
-No obstante, hoy en día, análisis de clusters (\textit{Clusters analysis}) o simplemente clustering son los términos más ampliamente aceptados y extendidos para referirse a tareas que involucran el descubrimiento de subgrupos dentro de un conjunto de elementos.
+These methods adopt different names that depend on the field in which they are applied: numerical taxonomy in biology, $Q$ analysis in psychology, or unsupervised pattern recognition in \acf{AI}.
+However, today, cluster analysis, or just clustering, are the most widely accepted and extended terms for tasks that involve the discovery of subgroups within a set of elements.
 
-En la mayoría de las aplicaciones del clustering, el objetivo es obtener una partición de los datos, en la que cada instancia u objeto pertenezca a un único cluster, y la unión de todos ellos contenga a todos los objetos individuales.
-Dicho esto, debe destacarse que, en algunas circunstancias, son aceptables soluciones en las que existe solapamiento entre clusters, así como el hecho de que puede no existir una partición aceptable de los datos.
+In most clustering applications, the objective is to obtain a data partition, in which each instance or object belongs to a single cluster, and the union of all of them contains all the individual objects.
+That said, it should be noted that, in some circumstances, solutions where there is overlap between clusters are acceptable, as well as the fact that there may not be an acceptable partition of the data.
 
-La manera más ampliamente extendida de representar la información sobre la que se debe aplicar clustering es una matriz $X$ de dimensión $n\times p$, en la que cada fila corresponde a una instancia u objeto a procesar, y cada columna corresponde a una de las variables que caracterizan dichas instancias u objetos.
-El término comúnmente aceptado para referirse a cada fila es el de \textit{vector de características}.
+In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
+In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
+A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
+In this study we will focus on partitional clustering.
+
+The most widely used way to represent the information on which clustering is applied is an array $X$ of dimension $n\times u$, in which each row corresponds to an instance or object to be processed, and each column corresponds to one of the variables that characterize those instances or objects.
+The commonly accepted term to refer to each row is \textit{features vector}, and the features vector set is referred as \textit{dataset}.
 
@@
 
-x_{1,1} & x_{1,2} & \cdots & x_{1,p} \\
+x_{1,1} & x_{1,2} & \cdots & x_{1,u} \\
 
@@
 
-x_{n,1} & \cdots & \cdots & x_{n,p} \\
+x_{n,1} & \cdots & \cdots & x_{n,u} \\
 
@@
 
-La entrada $x_{i,j}$ en $X$ se corresponde con el valor de la variable $j$-esima en la instancia $i$.
+The entry $x_{i,j}$ in $X$ corresponds to the value of the variable $j$th in the instance $i$.
+We will refer to the $i$th instance in $X$ as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$, which is a feature vector.
 
-Las variables en $X$ pueden ser una mezcla de atributos en un dominio continuo, discreto o categórico.
-Además, es posible que, en problemas reales, algunas entradas no estén disponibles.
-Esta mezcla de tipos de variables y los valores perdidos pueden complicar la tarea del clustering.
-Sin embargo, existen métodos para tratar estos casos, como la inferencia de valores perdidos o las transformaciones de dominio.
+We can define partitional clustering as the task of grouping the instances of the dataset $X$ into $k$ clusters, so that new information can be extracted from them.
+A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$.
+As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $\{c_i, \cdots, c_k\}$ to form a partition called $C$.
+The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement \cite{jain1999data}.
 
-En algunas aplicaciones, las filas de la matriz pueden contener medidas repetidas de la misma variable, aunque bajo diferentes condiciones, o en diferentes momentos, incluso en diferentes localizaciones espaciales.
-Un ejemplo de ello pueden ser las medidas de altura de un grupo de niños en un mismo mes a lo largo de diferentes años.
-Este tipo de datos posee una estructura que, de nuevo, puede complicar la tarea del clustering.
+Variables in $X$ can be a mixture of attributes in a continuous, discrete, or categorical domain.
+In addition, it is possible that, in real problems, some entries may not be available.
+This mix of variable types and missing values can complicate the clustering task.
+However, there are methods to deal with these cases, such as inference of missing values or domain transformations.
 
-Algunos métodos de clustering conllevan realizar transformaciones sobre la matriz $X$ para transformarla en una matriz de $n \times n$ en la que se almacenan medidas extraídas de la matriz $X$ que relacionen una instancia con todas las demás, como pueden ser la similitud, distancia o disimilitud.
+In some applications, matrix rows may contain repeated measurements of the same variable, although under different conditions, or at different times, even at different spatial locations.
+An example of this can be the height measurements of a group of children in the same month over different years.
+This type of data has a structure that, again, can complicate the task of clustering.
 
-El clustering es, dicho de manera simple, descubrimiento de grupos en datos, y no debe ser en ningún caso confundido con los métodos de discriminación o asignación, conocidos en el ámbito de la inteligencia artificial como aprendizaje supervisado.
-En ellos los grupos son conocidos a priori y el objetivo del análisis es obtener una regla de clasificación o clasificador que permita asignar nuevas instancias o individuos a uno de los grupos ya conocidos.
+Some clustering methods involve making transformations on the $X$ matrix to convert it into a $u\times u$ matrix in which measurements extracted from the $X$ matrix are stored that relate one instance to all the others, such as similarity, distance or dissimilarity.
 
-Una vez definida la estructura general de los métodos de clustering, estaría justificado preguntar, ¿qué es un cluster? La Sección \ref{QueEsCluster} intentará dar respuesta a esta pregunta.
+Clustering is, put simply, the discovery of groups in data, and should never be confused with methods of discrimination or assignment, known in \acs{AI} as Supervised Learning, as we have already mentioned in Section \ref{sec:TypesOfML}.
 
-\subsection{¿Qué es un cluster?} \label{QueEsCluster}
+Once the general structure of clustering methods has been defined, it would be justified to ask, what is a cluster? The Section \ref{sec:WhatIsCluster} will try to answer this question.
 
-Hasta este momento, los términos cluster, grupo y clase han sido empleados de una manera completamente intuitiva, sin necesidad alguna de definición formal, una prueba más de lo innato de estos conceptos en el ser humano.
-De hecho, dar una definición formal de cluster resulta una tarea, no sólo complicada, sino en muchas ocasiones poco útil.
-Bonner, por ejemplo, propuso en 1964 una definición de cluster completamente dependiente de la interpretación del usuario.
-En lo que a este autor respecta, un cluster es aquello que el usuario entiende como cluster sin haberle propuesto una definición formal del mismo.
+\subsection{What is a Cluster?} \label{sec:WhatIsCluster}
 
-Aunque la definición de Bonner es acertada en un amplio rango de situaciones, autores como Cormack, en 1971, o Gordon en 1999, proponen una definición algo más analítica desde el punto de vista matemático, definiendo un cluster en términos de cohesión interna (homogeneidad), y aislamiento externo (separación).
-La Figura \ref{fig:figure1} muestra de manera informal las propiedades descritas anteriormente, de forma que a cualquier observador le resultarán aparentes los clusters presentes en ella sin necesidad de una definición formal de cluster.
-Este hecho puede explicar por qué alcanzar una definición matemáticamente precisa de homogeneidad y separación puede llegar a ser innecesario.
+So far, the terms cluster, group and class have been used in a completely intuitive way, with no need for any formal definition, a further proof of the innate nature of these concepts in the human being.
+In fact, to give a formal definition of cluster is a task, not only complicated, but in many occasions not very useful.
+For example, in \cite{lance1967general} a cluster definition completely dependent on the user's interpretation is proposed by Bonner.
+As far as this definition is concerned, a cluster is what the user understands as a cluster without having proposed a formal definition of it.
 
-No queda completamente clara la manera en que las personas reconocen diferentes clusters cuando éstos son representados en un plano, pero una de las variables que con certeza influye es la distribución de distancias relativas entre los objetos o puntos.
+\begin{quotation}{\slshape
+		The value judgment of the user is the ultimate criterion for evaluating the meaning of these terms [cluster]. If using them produces an answer of value, no more need be asked of them.}
+	\begin{flushright}
+		\textbf{R.E. Bonner, On Some Clustering Techniques, 1964} 
+	\end{flushright}
+\end{quotation}
 
-\clearpage
+Although Bonner's definition is accurate in a wide range of situations, authors such as Cormack, in 1971 \cite{cormack1971review}, or Gordon in 1999 \cite{gordon:1999}, propose a somewhat more analytical definition from the mathematical point of view, defining a cluster in terms of internal cohesion---homogeneity---, and external isolation---separation.
+Figure \ref{fig:clusters_properties} informally illustrates the properties described above, so that any observer will find the clusters present in it apparent without the need for a formal cluster definition.
+This may explain why achieving a mathematically accurate definition of homogeneity and separation may become unnecessary.
 
@@
 	\myfloatalign
-	{\includegraphics[width=.3\linewidth]{imagenes/c2/TwoBasicsClusters}}
-	{\includegraphics[width=.3\linewidth]{imagenes/c2/MoonsBasics}}
-	{\includegraphics[width=.3\linewidth]{imagenes/c2/ThreeBasicClusters}}
-	\caption[Clusters con cohesión interna y/o aislamiento externo]{Clusters con cohesión interna y/o aislamiento externo.}\label{fig:figure1}
+	{\includegraphics[width=.3\linewidth]{gfx/Clustering/TwoBasicsClusters}}
+	{\includegraphics[width=.3\linewidth]{gfx/Clustering/MoonsBasics}}
+	{\includegraphics[width=.3\linewidth]{gfx/Clustering/ThreeBasicClusters}}
+	\caption[Clusters with internal cohesion and/or external isolation]{Clusters with internal cohesion and/or external isolation}\label{fig:clusters_properties}
 \end{figure}
 
-Por otra parte, como ya mencionamos anteriormente en esta sección, puede darse el caso de que en un conjunto de datos no exista una partición justificada.
-En la Figura \ref{fig:figure2} se muestra un conjunto de datos para el que la mayoría de observadores llegaría a la conclusión de que no existen grupos diferenciados, simplemente una nube de puntos uniformemente distribuidos.
-Idealmente, es de esperar que un método de clustering aplicado a este mismo conjunto de datos llegue a la misma conclusión.
+It is not entirely clear how people recognize different clusters when they are represented on a plane, but one of the variables that certainly influences is the distribution of relative distances between objects or points.
+
+On the other hand, as mentioned earlier in this section, it may be the case that there is no justified partition in a dataset.
+Figure \ref{fig:uniform_cloud} shows a dataset for which most observers would conclude that there are no distinct groups, just a cloud of uniformly distributed points.
+Ideally, one would expect a clustering method applied to this same dataset to reach the same conclusion.
 
@@
 	\centering
-	\includegraphics[scale=0.2]{imagenes/c2/rand.png} 
-	\caption{Distribución uniforme de puntos.}\label{fig:figure2}
+	\includegraphics[scale=0.2]{gfx/Clustering/rand.png} 
+	\caption{Cloud of uniformly distributed points.}\label{fig:uniform_cloud}
 \end{figure}
 
-Sin embargo, la mayoría de métodos de aprendizaje no supervisado darán como resultado un particionamiento uniforme como el que se muestra en la Figura \ref{fig:figure3}.
-El número de particiones encontradas dependerá del método aplicado, si bien en cualquier caso obtendremos un particionamiento uniforme.
+However, most unsupervised learning methods will result in uniform partitioning as shown in Figure \ref{fig:uniform_partition}.
+The number of partitions found will depend on the method applied, although uniform partitioning will be achieved in any case.
 
@@
 	\centering
-	\includegraphics[scale=0.2]{imagenes/c2/randClasif.png} 
-	\caption{Distribución uniforme de puntos clasificados.}\label{fig:figure3}
+	\includegraphics[scale=0.2]{gfx/Clustering/randClasif.png} 
+	\caption{Classified cloud of uniformly distributed points.}\label{fig:uniform_partition}
 \end{figure}
 
-El proceso de dividir una distribución homogénea de datos en diferentes grupos se conoce como disección, y tal proceso puede ser útil en ciertas circunstancias.
-Sin embargo, dado que en la mayoría de las ocasiones de aplicación real de métodos de clusters, no se conoce a priori la estructura de los datos, existe el riesgo de interpretar todas las soluciones en términos de existencia de subgrupos, lo que conllevaría la imposición de una estructura ficticia en datos en los que no hay estructura presente.
+The process of dividing a homogeneous distribution of data into different groups is known as dissection, and such a process may be useful in certain circumstances.
+However, since in most cases of actual application of cluster methods, the structure of the data is not known a priori, there is a risk of interpreting all the solutions in terms of the existence of subgroups, which would lead to the imposition of a fictitious structure on data in which there is no structure present.
 
-\section{Aplicaciones del clustering}
+\section{Applications of Clustering}
 
-Como ya se ha indicado, el problema general al que intenta dar solución el clustering está presente en muchas disciplinas: biología, botánica, medicina, psicología, geografía, marketing, procesamiento de imágenes, psiquiatría, arqueología, etc.
-En esta sección se presentan algunas de las aplicaciones del clustering relacionadas con los citados campos.
+As it has been already mentioned, the general problem clustering attempts to solve is present in many disciplines: biology, botany, medicine, psychology, geography, marketing, image processing, psychiatry, archaeology, etc.
+This section presents some of the clustering applications related to these various knowledge fields.
 
-\subsection{Aplicaciones en marketing}
+\subsection{Marketing}
 
-Dividir los clientes en grupos homogéneos es una de las tareas más frecuentes en marketing.
-Un director de marketing podría preguntarse cómo agrupar los posibles clientes según los beneficios potenciales del producto que intenta introducir en el mercado.
-Por otra parte, un analista de marketing podría estar interesado en agrupar las empresas según sus características financieras, para poder analizarlas y predecir sus estrategias de mercado.
+Dividing customers into homogeneous groups is one of the most common marketing tasks.
+A marketing manager might wonder how to group potential customers according to the potential benefits of the product he or she is trying to introduce to the market.
+On the other hand, a marketing analyst might be interested in grouping companies according to their financial characteristics, in order to be able to analyze them and predict their market strategies.
 
-Un ejemplo de aplicación del clustering en este ámbito fue publicado por Green et al.
-(1967).
-Así, con un gran número de ciudades disponibles para el análisis, debieron restringir los lugares en los que llevar a cabo sus estudios debido a motivos económicos.
-Para ello hicieron uso del análisis de clusters, clasificando las ciudades en pequeños grupos basándose en 14 características de las mismas, entre ellas el tamaño e ingresos medios \textit{per capita}.
-Dado que se esperaba que las ciudades incluidas en un mismo grupo fueran muy similares, escogieron una ciudad de cada uno de ellos para realizar sus estudios.
+An example of the application of clustering in this field was published in \cite{green1967cluster}.
+Thus, with a large number of cities available for analysis, they had to restrict the places in which to carry out their studies for economic reasons.
+For this they relied on cluster analysis, classifying cities into small groups based on 14 city characteristics, including size and average income \textit{per capita}.
+Since the cities in the same group were expected to be very similar, they chose one city from each of them to conduct their studies.
 
-Otra aplicación del análisis de clusters en el marketing fue descrita por Chakrapani (2004).
-En este caso, un fabricante de coches cree que comprar un coche deportivo no es una decisión basada sólo en capacidades económicas o edad, sino que es una decisión relacionada con el estilo de vida que llevan aquellos que deciden comprar un coche de estas características, frente a aquellos que no lo hacen.
-En consecuencia, el fabricante decide realizar un estudio, empleando análisis de clusters, que le permita identificar todas las características relacionadas con las personas que comprarían un coche deportivo, para así enfocar sus campañas de marketing a este sector específicamente.
+Another application of cluster analysis in marketing was described in \cite{chakrapani2004statistics}.
+In this case, a car manufacturer believes that buying a sports car is not a decision based solely on economic capabilities or age, but is a lifestyle decision made by those who decide to buy a sports car, as opposed to those who do not.
+Consequently, the manufacturer decides to carry out a study, using cluster analysis, which allows it to identify all the characteristics related to the people who would buy a sports car, in order to focus its marketing campaigns specifically on this sector.
 
-\subsection{Aplicaciones en astronomía}
+\subsection{Astronomy}
 
-Dado un conjunto de datos astronómicos, los investigadores quieren saber, por ejemplo, cuantas clases de estrellas hay presentes en ellos, basándose en algun criterio estadístico.
-Las preguntas más frecuentes dentro de este ámbito son: ¿Cuantos objetos estadísticamente diferentes están presentes en los datos y a qué clase debe ser asignado cada objeto?
-¿Aparecen clases de objetos previamente desconocidas?.
-El análisis de clusters puede ser aplicado para dar respuesta a estas cuestiones, ayudando a detectar objetos estadísticamente anómalos, así como a guiar el proceso de clasificación de los mismos.
-Algunos ejemplos incluyen el descubrimiento de quasars con alto corrimiento al rojo, quasars de tipo 2 (altamente luminosos, núcleos galácticos activos a menudo oscurecidos por polvo y gas), y enanas marrones.
+Given a set of astronomical data, researchers want to know, for example, how many classes of stars are present in them, based on some statistical criteria.
+The most frequently asked questions in this area are: How many statistically different objects are present in the data and to which class should each object be assigned?
+Do previously unknown classes of objects appear?
+Cluster analysis can be applied to answer these questions, helping to detect statistically anomalous objects, as well as to guide the process of classifying them.
+Examples include the discovery of high redshift quasars, type 2 quasars (highly luminous, active galactic nuclei often obscured by dust and gas), and brown dwarfs.
 
-Un ejemplo específico viene dado por el estudio de Faúndez-Abans et al.
-(1996), quienes aplicaron técnicas de clustering a datos sobre la composición química de 192 nebulosas planetarias.
-Se identificaron 6 grupos diferentes que eran similares en muchos aspectos a una clasificación previa de dichos objetos, pero que también mostraban diferencias interesantes que hasta ese momento los investigadores habían pasado por alto.
+In \cite{faundez1996classification} a specific example of the above can be found.
+Clustering techniques are applied to data on the chemical composition of 192 planetary nebulae.
+Six different groups were identified that were similar in many respects to a previous classification of such objects, but also showed interesting differences that researchers had so far missed.
 
-Un segundo ejemplo lo encontramos en el estudio de Celeux y Govaert (1992), quienes aplicaron clustering basado en distribuciones normales a un conjunto de 2370 estrellas, descritas por su velocidad relativa al núcleo galáctico y a la rotación galáctica.
-Usando un modelo de tres clusters, encontraron un cluster de gran tamaño y pequeño volumen, y dos de pequeño tamaño y gran volumen.
+Clustering based on normal distributions is applied to a set of 2370 stars in \cite{celeux1992classification}.
+They were described by their speed relative to the galactic nucleus and galactic rotation.
+Using a model of three clusters, they found one cluster of large size and small volume, and two of small size and large volume.
 
-\subsection{Aplicaciones en psiquiatría}
+\subsection{Psychiatry}
 
-Las enfermedades de la mente son a menudo más difíciles de diagnosticar que las enfermedades del cuerpo; es por ello que en el campo de la psiquiatría ha crecido el interés por las técnicas de análisis de clusters que permitan refinar, o incluso redefinir, las técnicas de diagnosis para este tipo de enfermedades.
-Gran parte de este trabajo involucra pacientes deprimidos, casos en los que el interés reside en distinguir entre dos tipos de depresión, la endógena (congénita), y la neurótica.
+Mind disorders are often more difficult to diagnose than body pathologies, which is why in the field of psychiatry there has been a growing interest in cluster analysis techniques to refine, or even redefine, diagnostic techniques for this type of disease.
+Much of this work involves depressed patients, cases in which the interest lies in distinguishing between two types of depression, endogenous (congenital) and neurotic.
 
-Pilowsky et al.
-(1968), por ejemplo, usando métodos desarrollados por otros autores, aplicaron técnicas de clustering a 200 pacientes en base a sus respuestas a un cuestionario sobre la depresión, junto a información sobre su sexo, edad, estado mental y enfermedad padecida.
-Éste es un claro ejemplo de variables de diferentes tipos incluidas en el mismo conjunto de datos.
-Uno de los grupos obtenidos como resultado de este estudio fue identificado como marcador de la depresión endógena.
+In \cite{pledger2008using} clustering techniques are applied to 200 patients based on their responses to a questionnaire about depression, along with information about their sex, age, mental state, and illness.
+This is a clear example of different types of variables included in the same data set.
+One of the groups obtained as a result of this study was identified as a marker of endogenous depression.
 
-El análisis de clusters también ha sido empleado para encontrar una clasificación de individuos que intentaron cometer suicidio, que podría sentar las bases para estudios posteriores sobre las causas y tratamientos del problema.
-Paykey y Rassaby (1978), por ejemplo, estudiaron 236 casos de suicidas fallidos registrados por el servicio de emergencias de una ciudad de los Estados Unidos de América.
-Del conjunto de posibles variables, 14 fueron seleccionadas como particularmente relevantes para la clasificación y, por tanto, fueron usadas en el análisis.
-Entre ellas se encontraban: edad, número de intentos de suicidio, gravedad de la depresión y grado de hostilidad, además de una serie de características demográficas.
-Al conjunto de datos resultante se le aplicó métodos de clustering; el resultado más significativo obtenido corresponde a una división en tres clusters bien definidos.
+Cluster analysis has also been used to find a classification of individuals who attempted suicide, which could lay the groundwork for further studies on the causes and treatments of the problem.
+In \cite{paykel1978classification} a study on 236 cases of failed suicides recorded by the emergency service of a city in the United States of America is presented.
+Of the set of possible variables, 14 were selected as particularly relevant for classification and, therefore, were used in the analysis.
+Among them were: age, number of suicide attempts, severity of depression and degree of hostility, in addition to a series of demographic characteristics.
+Clustering methods were applied to the resulting data set; the most significant result obtained corresponds to a division into three well-defined clusters.
 
@@
 
-Diariamente se recogen enormes cantidades de datos sobre la meteorología mundial.
-Explorar estos datos mediante técnicas de clustering puede aportar nuevos enfoques para la climatología y el medio ambiente.
+Huge amounts of global weather data are collected daily. Exploring these data using clustering techniques can bring new approaches to study climatology and the environment.
 
-Littmann (2000), por ejemplo, aplicó clustering a los datos recogidos sobre los cambios diarios en la presión superficial en la cuenca Mediterránea, y encontró 20 grupos que explicaban la varianza de las lluvias en las regiones centrales del Mediterráneo.
-Otro ejemplo viene de la mano de Liu y George (2005), quienes usaron el algoritmo K-medias difuso (\textit{Fuzzy K-means}, \acs{FKM}) con datos espaciotemporales de la climatología de las regiones del sur central de EEUU.
+In \cite{littmann2000empirical} clustering techniques are applied to data collected on daily changes in surface pressure in the Mediterranean basin, and found 20 groups explaining the variance of rainfall in the central Mediterranean regions.
+Another example can be found in \cite{liu2005mining}, in which the Fuzzy K-means algorithm is applied to spatial-temporal data of the climatology of the southern-central regions of the USA.
 
-\subsection{Aplicaciones en arqueología}
+\subsection{Archaeology}
 
-La arqueología es otra de las disciplinas en la que resulta útil la aplicación del clustering.
-La clasificación de los diferentes objetos encontrados en los yacimientos puede ayudar a descubrir su uso, los periodos a los que pertenecen, así como la población que los utilizó.
-De forma similar, el estudio de materiales fosilizados puede ayudar a revelar como vivieron las sociedades prehistóricas.
+Archaeology is another discipline in which clustering is useful.
+The classification of the different objects found in the deposits can help to discover their use, the periods to which they belong, as well as the population that used them.
+Similarly, studying fossilized materials can help reveal how prehistoric societies lived.
 
-Un ejemplo temprano de la aplicación de clustering a objetos arqueológicos viene dado por Hodson et al.
-(1966), quienes aplicaron técnicas de clustering a un grupo de broches que datan de la Edad de Hierro, encontrando una clasificación para los mismos de demostrada relevancia arqueológica.
-Otro ejemplo de la mano de Hodson (1971) es la aplicación del algoritmo K-medias (\acs{KM}, Apéndice \ref{ap:kmeans}) para construir una taxonomía de hachas de mano encontradas en las Islas Británicas.
-Las variables tenidas en cuenta para describir cada hacha incluyen longitud, anchura y valores en una escala que describen cómo de puntiaguda era la herramienta.
-El clustering dio como resultado dos grupos de hachas, uno formado por las pequeñas y delgadas, y otro formado por las grandes y gruesas.
+An early example of the application of clustering to archaeological objects is given in \cite{hodson1966some}, where clustering techniques are applied to a group of brooches dating from the Iron Age, finding a classification for them of proven archaeological relevance.
+Another example found in \cite{hodson1971numerical} is the application of the K-means algorithm to construct a taxonomy of hand axes found in the British Isles.
+The variables taken into account to describe each axe include length, width and pointedness.
+Clustering resulted in two groups of axes, one made up of small and thin axes and the other made up of large and thick axes.
 
-Respecto a materiales fosilizados, Sutton y Reinhard (1995) realizaron un estudio sobre 155 coprolitos encontrados en \textit{Antelope House}, un yacimiento prehistórico en el Cañón de Chelly en Arizona.
-El estudio arrojó como resultado una interpretación de las diferencias entre coprolitos basada en la alimentación.
+Regarding fossilized materials, in \cite{sutton1995cluster} a study of 155 coprolites found at Antelope House, a prehistoric site at Chelly Canyon in Arizona is conducted.
+The study resulted in a diet-based interpretation of the differences between coprolites.
 
@@
 
-Tiempos recientes están siendo testigo de un tremendo auge en el interés por la Bioinformática, acompañada por la biología molecular, ciencias de la computación, matemáticas y estadística.
-Tal aumento ha sido acelerado por la siempre creciente base de datos genómica y proteica, que son por sí mismas resultado de un grandísimo avance en las técnicas de secuenciación del ADN, medidas de expresión de los genes y compresión de las estructuras macromoleculares.
-La estadística ha resultado relevante en el estudio de la expresión de los genes.
-Los genes contenidos en el ADN de cada célula proporcionan las plantillas necesarias para la generación de las proteínas implicadas en la mayoría de los procesos estructurales y biomecánicos que tienen lugar en cada uno de nosotros.
-Sin embargo, aunque la mayoría de las células en los seres humanos contienen todos los complementos genéticos que componen el genoma humano, los genes se expresan de manera selectiva en cada célula dependiendo del tipo de la misma, del tejido y de las condiciones generales tanto dentro como fuera de la célula.
-La biología molecular ha puesto de manifiesto que la mayoría de los procesos en la vida de una célula están regulados por factores que afectan a la expresión de sus genes.
+Recent times are witnessing a tremendous boom in interest in Bioinformatics, complemented by molecular biology, computer science, mathematics and statistics.
+Such an increase has been accelerated by the ever-growing database of genomic and protein data, which are themselves the result of a major advance in DNA sequencing techniques, gene expression measurements, and compression of macromolecular structures.
+Statistics have been relevant in the study of gene expression.
+The genes contained in the DNA of each cell provide the necessary templates for the generation of the proteins involved in most of the structural and biomechanical processes that take place in each of us.
+However, although most cells in humans contain all the genetic components that make up the human genome, genes are expressed selectively in each cell depending on cell type, tissue and general conditions inside and outside the cell.
+Molecular biology has shown that most of the processes in a cell's life are regulated by factors that affect the expression of its genes.
 
-Como hemos visto, uno de los campos de investigación más activos hoy en día es el que estudia los procesos que regulan la expresión de los genes.
-Con el fin de almacenar la información relativa a esta área de estudio surgen los microarrays, (Cortesse, 2000).
-Desde el punto de vista del análisis de datos, una de las características relevantes en este tipo de información es que el número de atributos de cada instancia ($p$), supera con creces al número de instancias disponibles ($n$); conjuntos de datos como este son calificados como \textit{datos de alta dimensionalidad}.
+As we have seen, one of the most active fields of research today is the study of the processes that regulate gene expression.
+In order to store the information related to this area of study, microarrays arise \cite{cortese2000array}.
+From the data analysis point of view, one of the relevant features in this type of information is that the number of attributes of each instance ($u$) far exceeds the number of available instances ($n$); data sets such as this are qualified as \textit{high dimensionality data}.
 
-La mayoría de métodos estadísticos clásicos no pueden ser aplicados a este tipo de conjuntos de datos sin ser modificados de forma sustancial.
-Sin embargo, el análisis de clusters acepta bien tales conjuntos de datos y puede ser empleado para identificar grupos de genes con patrones de expresión similares, y dar respuesta a preguntas como por qué un gen se ve afectado por cierta enfermedad, o qué genes son responsables de enfermedades genéticas hereditarias.
+Most classical statistical methods cannot be applied to this type of dataset without being substantially modified.
+However, cluster analysis accepts such data sets well and can be used to identify groups of genes with similar expression patterns, and answer questions such as why a gene is affected by a certain disease, or which genes are responsible for hereditary genetic diseases.
 
-Un ejemplo de aplicación lo encontramos en el trabajo de Selinski e Ickstadt (2008), quienes usaron clustering sobre polimorfismos de nucleótidos simples para detectar diferencias entre enfermedades a nivel genético.
+An example of application is found in \cite{selinski2008cluster}, who used clustering on single-nucleotide polymorphisms to detect differences between diseases at the genetic level.
 
-\section{Resumen}
+\section{Recent Applications}
 
-El clustering consiste en la exploración de conjuntos de datos.
-Su objetivo es discernir si pueden o no ser resumidos de manera significativa, en términos de un número relativamente pequeño de grupos o clusters de objetos o individuos, que se parecen unos a otros y que se diferencian de los que se encuentran en otros clusters.
+Although the examples of practical clustering applications presented so far are illustrative, it should be noted that clustering techniques are applied in a multitude of modern problems.
+Since it is not humanly possible to cover all the practical uses of clustering, a brief list of some of them is presented here: cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
+\textcolor{red}{Poner mas ejemplos de aplicación recientes}.
+It seems clear that clustering is an extremely powerful tool, no only for the development of any branch of science, but also because it can be applied to directly help people though medicine or society management.
 
-Muchas ramas de la ciencia han hecho uso de las técnicas de clustering de manera exitosa, para avanzar en sus respectivos campos, y procesar grandes cantidades de datos, cuyo análisis sería impensable afrontar con técnicas tradicionales.
+\section{Summary}
+
+Clustering consists of exploring datasets.
+Its goal is to determine whether or not they can be summarized in a meaningful way, in terms of a relatively small number of groups or clusters of objects or individuals which resemble each other and differ from those found in other clusters.
+
+Many branches of science have successfully made use of clustering techniques to advance in their respective fields and to process large amounts of data, whose analysis would be unthinkable to confront with traditional techniques.
 

### 2019-06-19 Memoria/FrontBackmatter/Legality.tex
Escrito capitulo 2

\chapter*{}
\thispagestyle{empty}

%\noindent\rule[-1ex]{\textwidth}{2pt}\\[4.5ex]

Yo, \textbf{Germán González Almagro}, alumno de la titulación en ingeniería informática de la \textbf{Escuela Técnica Superior
de Ingenierías Informática y de Telecomunicación de la Universidad de Granada}, con DNI $\mathrm{76593910}$T, autorizo la
ubicación de la siguiente copia de mi Trabajo Fin de Grado en la biblioteca del Centro para que pueda ser
consultada por las personas que lo deseen.

\vspace{0.1cm}

\begin{flushleft}
	Granada, 17 de junio de 2018.
\end{flushleft}

\vspace{2.5cm}

\begin{flushright}

	\begin{tabular}{m{6cm}}
		\\ \hline
		\centering\textbf{Fdo. \myName} \\
	\end{tabular}

\end{flushright}

\chapter*{}
\thispagestyle{empty}

%\noindent\rule[-1ex]{\textwidth}{2pt}\\[4.5ex]

D. \textbf{\myProf}, Profesor del Departamento Ciencias de la Computación en Inteligencia Artificial de la Universidad de Granada.

\vspace{0.5cm}

\textbf{Informa:}

\vspace{0.5cm}

Que el presente trabajo, titulado \textit{\textbf{<<Clustering con Restricciones: un marco unificado>>}},
ha sido realizado bajo su supervisión por \textbf{Germán González Almagro}, y autorizamos la defensa de dicho trabajo ante el tribunal
que corresponda.

\vspace{0.5cm}

Y para que conste, expiden y firman el presente informe en Granada a 17 de junio de 2018.

\vspace{1cm}

\textbf{El director:}

\vspace{5cm}

\begin{table}
	\begin{minipage}{.4\textwidth}
		\centering
		\begin{tabular}{c}
			\\ \hline 
			\textbf{\myProf}\\
		\end{tabular}
	\end{minipage}
\end{table}

### 2019-06-22 Memoria/Chapters/ConstrainedClustering.tex
Capitulo tres de la memoria empezado

\chapter{Constrained Clustering}\label{ch:ConstrainedClustering}

Having introduced the basic concepts of clustering, the aim of this chapter is to define the concepts related to constrained clustering, as well as to give examples of its application and highlight the benefits and problems presented by its use. To this end we take as our main reference the survey presented in \cite{davidson2007survey}.

\section{Motivation for Constrained Clustering}

As we discussed in Chapter \ref{ch:IntroClustering}, unsupervised clustering methods are useful for structuring data referring to a particular area. An example of this can be found in text classification; In \cite{cohn2003semi} a problem proposed by Yahoo! is faced. The problem consist of, given a large number of text documents, grouping them according to a taxonomy in which documents with similar topics are nearby. For this, unsupervised clustering methods are useful, as the information on the problem initially available is limited. However, in \cite{wagstaff2001constrained} is shown that applying unsupervised clustering to certain problems, such as grouping GPS data in such a way that clusters define the lanes of a road, does not produce significant results, as the clusters obtained are far from the elongated shape that would be expected as a result. To tackle the problem, they introduced a new element in clustering, the instance-level constraints, which made it possible to include knowledge about the clusters that would guide the clustering methods to obtain the expected results. It was enough to indicate that the lanes of the road on which the vehicles circulate are four meters wide, and therefore any vehicle that is at a distance of more than 4 meters from another, perpendicular to the direction of displacement, must be located in a different cluster.

We are now in a new scenario: it is possible to incorporate additional information to the clustering process, in addition to that contained in the data set itself, to guide it in the formation of the partition and obtain more precise results. This places constrained clustering in the framework of semi-supervised learning, unlike traditional clustering methods which fall within the area of unsupervised clustering.

\section{Constraints Definition}

The new type of information that we incorporate to clustering is given in the form of instance-level constraints, that is, to specify if two instances ($x_i$ and $x_j$) from the dataset ($X$) must be in the same cluster or, on the contrary, they must be in separate clusters.

Constraints that indicate that two instances must be placed in the same cluster are called \acf{ML}, and are noted with $C_=(x_i,x_j)$. Similarly, constraints that specify that two instances must not be placed in the same cluster are called \acf{CL}, and are noted with $C_\neq(x_i,x_j)$ \cite{wagstaff2000clustering}.

Although they may seem simple, constraints defined as above have some very interesting properties. Must-link constraints are an example of an equivalence relationship, and therefore are symmetrical, reflexive and transitive, formalizing:

\begin{observation}
	\textbf{Must-link constraints are transitive.} Let $CC_i$ and $CC_j$ be connected components (completely connected subgraphs by \acs{ML} constraints), and let $x_i$ and $x_j$ be the instances in $CC_i$ and $CC_j$ respectively. Then $C_=(x,y): x \in CC_i, y \in CC_j \rightarrow C_=(a,b) \forall a,b: a\in CC_i, b \in CC_j$. \cite{davidson2007survey}
\end{observation}

\begin{observacion}
	\textbf{Cannot-link constraints can be entailed.} Let $CC_i$ and $CC_j$ be connected components (completely connected subgraphs by \acs{CL} constraints) and let $x_i$ and $x_j$ be the instances in $CC_i$ and $CC_j$ respectively. Then $C_\neq(x,y): x \in CC_i, y \in CC_j \rightarrow _\neq(a,b) \forall a,b: a\in CC_i, b \in CC_j$. \cite{davidson2007survey}
\end{observacion}

A clear example of constraint use can be found in clustering applications where there are distance measurement limitations, as in the case of GPS data. Thus, if we want that instances forming two clusters to be separated by a distance greater or equal to $\delta$, it is enough to set \acs{ML} constraints between all instances separated by less than $\delta$ units. Similarly, if we want the diameter of the clusters to be at most $\epsilon$, we must set \acs{CL} constraint between instances separated by more than $\epsilon$ units. Figure \ref{fig:DistanceConstraints} shows a graphical representation of these two types of constraints.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.45]{gfx/ConstClust/RestriccionesDeltaEpsilon.png} 
	\caption[\textit{delta} and \textit{epsilon} constraints]{\textit{delta} and \textit{epsilon} constraints \cite{davidson2007survey}.}\label{fig:DistanceConstraints}
\end{figure}

\section{Use of Constraints}

While supervised learning involves knowing the label associated with each instance, \acf{SSL} has only a subset of tagged instances. On the other hand, in many domains the available information refers to relations between instances, and not to the specific class to which they belong. Moreover, in interactive clustering systems, a user who is not expert in the domain of the problem will probably be able to provide information in the form of constraints such as \acf{ML} and \acf{CL}. \cite{cohn2003semi}\cite{davidson2007hierarchical}, rather than providing information on what particular class certain instances belong to.

Typically, constraints are incorporated into clustering problems in two ways. They can be used to alter the instance clusters assignment rule of the method in hand, so that the solution satisfies as many constraints as possible. Alternatively, it is possible to train the distance function used by the method based on the constraints, either before or during the application of the method. In any case, the initialization phase can take the constraints into account, so that instances associated with constraints \acf{ML} will be placed in the same cluster, and those between which there is a constraint \acf{CL} will be placed in different clusters. Based on this distinction, we identify two ways of approaching the problem, those based on constraints (\textit{constraint-based}), and those based on distances (\textit{distance-based}).

\subsection{Constraints-based Methods}

In constraint-based methods, the clustering method itself is modified in such a way that the available information is used to bias the search and obtain an appropriate data partition.

Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained}\cite{davidson2005agglomerative} and soft \cite{law2004clustering}\cite{basu2004active}\cite{segal2003discovering}\cite{davidson2005clustering}\cite{law2005model} constraints. Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}. For the purposes of this work, we will employ the latter. There are several techniques to obtain a partition based on constraints:

\begin{itemize}

	\item  Modify the objective function to include a penalty term for breaking constraints \cite{demiriz1999semi} \cite{davidson2005clustering}.

	\item Group instances with additional information obtained from a conditional distribution into an auxiliary space \cite{sinkkonen2000semisupervised}.

	\item Enforcing all constraints to be satisfied by modifying the the instance clusters assignment rule \cite{wagstaff2001constrained}.

	\item Initialize clusters based on constraints inferred from a set of labeled instances \cite{basu2002semi}.

\end{itemize}

Figure \ref{fig:ConstOverDataset} shows a data set together with its associated constraints. Figure \ref{fig:ClusteringSatAllConst} proposes a possible clustering that satisfies all constraints.

\begin{figure}[bth]
	\myfloatalign
	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/InputInstancesAndConst1}
	\caption[Constraints on a dataset.]{Constraints on a dataset. \cite{davidson2007survey}} \label{fig:ConstOverDataset}
	}
	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/ClusteringSatAll}
	\caption[Clustering satisfying all constraints.]{Clustering satisfying all constraints. \cite{davidson2007survey}} \label{fig:ClusteringSatAllConst}
	}
\end{figure}

\subsection{Distance-based Methods}

In distance-based approaches, classic clustering methods that use a distance measurement are employed, so that the distance measurement is modified to incorporate the constraints. In this context, satisfying constraints means that instances related to constraints \acf{ML} are placed together in the space, and those related by \acf{CL} are separated.

Figure \ref{fig:MetricLearned} shows a possible clustering based on a metric learned from the constraints specified in Figure \ref{fig:ConstOverDataset2}. It should be noted that in Figure \ref{fig:MetricLearned} the space in which the data is located has been compressed on the vertical axis and widened on the horizontal axis to match the learned distance metric.

\begin{figure}[bth]
	\myfloatalign
	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/InputInstancesAndConst2}
	\caption[Constraints on a dataset.]{Constraints on a dataset. \cite{davidson2007survey}} \label{fig:ConstOverDataset2}
	}
	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/MetricaAprendida}
	\caption[Clustering based on a metric learned based on constraints.]{Clustering based on a metric learned based on constraints. \cite{davidson2007survey}} \label{fig:MetricLearned}
	}
\end{figure}

\section{Constrained Clustering Applications} \label{sec:CCApplications}

This section shows some application cases in which constrained clustering has turned out to be a more useful tool than unsupervised clustering. For each case we will analyze how the constraints were obtained and how they improve the results in the resulting clustering.

\subsection{Image Analysis}

Figure \ref{fig:CMUFacesDatabase} shows a sample from the CMU (Carnegie Mellon University) faces dataset, where the task is to group faces based on different criteria. In this case, the goal is to group the faces according to their orientation.

\clearpage

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Profile faces.]
	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Caras1}} \quad
	\subfloat[Front faces.]
	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Caras3}} \quad
	\subfloat[Faces up.]
	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Caras2}} \quad
	\caption[CMU database faces.]{CMU database faces. \cite{davidson2007survey}}\label{fig:CMUFacesDatabase}
\end{figure}

The method used to obtain the constraints is one of the most popular in the literature: set the number of clusters of the resulting partition equal to the number of classes in the database, and generate the constraints from a subset of tagged instances; that is, if two instances have different labels, set a \acf{CL} constraint between them, otherwise one of \acf{ML} type. Thus, between the images shown in Figure \ref{fig:CMUFacesDatabase} \acf{CL} constraints are set, since, although they belong to the same person, they do not have the same orientation.

\begin{figure}[bth]
	\myfloatalign
	{\includegraphics[width=.35\linewidth]{gfx/ConstClust/AnalisisImagenes/CarasDifOr1}} \quad
	{\includegraphics[width=.35\linewidth]{gfx/ConstClust/AnalisisImagenes/CarasDifOr2}}
	\caption[\acs{CL} constraints between faces of the same person.]{\acs{CL} constraints between faces of the same person. \cite{davidson2007survey}}\label{fig:figure10}
\end{figure}

Figure \ref{fig:AiboRoboRobtoClustSys} shows another set of image data on which constrained clustering techniques are applied. In this case, the task is to perform object recognition to incorporate the method into the navigation system of the Aibo robot \cite{davidson2005clustering}. Distance constraints such as $\delta$ and $\epsilon$ are used as described in Figure \ref{fig:DistanceConstraints}. In this way well differentiated clusters are obtained and therefore they are useful for the path-finding techniques performed by the robot during navigation.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[Imagen original]
	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo1}} \quad
	\subfloat[Clustering sin restricciones]
	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo2}} \quad
	\subfloat[Clustering con restricciones]
	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo3}} \quad
	\caption[Clustering method used in Aibo Robot navigation system.]{Clustering method used in Aibo Robot navigation system. \cite{davidson2007survey}\cite{davidson2005clustering}}\label{fig:AiboRobtoClustSys}
\end{figure}

\subsection{Video Analysis}

Video databases are one example where constraints can be generated directly from the data domain, especially if space-time data information is available \cite{yan2006discriminative}. In time sequenced data it is possible to set \acf{ML} constraints between groups of pixels of frames close in time. This is especially useful when the task is to perform object recognition based on clustering and segmentation. It is also possible to add \acf{CL} constraints to image segment in the same snapshot, as there is a low probability that they are associated with the same object after segmentation is done. In fact, in the video analysis domain problems there are a variety of constraint extraction methods \cite{yan2006discriminative}, Figure \ref{fig:ConstExtractionVideo} shows some examples.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[]
	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoA}}
	\quad
	\subfloat[]
	{\includegraphics[width=.225\linewidth]{gfx/ConstClust/Videos/VideoB}} \quad
	\subfloat[]
	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoC}}
	\quad
	\subfloat[]
	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoD}}
	\caption[Constraints extraction in video data.]{Constraints extraction in video data. \cite{yan2006discriminative} \cite{davidson2007survey}}\label{fig:ConstExtractionVideo}
\end{figure}

In Figure \ref{fig:ConstExtractionVideo}, image (a) corresponds to constraints extracted from tracking a person over a period of time, (b) corresponds to spatial constraints associating two objects located in the same frame, image (c) corresponds to restrictions obtained through facial recognition and (d) to those provided by the user.

With so many constraint extraction methods, one might ask: what happens if too many constraints are imposed? Does this make the problem over-constrained? In Section \ref{sec:CCDisadvantages} we will address these questions.

\subsection{Biological Data}

In gene clustering based on micro-array data, genes are represented by their expression profile in different experiments and grouped using different methods, in this case constrained clustering methods. Figure \ref{fig:GeneticApp} shows an example: these are \acf{ML} constraints set between genes based on co-occurrence data stored in the protein interaction database, which contains information about which genes (and their associated proteins) are present in the same cell processes \cite{xenarios2001dip}. This information can be used to improve results provided by clustering methods. \cite{segal2003discovering}.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.3]{gfx/ConstClust/Genetica/Genes} 
	\caption[Gene clustering based on micro-array data.]{Gene clustering based on micro-array data. \cite{davidson2007survey}}\label{fig:GeneticApp}
\end{figure}

\subsection{Text Analysis}

In content classification tasks, the goal is to automatically split large amounts of documents into groups or clusters. In this case it is possible to extract constraints from multiple auxiliary resources. For example, if two documents are in the same directory, a \acf{ML} constraints could be set between them. This way it is possible to adjust the resulting clustering to suit a particular criterion, such as creating a hierarchy of documents similar to the way they are organized in the input directories structure.

\subsection{Web Data}

El clustering con restricciones resulta de gran utilidad en el procesamiento de datos de búsqueda en páginas web. Aquí, el objetivo es agrupar, de manera automática, los resultados de una consulta ambigua en el motor de búsqueda en clusters de \acs{URL}s, que se refieran al concepto introducido como consulta en diferentes contextos. En este ámbito es posible extraer las restricciones a partir de búsquedas realizadas anteriormente por los usuarios, de manera que se establece una restricción de tipo \acf{ML} entre \acs{URL}s visitadas en la misma sesión de usuario. Aplicar clustering utilizando estas restricciones puede ayudar a sesgar el resultado de las búsquedas hacia las preferencias del usuario.

\subsection{Aplicaciones en datos de audio}

Constrained clustering is very useful in the processing of search data on web pages. Here, the aim is to automatically group the results of an ambiguous query in the search engine into clusters of \acs{URL}s, which refer to the concept introduced as a query in different contexts. In this area it is possible to extract constraints from searches previously performed by users, so that a \acf{ML} constraint is set between visited \acs{URL}s in the same user session. Clustering using these constraints can help biasing search results towards user preferences.

\subsection{Aplicaciones en datos de GPS} \label{EjemploGPS}

As mentioned at the beginning of Chapter \ref{ch:ConstrainedClustering}, constrained clustering can be applied to \acs{GPS} data to identify the lane in which each vehicle is running, as shown in Figure \ref{fig:GPSClustData}. Each instance is represented by the position it occupies on the road in two-dimensional Cartesian coordinates $(x,y)$, obtained from \acs{GPS} data. Figure \ref{fig:figure15} graphically shows this representation of the data (note that multiple instances can refer to the same vehicle at different times).

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.3]{gfx/ConstClust/GPS/Coches} 
	\caption[GPS data use.]{GPS data use. \cite{davidson2007survey} \cite{wagstaff2001constrained}}\label{fig:GPSClustData}
\end{figure}

In this domain, real clusters have an elongated shape on the horizontal axis and are aligned perpendicular to the motion direction. To get the resulting clusters to have this shape, we can make use of the constraints. We will set a \acf{CL} constraint between those instances more than 4 meters perpendicularly to the motion direction (since the lanes have a maximum width of 4 meters), and \acf{ML} constraints between those instances that present continuity in the motion direction axis, since it is likely that the vehicles they represent are in the same lane. This clustering model has proven to be very useful in real-time navigation \cite{wagstaff2001constrained}, allowing the user to be notified when to change lanes, or when not to leave them.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.32]{gfx/ConstClust/GPS/Instancias} 
	\caption[Resulting Clustering from GPS data with no constraints.]{Resulting Clustering from GPS data with no constraints. \cite{davidson2007survey} \cite{wagstaff2001constrained}}\label{fig:figure15}
\end{figure}

\section{Beneficios del uso de restricciones}

Encontramos dos beneficios principales en el uso de restricciones:

\begin{itemize}

	\item Incremento de la exactitud en las predicciones de las etiquetas al generar restricciones en base a un subconjunto de instancias etiquetadas.

	\item Obtención de clusters con geometría adaptable a cada problema.

\end{itemize}

A continuación se analizan estos dos beneficios:

Dado $X = \{x_1 \cdots x_u\}$, un gran conjunto de instancias no etiquetadas, y $L = \{(x_{u+1}, y_{u+1})\cdots (x_{u+l}, y_{u+l})\}$, un pequeño conjunto de instancias etiquetadas, es común escoger dos elementos de $L$ (con reemplazamiento) y establecer una restricción \acs{ML} entre ellos si pertenecen a la misma clase o, en caso contrario, una de tipo \acs{CL}. Un método apropiado para evaluar los resultados ofrecidos por un método de clustering es medir el nivel de exactitud de éste a la hora de predecir las etiquetas del conjunto $X$. Esto normalmente requiere que se especifique el número de clusters deseados igual al número de clases conocidas en $X$ ($K = K^*$). Para medir la exactitud se emplean métodos como \textit{Rand Index} \cite{rand1971objective}.

El trabajo de Wagstaff y Cardie \cite{wagstaff2000clustering}, en el que generaban las restricciones de la manera descrita anteriormente, demostraba que, cuando se realiza un promedio de la exactitud de las predicciones obtenidas con algoritmos de clustering con restricciones, variando estas últimas entre experimentos, se obtienen resultados hasta un 20\% mejores que con las técnicas clásicas.

\begin{observacion}

	\textbf{El uso de restricciones, en promedio, incrementa la precisión.}
	El rendimiento de un método al predecir etiquetas aumenta cuando se promedia empleando numerosos conjuntos de restricciones diferentes. \cite{davidson2007survey}
	\label{ob:observacion33}

\end{observacion}

Esta regla, sin embargo, no es cierta en todos los casos, pues en conjuntos de datos como \textit{Tic-Tac-Toe Endgame}, no se consigue ningún incremento en las predicciones sea cual sea el número de restricciones empleadas. La explicación dada por los autores citados para estas excepciones se basa en que establecer $K = K^*$ no es apropiado en este caso.

El otro beneficio que reporta el uso de restricciones es la posibilidad de obtener clusters con la geometría deseada, como el ejemplo de aplicar clustering a datos \acs{GPS}, analizado en la Sección \ref{EjemploGPS} de este trabajo.

\section{Problemas del uso de restricciones} \label{Problemas}

Aunque, tal y como hemos comprobado, la incorporación de restricciones a los métodos de clustering reporta beneficios en algunas aplicaciones, existen dos inconvenientes principales que se exponen a continuación, así como posibles soluciones a los mismos.

\subsection{El problema de la factibilidad}

La introducción de restricciones en el clustering cambia el problema al que éste da solución, que pasa a ser: \textit{Encontrar la mejor partición que satisfaga todas las restricciones}. De esta manera, si aquellas no están bien especificadas o si los métodos de extracción son inadecuados, podemos encontrar que las restricciones se contradicen, lo que deriva en que no existe una asignación de instancias a clusters que las satisfaga todas. Por ejemplo, no existe asignación que satisfaga las restricciones $ML(x_1,x_2)$ y $CL(x_1,x_2)$, independientemente del valor de $K$. Lo mismo sucede para $K = 2$, y las restricciones $CL(x_1, x_2)$, $CL(x_2, x_3)$ y $CL(x_1, x_3)$. Formalizando, el problema de la factibilidad para problemas de clustering (no jerárquico) con restricciones viene definido por:

\begin{definicion}

	\textbf{Problema de la factibilidad para clustering con restricciones:} Dado un conjunto de datos $X$, un conjunto de restricciones $R$, un umbral superior $K_l$ y un umbral superior $K_u$ para el número de clusters resultantes, ¿Existe una partición de $X$ en bloques tal que $K_l \le K \le K_u$ y todas las restricciones en $R$ se satisfacen? \cite{davidson2005clustering} \cite{davidson2007survey}

\end{definicion}

La complejidad teórica del problema dependerá del tipo de restricciones que se combinen en él. La Tabla \ref{tab:tabla1} presenta, de manera resumida, la complejidad esperada en cada caso.

\begin{table}[h]
	\centering
	\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{10pt}
	\renewcommand{\arraystretch}{1}

	\rowcolors{2}{gray!25}{white}
	\begin{tabular}{ >{\centering\arraybackslash}m{4cm}  >{\centering\arraybackslash}m{4cm} }
		\hline
		\rowcolor{black}
		\multicolumn{2}{c}{\bf \color{white}{Complejidad del clustering con restricciones}}\\
		\hline
		\rowcolor{gray!50}
		\textbf{Restricciones} & \textbf{Complejidad} \\
		Restricciones $\delta$ & $\mathbf{P}$ \\
		Restricciones $\epsilon$ & $\mathbf{P}$ \\
		\acs{ML} y $\delta$ & $\mathbf{P}$ \\
		\acs{ML} y $\epsilon$ & $\mathbf{NP}$-completo \\
		$\epsilon$ y $\delta$ & $\mathbf{P}$ \\
		\acs{CL} y otra & $\mathbf{NP}$-completo \\
		\hline

	\end{tabular}
	\caption[Complejidad del problema del clustering en función del tipo de restricciones]{Complejidad del problema del clustering en función del tipo de restricciones. \cite{davidson2007survey}}
	\label{tab:tabla1}
\end{table}

Tal y como queda reflejado en la Tabla \ref{tab:tabla1}, la utilización de restricciones \acf{CL} eleva el nivel de complejidad del clustering a $\mathbf{NP}$-completo y, por tanto, el problema del clustering con restricciones es intratable. De manera intuitiva puede entenderse fácilmente que, si encontrar una sola partición que satisfaga las restricciones es un problema complejo, más complejo es aún encontrar la mejor.

\begin{observacion}

	\textbf{Saber que existe una solución factible no nos ayuda a encontrarla.} Las consecuencias de este resultado sobre la complejidad del clustering con restricciones implican que, aun en caso de que exista una partición factible, no será fácil de encontrar, hablando en términos de complejidad algorítmica. \cite{davidson2007survey}
	\label{ob:observacion34}

\end{observacion}

Los autores Wagstaff \cite{wagstaff2002intelligent} y Davidson y Ravi \cite{davidson2007hierarchical} muestran que aun especificando el número de clusters de salida igual al de clases verdaderas ($K = K*$), cosa que garantiza que existe una solución factible, algoritmos simples como la adaptación de K-medias (\acs{KM}, Apéndice \ref{ap:kmeans}) al clustering con restricciones (\textit{COP-K-means} \cite{wagstaff2001constrained}, Sección \ref{copkm}), pueden no converger debido al problema de la factibilidad.

\subsection{El problema de la utilidad de conjuntos de restricciones} \label{ProbRestr}

En el clustering con restricciones se asume que éstas son indicaciones que guían al algoritmo para encontrar la partición de los datos deseada. Entonces, está justificado pensar que, de cuanta más información adicional (restricciones) dispongamos, más cercano estará el resultado que obtengamos al que buscamos, tal y como la Observación \ref{ob:observacion33} afirmaba. Sin embargo, y a pesar de lo dispuesto en dicha observación, encontramos casos en los que, aun generando las restricciones sin ruido y en base a las etiquetas verdaderas, existen conjuntos de restricciones que, lejos de mejorar los resultados, los empeoran considerablemente \cite{davidson2006proceedings}. Esto parece estar en desacuerdo con la Observación \ref{ob:observacion33}, sin embargo, recordemos que en ella se hace referencia al caso medio, y no a casos particulares.

\begin{observacion}

	\textbf{Conjuntos de restricciones particulares pueden causar efectos adversos}. Algunos conjuntos de restricciones generados en base a las etiquetas verdaderas y libres de ruido pueden resultar en una pérdida de precisión a la hora de predecir esas mismas etiquetas. \cite{davidson2007survey}

\end{observacion}

\subsection{Soluciones al problema de la factibilidad} \label{problemaFactib}

El problema de la factibilidad puede ser abordado de varias maneras. La más inmediata quizá sea mantener el número de restricciones bajo, en proporción al número de instancias totales, para minimizar la probabilidad de que surjan inconsistencias. Sin embargo, no poder aumentar el número de restricciones si el problema lo requiere no es el escenario ideal. Por ello se debe poner interés en analizar cuando un problema pasa a estar sobrerestringido, ya que, como hemos estudiamos en la Sección \ref{ProbRestr}, incluso generando las restricciones en base a las etiquetas verdaderas, algoritmos como COP-K-medias (Sección \ref{copkm}) dejan de ser efectivos conforme aumenta el número de restricciones a satisfacer, incluso reiniciado de manera aleatoria el algoritmo varias veces.

El fenómeno de la sobrerestricción de problemas mediante el uso de restricciones \acf{CL} está íntimamente relacionado con el problema del coloreado de grafos; de hecho, ha sido demostrado que éste es equivalente al problema del clustering con restricciones \acs{CL} \cite{davidson2006identifying}. Así, encontramos que resolver un problema con restricciones \acs{CL} mediante algoritmos como COP-K-medias es, a efectos prácticos, resolver el problema del coloreado de grafos.

\begin{observacion}

	\textbf{El problema del clustering con restricciones \acs{CL} es análogo al problema del coloreado de grafos.} \cite{davidson2007survey}

\end{observacion}

Este resultado permite trasladar muchas de las propiedades del problema del coloreado de grafos al problema de clustering con restricciones. Por ejemplo, el teorema de Brook establece que el coloreado de grafos es sencillo cuando el número de colores disponibles ($K$ en nuestro caso), es mayor que el máximo grado del grafo.

\begin{observacion}

	\textbf{El teorema de Brook es aplicable al problema del clustering con restricciones.}
	Si $ K > $ (Mayor número de restricciones \acs{CL} sobre una instancia), entonces siempre existirá una partición factible. \cite{davidson2007survey} \label{ob:observacion37}

\end{observacion}

Con esto, y aunque la Observación \ref{ob:observacion34} indique lo contrario, cuando el problema del clustering cumple la condición expuesta en la Observación \ref{ob:observacion37}, podemos garantizar que siempre se encontrará una solución al problema en tiempo polinómico. Para asegurar la condición de Brook, es posible construir el conjunto de restricciones de manera que ninguna instancia tome partido en más de $K$ restricciones \acf{CL}. \cite{davidson2006identifying}

\subsection{Soluciones al problema de la utilidad de conjuntos de restricciones}

La solución al problema es simple: identificar aquellos conjuntos de restricciones verdaderamente útiles. Sin embargo, esto involucra aplicar algún tipo de métrica que permita evaluar cuándo un conjunto de restricciones dado cumple esta condición. A tal fin, Davidson, Wagstaff y Basu propusieron dos medidas: informatividad y coherencia.\\

La \textbf{informatividad} es una medida referida a la cantidad de información presente en el conjunto de restricciones que el algoritmo no puede determinar por sí mismo. Por ejemplo, en la Figura \ref{fig:figure16}, un algoritmo como COP-K-medias (Sección \ref{copkm}) se vería inclinado a agrupar instancias cercanas en el espacio y colocar en clusters separados aquellas que se encuentren lejanas; sin embargo, las restricciones sesgan el espacio de soluciones evitando que esto suceda.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.4]{gfx/ConstClust/Inform/Inform} 
	\caption[Ejemplo de conjunto de restricciones informativo.]{Ejemplo de conjunto de restricciones informativo. \cite{davidson2007survey}}\label{fig:figure16}
\end{figure}

La informatividad se estima utilizando el conjunto de restricciones como un conjunto de test, de manera que se mide la habilidad del algoritmo para predecir las restricciones presentes en él. Formalizando, dado un conjunto de restricciones $R$ y un algoritmo $A$, obtenemos la partición $P_A$ aplicando el algoritmo al conjunto de datos de entrada especificando el conjunto de restricciones vacío. Calculamos entonces la fracción de las restricciones incumplidas por $P_A$ \cite{davidson2007survey}:

\begin{equation}
I_A(R) = \frac{1}{|R|}\left[ \sum_{r \in R} unsat(r, P_A) \right] 
\end{equation}

\clearpage

Por otra parte, la \textbf{coherencia} mide el grado de concordancia dentro del propio conjunto de restricciones respecto a una métrica dada ($D$). Por ejemplo, la Figura \ref{fig:figure17} muestra dos restricciones paralelas y muy cercanas, pero de distinto tipo. Es en casos como este en los que se da una contradicción, ya que las restricciones \acf{ML} indican que la distancia entre las instancias involucradas en ellas es pequeña, mientras que las de tipo \acf{CL} deben indicar lo contrario.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.4]{gfx/ConstClust/Coherencia/Coher1}
	\caption[Ejemplo de conjunto de restricciones contradictorio.]{Ejemplo de conjunto de restricciones contradictorio. \cite{davidson2007survey}}\label{fig:figure17}
\end{figure}

Entonces, la medida de coherencia viene dada por el grado de solapamiento que presentan las restricciones al interpretarlas como vectores en el espacio y proyectarlas sobre uno de los ejes, tal y como se muestra en la Figura \ref{fig:figure18}.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.4]{gfx/ConstClust/Coherencia/Coher2}
	\caption[Representación de la medida de coherencia.]{Representación de la medida de coherencia. \cite{davidson2007survey}}\label{fig:figure18}
\end{figure}

\section{Resumen}

El clustering con restricciones incorpora nueva información al problema del clustering original, esta viene dada en forma de especificaciones de pertenencia al mismo o a diferentes clusters sobre parejas de instancias. Las restricciones, ya sean \acf{ML}, \acf{CL}, o restricciones de distancia, son utilizadas para guiar al método de clustering que apliquemos al conjunto de datos en cuestión, en la búsqueda de la partición resultado.

Los método de clustering derivados de este concepto han demostrado ser de gran utilidad en múltiples ámbitos, así como presenta problemas que pueden ser subsanados estudiando en profundidad las restricciones a emplear para resolver cada problema.

### 2019-06-22 Memoria/Chapters/Introduction.tex
Escrito capitulo 2
Capitulo tres de la memoria empezado

@@
 Although there is a major impediment to the development of such a method: we do not know if it exists!
-The search for a method to avoid the disadvantages of the human learning process is what drives research in the science we know today as \acf{ML}.
+The search for a method to avoid the disadvantages of the human learning process is what drives research in the science we know today as \textit{machine learning}.
 
@@
 
-\acs{ML} is a field derived from computer science, and a branch of \acf{AI}, whose ultimate goal is to develop methods that make it possible for machines to learn.
-In other words, the techniques embedded in \acs{ML} aim to develop programs that guide computers to learn from a set of examples.
-\acs{ML} is part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work and help them.
+Machine learning is a field derived from computer science, and a branch of \acf{AI}, whose ultimate goal is to develop methods that make it possible for machines to learn.
+In other words, the techniques embedded in machine learning aim to develop programs that guide computers to learn from a set of examples.
+Machine learning is part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work and help them.
 
-\subsection{Types of Machine Learning}
+\subsection{Types of Machine Learning} \label{sec:TypesOfML}
 
-There are several approaches within the field of \acf{ML}, these are classified according to the information that is provided to the machine for it to learn.
+There are several approaches within the field of machine learning, these are classified according to the information that is provided to the machine for it to learn.
 Thus, following the classification proposed in \cite{ayodele2010types}, we find four different types of learning:

### 2019-06-23 Memoria/Chapters/ConstrainedClustering.tex
Terminada redaccion capitulo 3

@@
 Constraints that indicate that two instances must be placed in the same cluster are called \acf{ML}, and are noted with $C_=(x_i,x_j)$.
-Similarly, constraints that specify that two instances must not be placed in the same cluster are called \acf{CL}, and are noted with $C_\neq(x_i,x_j)$ \cite{wagstaff2000clustering}.
+Similarly, constraints that specify that two instances must not be placed in the same cluster are called \acf{CL}, and are noted with $C_{\neq}(x_i,x_j)$ \cite{wagstaff2000clustering}.
 
@@
 
-\begin{observacion}
+\begin{observation}
 	\textbf{Cannot-link constraints can be entailed.} Let $CC_i$ and $CC_j$ be connected components (completely connected subgraphs by \acs{CL} constraints) and let $x_i$ and $x_j$ be the instances in $CC_i$ and $CC_j$ respectively.
-Then $C_\neq(x,y): x \in CC_i, y \in CC_j \rightarrow _\neq(a,b) \forall a,b: a\in CC_i, b \in CC_j$.
+Then $C_{\neq}(x,y): x \in CC_i, y \in CC_j \rightarrow _{\neq}(a,b) \forall a,b: a\in CC_i, b \in CC_j$.
 \cite{davidson2007survey}
-\end{observacion}
+\end{observation}
 
@@
 The method used to obtain the constraints is one of the most popular in the literature: set the number of clusters of the resulting partition equal to the number of classes in the database, and generate the constraints from a subset of tagged instances; that is, if two instances have different labels, set a \acf{CL} constraint between them, otherwise one of \acf{ML} type.
-Thus, between the images shown in Figure \ref{fig:CMUFacesDatabase} \acf{CL} constraints are set, since, although they belong to the same person, they do not have the same orientation.
+Thus, between the images shown in Figure \ref{fig:FacesDatabaseCL} \acf{CL} constraints are set, since, although they belong to the same person, they do not have the same orientation.
 
@@
 	{\includegraphics[width=.35\linewidth]{gfx/ConstClust/AnalisisImagenes/CarasDifOr2}}
-	\caption[\acs{CL} constraints between faces of the same person.]{\acs{CL} constraints between faces of the same person. \cite{davidson2007survey}}\label{fig:figure10}
+	\caption[\acs{CL} constraints between faces of the same person.]{\acs{CL} constraints between faces of the same person. \cite{davidson2007survey}}\label{fig:FacesDatabaseCL}
 \end{figure}
 
-Figure \ref{fig:AiboRoboRobtoClustSys} shows another set of image data on which constrained clustering techniques are applied.
+Figure \ref{fig:AiboRobotClustSys} shows another set of image data on which constrained clustering techniques are applied.
 In this case, the task is to perform object recognition to incorporate the method into the navigation system of the Aibo robot \cite{davidson2005clustering}.
@@
 	\caption[Clustering method used in Aibo Robot navigation system.]{Clustering method used in Aibo Robot navigation system.
-\cite{davidson2007survey}\cite{davidson2005clustering}}\label{fig:AiboRobtoClustSys}
+\cite{davidson2007survey}\cite{davidson2005clustering}}\label{fig:AiboRobotClustSys}
 \end{figure}
@@
 Does this make the problem over-constrained?
-In Section \ref{sec:CCDisadvantages} we will address these questions.
+In Section \ref{sec:ConstraintsProblems} we will address these questions.
 
@@
 
-\subsection{Aplicaciones en datos de GPS} \label{EjemploGPS}
+\subsection{Aplicaciones en datos de GPS} \label{sec:GPSApp}
 
@@
 
-In this domain, real clusters have an elongated shape on the horizontal axis and are aligned perpendicular to the motion direction.
+In this domain, real clusters have an elongated shape on the horizontal axis and are aligned perpendicularly to the motion direction.
 To get the resulting clusters to have this shape, we can make use of the constraints.
@@
 
-\section{Beneficios del uso de restricciones}
+\subsection{Recent Applications}
 
-Encontramos dos beneficios principales en el uso de restricciones:
+The named constrained clustering applications are the most studied in the literature, but it is worth noting that is has been applied to many other fields in recent times.
+The following is a list of some of those applications: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, among others.
+\textcolor{red}{Poner mas ejemplos de aplicación recientes}
+
+\section{Benefits of Using Constraints}
+
+Two main benefits are found in the use of constraints:
 
@@
 
-	\item Incremento de la exactitud en las predicciones de las etiquetas al generar restricciones en base a un subconjunto de instancias etiquetadas.
+	\item Increased accuracy in label predictions by generating constraints based on a subset of labeled instances.
 
-	\item Obtención de clusters con geometría adaptable a cada problema.
+	\item Obtaining clusters with adaptable geometry to each problem.
 
@@
 
-A continuación se analizan estos dos beneficios:
+These two benefits are discussed below:
 
-Dado $X = \{x_1 \cdots x_u\}$, un gran conjunto de instancias no etiquetadas, y $L = \{(x_{u+1}, y_{u+1})\cdots (x_{u+l}, y_{u+l})\}$, un pequeño conjunto de instancias etiquetadas, es común escoger dos elementos de $L$ (con reemplazamiento) y establecer una restricción \acs{ML} entre ellos si pertenecen a la misma clase o, en caso contrario, una de tipo \acs{CL}.
-Un método apropiado para evaluar los resultados ofrecidos por un método de clustering es medir el nivel de exactitud de éste a la hora de predecir las etiquetas del conjunto $X$.
-Esto normalmente requiere que se especifique el número de clusters deseados igual al número de clases conocidas en $X$ ($K = K^*$).
-Para medir la exactitud se emplean métodos como \textit{Rand Index} \cite{rand1971objective}.
+Given $X = \{x_1 \cdots x_u\}$, a large set of unlabeled instances, and $L = \{(x_{u+1}, y_{u+1}) \cdots (x_{u+l}, y_{u+l})\}$, a small set of labeled instances, it is common to choose two elements from $L$ (with replacement) and set a \acs{ML} constraint between them if they belong to the same class, or a \acs{CL} constraints otherwise.
+An appropriate method to evaluate the results provided by a clustering method is to measure the level of accuracy of this when predicting dataset $X$ labels.
+This normally requires specifying the number of desired clusters equal to the number of known classes in $X$ ($k = k^*$).
+Methods such as the RandIndex \cite{rand1971objective} or its variants, which will be discussed later, are used to measure the degree of agreement between two given partitions (clusterings).
 
-El trabajo de Wagstaff y Cardie \cite{wagstaff2000clustering}, en el que generaban las restricciones de la manera descrita anteriormente, demostraba que, cuando se realiza un promedio de la exactitud de las predicciones obtenidas con algoritmos de clustering con restricciones, variando estas últimas entre experimentos, se obtienen resultados hasta un 20\% mejores que con las técnicas clásicas.
+In \cite{wagstaff2000clustering}, where constraints are generated as described above, it is shown that, when an average of the accuracy of the predictions obtained with constrained clustering algorithms is made, varying the latter between experiments, results are obtained up to 20\% better than with classical clustering techniques.
 
-\begin{observacion}
+\begin{observation}
 
-	\textbf{El uso de restricciones, en promedio, incrementa la precisión.}
-	El rendimiento de un método al predecir etiquetas aumenta cuando se promedia empleando numerosos conjuntos de restricciones diferentes. \cite{davidson2007survey}
-	\label{ob:observacion33}
+	\textbf{Using constraints increases accuracy on average.}
+	The performance of a method in predicting labels increases when averaged using numerous different constraint sets. \cite{davidson2007survey}
+	\label{ob:AccuracyIncrease}
 
-\end{observacion}
+\end{observation}
 
-Esta regla, sin embargo, no es cierta en todos los casos, pues en conjuntos de datos como \textit{Tic-Tac-Toe Endgame}, no se consigue ningún incremento en las predicciones sea cual sea el número de restricciones empleadas.
-La explicación dada por los autores citados para estas excepciones se basa en que establecer $K = K^*$ no es apropiado en este caso.
+This rule, however, is not true in all cases, as in data sets such as \textit{Tic-Tac-Toe Endgame}, no increase in predictions is achieved regardless of the number of constraints used.
+A possible explanation for these exceptions is based on the fact that setting $k = k^*$ is not appropriate in these cases.
 
-El otro beneficio que reporta el uso de restricciones es la posibilidad de obtener clusters con la geometría deseada, como el ejemplo de aplicar clustering a datos \acs{GPS}, analizado en la Sección \ref{EjemploGPS} de este trabajo.
+The other benefit of using constraints is the possibility of obtaining clusters with the desired geometry, such as the example of \acs{GPS} data clustering, analyzed in the Section \ref{sec:GPSApp}.
 
-\section{Problemas del uso de restricciones} \label{Problemas}
+\section{Problems of Using Constraints} \label{sec:ConstraintsProblems}
 
-Aunque, tal y como hemos comprobado, la incorporación de restricciones a los métodos de clustering reporta beneficios en algunas aplicaciones, existen dos inconvenientes principales que se exponen a continuación, así como posibles soluciones a los mismos.
+Although, as we have seen, the incorporation of constraints to clustering methods brings benefits in some applications, there are two main drawbacks that are discussed below, as well as possible solutions to them.
 
-\subsection{El problema de la factibilidad}
+\subsection{The Feasibility Problem}
 
-La introducción de restricciones en el clustering cambia el problema al que éste da solución, que pasa a ser: \textit{Encontrar la mejor partición que satisfaga todas las restricciones}.
-De esta manera, si aquellas no están bien especificadas o si los métodos de extracción son inadecuados, podemos encontrar que las restricciones se contradicen, lo que deriva en que no existe una asignación de instancias a clusters que las satisfaga todas.
-Por ejemplo, no existe asignación que satisfaga las restricciones $ML(x_1,x_2)$ y $CL(x_1,x_2)$, independientemente del valor de $K$.
-Lo mismo sucede para $K = 2$, y las restricciones $CL(x_1, x_2)$, $CL(x_2, x_3)$ y $CL(x_1, x_3)$.
-Formalizando, el problema de la factibilidad para problemas de clustering (no jerárquico) con restricciones viene definido por:
+Introducing constraints in clustering changes the problem it solves, which becomes: \textit{Find the best partition that satisfies all constraints}.
+This way, if constraints are not properly specified or if the extraction methods are inadequate, we can find that the constraints contradict each other, which means that there is no partition that satisfies them all.
+For instance, there is no partition that satisfies the constraints $C_=(x_i,x_j)$ and $C_{\neq}(x_i,x_j)$, regardless of the value of $k$.
+The same happens for $k = 2$, and the constraints $C_{\neq}(x_i, x_j)$, $C_{\neq}(x_j, x_q)$ and $CL(x_i, x_q)$.
+Definition \ref{def:FeasibilityProblem} formalizes the feasibility problem for non-hierarchical instance-level constrained clustering:
 
-\begin{definicion}
+\begin{definition}
 
-	\textbf{Problema de la factibilidad para clustering con restricciones:} Dado un conjunto de datos $X$, un conjunto de restricciones $R$, un umbral superior $K_l$ y un umbral superior $K_u$ para el número de clusters resultantes, ¿Existe una partición de $X$ en bloques tal que $K_l \le K \le K_u$ y todas las restricciones en $R$ se satisfacen?
+	\textbf{Feasibility problem for non-hierarchical instance-level constrained clustering:} Given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
 \cite{davidson2005clustering} \cite{davidson2007survey}
 
-\end{definicion}
+	\label{def:FeasibilityProblem}
 
-La complejidad teórica del problema dependerá del tipo de restricciones que se combinen en él. La Tabla \ref{tab:tabla1} presenta, de manera resumida, la complejidad esperada en cada caso.
+\end{definition}
 
-\begin{table}[h]
+ The theoretical complexity of a constrained clustering problem depends on the type of constraints combined in it. Table \ref{tab:CCComplexity} summarizes the expected complexity in each case.
+
+\begin{table}[!h]
 	\centering
-	\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{10pt}
-	\renewcommand{\arraystretch}{1}
-
-	\rowcolors{2}{gray!25}{white}
+	%\setlength{\arrayrulewidth}{1mm}
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.2}
+	%\resizebox{\textwidth}{!}{
 	\begin{tabular}{ >{\centering\arraybackslash}m{4cm}  >{\centering\arraybackslash}m{4cm} }
 		\hline
-		\rowcolor{black}
-		\multicolumn{2}{c}{\bf \color{white}{Complejidad del clustering con restricciones}}\\
-		\hline
-		\rowcolor{gray!50}
-		\textbf{Restricciones} & \textbf{Complejidad} \\
-		Restricciones $\delta$ & $\mathbf{P}$ \\
-		Restricciones $\epsilon$ & $\mathbf{P}$ \\
-		\acs{ML} y $\delta$ & $\mathbf{P}$ \\
-		\acs{ML} y $\epsilon$ & $\mathbf{NP}$-completo \\
-		$\epsilon$ y $\delta$ & $\mathbf{P}$ \\
-		\acs{CL} y otra & $\mathbf{NP}$-completo \\
+		\textbf{Constraints} & \textbf{Complexity} \\
+		$\delta$ & $\mathbf{P}$ \\
+		$\epsilon$ & $\mathbf{P}$ \\
+		\acs{ML} and $\delta$ & $\mathbf{P}$ \\
+		\acs{ML} and $\epsilon$ & $\mathbf{NP}$-complete \\
+		$\epsilon$ and $\delta$ & $\mathbf{P}$ \\
+		\acs{CL} and other & $\mathbf{NP}$-complete \\
 		\hline
 
-	\end{tabular}
-	\caption[Complejidad del problema del clustering en función del tipo de restricciones]{Complejidad del problema del clustering en función del tipo de restricciones. \cite{davidson2007survey}}
-	\label{tab:tabla1}
+	\end{tabular}%}
+	\caption[Complexity of the constrained clustering problem depending on the type of constraints.]{Complexity of the constrained clustering problem depending on the type of constraints.
+\cite{davidson2007survey}}
+\label{tab:CCComplexity}
 \end{table}
 
-Tal y como queda reflejado en la Tabla \ref{tab:tabla1}, la utilización de restricciones \acf{CL} eleva el nivel de complejidad del clustering a $\mathbf{NP}$-completo y, por tanto, el problema del clustering con restricciones es intratable.
-De manera intuitiva puede entenderse fácilmente que, si encontrar una sola partición que satisfaga las restricciones es un problema complejo, más complejo es aún encontrar la mejor.
+As shown in Table \ref{tab:CCComplexity}, the use of \acf{CL} constraints increases the constrained clustering complexity level to $\mathbf{NP}$-complete and therefore constrained clustering is intractable.
+Intuitively it can be easily understood that if finding a single partition that satisfies the constraints is a complex problem, finding the best of them is harder.
 
-\begin{observacion}
+\begin{observation}
 
-	\textbf{Saber que existe una solución factible no nos ayuda a encontrarla.} Las consecuencias de este resultado sobre la complejidad del clustering con restricciones implican que, aun en caso de que exista una partición factible, no será fácil de encontrar, hablando en términos de complejidad algorítmica.
-\cite{davidson2007survey}
-	\label{ob:observacion34}
+	\textbf{Knowing a feasible solution exists does not help us find it.} The consequences of this result on the constrained clustering complexity imply that, even if there is a feasible partition, it will not be easy to find, speaking in terms of algorithmic complexity \cite{davidson2007survey}.
+	\label{ob:FeasibleSolution}
 
-\end{observacion}
+\end{observation}
 
-Los autores Wagstaff \cite{wagstaff2002intelligent} y Davidson y Ravi \cite{davidson2007hierarchical} muestran que aun especificando el número de clusters de salida igual al de clases verdaderas ($K = K*$), cosa que garantiza que existe una solución factible, algoritmos simples como la adaptación de K-medias (\acs{KM}, Apéndice \ref{ap:kmeans}) al clustering con restricciones (\textit{COP-K-means} \cite{wagstaff2001constrained}, Sección \ref{copkm}), pueden no converger debido al problema de la factibilidad.
+In \cite{wagstaff2002intelligent} and \cite{davidson2007hierarchical} is shown that, even in the number of clusters in the output partition is set equal to the number of classes in the dataset ($k = k*$), which guarantees that a feasible solution does exists, simple algorithms like the COP-kmeans \cite{wagstaff2001constrained} may not converge due to the feasibility problem.
 
-\subsection{El problema de la utilidad de conjuntos de restricciones} \label{ProbRestr}
+\subsection{The Constraint Set Utility Problem} \label{sec:UtilityProblem}
 
-En el clustering con restricciones se asume que éstas son indicaciones que guían al algoritmo para encontrar la partición de los datos deseada.
-Entonces, está justificado pensar que, de cuanta más información adicional (restricciones) dispongamos, más cercano estará el resultado que obtengamos al que buscamos, tal y como la Observación \ref{ob:observacion33} afirmaba.
-Sin embargo, y a pesar de lo dispuesto en dicha observación, encontramos casos en los que, aun generando las restricciones sin ruido y en base a las etiquetas verdaderas, existen conjuntos de restricciones que, lejos de mejorar los resultados, los empeoran considerablemente \cite{davidson2006proceedings}.
-Esto parece estar en desacuerdo con la Observación \ref{ob:observacion33}, sin embargo, recordemos que en ella se hace referencia al caso medio, y no a casos particulares.
+In constrained clustering it is assumed that constraints are a guide for an algorithm to find de desired data partition.
+So, it is reasonable to think that the more additional information (constraints) we have available, the closer the result we get to the one we are looking for, as Observation \ref{ob:AccuracyIncrease} stated.
+However, and despite the stipulations of this observation, we find cases in which, even generating the constraints without noise and on the basis of the true labels, there are constraint sets that, far from improving the results, worsen them considerably \cite{davidson2006proceedings}.
+This seems to disagree with the Observation \ref{ob:AccuracyIncrease}, however, let us remember that it refers to the average case, and not to particular cases.
 
-\begin{observacion}
+\begin{observation}
 
-	\textbf{Conjuntos de restricciones particulares pueden causar efectos adversos}.
-Algunos conjuntos de restricciones generados en base a las etiquetas verdaderas y libres de ruido pueden resultar en una pérdida de precisión a la hora de predecir esas mismas etiquetas.
-\cite{davidson2007survey}
+	\textbf{Individual constraint sets can have adverse effects}.
+Some constraint sets generated on the same ground truth labels that evaluate the clusters can cause a loss of accuracy at predicting those very labels \cite{davidson2007survey}.
 
-\end{observacion}
+\end{observation}
 
-\subsection{Soluciones al problema de la factibilidad} \label{problemaFactib}
+\subsection{Solutions for the Feasibility Problem} \label{sec:FeasibilityProblemSolutions}
 
-El problema de la factibilidad puede ser abordado de varias maneras.
-La más inmediata quizá sea mantener el número de restricciones bajo, en proporción al número de instancias totales, para minimizar la probabilidad de que surjan inconsistencias.
-Sin embargo, no poder aumentar el número de restricciones si el problema lo requiere no es el escenario ideal.
-Por ello se debe poner interés en analizar cuando un problema pasa a estar sobrerestringido, ya que, como hemos estudiamos en la Sección \ref{ProbRestr}, incluso generando las restricciones en base a las etiquetas verdaderas, algoritmos como COP-K-medias (Sección \ref{copkm}) dejan de ser efectivos conforme aumenta el número de restricciones a satisfacer, incluso reiniciado de manera aleatoria el algoritmo varias veces.
+The feasibility problem can be tackled in several ways.
+The most immediate may be to keep the number of constraints low, in proportion to the number of total instances, to minimize the likelihood of inconsistencies.
+However, not being able to increase the number of constraints if the problem requires it is not the ideal scenario.
+Therefore, interest should be given to analyzing when a problem becomes over-constrained, since, as we have studied in Section \ref{sec:ConstraintsProblems}, even generating constraints based on the ground truth, algorithms such as COP-kmeans are no longer effective as the number of constraints to be satisfied increases, even if the algorithm is randomly restarted several times.
 
-El fenómeno de la sobrerestricción de problemas mediante el uso de restricciones \acf{CL} está íntimamente relacionado con el problema del coloreado de grafos; de hecho, ha sido demostrado que éste es equivalente al problema del clustering con restricciones \acs{CL} \cite{davidson2006identifying}.
-Así, encontramos que resolver un problema con restricciones \acs{CL} mediante algoritmos como COP-K-medias es, a efectos prácticos, resolver el problema del coloreado de grafos.
+The problem over-constraining phenomenon through the use of \acf{CL} constraints is intimately related to the graph coloring problem; in fact, it has been proven that it is equivalent to constrained clustering with \acf{CL} \cite{davidson2006identifying}.
+Thus, we find that solving a problem with \acf{CL} constraints using algorithms such as COP-K-medias is, for practical purposes, solving the graph coloring problem.
 
-\begin{observacion}
+\begin{observation}
 
-	\textbf{El problema del clustering con restricciones \acs{CL} es análogo al problema del coloreado de grafos.} \cite{davidson2007survey}
+	\textbf{Constrained clustering with \acs{CL} is analogous to the graph coloring problem.} \cite{davidson2007survey}
 
-\end{observacion}
+\end{observation}
 
-Este resultado permite trasladar muchas de las propiedades del problema del coloreado de grafos al problema de clustering con restricciones.
-Por ejemplo, el teorema de Brook establece que el coloreado de grafos es sencillo cuando el número de colores disponibles ($K$ en nuestro caso), es mayor que el máximo grado del grafo.
+This result allows you to transfer many of the properties of the graph coloring problem to the constrained clustering problem.
+For example, Brook's theorem states that graph coloring is simple when the number of available colors ($k$ in our case) is greater than the maximum degree of the graph.
 
-\begin{observacion}
+\begin{observation}
 
-	\textbf{El teorema de Brook es aplicable al problema del clustering con restricciones.}
-	Si $ K > $ (Mayor número de restricciones \acs{CL} sobre una instancia), entonces siempre existirá una partición factible. \cite{davidson2007survey} \label{ob:observacion37}
+	\textbf{Brook's theorem applies to the constrained clustering.}
+	If $ k > $ (Most \acs{CL} constraints on an instance), then there always exists a feasible partition. \cite{davidson2007survey} \label{ob:BrooksTheorem}
 
-\end{observacion}
+\end{observation}
 
-Con esto, y aunque la Observación \ref{ob:observacion34} indique lo contrario, cuando el problema del clustering cumple la condición expuesta en la Observación \ref{ob:observacion37}, podemos garantizar que siempre se encontrará una solución al problema en tiempo polinómico.
-Para asegurar la condición de Brook, es posible construir el conjunto de restricciones de manera que ninguna instancia tome partido en más de $K$ restricciones \acf{CL}.
-\cite{davidson2006identifying}
+With this, and although Observation \ref{ob:FeasibleSolution} indicates differently, when constrained clustering meets the condition set out in Observation \ref{ob:BrooksTheorem}, we can guarantee that a solution will always be found in polynomial time.
+To ensure Brook's condition, it is possible to build the constraint set so that no instance takes a side in more than $K$ \acf{CL} constraints \cite{davidson2006identifying}.
 
-\subsection{Soluciones al problema de la utilidad de conjuntos de restricciones}
+\subsection{Solutions for the Constraint Set Utility Problem}
 
-La solución al problema es simple: identificar aquellos conjuntos de restricciones verdaderamente útiles.
-Sin embargo, esto involucra aplicar algún tipo de métrica que permita evaluar cuándo un conjunto de restricciones dado cumple esta condición.
-A tal fin, Davidson, Wagstaff y Basu propusieron dos medidas: informatividad y coherencia.\\
+The solution to this problem is simple: identify those truly useful constraint sets.
+However, this involves applying some kind of metric that allows us to evaluate when a given set of constraints meets this condition.
+To this end two measures are proposed: informativity and coherence.
 
-La \textbf{informatividad} es una medida referida a la cantidad de información presente en el conjunto de restricciones que el algoritmo no puede determinar por sí mismo.
-Por ejemplo, en la Figura \ref{fig:figure16}, un algoritmo como COP-K-medias (Sección \ref{copkm}) se vería inclinado a agrupar instancias cercanas en el espacio y colocar en clusters separados aquellas que se encuentren lejanas; sin embargo, las restricciones sesgan el espacio de soluciones evitando que esto suceda.
+The \textbf{informativity} measure refers to the amount of information present in the constraint set that an algorithm cannot determine for itself.
+For example, in Figure \ref{fig:Informativity}, an algorithm such as COP-kmeans would tend to group nearby instances and place those that are far away in separate clusters; however, constraints bias the solution space preventing this from happening.
 
@@
 	\includegraphics[scale=0.4]{gfx/ConstClust/Inform/Inform} 
-	\caption[Ejemplo de conjunto de restricciones informativo.]{Ejemplo de conjunto de restricciones informativo. \cite{davidson2007survey}}\label{fig:figure16}
+	\caption[Informative constraint set.]{Informative constraint set \cite{davidson2007survey}}\label{fig:Informativity}.
 \end{figure}
 
-La informatividad se estima utilizando el conjunto de restricciones como un conjunto de test, de manera que se mide la habilidad del algoritmo para predecir las restricciones presentes en él.
-Formalizando, dado un conjunto de restricciones $R$ y un algoritmo $A$, obtenemos la partición $P_A$ aplicando el algoritmo al conjunto de datos de entrada especificando el conjunto de restricciones vacío.
-Calculamos entonces la fracción de las restricciones incumplidas por $P_A$ \cite{davidson2007survey}:
+Informativity is estimated using the constraint set as a test set, so the ability of the algorithm to predict the constraints present in it is measured.
+Formalizing, given a set of constraints $CS$ and a constrained clustering algorithm $A$, we get the partition $P_A$ by applying the algorithm to the input dataset specifying an empty constraint set.
+We can compute the number of unsatisfied constraints in $P_A$ a in Equation \ref{eq:Informativity} \cite{davidson2007survey}.
 
@@
 I_A(R) = \frac{1}{|R|}\left[ \sum_{r \in R} unsat(r, P_A) \right] 
+\label{eq:Informativity}
 \end{equation}
 
-\clearpage
-
-Por otra parte, la \textbf{coherencia} mide el grado de concordancia dentro del propio conjunto de restricciones respecto a una métrica dada ($D$).
-Por ejemplo, la Figura \ref{fig:figure17} muestra dos restricciones paralelas y muy cercanas, pero de distinto tipo.
-Es en casos como este en los que se da una contradicción, ya que las restricciones \acf{ML} indican que la distancia entre las instancias involucradas en ellas es pequeña, mientras que las de tipo \acf{CL} deben indicar lo contrario.
+\textbf{Coherence} measures the degree of agreement within the set of constraints itself with respect to a given metric ($D$).
+For instance, Figure \ref{fig:Coherence} shows two parallel and very close constraints of different type.
+It is in cases like this that the contradiction occurs, as the \acf{ML} constraints indicate that distance between instances involved in them is small, while \acf{CL}s indicate the opposite.
 
@@
 	\includegraphics[scale=0.4]{gfx/ConstClust/Coherencia/Coher1}
-	\caption[Ejemplo de conjunto de restricciones contradictorio.]{Ejemplo de conjunto de restricciones contradictorio. \cite{davidson2007survey}}\label{fig:figure17}
+	\caption[Non-coherent constraint set.]{Non-coherent constraint set. \cite{davidson2007survey}}\label{fig:Coherence}
 \end{figure}
 
-Entonces, la medida de coherencia viene dada por el grado de solapamiento que presentan las restricciones al interpretarlas como vectores en el espacio y proyectarlas sobre uno de los ejes, tal y como se muestra en la Figura \ref{fig:figure18}.
+Then, the coherence measure is given by the degree of overlap that constraints produce by interpreting them as vectors in space and projecting them along one of the axes, as shown in Figure \ref{fig:CoherenceOverlap}.
 
@@
 	\includegraphics[scale=0.4]{gfx/ConstClust/Coherencia/Coher2}
-	\caption[Representación de la medida de coherencia.]{Representación de la medida de coherencia. \cite{davidson2007survey}}\label{fig:figure18}
+	\caption[Coherence measure representation.]{Coherence measure representation. \cite{davidson2007survey}}\label{fig:CoherenceOverlap}
 \end{figure}
 
-\section{Resumen}
+\section{Summary}
 
-El clustering con restricciones incorpora nueva información al problema del clustering original, esta viene dada en forma de especificaciones de pertenencia al mismo o a diferentes clusters sobre parejas de instancias.
-Las restricciones, ya sean \acf{ML}, \acf{CL}, o restricciones de distancia, son utilizadas para guiar al método de clustering que apliquemos al conjunto de datos en cuestión, en la búsqueda de la partición resultado.
+Constrained clustering adds a new type of information to the original clustering problem, which is given in the form of specifications of belonging to the same or to different clusters on pairs of instances.
+Constraints, whether if they are the pairwise \acf{ML} and \acf{CL} constraints, or the distance-based $\delta$ and $\epsilon$ constraints, they are used to guide the clustering method we apply to a dataset in the search for a result partition that meets certain desired characteristics.
 
-Los método de clustering derivados de este concepto han demostrado ser de gran utilidad en múltiples ámbitos, así como presenta problemas que pueden ser subsanados estudiando en profundidad las restricciones a emplear para resolver cada problema.
+Clustering methods derived from this concept have proved to be very useful in many areas, as well as they present some problems that can be solved by deeply studying the constraints to be used to solve each individual problem.

### 2019-06-24 Memoria/Chapters/Clustering.tex
Capitulo tres de la memoria empezado
Terminado capitulo 3

@@
-\chapter{Short introduction to Clustering}\label{ch:Breve introducción al Clustering}
+\chapter{Short Introduction to Clustering}\label{ch:IntroClustering}
 
@@
 
-\section{Reasons for classifying}
+\section{Reasons for Classifying}
 
@@
 In this way, the processes of extracting useful information and applying it are simplified.
-Thus, if we are able to meaningfully divide a large dataset into subsets or groups, we can extract information common to all elements of the subset and provide a precise description that encompasses them.
+Thus, if we are able to meaningfully divide a large dataset into subsets or groups, we can extract information common to elements of one of these subsets and provide a precise description that encompasses them.
 
@@
 
-It is important to note that, in most cases, there is no single classification criterion for the same data set, but rather a wide variety of data.
+It is important to note that, in most cases, there is no single classification criterion for the same dataset, but rather a wide variety of data.
 In the case of individuals, they could be classified, for example, on the basis of their income, or according to the amount of calories they consume over a defined period of time.
@@
 
-Numerical methods for clustering arise in branches of the natural sciences, such as biology or zoology, in an attempt to eliminate the subjectivity implicit in the classification process that is triggered by the discovery of a new species.
+Numerical methods for clustering arise in branches of natural sciences, such as biology or zoology, in an attempt to eliminate the subjectivity implicit in the classification process triggered by the discovery of a new species.
 The aim is to provide a non-subjective and stable method for classifying and grouping elements.
@@
 
-In most clustering applications, the objective is to obtain a data partition, in which each instance or object belongs to a single cluster, and the union of all of them contains all the individual objects.
-That said, it should be noted that, in some circumstances, solutions where there is overlap between clusters are acceptable, as well as the fact that there may not be an acceptable partition of the data.
+In most clustering applications, the goal is to obtain a data partition, in which each instance or object belongs to a single cluster, and the union of all of them contains all the individual objects.
+That said, it should be noted that, in some circumstances, solutions where there is overlap between clusters are acceptable, as well as there may not be an acceptable data partition.
 
@@
 A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
-In this study we will focus on partitional clustering.
+In this work we will focus on partitional clustering.
 
-The most widely used way to represent the information on which clustering is applied is an array $X$ of dimension $n\times u$, in which each row corresponds to an instance or object to be processed, and each column corresponds to one of the variables that characterize those instances or objects.
+The most common way to represent the information on which clustering is applied is an array $X$ of dimension $n\times u$, in which each row corresponds to an instance or object to be processed, and each column corresponds to one of the variables that characterize those instances or objects.
 The commonly accepted term to refer to each row is \textit{features vector}, and the features vector set is referred as \textit{dataset}.
@@
 As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $\{c_i, \cdots, c_k\}$ to form a partition called $C$.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement \cite{jain1999data}.
+The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot,\cdot)$ \cite{jain1999data}.
 
@@
 
-Some clustering methods involve making transformations on the $X$ matrix to convert it into a $u\times u$ matrix in which measurements extracted from the $X$ matrix are stored that relate one instance to all the others, such as similarity, distance or dissimilarity.
+Some clustering methods involve making transformations on the $X$ matrix to convert it into a $n \times n$ matrix in which measurements extracted from the $X$ matrix are stored that relate one instance to all the others, such as similarity, distance or dissimilarity.
 
@@
 
-Once the general structure of clustering methods has been defined, it would be justified to ask, what is a cluster? The Section \ref{sec:WhatIsCluster} will try to answer this question.
+Once the general structure of clustering methods has been defined, it would be justified to ask, what is a cluster? Section \ref{sec:WhatIsCluster} will try to answer this question.
 
@@
 So far, the terms cluster, group and class have been used in a completely intuitive way, with no need for any formal definition, a further proof of the innate nature of these concepts in the human being.
-In fact, to give a formal definition of cluster is a task, not only complicated, but in many occasions not very useful.
-For example, in \cite{lance1967general} a cluster definition completely dependent on the user's interpretation is proposed by Bonner.
+In fact, giving a formal definition of cluster is a task, not only complicated, but in many occasions not very useful.
+For example, in \cite{lance1967general} a cluster definition completely dependent on the user's interpretation is proposed.
 As far as this definition is concerned, a cluster is what the user understands as a cluster without having proposed a formal definition of it.
@@
 
-Although Bonner's definition is accurate in a wide range of situations, authors such as Cormack, in 1971 \cite{cormack1971review}, or Gordon in 1999 \cite{gordon:1999}, propose a somewhat more analytical definition from the mathematical point of view, defining a cluster in terms of internal cohesion---homogeneity---, and external isolation---separation.
-Figure \ref{fig:clusters_properties} informally illustrates the properties described above, so that any observer will find the clusters present in it apparent without the need for a formal cluster definition.
+Although the definition above is accurate in a wide range of situations, a somewhat more analytical definition from the mathematical point of view, defining a cluster in terms of internal cohesion---homogeneity---, and external isolation---separation is proposed in \cite{cormack1971review} and \cite{gordon:1999}.
+Figure \ref{fig:ClustersProperties} informally illustrates the properties described above, so that any observer will find the clusters present in it apparent without the need for a formal cluster definition.
 This may explain why achieving a mathematically accurate definition of homogeneity and separation may become unnecessary.
@@
 	{\includegraphics[width=.3\linewidth]{gfx/Clustering/ThreeBasicClusters}}
-	\caption[Clusters with internal cohesion and/or external isolation]{Clusters with internal cohesion and/or external isolation}\label{fig:clusters_properties}
+	\caption[Clusters with internal cohesion and/or external isolation]{Clusters with internal cohesion and/or external isolation}\label{fig:ClustersProperties}
 \end{figure}
@@
 The process of dividing a homogeneous distribution of data into different groups is known as dissection, and such a process may be useful in certain circumstances.
-However, since in most cases of actual application of cluster methods, the structure of the data is not known a priori, there is a risk of interpreting all the solutions in terms of the existence of subgroups, which would lead to the imposition of a fictitious structure on data in which there is no structure present.
+However, since in most cases of actual application of cluster methods, the structure of the data is not known a priori, there is risk of interpreting all solutions in terms of the existence of subgroups, which would lead to the imposition of a fictitious structure on data in which there is no structure present.
 
@@
 As it has been already mentioned, the general problem clustering attempts to solve is present in many disciplines: biology, botany, medicine, psychology, geography, marketing, image processing, psychiatry, archaeology, etc.
-This section presents some of the clustering applications related to these various knowledge fields.
+This section presents some of the clustering applications related to these and others knowledge fields.
 
@@
 An example of the application of clustering in this field was published in \cite{green1967cluster}.
-Thus, with a large number of cities available for analysis, they had to restrict the places in which to carry out their studies for economic reasons.
-For this they relied on cluster analysis, classifying cities into small groups based on 14 city characteristics, including size and average income \textit{per capita}.
-Since the cities in the same group were expected to be very similar, they chose one city from each of them to conduct their studies.
+With a large number of cities available for analysis, they had to restrict the places in which to carry out their studies for economic reasons.
+To this end they relied on cluster analysis, classifying cities into small groups based on 14 city features, including size and average income \textit{per capita}.
+Since cities in the same group were expected to be very similar, they chose one city from each of them to conduct their studies.
 
@@
 In this case, a car manufacturer believes that buying a sports car is not a decision based solely on economic capabilities or age, but is a lifestyle decision made by those who decide to buy a sports car, as opposed to those who do not.
-Consequently, the manufacturer decides to carry out a study, using cluster analysis, which allows it to identify all the characteristics related to the people who would buy a sports car, in order to focus its marketing campaigns specifically on this sector.
+Consequently, the manufacturer decides to carry out a study, using cluster analysis, which allows it to identify all features related to the people who would buy a sports car, in order to focus its marketing campaigns specifically on this sector.
 
@@
 Given a set of astronomical data, researchers want to know, for example, how many classes of stars are present in them, based on some statistical criteria.
-The most frequently asked questions in this area are: How many statistically different objects are present in the data and to which class should each object be assigned?
+The most frequently asked questions in this area are: how many statistically different objects are present in the data and to which class should each object be assigned?
 Do previously unknown classes of objects appear?
@@
 Clustering techniques are applied to data on the chemical composition of 192 planetary nebulae.
-Six different groups were identified that were similar in many respects to a previous classification of such objects, but also showed interesting differences that researchers had so far missed.
+Six different groups were identified, which were similar in many respects to a previous classification of such objects, but also showed interesting differences that researchers had so far missed.
 
 Clustering based on normal distributions is applied to a set of 2370 stars in \cite{celeux1992classification}.
-They were described by their speed relative to the galactic nucleus and galactic rotation.
+They were described by their relative speed to the galactic nucleus and galactic rotation.
 Using a model of three clusters, they found one cluster of large size and small volume, and two of small size and large volume.
@@
 
-Mind disorders are often more difficult to diagnose than body pathologies, which is why in the field of psychiatry there has been a growing interest in cluster analysis techniques to refine, or even redefine, diagnostic techniques for this type of disease.
+Mind disorders are often more difficult to diagnose than body pathologies, which is why there has been a growing interest in cluster analysis techniques to refine, or even redefine, diagnostic techniques for this type of disease in the field of psychiatry.
 Much of this work involves depressed patients, cases in which the interest lies in distinguishing between two types of depression, endogenous (congenital) and neurotic.
@@
 In \cite{pledger2008using} clustering techniques are applied to 200 patients based on their responses to a questionnaire about depression, along with information about their sex, age, mental state, and illness.
-This is a clear example of different types of variables included in the same data set.
+This is a clear example of different types of variables included in the same dataset.
 One of the groups obtained as a result of this study was identified as a marker of endogenous depression.
@@
 In \cite{paykel1978classification} a study on 236 cases of failed suicides recorded by the emergency service of a city in the United States of America is presented.
-Of the set of possible variables, 14 were selected as particularly relevant for classification and, therefore, were used in the analysis.
-Among them were: age, number of suicide attempts, severity of depression and degree of hostility, in addition to a series of demographic characteristics.
-Clustering methods were applied to the resulting data set; the most significant result obtained corresponds to a division into three well-defined clusters.
+From the set of available variables, 14 were selected as particularly relevant for classification and, therefore, were used in the analysis.
+Among them were: age, number of suicide attempts, severity of depression and degree of hostility, in addition to a series of demographic features.
+Clustering methods were applied to the resulting dataset; the most significant result obtained corresponds to a division into three well-defined clusters.
 
-\subsection{Aplicaciones en meteorología y climatología}
+\subsection{Meteorology and Climatology}
 
-Huge amounts of global weather data are collected daily. Exploring these data using clustering techniques can bring new approaches to study climatology and the environment.
+Huge amounts of global weather data are collected daily. Exploring these data using clustering techniques can bring new approaches to study climatology and environmental phenomena.
 
-In \cite{littmann2000empirical} clustering techniques are applied to data collected on daily changes in surface pressure in the Mediterranean basin, and found 20 groups explaining the variance of rainfall in the central Mediterranean regions.
-Another example can be found in \cite{liu2005mining}, in which the Fuzzy K-means algorithm is applied to spatial-temporal data of the climatology of the southern-central regions of the USA.
+In \cite{littmann2000empirical} clustering techniques are applied to data collected on daily changes in surface pressure in the Mediterranean basin, 20 groups explaining the variance of rainfall in the central Mediterranean regions were found.
+Another example can be found in \cite{liu2005mining}, in which the Fuzzy K-means algorithm is applied to spatial-temporal data of the USA southern-central regions climatology.
 
@@
 Archaeology is another discipline in which clustering is useful.
-The classification of the different objects found in the deposits can help to discover their use, the periods to which they belong, as well as the population that used them.
+The classification of different objects found in deposits can help to discover their use, the periods to which they belong, as well as the population that used them.
 Similarly, studying fossilized materials can help reveal how prehistoric societies lived.
@@
 
-\subsection{Aplicaciones en bioinformática y genética}
+\subsection{Bioinformatics and Genetics}
 
@@
 In order to store the information related to this area of study, microarrays arise \cite{cortese2000array}.
-From the data analysis point of view, one of the relevant features in this type of information is that the number of attributes of each instance ($u$) far exceeds the number of available instances ($n$); data sets such as this are qualified as \textit{high dimensionality data}.
+From the data analysis point of view, one of the relevant features in this type of information is that the number of attributes of each instance ($u$) far exceeds the number of available instances ($n$); datasets featuring this property are called \textit{high dimensionality datasets}.
 
-Most classical statistical methods cannot be applied to this type of dataset without being substantially modified.
-However, cluster analysis accepts such data sets well and can be used to identify groups of genes with similar expression patterns, and answer questions such as why a gene is affected by a certain disease, or which genes are responsible for hereditary genetic diseases.
+Most classic statistical methods cannot be applied to this type of dataset without being substantially modified.
+However, cluster analysis accepts such datasets and can be used to identify groups of genes with similar expression patterns.
+This allows us to answer questions such as why a gene is affected by a certain disease, or which genes are responsible for hereditary genetic diseases.
 
@@
 Although the examples of practical clustering applications presented so far are illustrative, it should be noted that clustering techniques are applied in a multitude of modern problems.
-Since it is not humanly possible to cover all the practical uses of clustering, a brief list of some of them is presented here: cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
-\textcolor{red}{Poner mas ejemplos de aplicación recientes}.
-It seems clear that clustering is an extremely powerful tool, no only for the development of any branch of science, but also because it can be applied to directly help people though medicine or society management.
+Since it is not humanly possible to cover all the practical uses of clustering, a brief list of some of them is presented here: computer vision-guided traffic management \cite{kumaran2019computer}, near real-time detection of
+spatial clusters from large position data streams generated
+by moving objects \cite{junior2019dg2cep}, wireless sensor networks management \cite{wan2019similarity}, internet of things \cite{aranzazu2019anchor}, detecting taxi movements \cite{ibrahim2019detecting}, \textit{big data} distributed multi-target regression \cite{corizzo2019dencast},  cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
+
+It seems clear that clustering is an extremely powerful tool, no only for the development of any branch of science, but also because it can be applied to  many real-world problems and directly help people.
 
@@
 
-Many branches of science have successfully made use of clustering techniques to advance in their respective fields and to process large amounts of data, whose analysis would be unthinkable to confront with traditional techniques.
+Many branches of science have successfully made use of clustering techniques to advance in their respective fields and to process large amounts of data, whose analysis would be unthinkable to tackle with traditional techniques.
 

### 2019-06-24 Memoria/Chapters/ConstrainedClustering.tex
Terminado capitulo 3

@@
 
-Having introduced the basic concepts of clustering, the aim of this chapter is to define the concepts related to constrained clustering, as well as to give examples of its application and highlight the benefits and problems presented by its use.
+Once the basics concepts about clustering have been introduced, we can focus im defining the concepts related to constrained clustering, as well as to give examples of its application and highlight the benefits and problems presented by its use.
 To this end we take as our main reference the survey presented in \cite{davidson2007survey}.
@@
 An example of this can be found in text classification; In \cite{cohn2003semi} a problem proposed by Yahoo!
-is faced.
-The problem consist of, given a large number of text documents, grouping them according to a taxonomy in which documents with similar topics are nearby.
-For this, unsupervised clustering methods are useful, as the information on the problem initially available is limited.
-However, in \cite{wagstaff2001constrained} is shown that applying unsupervised clustering to certain problems, such as grouping GPS data in such a way that clusters define the lanes of a road, does not produce significant results, as the clusters obtained are far from the elongated shape that would be expected as a result.
-To tackle the problem, they introduced a new element in clustering, the instance-level constraints, which made it possible to include knowledge about the clusters that would guide the clustering methods to obtain the expected results.
-It was enough to indicate that the lanes of the road on which the vehicles circulate are four meters wide, and therefore any vehicle that is at a distance of more than 4 meters from another, perpendicular to the direction of displacement, must be located in a different cluster.
+is addressed.
+The problem consist of, given a large number of text documents, grouping them according to a taxonomy in which documents with similar topics are conceptually nearby.
+To solve this problem, unsupervised clustering methods are useful, as the information on the problem initially available is limited.
+However, in \cite{wagstaff2001constrained} is shown that applying unsupervised clustering to certain problems, such as grouping \acf{GPS} data in such a way that clusters define the lanes of a road, does not produce significant results, as the clusters obtained are far from the elongated shape that would be expected as a result.
+To tackle the problem, they introduced a new element in clustering, the constraints, which made it possible to include domain knowledge that would guide the clustering methods to obtain the expected results.
+It was enough to indicate that the lanes of the road on which the vehicles circulate are four meters wide, and therefore any vehicle that is at a distance of more than 4 meters from another, perpendicular to the motion direction, must be located in a different cluster.
 
-We are now in a new scenario: it is possible to incorporate additional information to the clustering process, in addition to that contained in the data set itself, to guide it in the formation of the partition and obtain more precise results.
-This places constrained clustering in the framework of semi-supervised learning, unlike traditional clustering methods which fall within the area of unsupervised clustering.
+We are now in a new scenario: it is possible to incorporate additional information to the clustering process, in addition to that contained in the dataset itself, to guide it in the formation of the output partition and obtain more accurate results.
+This new approach is known as constrained clustering, a \acf{SSL} paradigm, unlike traditional clustering methods which fall within the area of unsupervised learning.
 
@@
 Although they may seem simple, constraints defined as above have some very interesting properties.
-Must-link constraints are an example of an equivalence relationship, and therefore are symmetrical, reflexive and transitive, formalizing:
+\acs{ML} constraints are an example of an equivalence relationship, and therefore are symmetrical, reflexive and transitive.
+Observation \ref{ob:MLConstraintsRelation} formalizes this concept.
+On the other hand, and although \acs{CL} do not constitute an equivalence relation, they feature other properties, formalized in Observation \ref{ob:CLConstraintsRelation}.
 
@@
 	\textbf{Must-link constraints are transitive.} Let $CC_i$ and $CC_j$ be connected components (completely connected subgraphs by \acs{ML} constraints), and let $x_i$ and $x_j$ be the instances in $CC_i$ and $CC_j$ respectively.
-Then $C_=(x,y): x \in CC_i, y \in CC_j \rightarrow C_=(a,b) \forall a,b: a\in CC_i, b \in CC_j$.
-\cite{davidson2007survey}
+Then $C_=(x_i,x_j): x_i \in CC_i, x_j \in CC_j \rightarrow C_=(x_a,x_b) \forall x_a, x_b: x_a\in CC_i, x_b \in CC_j$ \cite{davidson2007survey}.
+	\label{ob:MLConstraintsRelation}
 \end{observation}
@@
 	\textbf{Cannot-link constraints can be entailed.} Let $CC_i$ and $CC_j$ be connected components (completely connected subgraphs by \acs{CL} constraints) and let $x_i$ and $x_j$ be the instances in $CC_i$ and $CC_j$ respectively.
-Then $C_{\neq}(x,y): x \in CC_i, y \in CC_j \rightarrow _{\neq}(a,b) \forall a,b: a\in CC_i, b \in CC_j$.
-\cite{davidson2007survey}
+Then $C_{\neq}(x_i, x_j): x_i \in CC_i, x_j \in CC_j \rightarrow _{\neq}(x_a, x_b) \forall x_a, x_b: x_a\in CC_i, x_b \in CC_j$ \cite{davidson2007survey}
+	\label{ob:CLConstraintsRelation}.
 \end{observation}
 
-A clear example of constraint use can be found in clustering applications where there are distance measurement limitations, as in the case of GPS data.
+A clear example of constraint use can be found in clustering applications where there are distance measurement limitations, as in the case of \acs{GPS} data.
 Thus, if we want that instances forming two clusters to be separated by a distance greater or equal to $\delta$, it is enough to set \acs{ML} constraints between all instances separated by less than $\delta$ units.
@@
 
-While supervised learning involves knowing the label associated with each instance, \acf{SSL} has only a subset of tagged instances.
+While supervised learning involves knowing the label associated with each instance, \acs{SSL} has only a subset of labeled instances.
 On the other hand, in many domains the available information refers to relations between instances, and not to the specific class to which they belong.
-Moreover, in interactive clustering systems, a user who is not expert in the domain of the problem will probably be able to provide information in the form of constraints such as \acf{ML} and \acf{CL}.
-\cite{cohn2003semi}\cite{davidson2007hierarchical}, rather than providing information on what particular class certain instances belong to.
+Moreover, in interactive clustering systems, a user who is not expert in the problem domain will probably be able to provide information in the form of constraints such as \acs{ML} and \acs{CL}, rather than providing information on what particular class certain instances belong to \cite{cohn2003semi,davidson2007hierarchical}.
 
@@
 Alternatively, it is possible to train the distance function used by the method based on the constraints, either before or during the application of the method.
-In any case, the initialization phase can take the constraints into account, so that instances associated with constraints \acf{ML} will be placed in the same cluster, and those between which there is a constraint \acf{CL} will be placed in different clusters.
-Based on this distinction, we identify two ways of approaching the problem, those based on constraints (\textit{constraint-based}), and those based on distances (\textit{distance-based}).
+In any case, the initialization phase can take the constraints into account, so that instances associated with constraints \acs{ML} will be placed in the same cluster, and those between which there is a constraint \acs{CL} will be placed in different clusters.
+Based on this distinction, we identify two ways of approaching the problem, those based on explicit constraints (\textit{constraints-based}), and those based on distances (\textit{distance-based}).
 
@@
 
-In constraint-based methods, the clustering method itself is modified in such a way that the available information is used to bias the search and obtain an appropriate data partition.
+In constraints-based methods, the clustering method itself is modified in such a way that the available information is used to bias the search and obtain an appropriate data partition.
 
-Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained}\cite{davidson2005agglomerative} and soft \cite{law2004clustering}\cite{basu2004active}\cite{segal2003discovering}\cite{davidson2005clustering}\cite{law2005model} constraints.
+Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained,davidson2005agglomerative} and soft \cite{law2004clustering,basu2004active,segal2003discovering,davidson2005clustering,law2005model} constraints.
 Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}.
@@
 
-	\item  Modify the objective function to include a penalty term for breaking constraints \cite{demiriz1999semi} \cite{davidson2005clustering}.
+	\item  Modifying the objective function to include a penalty term for breaking constraints \cite{demiriz1999semi,davidson2005clustering}.
 
-	\item Group instances with additional information obtained from a conditional distribution into an auxiliary space \cite{sinkkonen2000semisupervised}.
+	\item Grouping instances with additional information obtained from a conditional distribution into an auxiliary space \cite{sinkkonen2000semisupervised}.
 
@@
 
-Figure \ref{fig:ConstOverDataset} shows a data set together with its associated constraints. Figure \ref{fig:ClusteringSatAllConst} proposes a possible clustering that satisfies all constraints.
+Figure \ref{fig:ConstOverDataset} shows a dataset together with its associated constraints. Figure \ref{fig:ClusteringSatAllConst} proposes a possible clustering that satisfies all constraints.
 
@@
 	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/InputInstancesAndConst1}
-	\caption[Constraints on a dataset.]{Constraints on a dataset. \cite{davidson2007survey}} \label{fig:ConstOverDataset}
+	\caption[Constraints on a dataset.]{Constraints on a dataset \cite{davidson2007survey}.} \label{fig:ConstOverDataset}
 	}
 	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/ClusteringSatAll}
-	\caption[Clustering satisfying all constraints.]{Clustering satisfying all constraints. \cite{davidson2007survey}} \label{fig:ClusteringSatAllConst}
+	\caption[Clustering satisfying all constraints.]{Clustering satisfying all constraints \cite{davidson2007survey}.} \label{fig:ClusteringSatAllConst}
 	}
@@
 In distance-based approaches, classic clustering methods that use a distance measurement are employed, so that the distance measurement is modified to incorporate the constraints.
-In this context, satisfying constraints means that instances related to constraints \acf{ML} are placed together in the space, and those related by \acf{CL} are separated.
+In this context, satisfying constraints means that instances related to constraints \acs{ML} are placed together in the space, and those related by \acs{CL} are separated.
 
 Figure \ref{fig:MetricLearned} shows a possible clustering based on a metric learned from the constraints specified in Figure \ref{fig:ConstOverDataset2}.
-It should be noted that in Figure \ref{fig:MetricLearned} the space in which the data is located has been compressed on the vertical axis and widened on the horizontal axis to match the learned distance metric.
+It should be noted that in Figure \ref{fig:MetricLearned} the data space has been compressed on the vertical axis and widened on the horizontal axis to match the learned distance metric.
 
@@
 	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/InputInstancesAndConst2}
-	\caption[Constraints on a dataset.]{Constraints on a dataset. \cite{davidson2007survey}} \label{fig:ConstOverDataset2}
+	\caption[Constraints on a dataset.]{Constraints on a dataset \cite{davidson2007survey}.} \label{fig:ConstOverDataset2}
 	}
 	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/MetricaAprendida}
-	\caption[Clustering based on a metric learned based on constraints.]{Clustering based on a metric learned based on constraints. \cite{davidson2007survey}} \label{fig:MetricLearned}
+	\caption[Clustering based on a metric learned based on constraints.]{Clustering based on a metric learned based on constraints \cite{davidson2007survey}.} \label{fig:MetricLearned}
 	}
@@
 	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Caras2}} \quad
-	\caption[CMU database faces.]{CMU database faces. \cite{davidson2007survey}}\label{fig:CMUFacesDatabase}
+	\caption[CMU faces database.]{CMU faces database \cite{davidson2007survey}.}\label{fig:CMUFacesDatabase}
 \end{figure}
 
-The method used to obtain the constraints is one of the most popular in the literature: set the number of clusters of the resulting partition equal to the number of classes in the database, and generate the constraints from a subset of tagged instances; that is, if two instances have different labels, set a \acf{CL} constraint between them, otherwise one of \acf{ML} type.
-Thus, between the images shown in Figure \ref{fig:FacesDatabaseCL} \acf{CL} constraints are set, since, although they belong to the same person, they do not have the same orientation.
+The method used to obtain the constraints is one of the most popular in the literature: set the number of clusters of the resulting partition equal to the number of classes in the database, and generate the constraints from a subset of labeled instances; that is, if two instances have different labels, set a \acs{CL} constraint between them, otherwise one of \acs{ML} type.
+Thus, between the images shown in Figure \ref{fig:FacesDatabaseCL} \acs{CL} constraints are set, since, although they belong to the same person, they do not have the same orientation.
 
@@
 	{\includegraphics[width=.35\linewidth]{gfx/ConstClust/AnalisisImagenes/CarasDifOr2}}
-	\caption[\acs{CL} constraints between faces of the same person.]{\acs{CL} constraints between faces of the same person. \cite{davidson2007survey}}\label{fig:FacesDatabaseCL}
+	\caption[\acs{CL} constraints between faces of the same person.]{\acs{CL} constraints between faces of the same person \cite{davidson2007survey}.}\label{fig:FacesDatabaseCL}
 \end{figure}
@@
 	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo3}} \quad
-	\caption[Clustering method used in Aibo Robot navigation system.]{Clustering method used in Aibo Robot navigation system.
-\cite{davidson2007survey}\cite{davidson2005clustering}}\label{fig:AiboRobotClustSys}
+	\caption[Clustering method used in Aibo robot navigation system.]{Clustering method used in Aibo robot navigation system \cite{davidson2007survey,davidson2005clustering}.}\label{fig:AiboRobotClustSys}
 \end{figure}
@@
 Video databases are one example where constraints can be generated directly from the data domain, especially if space-time data information is available \cite{yan2006discriminative}.
-In time sequenced data it is possible to set \acf{ML} constraints between groups of pixels of frames close in time.
+In time sequenced data it is possible to set \acs{ML} constraints between groups of pixels of frames close in time.
 This is especially useful when the task is to perform object recognition based on clustering and segmentation.
-It is also possible to add \acf{CL} constraints to image segment in the same snapshot, as there is a low probability that they are associated with the same object after segmentation is done.
+It is also possible to add \acs{CL} constraints to image segment in the same snapshot, as there is a low probability that they are associated with the same object after segmentation is done.
 In fact, in the video analysis domain problems there are a variety of constraint extraction methods \cite{yan2006discriminative}, Figure \ref{fig:ConstExtractionVideo} shows some examples.
@@
 	\subfloat[]
-	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoA}}
+	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoA}
+	\label{fig:ConstExtractionVideoA}}
 	\quad
 	\subfloat[]
-	{\includegraphics[width=.225\linewidth]{gfx/ConstClust/Videos/VideoB}} \quad
+	{\includegraphics[width=.225\linewidth]{gfx/ConstClust/Videos/VideoB}
+	\label{fig:ConstExtractionVideoB}} \quad
 	\subfloat[]
-	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoC}}
+	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoC}
+	\label{fig:ConstExtractionVideoC}}
 	\quad
 	\subfloat[]
-	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoD}}
-	\caption[Constraints extraction in video data.]{Constraints extraction in video data. \cite{yan2006discriminative} \cite{davidson2007survey}}\label{fig:ConstExtractionVideo}
+	{\includegraphics[width=.4\linewidth]{gfx/ConstClust/Videos/VideoD}
+	\label{fig:ConstExtractionVideoD}}
+	\caption[Constraints extraction in video data.]{Constraints extraction in video data \cite{yan2006discriminative,davidson2007survey}.}\label{fig:ConstExtractionVideo}
 \end{figure}
 
-In Figure \ref{fig:ConstExtractionVideo}, image (a) corresponds to constraints extracted from tracking a person over a period of time, (b) corresponds to spatial constraints associating two objects located in the same frame, image (c) corresponds to restrictions obtained through facial recognition and (d) to those provided by the user.
+In Figure \ref{fig:ConstExtractionVideo}, image \ref{fig:ConstExtractionVideoA} corresponds to constraints extracted from tracking a person over a period of time, \ref{fig:ConstExtractionVideoB} corresponds to spatial constraints associating two objects located in the same frame, image \ref{fig:ConstExtractionVideoC} corresponds to restrictions obtained through facial recognition and \ref{fig:ConstExtractionVideoD} to those provided by the user.
 
@@
 
-In gene clustering based on micro-array data, genes are represented by their expression profile in different experiments and grouped using different methods, in this case constrained clustering methods.
-Figure \ref{fig:GeneticApp} shows an example: these are \acf{ML} constraints set between genes based on co-occurrence data stored in the protein interaction database, which contains information about which genes (and their associated proteins) are present in the same cell processes \cite{xenarios2001dip}.
+In gene clustering based on micro-array data, genes are represented by their expression profile in different experiments and grouped using different constrained clustering methods.
+Figure \ref{fig:GeneticApp} shows an example: these are \acs{ML} constraints set between genes based on co-occurrence data stored in the protein interaction database, which contains information about which genes (and their associated proteins) are present in the same cell processes \cite{xenarios2001dip}.
 This information can be used to improve results provided by clustering methods.
@@
 	\includegraphics[scale=0.3]{gfx/ConstClust/Genetica/Genes} 
-	\caption[Gene clustering based on micro-array data.]{Gene clustering based on micro-array data. \cite{davidson2007survey}}\label{fig:GeneticApp}
+	\caption[Gene clustering based on micro-array data.]{Gene clustering based on micro-array data \cite{davidson2007survey}.}\label{fig:GeneticApp}
 \end{figure}
@@
 In this case it is possible to extract constraints from multiple auxiliary resources.
-For example, if two documents are in the same directory, a \acf{ML} constraints could be set between them.
+For example, if two documents are in the same directory, a \acs{ML} constraints could be set between them.
 This way it is possible to adjust the resulting clustering to suit a particular criterion, such as creating a hierarchy of documents similar to the way they are organized in the input directories structure.
@@
 
-El clustering con restricciones resulta de gran utilidad en el procesamiento de datos de búsqueda en páginas web.
-Aquí, el objetivo es agrupar, de manera automática, los resultados de una consulta ambigua en el motor de búsqueda en clusters de \acs{URL}s, que se refieran al concepto introducido como consulta en diferentes contextos.
-En este ámbito es posible extraer las restricciones a partir de búsquedas realizadas anteriormente por los usuarios, de manera que se establece una restricción de tipo \acf{ML} entre \acs{URL}s visitadas en la misma sesión de usuario.
-Aplicar clustering utilizando estas restricciones puede ayudar a sesgar el resultado de las búsquedas hacia las preferencias del usuario.
+Constrained clustering is very useful in web search data processing.
+Here, the aim is to automatically group the results of an ambiguous query in the search engine into clusters of \acs{URL}s, which refer to the concept introduced as a query in different contexts.
+In this area it is possible to extract the constraints from searches previously performed by users, so that a \acs{ML} constraints is set between \acs{URL}s visited in the same user session.
+Clustering using these constraints can help bias the result of searches towards user's preferences.
 
-\subsection{Aplicaciones en datos de audio}
+\subsection{Audio Data}
 
@@
 Here, the aim is to automatically group the results of an ambiguous query in the search engine into clusters of \acs{URL}s, which refer to the concept introduced as a query in different contexts.
-In this area it is possible to extract constraints from searches previously performed by users, so that a \acf{ML} constraint is set between visited \acs{URL}s in the same user session.
+In this area it is possible to extract constraints from searches previously performed by users, so that a \acs{ML} constraint is set between visited \acs{URL}s in the same user session.
 Clustering using these constraints can help biasing search results towards user preferences.
 
-\subsection{Aplicaciones en datos de GPS} \label{sec:GPSApp}
+\subsection{GPS Data} \label{sec:GPSApp}
 
@@
 	\includegraphics[scale=0.3]{gfx/ConstClust/GPS/Coches} 
-	\caption[GPS data use.]{GPS data use. \cite{davidson2007survey} \cite{wagstaff2001constrained}}\label{fig:GPSClustData}
+	\caption[GPS data use.]{\acs{GPS} data use. \cite{davidson2007survey} \cite{wagstaff2001constrained}}\label{fig:GPSClustData}
 \end{figure}
@@
 To get the resulting clusters to have this shape, we can make use of the constraints.
-We will set a \acf{CL} constraint between those instances more than 4 meters perpendicularly to the motion direction (since the lanes have a maximum width of 4 meters), and \acf{ML} constraints between those instances that present continuity in the motion direction axis, since it is likely that the vehicles they represent are in the same lane.
+We will set a \acs{CL} constraint between those instances more than four meters apart perpendicularly to the motion direction (since the lanes have a maximum width of four meters), and \acs{ML} constraints between those instances that present continuity in the motion direction axis, since it is likely that the vehicles they represent are in the same lane.
 This clustering model has proven to be very useful in real-time navigation \cite{wagstaff2001constrained}, allowing the user to be notified when to change lanes, or when not to leave them.
@@
 	\includegraphics[scale=0.32]{gfx/ConstClust/GPS/Instancias} 
-	\caption[Resulting Clustering from GPS data with no constraints.]{Resulting Clustering from GPS data with no constraints. \cite{davidson2007survey} \cite{wagstaff2001constrained}}\label{fig:figure15}
+	\caption[Clustering of GPS data with no constraints.]{Clustering of \acs{GPS} data with no constraints \cite{davidson2007survey,wagstaff2001constrained}.}\label{fig:figure15}
 \end{figure}
@@
 
-The named constrained clustering applications are the most studied in the literature, but it is worth noting that is has been applied to many other fields in recent times.
-The following is a list of some of those applications: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, among others.
-\textcolor{red}{Poner mas ejemplos de aplicación recientes}
+The already mentioned constrained clustering applications are the most studied in the literature, although the reader may have the impression that constrained clustering is no longer used in real-world problems.
+To prevent that we provide a list of some more recent applications of constrained clustering to modern problems: advanced robotics navigation systems \cite{semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, semantically similar relations identification \cite{wang2015constrained}, cooperative team building \cite{yang2014team}, video face clustering \cite{zhou2014video}, image clustering \cite{habashi2017semi} and storage location assignment in warehouses \cite{yang2016constrained} among others.
 
@@
 
-Given $X = \{x_1 \cdots x_u\}$, a large set of unlabeled instances, and $L = \{(x_{u+1}, y_{u+1}) \cdots (x_{u+l}, y_{u+l})\}$, a small set of labeled instances, it is common to choose two elements from $L$ (with replacement) and set a \acs{ML} constraint between them if they belong to the same class, or a \acs{CL} constraints otherwise.
-An appropriate method to evaluate the results provided by a clustering method is to measure the level of accuracy of this when predicting dataset $X$ labels.
+Given $X = \{x_1 \cdots x_n\}$, a large set of unlabeled instances, and $L = \{(x_{n+1}, l_{n+1}) \cdots (x_{n+|L|}, l_{n+|L|})\}$, a small set of labeled instances, it is common to choose two elements from $L$ (with replacement) and set a \acs{ML} constraint between them if they belong to the same class, or a \acs{CL} constraints otherwise.
+An appropriate method to evaluate the results provided by a clustering method is to measure the level of accuracy it achieves when predicting dataset $X$ labels.
 This normally requires specifying the number of desired clusters equal to the number of known classes in $X$ ($k = k^*$).
@@
 
-In \cite{wagstaff2000clustering}, where constraints are generated as described above, it is shown that, when an average of the accuracy of the predictions obtained with constrained clustering algorithms is made, varying the latter between experiments, results are obtained up to 20\% better than with classical clustering techniques.
+In \cite{wagstaff2000clustering}, where constraints are generated as described above, it is shown that accuracy is increased up to a 20\% over classic methods when averaging results over different constraint sets.
 
@@
 
-	\textbf{Using constraints increases accuracy on average.}
-	The performance of a method in predicting labels increases when averaged using numerous different constraint sets. \cite{davidson2007survey}
+	\textbf{Using constraints increases average accuracy.}
+	The performance of a method in predicting labels increases when averaged using numerous different constraint sets \cite{davidson2007survey}.
 	\label{ob:AccuracyIncrease}
@@
 
-This rule, however, is not true in all cases, as in data sets such as \textit{Tic-Tac-Toe Endgame}, no increase in predictions is achieved regardless of the number of constraints used.
-A possible explanation for these exceptions is based on the fact that setting $k = k^*$ is not appropriate in these cases.
+This rule, however, is not true in all cases, as in datasets such as \textit{Tic-Tac-Toe Endgame}, no increase in predictions is achieved regardless of the number of constraints used.
+A possible explanation for these exceptions is based on the fact that setting $k = k^*$ may not be appropriate in these cases.
 
@@
 
-Although, as we have seen, the incorporation of constraints to clustering methods brings benefits in some applications, there are two main drawbacks that are discussed below, as well as possible solutions to them.
+Although, as we have seen, incorporating constraints to clustering methods brings benefits in some applications, there are two main drawbacks that are discussed below, as well as possible solutions to them.
 
@@
 For instance, there is no partition that satisfies the constraints $C_=(x_i,x_j)$ and $C_{\neq}(x_i,x_j)$, regardless of the value of $k$.
-The same happens for $k = 2$, and the constraints $C_{\neq}(x_i, x_j)$, $C_{\neq}(x_j, x_q)$ and $CL(x_i, x_q)$.
+The same happens for $k = 2$, and the constraints $C_{\neq}(x_i, x_j)$, $C_{\neq}(x_j, x_q)$ and $C_{\neq}(x_i, x_q)$.
 Definition \ref{def:FeasibilityProblem} formalizes the feasibility problem for non-hierarchical instance-level constrained clustering:
@@
 	\textbf{Feasibility problem for non-hierarchical instance-level constrained clustering:} Given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
-\cite{davidson2005clustering} \cite{davidson2007survey}
+\cite{davidson2005clustering,davidson2007survey}
 
@@
 	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
+	%\setlength{\tabcolsep}{7pt}
+	%\renewcommand{\arraystretch}{1}
 	%\resizebox{\textwidth}{!}{
-	\begin{tabular}{ >{\centering\arraybackslash}m{4cm}  >{\centering\arraybackslash}m{4cm} }
+	\begin{tabular}{ >{\centering\arraybackslash}m{3cm}  >{\centering\arraybackslash}m{3cm} }
 		\hline
 		\textbf{Constraints} & \textbf{Complexity} \\
+		\hline
 		$\delta$ & $\mathbf{P}$ \\
@@
 	\end{tabular}%}
-	\caption[Complexity of the constrained clustering problem depending on the type of constraints.]{Complexity of the constrained clustering problem depending on the type of constraints.
-\cite{davidson2007survey}}
+	\caption[Complexity of the constrained clustering problem.]{Complexity of the constrained clustering problem \cite{davidson2007survey}.}
 \label{tab:CCComplexity}
@@
 
-As shown in Table \ref{tab:CCComplexity}, the use of \acf{CL} constraints increases the constrained clustering complexity level to $\mathbf{NP}$-complete and therefore constrained clustering is intractable.
+As shown in Table \ref{tab:CCComplexity}, the use of \acs{CL} constraints increases the constrained clustering complexity level to $\mathbf{NP}$-complete and therefore constrained clustering is intractable.
 Intuitively it can be easily understood that if finding a single partition that satisfies the constraints is a complex problem, finding the best of them is harder.
@@
 
-In \cite{wagstaff2002intelligent} and \cite{davidson2007hierarchical} is shown that, even in the number of clusters in the output partition is set equal to the number of classes in the dataset ($k = k*$), which guarantees that a feasible solution does exists, simple algorithms like the COP-kmeans \cite{wagstaff2001constrained} may not converge due to the feasibility problem.
+In \cite{wagstaff2002intelligent} and \cite{davidson2007hierarchical} is shown that, even in the number of clusters in the output partition is set equal to the number of classes in the dataset ($k = k*$), which guarantees that a feasible solution does exists, simple algorithms like COP-kmeans \cite{wagstaff2001constrained} may not converge due to the feasibility problem.
 
@@
 
-The problem over-constraining phenomenon through the use of \acf{CL} constraints is intimately related to the graph coloring problem; in fact, it has been proven that it is equivalent to constrained clustering with \acf{CL} \cite{davidson2006identifying}.
-Thus, we find that solving a problem with \acf{CL} constraints using algorithms such as COP-K-medias is, for practical purposes, solving the graph coloring problem.
+The problem over-constraining phenomenon through the use of \acs{CL} constraints is intimately related to the graph coloring problem.
+It has been proven that constrained clustering with \acs{CL} is equivalent to the graph coloring problem \cite{davidson2006identifying}.
+Thus, we find that solving a problem with \acs{CL} constraints using algorithms such as COP-Kmeans is, for practical purposes, solving the graph coloring problem.
 
@@
 
-	\textbf{Constrained clustering with \acs{CL} is analogous to the graph coloring problem.} \cite{davidson2007survey}
+	\textbf{Constrained clustering with \acs{CL} is analogous to the graph coloring problem} \cite{davidson2007survey}.
 
@@
 	\textbf{Brook's theorem applies to the constrained clustering.}
-	If $ k > $ (Most \acs{CL} constraints on an instance), then there always exists a feasible partition. \cite{davidson2007survey} \label{ob:BrooksTheorem}
+	If $ k > $ (Most \acs{CL} constraints on an instance), then there always exists a feasible partition \cite{davidson2007survey} \label{ob:BrooksTheorem}.
 
@@
 With this, and although Observation \ref{ob:FeasibleSolution} indicates differently, when constrained clustering meets the condition set out in Observation \ref{ob:BrooksTheorem}, we can guarantee that a solution will always be found in polynomial time.
-To ensure Brook's condition, it is possible to build the constraint set so that no instance takes a side in more than $K$ \acf{CL} constraints \cite{davidson2006identifying}.
+To ensure Brook's condition, it is possible to build the constraint set so that no instance takes a side in more than $k$ \acs{CL} constraints \cite{davidson2006identifying}.
 
@@
 	\includegraphics[scale=0.4]{gfx/ConstClust/Inform/Inform} 
-	\caption[Informative constraint set.]{Informative constraint set \cite{davidson2007survey}}\label{fig:Informativity}.
+	\caption[Informative constraint set.]{Informative constraint set \cite{davidson2007survey}.}\label{fig:Informativity}
 \end{figure}
@@
 \begin{equation}
-I_A(R) = \frac{1}{|R|}\left[ \sum_{r \in R} unsat(r, P_A) \right] 
+I_A(R) = \frac{1}{|R|}\left[ \sum_{r \in R} \text{unsat}(r, P_A) \right].
 \label{eq:Informativity}
@@
 
-\textbf{Coherence} measures the degree of agreement within the set of constraints itself with respect to a given metric ($D$).
+\textbf{Coherence} measures the degree of agreement within the set of constraints itself with respect to a given metric.
 For instance, Figure \ref{fig:Coherence} shows two parallel and very close constraints of different type.
-It is in cases like this that the contradiction occurs, as the \acf{ML} constraints indicate that distance between instances involved in them is small, while \acf{CL}s indicate the opposite.
+It is in cases like this that the contradiction occurs, as the \acs{ML} constraints indicate that distance between instances involved in them is small, while \acs{CL}s indicate the opposite.
 
@@
 	\includegraphics[scale=0.4]{gfx/ConstClust/Coherencia/Coher1}
-	\caption[Non-coherent constraint set.]{Non-coherent constraint set. \cite{davidson2007survey}}\label{fig:Coherence}
+	\caption[Non-coherent constraint set.]{Non-coherent constraint set \cite{davidson2007survey}.}\label{fig:Coherence}
 \end{figure}
@@
 Constrained clustering adds a new type of information to the original clustering problem, which is given in the form of specifications of belonging to the same or to different clusters on pairs of instances.
-Constraints, whether if they are the pairwise \acf{ML} and \acf{CL} constraints, or the distance-based $\delta$ and $\epsilon$ constraints, they are used to guide the clustering method we apply to a dataset in the search for a result partition that meets certain desired characteristics.
+Constraints, whether if they are the pairwise \acf{ML} and \acf{CL} constraints, or the distance-based $\delta$ and $\epsilon$ constraints, they are used to guide the clustering method we apply to a dataset in the search for a result partition that meets certain desired properties.
 
-Clustering methods derived from this concept have proved to be very useful in many areas, as well as they present some problems that can be solved by deeply studying the constraints to be used to solve each individual problem.
+Clustering methods derived from this concept have proved to be advantageous in many areas, as well as they present some problems that can be solved by deeply studying the constraints to be used to solve each individual problem.

### 2019-06-24 Memoria/Chapters/Introduction.tex
Terminado capitulo 3

@@
 
-One of the most basic and primitive abilities that human beings are endowed with is that of grouping similar objects to produce a classification that is useful to them.
+One of the most basic and primitive abilities that human beings are endowed with is being able to group similar objects to produce a classification that is useful to them.
 For example, they should have been able to realize which objects were edible, which were poisonous, and which would try to kill them.
@@
 Just as classification is a basic skill for people in their daily lives, it is also essential in most branches of science.
-In biology, for example, the classification of different types of organisms has been the subject of study since the beginning of their existence.
-Aristotle built an elaborated animal classification system that divided all creatures into two groups, those with red blood and those without.
+In biology, for example, the classification of different types of organisms has been the subject of study since the beginning of its existence.
+Aristotle built an elaborated animal classification system, dividing all creatures into two groups: red-blooded and not red-blooded.
 Later he proposed a subdivision that classified them according to the way in which new individuals came into the world, either alive, by means of eggs, chrysalis, etc.
@@
 	\includegraphics[scale=0.35]{gfx/Intro/HR_diagram} 
-	\caption[Hertzsprung-Russell starts classification diagram]{Hertzsprung-Russell starts classification diagram}\label{fig:intro_HRdiagram}
+	\caption[Hertzsprung-Russell stars classification diagram.]{Hertzsprung-Russell stars classification diagram.}\label{fig:intro_HRdiagram}
 \end{figure}
 
-\section{The learning task}
+\section{The Learning Task}
 
@@
 
-One of the major differences between human learning and machine learning, and we must admit that in this the machines again lead us, is the way in which knowledge is transferred.
+One of the major differences between human learning and machine learning, and we must admit that to this respect machines again lead us, is the way in which knowledge is transferred.
 In machines it is enough to copy and paste, that is to say, once a program performs correctly the task for which it has been designed in a computer, it performs it in all of them.

### 2019-06-26 Memoria/Chapters/ExperimentalResults.tex
Nuevos capitulos en memoria y correcciones en papers

\chapter{Experimental Results} \label{ch:ExperimentalResults}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 0.022 & 5.454 & 383.642 && \textbf{0.086} & 5.091 & 366.491 \\
			Breast Cancer & \textbf{0.395} & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
			Bupa & \textbf{0.281} & 14.151 & 1420.117 && 0.104 & 16.672 & 1409.487 \\
			Circles & \textbf{0.166} & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
			Ecoli & 0.009 & 20.321 & 1169.798 && \textbf{0.020} & 12.763 & 1112.981 \\
			Glass & \textbf{0.016} & 8.658 & 735.507 && 0.013 & 4.848 & 694.962 \\
			Haberman & 0.047 & 15.699 & 1155.985 && \textbf{0.084} & 15.441 & 1146.023 \\
			Hayesroth & \textbf{0.027} & 1.500 & 550.117 && 0.025 & 2.000 & 529.424 \\
			Ionosphere & 0.125 & 17.587 & 1894.390 && \textbf{0.332} & 12.603 & 1872.308 \\
			Iris & 0.238 & 0.190 & 505.112 && \textbf{0.342} & 0.000 & 483.584 \\
			Monk2 & \textbf{0.421} & 14.926 & 1848.543 && 0.270 & 16.385 & 1821.208 \\
			Moons & \textbf{0.277} & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
			Saheart & 0.147 & 20.555 & 2135.624 && \textbf{0.265} & 16.725 & 2095.554 \\
			Sonar & \textbf{0.124} & 7.619 & 1106.039 && 0.116 & 8.952 & 1081.994 \\
			Soybean & 0.496 & 0.000 & 176.333 && \textbf{0.515} & 0.000 & 151.163 \\
			Spectfheart & 0.334 & 9.573 & 1451.235 && \textbf{0.343} & 8.604 & 1405.243 \\
			Spiral & 0.058 & 15.402 & 1098.622 && \textbf{0.114} & 13.793 & 1099.747 \\
			Tae & 0.018 & 0.833 & 520.855 && \textbf{0.024} & 1.333 & 508.988 \\
			Vowel & \textbf{0.003} & 13.330 & 4245.672 && 0.002 & 11.523 & 4290.154 \\
			Zoo & \textbf{0.105} & 1.091 & 336.345 && 0.098 & 2.182 & 317.674 \\
			\hline
			Mean & \textbf{0.161} & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
	\label{tab:results10}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 0.296 & 6.167 & 381.625 && \textbf{0.332} & 5.833 & 362.106 \\
			Breast Cancer & 0.772 & 10.172 & 3854.911 && \textbf{0.911} & 3.310 & 3766.402 \\
			Bupa & \textbf{0.840} & 5.973 & 1463.152 && 0.801 & 5.641 & 1437.639 \\
			Circles & 0.828 & 6.101 & 1136.484 && \textbf{0.982} & 0.404 & 1108.928 \\
			Ecoli & 0.015 & 24.314 & 1200.463 && \textbf{0.025} & 19.090 & 1127.568 \\
			Glass & 0.017 & 17.500 & 745.987 && \textbf{0.031} & 11.743 & 722.009 \\
			Haberman & \textbf{0.866} & 4.677 & 1209.471 && 0.796 & 5.314 & 1147.479 \\
			Hayesroth & \textbf{0.107} & 7.536 & 545.861 && 0.099 & 8.261 & 520.702 \\
			Ionosphere & \textbf{0.803} & 7.722 & 1986.543 && 0.591 & 11.974 & 1944.085 \\
			Iris & \textbf{0.412} & 3.873 & 506.548 && 0.260 & 6.087 & 486.723 \\
			Monk2 & 0.751 & 10.827 & 1949.305 && \textbf{0.945} & 2.067 & 1920.361 \\
			Moons & 0.958 & 1.334 & 1133.858 && \textbf{0.995} & 0.020 & 1113.687 \\
			Saheart & 0.749 & 10.940 & 2281.992 && \textbf{0.945} & 1.839 & 2225.210 \\
			Sonar & 0.800 & 5.040 & 1124.700 && \textbf{0.890} & 2.177 & 1069.862 \\
			Soybean & \textbf{0.428} & 0.000 & 170.674 && 0.422 & 0.000 & 155.948 \\
			Spectfheart & \textbf{0.924} & 2.854 & 1561.040 && 0.778 & 5.732 & 1485.260 \\
			Spiral & 0.847 & 5.677 & 1133.962 && \textbf{0.949} & 1.495 & 1102.920 \\
			Tae & \textbf{0.119} & 7.115 & 522.088 && 0.066 & 8.380 & 513.289 \\
			Vowel & \textbf{0.002} & 14.486 & 4807.165 && 0.001 & 13.341 & 4867.849 \\
			Zoo & \textbf{0.157} & 3.500 & 333.984 && 0.128 & 2.834 & 315.131 \\
			\hline
			Mean & 0.483 & 9.456 & 1784.931 && \textbf{0.507} & 6.997 & 1751.368 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
	\label{tab:results15}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
			Bupa & 0.936 & 2.975 & 1529.151 && \textbf{0.993} & 0.281 & 1508.772 \\
			Circles & 0.935 & 2.814 & 1180.268 && \textbf{1.000} & 0.000 & 1170.452 \\
			Ecoli & 0.028 & 25.716 & 1267.419 && \textbf{0.047} & 21.466 & 1277.903 \\
			Glass & 0.038 & 23.610 & 766.861 && \textbf{0.082} & 16.412 & 796.162 \\
			Haberman & 0.943 & 2.073 & 1266.131 && \textbf{0.997} & 0.138 & 1235.441 \\
			Hayesroth & \textbf{0.548} & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
			Ionosphere & 0.902 & 4.314 & 2051.251 && \textbf{0.993} & 0.258 & 2019.558 \\
			Iris & \textbf{0.691} & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
			Monk2 & 0.855 & 6.025 & 2055.425 && \textbf{0.991} & 0.379 & 2013.468 \\
			Moons & 0.937 & 2.689 & 1177.848 && \textbf{0.995} & 0.158 & 1157.222 \\
			Saheart & 0.881 & 5.250 & 2402.513 && \textbf{0.981} & 0.706 & 2353.931 \\
			Sonar & 0.992 & 0.186 & 1138.769 && \textbf{1.000} & 0.000 & 1127.020 \\
			Soybean & 0.416 & 0.000 & 171.281 && \textbf{0.563} & 0.000 & 153.406 \\
			Spectfheart & 0.953 & 1.677 & 1602.229 && \textbf{1.000} & 0.000 & 1526.925 \\
			Spiral & 0.921 & 3.604 & 1178.580 && \textbf{0.998} & 0.091 & 1155.194 \\
			Tae & 0.351 & 9.807 & 527.292 && \textbf{0.392} & 9.419 & 508.610 \\
			Vowel & 0.002 & 14.682 & 5450.331 && \textbf{0.003} & 13.766 & 5514.052 \\
			Zoo & \textbf{0.176} & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
			\hline
			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
	\label{tab:results20}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
			Breast Cancer & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
			Circles & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
			Ecoli & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
			Glass & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
			Haberman & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
			Hayesroth & 0.031 & 0.027 & - & \textbf{0.106} & 0.097 & 0.054 & 0.139 \\
			Ionosphere & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
			Iris & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
			Monk2 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
			Moons & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
			Saheart & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
			Sonar & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
			Spectfheart & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
			Spiral & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
			Tae & \textbf{0.031} & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
			Vowel & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
			Zoo & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
			\hline
			Mean & \textbf{0.474} & 0.158 & -0.510 & 0.226 & 0.212 & 0.175 & -0.186 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{10}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results10}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
			Breast Cancer & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
			Bupa & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
			Circles & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
			Ecoli & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
			Glass & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
			Haberman & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
			Hayesroth & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
			Ionosphere & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
			Iris & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
			Monk2 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
			Moons & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
			Saheart & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
			Sonar & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
			Spectfheart & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
			Spiral & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
			Tae & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
			Zoo & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
			\hline
			Mean & \textbf{0.641} & 0.490 & -0.070 & 0.225 & 0.233 & 0.258 & -0.196 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{15}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results15}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
			Breast Cancer & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
			Glass & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
			Haberman & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
			Iris & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
			Saheart & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
			Sonar & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
			Spectfheart & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
			Spiral & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
			Tae & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
			Zoo & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
			\hline
			Mean & \textbf{0.733} & 0.642 & 0.064 & 0.223 & 0.285 & 0.344 & -0.187 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{20}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results20}
\end{table}

### 2019-06-26 Memoria/Chapters/ExperimentalSetup.tex
Nuevos capitulos en memoria y correcciones en papers

--- 2019-06-14 Papers/DILS/paperDILS.tex
@@
-\documentclass[review]{elsarticle}
+\chapter{Experimental Setup} \label{ch:ExperimentalSetup}
 
-\usepackage{graphics, float, url}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{lineno,hyperref}
-\usepackage{mathtools}
-\usepackage{multirow}
-\usepackage{adjustbox}
-\usepackage{chngpage}
-\usepackage{setspace}
-\usepackage{amsfonts}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage[ruled, lined, onelanguage, linesnumbered]{algorithm2e}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\newcommand{\myfloatalign}{\centering}
-\modulolinenumbers[5]
+In this section we present the experimental setup used to compare the two new proposals introduced in Part \ref{pt:NewProposals} with the current state of the art.
+We will start by briefly introducing the state-of-the-art methods.
+Afterwards we specify the dataset used in experimentation, as well as the constraint sets and the results evaluation method.
+We also include the required specifications to replicate the experiments.
 
-\usepackage{epstopdf}
-\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
-\AppendGraphicsExtensions{.tiff}
-
-\journal{Computers and Operations Research}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-\newtheorem{definition}{Definition}
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Improving instance-level constrained clustering through Dual Iterative Local Search}
-
-\author[mymainaddress]{Germ\'an Gonz\'alez-Almagro\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{germangalmagro@ugr.es}
-
-\author[mymainaddress]{Juli\'an Luengo}
-
-\author[mysecondaddress]{Jos\'e-Ram\'on Cano}
-
-\author[mymainaddress]{Salvador Garc\'ia}
-
-\address[mymainaddress]{DaSCI Andalusian Institute of Data Science and Computational Intelligence, University of Granada, Spain}
-
-\address[mysecondaddress]{Dept. of Computer Science, EPS of Linares, University of Ja\'en, Campus Cient\'ifico Tecnol\'ogico de Linares, Cintur\'on Sur S/N, Linares 23700, Ja\'en, Spain}
-
-\begin{abstract}
-Clustering has always been a powerful tool in knowledge discovery.
-Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning.
-This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to provide quiality results for the constrained clustering problem.
-We will compare the results obtained by this proposal to those obtained by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
-\end{abstract}
-
-\begin{keyword}
-constrained clustering, instance-level, must-link, cannot-link, dual iterative local search.
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
-
-\section{Introduction} \label{sec:Intro}
-
-One of the most widely known and studied data analysis problems is clustering.
-It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
-Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics,\cite{Everitt:2009:CA:1538772} cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
-
-In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
-In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
-A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
-In this paper we will focus on partitional clustering.
-
-We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them.
-A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features.
-More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$.
-A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$.
-As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement \cite{jain1999data}.
-
-Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
-We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}.
-The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
-The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{triguero2015self}.
-
-Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
-It has been successfully applied in many knowledge fields, among which it is worth mentioning: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, and lane finding in GPS data \cite{wagstaff2001constrained} among others.
-
-Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level \cite{bradley2000constrained}, instance-level \cite{davidson2007survey} and feature-level constrained clustering \cite{schmidt2011clustering}.
-Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.
-
-In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints.
-In one hand, pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
-On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure \cite{davidson2007survey}.
-This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
-
-Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained} and soft \cite{law2004clustering} constraints.
-Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}.
-For the purposes of this paper, we will employ the latter.
-
-Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem.
-Therefore, the incorporation of constraints may modify the complexity of the clustering problem, depending on the type of constraints used.
-As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete \cite{davidson2005clustering}.
-
-The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it.
-As mentioned earlier, it is a difficult problem to solve, so metaheuristic are presented as a promising option to find quality approximate solutions.
-Metaheuristic algorithms are designed to explore the solution space associated with a problem guided by a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
-In this type of approaches the key is to find a good diversification-intensification tradeoff.
-This is the reason why the classic Local Search (LS) heuristic algorithm derived in the Iterative Local Search (ILS) metaheuristic algorithm, which introduces periodics perturbations in the exploration of the solution space to avoid local optima \cite{lourencco2010iterated}.
-ILS---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
-ILS is also very commonly used to design hybrid methods with good diversification-intensification tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
-
-In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptative and guided way.
-We apply it to the constrained clustering problem, prooving that it constitute a good approach to obtain quality results.
-
-Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and the state-of-the-art.
-In Sections \ref{sec:DILS} and \ref{sec:DILS_CC} the DILS algorithm and its constrained clustering formulation will be reviewed.
-Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
-Finally, in Section \ref{sec:conclusiones} conclusions are discussed.
-
-\section{Background} \label{sec:background}
-
-In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering (Section \ref{sec:BackSOTA}).
-
-\subsection{Constrained Clustering} \label{sec:BackCC}
-
-In most clustering applications it is common to have some kind of information about the dataset to be analyzed.
-In pairwise instance-level constrained clustering this information is given in the form of pairs of instances.
-A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster.
-It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms.
-We can now formalize the two type of constraints mentioned:
-
-\begin{itemize}
-
-	\item Must-link constraints $C_=(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
-
-	\item Cannot-link constraints $C_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.
-
-\end{itemize}
-
-The goal of constrained clustering is to find a partition (or clustering) of $k$ clusters $C = \{c_1, \cdots, c_k\}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set.
-As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $n = |X| = \sum_{i = 1}^{k} |c_i|$.
-
-Knowing how a constraint is defined, ML constraints are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric.
-This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified.
-In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
-
-It can also be proven that CL constraints do not constitute an equivalence relation.
-However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $C_{\neq}(x_a,x_b)$, then it is also true that $C_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
-
-\subsection{The Feasibility Problem} \label{sec:BackFeas}
-
-Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem.
-The feasibility problem for non-hierarchical instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
-
-\begin{definition}
-
-	\textbf{Feasibility Problem}: given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
-\cite{davidson2005clustering}
-	\label{def1}
-
-\end{definition}
-
-In \cite{davidson2005clustering} it is proven that, when $k_l = 1$ and $k_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem (it is also proven that it is not harder, so both have the same complexity).
-Table \ref{tab:feasibility} shows the complexity of the feasibility for different types of constraints.
-
-\begin{table}[!h]
-	\centering
-	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\resizebox{\textwidth}{!}{
-		\begin{tabular}{c c}
-			\hline
-			Constraints & Complexity \\
-			\hline
-			Must-Link & $\mathbf{P}$\\
-			Cannot-Link & $\mathbf{NP}$-complete\\
-			ML and CL & $\mathbf{NP}$-complete\\
-			\hline
-
-		\end{tabular}%}
-	\caption{Feasibility problem complexity \cite{davidson2005clustering}.}
-	\label{tab:feasibility}
-\end{table}
-
-These complexity results show that the feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too.
-For more details on the complexity of constrained clustering see \cite{davidson2005clustering}.
-
-Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why metaheuristic algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
-
-\subsection{State-of-the-art Methods} \label{sec:BackSOTA}
+\subsection{State-of-the-art Methods} \label{sec:SOTA}
 
@@
 It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
-This method was named COP-kmeans, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
+This method was named \acf{COPKM}, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
 
-In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
+In \cite{antoine2012cecm} \acf{CECM}, a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
 The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
@@
 The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity.
-Linear CVQE (LCVQE) introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex.
+\acf{LCVQE} introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex.
 The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.
 
-Two Views Clustering (TVClust) and Relation Dirichlet Process - Means (RDPM) were proposed in \cite{khashabi2015clustering}.
-TVClust is able to incorporate the constraints into the clustering problem by making a soft interpretation of them.
+\acf{TVClust} and \acf{RDPM} were proposed in \cite{khashabi2015clustering}.
+\acs{TVClust} is able to incorporate the constraints into the clustering problem by making a soft interpretation of them.
 The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
-Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
+Using this model as a basis, the authors derive the \acs{RDPM} deterministic algorithm.
 This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($k$) does not need to be specified.
 
-An adaptation of the Biased Random-Key Genetic Algorithm (BRKGA) \cite{gonccalves2011biased}---a variant of the Random-Key genetic algorithm \cite{bean1994genetic}---for the constrained clustering problem is proposed in \cite{de2017comparison}.
-It uses a genetic algorithm, combined with an LS procedure and guided by a penalty-style fitness function, to obtain a partition of the input dataset.
-It constitutes one of the few approaches to constrained clustering from the point of view of the evolutionary computing.
+\section{Datasets}
 
-\clearpage
-
-\section{The Dual Iterative Local Search Method} \label{sec:DILS}
-
-The Dual Iterative Local Search (DILS) is a new variant of the classic ILS method.
-Its goal is to perform a search in the solution space to find the solution with the best fitness function $f$ value by introducing diversity in an adaptative and guided way.
-While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-introducing methods and to avoid local optima.
-The individual $m_b$ is the one providing the best fitness value, so $m_w$ provides the worst fitness value.
-
-In each generation $G$ of the optimization process DILS builds a new individual by applying a recombination operator to $m_b$ and $m_w$.
-After that, it applies a strong mutation operator to the newly generated individual.
-This is the invidual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
-The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the acceptance criterion for DILS.
-Equation \ref{eq1} displays the expression for this criterion in the case of a minimization problem.
-
-\begin{equation}
-m_{w,G} = \left\{ \begin{array}{lc}
-m_t &   \text{if} \;\; f(m_t) < f(m_w)\\
-m_{w,G} &  \text{otherwise}
-\end{array}
-\right..
-\label{eq1}
-\end{equation}
-
-Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solution space.
-To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory $\{m_b, m_w\}$, Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.
-
-\begin{equation}
-m_{w,G+1} = \left\{ \begin{array}{lc}
-\text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
-m_{w,G} &  \text{otherwise}
-\end{array}
-\right.,
-\label{eq2}
-\end{equation}
-
-\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
-This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via recombination operator) as the best parent.
-Figure \ref{img:DILS_diag} summarizes the DILS optimization process.
-
-\begin{figure}[!h]
-	\centering
-	\includegraphics[scale=0.25]{Figures/DILSDiagram.png}
-	\caption{Diagram summarizing DILS optimization process.}\label{img:DILS_diag}
-\end{figure}
-
-It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reasigned to the true best and worst individual respectively.
-This way, at the point where the reinitialization criterion is tested, it is possible that $f(m_w)$ provides a better value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
-
-This leads us to note the influence of the parameter $\xi$.
-In one hand, if we are dealing with a minimization problem and we set $\xi > 0$, then the true worst individual is only reinitializatized when the trial individual $m_t$ replaced $m_w$ and is also better than $m_b$ by a margin.
-On the other hand, and also for the minimization case, but this time with $\xi < 0$, we allow DILS to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
-It comes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
-Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$, and we recommend $[-0.5, 0.5]$.
-To summarize, the closer $\xi$ is to 1 the more restrictive the reinitialization criterion is---in the minimization case.
-It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages.
-The reason being is that the fitness value of the best individual $f(m_b)$ never deteriorate ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it become harder for $m_b$ to get better o close to it.
-Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.
-
-\begin{algorithm}
-	\SetNlSty{textbf}{[}{]}
-	\SetNlSkip{0.5em}
-	\setstretch{1.2}
-	\SetKwFunction{RandInit}{RandInit}
-	\SetKwFunction{Best}{Best}
-	\SetKwFunction{Worst}{Worst}
-	\SetKwFunction{Recombination}{Recombination}
-	\SetKwFunction{Mutation}{Mutation}
-	\SetKwFunction{LocalSearch}{LocalSearch}
-	\KwIn{Probability for recombination operator $p_r$, segment size for mutation operator $p_s$, reinitialization method tolerance $\xi$.}
-	\tcp{Initialization phase}
-	$G \leftarrow 0$\\
-	$m_b \leftarrow$ \RandInit{};
-	$m_w \leftarrow$ \RandInit{}\\
-	\tcp{Main loop}
-	\While{Termination criteria are not met}{
-		\tcp{Find best and worst individual}
-		$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $};
-		$m_w \leftarrow $ \Worst{$ \{m_b, m_w\} $}\\
-		\tcp{Generate new individual}
-		$m_t \leftarrow $ \Recombination{$m_b, \; m_w$}\\
-		$m_t \leftarrow $ \Mutation{$m_t$}\\
-		\tcp{Improve new individual}
-		$m_t \leftarrow $ \LocalSearch{$m_t$}\\
-		\tcp{Apply replacement operator}
-		\If{$f(m_t) < f(m_w)$}{
-
-			$m_w \leftarrow m_t$\\
-
-		}
-		\tcp{Check restart criterion}
-		\If{$f(m_b) - f(m_w) > f(m_b) \times \xi$}{
-
-			$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
-			$m_w \leftarrow$ \RandInit{}\\
-
-		}
-
-	$G \leftarrow G + 1$
-}
-\caption{DILS}\label{alg:DILS}
-\end{algorithm}
-
-For the recombination and mutation operators that appear in Algorithm \ref{alg:DILS}, we use the uniform recombination and the segment mutation operator respectively.
-The uniform recombination consists of selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}.
-The segment mutation operator consists of replacing a fixed size segment from the features of the individual for randomly generated features CITA.
-The size of the segment $p_s$ must be fixed at the beginning of the optimization process.
-
-It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems.
-In addition, the LS procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation matters and termination criteria.
-
-\section{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}
-
-In this section, we present an optimization scheme for the DILS algorithm when applied to constrained clustering.
-We will call this new approach DILS$_{CC}$.
-To begin with, we will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---being $n$ the number of instances of the dataset.
-Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, so each individual $m$ is a solution to the clustering problem.
-Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to intially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.
-
-Regarding the fitness function $f$, for each individual $m$ its fitness value $f_m$ can be calculated as in Equation \ref{eq3}.
-
-\begin{equation}
-f_m = z_m * (\text{infeasibility}_m + 1),
-\label{eq3}
-\end{equation}
-
-\noindent where $\text{infeasibility}_m$ is the number of non-satisfied constraints and $z_m$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq4}.
-
-\begin{equation}
-z_m = \sum_{c_i \in C_m} \left[ \frac{\sum_{x_a, x_b \in c_i} d^2(x_a,x_b)}{|c_i|}\right].
-\label{eq4}
-\end{equation}
-
-We also need to specify an LS procedure for DILS to be able to improve the mutant individual to obtain the trial vector $m_t$.
-Algorithm \ref{alg:LS} shows the LS used in DILS$_{CC}$.
-Note that, in order to not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
-We also stablish the maximum number of neighbors $p_m$ that can be generated in each call, this is done to find a better diversification-intensification tradeoff.
-
-\begin{algorithm}
-	\SetNlSty{textbf}{[}{]}
-	\SetNlSkip{0.5em}
-	\setstretch{1.2}
-	\SetKwFunction{RandomShuffle}{RandomShuffle}
-	\SetKwFunction{Rand}{Rand}
-	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $p_m$.}
-	\BlankLine
-	\While{$improvement$ {\bf and} $generated < p_m$}{
-		$improvement \leftarrow$ \texttt{false} \\
-		\tcp{Select random index (object) from $m$}
-		$i \leftarrow $ \Rand{$1$,$n$} \\
-		\tcp{Random shuffle labels set}
-		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
-		\For{$l \in RSL$}{
-			$m^\prime \leftarrow m$\\
-			\tcc{Move object $i$ from $m$ to the cluster associated with label $l$}
-			$m^\prime_i \leftarrow l$\\
-			\If{$f(m^\prime) < f(m)$}{
-				$m \leftarrow m^\prime$\\
-				$improvement \leftarrow$ \texttt{true} \\
-			}
-			$generated \leftarrow generated + 1$\\
-		}
-	}
-	\BlankLine
-	\KwRet ($m$)
-
-	\caption{Local Search}\label{alg:LS}
-\end{algorithm}
-
-\clearpage
-
-\section{Experimental Setup} \label{sec:expSetup}
-
-For our experiments we will compare the results obtained by DILS$_{CC}$ and state-of-the-art methods over 25 datasets and 3 constraint sets for each one of them.
+For our experiments we will compare the results obtained by \acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$ with those obtained by state-of-the-art methods and \acs{BRKGA}$_{CC}$ over 20 datasets and 3 constraint sets for each one of them.
 Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
@@
 		Hayesroth & 160 & 3 & 4 \\
-		Heart & 270 & 2 & 13 \\
 		Ionosphere & 351 & 2 & 33 \\
 		Iris & 150 & 3 & 4 \\
-		Led7Digit & 500 & 10 & 7 \\
 		Monk2 & 432 & 2 & 6 \\
 		Moons & 300 & 2 & 2 \\
-		Movement Libras & 360 & 15 & 90 \\
-		Newthyroid & 215 & 3 & 5 \\
 		Saheart & 462 & 2 & 9 \\
@@
 		Vowel & 990 & 11 & 13 \\
-		Wine & 178 & 3 & 13 \\
 		Zoo & 101 & 7 & 16 \\
@@
 
-\subsection{Constraint Generation} \label{sec:ConstGent}
+\section{Constraint Generation} \label{sec:ConstGent}
 
@@
 	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
+	%\setlength{\tabcolsep}{7pt}
+	%\renewcommand{\arraystretch}{1}
 	%\begin{adjustwidth}{-1in}{-1in}
@@
 			Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
-		Heart & 178 & 173 && 396 & 424 && 744 & 687 \\
 			Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
 			Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
-		Led7Digit & 126 & 1099 && 267 & 2508 && 460 & 4490 \\
 			Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
 			Moons & 200 & 235 && 494 & 496 && 900 & 870 \\
-		Movement Libras & 27 & 603 && 112 & 1319 && 158 & 2398 \\
-		Newthyroid & 108 & 123 && 270 & 258 && 449 & 454 \\
+
 			Saheart & 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
@@
 			Vowel & 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
-		Wine & 49 & 104 && 121 & 230 && 217 & 413 \\
 			Zoo & 21 & 34 && 29 & 91 && 41 & 169 \\
@@
 
-\clearpage
-
 \subsection{Evaluation Method} \label{sec:EvalMet}
@@
 Since we have the true labels associated to each of the datasets, we can use them to evaluate the results provided by each method.
-We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
+We will use the \acf{ARI} to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
 The basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
@@
 
-The ARI is a corrected-for-chance version of the Rand Index.
+The \acs{ARI} is a corrected-for-chance version of the Rand Index.
 This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline.
-The ARI is computed as in Equation \eqref{eq16}.
+The \acs{ARI} is computed as in Equation \eqref{eq16}.
 
@@
 \noindent where Maximum Index is expected to be 1 and Expected Index is the already mentioned expected degree of similarity with a random model.
-It is easy to see that $\text{ARI}(C_1, C_2) \in [-1,1]$, such that an ARI value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $\text{Rand}(C_1, C_2)$ is less than expected when comparing with random partitions.
-To summarize, the higher the ARI, the greater the degree of similarity between $C_1$ and $C_2$.
-For more details on ARI see \cite{hubert1985comparing}.
+It is easy to see that $\text{ARI}(C_1, C_2) \in [-1,1]$, such that an \acs{ARI} value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $\text{Rand}(C_1, C_2)$ is less than expected when comparing with random partitions.
+To summarize, the higher the \acs{ARI}, the greater the degree of similarity between $C_1$ and $C_2$.
+For more details on \acs{ARI} see \cite{hubert1985comparing}.
 
 Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper.
-To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth labels.
+To accomplish this task we just set one of the two partitions given to compute \acs{ARI} as the ground truth labels.
 
-\subsection{Validation of results} \label{sec:ValidtnMethod}
+\section{Validation of results} \label{sec:ValidtnMethod}
 
@@
 In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$.
-Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
-$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$
+Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form: \\
 
-\subsection{Calibration}
+\noindent\resizebox{\textwidth}{!}{$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$}
 
-Table \ref{tab:params} shows a summary of the parameter setup used for the DILS$_{CC}$.
+\section{Calibration}
+
+Table \ref{tab:paramsSHADE} shows a summary of the parameter setup used for both \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$ algorithms.
+Since both of these methods are nature-inspired population-based optimization schemes it is worth going in depth in they performance similarities and differences.
+All parameters have been set with the aid to achieve a fair comparison, which will be later presented in Chapter \ref{ch:ExperimentalResults}.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.4}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
+			\hline
+			Parameter & Meaning & \acs{BRKGA}$_{CC}$ & \acs{SHADE}$_{CC}$ \\
+			\hline
+			$|P|$ & Population size & 100 & 100 \\
+			Evals & Fitness function evaluations & 300000 & 300000 \\
+			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
+			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
+			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
+			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
+			\hline
+
+	\end{tabular}}
+	%\end{adjustwidth}
+
+	\caption[Parameters setup used for BRKGA$_{CC}$ and SHADE$_{CC}$]{Parameters setup used for \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.}
+	\label{tab:paramsSHADE}
+\end{table}
+
+Table \ref{tab:paramsDILS} shows a summary of the parameter setup used for the \acs{DILS}$_{CC}$ method.
 
@@
 
-	\caption{Parameters setup used for DILS$_{CC}$.}
-	\label{tab:params}
+	\caption[Parameters setup used for DILS$_{CC}$]{Parameters setup used for \acs{DILS}$_{CC}$.}
+	\label{tab:paramsDILS}
 \end{table}
 
-The stop criterion is given by the number of evaluations of the fitness function $f$, which at most will be 300000.
-The implementation for DILS$_{CC}$ can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
-
-To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
-The implementation for the BRKGA algorith can be found in the same link as DILS$_{CC}$.
-In all cases the value for $k$ is equal to the number of classes displayed in Table \ref{tab:datasets}.
-
-\textcolor{red}{Meter todos los algoritmos en el mismo repo para no tener que poner tantos enlaces a github? Poner los enlaces como notas a pie de pagina tambien?}
+In all three cases---\acs{BRKGA}$_{CC}$, \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$---the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:SOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA}.
 
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabular}{cl}
+		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} c}
 			\hline
-			Method name & Parameters name and values \\
+			Common Parameters & Meaning & Value \\
 			\hline
-			BRKGA & Evals = $300000$; $|P| = 100$; $P_e = P_m = 0.2 * |P|$; $p_{\text{inherit}} = 50\%$ \\
-			COPKM & \texttt{max\_iter} = $300$; \texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''} \\
-			CECM & \texttt{max\_iter} = $300$; $\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''} \\
-			LCVQE & \texttt{max\_iter} = $300$; \texttt{initial\_centroids} = $\emptyset$ \\
-			RDPM & \texttt{max\_iter} = $300$; $\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset. \\
-			TVClust & \texttt{max\_iter} = $300$; $\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$ \\
+			\texttt{max\_iter} & Maximum number of iterations & 300 \\
+			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
+			\hline
+			\hline
+			Specific Parameters & \multicolumn{2}{l}{Name and Value} \\
+			\hline
+			\acs{COPKM} & \multicolumn{2}{l}{\texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''}} \\
+			\acs{CECM} & \multicolumn{2}{l}{$\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''}} \\
+			\acs{LCVQE} & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
+			\acs{RDPM} & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset} \\
+			\acs{TVClust} & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
 			\hline
@@
 Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
-
-\section{Experimental Results} \label{sec:results}
-
-In this section we present Tables from \ref{tab:results10} to \ref{tab:results20}, which display the results obtained by DILS$_{CC}$ and six state-of-the-art methods to be compared for each dataset and constraint set.
-
-Since some of the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
-To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
-
-It should be noted that there are some missing results in these tables.
-In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
-It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
-In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
-			Breast Cancer & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
-			Circles & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
-			Ecoli & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
-			Glass & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
-			Haberman & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
-			Hayesroth & 0.031 & 0.027 & - & \textbf{0.106} & 0.097 & 0.054 & 0.139 \\
-			Heart & \textbf{0.793} & 0.483 & - & 0.027 & 0.036 & 0.423 & -0.003 \\
-			Ionosphere & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
-			Iris & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
-			Led7Digit & 0.016 & 0.009 & \textbf{0.497} & 0.425 & 0.369 & 0.316 & - \\
-			Monk2 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
-			Moons & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
-			Movement Libras & 0.022 & 0.004 & 0.285 & \textbf{0.293} & 0.256 & 0.000 & - \\
-			Newthyroid & 0.028 & 0.017 & - & 0.579 & 0.289 & \textbf{0.846} & 0.002 \\
-			Saheart & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
-			Sonar & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
-			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
-			Spectfheart & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
-			Spiral & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
-			Tae & \textbf{0.031} & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
-			Vowel & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
-			Wine & 0.334 & 0.120 & - & 0.360 & \textbf{0.368} & 0.281 & 0.004 \\
-			Zoo & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
-			\hline
-			Mean & \textbf{0.474} & 0.158 & -0.510 & 0.226 & 0.212 & 0.175 & -0.186 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{10}$ by DILS$_{CC}$ and the state-of-the-art methods.}
-	\label{tab:results10}
-\end{table}
-
-Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-We can note that DILS$_{CC}$ represents a consistent improvement over all other 5 methods.
-This is due to the fact that the intensification capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lower level of constraint-based information we work with.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Breast Cancer & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
-			Bupa & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
-			Circles & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Ecoli & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
-			Glass & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
-			Haberman & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
-			Hayesroth & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
-			Heart & \textbf{1.000} & 0.692 & \textbf{1.000} & 0.025 & 0.033 & 0.442 & 0.000 \\
-			Ionosphere & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
-			Iris & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
-			Led7Digit & 0.020 & 0.004 & - & 0.425 & \textbf{0.490} & 0.324 & - \\
-			Monk2 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
-			Moons & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
-			Movement Libras & 0.025 & 0.006 & -0.742 & \textbf{0.293} & 0.255 & 0.000 & - \\
-			Newthyroid & 0.440 & 0.320 & - & 0.568 & 0.356 & \textbf{0.923} & 0.052 \\
-			Saheart & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
-			Sonar & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
-			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
-			Spectfheart & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
-			Spiral & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
-			Tae & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
-			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
-			Wine & \textbf{0.725} & 0.528 & - & 0.360 & 0.367 & 0.348 & -0.001 \\
-			Zoo & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
-			\hline
-			Mean & \textbf{0.641} & 0.490 & -0.070 & 0.225 & 0.233 & 0.258 & -0.196 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{15}$ by DILS$_{CC}$ and the state-of-the-art methods.}
-	\label{tab:results15}
-\end{table}
-
-Results obtained with the $SC_{15}$ constraint set are presented in Table \ref{tab:results15}.
-We observe how DILS$_{CC}$ and COPKM are able to obtain ARI values equal to 1, meaning that the resulting partitions of the datasets perfectly match the true partitions.
-The fact that DILS$_{CC}$ is always able to output a partition of the datasets is worth the advantage over COPKM.
-It is even the case that COPKM is not able to provide an output partition while DILS finds the optimal partition or very close to the optimal partition for the same datasets.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Breast Cancer & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
-			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
-			Glass & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
-			Haberman & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
-			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Heart & \textbf{1.000} & 0.942 & \textbf{1.000} & 0.025 & 0.033 & 1.000 & 0.000 \\
-			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
-			Iris & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
-			Led7Digit & 0.017 & 0.007 & - & 0.425 & \textbf{0.571} & 0.327 & - \\
-			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
-			Movement Libras & 0.018 & 0.006 & - & \textbf{0.293} & \textbf{0.293} & 0.000 & - \\
-			Newthyroid & 0.842 & 0.848 & - & 0.568 & 0.393 & \textbf{0.933} & 0.004 \\
-			Saheart & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
-			Sonar & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
-			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
-			Spectfheart & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
-			Spiral & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
-			Tae & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
-			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
-			Wine & 0.916 & \textbf{0.937} & - & 0.360 & 0.363 & 0.380 & 0.015 \\
-			Zoo & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
-			\hline
-			Mean & \textbf{0.733} & 0.642 & 0.064 & 0.223 & 0.285 & 0.344 & -0.187 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{20}$ by DILS$_{CC}$ and the state-of-the-art methods.}
-	\label{tab:results20}
-\end{table}
-
-In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how DILS$_{CC}$ scales the quality of its results with the number of restrictions, as well as methods such as BRKGA or COPKM, the latter being more limited in other aspects.
-It is also woth noting how DILS$_{CC}$ provides the better average result for each of the three constraint set analyzed in this paper.
-
-\clearpage
-
-\section{Statistical Analysis of Results} \label{sec:analisis}
-
-With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
-This way we can statistically determine whether DILS$_{CC}$ represents a significant improvement over previous proposals.
-
-Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with DILS$_{CC}$ as sample $B$.
-
-One of the major advantages of the Bayesian sign test is that we can obtain a very illustrative visual representation of its results.
-We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
-With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle.
-In order to find out where a certain triplet will be placed within the triangle, we will take each of its three values and draw a parallel line to the opposing side of the corresponding vertex; the separation between a triangle side and its parallel line will be proportional to the associated triplet value, so that the higher the value, the closer the line will be to the vertex.
-The location where the three lines intersect is where we draw a point.
-Since the values of every triplet describe a probability distribution and therefore they must add up to one, we can be sure that all triplets will lie in some point within the triangle.
-The color indicates the density of points in a given region, with yellow representing a high density and red a low density.
-
-\begin{figure}[ht!]
-	\centering
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/DILSvsBRKGA.pdf}
-		\caption{BRKGA}
-		\label{fig:DILSvsBRKGA}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/DILSvsCOPKM.pdf}
-		\caption{COPKM}
-		\label{fig:DILSvsCOPKM}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/DILSvsLCVQE.pdf}
-		\caption{LCVQE}
-		\label{fig:DILSvsLCVQE}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/DILSvsTVClust.pdf}
-		\caption{TVClust}
-		\label{fig:DILSvsTVClust}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/DILSvsRDPM.pdf}
-		\caption{RDPM}
-		\label{fig:DILSvsRDPM}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/DILSvsCEKM.pdf}
-		\caption{CEKM}
-		\label{fig:DILSvsCEKM}
-	\end{subfigure}
-	\caption{Heat diagrams comparing DILS with the state-of-the-art methods.}
-	\label{fig:heatDiag}
-\end{figure}
-
-Figure \ref{fig:heatDiag} shows all six heatmaps obtained when applying the Bayesian sign test to DILS$_{CC}$ and the state-of-the-art methods.
-All heatmaps suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle.
-Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strog evidence that DILS$_{CC}$ does represent a significan improvement over all other six methods, being te results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.
-
-\clearpage
-
-\section{Conclusions} \label{sec:conclusiones}
-
-In this paper we proposed the DILS heuristic method, along with its application to the SSL constrained clustering problem, which we called DILS$_{CC}$.
-Focusing on the instance-level ML and CL constraints, DILS$_{CC}$ has proven that heuristic techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
-
-Supported by the Bayesian statistical tests, we were able to objectively prove that the DILS$_{CC}$ approach is significantly better than the state-of-the-art methods.
-DILS$_{CC}$ has proven to be better when considering the quality and availability of the solutions, especially in cases where large sets of constraints are analyzed.
-
-\section*{Acknowledgements}
-
-Our work has been supported by the research project TIN2017-89517-P and PP2016.PRI.I.02.
-
-\section*{Compliance with ethical standards}
-
-\section*{Declaration of competing interest}
-
-The authors declare that there is no conflict of interest.
-
-\clearpage
-
-\section*{References}
-
-\bibliography{mybibfile}
-
-\end{document}
+All parameters described above are ment to be used for the implementation that can be found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}.

### 2019-06-26 Memoria/Chapters/NewProposals.tex
Nuevos capitulos en memoria y correcciones en papers

--- 2019-06-12 Papers/SHADE/paperSHADE.tex
@@
-\documentclass[review]{elsarticle}
-
-\usepackage{graphics, float, url}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{lineno,hyperref}
-\usepackage{mathtools}
-\usepackage{multirow}
-\usepackage{adjustbox}
-\usepackage{chngpage}
-\usepackage{setspace}
-\usepackage{amsfonts}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage[ruled, lined, onelanguage, linesnumbered]{algorithm2e}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-\newcommand{\myfloatalign}{\centering}
-\modulolinenumbers[5]
-
-\usepackage{epstopdf}
-\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
-\AppendGraphicsExtensions{.tiff}
-
-\journal{Applied Soft Computing}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-\newtheorem{definition}{Definition}
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Enhancing instance-level constrained clustering through differential evolution}
-
-\author[mymainaddress]{Germ\'an Gonz\'alez-Almagro\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{germangalmagro@ugr.es}
-
-\author[mymainaddress]{Juli\'an Luengo}
-
-\author[mysecondaddress]{Jos\'e-Ram\'on Cano}
-
-\author[mymainaddress]{Salvador Garc\'ia}
-
-\address[mymainaddress]{DaSCI Andalusian Institute of Data Science and Computational Intelligence, University of Granada, Spain}
-
-\address[mysecondaddress]{Dept. of Computer Science, EPS of Linares, University of Ja\'en, Campus Cient\'ifico Tecnol\'ogico de Linares, Cintur\'on Sur S/N, Linares 23700, Ja\'en, Spain}
-
-\begin{abstract}
-Clustering has always been a powerful tool in knowledge discovery.
-Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering.
-This technique is a generalization of traditional clustering that considers additional information encoded by constraints.
-Constraints can be given in the form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose the first application of Differential Evolution to the constrained clustering problem, which has proven to produce a better exploration-exploitation trade-off when comparing with previous approaches.
-We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
-\end{abstract}
-
-\begin{keyword}
-constrained clustering, instance-level, must-link, cannot-link, differential evolution.
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
-
-\section{Introduction} \label{sec:Intro}
-
-One of the most widely known and studied data analysis problems is clustering.
-It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
-Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics,\cite{Everitt:2009:CA:1538772} cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
-
-In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
-In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
-A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
-In this paper we will focus on partitional clustering.
-
-We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them.
-A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features.
-More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$.
-A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$.
-As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$.
-The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$ \cite{jain1999data}.
-
-Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
-We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}.
-The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
-The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{triguero2015self}.
-
-Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
-It has been successfully applied in many knowledge fields, among which it is worth mentioning: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, and lane finding in GPS data \cite{wagstaff2001constrained} among others.
-
-Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level \cite{bradley2000constrained}, instance-level \cite{davidson2007survey} and feature-level constrained clustering \cite{schmidt2011clustering}.
-Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.
-
-In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints.
-In one hand, pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
-On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure \cite{davidson2007survey}.
-This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
-
-Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained} and soft \cite{law2004clustering} constraints.
-Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}.
-For the purposes of this paper, we will employ the latter.
+\chapter{New Proposals}\label{ch:NewProposals}
 
 Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem.
-Therefore, the incorporation of constraints may modify the complexity of the clustering problem, depending on the type of constraints used.
-As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete \cite{davidson2005clustering}.
+As we have already mentioned in Section \ref{sec:FeasibilityProblem}, the incorporation of constraints may modify the complexity of the clustering problem, making it $\mathbf{NP}$-complete and therefore intractable \cite{davidson2005clustering}.
+Bearing this in mind, in this chapter we propose two novel approaches to find approximate solutions to the constrained clustering problem.
+In Section \ref{sec:CCSHADE} a \acf{DE} based approach is presented, while in Section \ref{sec:DILS} a new \acf{LS} variant is discussed, as well as its application to constrained clustering.
+
+\section{Constrained Clustering Through Differential Evolution} \label{sec:CCSHADE}
 
@@
 Nature-inspired algorithms are designed to emulate natural optimization phenomena, such as evolution, collective behavior of animals, physics laws or even human being-related processes.
-There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the Biased Random-key Genetic Algorithm (BRKGA) presented in \cite{de2017comparison}.
+There have been attempts to solve the constrained clustering problem with nature-inspired algorithms, such as the adaptation of the \acf{BRKGA} presented in \cite{de2017comparison}.
 Swarm-based methods have also been applied to constrained clustering, such as the one presented in \cite{xu2013improving}.
 
-Differential Evolution (DE) is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}.
-In this paper we propose the first DE application to find quality solutions for the constrained clustering problem.
-In particular, we will take as basis for our proposal one of the DE variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
-We will make use of the Random-key concept from BRKGA, along with proposing a new fitness function, to build the already mentioned DE variant.
-We will properly compare the results obtained by our new method with previous nature-inspired algorithms, as well as with the constrained clustering state-of-the-art.
-
-Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and DE.
-In Section \ref{sec:brkga} the BRKGA algorithm and its adaptation for constrained clustering will be briefly reviewed.
-Afterward, the SHADE variant of DE algorithm will be presented in Section \ref{sec:SHADE}.
-In Section \ref{sec:SHADEadapt} the scheme that allow us to apply SHADE to constrained clustering, including the proposal of a new fitness function, will be presented.
-Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
-Finally, in Section \ref{sec:conclusiones} conclusions are discussed.
-
-\section{Background} \label{sec:background}
-
-In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering (Section \ref{sec:BackSOTA}).
-We will also present the basis of the DE optimization method (Section \ref{sec:BackDE}).
-
-\subsection{Constrained Clustering} \label{sec:BackCC}
-
-In most clustering applications it is common to have some kind of information about the dataset to be analyzed.
-In pairwise instance-level constrained clustering this information is given in the form of pairs of instances.
-A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster.
-It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms.
-We can now formalize the two type of constraints mentioned:
-
-\begin{itemize}
-
-	\item Must-link constraints $C_=(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.
-
-	\item Cannot-link constraints $C_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.
-
-\end{itemize}
-
-The goal of constrained clustering is to find a partition (or clustering) of $k$ clusters $C = \{c_1, \cdots, c_k\}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set.
-As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $n = |X| = \sum_{i = 1}^{k} |c_i|$.
-
-Knowing how a constraint is defined, ML constraints are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric.
-This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified.
-In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
-
-It can also be proven that CL constraints do not constitute an equivalence relation.
-However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $C_{\neq}(x_a,x_b)$, then it is also true that $C_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.
-
-\subsection{The Feasibility Problem} \label{sec:BackFeas}
-
-Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem.
-The feasibility problem for non-hierarchical instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
-
-\begin{definition}
-
-	\textbf{Feasibility Problem}: given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied?
-\cite{davidson2005clustering}
-	\label{def1}
-
-\end{definition}
-
-In \cite{davidson2005clustering} it is proven that, when $k_l = 1$ and $k_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem (it is also proven that it is not harder, so both have the same complexity).
-Table \ref{tab:feasibility} shows the complexity of the feasibility for different types of constraints.
-
-\begin{table}[!h]
-	\centering
-	%\setlength{\arrayrulewidth}{1mm}
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\resizebox{\textwidth}{!}{
-		\begin{tabular}{c c}
-			\hline
-			Constraints & Complexity \\
-			\hline
-			Must-Link & $\mathbf{P}$\\
-			Cannot-Link & $\mathbf{NP}$-complete\\
-			ML and CL & $\mathbf{NP}$-complete\\
-			\hline
-
-		\end{tabular}%}
-	\caption{Feasibility problem complexity \cite{davidson2005clustering}.}
-	\label{tab:feasibility}
-\end{table}
-
-These complexity results show that the feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too.
-For more details on the complexity of constrained clustering see \cite{davidson2005clustering}.
-
-Intractable problems are hard to solve with deterministic and exact methods.
-That is the reason why nature-inspired algorithms constitute good approaches to find quality solutions to the constrained clustering problem.
-
-\subsection{State-of-the-art Methods} \label{sec:BackSOTA}
-
-The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}.
-It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
-This method was named COP-kmeans, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
-
-In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
-The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
-This method redefines constraints from the point of view of belief functions and includes them in the cost function.
-
-A modification of the Constrained Vector Quantization Error algorithm (CVQE \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}.
-The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity.
-Linear CVQE (LCVQE) introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex.
-The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.
-
-Two Views Clustering (TVClust) and Relation Dirichlet Process - Means (RDPM) were proposed in \cite{khashabi2015clustering}.
-TVClust is able to incorporate the constraints into the clustering problem by making a soft interpretation of them.
-The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations.
-Using this model as a basis, the authors derive the deterministic algorithm RDP-means.
-This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($k$) does not need to be specified.
-
-\subsection{Differential Evolution} \label{sec:BackDE}
+\acs{DE} is an evolution-based algorithm that has proven to be excellent in real-domain problem solving \cite{das2011differential}.
+In this section we propose the first \acs{DE} application to find quality solutions for the constrained clustering problem.
+In particular, we will take as basis for our proposal one of the \acs{DE} variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
+We will make use of the Random-key concept from \acs{BRKGA}, along with proposing a new fitness function, to build the already mentioned \acs{DE} application.
+
+\subsection{A Glimpse to Differential Evolution}
 
@@
 
-DE, which emerged in \cite{storn1997differential}, is a good example of evolution-based methods.
-From that point on, the reputation of DE was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
-In particular, DE excelled in real-coded optimization problems.
-A highly detailed analysis of DE and its variants can be found in \cite{das2011differential}.
-
-DE arose as a variant of population-based evolutionary algorithms, its goal being to perform an intelligent search in the solution space of the problem in the most optimized and effective way possible.
-
-Like most evolutionary algorithms, DE uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,v]}\}$, with $v$ being the number of features of each individual.
+\acf{DE}, which emerged in \cite{storn1997differential}, is a good example of evolution-based methods.
+From that point on, the reputation of \acs{DE} was consolidated in conferences and competitions where it was able to obtain competitive results that rivaled those achieved by the state-of-the-art.
+In particular, \acs{DE} excelled in real-coded optimization problems.
+A highly detailed analysis of \acs{DE} and its variants can be found in \cite{das2011differential}.
+
+\acs{DE} arose as a variant of population-based evolutionary algorithms, its goal being to perform an intelligent search in the solution space of the problem in the most optimized and effective way possible.
+
+Like most evolutionary algorithms, \acs{DE} uses a population of individuals $P$, where each individual $p_i$ is a vector of real values $p_i = \{p_{[i,1]},\cdots,p_{[i,v]}\}$, with $v$ being the number of features of each individual.
 Each $p_i$ serves as parameter for the function to be optimized.
@@
 To guide the optimization process, we use a fitness function.
-The task of DE is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^{v} \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
+The task of \acs{DE} is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^{v} \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
 
@@
 	\centering
-	\includegraphics[scale=0.3]{Figures/DEop.png}
-	\caption{The optimization process of DE \cite{das2011differential}.}\label{img:DE}
+	\includegraphics[scale=0.3]{gfx/NewProp/SHADE/DEop.png}
+	\caption[The optimization process of DE.]{The optimization process of \acs{DE} \cite{das2011differential}.}\label{fig:DE}
 \end{figure}
 
-The optimization process of DE the works through a simple cycle of stages, presented in Figure \ref{img:DE}, that we can summarize as follows:
+The optimization process of \acs{DE} the works through a simple cycle of stages, presented in Figure \ref{fig:DE}, that we can summarize as follows:
 
@@
 
-	\item \textbf{Initialization of Vectors}: DE begins with a randomly initialized population of a given number of $v$-dimensional arrays of parameters.
+	\item \textbf{Initialization of Vectors}: \acs{DE} begins with a randomly initialized population of a given number of $v$-dimensional arrays of parameters.
 These vectors constitute candidate solutions for the problem to be solved.
 
-	\item \textbf{Difference-vector-based mutation}: DE introduces diversity into the population via the mutation operator.
+	\item \textbf{Difference-vector-based mutation}: \acs{DE} introduces diversity into the population via the mutation operator.
 This operator is applied to each individual of the population, known in this context as the parent vector.
@@
 
-Even though DE was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
-We will focus on the SHADE variant, which adaptively optimizes some of the parameters of DE.
+Even though \acs{DE} was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
+We will focus on the SHADE variant, which adaptively optimizes some of the parameters of \acs{DE}.
 To carry out this task SHADE uses a historic record of values for these parameters that allows it to set them in favor of the optimization process, based on how good the algorithm behaved with certain parameter values in the past (see Section \ref{sec:SHADE}).
@@
 
-Genetic Algorithms (GAs) emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
+\acp{GA} emerged as one of the first metaheuristics inspired by the concept of natural selection and evolution.
 It is one of the most studied and successful evolutionary algorithms, thanks in part to its simplicity and ease of implementation \cite{fausto2019ants, goldberg1989genetic}.
 
-In a GA, a population $P$ of a given number of solutions $p_i = \{p_{[i,1]}, \cdots, p_{[i,v]}\}$ is first initialized; these solutions are also called chromosomes or individuals.
+In a \acs{GA}, a population $P$ of a given number of solutions \sloppy $p_i = \{p_{[i,1]}, \cdots, p_{[i,v]}\}$ is first initialized; these solutions are also called chromosomes or individuals.
 Each element of a chromosome is called gene and the value it takes is known as allele.
-To simulate the natural processes of evolution and natural selection, the set of evolutionary operators that GAs make use of is similar to that used by DE: crossover, mutation and selection.
+To simulate the natural processes of evolution and natural selection, the set of evolutionary operators that \acsp{GA} make use of is similar to that used by \acs{DE}: crossover, mutation and selection.
 The crossover operator consists in combining two different chromosomes (parents) from the population and obtaining a new chromosome (offspring) which inherits alleles for its genes from both parents.
-The mutation operator is used by GAs to avoid local optima, and it consists in introducing random changes into the population to favor diversity.
+The mutation operator is used by \acsp{GA} to avoid local optima, and it consists in introducing random changes into the population to favor diversity.
 The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function that assigns a value to each chromosome to assess its quality as a solution for the problem.
-GAs apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal \cite{fausto2019ants}.
-
-The Random-Key Genetic Algorithm (RKGA) emerged in \cite{bean1994genetic} as a solid approach to high constrained problems.
+\acsp{GA} apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal \cite{fausto2019ants}.
+
+\acf{RKGA} emerged in \cite{bean1994genetic} as a solid approach to high constrained problems.
 It uses vectors of random numbers to represent solutions to the problem (individuals), which are used as keys to obtain the actual solutions.
@@
 A decoder is a deterministic procedure that allows us to map random-key vectors to the literal solution space.
-We can build a decoder that always provides as a result a feasible solution to the problem we want to solve, thus avoiding the feasibility problem that we found in classic GAs.
-
-The Biased Random-key Genetic Algorithm (BRKGA) was first proposed in \cite{gonccalves2011biased} as a variant of the RKGA.
-In the BRKGA optimization process, first, the population $P$ of $|P|$ vectors of random-keys is initialized; this will be the initial generation.
+We can build a decoder that always provides as a result a feasible solution to the problem we want to solve, thus avoiding the feasibility problem that we found in classic \acsp{GA}.
+
+The \acf{BRKGA} was first proposed in \cite{gonccalves2011biased} as a variant of the \acs{RKGA}.
+In the \acs{BRKGA} optimization process, first, the population $P$ of $|P|$ vectors of random-keys is initialized; this will be the initial generation.
 Each random-key vector $p_i$ has $v$ random-keys in it.
@@
 
-It should be noted that in BRKGA there is a clear differentiation between problem-dependent and problem-independent elements.
+It should be noted that in \acs{BRKGA} there is a clear differentiation between problem-dependent and problem-independent elements.
 The independent part of the algorithm has no knowledge of the problem it solves, it is limited to performing the operations related to the population optimization process.
 The dependent part of the problem is composed of the decoder and the fitness function, which allow us to obtain actual solutions for the problem and its evaluation.
-Therefore, to specify a BRKGA optimization scheme it is only necessary to define the decoder and the fitness function \cite{gonccalves2011biased}.
+Therefore, to specify a \acs{BRKGA} optimization scheme it is only necessary to define the decoder and the fitness function \cite{gonccalves2011biased}.
 
@@
 
-An adaptation of BRKGA for constrained clustering was proposed in \cite{de2017comparison}, we will refer to it as BRKGA$_{CC}$.
+An adaptation of \acs{BRKGA} for constrained clustering was proposed in \cite{de2017comparison}, we will refer to it as \acs{BRKGA}$_{CC}$.
 
@@
 	\end{tabular}}
-	\caption{Random-key decodification example \cite{de2017comparison}.}
+	\caption[Random-key decodification example.]{Random-key decodification example \cite{de2017comparison}.}
 	\label{tab:decodingrk}
@@
 
-In \cite{de2017comparison}, the authors also added a new element to BRKGA, a local search procedure.
+In \cite{de2017comparison}, the authors also added a new element to \acs{BRKGA}, an \acs{LS} procedure.
 This local optimization method is applied to each decoded individual in each generation, but its results are not transferred to the individual, so that diversity is maintained.
-Thus, the individuals produced by the local search are only used to update the best solution found so far if needed.
-For more details on BRKGA$_{CC}$ see \cite{de2017comparison}.
+Thus, the individuals produced by the \acs{LS} are only used to update the best solution found so far if needed.
+Figure \ref{fig:BRKGA_CC} shows a visual representation of the overall BRKGA optimization process for constrained clustering.
+For more details on \acs{BRKGA}$_{CC}$ see \cite{de2017comparison}.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.4]{gfx/NewProp/SHADE/BRKGA_Diagram.jpg} 
+	\caption[BRKGA$_{CC}$ optimization process for constrained clustering.]{\acs{BRKGA}$_{CC}$ optimization process for constrained clustering. \cite{de2017comparison}.}\label{fig:BRKGA_CC}
+\end{figure}
 
@@
 
-Success History Based Adaptive Differential Evolution (SHADE) is an optimization method based on differential evolution.
-Proposed in \cite{tanabe2013success}, it was an improvement for the JADE model, presented in \cite{zhang2009jade}.
-
-Unlike JADE, which only considers the parameters used in the previous generation, SHADE uses a historical record of successful parameters as a mechanism for adapting parameters involved in the process of creating new generations.
-
-Since SHADE is an evolution-based algorithm, it shares elements with BRKGA (Section \ref{sec:brkga}), so we will use a similar notation to define its elements.
+\acf{SHADE} is an optimization method based on \acs{DE}. Proposed in \cite{tanabe2013success}, it was an improvement for the JADE model, presented in \cite{zhang2009jade}.
+
+Unlike JADE, which only considers the parameters used in the previous generation, \acs{SHADE} uses a historical record of successful parameters as a mechanism for adapting parameters involved in the process of creating new generations.
+
+Since \acs{SHADE} is an evolution-based algorithm, it shares elements with \acs{BRKGA} (Section \ref{sec:brkga}), so we will use a similar notation to define its elements.
 We will note the population as $P_G$, formed by $|P|$ individuals with the form $p_{[i,G]} = \{p_{[i,G,1]}, \cdots, p_{[i,G,v]}\}$, with $G$ indicating the generation to which $p_i$ belongs.
@@
 
-As in differential evolution, SHADE employs mutant generation, crossover and selection operators to explore the space of solutions and bring in new generations of individuals.
-
-The mutation strategy used by SHADE is known as current-to-$p$best/1, and the expression that generates new individuals is described in Equation \eqref{eq3}.
+As in \acs{DE}, \acs{SHADE} employs mutant generation, crossover and selection operators to explore the space of solutions and bring in new generations of individuals.
+
+The mutation strategy used by \acs{SHADE} is known as current-to-$p$best/1, and the expression that generates new individuals is described in Equation \eqref{eq3}.
 
@@
 After generating the mutant vector $m_{[i,G]}$, it is combined with the parent $p_{[i,G]}$ vector by means of a crossover operator; the result is the trial vector $t_{[i,G]} = \{t_{[i,G,1]}, \cdots, t_{[i,G,v]}\}$.
-SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.
+\acs{SHADE} uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.
 
@@
 
-The SHADE selection operator determines the survivors individuals for the next generation.
+The \acs{SHADE} selection operator determines the survivors individuals for the next generation.
 It compares each parent $x_{[i,G]}$ with the trial vector $t_{[i,G,j]}$, generated based on it.
@@
 
-To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{[i,G]}$ who were replaced by their associated trial vector $t_{[i,G]}$ are saved.
+To maintain diversity, \acs{SHADE} can make use of an external archive $A$, in which parents $p_{[i,G]}$ who were replaced by their associated trial vector $t_{[i,G]}$ are saved.
 To make use of the archive, we will consider that the individual $p_{[r2,G]}$ from Equation \eqref{eq3} is selected from $P \cup A$.
@@
 The parameters $p\text{best}$, $F_i$ and $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process.
-These are the parameters that SHADE adaptively optimizes using the mentioned history record.
+These are the parameters that \acs{SHADE} adaptively optimizes using the mentioned history record.
 
@@
 
-The SHADE method stores in memory a structure $M$ with $|M|$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:SHADEmemory}.
+The \acs{SHADE} method stores in memory a structure $M$ with $|M|$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:SHADEmemory}.
 This structure is initialized with the value $0.5$ for all entries.
@@
 	\end{tabular}}
-	\caption{Historical memory $M_{CR}$, $M_{F}$ used by SHADE \cite{tanabe2013success}.}
+	\caption[Historical memory $M_{CR}$, $M_{F}$ used by SHADE.]{Historical memory $M_{CR}$, $M_{F}$ used by \acs{SHADE} \cite{tanabe2013success}.}
 	\label{tab:SHADEmemory}
@@
 
-SHADE keeps two auxiliary sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent.
+\acs{SHADE} keeps two auxiliary sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent.
 At the end of each generation, the content of the memory $M$ is updated following Equations \eqref{eq8} and \eqref{eq9}.
@@
 
-In this section, we present an optimization scheme for the SHADE algorithm when applied to constrained clustering.
-We will call this new approach SHADE$_{CC}$.
+In this section, we present an optimization scheme for the \acs{SHADE} algorithm when applied to constrained clustering.
+We will call this new approach \acs{SHADE}$_{CC}$.
 To begin with, we will use random-keys to create the individuals of the population; this way, each individual is defined as a vector of $n$ random-keys ($v = n$).
@@
 To make these concepts clear we consider a toy dataset and three partitions over it.
-The dataset from Figure \ref{img:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
+The dataset from Figure \ref{fig:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
 Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($\text{infeasibility} = 0$), whereas $C_2$ does violate it ($\text{infeasibility} = 1$).
@@
 	\myfloatalign
-	{\includegraphics[width=.45\linewidth]{Figures/Dataset}} \quad
+	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/Dataset}} \quad
 	%\subfloat[Imagen original]
-	{\includegraphics[width=.45\linewidth]{Figures/C0}} \quad
+	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C0}} \quad
 	%\subfloat[Clustering sin restricciones]
-	{\includegraphics[width=.45\linewidth]{Figures/C1}} \quad
+	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C1}} \quad
 	%\subfloat[Clustering con restricciones]
-	{\includegraphics[width=.45\linewidth]{Figures/C2}} \quad
-	\caption{Three partitions over a toy dataset.}
-	\label{img:toydatasets}
+	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C2}} \quad
+	\caption[Three partitions over a toy dataset.]{Three partitions over a toy dataset.}
+	\label{fig:toydatasets}
 \end{figure}
 
-Table \ref{tab:fitnessfunctions} shows the fitness values for each partition from Figure \ref{img:toydatasets}.
+Table \ref{tab:fitnessfunctions} shows the fitness values for each partition from Figure \ref{fig:toydatasets}.
 We refer to the function presented in Equation \eqref{eq1} as $f_1$ and to the function presented in Equation \eqref{eq14} as $f_2$.
@@
 
-	\caption{Expression and value of fitness functions over three partitions ($\mu = 10$).}
+	\caption[Expression and value of fitness functions over three partitions ($\mu = 10$).]{Expression and value of fitness functions over three partitions ($\mu = 10$).}
 	\label{tab:fitnessfunctions}
@@
 The percentage of individuals considered as elite must be determined for each case.
-
-Algorithm \ref{alg:SHADE} describes the overall SHADE$_{CC}$ optimization process, including the local search procedure presented in Algorithm \ref{alg:LS}.
+Figure \ref{fig:SHADE_CC} shows a representation of the \acs{SHADE}$_{CC}$ optimization process.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.4]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
+	\caption[SHADE$_{CC}$ optimization process for constrained clustering.]{\acs{SHADE}$_{CC}$ optimization process for constrained clustering.}\label{fig:SHADE_CC}
+\end{figure}
+
+Algorithm \ref{alg:SHADE} describes the overall \acs{SHADE}$_{CC}$ optimization process, including the \acs{LS} procedure presented in Algorithm \ref{alg:SHADE_LS}.
 The goal of this procedure is to locally improve solutions (individuals from $P$) in a non-exhaustive way.
@@
 			$p_{[i,G]} \leftarrow$ rand$[N/2, 0.2]$\\
-			\tcp{Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin}
+			Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin\\
 
@@
 	}
-	\caption{SHADE$_{CC}$}\label{alg:SHADE}
+	\caption[SHADE$_{CC}$]{\acs{SHADE}$_{CC}$}\label{alg:SHADE}
 \end{algorithm}
@@
 	\BlankLine
-	\KwRet ($S$)
-
-\caption{Local Search}\label{alg:LS}
+	\KwRet{$S$}
+
+	\caption{Local Search}\label{alg:SHADE_LS}
 \end{algorithm}
 
-\clearpage
-
-\section{Experimental Setup} \label{sec:expSetup}
-
-For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraint sets for each one of them.
-Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
-\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/SHADE_CC}{GitHub}.
-Table \ref{tab:datasets} displays a summary of every dataset.
-
-\begin{table}[!h]
-	\centering
-	%\setlength{\arrayrulewidth}{1mm}
-	%\setlength{\tabcolsep}{5pt}
-	%\renewcommand{\arraystretch}{1.2}
-	%\resizebox{\textwidth}{!}{
-	\small
-	\begin{tabular}{l c c c}
-		\hline
-		Name & No. Instances & No. Classes & No. Features \\
-		\hline
-		Appendicitis & 106 & 2 & 7 \\
-		Breast Cancer & 569 & 2 & 30 \\
-		Bupa & 345 & 2 & 6 \\
-		Circles & 300 & 2 & 2 \\
-		Ecoli & 336 & 8 & 7 \\
-		Glass & 214 & 6 & 9 \\
-		Haberman & 306 & 2 & 3 \\
-		Hayesroth & 160 & 3 & 4 \\
-		Ionosphere & 351 & 2 & 33 \\
-		Iris & 150 & 3 & 4 \\
-		Led7Digit & 500 & 10 & 7 \\
-		Monk2 & 432 & 2 & 6 \\
-		Moons & 300 & 2 & 2 \\
-		Movement Libras & 360 & 15 & 90 \\
-		Pima & 768 & 2 & 8 \\
-		Saheart & 462 & 2 & 9 \\
-		Sonar & 208 & 2 & 60 \\
-		Soybean & 47 & 4 & 35 \\
-		Spectfheart & 267 & 2 & 44 \\
-		Spiral & 300 & 2 & 2 \\
-		Tae & 151 & 3 & 5 \\
-		Vehicle & 846 & 4 & 18 \\
-		Vowel & 990 & 11 & 13 \\
-		Wdbc & 569 & 2 & 30 \\
-		Zoo & 101 & 7 & 16 \\
-		\hline
-
-	\end{tabular}%}
-	\caption{Summary of datasets used for the experiments.}
-	\label{tab:datasets}
-\end{table}
-
-Classification datasets are commonly used in the literature to test constrained clustering algorithms; they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
-They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).
-
-\subsection{Constraint Generation} \label{sec:ConstGent}
-
-Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets.
-This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
-
-We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with three small percentages of the size of the dataset: 10\%, 15\% and 20\%.
-With $n_f$ being the fraction of the size of the dataset associated with each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
-
-The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
-Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-	\begin{tabular}{lcc c cc c cc}
-		\hline
-		\multirow{2}{*}{Dataset} &
-		\multicolumn{2}{c}{$CS_{10}$} && \multicolumn{2}{c}{$CS_{15}$} && \multicolumn{2}{c}{$CS_{20}$} \\
-		\cline{2-3} \cline{5-6} \cline{8-9}
-		& ML & CL && ML & CL && ML & CL \\
-		\hline
-		Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
-		Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
-		Bupa & 323 & 272 && 699 & 627 && 1201 & 1145 \\
-		Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
-		Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
-		Glass & 52 & 179 && 139 & 389 && 259 & 644 \\
-		Haberman & 304 & 161 && 634 & 401 && 1135 & 756 \\
-		Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
-		Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
-		Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
-		Led7Digit & 126 & 1099 && 267 & 2508 && 460 & 4490 \\
-		Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
-		Moons & 200 & 235 && 494 & 496 && 900 & 870 \\
-		Movement Libras & 27 & 603 && 112 & 1319 && 158 & 2398 \\
-		Pima & 1604 & 1322 && 3595 & 3075 && 6452 & 5329 \\
-		Saheart & 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
-		Sonar & 100 & 110 && 245 & 251 && 436 & 425 \\
-		Soybean & 4 & 6 && 6 & 22 && 12 & 33 \\
-		Spectfheart & 233 & 118 && 543 & 277 && 965 & 466 \\
-		Spiral & 224 & 211 && 487 & 503 && 918 & 852 \\
-		Tae & 40 & 80 && 82 & 171 && 151 & 314 \\
-		Vehicle & 874 & 2696 && 1955 & 6046 && 3589 & 10776 \\
-		Vowel & 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
-		Wdbc & 840 & 756 && 1925 & 1730 && 3472 & 2969 \\
-		Zoo & 21 & 34 && 29 & 91 && 41 & 169 \\
-		\hline
-
-	\end{tabular}}
-	%\end{adjustwidth}
-
-	\caption{Number of constraints used in experiments.}
-	\label{tab:constraints}
-\end{table}
-
-Note that the greater the number of classes in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
-This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.
-
-\clearpage
-
-\subsection{Evaluation Method} \label{sec:EvalMet}
-
-Since we have the true labels associated to each of the datasets, we can use them to evaluate the results provided by each method.
-We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
-The basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$.
-$C_1$ and $C_2$ are viewed as collections of $n(n - 1)/2$ pairwise decisions \cite{rand1971objective}.
-
-For each pair of instances $x_i$ and $x_j$ in $X$, a partition assigns them to the same cluster or to different clusters.
-We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in both $C_1$ and $C_2$, and $b$ as the opposite event ($x_i$ and $x_j$ are in different clusters in $C_1$ and $C_2$).
-Then, the degree of similarity between $C_1$ and $C_2$ is calculated as in Equation \eqref{eq15}.
-
-\begin{equation}
-\text{Rand}(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
-\label{eq15}
-\end{equation}
-
-The ARI is a corrected-for-chance version of the Rand Index.
-This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline.
-The ARI is computed as in Equation \eqref{eq16}.
-
-\begin{equation}
-\text{ARI}(C_1, C_2) = \frac{\text{Rand}(C_1, C_2) - \text{Expected Index}}{\text{Maximum Index} - \text{Expected Index}},
-\label{eq16}
-\end{equation}
-
-\noindent where Maximum Index is expected to be 1 and Expected Index is the already mentioned expected degree of similarity with a random model.
-It is easy to see that $\text{ARI}(C_1, C_2) \in [-1,1]$, such that an ARI value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $\text{Rand}(C_1, C_2)$ is less than expected when comparing with random partitions.
-To summarize, the higher the ARI, the greater the degree of similarity between $C_1$ and $C_2$.
-For more details on ARI see \cite{hubert1985comparing}.
-
-Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper.
-To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth labels.
-
-\subsection{Validation of results} \label{sec:ValidtnMethod}
-
-In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic Null Hypothesis Statistical Tests (NHST).
-In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
-\textit{"In a nutshell: NHST do not answer the question we ask"}.
-To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
-
-To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
-Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
-This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.
-
-Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate.
-As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
-
-Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
-That is: What can we conclude when NHST do not reject the null hypothesis?
-We can not infer anything since NHST can not provide evidence in its favor.
-
-Finally, there are two other problems that researchers face when performing NHST.
-The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results.
-The second one is the need to previously formalize the intentions of the sampling of the results, which are usually fixed a posteriori; this could lead to a misreading of said results.
-
-As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST.
-In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test.
-To make use of it we will employ the R package \texttt{rNPBST}, whose documentation and guide can be found in \cite{carrasco2017rnpbst}.
-
-We will use the Bayesian sign test to validate our results, which is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution.
-To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$.
-In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$.
-Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form:
-$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$
-
-\subsection{Calibration}
-
-Table \ref{tab:params} shows a summary of the parameter setup used for both BRKGA$_{CC}$ and SHADE$_{CC}$ algorithms.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
-			\hline
-			Parameter & Meaning & BRKGA$_{CC}$ & SHADE$_{CC}$ \\
-			\hline
-			$|P|$ & Population size & 100 & 100 \\
-			Evals & Fitness function evaluations & 300000 & 300000 \\
-			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
-			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
-			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
-			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Parameters setup used for BRKGA$_{CC}$ and SHADE$_{CC}$.}
-	\label{tab:params}
-\end{table}
-
-In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
-A Python implementation for both algorithms can be found at \href{https://github.com/GermangUgr/SHADE\_CC}{GitHub}
-
-To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} c}
-			\hline
-			Common Parameters & Meaning & Value \\
-			\hline
-			\texttt{max\_iter} & Maximum number of iterations & 300 \\
-			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
-			\hline
-			\hline
-			Specific Parameters & \multicolumn{2}{l}{Name and Value} \\
-			\hline
-			COPKM & \multicolumn{2}{l}{\texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''}} \\
-			CECM & \multicolumn{2}{l}{$\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''}} \\
-			LCVQE & \multicolumn{2}{l}{\texttt{initial\_centroids} = $\emptyset$} \\
-			RDPM & \multicolumn{2}{l}{$\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset} \\
-			TVClust & \multicolumn{2}{l}{$\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$} \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Parameters setup used for the state-of-the-art algorithms.}
-	\label{tab:paramsSOTA}
-\end{table}
-
-Parameter values have been assigned following the guidelines of the original creators of the different proposals.
-Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible.
-Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible.
-Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
-
-\section{Experimental Results} \label{sec:results}
-
-In this section we present Tables from \ref{tab:results10} to \ref{tab:results20SOTA}, which display the results obtained by the methods to be compared for each dataset and constraint set.
-
-Since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
-To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
-
-\subsection{Comparing BRKGA$_{CC}$ with SHADE$_{CC}$}
-
-In Tables from \ref{tab:results10} to \ref{tab:results20} the ARI columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{l ccc c ccc}
-			\hline
-			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
-			\cline{2-4} \cline{6-8}
-			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
-			\hline
-			Appendicitis & 0.022 & 5.454 & 383.642 && \textbf{0.086} & 5.091 & 366.491 \\
-			Breast Cancer & \textbf{0.395} & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
-			Bupa & \textbf{0.281} & 14.151 & 1420.117 && 0.104 & 16.672 & 1409.487 \\
-			Circles & \textbf{0.166} & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
-			Ecoli & 0.009 & 20.321 & 1169.798 && \textbf{0.020} & 12.763 & 1112.981 \\
-			Glass & \textbf{0.016} & 8.658 & 735.507 && 0.013 & 4.848 & 694.962 \\
-			Haberman & 0.047 & 15.699 & 1155.985 && \textbf{0.084} & 15.441 & 1146.023 \\
-			Hayesroth & \textbf{0.027} & 1.500 & 550.117 && 0.025 & 2.000 & 529.424 \\
-			Ionosphere & 0.125 & 17.587 & 1894.390 && \textbf{0.332} & 12.603 & 1872.308 \\
-			Iris & 0.238 & 0.190 & 505.112 && \textbf{0.342} & 0.000 & 483.584 \\
-			Led7Digit & \textbf{0.009} & 11.347 & 1786.943 && 0.004 & 8.866 & 1784.708 \\
-			Monk2 & \textbf{0.421} & 14.926 & 1848.543 && 0.270 & 16.385 & 1821.208 \\
-			Moons & \textbf{0.277} & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
-			Movement Libras & \textbf{0.004} & 4.159 & 1508.279 && 0.003 & 3.396 & 1519.185 \\
-			Pima & 0.236 & 25.195 & 4072.087 && \textbf{0.275} & 22.468 & 3990.383 \\
-			Saheart & 0.147 & 20.555 & 2135.624 && \textbf{0.265} & 16.725 & 2095.554 \\
-			Sonar & \textbf{0.124} & 7.619 & 1106.039 && 0.116 & 8.952 & 1081.994 \\
-			Soybean & 0.496 & 0.000 & 176.333 && \textbf{0.515} & 0.000 & 151.163 \\
-			Spectfheart & 0.334 & 9.573 & 1451.235 && \textbf{0.343} & 8.604 & 1405.243 \\
-			Spiral & 0.058 & 15.402 & 1098.622 && \textbf{0.114} & 13.793 & 1099.747 \\
-			Tae & 0.018 & 0.833 & 520.855 && \textbf{0.024} & 1.333 & 508.988 \\
-			Vehicle & 0.005 & 28.515 & 4366.104 && \textbf{0.006} & 22.151 & 4352.663 \\
-			Vowel & \textbf{0.003} & 13.330 & 4245.672 && 0.002 & 11.523 & 4290.154 \\
-			Wdbc & \textbf{0.478} & 16.316 & 3596.517 && 0.357 & 17.093 & 3528.029 \\
-			Zoo & \textbf{0.105} & 1.091 & 336.345 && 0.098 & 2.182 & 317.674 \\
-			\hline
-			Mean & \textbf{0.161} & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
-	\label{tab:results10}
-\end{table}
-
-Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-We can note that SHADE$_{CC}$ represents a consistent improvement over BRKGA$_{CC}$ in both Unsat and Time, while in terms of ARI both methods are similar.
-This is because the penalty applied to the fitness function of both methods has little presence when operating with a small constraint set, and since they are population-based and heuristic-guided methods, they have similar exploration and exploitation capabilities.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{l ccc c ccc}
-			\hline
-			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
-			\cline{2-4} \cline{6-8}
-			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
-			\hline
-			Appendicitis & 0.296 & 6.167 & 381.625 && \textbf{0.332} & 5.833 & 362.106 \\
-			Breast Cancer & 0.772 & 10.172 & 3854.911 && \textbf{0.911} & 3.310 & 3766.402 \\
-			Bupa & \textbf{0.840} & 5.973 & 1463.152 && 0.801 & 5.641 & 1437.639 \\
-			Circles & 0.828 & 6.101 & 1136.484 && \textbf{0.982} & 0.404 & 1108.928 \\
-			Ecoli & 0.015 & 24.314 & 1200.463 && \textbf{0.025} & 19.090 & 1127.568 \\
-			Glass & 0.017 & 17.500 & 745.987 && \textbf{0.031} & 11.743 & 722.009 \\
-			Haberman & \textbf{0.866} & 4.677 & 1209.471 && 0.796 & 5.314 & 1147.479 \\
-			Hayesroth & \textbf{0.107} & 7.536 & 545.861 && 0.099 & 8.261 & 520.702 \\
-			Ionosphere & \textbf{0.803} & 7.722 & 1986.543 && 0.591 & 11.974 & 1944.085 \\
-			Iris & \textbf{0.412} & 3.873 & 506.548 && 0.260 & 6.087 & 486.723 \\
-			Led7Digit & 0.004 & 12.836 & 1902.272 && \textbf{0.007} & 10.919 & 1897.511 \\
-			Monk2 & 0.751 & 10.827 & 1949.305 && \textbf{0.945} & 2.067 & 1920.361 \\
-			Moons & 0.958 & 1.334 & 1133.858 && \textbf{0.995} & 0.020 & 1113.687 \\
-			Movement Libras & \textbf{0.006} & 8.679 & 1584.028 && 0.001 & 8.889 & 1590.321 \\
-			Pima & 0.481 & 22.051 & 4451.193 && \textbf{0.762} & 9.928 & 4402.693 \\
-			Saheart & 0.749 & 10.940 & 2281.992 && \textbf{0.945} & 1.839 & 2225.210 \\
-			Sonar & 0.800 & 5.040 & 1124.700 && \textbf{0.890} & 2.177 & 1069.862 \\
-			Soybean & \textbf{0.428} & 0.000 & 170.674 && 0.422 & 0.000 & 155.948 \\
-			Spectfheart & \textbf{0.924} & 2.854 & 1561.040 && 0.778 & 5.732 & 1485.260 \\
-			Spiral & 0.847 & 5.677 & 1133.962 && \textbf{0.949} & 1.495 & 1102.920 \\
-			Tae & \textbf{0.119} & 7.115 & 522.088 && 0.066 & 8.380 & 513.289 \\
-			Vehicle & \textbf{0.089} & 28.044 & 4806.559 && 0.018 & 26.856 & 4749.812 \\
-			Vowel & \textbf{0.002} & 14.486 & 4807.165 && 0.001 & 13.341 & 4867.849 \\
-			Wdbc & 0.805 & 8.985 & 3829.407 && \textbf{0.930} & 2.801 & 3750.711 \\
-			Zoo & \textbf{0.157} & 3.500 & 333.984 && 0.128 & 2.834 & 315.131 \\
-			\hline
-			Mean & 0.483 & 9.456 & 1784.931 && \textbf{0.507} & 6.997 & 1751.368 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
-	\label{tab:results15}
-\end{table}
-
-Table \ref{tab:results15} presents the results obtained with the $CS_{15}$ constraint set.
-While conclusions for Unsat and Time remain unchanged, we see a clear difference as far as ARI is concerned.
-We observe that both methods already report ARI results above 0.7, which in an ARI ranking is considered to be an acceptable result.
-We can even observe that the SHADE$_{CC}$ method provides an ARI higher than 0.9 for 7 datasets, compared to BRKGA$_{CC}$ that only obtains similar results in 2 datasets.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{l ccc c ccc}
-			\hline
-			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
-			\cline{2-4} \cline{6-8}
-			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
-			\hline
-			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
-			Bupa & 0.936 & 2.975 & 1529.151 && \textbf{0.993} & 0.281 & 1508.772 \\
-			Circles & 0.935 & 2.814 & 1180.268 && \textbf{1.000} & 0.000 & 1170.452 \\
-			Ecoli & 0.028 & 25.716 & 1267.419 && \textbf{0.047} & 21.466 & 1277.903 \\
-			Glass & 0.038 & 23.610 & 766.861 && \textbf{0.082} & 16.412 & 796.162 \\
-			Haberman & 0.943 & 2.073 & 1266.131 && \textbf{0.997} & 0.138 & 1235.441 \\
-			Hayesroth & \textbf{0.548} & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
-			Ionosphere & 0.902 & 4.314 & 2051.251 && \textbf{0.993} & 0.258 & 2019.558 \\
-			Iris & \textbf{0.691} & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
-			Led7Digit & \textbf{0.007} & 13.653 & 2053.264 && 0.005 & 12.170 & 2066.223 \\
-			Monk2 & 0.855 & 6.025 & 2055.425 && \textbf{0.991} & 0.379 & 2013.468 \\
-			Moons & 0.937 & 2.689 & 1177.848 && \textbf{0.995} & 0.158 & 1157.222 \\
-			Movement Libras & \textbf{0.006} & 8.381 & 1654.159 && 0.004 & 7.676 & 1645.174 \\
-			Pima & 0.753 & 11.544 & 4855.446 && \textbf{0.899} & 4.419 & 4820.684 \\
-			Saheart & 0.881 & 5.250 & 2402.513 && \textbf{0.981} & 0.706 & 2353.931 \\
-			Sonar & 0.992 & 0.186 & 1138.769 && \textbf{1.000} & 0.000 & 1127.020 \\
-			Soybean & 0.416 & 0.000 & 171.281 && \textbf{0.563} & 0.000 & 153.406 \\
-			Spectfheart & 0.953 & 1.677 & 1602.229 && \textbf{1.000} & 0.000 & 1526.925 \\
-			Spiral & 0.921 & 3.604 & 1178.580 && \textbf{0.998} & 0.091 & 1155.194 \\
-			Tae & 0.351 & 9.807 & 527.292 && \textbf{0.392} & 9.419 & 508.610 \\
-			Vehicle & \textbf{0.144} & 28.781 & 5189.066 && 0.044 & 29.008 & 5284.288 \\
-			Vowel & 0.002 & 14.682 & 5450.331 && \textbf{0.003} & 13.766 & 5514.052 \\
-			Wdbc & 0.823 & 8.281 & 4019.683 && \textbf{0.972} & 1.068 & 3983.564 \\
-			Zoo & \textbf{0.176} & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
-			\hline
-			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
-			\hline
-
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
-	\label{tab:results20}
-\end{table}
-
-It is in Table \ref{tab:results20}, which gathers results obtained for the constraint set $CS_{20}$, that we observe major differences between the two methods.
-We find that SHADE$_{CC}$ is capable of producing clusters identical to the original ones for 4 datasets, regardless the single dataset for which BRKGA$_{CC}$ achieves it.
-Let us remember that an ARI of 1 represents a perfect match between the two partitions given to compute it.
-This improvement in ARI is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, SHADE$_{CC}$ outperforms BRKGA$_{CC}$.
-Conclusions on execution time remain the same.
-
-\subsection{Comparing SHADE$_{CC}$ with the state-of-the-art} \label{sec:SHADEvsSOTA}
-
-In this section we present a brief comparison between the proposed SHADE$_{CC}$ method and some of the state-of-the-art methods.
-Tables from \ref{tab:results10SOTA} to \ref{tab:results20SOTA} show the results obtained by these methods.
-To shorten the presented results this time we will only compare the ARI.
-
-It should be noted that there are some missing results in these tables.
-In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
-It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
-In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lcccccc}
-			\hline
-			Dataset & SHADE$_{CC}$ & COPKM & LCVQE & RDPM & TVClust & CECM \\
-			\hline
-			Appendicitis & 0.086 & - & \textbf{0.335} & 0.316 & 0.025 & -0.005 \\
-			Breast Cancer & 0.159 & -0.604 & 0.486 & \textbf{0.502} & 0.000 & 0.000 \\
-			Bupa & \textbf{0.104} & - & -0.005 & -0.005 & -0.004 & -0.011 \\
-			Circles & 0.063 & - & -0.003 & \textbf{0.162} & 0.137 & 0.133 \\
-			Ecoli & 0.020 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
-			Glass & 0.013 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
-			Haberman & 0.084 & - & -0.002 & 0.127 & \textbf{0.218} & -0.004 \\
-			Hayesroth & 0.025 & - & 0.106 & 0.097 & 0.054 & \textbf{0.139} \\
-			Ionosphere & \textbf{0.332} & - & 0.168 & 0.197 & 0.000 & 0.030 \\
-			Iris & 0.342 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
-			Led7Digit & 0.004 & \textbf{0.497} & 0.425 & 0.369 & 0.316 & - \\
-			Monk2 & 0.270 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
-			Moons & 0.272 & - & 0.241 & 0.319 & \textbf{0.785} & 0.092 \\
-			Movement Libras & 0.003 & 0.285 & \textbf{0.293} & 0.256 & 0.000 & - \\
-			Pima & \textbf{0.275} & - & 0.076 & 0.075 & 0.145 & -0.008 \\
-			Saheart & 0.265 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
-			Sonar & \textbf{0.116} & - & 0.004 & 0.013 & 0.000 & 0.000 \\
-			Soybean & 0.515 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
-			Spectfheart & \textbf{0.343} & - & -0.107 & -0.114 & 0.000 & 0.050 \\
-			Spiral & \textbf{0.114} & - & -0.003 & 0.012 & 0.034 & -0.002 \\
-			Tae & 0.024 & - & 0.009 & -0.000 & \textbf{0.067} & 0.000 \\
-			Vehicle & 0.006 & - & 0.121 & 0.081 & \textbf{0.280} & - \\
-			Vowel & 0.002 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
-			Wdbc & 0.357 & - & 0.486 & \textbf{0.502} & 0.000 & 0.000 \\
-			Zoo & 0.098 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
-			\hline
-			Mean  & 0.156 & -0.51 & \textbf{0.215} & 0.211 & 0.13 & -0.226 \\
-			\hline
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
-	\label{tab:results10SOTA}
-\end{table}
-
-Table \ref{tab:results10SOTA} presents the results obtained for the $CS_{10}$ constraint set.
-LCVQE, RDPM and TVClust methods stand out, achieving results as competitive as SHADE$_{CC}$ with a low number of constraints.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lcccccc}
-			\hline
-			Dataset & SHADE$_{CC}$ & COPKM & LCVQE & RDPM & TVClust & CECM \\
-			\hline
-			Appendicitis & \textbf{0.332} & - & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Breast Cancer & 0.911 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
-			Bupa & 0.801 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
-			Circles & 0.982 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Ecoli & 0.025 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
-			Glass & 0.031 & - & 0.280 & 0.253 & 0.229 & - \\
-			Haberman & 0.796 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
-			Hayesroth & 0.099 & - & 0.106 & 0.107 & \textbf{0.150} & 0.135 \\
-			Ionosphere & 0.591 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
-			Iris & 0.260 & - & \textbf{0.730} & 0.547 & 0.421 & 0.684 \\
-			Led7Digit & 0.007 & - & 0.425 & \textbf{0.490} & 0.324 & - \\
-			Monk2 & 0.945 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
-			Moons & 0.995 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
-			Movement Libras & 0.001 & -0.742 & \textbf{0.293} & 0.255 & 0.000 & - \\
-			Pima & \textbf{0.762} & - & 0.076 & 0.075 & 1.000 & 0.000 \\
-			Saheart & 0.945 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
-			Sonar & \textbf{0.890} & - & 0.004 & 0.019 & 0.000 & 0.000 \\
-			Soybean & 0.422 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
-			Spectfheart & 0.778 & \textbf{0.983} & -0.107 & -0.117 & 0.000 & -0.070 \\
-			Spiral & \textbf{0.949} & - & -0.003 & 0.014 & 0.006 & 0.051 \\
-			Tae & \textbf{0.066} & - & 0.008 & -0.004 & 0.062 & 0.000 \\
-			Vehicle & 0.018 & \textbf{1.000} & 0.121 & 0.081 & 0.576 & - \\
-			Vowel & 0.001 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
-			Wdbc & 0.930 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Zoo & 0.128 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
-			\hline
-			Mean  & \textbf{0.507} & 0.01 & 0.213 & 0.229 & 0.253 & -0.238 \\
-			\hline
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
-	\label{tab:results15SOTA}
-\end{table}
-
-In one hand, Table \ref{tab:results15SOTA}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by SHADE$_{CC}$.
-On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by SHADE$_{CC}$, with the exception of COPKM, which is able to reach the optimum for some of the cases in which it finds a solution.
-
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.2}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lcccccc}
-			\hline
-			Dataset & SHADE$_{CC}$ & COPKM & LCVQE & RDPM & TVClust & CECM \\
-			\hline
-			Appendicitis & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Breast Cancer & 0.977 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.993 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & \textbf{1.000} & \textbf{1.000} & -0.003 & 0.629 & \textbf{1.000} & 0.138 \\
-			Ecoli & 0.047 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
-			Glass & 0.082 & - & 0.271 & \textbf{0.287} & 0.218 & - \\
-			Haberman & 0.997 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
-			Hayesroth & \textbf{0.531} & - & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Ionosphere & 0.993 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
-			Iris & 0.553 & - & \textbf{0.708} & 0.540 & 0.585 & 0.684 \\
-			Led7Digit & 0.005 & - & 0.425 & \textbf{0.571} & 0.327 & - \\
-			Monk2 & 0.991 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & 0.995 & \textbf{1.000} & 0.241 & 0.831 & \textbf{1.000} & 0.180 \\
-			Movement Libras & 0.004 & - & \textbf{0.293} & \textbf{0.293} & 0.000 & - \\
-			Pima & 0.899 & \textbf{1.000} & 0.076 & 0.076 & \textbf{1.000} & -0.005 \\
-			Saheart & 0.981 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
-			Sonar & \textbf{1.000} & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
-			Soybean & 0.563 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
-			Spectfheart & \textbf{1.000} & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
-			Spiral & 0.998 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
-			Tae & \textbf{0.392} & - & 0.008 & 0.000 & 0.035 & 0.000 \\
-			Vehicle & 0.044 & \textbf{1.000} & 0.122 & 0.081 & 0.689 & - \\
-			Vowel & 0.003 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
-			Wdbc & 0.972 & \textbf{1.000} & 0.486 & 0.522 & 0.000 & 0.000 \\
-			Zoo & 0.153 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
-			\hline
-			Mean  & \textbf{0.647} & 0.224 & 0.212 & 0.28 & 0.319 & -0.228 \\
-			\hline
-		\end{tabular}}
-		%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and the state-of-the-art.}
-	\label{tab:results20SOTA}
-\end{table}
-
-Table \ref{tab:results20SOTA} presents the results obtained for $CS_{20}$.
-For SHADE$_{CC}$ we continue to observe its progression in accuracy.
-As far as the state-of-the-art methods are concerned, we find a general improvement in ARI, except for CECM.
-Even with this, it is SHADE$_{CC}$ and COPKM that seem to obtain the highest quality results, with the difference that SHADE$_{CC}$ is able to produce results in all cases.
-
-\clearpage
-
-\section{Statistical Analysis of Results} \label{sec:analisis}
-
-With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
-This way we can statistically determine whether SHADE$_{CC}$ represents a significant improvement over previous proposals.
-
-Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by BRKGA$_{CC}$ as sample $A$, and to the results obtained with SHADE$_{CC}$ as sample $B$.
-
-One of the major advantages of the Bayesian sign test is that we can obtain a very illustrative visual representation of its results.
-We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
-With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle.
-In order to find out where a certain triplet will be placed within the triangle, we will take each of its three values and draw a parallel line to the opposing side of the corresponding vertex; the separation between a triangle side and its parallel line will be proportional to the associated triplet value, so that the higher the value, the closer the line will be to the vertex.
-The location where the three lines intersect is where we draw a point.
-Since the values of every triplet describe a probability distribution and therefore they must add up to one, we can be sure that all triplets will lie in some point within the triangle.
-The color indicates the density of points in a given region, with yellow representing a high density and red a low density.
-
-\begin{figure}[ht!]
-	\centering
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/BayesSignoARI.pdf}
-		\caption{ARI}
-		\label{fig:bayesARI}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/BayesSignoTime.pdf}
-		\caption{Time}
-		\label{fig:bayesTime}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/BayesSignoUnsat.pdf}
-		\caption{Unsat}
-		\label{fig:bayesUnsat}
-	\end{subfigure}
-	\caption{Heat diagrams for the comparison with BRKGA$_{CC}$.}
+\section{Constrained Clustering Through Local Search} \label{sec:DILS}
+
+Following the stream of thought introduced at the beginning of this chapter, it is justified to think that metaheuristics constitute a promising approach to fin approximate quality solutions to the tough challenge that constrained clustering constitutes.
+Metaheuristic algorithms are designed to explore the solution space of a problem with the guidance of a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
+In this type of approaches the key is to find a good exploration-exploitation tradeoff.
+This is the reason why the classic \acf{LS} heuristic algorithm derived in the \acf{ILS} metaheuristic algorithm, which introduces periodic perturbations in the exploration of the solution space to escape local optima \cite{lourencco2010iterated}.
+\acs{ILS}---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
+\acs{ILS} is also very commonly used in the design of hybrid methods with good exploration-exploitation tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
+In this section we propose a new variant of the \acs{ILS} method, the \acs{DILS}, which introduces diversity in the exploration of the solution space in an adaptive and guided way.
+
+\subsection{Short Introduction to Iterated Local Search}
+
+Before defining the \acs{ILS}---from now on \acs{ILS}---,it is necessary to understand the \acs{LS} and the motivation that leads to designing methods with greater exploratory capacity.
+
+An \acs{LS} is a procedure that consists of performing an heuristics-guided iterative improvement.
+Taking as initial state a well defined solution to the problem, the \acs{LS} uses a neighbor generation operator to explore new solutions.
+When a solution better than the current one is found, it replaces the current one.
+In order to evaluate the solutions, the \acs{LS} uses a function that assigns a value to each one of them, this is known as a target function.
+***This value will be better the closer the solution that evaluates the optimal solution.***
+
+Both the neighbor generation operator and the target function used by the \acs{LS} must be defined for each problem.
+The target function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem.
+The neighbor generation operator takes as base a solution to the problem $s$, and generates a new solution $s^\prime$ by modifying some of its components in a local way.
+
+The Algorithm \ref{alg:BasicLS} presents a basic scheme for an \acs{LS} procedure.
+It is worth noting that the condition of line 4 must be adapted if the problem to be solved is one of maximization or minimization.
+In addition, the general scheme considers that the optimization process ends when the current solution is not improved, however the stop condition can be adapted to each problem.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{GenerateNeighbor}{GenerateNeighbor}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Initial solution $s$}
+	\BlankLine
+	\While{$improvement$}{
+		$improvement \leftarrow$ \texttt{false} \\
+
+		$s^\prime \leftarrow $ \GenerateNeighbor{$s$}
+
+		\If{$f(s^\prime) < f(s)$}{
+
+			$s \leftarrow s^\prime$\\
+			$improvement \leftarrow$ \texttt{true} \\
+
+		}	
+	}
+	\BlankLine
+	\KwRet ($s$)
+
+	\caption{Basic Local Search}\label{alg:BasicLS}
+\end{algorithm}
+
+With the \acs{LS} defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$.
+Figure \ref{img:LS} shows a pictorical representation of the above.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{gfx/NewProp/DILS/LS.png}
+	\caption[Local optimum achieved by an LS procedure]{Local optimum achieved by an \acs{LS} procedure}\label{img:LS}
 \end{figure}
 
-In one hand, figure \ref{fig:bayesARI} shows the heatmap associated with the ARI measurement.
-The diagram suggests that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle.
-It also indicates that in some cases there are no significant differences between $A$ and $B$, since we also find triplets in the rope area.
-Bearing this in mind, and considering that ARI is a measure to maximize, we can say that, for cases where there are significant differences, SHADE$_{CC}$ performs better than BRKGA$_{CC}$.
-On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} which show the heatmaps obtained for Time and Unsat measurements, are very similar.
-In both, all triplets are represented in the lower right third of the triangle, indicating that the Bayesian sign test assigns a high probability to $A - B > 0$.
-With this, and bearing in mind that both measures must be minimized, we conclude that SHADE$_{CC}$ represents a solid improvement over BRKGA$_{CC}$ as far as this measures are concerned.
-
-Regarding the comparison of SHADE$_{CC}$ with the state-of-the-art, we will keep the results obtained by SHADE$_{CC}$ as $B$, and note the results obtained by the state-of-the-art methods as $A$.
-
-This way, we apply the process described earlier in this section to obtain the heatmaps corresponding to the comparison of SHADE$_{CC}$ with each of the 5 state-of-the-art methods.
-This time the only measure we compare is ARI.
-Figures \ref{fig:bayesCOPKM} to \ref{fig:bayesCECM} show said comparison.
-
-\begin{figure}[ht!]
-	\centering
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriCOPKM.pdf}
-		\caption{SHADE$_{CC}$ vs COPKM}
-		\label{fig:bayesCOPKM}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriLCVQE.pdf}
-		\caption{SHADE$_{CC}$ vs LCVQE}
-		\label{fig:bayesLCVQE}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriRDPM.pdf}
-		\caption{SHADE$_{CC}$ vs RDPM}
-		\label{fig:bayesRDPM}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriTVClust.pdf}
-		\caption{SHADE$_{CC}$ vs TVClust}
-		\label{fig:bayesTVClust}
-	\end{subfigure}
-	\begin{subfigure}{.45\textwidth}
-		\includegraphics[width=1\linewidth]{Figures/AriCECM.pdf}
-		\caption{SHADE$_{CC}$ vs CECM}
-		\label{fig:bayesCECM}
-	\end{subfigure}
-	\caption{Heat diagrams for the comparison with the state-of-the-art.}
+It is therefore necessary to appeal to techniques that introduce diversity in the search to avoid the problem of local optimal. This is exactly what \acs{ILS} is trying to do.
+
+With \acs{ILS}, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^\prime$ that will not be a local optimum and will be far enough away from $s^*$.
+After that, we will apply an \acs{LS} procedure to find a new local optimal $s^{*\prime}$.
+Figure \ref{img:ILS} shows a representation of the above concepts.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{gfx/NewProp/DILS/ILS.png}
+	\caption[ILS perturbation over LS local optima]{\acs{ILS} perturbation over \acs{LS} local optima}\label{img:ILS}
 \end{figure}
 
-We see very similar diagrams for COPKM (\ref{fig:bayesCOPKM}), LVCQE (\ref{fig:bayesLCVQE}), RDPM (\ref{fig:bayesRDPM}) and TVClust (\ref{fig:bayesTVClust}).
-In all of them most of the triplets are represented in the lower middle region of the diagram, with a slight shift of the point cloud to the left.
-That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---, but the advantage is for one or the other depending on the case, being SHADE$_{CC}$ the one that gets it in most of them.
-However, this does not apply in the case of CECM (\ref{fig:bayesCECM}), where the advantage is indisputably for SHADE$_{CC}$.
-
-\section{Conclusions} \label{sec:conclusiones}
-
-In this paper we proposed SHADE$_{CC}$, the first application of a DE-based algorithm to the SSL constrained clustering problem.
-Focusing on the instance-level ML and CL constraints, SHADE$_{CC}$ has proven that nature-inspired evolution-based techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
-
-Supported by the Bayesian statistical tests, we were able to objectively prove that the SHADE$_{CC}$ approach is significantly better than previous evolution-based proposals.
-As far as the state-of-the-art is concerned, the SHADE$_{CC}$ approach has proven to be equivalent or better when considering the quality of the solutions, especially in cases where large sets of constraints are analyzed.
-
-\section*{Acknowledgements}
-
-Our work has been supported by the research project TIN2017-89517-P and PP2016.PRI.I.02.
-
-\section*{Compliance with ethical standards}
-
-\section*{Declaration of competing interest}
-
-The authors declare that there is no conflict of interest.
-
-\clearpage
-
-\section*{References}
-
-\bibliography{mybibfile}
-
-\end{document}
+It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^\prime$ is $s^*$ and therefore the solution space will not be explored effectively.
+On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.
+
+The method for generating the perturbation applied to the local optimum $s^*$ may vary depending on the problem, and may be based on previous solutions.
+In this way, it is possible to keep a history of the solutions reached so far, so they can be used when applying the perturbation.
+Algorithm \ref{alg:ILS} summarizes the \acs{ILS} overall optimization process.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{LocalSearch}{LocalSearch}
+	\SetKwFunction{Perturbation}{Perturbation}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Initial solution $s$}
+	\BlankLine
+	$s^* \leftarrow$ \LocalSearch{$s$}\\
+	\While{Termination criteria are not met}{
+
+		$s^\prime \leftarrow $ \Perturbation{$s^*$, $history$}\\
+		$s^{*\prime} \leftarrow$ \LocalSearch{$s^\prime$}
+
+		\If{$f(s^{*\prime}) < f(s^*)$}{
+
+			$s^* \leftarrow s^{*\prime}$
+
+		}	
+	}
+	\BlankLine
+	\KwRet ($s*$)
+
+	\caption{Basic Iterated Local Search}\label{alg:ILS}
+\end{algorithm}
+
+In the \acs{ILS} scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only use the difference in the cost of the solutions to be compared to choose between them.
+This is the most simple scheme and does not depends on previous solutions recorded in the history.
+
+\section{The Dual Iterative Local Search Method}
+
+The \acf{DILS} is a new variant of the classic \acs{ILS} method.
+Its goal is to perform a search in the solution space to find the solution with the best fitness value (given by fitness function $f$) by introducing diversity in an adaptive and guided way.
+While \acs{ILS} works with a single individual, \acs{DILS} keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima.
+The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value.
+
+In each generation $G$ of the optimization process \acs{DILS} builds a new individual by applying a recombination operator to $m_b$ and $m_w$.
+After that, it applies a strong mutation operator to the newly generated individual.
+This is the individual \acs{DILS} applies the \acs{LS} procedure to in order to improve its fitness value, resulting in the trial individual $m_t$.
+The trial individual $m_t$ replaces the worst of its parents, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for \acs{DILS}.
+Equation \ref{eq2.1} displays the expression for this criterion in the case of a minimization problem.
+
+\begin{equation}
+m_{w,G} = \left\{ \begin{array}{lc}
+m_t &   \text{if} \;\; f(m_t) < f(m_w)\\
+m_{w,G} &  \text{otherwise}
+\end{array}
+\right..
+\label{eq2.1}
+\end{equation}
+
+Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solution space.
+To avoid it \acs{DILS} implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory ($\{m_b, m_w\}$).
+Equation \ref{eq2.2} shows the reinitialization criterion for a minimization problem.
+
+\begin{equation}
+m_{w,G+1} = \left\{ \begin{array}{lc}
+\text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
+m_{w,G} &  \text{otherwise}
+\end{array}
+\right.,
+\label{eq2.2}
+\end{equation}
+
+\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
+This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best parent.
+Figure \ref{img:DILS_diag} summarizes the \acs{DILS} optimization process.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[scale=0.25]{gfx/NewProp/DILS/DILSDiagram.jpg}
+	\caption{Diagram summarizing \acs{DILS} optimization process.}\label{img:DILS_diag}
+\end{figure}
+
+It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively.
+This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
+
+This highlights the influence of the parameter $\xi$.
+Let us assume again a minimization problem.
+On one hand, if $\xi > 0$, then the true worst individual is only reinitialized when $m_t$ replaces $m_w$ and is also better than $m_b$ by a margin.
+On the other hand, if $\xi < 0$, we allow \acs{DILS} to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
+It becomes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
+Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$, and we recommend $[-0.5, 0.5]$.
+To summarize, the closer $\xi$ is to 1 the more restrictive the reinitialization criterion is---in the minimization case.
+It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages.
+The reason for this is that the fitness value of the best individual $f(m_b)$ never worsens ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it becomes increasingly harder for $m_w$ to achieve similar or better scores.
+Algorithm \ref{alg:DILS} summarizes the overall \acs{DILS} optimization process.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{RandInit}{RandInit}
+	\SetKwFunction{Best}{Best}
+	\SetKwFunction{Worst}{Worst}
+	\SetKwFunction{Recombination}{Recombination}
+	\SetKwFunction{Mutation}{Mutation}
+	\SetKwFunction{LocalSearch}{LocalSearch}
+	\KwIn{Probability for recombination operator $p_r$, segment size for mutation operator $p_s$, reinitialization method tolerance $\xi$.}
+	\tcp{Initialization phase}
+	$G \leftarrow 0$\\
+	$m_b \leftarrow$ \RandInit{};
+	$m_w \leftarrow$ \RandInit{}\\
+	\tcp{Main loop}
+	\While{Termination criteria are not met}{
+		\tcp{Find best and worst individual}
+		$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $};
+		$m_w \leftarrow $ \Worst{$ \{m_b, m_w\} $}\\
+		\tcp{Generate new individual}
+		$m_t \leftarrow $ \Recombination{$m_b, \; m_w$}\\
+		$m_t \leftarrow $ \Mutation{$m_t$}\\
+		\tcp{Improve new individual}
+		$m_t \leftarrow $ \LocalSearch{$m_t$}\\
+		\tcp{Apply replacement operator}
+		\If{$f(m_t) < f(m_w)$}{
+
+			$m_w \leftarrow m_t$\\
+
+		}
+		\tcp{Check restart criterion}
+		\If{$f(m_b) - f(m_w) > f(m_b) \times \xi$}{
+
+			$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
+			$m_w \leftarrow$ \RandInit{}\\
+
+		}
+
+		$G \leftarrow G + 1$
+	}
+
+	$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
+	\KwRet{$m_b$}
+	\caption{DILS}\label{alg:DILS}
+\end{algorithm}
+
+For the recombination and mutation operators that appear in Algorithm \ref{alg:DILS}, we use the uniform recombination and the segment mutation operator respectively.
+The uniform recombination consists in selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}.
+The segment mutation operator consists in replacing a fixed-size segment from the features of the individual with randomly generated features.
+The size of the segment $p_s$ must be fixed at the beginning of the optimization process.
+
+It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of \acs{DILS} for maximization problems.
+In addition, the \acs{LS} procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation details and termination criteria.
+
+\section{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}
+
+In this section, we present an optimization scheme for the \acs{DILS} algorithm when applied to constrained clustering.
+We will call this new approach \acs{DILS}$_{CC}$.
+We will use an integer-based representation for all the individuals \acs{DILS}$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---with $n$ being the number of instances of the dataset.
+Each integer $m_i \in [0,k)$ in $m$ represents the label of the cluster that instance $x_i$ is assigned to, and therefore each individual $m$ is a solution to the clustering problem.
+Additionally, in order not to bias the exploration of the solution space, we will use random initialization to set $m_b$ and $m_w$; this procedure will also be used in the reinitialization of $m_w$.
+
+Regarding the fitness function $f$, for each individual $m$ its fitness value $f_m$ can be calculated as in Equation \ref{eq2.3}.
+
+\begin{equation}
+f_m = z_m * (\text{infeasibility}_m + 1),
+\label{eq2.3}
+\end{equation}
+
+\noindent where $\text{infeasibility}_m$ is the number of non-satisfied constraints and $z_m$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2.4}.
+
+\begin{equation}
+z_m = \sum_{c_i \in C_m} \left[ \frac{\sum_{x_a, x_b \in c_i} d^2(x_a,x_b)}{|c_i|}\right].
+\label{eq2.4}
+\end{equation}
+
+We also need to specify an \acs{LS} procedure for \acs{DILS} to be able to improve the mutant individual to obtain the trial vector $m_t$.
+Algorithm \ref{alg:DISL_LS} shows the \acs{LS} used in \acs{DILS}$_{CC}$.
+Note that, in order not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
+This \acs{LS} procedure is similar to the one presented in Algorithm \ref{alg:SHADE_LS}, being the mos relevant difference the introduction of the maximum number of neighbors $p_m$ that can be generated in each call; this is done to ensure a better exploration-exploitation tradeoff.
+
+\begin{algorithm}
+	\SetNlSty{textbf}{[}{]}
+	\SetNlSkip{0.5em}
+	\setstretch{1.2}
+	\SetKwFunction{RandomShuffle}{RandomShuffle}
+	\SetKwFunction{Rand}{Rand}
+	\SetKwRepeat{Do}{do}{while}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $p_m$.}
+	\BlankLine
+	\While{$improvement$ {\bf and} $generated < p_m$}{
+		$improvement \leftarrow$ \texttt{false} \\
+		\tcp{Select random index (object) from $m$}
+		$i \leftarrow $ \Rand{$1$,$n$} \\
+		\tcp{Random shuffle labels set}
+		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
+		\For{$l \in RSL$}{
+			$m^\prime \leftarrow m$\\
+			\tcc{Move object $i$ from $m$ to the cluster associated with label $l$}
+			$m^\prime_i \leftarrow l$\\
+			\If{$f(m^\prime) < f(m)$}{
+				$m \leftarrow m^\prime$\\
+				$improvement \leftarrow$ \texttt{true} \\
+			}
+			$generated \leftarrow generated + 1$\\
+		}
+	}
+	\BlankLine
+	\KwRet{$m$}
+
+	\caption{Local Search}\label{alg:DISL_LS}
+\end{algorithm}

### 2019-06-26 Papers/DILS/paperDILS.tex
Nuevos capitulos en memoria y correcciones en papers

@@
 Clustering has always been a powerful tool in knowledge discovery.
-Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning.
-This new information can be given in form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
-We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to provide quality results for the constrained clustering problem.
-We will compare the results obtained by this proposal to those obtained by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
+Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering.
+This technique is a generalization of traditional clustering that considers additional information encoded by constraints.
+Constraints can be given in the form of instance-level
+must-link and cannot-link constraints, which this paper focuses on.
+We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to produce quality results for the constrained clustering problem.
+We compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them.
-A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features.
+A dataset $X$ is composed of $n$ instances, each one of them described by $u$ features.
 More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$.
@@
 The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it.
-As mentioned earlier, it is a difficult problem to solve, so metaheuristic are presented as a promising option to find quality approximate solutions.
-Metaheuristic algorithms are designed to explore the solution space associated with a problem guided by a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
+As mentioned earlier, its resolution poses a tough challenge, and metaheuristics are presented as a promising option to find quality approximate solutions.
+Metaheuristic algorithms are designed to explore the solution space of a problem with the guidance of a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
 In this type of approaches the key is to find a good exploration-exploitation tradeoff.
-This is the reason why the classic Local Search (LS) heuristic algorithm derived in the Iterative Local Search (ILS) metaheuristic algorithm, which introduces periodics perturbations in the exploration of the solution space to avoid local optima \cite{lourencco2010iterated}.
+This is the reason why the classic Local Search (LS) heuristic algorithm derived in the Iterative Local Search (ILS) metaheuristic algorithm, which introduces periodic perturbations in the exploration of the solution space to escape local optima \cite{lourencco2010iterated}.
 ILS---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
-ILS is also very commonly used to design hybrid methods with good exploration-exploitation tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
+ILS is also very commonly used in the design of hybrid methods with good exploration-exploitation tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
 
 In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptive and guided way.
-We apply it to the constrained clustering problem, prooving that it constitute a good approach to obtain quality results.
+We apply it to the constrained clustering problem, proving that it constitutes a competitive approach to obtain quality results.
 
 Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and the state-of-the-art.
-In Sections \ref{sec:DILS} and \ref{sec:DILS_CC} the DILS algorithm and its constrained clustering formulation will be reviewed.
-Sections from \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, results and their analysis respectively.
-Finally, in Section \ref{sec:conclusiones} conclusions are discussed.
+In Sections \ref{sec:DILS} and \ref{sec:DILS_CC} the DILS algorithm and its formulation for constrained clustering will be reviewed.
+Sections \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, the results and their analysis, respectively.
+Finally, in Section \ref{sec:conclusiones} the conclusions are discussed.
 
@@
 It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
-This method was named COP-kmeans, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
+This method, named COP-kmeans, introduces a modification to the assignment rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
 
-In \cite{antoine2012cecm} Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
+Within the fuzzy clustering family of methods, Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm, is proposed in \cite{antoine2012cecm}.
 The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
@@
 It uses a genetic algorithm, combined with an LS procedure and guided by a penalty-style fitness function, to obtain a partition of the input dataset.
-It constitutes one of the few approaches to constrained clustering from the point of view of the evolutionary computing.
+It constitutes one of the few approaches to constrained clustering from the point of view of evolutionary computing.
 
@@
 The Dual Iterative Local Search (DILS) is a new variant of the classic ILS method.
-Its goal is to perform a search in the solution space to find the solution with the best fitness function $f$ value by introducing diversity in an adaptive and guided way.
-While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-introducing methods and to avoid local optima.
-The individual $m_b$ is the one providing the best fitness value, so $m_w$ provides the worst fitness value.
+Its goal is to perform a search in the solution space to find the solution with the best fitness value (given by fitness function $f$) by introducing diversity in an adaptive and guided way.
+While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima.
+The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value.
 
@@
 After that, it applies a strong mutation operator to the newly generated individual.
-This is the individual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trail individual $m_t$.
-The trial individual $m_t$ replace the worst of its parents $m_w$ if it provides a better fitness value, this operation makes the acceptance criterion for DILS.
+This is the individual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trial individual $m_t$.
+The trial individual $m_t$ replaces the worst of its parents, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for DILS.
 Equation \ref{eq1} displays the expression for this criterion in the case of a minimization problem.
@@
 Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solution space.
-To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory $\{m_b, m_w\}$, Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.
+To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory ($\{m_b, m_w\}$).
+Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.
 
@@
 \noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
-This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via recombination operator) as the best parent.
+This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best parent.
 Figure \ref{img:DILS_diag} summarizes the DILS optimization process.
@@
 
-It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reasigned to the true best and worst individual respectively.
-This way, at the point where the reinitialization criterion is tested, it is possible that $f(m_w)$ provides a better value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
+It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively.
+This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
 
-This leads us to note the influence of the parameter $\xi$.
-In one hand, if we are dealing with a minimization problem and we set $\xi > 0$, then the true worst individual is only reinitializatized when the trial individual $m_t$ replaced $m_w$ and is also better than $m_b$ by a margin.
-On the other hand, and also for the minimization case, but this time with $\xi < 0$, we allow DILS to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
-It comes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
+This highlights the influence of the parameter $\xi$.
+Let us assume again a minimization problem.
+On one hand, if $\xi > 0$, then the true worst individual is only reinitialized when $m_t$ replaces $m_w$ and is also better than $m_b$ by a margin.
+On the other hand, if $\xi < 0$, we allow DILS to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
+It becomes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
 Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$, and we recommend $[-0.5, 0.5]$.
@@
 It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages.
-The reason being is that the fitness value of the best individual $f(m_b)$ never deteriorate ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it become harder for $m_b$ to get better o close to it.
+The reason for this is that the fitness value of the best individual $f(m_b)$ never worsens ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it becomes increasingly harder for $m_w$ to achieve similar or better scores.
 Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.
@@
 For the recombination and mutation operators that appear in Algorithm \ref{alg:DILS}, we use the uniform recombination and the segment mutation operator respectively.
-The uniform recombination consists of selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}.
-The segment mutation operator consists of replacing a fixed size segment from the features of the individual for randomly generated features.
+The uniform recombination consists in selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}.
+The segment mutation operator consists in replacing a fixed-size segment from the features of the individual with randomly generated features.
 The size of the segment $p_s$ must be fixed at the beginning of the optimization process.
@@
 It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems.
-In addition, the LS procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation matters and termination criteria.
+In addition, the LS procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation details and termination criteria.
 
@@
 We will call this new approach DILS$_{CC}$.
-To begin with, we will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---being $n$ the number of instances of the dataset.
-Each integer $m_i \in [0,k)$ in $m$ represent the label of the cluster that instance $x_i$ belongs to, so each individual $m$ is a solution to the clustering problem.
-Additionally, in order to not bias the exploration of the solution space, we will use a random initialization to initially set $m_b$ and $m_w$, this procedure will also be used when $m_w$ must be reinitialized.
+We will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---with $n$ being the number of instances of the dataset.
+Each integer $m_i \in [0,k)$ in $m$ represents the label of the cluster that instance $x_i$ is assigned to, and therefore each individual $m$ is a solution to the clustering problem.
+Additionally, in order not to bias the exploration of the solution space, we will use random initialization to set $m_b$ and $m_w$; this procedure will also be used in the reinitialization of $m_w$.
 
@@
 Algorithm \ref{alg:LS} shows the LS used in DILS$_{CC}$.
-Note that, in order to not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
-We also stablish the maximum number of neighbors $p_m$ that can be generated in each call, this is done to find a better exploration-exploitation tradeoff.
+Note that, in order not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$.
+We also establish the maximum number of neighbors $p_m$ that can be generated in each call; this is done to ensure a better exploration-exploitation tradeoff.
 
@@
 Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-We can note that DILS$_{CC}$ represents a consistent improvement over all other 5 methods.
-This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lower level of constraint-based information we work with.
+It is possible to notice that DILS$_{CC}$ represents a consistent improvement over all other 5 methods.
+This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lowest level of constraint-based information we work with.
 
@@
 
-Results obtained with the $SC_{15}$ constraint set are presented in Table \ref{tab:results15}.
+The results obtained with the $SC_{15}$ constraint set are presented in Table \ref{tab:results15}.
 We observe how DILS$_{CC}$ and COPKM are able to obtain ARI values equal to 1, meaning that the resulting partitions of the datasets perfectly match the true partitions.
-The fact that DILS$_{CC}$ is always able to output a partition of the datasets is worth the advantage over COPKM.
-It is even the case that COPKM is not able to provide an output partition while DILS finds the optimal partition or very close to the optimal partition for the same datasets.
+The fact that DILS$_{CC}$ is always able to output a partition of the datasets is a noteworthy advantage over COPKM.
+Even in the cases where COPKM is unable to produce a partition, DILS often finds the optimal or a near-optimal partition.
 
@@
 
-In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how DILS$_{CC}$ scales the quality of its results with the number of restrictions, as well as methods such as BRKGA or COPKM, the latter being more limited in other aspects.
-It is also worth noting how DILS$_{CC}$ provides the better average result for each of the three constraint set analyzed in this paper.
+In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how the quality of the results of DILS$_{CC}$ scales with the number of constraints, as well as that of methods such as BRKGA or COPKM (the latter being more limited in other aspects).
+It is also worth noting how DILS$_{CC}$ provides the best average ARI for each of the three constraint sets analyzed in this paper.
 
@@
 All heatmaps suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle.
-Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that DILS$_{CC}$ does represent a significant improvement over all other six methods, being te results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.
+Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that DILS$_{CC}$ does represent a significant improvement over all other six methods, being the results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.
 
@@
 In this paper we proposed the DILS heuristic method, along with its application to the SSL constrained clustering problem, which we called DILS$_{CC}$.
-Focusing on the instance-level ML and CL constraints, DILS$_{CC}$ has proven that heuristic techniques constitute a good approach to constrained clustering, being able to scale the quality of the results in a way directly proportional to the number of constraints.
+Focusing on instance-level ML and CL constraints, DILS$_{CC}$ has proven that heuristic techniques constitute a competitive approach to constrained clustering, being able to scale the quality of the results in a directly proportional way to the number of constraints.
 

### 2019-06-26 Papers/SHADE/paperSHADE.tex
Empezada memoria TFM
Nuevos capitulos en memoria y correcciones en papers

@@
 This technique is a generalization of traditional clustering that considers additional information encoded by constraints.
-Constraints can be given in the form of instance-level must-link and cannot-link constraints, in which this paper focuses on.
+Constraints can be given in the form of instance-level must-link and cannot-link constraints, which this paper focuses on.
 We propose the first application of Differential Evolution to the constrained clustering problem, which has proven to produce a better exploration-exploitation trade-off when comparing with previous approaches.
@@
 	\BlankLine
-	\KwRet ($S$)
+	\KwRet{$S$}
 
@@
 			\hline
+			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
 			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\

### 2019-06-28 Memoria/Chapters/Clustering.tex
Correcciones y nuevas secciones

@@
 For this purpose we will follow the study carried out in \cite{Everitt:2009:CA:1538772}.
+In Section  \ref{sec:ReasonsClassif} we present the reasons that lead humans to develop automatic classification methods.
+The numerical formulation for the latter is given in Section \ref{sec:NumMethClassif}.
+We give an intuitive definition for ``cluster'' in Section \ref{sec:WhatIsCluster}.
+Lastly a list of clustering application is presented in Section \ref{sec:ClusteringApplications}.
 
-\section{Reasons for Classifying}
+\section{Reasons for Classifying} \label{sec:ReasonsClassif}
 
@@
 
-\section{Numerical Methods of Classification}
+\section{Numerical Methods of Classification} \label{sec:NumMethClassif}
 
@@
 
-\subsection{What is a Cluster?} \label{sec:WhatIsCluster}
+\section{What is a Cluster?} \label{sec:WhatIsCluster}
 
@@
 
-\section{Applications of Clustering}
+\section{Applications of Clustering} \label{sec:ClusteringApplications}
 

### 2019-06-28 Memoria/Chapters/ConstrainedClustering.tex
Nuevos capitulos en memoria y correcciones en papers
Correcciones y nuevas secciones

@@
 To this end we take as our main reference the survey presented in \cite{davidson2007survey}.
+We will start by describing the motivation for constrained clustering in Section \ref{sec:CCMotiv}.
+We formalize the concept of constraints and discuss their use in Sections \ref{sec:ConstDef} and \ref{sec:ConstUse} respectively.
+Afterwards a list of constrained clustering application is presented in Section \ref{sec:CCApplications}.
+Benefits and problems, as well as solutions for these problems, are discussed in Sections \ref{sec:ConstraintsBenefits} and \ref{sec:ConstraintsProblems} respectively.
 
-\section{Motivation for Constrained Clustering}
+\section{Motivation for Constrained Clustering} \label{sec:CCMotiv}
 
@@
 
-\section{Constraints Definition}
+\section{Constraints Definition} \label{sec:ConstDef}
 
@@
 
-\section{Use of Constraints}
+\section{Use of Constraints} \label{sec:ConstUse}
 
@@
 
-Figure \ref{fig:ConstOverDataset} shows a dataset together with its associated constraints. Figure \ref{fig:ClusteringSatAllConst} proposes a possible clustering that satisfies all constraints.
+Figure \ref{fig:ConstClustOverDataset} shows a dataset together with its associated constraints and proposes a possible clustering that satisfies all constraints.
 
@@
 	\myfloatalign
-	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/InputInstancesAndConst1}
-	\caption[Constraints on a dataset.]{Constraints on a dataset \cite{davidson2007survey}.} \label{fig:ConstOverDataset}
-	}
-	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/ClusteringSatAll}
-	\caption[Clustering satisfying all constraints.]{Clustering satisfying all constraints \cite{davidson2007survey}.} \label{fig:ClusteringSatAllConst}
-	}
+	\subfloat[Constraints over a dataset.]
+	{\includegraphics[width=.45\linewidth]{gfx/ConstClust/InputInstancesAndConst1}}
+	\hspace{1cm}
+	\subfloat[Clustering satisfying all constraints.]
+	{\includegraphics[width=.45\linewidth]{gfx/ConstClust/ClusteringSatAll}}
+
+	\caption[Constraints and clustering over a dataset]{Constraints and clustering over a dataset \cite{davidson2007survey}.} \label{fig:ConstClustOverDataset}
 \end{figure}
@@
 
-Figure \ref{fig:MetricLearned} shows a possible clustering based on a metric learned from the constraints specified in Figure \ref{fig:ConstOverDataset2}.
+Figure \ref{fig:ConstrainsAndMetricLearned} shows a possible clustering based on a metric learned from the constraints.
 It should be noted that in Figure \ref{fig:MetricLearned} the data space has been compressed on the vertical axis and widened on the horizontal axis to match the learned distance metric.
@@
 	\myfloatalign
-	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/InputInstancesAndConst2}
-	\caption[Constraints on a dataset.]{Constraints on a dataset \cite{davidson2007survey}.} \label{fig:ConstOverDataset2}
-	}
-	{\includegraphics[width=.6\linewidth]{gfx/ConstClust/MetricaAprendida}
-	\caption[Clustering based on a metric learned based on constraints.]{Clustering based on a metric learned based on constraints \cite{davidson2007survey}.} \label{fig:MetricLearned}
-	}
+	\subfloat[Constraints over a dataset.]
+	{\includegraphics[width=.45\linewidth]{gfx/ConstClust/InputInstancesAndConst2}} 
+	\hspace{0.5cm}
+	\subfloat[Clustering based on a metric learned from constraints.]
+	{\includegraphics[width=.45\linewidth]{gfx/ConstClust/MetricaAprendida}
+	\label{fig:MetricLearned}}
+	\caption[Clustering based on a metric learned from constraints.]{Clustering based on a metric learned based on constraints \cite{davidson2007survey}.} \label{fig:ConstrainsAndMetricLearned}
 \end{figure}
@@
 Figure \ref{fig:CMUFacesDatabase} shows a sample from the CMU (Carnegie Mellon University) faces dataset, where the task is to group faces based on different criteria.
-In this case, the goal is to group the faces according to their orientation.
-
-\clearpage
+In this case, the goal is to group faces according to their orientation.
 
@@
 
-\section{Benefits of Using Constraints}
+\section{Benefits of Using Constraints} \label{sec:ConstraintsBenefits}
 
@@
 
-\subsection{The Feasibility Problem}
+\subsection{The Feasibility Problem} \label{sec:FeasibilityProblem}
 

### 2019-06-28 Memoria/Chapters/ExperimentalResults.tex
Correcciones y nuevas secciones

@@
 \chapter{Experimental Results} \label{ch:ExperimentalResults}
+
+This chapter presents the results obtained by the two new proposals introduced in this work (in Chapter \ref{ch:NewProposals}) and those obtained by the state-of-the-art.
+Being the mentioned its main purpose, we firstly introduce a brief study on the influence of the parameter $\xi$ in the \acs{DILS} method in Section \ref{sec:XiInfl}, so the reader gets a better comprehension of theory presented in Section \ref{sec:DILS}.
+Afterwards we will carry out an in-depth comparison between \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$ in Section \ref{sec:BRKGAvsSHADE}, as both are population-based and heuristic-guided methods.
+Finally, in Section \ref{sec:NewPropvsSOTA}  we will present a more general comparison that will include all state-of-the-art methods and the new proposals.
+
+It is worth noting that, since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
+To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+
+\section{On the Influence of $\xi$} \label{sec:XiInfl}
+
+Parameter $\xi$ from \acs{DILS} plays a key role in the development of its optimization process.
+It was theoretically introduced in Section \ref{sec:DILS}.
+The aim of this section is to prove its influence in the \acs{DILS} constrained clustering application (\acs{DILS}$_{CC}$).
+With this in mind we set up the following experiment: apply \acs{DILS}$_{CC}$ to the Iris dataset and its associated $CS_{10}$ constraint set and save the number of worst individual ($m_w$) reinitializations in each generations.
+We have to keep in mind that \acs{DILS}$_{CC}$ is a non deterministic algorithm, due to its \acf{ILS} procedure.
+To mitigate the effect this could have of our study we will repeat the experiment 20 times, with the objective of analyzing average results and standard deviation.
+
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[Restarts per generation with $\xi \le 0$]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMenor.pdf}} 
+	\hspace{1cm}
+	\subfloat[Restarts per generation with $\xi \ge 0$]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMayor.pdf}}
+	\caption{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
+	\label{fig:RestartsLines}
+\end{figure}
+
+Figure \ref{fig:RestartsLines} shows the number of total reinitializations per generation for $\xi \in [-0.5,0.5]$.
+Note that the $y$ axis is on a continuous scale due to the averaging process.
+As we expected this number grows in a logarithmic way with respect to the number of generations, which means that the algorithm tends to converge if a sufficient number of generations.
+Also it is clear that the higher $\xi$ the less reinitializations per generations are done.
+
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[Hisogram showing values]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/HistValues.pdf}} 
+	\hspace{1cm}
+	\subfloat[Hisogram showing variance]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/HistVar.pdf}}
+	\caption{Mean number of reinitializations histogram for $\xi \in [-0.5,0.5]$.}
+	\label{fig:RestartsHist}
+\end{figure}
+
+Figure \ref{fig:RestartsHist} aims to present a more clear visual representation of the average number of reinitializations and the standard deviation found in the total number of reinitializations for the 20 runs and for each $\xi$ value.
+We observe how the average number of reinitializations grows in an irregular way as $\xi$ decreases.
+This effect would tend to lessen as we increase the size of the sample runs increases.
+To put it clear: the more runs the more regular the grow of the average number reinitializations would be.
+Regarding the variance, we can clearly see how the more permissive the reinitilization rule is the higher it gets.
+Which is exactly what we would expect in a non deterministic method such as \acs{DILS}$_{CC}$.
+
+We can conclude that the parameter $\xi$ has the expected influence over the \acs{DILS}$_{CC}$ optimization process.
+Note that in a problem in which the \acs{LS} procedure used is deterministic we would also find differences between consecutive runs over the same datasets, given that the \acs{DILS} general optimization scheme uses stochastic tools---such as the recombination and mutation operators---to make sure that diversity is introduced in the search of the global optima for the problem we apply it to.
+
+\section{Comparing BRKGA$_{CC}$ with SHADE$_{CC}$} \label{sec:BRKGAvsSHADE}
+
+Tables from \ref{tab:resultsBRKGAvsSHADE10} to \ref{tab:resultsBRKGAvsSHADE20} present the comparison between \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.
+In them the \acs{ARI} columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
+
+\clearpage
 
@@
 			\multirow{2}{*}{Dataset} &
-			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
+			\multicolumn{3}{c}{\acs{BRKGA}$_{CC}$} &&  \multicolumn{3}{c}{\acs{SHADE}$_{CC}$} \\
 			\cline{2-4} \cline{6-8}
-			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
+			& \acs{ARI} & Unsat(\%) & Time(s) && \acs{ARI} & Unsat(\%) & Time(s) \\
 			\hline
@@
 			\hline
-			Mean & \textbf{0.161} & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
+			Mean & \textbf{0.165} & 10.437 & 1328.026 && 0.162 & 9.837 & 1306.276 \\
 			\hline
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
-	\label{tab:results10}
+	\caption[Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.]{Experimental results obtained for $CS_{10}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
+	\label{tab:resultsBRKGAvsSHADE10}
 \end{table}
+
+Table \ref{tab:resultsBRKGAvsSHADE10} shows the results for the $SC_{10}$ constraint set.
+We can note that \acs{SHADE}$_{CC}$ represents a consistent improvement over \acs{BRKGA}$_{CC}$ in both Unsat and Time, while in terms of \acs{ARI} both methods are similar.
+This is because the penalty applied to the fitness function of both methods has little presence when operating with a small constraint set, and since they are population-based and heuristic-guided methods, they have similar exploration and exploitation capabilities.
 
@@
 			\hline
-			Mean & 0.483 & 9.456 & 1784.931 && \textbf{0.507} & 6.997 & 1751.368 \\
+			Mean & 0.535 & 7.79 & 1402.49 && \textbf{0.547} & 5.777 & 1369.658 \\
 			\hline
@@
 
-	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
-	\label{tab:results15}
+	\caption[Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.]{Experimental results obtained for $CS_{15}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
+	\label{tab:resultsBRKGAvsSHADE15}
 \end{table}
+
+Table \ref{tab:resultsBRKGAvsSHADE15} presents the results obtained with the $CS_{15}$ constraint set.
+While conclusions for Unsat and Time remain unchanged, we see a clear difference as far as \acs{ARI} is concerned.
+We observe that both methods already report \acs{ARI} results above 0.7, which in an \acs{ARI} ranking is considered to be an acceptable result.
+We can even observe that the \acs{SHADE}$_{CC}$ method provides an \acs{ARI} higher than 0.9 for 7 datasets, compared to \acs{BRKGA}$_{CC}$ that only obtains similar results in 2 datasets.
 
@@
 			\hline
-			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
+			Mean & 0.666 & 6.575 & 1481.007 && \textbf{0.713} & 4.087 & 1461.155 \\
 			\hline
@@
 
-	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
-	\label{tab:results20}
+	\caption[Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.]{Experimental results obtained for $CS_{20}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
+	\label{tab:resultsBRKGAvsSHADE20}
 \end{table}
+
+It is in Table \ref{tab:resultsBRKGAvsSHADE20}, which gathers results obtained for the constraint set $CS_{20}$, that we observe major differences between the two methods.
+We find that \acs{SHADE}$_{CC}$ is capable of producing clusters identical to the original ones for 4 datasets, regardless the single dataset for which \acs{BRKGA}$_{CC}$ achieves it.
+Let us remember that an \acs{ARI} of 1 represents a perfect match between the two partitions given to compute it.
+This improvement in \acs{ARI} is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, \acs{SHADE}$_{CC}$ outperforms \acs{BRKGA}$_{CC}$.
+Conclusions on execution time remain the same.
+
+\newpage
+
+\section{Comparing SHADE$_{CC}$ and DILS$_{CC}$ with The-State-Of-The-Art} \label{sec:NewPropvsSOTA}
+
+In this section we present Tables from \ref{tab:resultsSOTA10} to \ref{tab:resultsSOTA20}, which display the results obtained by the new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---and six state-of-the-art methods---including \acs{BRKGA}$_{CC}$---to be compared for each dataset and constraint set.
+
+It should be noted that there are some missing results in these tables.
+In the case of the \acs{COPKM} algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
+It is possible that \acs{COPKM} cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
+In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
+We establish that in these cases the \acs{ARI} value considered for the mean calculation---and the forthcoming statistical analysis of results---is -1.000, which is the worst value the \acs{ARI} measure can deliver.
+
+\begin{table}[!h]
+	\centering
+	\setlength{\tabcolsep}{7pt}
+	\renewcommand{\arraystretch}{1.2}
+	%\begin{adjustwidth}{-1in}{-1in}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{lcccccccc}
+			\hline
+			Dataset & \acs{SHADE}$_{CC}$ & \acs{DILS}$_{CC}$ & \acs{BRKGA}$_{CC}$ & \acs{COPKM} & \acs{LCVQE} & \acs{RDPM} & \acs{TVClust} & \acs{CECM} \\
+			\hline
+			Appendicitis & 0.086 & 0.611 & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
+			Breast Cancer & 0.159 & 0.755 & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.104 & 0.870 & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
+			Circles & 0.063 & 0.798 & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
+			Ecoli & 0.020 & 0.069 & 0.009 & - & 0.387 & 0.417 & 0.265 & - \\
+			Glass & 0.013 & 0.009 & 0.016 & 0.184 & 0.268 & 0.197 & 0.211 & - \\
+			Haberman & 0.084 & 0.638 & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
+			Hayesroth & 0.025 & 0.031 & 0.027 & - & 0.106 & 0.097 & 0.054 & 0.139 \\
+			Ionosphere & 0.332 & 0.792 & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
+			Iris & 0.342 & 0.621 & 0.238 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
+			Monk2 & 0.270 & 0.815 & 0.421 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
+			Moons & 0.272 & 0.968 & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
+			Saheart & 0.265 & 0.800 & 0.147 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
+			Sonar & 0.116 & 0.649 & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
+			Soybean & 0.515 & 0.304 & 0.496 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
+			Spectfheart & 0.343 & 0.896 & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
+			Spiral & 0.114 & 0.856 & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
+			Tae & 0.024 & 0.031 & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
+			Vowel & 0.002 & 0.002 & 0.003 & - & 0.063 & -0.003 & 0.067 & - \\
+			Zoo & 0.098 & 0.151 & 0.105 & 0.715 & 0.666 & 0.412 & 0.335 & - \\
+			\hline
+			Mean & 0.162 & \textbf{0.533} & 0.165 & -0.527 & 0.199 & 0.199 & 0.125 & -0.132 \\
+			\hline
+
+	\end{tabular}}
+	%\end{adjustwidth}
+
+	\caption{Experimental results obtained for $CS_{10}$ by the new proposals and the state-of-the-art methods.}
+	\label{tab:resultsSOTA10}
+\end{table}
+
+Table \ref{tab:resultsSOTA10} presents the results obtained for the $CS_{10}$ constraint set.
+\acs{LCVQE}, \acs{RDPM} and \acs{TVClust} methods stand out of \acs{SHADE}$_{CC}$, achieving results as competitive as it with a low number of constraints.
+Nevertheless \acs{DILS}$_{CC}$ represents a consistent improvement over all other 6 state-of-the-art methods.
+This is due to the fact that the exploitation capacity of \acs{DILS}$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lowest level of constraint-based information we work with.
 
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
+		\begin{tabular}{lcccccccc}
 			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			Dataset & \acs{SHADE}$_{CC}$ & \acs{DILS}$_{CC}$ & \acs{BRKGA}$_{CC}$ & \acs{COPKM} & \acs{LCVQE} & \acs{RDPM} & \acs{TVClust} & \acs{CECM} \\
 			\hline
-			Appendicitis & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
-			Breast Cancer & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
-			Circles & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
-			Ecoli & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
-			Glass & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
-			Haberman & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
-			Hayesroth & 0.031 & 0.027 & - & \textbf{0.106} & 0.097 & 0.054 & 0.139 \\
-			Ionosphere & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
-			Iris & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
-			Monk2 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
-			Moons & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
-			Saheart & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
-			Sonar & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
-			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
-			Spectfheart & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
-			Spiral & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
-			Tae & \textbf{0.031} & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
-			Vowel & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
-			Zoo & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
+			Appendicitis & 0.332 & 0.957 & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
+			Breast Cancer & 0.911 & 0.788 & 0.772 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Bupa & 0.801 & 0.984 & 0.840 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
+			Circles & 0.982 & 1.000 & 0.828 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
+			Ecoli & 0.025 & 0.134 & 0.015 & - & 0.387 & 0.372 & 0.719 & - \\
+			Glass & 0.031 & 0.057 & 0.017 & - & 0.280 & 0.253 & 0.229 & - \\
+			Haberman & 0.796 & 1.000 & 0.866 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
+			Hayesroth & 0.099 & 0.386 & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
+			Ionosphere & 0.591 & 0.970 & 0.803 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
+			Iris & 0.260 & 0.850 & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
+			Monk2 & 0.945 & 0.902 & 0.751 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
+			Moons & 0.995 & 1.000 & 0.958 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Saheart & 0.945 & 0.862 & 0.749 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
+			Sonar & 0.890 & 0.981 & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
+			Soybean & 0.422 & 0.469 & 0.428 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
+			Spectfheart & 0.778 & 1.000 & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
+			Spiral & 0.949 & 1.000 & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
+			Tae & 0.066 & 0.299 & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
+			Vowel & 0.001 & 0.003 & 0.002 & - & 0.063 & -0.003 & 0.073 & - \\
+			Zoo & 0.128 & 0.185 & 0.157 & 0.435 & 0.642 & 0.450 & 0.353 & - \\
 			\hline
-			Mean & \textbf{0.474} & 0.158 & -0.510 & 0.226 & 0.212 & 0.175 & -0.186 \\
+			Mean & 0.547 & \textbf{0.691} & 0.535 & 0.05 & 0.197 & 0.216 & 0.221 & -0.148 \\
 			\hline
@@
 
-	\caption{Experimental results obtained for $CS_{10}$ by DILS$_{CC}$ and the state-of-the-art methods.}
-	\label{tab:results10}
+	\caption{Experimental results obtained for $CS_{15}$ by the new proposals and the state-of-the-art methods.}
+	\label{tab:resultsSOTA15}
 \end{table}
+
+In one hand, Table \ref{tab:resultsSOTA15}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by \acs{SHADE}$_{CC}$.
+On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by \acs{SHADE}$_{CC}$, with the exception of \acs{COPKM}.
+Also we observe how \acs{DILS}$_{CC}$ and \acs{COPKM} are able to obtain \acs{ARI} values equal to 1, meaning that the resulting partitions of the datasets perfectly match the true partitions.
+The fact that \acs{DILS}$_{CC}$ is always able to output a partition of the datasets is a noteworthy advantage over \acs{COPKM}.
+Even in the cases where \acs{COPKM} is unable to produce a partition, \acs{DILS} often finds the optimal or a near-optimal partition.
 
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
+		\begin{tabular}{lcccccccc}
 			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			Dataset & \acs{SHADE}$_{CC}$ & \acs{DILS}$_{CC}$ & \acs{BRKGA}$_{CC}$ & \acs{COPKM} & \acs{LCVQE} & \acs{RDPM} & \acs{TVClust} & \acs{CECM} \\
 			\hline
-			Appendicitis & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Breast Cancer & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
-			Bupa & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
-			Circles & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Ecoli & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
-			Glass & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
-			Haberman & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
-			Hayesroth & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
-			Ionosphere & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
-			Iris & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
-			Monk2 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
-			Moons & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
-			Saheart & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
-			Sonar & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
-			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
-			Spectfheart & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
-			Spiral & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
-			Tae & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
-			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
-			Zoo & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
+			Appendicitis & 1.000 & 1.000 & 1.000 & - & 0.305 & 0.331 & 0.012 & -0.006 \\
+			Breast Cancer & 0.977 & 0.807 & 0.807 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.993 & 0.991 & 0.936 & 1.000 & -0.005 & -0.007 & -0.006 & 0.000 \\
+			Circles & 1.000 & 1.000 & 0.935 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
+			Ecoli & 0.047 & 0.201 & 0.028 & - & 0.387 & 0.459 & 0.763 & - \\
+			Glass & 0.082 & 0.322 & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
+			Haberman & 0.997 & 1.000 & 0.943 & 1.000 & -0.002 & 0.106 & 0.737 & 0.000 \\
+			Hayesroth & 0.531 & 0.812 & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
+			Ionosphere & 0.993 & 0.982 & 0.902 & 1.000 & 0.173 & 0.309 & 0.000 & 0.115 \\
+			Iris & 0.553 & 0.968 & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
+			Monk2 & 0.991 & 0.904 & 0.855 & 1.000 & 0.072 & 0.253 & -0.002 & 0.160 \\
+			Moons & 0.995 & 1.000 & 0.937 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
+			Saheart & 0.981 & 0.881 & 0.881 & 1.000 & 0.018 & 0.026 & 1.000 & -0.006 \\
+			Sonar & 1.000 & 1.000 & 0.992 & 1.000 & 0.004 & 0.127 & 0.000 & 0.003 \\
+			Soybean & 0.563 & 0.564 & 0.416 & -0.218 & 0.545 & 0.631 & 0.000 & -0.016 \\
+			Spectfheart & 1.000 & 1.000 & 0.953 & 1.000 & -0.107 & -0.112 & 0.000 & -0.054 \\
+			Spiral & 0.998 & 1.000 & 0.921 & 1.000 & -0.003 & 0.011 & 0.006 & 0.045 \\
+			Tae & 0.392 & 0.892 & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
+			Vowel & 0.003 & 0.003 & 0.002 & - & 0.063 & -0.003 & 0.071 & - \\
+			Zoo & 0.153 & 0.203 & 0.176 & 0.821 & 0.642 & 0.439 & 0.335 & - \\
 			\hline
-			Mean & \textbf{0.641} & 0.490 & -0.070 & 0.225 & 0.233 & 0.258 & -0.196 \\
+			Mean & 0.713 & \textbf{0.777} & 0.666 & 0.23 & 0.195 & 0.273 & 0.298 & -0.135 \\
 			\hline
@@
 
-	\caption{Experimental results obtained for $CS_{15}$ by DILS$_{CC}$ and the state-of-the-art methods.}
-	\label{tab:results15}
+	\caption{Experimental results obtained for $CS_{20}$ by the new proposals and the state-of-the-art methods.}
+	\label{tab:resultsSOTA20}
 \end{table}
 
-\begin{table}[!h]
-	\centering
-	\setlength{\tabcolsep}{7pt}
-	\renewcommand{\arraystretch}{1.4}
-	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{lccccccc}
-			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
-			\hline
-			Appendicitis & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Breast Cancer & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
-			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
-			Glass & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
-			Haberman & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
-			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
-			Iris & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
-			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
-			Saheart & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
-			Sonar & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
-			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
-			Spectfheart & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
-			Spiral & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
-			Tae & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
-			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
-			Zoo & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
-			\hline
-			Mean & \textbf{0.733} & 0.642 & 0.064 & 0.223 & 0.285 & 0.344 & -0.187 \\
-			\hline
+Table \ref{tab:resultsSOTA20} presents the results obtained for $CS_{20}$.
+For \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$ we continue to observe its progression in accuracy, as well as for methods such as \acs{BRKGA}$_{CC}$ or \acs{COPKM} (the latter being more limited in other aspects).
+As far as the state-of-the-art methods are concerned, we find a general improvement in \acs{ARI}, except for CECM.
+Even with this, it is \acs{SHADE}$_{CC}$ and \acs{COPKM} that seem to obtain the highest quality results, with the difference that \acs{SHADE}$_{CC}$ is able to produce results in all cases.
+It is also worth noting how \acs{DILS}$_{CC}$ provides the best average \acs{ARI} for each of the three constraint sets analyzed in this paper.
 
-	\end{tabular}}
-	%\end{adjustwidth}
-
-	\caption{Experimental results obtained for $CS_{20}$ by DILS$_{CC}$ and the state-of-the-art methods.}
-	\label{tab:results20}
-\end{table}
+Regarding the comparison between the two new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---, we have seen how both of them represent an improvement over previous constrained clustering approaches, being \acs{DILS}$_{CC}$ the best of the two.
+This is due to the fact that a real-based problem representation does not provide any other advantage that the possibility to apply real-coding evolutionary methods.
+This way, given that the \acs{SHADE} method proved to be the best approach to real optimization problems we expected it to outperform \acs{BRKGA}, though \acs{SHADE} imposes an arithmetic relationship between individuals---through the mutation operator---which is hard to understand in the clustering domain.
+This is, applying an arithmetic operator to two labels does not provide easily understandable combination of them.
+Judging from the raw results, and with its straight integer-based problem representation and its strong exploitation capability \acs{DILS}$_{CC}$ stands out as the best of all methods we obtained results for.

### 2019-06-28 Memoria/Chapters/ExperimentalSetup.tex
Correcciones y nuevas secciones

@@
 
-In this section we present the experimental setup used to compare the two new proposals introduced in Part \ref{pt:NewProposals} with the current state of the art.
-We will start by briefly introducing the state-of-the-art methods.
-Afterwards we specify the dataset used in experimentation, as well as the constraint sets and the results evaluation method.
-We also include the required specifications to replicate the experiments.
+In this chapter we present the experimental setup used to compare the two new proposals introduced in Chapter \ref{ch:NewProposals} with the current state of the art.
+We will start by briefly introducing the state-of-the-art methods in Section \ref{sec:SOTA}.
+Afterwards we specify the dataset used in experimentation (Section \ref{sec:Datasets}), as well as the constraint sets (Section \ref{sec:ConstGen}) and the results evaluation method (Section \ref{sec:EvalMet}).
+In Section \ref{sec:ValidtnMethod} we introduce the validation method we will apply to the results.
+We also include the required specifications to replicate the experiments in Section \ref{sec:Calibration}.
 
-\subsection{State-of-the-art Methods} \label{sec:SOTA}
+\section{State-of-the-art Methods} \label{sec:SOTA}
 
@@
 
-\section{Datasets}
+\section{Datasets} \label{sec:Datasets}
 
@@
 
-Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}).
+Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGen}).
 They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).
 
-\section{Constraint Generation} \label{sec:ConstGent}
+\section{Constraint Generation} \label{sec:ConstGen}
 
@@
 We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with three small percentages of the size of the dataset: 10\%, 15\% and 20\%.
-With $n_f$ being the fraction of the size of the dataset associated with each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
-
-The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
+With $n_f$ being the fraction of the size of the dataset associated with each of these percentages, the formula $(n_f(n_f-1))/2$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.
 Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.
@@
 
+The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
 Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
 This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.
+Figure \ref{fig:IrisConst} displays a visual representation of the constraint sets generated for the Iris dataset.
 
-\subsection{Evaluation Method} \label{sec:EvalMet}
+\begin{figure}[bth]
+	\myfloatalign
+	\subfloat[$CS_{10}$ \acs{ML} constraints]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML10.png}} 
+	\subfloat[$CS_{10}$ \acs{CL} constraints]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL10.png}} \quad
+	\subfloat[$CS_{15}$ \acs{ML} constraints]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML15.png}} 
+	\subfloat[$CS_{15}$ \acs{CL} constraints]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL15.png}} \quad
+	\subfloat[$CS_{20}$ \acs{ML} constraints]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML20.png}} 
+	\subfloat[$CS_{20}$ \acs{CL} constraints]
+	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL20.png}} 
+	\caption{Visual representation of the three constraints sets the Iris dataset.}
+	\label{fig:IrisConst}
+\end{figure}
 
-Since we have the true labels associated to each of the datasets, we can use them to evaluate the results provided by each method.
+\section{Evaluation Method} \label{sec:EvalMet}
+
+Since we have the true labels associated with each of the datasets, we can use them to evaluate the results provided by each method.
 We will use the \acf{ARI} to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
@@
 
-In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic Null Hypothesis Statistical Tests (NHST).
-In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in.
-\textit{"In a nutshell: NHST do not answer the question we ask"}.
-To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
+In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic \acf{NHST}.
+In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of \acs{NHST}, and a new model is proposed for carrying out comparisons researchers are interested in.
+\textit{"In a nutshell: \acs{NHST} do not answer the question we ask"}.
+To put it clear, the disadvantages of the \acs{NHST} that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
 
-To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
-Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
+To start with, \acs{NHST} do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
+Another pitfall of \acs{NHST} is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
 This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.
 
-Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate.
-As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
+Also, \acs{NHST} do not provide information about the magnitude of the effects and the uncertainty of its estimate.
+As a consequence, \acs{NHST} may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.
 
-Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis!
-That is: What can we conclude when NHST do not reject the null hypothesis?
-We can not infer anything since NHST can not provide evidence in its favor.
+Furthermore, and this is a situation that all researchers have faced, \acs{NHST} do not provide any information about the null hypothesis!
+That is: What can we conclude when \acs{NHST} do not reject the null hypothesis?
+We can not infer anything since \acs{NHST} can not provide evidence in its favor.
 
-Finally, there are two other problems that researchers face when performing NHST.
+Finally, there are two other problems that researchers face when performing \acs{NHST}.
 The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results.
@@
 
-As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST.
+As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of \acs{NHST}.
 In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test.
@@
 
-\section{Calibration}
+\section{Calibration} \label{sec:Calibration}
 
@@
 	%\begin{adjustwidth}{-1in}{-1in}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c m{5cm} cc}
+	%\resizebox{\textwidth}{!}{
+		\begin{tabular}{>{\centering\arraybackslash}c >{\arraybackslash}L{5cm} cc}
 			\hline
@@
 
-	\end{tabular}}
+	\end{tabular}%}
 	%\end{adjustwidth}
@@
 \end{table}
+
+\newpage
 
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c m{6cm} c}
+		\begin{tabular}{c L{6cm} c}
 			\hline
@@
 To compare with the state-of-the-art methods mentioned in Section \ref{sec:SOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA}.
+
+\clearpage
 

### 2019-06-29 Memoria/Chapters/AnalysisResults.tex
Escrito analisis de resultados en memoria

\chapter{Empirical Analysis of Results} \label{ch:AnalysisResults}

With the results obtained by all methods for a total of 60 different datasets---the 20 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them. This way we can statistically determine whether the new proposals---\acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$---represent a significant improvement over previous proposals.

One of the major advantages of the Bayesian sign test---introduced in Section \ref{sec:ValidtnMethod}---is that we can obtain a very illustrative visual representation of its results. We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates. With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle. In order to find out where a certain triplet will be placed within the triangle, we will take each of its three values and draw a parallel line to the opposing side of the corresponding vertex; the separation between a triangle side and its parallel line will be proportional to the associated triplet value, so that the higher the value, the closer the line will be to the vertex. The location where the three lines intersect is where we draw a point. Since the values of every triplet describe a probability distribution and therefore they must add up to one, we can be sure that all triplets will lie in some point within the triangle. The color indicates the density of points in a given region, with yellow representing a high density and red a low density.

\section{Testing SHADE$_{CC}$} \label{sec:TestSHADE}

In this section we will determine if there are significant differences between \acs{SHADE}$_{CC}$ and all 6 state-of-the-art-method. To do so we will perform an in-depth statistical comparison between \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$, as well as a more general comparison concerning the state-of-the-art methods.

Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by \acs{BRKGA}$_{CC}$ as sample $A$, and to the results obtained with \acs{SHADE}$_{CC}$ as sample $B$.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[\acs{ARI}]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_ARI.pdf}\label{fig:bayesARI}} \quad
	\subfloat[Time]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_TIME.pdf}\label{fig:bayesTime}}
	\subfloat[Unsat]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_UNSAT.pdf}\label{fig:bayesUnsat}}
	\caption[Heatmaps for the comparison of SHADE$_{CC}$ with BRKGA$_{CC}$.]{Heatmaps for the comparison of \acs{SHADE}$_{CC}$ with \acs{BRKGA}$_{CC}$.}
	\label{fig:HeatmapsSHADEvsBRKGA}
\end{figure}

In one hand, figure \ref{fig:bayesARI} shows the heatmap associated with the \acs{ARI} measurement. The diagram suggests that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle. It also indicates that in some cases there are no significant differences between $A$ and $B$, since we also find triplets in the rope area. Bearing this in mind, and considering that \acs{ARI} is a measure to maximize, we can say that, for cases where there are significant differences, \acs{SHADE}$_{CC}$ performs better than \acs{BRKGA}$_{CC}$. On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} which show the heatmaps obtained for Time and Unsat measurements, are very similar. In both, all triplets are represented in the lower right third of the triangle, indicating that the Bayesian sign test assigns a high probability to $A - B > 0$. With this, and bearing in mind that both measures must be minimized, we conclude that \acs{SHADE}$_{CC}$ represents a solid improvement over \acs{BRKGA}$_{CC}$ as far as this measures are concerned.

Regarding the comparison of \acs{SHADE}$_{CC}$ with the state-of-the-art, we will keep the results obtained by \acs{SHADE}$_{CC}$ as $B$, and note the results obtained by the state-of-the-art methods as $A$.

This way, we apply the process described earlier in this section to obtain the heatmaps corresponding to the comparison of \acs{SHADE}$_{CC}$ with each of the remaining 5 state-of-the-art methods. This time the only measure we compare is \acs{ARI}. Figure \ref{fig:HeatmapsSHADEvsSOTA} show said comparison.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[\acs{COPKM} vs \acs{SHADE}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/COPKMvsSHADE.pdf}\label{fig:COPKMvsSHADE}}
	\subfloat[\acs{LCVQE} vs \acs{SHADE}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/LCVQEvsSHADE.pdf}\label{fig:LCVQEvsSHADE}} \quad
	\subfloat[\acs{RDPM} vs \acs{SHADE}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/RDPMvsSHADE.pdf}\label{fig:RDPMvsSHADE}}
	\subfloat[\acs{TVClust} vs \acs{SHADE}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/TVClustvsSHADE.pdf}\label{fig:TVClustvsSHADE}} \quad
	\subfloat[\acs{CECM} vs \acs{SHADE}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/CEKMvsSHADE.pdf}\label{fig:CECMvsSHADE}}
	\caption[Heatmaps for the comparison of SHADE$_{CC}$ with the state-of-the-art.]{Heatmaps for the comparison of \acs{SHADE}$_{CC}$ with the state-of-the-art.}
	\label{fig:HeatmapsSHADEvsSOTA}
\end{figure}

We see very similar diagrams for \acs{COPKM} (\ref{fig:COPKMvsSHADE}), LVCQE (\ref{fig:LCVQEvsSHADE}), \acs{RDPM} (\ref{fig:RDPMvsSHADE}) and \acs{TVClust} (\ref{fig:TVClustvsSHADE}). In all of them most of the triplets are represented in the lower middle region of the diagram, with a slight shift of the point cloud to the left. That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---, but the advantage is for one or the other depending on the case, being \acs{SHADE}$_{CC}$ the one that gets it in most of them. However, this does not apply in the case of \acs{CECM} (\ref{fig:CECMvsSHADE}), where the advantage is indisputably for \acs{SHADE}$_{CC}$.

\section{Testing DILS$_{CC}$} \label{sec:TestDILS}

In this case we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with \acs{DILS}$_{CC}$ as sample $B$.

\begin{figure}[bth]
	\myfloatalign
	\subfloat[\acs{BRKGA}$_{CC}$ vs \acs{DILS}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsDILS.pdf}\label{fig:BRKGAvsDILS}}
	\subfloat[\acs{COPKM} vs \acs{DILS}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/COPKMvsDILS.pdf}\label{fig:COPKMvsDILS}} \quad
	\subfloat[\acs{LCVQE} vs \acs{DILS}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/LCVQEvsDILS.pdf}\label{fig:LCVQEvsDILS}}
	\subfloat[\acs{RDPM} vs \acs{DILS}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/RDPMvsDILS.pdf}\label{fig:RDPMvsDILS}} \quad
	\subfloat[\acs{TVClust} vs \acs{DILS}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/TVClustvsDILS.pdf}\label{fig:TVClustvsDILS}}
	\subfloat[\acs{CECM} vs \acs{DILS}$_{CC}$]
	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/CEKMvsDILS.pdf}\label{fig:CECMvsDILS}}
	\caption[Heatmaps for the comparison of DILS$_{CC}$ with the state-of-the-art.]{Heatmaps for the comparison of \acs{DILS}$_{CC}$ with the state-of-the-art.}
	\label{fig:HeatmapsDILSvsSOTA}
\end{figure}

Figure \ref{fig:heatDiag} shows all six heatmaps obtained when applying the Bayesian sign test to \acs{DILS}$_{CC}$ and the state-of-the-art methods. Diagrams for \acs{BRKGA}$_{CC}$ (\ref{fig:BRKGAvsDILS}), \acs{LCVQE} (\ref{fig:LCVQEvsDILS}), \acs{RDPM} (\ref{fig:RDPMvsDILS}), \acs{TVClust} (\ref{fig:TVClustvsDILS}) and \acs{CECM} (\ref{fig:CECMvsDILS}) are very similar. They suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle. Bearing this in mind, and considering that \acs{ARI} is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that \acs{DILS}$_{CC}$ does represent a significant improvement over the named 5 methods, being the results offered by the latter non equivalent to those offered by \acs{DILS}$_{CC}$ and lower in quality. In the case of \acs{COPKM} (Figure \ref{fig:BRKGAvsDILS}) we find the cloud of triplets closer to the lower-middle of the diagram, indicating that results offered by \acs{COPKM} and \acs{DILS}$_{CC}$ remain being significantly different but in some cases the advantage would be for \acs{COPKM}. However, in most cases \acs{DILS}$_{CC}$ will offer better results.

\section{SHADE$_{CC}$ or DILS$_{CC}$?} \label{sec:SHADEorDILS}

Regarding the comparison between the two new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---, we have seen how both of them represent an improvement over previous constrained clustering approaches. In Section \ref{sec:NewPropvsSOTA} we saw how \acs{DILS}$_{CC}$ seemed to offer better results than \acs{SHADE}$_{CC}$, but it is worth empirically analyze this conclusions through the Bayesian sign test. We will refer to the results obtained \acs{SHADE}$_{CC}$ $A$, and to the results obtained with \acs{DILS}$_{CC}$ as sample $B$. Figure \ref{fig:SHADEvsDILS} shows the heatmap obtained with this setup.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.6]{gfx/AnalysisResults/SHADEvsDILS.pdf}
	\caption[Heatmap for the comparison of SHADE$_{CC}$ with DILS$_{CC}$]{Heatmap for the comparison of \acs{SHADE}$_{CC}$ with \acs{DILS}$_{CC}$}\label{fig:SHADEvsDILS}
\end{figure}

As we could expect, the test assigns a high probability to $A - B < 0$, being the most of the triplets represented in the lower-left third of the triangle. This indicates that there are significant differences between both methods in most cases, and the test gives a clear lead to \acs{DILS}$_{CC}$. We conclude that the reason behind this is that a real-based problem representation does not provide any other advantage that the possibility to apply real-coding evolutionary methods. This way, given that the \acs{SHADE} method proved to be the best approach to real optimization problems we expected it to outperform \acs{BRKGA}, though \acs{SHADE} imposes an arithmetic relationship between individuals---through the mutation operator---which is hard to understand in the clustering domain. This is, applying an arithmetic operator to two labels does not provide easily understandable combination of them. Judging from the raw results and the empirical analysis, and with its straight integer-based problem representation and its strong exploitation capability \acs{DILS}$_{CC}$ stands out as the best of all methods we obtained results for.

### 2019-06-29 Memoria/Chapters/ExperimentalResults.tex
Escrito analisis de resultados en memoria

@@
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMayor.pdf}}
-	\caption{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
+	\caption[Visual representation of the influence of $\xi$ over the DILS$_{CC}$ optimization process.]{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
 	\label{fig:RestartsLines}
@@
 
-We can conclude that the parameter $\xi$ has the expected influence over the \acs{DILS}$_{CC}$ optimization process.
+We can conclude that parameter $\xi$ has the expected influence over the \acs{DILS}$_{CC}$ optimization process.
 Note that in a problem in which the \acs{LS} procedure used is deterministic we would also find differences between consecutive runs over the same datasets, given that the \acs{DILS} general optimization scheme uses stochastic tools---such as the recombination and mutation operators---to make sure that diversity is introduced in the search of the global optima for the problem we apply it to.
@@
 
-Regarding the comparison between the two new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---, we have seen how both of them represent an improvement over previous constrained clustering approaches, being \acs{DILS}$_{CC}$ the best of the two.
-This is due to the fact that a real-based problem representation does not provide any other advantage that the possibility to apply real-coding evolutionary methods.
-This way, given that the \acs{SHADE} method proved to be the best approach to real optimization problems we expected it to outperform \acs{BRKGA}, though \acs{SHADE} imposes an arithmetic relationship between individuals---through the mutation operator---which is hard to understand in the clustering domain.
-This is, applying an arithmetic operator to two labels does not provide easily understandable combination of them.
-Judging from the raw results, and with its straight integer-based problem representation and its strong exploitation capability \acs{DILS}$_{CC}$ stands out as the best of all methods we obtained results for.
+\newpage
+
+All comparisons carried out in this sections are merely supported by the raw results.
+To prove whether the new proposals represent an improvement over the state-of-the-art methods it is necessary to perform an statistic-based empirical analysis of the results.
+The last chapter (Chapter \ref{ch:AnalysisResults}) of Part \ref{pt:Experiments} will address this question.

### 2019-06-29 Memoria/Chapters/Introduction.tex
Escrito analisis de resultados en memoria

@@
 As we have already mentioned, it is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
-We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}.
+We can divide \acs{SSL} methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}.
 The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
@@
 
-\section{Objetivos}
-
-\section{Estructura}
+\section{Structure}

### 2019-06-29 Memoria/Chapters/NewProposals.tex
Correcciones y nuevas secciones
Escrito analisis de resultados en memoria

@@
 As we have already mentioned in Section \ref{sec:FeasibilityProblem}, the incorporation of constraints may modify the complexity of the clustering problem, making it $\mathbf{NP}$-complete and therefore intractable \cite{davidson2005clustering}.
-Bearing this in mind, in this chapter we propose two novel approaches to find approximate solutions to the constrained clustering problem.
+Bearing this in mind, in this chapter we propose two new approaches to find approximate solutions to the constrained clustering problem.
 In Section \ref{sec:CCSHADE} a \acf{DE} based approach is presented, while in Section \ref{sec:DILS} a new \acf{LS} variant is discussed, as well as its application to constrained clustering.
@@
 
-\section{Brief Review of the BRKGA Algorithm} \label{sec:brkga}
+\subsection{Brief Review of the BRKGA Algorithm} \label{sec:brkga}
 
@@
 	\centering
-	\includegraphics[scale=0.4]{gfx/NewProp/SHADE/BRKGA_Diagram.jpg} 
+	\includegraphics[scale=0.31]{gfx/NewProp/SHADE/BRKGA_Diagram.jpg} 
 	\caption[BRKGA$_{CC}$ optimization process for constrained clustering.]{\acs{BRKGA}$_{CC}$ optimization process for constrained clustering. \cite{de2017comparison}.}\label{fig:BRKGA_CC}
@@
 
-\section{SHADE optimization scheme for constrained clustering} \label{sec:SHADEadapt}
+\subsection{SHADE optimization scheme for constrained clustering} \label{sec:SHADEadapt}
 
@@
 	\myfloatalign
-	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/Dataset}} \quad
+	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/Dataset}} \quad
 	%\subfloat[Imagen original]
-	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C0}} \quad
+	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C0}} \quad
 	%\subfloat[Clustering sin restricciones]
-	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C1}} \quad
+	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C1}} \quad
 	%\subfloat[Clustering con restricciones]
-	{\includegraphics[width=.45\linewidth]{gfx/NewProp/SHADE/C2}} \quad
+	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C2}} \quad
 	\caption[Three partitions over a toy dataset.]{Three partitions over a toy dataset.}
@@
 	\centering
-	\includegraphics[scale=0.4]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
+	\includegraphics[scale=0.31]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
 	\caption[SHADE$_{CC}$ optimization process for constrained clustering.]{\acs{SHADE}$_{CC}$ optimization process for constrained clustering.}\label{fig:SHADE_CC}
@@
 
+\newpage
+
 \begin{algorithm}
@@
 
-	\caption{Local Search}\label{alg:SHADE_LS}
+	\caption[SHADE$_{CC}$ Local Search]{\acs{SHADE}$_{CC}$ Local Search}\label{alg:SHADE_LS}
 \end{algorithm}
@@
 
-\subsection{Short Introduction to Iterated Local Search}
+\subsection{A Glimpse to Iterated Local Search}
 
@@
 
-\section{The Dual Iterative Local Search Method}
+\subsection{The Dual Iterative Local Search Method}
 
@@
 While \acs{ILS} works with a single individual, \acs{DILS} keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima.
-The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value.
+The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value at the end of each stage (generation) of the optimization process.
 
@@
 This is the individual \acs{DILS} applies the \acs{LS} procedure to in order to improve its fitness value, resulting in the trial individual $m_t$.
-The trial individual $m_t$ replaces the worst of its parents, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for \acs{DILS}.
+The trial individual $m_t$ replaces the worst of its predecessors, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for \acs{DILS}.
 Equation \ref{eq2.1} displays the expression for this criterion in the case of a minimization problem.
@@
 \noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
-This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best parent.
+This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best predecessor.
 Figure \ref{img:DILS_diag} summarizes the \acs{DILS} optimization process.
@@
 	\centering
-	\includegraphics[scale=0.25]{gfx/NewProp/DILS/DILSDiagram.jpg}
+	\includegraphics[scale=0.27]{gfx/NewProp/DILS/DILSDiagram.jpg}
 	\caption{Diagram summarizing \acs{DILS} optimization process.}\label{img:DILS_diag}
@@
 Algorithm \ref{alg:DILS} summarizes the overall \acs{DILS} optimization process.
+
+\newpage
 
@@
 
-\section{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}
+\subsection{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}
 
-In this section, we present an optimization scheme for the \acs{DILS} algorithm when applied to constrained clustering.
+In this section, we present the application scheme of \acs{DILS} to constrained clustering.
 We will call this new approach \acs{DILS}$_{CC}$.
@@
 
-	\caption{Local Search}\label{alg:DISL_LS}
+	\caption[DILS$_{CC}$ Local Search]{\acs{DILS}$_{CC}$ Local Search}\label{alg:DISL_LS}
 \end{algorithm}

### 2019-06-29 Memoria/FrontBackmatter/Colophon.tex
Escrito analisis de resultados en memoria

\pagestyle{empty}

\hfill

\vfill

\pdfbookmark[0]{Colophon}{colophon}
\section*{Colophon}
This document was typeset using the typographical look-and-feel \texttt{classicthesis} developed by Andr\'e Miede and Ivo Pletikosić.
The style was inspired by Robert Bringhurst's seminal book on typography ``\emph{The Elements of Typographic Style}''.
\texttt{classicthesis} is available for both \LaTeX\ and \mLyX:
\begin{center}
\url{https://bitbucket.org/amiede/classicthesis/}
\end{center}

\bigskip

\noindent\finalVersionString

### 2019-06-29 Memoria/FrontBackmatter/Declaration.tex
Escrito analisis de resultados en memoria

@@
 \pdfbookmark[0]{Declaration}{declaration}
-\chapter*{Declaration}
+\chapter*{Declaración de Originalidad}
 \thispagestyle{empty}
-\textbf{JAVIER LEÓN PALOMARES} con DNI $77138857$P, actuando en su propio nombre y derecho, \\
-
-\textbf{DECLARA, BAJO SU RESPONSABILIDAD:} \\
-
-Que el Trabajo Fin de Máster presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con fecha \_\_ de julio de 2019, titulado <<\textit{\_\_}>> es original, no es copia ni adaptación de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona.
-\\
-
-Y para que así conste a los efectos oportunos, firmo la presente en Granada, a \_\_ de julio de 2019.
 
@@
 
-%\noindent\textit{\myLocation, \myTime}
+\noindent \textbf{D. GERMÁN GONZÁLEZ ALMAGRO} con DNI $76593910$T, actuando en su propio nombre y derecho,\\
+
+\noindent \textbf{DECLARA, BAJO SU RESPONSABILIDAD}:\\
+
+\noindent Que el Trabajo Fin de Grado presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicaciones de la Universidad de  Granada, con fecha 17 de Junio de 2018, titulado <<\textit{Constrained Clustering: New Proposals}>> es original, no es copia de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona.\\
+
+\noindent Y para que así conste a los efectos oportunos, firmo la presente en Granada, a 17 de junio de 2018.
+
+\vspace{3cm}
 
@@
 \begin{flushright}
-    \begin{tabular}{m{5cm}}
+    \begin{tabular}{m{6cm}}
         \\ \hline
-        \centering\myName \\
+        \centering\textbf{Fdo. \myName} \\
     \end{tabular}
 \end{flushright}
-
-Asimismo, yo, Javier León Palomares, alumno del Máster en Ciencia de Datos e Ingeniería de Computadores de la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con DNI $77138857$P, autorizo la ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del centro para que pueda ser consultada por las personas que lo deseen.
-
-\begin{flushright}
-    \begin{tabular}{m{5cm}}
-        \\ \hline
-        \centering\myName \\
-    \end{tabular}
-\end{flushright}

### 2019-06-29 Memoria/FrontBackmatter/Declaration2.tex
Escrito analisis de resultados en memoria

%*******************************************************
% Declaration
%*******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Declaration}{declaration}
\chapter*{Declaration}
\thispagestyle{empty}
\textbf{GERMÁN GONZÁLEZ ALMAGRO} con DNI $76593910$T, actuando en su propio nombre y derecho, \\

\textbf{DECLARA, BAJO SU RESPONSABILIDAD:} \\

Que el Trabajo Fin de Máster presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con fecha \_\_ de julio de 2019, titulado <<\textit{\_\_}>> es original, no es copia ni adaptación de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona. \\

Y para que así conste a los efectos oportunos, firmo la presente en Granada, a \_\_ de julio de 2019.

\bigskip

%\noindent\textit{\myLocation, \myTime}

\smallskip

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

Asimismo, yo, Germán González Almagro, alumno del Máster en Ciencia de Datos e Ingeniería de Computadores de la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicación de la Universidad de Granada, con DNI con DNI $76593910$T, autorizo la ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del centro para que pueda ser consultada por las personas que lo deseen.

\begin{flushright}
    \begin{tabular}{m{5cm}}
        \\ \hline
        \centering\myName \\
    \end{tabular}
\end{flushright}

### 2019-06-29 Memoria/FrontBackmatter/Legality.tex
Escrito analisis de resultados en memoria

@@
 de Ingenierías Informática y de Telecomunicación de la Universidad de Granada}, con DNI $\mathrm{76593910}$T, autorizo la
-ubicación de la siguiente copia de mi Trabajo Fin de Grado en la biblioteca del Centro para que pueda ser
+ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del Centro para que pueda ser
 consultada por las personas que lo deseen.
@@
 
-Que el presente trabajo, titulado \textit{\textbf{<<Clustering con Restricciones: un marco unificado>>}},
-ha sido realizado bajo su supervisión por \textbf{Germán González Almagro}, y autorizamos la defensa de dicho trabajo ante el tribunal
+Que el presente trabajo, titulado \textit{\textbf{<<Constrained Clustering: New Proposals>>}},
+ha sido realizado bajo su supervisión por \textbf{Germán González Almagro}, y autorizo la defensa de dicho trabajo ante el tribunal
 que corresponda.
@@
 
-Y para que conste, expiden y firman el presente informe en Granada a 17 de junio de 2018.
+Y para que conste, expido y firmo el presente informe en Granada a 17 de junio de 2018.
 

### 2019-06-30 Memoria/Chapters/AnalysisResults.tex
Acabada memoria excepto conclusiones y trabajo futuro

@@
 	\subfloat[\acs{ARI}]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_ARI.pdf}\label{fig:bayesARI}} \quad
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_ARI.pdf}\label{fig:bayesARI}} \quad
 	\subfloat[Time]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_TIME.pdf}\label{fig:bayesTime}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_TIME.pdf}\label{fig:bayesTime}}
 	\subfloat[Unsat]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_UNSAT.pdf}\label{fig:bayesUnsat}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_UNSAT.pdf}\label{fig:bayesUnsat}}
 	\caption[Heatmaps for the comparison of SHADE$_{CC}$ with BRKGA$_{CC}$.]{Heatmaps for the comparison of \acs{SHADE}$_{CC}$ with \acs{BRKGA}$_{CC}$.}
@@
 	\subfloat[\acs{COPKM} vs \acs{SHADE}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/COPKMvsSHADE.pdf}\label{fig:COPKMvsSHADE}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/COPKMvsSHADE.pdf}\label{fig:COPKMvsSHADE}}
 	\subfloat[\acs{LCVQE} vs \acs{SHADE}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/LCVQEvsSHADE.pdf}\label{fig:LCVQEvsSHADE}} \quad
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/LCVQEvsSHADE.pdf}\label{fig:LCVQEvsSHADE}} \quad
 	\subfloat[\acs{RDPM} vs \acs{SHADE}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/RDPMvsSHADE.pdf}\label{fig:RDPMvsSHADE}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/RDPMvsSHADE.pdf}\label{fig:RDPMvsSHADE}}
 	\subfloat[\acs{TVClust} vs \acs{SHADE}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/TVClustvsSHADE.pdf}\label{fig:TVClustvsSHADE}} \quad
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/TVClustvsSHADE.pdf}\label{fig:TVClustvsSHADE}} \quad
 	\subfloat[\acs{CECM} vs \acs{SHADE}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/CEKMvsSHADE.pdf}\label{fig:CECMvsSHADE}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/CEKMvsSHADE.pdf}\label{fig:CECMvsSHADE}}
 	\caption[Heatmaps for the comparison of SHADE$_{CC}$ with the state-of-the-art.]{Heatmaps for the comparison of \acs{SHADE}$_{CC}$ with the state-of-the-art.}
@@
 	\subfloat[\acs{BRKGA}$_{CC}$ vs \acs{DILS}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/BRKGAvsDILS.pdf}\label{fig:BRKGAvsDILS}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsDILS.pdf}\label{fig:BRKGAvsDILS}}
 	\subfloat[\acs{COPKM} vs \acs{DILS}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/COPKMvsDILS.pdf}\label{fig:COPKMvsDILS}} \quad
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/COPKMvsDILS.pdf}\label{fig:COPKMvsDILS}} \quad
 	\subfloat[\acs{LCVQE} vs \acs{DILS}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/LCVQEvsDILS.pdf}\label{fig:LCVQEvsDILS}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/LCVQEvsDILS.pdf}\label{fig:LCVQEvsDILS}}
 	\subfloat[\acs{RDPM} vs \acs{DILS}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/RDPMvsDILS.pdf}\label{fig:RDPMvsDILS}} \quad
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/RDPMvsDILS.pdf}\label{fig:RDPMvsDILS}} \quad
 	\subfloat[\acs{TVClust} vs \acs{DILS}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/TVClustvsDILS.pdf}\label{fig:TVClustvsDILS}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/TVClustvsDILS.pdf}\label{fig:TVClustvsDILS}}
 	\subfloat[\acs{CECM} vs \acs{DILS}$_{CC}$]
-	{\includegraphics[width=0.45\linewidth]{gfx/AnalysisResults/CEKMvsDILS.pdf}\label{fig:CECMvsDILS}}
+	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/CEKMvsDILS.pdf}\label{fig:CECMvsDILS}}
 	\caption[Heatmaps for the comparison of DILS$_{CC}$ with the state-of-the-art.]{Heatmaps for the comparison of \acs{DILS}$_{CC}$ with the state-of-the-art.}
@@
 
-Figure \ref{fig:heatDiag} shows all six heatmaps obtained when applying the Bayesian sign test to \acs{DILS}$_{CC}$ and the state-of-the-art methods.
+Figure \ref{fig:HeatmapsDILSvsSOTA} shows all six heatmaps obtained when applying the Bayesian sign test to \acs{DILS}$_{CC}$ and the state-of-the-art methods.
 Diagrams for \acs{BRKGA}$_{CC}$ (\ref{fig:BRKGAvsDILS}), \acs{LCVQE} (\ref{fig:LCVQEvsDILS}), \acs{RDPM} (\ref{fig:RDPMvsDILS}), \acs{TVClust} (\ref{fig:TVClustvsDILS}) and \acs{CECM} (\ref{fig:CECMvsDILS}) are very similar.

### 2019-06-30 Memoria/Chapters/ConclusionsAndFutureWork.tex
Acabada memoria excepto conclusiones y trabajo futuro

\chapter{Conclusions and Future Work}\label{ch:ConclFW}

### 2019-06-30 Memoria/Chapters/ExperimentalResults.tex
Acabada memoria excepto conclusiones y trabajo futuro

@@
 This chapter presents the results obtained by the two new proposals introduced in this work (in Chapter \ref{ch:NewProposals}) and those obtained by the state-of-the-art.
-Being the mentioned its main purpose, we firstly introduce a brief study on the influence of the parameter $\xi$ in the \acs{DILS} method in Section \ref{sec:XiInfl}, so the reader gets a better comprehension of theory presented in Section \ref{sec:DILS}.
+Being the mentioned the main purpose of this chapter, we firstly introduce a brief study on the influence of the parameter $\xi$ in the \acs{DILS} method in Section \ref{sec:XiInfl}, so the reader gets a better comprehension of theory presented in Section \ref{sec:DILS}.
 Afterwards we will carry out an in-depth comparison between \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$ in Section \ref{sec:BRKGAvsSHADE}, as both are population-based and heuristic-guided methods.
@@
 			\hline
-			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
+			Appendicitis & \textbf{1.000} & 0.000 & 389.905 && \textbf{1.000} & 0.000 & 371.579 \\
 			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
@@
 			\hline
-			Appendicitis & 0.086 & 0.611 & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
-			Breast Cancer & 0.159 & 0.755 & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.104 & 0.870 & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
-			Circles & 0.063 & 0.798 & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
-			Ecoli & 0.020 & 0.069 & 0.009 & - & 0.387 & 0.417 & 0.265 & - \\
-			Glass & 0.013 & 0.009 & 0.016 & 0.184 & 0.268 & 0.197 & 0.211 & - \\
-			Haberman & 0.084 & 0.638 & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
-			Hayesroth & 0.025 & 0.031 & 0.027 & - & 0.106 & 0.097 & 0.054 & 0.139 \\
-			Ionosphere & 0.332 & 0.792 & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
-			Iris & 0.342 & 0.621 & 0.238 & -0.285 & 0.730 & 0.607 & 0.244 & 0.684 \\
-			Monk2 & 0.270 & 0.815 & 0.421 & 0.982 & 0.072 & 0.094 & -0.002 & 0.007 \\
-			Moons & 0.272 & 0.968 & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
-			Saheart & 0.265 & 0.800 & 0.147 & 0.974 & 0.018 & 0.020 & 0.068 & 0.000 \\
-			Sonar & 0.116 & 0.649 & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
-			Soybean & 0.515 & 0.304 & 0.496 & 0.503 & 0.545 & 0.621 & 0.000 & 0.244 \\
-			Spectfheart & 0.343 & 0.896 & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
-			Spiral & 0.114 & 0.856 & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
-			Tae & 0.024 & 0.031 & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
-			Vowel & 0.002 & 0.002 & 0.003 & - & 0.063 & -0.003 & 0.067 & - \\
-			Zoo & 0.098 & 0.151 & 0.105 & 0.715 & 0.666 & 0.412 & 0.335 & - \\
+			Appendicitis & 0.086 & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
+			Breast Cancer & 0.159 & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.104 & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
+			Circles & 0.063 & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
+			Ecoli & 0.020 & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
+			Glass & 0.013 & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
+			Haberman & 0.084 & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
+			Hayesroth & 0.025 & 0.031 & 0.027 & - & 0.106 & 0.097 & 0.054 & \textbf{0.139} \\
+			Ionosphere & 0.332 & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
+			Iris & 0.342 & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
+			Monk2 & 0.270 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
+			Moons & 0.272 & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
+			Saheart & 0.265 & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
+			Sonar & 0.116 & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
+			Soybean & 0.515 & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
+			Spectfheart & 0.343 & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
+			Spiral & 0.114 & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
+			Tae & 0.024 & 0.031 & 0.018 & - & 0.009 & -0.000 & \textbf{0.067} & 0.000 \\
+			Vowel & 0.002 & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
+			Zoo & 0.098 & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
 			\hline
@@
 			\hline
-			Appendicitis & 0.332 & 0.957 & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
-			Breast Cancer & 0.911 & 0.788 & 0.772 & 1.000 & 0.486 & 0.502 & 0.000 & 0.037 \\
-			Bupa & 0.801 & 0.984 & 0.840 & 1.000 & -0.005 & -0.007 & -0.007 & 0.001 \\
-			Circles & 0.982 & 1.000 & 0.828 & 1.000 & -0.003 & 0.375 & 0.973 & 0.105 \\
-			Ecoli & 0.025 & 0.134 & 0.015 & - & 0.387 & 0.372 & 0.719 & - \\
-			Glass & 0.031 & 0.057 & 0.017 & - & 0.280 & 0.253 & 0.229 & - \\
-			Haberman & 0.796 & 1.000 & 0.866 & 1.000 & -0.002 & 0.075 & 0.219 & -0.050 \\
-			Hayesroth & 0.099 & 0.386 & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
-			Ionosphere & 0.591 & 0.970 & 0.803 & 1.000 & 0.178 & 0.212 & 0.000 & 0.057 \\
-			Iris & 0.260 & 0.850 & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
-			Monk2 & 0.945 & 0.902 & 0.751 & 1.000 & 0.072 & 0.170 & -0.002 & 0.007 \\
-			Moons & 0.995 & 1.000 & 0.958 & 1.000 & 0.241 & 0.436 & 0.987 & 0.095 \\
-			Saheart & 0.945 & 0.862 & 0.749 & 1.000 & 0.020 & 0.037 & 0.221 & 0.000 \\
-			Sonar & 0.890 & 0.981 & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
-			Soybean & 0.422 & 0.469 & 0.428 & 0.584 & 0.545 & 0.605 & 0.000 & 0.000 \\
-			Spectfheart & 0.778 & 1.000 & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
-			Spiral & 0.949 & 1.000 & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
-			Tae & 0.066 & 0.299 & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
-			Vowel & 0.001 & 0.003 & 0.002 & - & 0.063 & -0.003 & 0.073 & - \\
-			Zoo & 0.128 & 0.185 & 0.157 & 0.435 & 0.642 & 0.450 & 0.353 & - \\
+			Appendicitis & 0.332 & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
+			Breast Cancer & 0.911 & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
+			Bupa & 0.801 & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
+			Circles & 0.982 & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
+			Ecoli & 0.025 & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
+			Glass & 0.031 & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
+			Haberman & 0.796 & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
+			Hayesroth & 0.099 & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
+			Ionosphere & 0.591 & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
+			Iris & 0.260 & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
+			Monk2 & 0.945 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
+			Moons & 0.995 & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
+			Saheart & 0.945 & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
+			Sonar & 0.890 & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
+			Soybean & 0.422 & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
+			Spectfheart & 0.778 & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
+			Spiral & 0.949 & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
+			Tae & 0.066 & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
+			Vowel & 0.001 & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
+			Zoo & 0.128 & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
 			\hline
@@
 			\hline
-			Appendicitis & 1.000 & 1.000 & 1.000 & - & 0.305 & 0.331 & 0.012 & -0.006 \\
-			Breast Cancer & 0.977 & 0.807 & 0.807 & 1.000 & 0.486 & 0.502 & 0.000 & 0.000 \\
-			Bupa & 0.993 & 0.991 & 0.936 & 1.000 & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & 1.000 & 1.000 & 0.935 & 1.000 & -0.003 & 0.629 & 1.000 & 0.138 \\
-			Ecoli & 0.047 & 0.201 & 0.028 & - & 0.387 & 0.459 & 0.763 & - \\
-			Glass & 0.082 & 0.322 & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
-			Haberman & 0.997 & 1.000 & 0.943 & 1.000 & -0.002 & 0.106 & 0.737 & 0.000 \\
-			Hayesroth & 0.531 & 0.812 & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Ionosphere & 0.993 & 0.982 & 0.902 & 1.000 & 0.173 & 0.309 & 0.000 & 0.115 \\
-			Iris & 0.553 & 0.968 & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
-			Monk2 & 0.991 & 0.904 & 0.855 & 1.000 & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & 0.995 & 1.000 & 0.937 & 1.000 & 0.241 & 0.831 & 1.000 & 0.180 \\
-			Saheart & 0.981 & 0.881 & 0.881 & 1.000 & 0.018 & 0.026 & 1.000 & -0.006 \\
-			Sonar & 1.000 & 1.000 & 0.992 & 1.000 & 0.004 & 0.127 & 0.000 & 0.003 \\
-			Soybean & 0.563 & 0.564 & 0.416 & -0.218 & 0.545 & 0.631 & 0.000 & -0.016 \\
-			Spectfheart & 1.000 & 1.000 & 0.953 & 1.000 & -0.107 & -0.112 & 0.000 & -0.054 \\
-			Spiral & 0.998 & 1.000 & 0.921 & 1.000 & -0.003 & 0.011 & 0.006 & 0.045 \\
-			Tae & 0.392 & 0.892 & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
-			Vowel & 0.003 & 0.003 & 0.002 & - & 0.063 & -0.003 & 0.071 & - \\
-			Zoo & 0.153 & 0.203 & 0.176 & 0.821 & 0.642 & 0.439 & 0.335 & - \\
+			Appendicitis & \textbf{1.000} & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
+			Breast Cancer & 0.977 & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
+			Bupa & 0.993 & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
+			Circles & \textbf{1.000} & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & \textbf{1.000} & 0.138 \\
+			Ecoli & 0.047 & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
+			Glass & 0.082 & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
+			Haberman & 0.997 & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
+			Hayesroth & 0.531 & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
+			Ionosphere & 0.993 & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
+			Iris & 0.553 & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
+			Monk2 & 0.991 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
+			Moons & 0.995 & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & \textbf{1.000} & 0.180 \\
+			Saheart & 0.981 & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
+			Sonar & \textbf{1.000} & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
+			Soybean & 0.563 & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
+			Spectfheart & \textbf{1.000} & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
+			Spiral & 0.998 & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
+			Tae & 0.392 & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
+			Vowel & 0.003 & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
+			Zoo & 0.153 & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
 			\hline

### 2019-06-30 Memoria/Chapters/ExperimentalSetup.tex
Acabada memoria excepto conclusiones y trabajo futuro

@@
 \begin{equation}
-\text{Rand}(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
+\text{Rand}(C_1, C_2) = \frac{a + b}{n(n - 1)/2}.
 \label{eq15}
@@
 \begin{equation}
-\text{ARI}(C_1, C_2) = \frac{\text{Rand}(C_1, C_2) - \text{Expected Index}}{\text{Maximum Index} - \text{Expected Index}},
+\text{ARI}(C_1, C_2) = \frac{\text{Rand}(C_1, C_2) - \text{Expected Index}}{\text{Maximum Index} - \text{Expected Index}}\;\;,
 \label{eq16}
@@
 
-In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic \acf{NHST}.
+In order to validate the results which will be presented in Chapter \ref{ch:ExperimentalResults} we will use Bayesian statistical tests, instead of the classic \acf{NHST}.
 In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of \acs{NHST}, and a new model is proposed for carrying out comparisons researchers are interested in.
@@
 Table \ref{tab:paramsDILS} shows a summary of the parameter setup used for the \acs{DILS}$_{CC}$ method.
+All parameters related to the new proposals---\acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$---have been set experimentally, no theoretical or in-depth experimental analysis have been carried out to find an optimal tune for them.
 
@@
 
-In all three cases---\acs{BRKGA}$_{CC}$, \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$---the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
+In all three cases---\acs{BRKGA}$_{CC}$, \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$---the stopping criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
 To compare with the state-of-the-art methods mentioned in Section \ref{sec:SOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA}.
@@
 Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
-All parameters described above are ment to be used for the implementation that can be found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}.
+All parameters described above are meant to be used for the Python implementation that can be found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}, which will be used to obtain our results.

### 2019-06-30 Memoria/Chapters/Introduction.tex
Acabada memoria excepto conclusiones y trabajo futuro

@@
 
+\section{Objectives}
+
+The main objective of this work is to provide an overview of the concepts and particularities related to constrained clustering.
+Some of these concepts must be addressed with sufficient depth, in order for the reader to be able to easily understand the new approaches to constrained clustering contained in this work.
+It is also of our interest to provide an example of proper experimental methodology to obtain relevant and objectivity-free conclusions.
+
 \section{Structure}
+
+Regarding the organization of this work, background knowledge concerning classic clustering and constrained clustering will be presented in Chapters \ref{ch:IntroClustering} and \ref{ch:ConstrainedClustering} respectively.
+In Chapter \ref{ch:NewProposals} two new approaches to the constrained clustering problem are discussed.
+The experimental setup to carry out the experiments to test them and their results are presented in Chapters \ref{ch:ExperimentalSetup} and \ref{ch:ExperimentalResults} respectively.
+Afterward an empirical analysis of these results is performed in Chapter \ref{ch:AnalysisResults}.
+Finally conclusions and future work are discussed in Chapter \ref{ch:ConclFW}.

### 2019-06-30 Memoria/FrontBackmatter/Abstract.tex
Acabada memoria excepto conclusiones y trabajo futuro

@@
 
+Clustering has always been a powerful tool in knowledge discovery.
+Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering.
+This technique is a generalization of traditional clustering that considers additional information encoded by constraints.
+Constraints can be given in the form of instance-level must-link and cannot-link constraints, which this paper focuses on.
+We propose the first application of Differential Evolution to the constrained clustering problem, as well as a new Local Search variant and its application to the problem at hands.
+We will compare the results obtained by these new proposals to those obtained by the state-of-the-art algorithms on twenty datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.\\
+
+\noindent\textbf{Keywords:} constrained clustering, instance-level, must-link, cannot-link, differential evolution, local search.
+
 \endgroup

### 2019-06-30 Memoria/FrontBackmatter/Resumen.tex
Acabada memoria excepto conclusiones y trabajo futuro

@@
 
+El clustering siempre ha sido una potente herramienta en tareas de descubrimiento de conocimiento.
+Tradicionalmente en el marco del apredizaje no supervisado, recibió atención renovada cuando se le incorporaron nuevos tipos de información, dando lugar a un nuevo tipo de aprendizaje semisupervisado: el clustering con restricciones.
+Esta técnica es una generalicación del clustering clásico para considerar información adicional, dada en forma de restricciones.
+Estas pueden venir dadas en forma de restricciones a nivel de instancia tipo \textit{must-link} y \textit{cannot-link}, en las que este trabajo se centra.
+Proponemos la primera aplicación de Differential Evolution al clustering con restricciones, así como una nueva variante de la Búsqueda Local, tanto su esquema de optimización general como su aplicación al clustering con restricciones.
+Compararemos los resultados obtenidos por estas nuevas propuestas con los obtenidos por el estado del arte sobre veinte conjuntos de datos con niveles incrementales de infomación basada e restricciones, apoyando nuestras conclusiones con la ayuda de los test Bayesianos.
+\\
+
+\noindent\textbf{Keywords:} Clustering con restricciones, restricciones a nivel de instancia, \textit{must-link}, \textit{cannot-link}, \textit{differential evolution}, búsqueda local.
+
 \endgroup

### 2019-07-06 Memoria/Chapters/AnalysisResults.tex
Cambios en memoria y en paper DILS

@@
 
-With the results obtained by all methods for a total of 60 different datasets---the 20 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them.
-This way we can statistically determine whether the new proposals---\acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$---represent a significant improvement over previous proposals.
+With the results obtained by all methods for a total of sixty different datasets---the twenty datasets in combination with the three constraint sets for each one of them---we can perform an empirical analysis of them.
+This way, we can statistically determine whether the new proposals {SHADE}$_{CC}$ and \acs{DILS}$_{CC}$ represent a significant improvement over previous proposals.
+To this end, we make use of Bayesian sign tests.
 
-One of the major advantages of the Bayesian sign test---introduced in Section \ref{sec:ValidtnMethod}---is that we can obtain a very illustrative visual representation of its results.
+One of the major advantages of the Bayesian sign test (introduced in Section \ref{sec:ValidtnMethod}) is that we can obtain a very illustrative visual representation of its results.
 We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates.
@@
 
-\section{Testing SHADE$_{CC}$} \label{sec:TestSHADE}
+\section[Testing \acsfont{SHADE}$_{CC}$]{Testing SHADE$_{CC}$} \label{sec:TestSHADE}
 
-In this section we will determine if there are significant differences between \acs{SHADE}$_{CC}$ and all 6 state-of-the-art-method.
-To do so we will perform an in-depth statistical comparison between \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$, as well as a more general comparison concerning the state-of-the-art methods.
+In this section we determine if there are significant differences between \acs{SHADE}$_{CC}$ and all six state-of-the-art-method.
+To do so we perform an in-depth statistical comparison between \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$, as well as a more general comparison concerning the state-of-the-art methods.
 
-Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by \acs{BRKGA}$_{CC}$ as sample $A$, and to the results obtained with \acs{SHADE}$_{CC}$ as sample $B$.
+Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we refer to the results obtained by \acs{BRKGA}$_{CC}$ as sample $A$, and to the results obtained with \acs{SHADE}$_{CC}$ as sample $B$.
 
@@
 	\myfloatalign
-	\subfloat[\acs{ARI}]
+	\subfloat[\acs{ARI}.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_ARI.pdf}\label{fig:bayesARI}} \quad
-	\subfloat[Time]
+	\subfloat[Time.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_TIME.pdf}\label{fig:bayesTime}}
-	\subfloat[Unsat]
+	\subfloat[Unsat.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsSHADE_UNSAT.pdf}\label{fig:bayesUnsat}}
-	\caption[Heatmaps for the comparison of SHADE$_{CC}$ with BRKGA$_{CC}$.]{Heatmaps for the comparison of \acs{SHADE}$_{CC}$ with \acs{BRKGA}$_{CC}$.}
+	\caption{Heatmaps comparing \acs{SHADE}$_{CC}$ with \acs{BRKGA}$_{CC}$.}
 	\label{fig:HeatmapsSHADEvsBRKGA}
@@
 
-In one hand, figure \ref{fig:bayesARI} shows the heatmap associated with the \acs{ARI} measurement.
+On one hand, Figure \ref{fig:bayesARI} shows the heatmap associated with the \acs{ARI} measurement.
 The diagram suggests that the Bayesian sign test assigns a high probability to $A - B < 0$, since most triplets are represented in the lower left third of the triangle.
@@
 Bearing this in mind, and considering that \acs{ARI} is a measure to maximize, we can say that, for cases where there are significant differences, \acs{SHADE}$_{CC}$ performs better than \acs{BRKGA}$_{CC}$.
-On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat} which show the heatmaps obtained for Time and Unsat measurements, are very similar.
+On the other hand, Figures \ref{fig:bayesTime} and \ref{fig:bayesUnsat}, which show the heatmaps obtained for Time and Unsat measurements, are very similar.
 In both, all triplets are represented in the lower right third of the triangle, indicating that the Bayesian sign test assigns a high probability to $A - B > 0$.
-With this, and bearing in mind that both measures must be minimized, we conclude that \acs{SHADE}$_{CC}$ represents a solid improvement over \acs{BRKGA}$_{CC}$ as far as this measures are concerned.
+With this, and bearing in mind that both measures must be minimized, we conclude that \acs{SHADE}$_{CC}$ represents a solid improvement over \acs{BRKGA}$_{CC}$ as far as these measures are concerned.
 
-Regarding the comparison of \acs{SHADE}$_{CC}$ with the state-of-the-art, we will keep the results obtained by \acs{SHADE}$_{CC}$ as $B$, and note the results obtained by the state-of-the-art methods as $A$.
+Regarding the comparison of \acs{SHADE}$_{CC}$ with the state-of-the-art, we keep the results obtained by \acs{SHADE}$_{CC}$ as $B$, and note the results obtained by the state-of-the-art methods as $A$.
 
-This way, we apply the process described earlier in this section to obtain the heatmaps corresponding to the comparison of \acs{SHADE}$_{CC}$ with each of the remaining 5 state-of-the-art methods.
+This way, we apply the process described earlier in this chapter to obtain the heatmaps corresponding to the comparison of \acs{SHADE}$_{CC}$ with each of the remaining five state-of-the-art methods.
 This time the only measure we compare is \acs{ARI}.
-Figure \ref{fig:HeatmapsSHADEvsSOTA} show said comparison.
+Figure \ref{fig:HeatmapsSHADEvsSOTA} shows said comparison.
 
@@
 	\myfloatalign
-	\subfloat[\acs{COPKM} vs \acs{SHADE}$_{CC}$]
+	\subfloat[\acs{COPKM} vs \acs{SHADE}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/COPKMvsSHADE.pdf}\label{fig:COPKMvsSHADE}}
-	\subfloat[\acs{LCVQE} vs \acs{SHADE}$_{CC}$]
+	\subfloat[\acs{LCVQE} vs \acs{SHADE}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/LCVQEvsSHADE.pdf}\label{fig:LCVQEvsSHADE}} \quad
-	\subfloat[\acs{RDPM} vs \acs{SHADE}$_{CC}$]
+	\subfloat[\acs{RDPM} vs \acs{SHADE}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/RDPMvsSHADE.pdf}\label{fig:RDPMvsSHADE}}
-	\subfloat[\acs{TVClust} vs \acs{SHADE}$_{CC}$]
+	\subfloat[\acs{TVClust} vs \acs{SHADE}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/TVClustvsSHADE.pdf}\label{fig:TVClustvsSHADE}} \quad
-	\subfloat[\acs{CECM} vs \acs{SHADE}$_{CC}$]
+	\subfloat[\acs{CECM} vs \acs{SHADE}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/CEKMvsSHADE.pdf}\label{fig:CECMvsSHADE}}
-	\caption[Heatmaps for the comparison of SHADE$_{CC}$ with the state-of-the-art.]{Heatmaps for the comparison of \acs{SHADE}$_{CC}$ with the state-of-the-art.}
+	\caption{Heatmaps comparing \acs{SHADE}$_{CC}$ with the state-of-the-art.}
 	\label{fig:HeatmapsSHADEvsSOTA}
@@
 In all of them most of the triplets are represented in the lower middle region of the diagram, with a slight shift of the point cloud to the left.
-That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---, but the advantage is for one or the other depending on the case, being \acs{SHADE}$_{CC}$ the one that gets it in most of them.
+That is indicative that there are indeed significant differences between the two methods---since there are no triplets with representation in the rope area---but the advantage is for one or the other depending on the case; \acs{SHADE}$_{CC}$ is usually the one that gets it in most of them.
 However, this does not apply in the case of \acs{CECM} (\ref{fig:CECMvsSHADE}), where the advantage is indisputably for \acs{SHADE}$_{CC}$.
 
-\section{Testing DILS$_{CC}$} \label{sec:TestDILS}
+\section[Testing \acsfont{DILS}$_{CC}$]{Testing DILS$_{CC}$} \label{sec:TestDILS}
 
-In this case we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with \acs{DILS}$_{CC}$ as sample $B$.
+In this case we refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with \acs{DILS}$_{CC}$ as sample $B$.
 
@@
 	\myfloatalign
-	\subfloat[\acs{BRKGA}$_{CC}$ vs \acs{DILS}$_{CC}$]
+	\subfloat[\acs{BRKGA}$_{CC}$ vs \acs{DILS}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/BRKGAvsDILS.pdf}\label{fig:BRKGAvsDILS}}
-	\subfloat[\acs{COPKM} vs \acs{DILS}$_{CC}$]
+	\subfloat[\acs{COPKM} vs \acs{DILS}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/COPKMvsDILS.pdf}\label{fig:COPKMvsDILS}} \quad
-	\subfloat[\acs{LCVQE} vs \acs{DILS}$_{CC}$]
+	\subfloat[\acs{LCVQE} vs \acs{DILS}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/LCVQEvsDILS.pdf}\label{fig:LCVQEvsDILS}}
-	\subfloat[\acs{RDPM} vs \acs{DILS}$_{CC}$]
+	\subfloat[\acs{RDPM} vs \acs{DILS}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/RDPMvsDILS.pdf}\label{fig:RDPMvsDILS}} \quad
-	\subfloat[\acs{TVClust} vs \acs{DILS}$_{CC}$]
+	\subfloat[\acs{TVClust} vs \acs{DILS}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/TVClustvsDILS.pdf}\label{fig:TVClustvsDILS}}
-	\subfloat[\acs{CECM} vs \acs{DILS}$_{CC}$]
+	\subfloat[\acs{CECM} vs \acs{DILS}$_{CC}$.]
 	{\includegraphics[width=0.5\linewidth]{gfx/AnalysisResults/CEKMvsDILS.pdf}\label{fig:CECMvsDILS}}
-	\caption[Heatmaps for the comparison of DILS$_{CC}$ with the state-of-the-art.]{Heatmaps for the comparison of \acs{DILS}$_{CC}$ with the state-of-the-art.}
+	\caption{Heatmaps comparing \acs{DILS}$_{CC}$ with the state-of-the-art.}
 	\label{fig:HeatmapsDILSvsSOTA}
@@
 
-Figure \ref{fig:HeatmapsDILSvsSOTA} shows all six heatmaps obtained when applying the Bayesian sign test to \acs{DILS}$_{CC}$ and the state-of-the-art methods.
+Figure \ref{fig:HeatmapsDILSvsSOTA} shows all six heatmaps obtained after applying the Bayesian sign test to \acs{DILS}$_{CC}$ and the state-of-the-art methods.
 Diagrams for \acs{BRKGA}$_{CC}$ (\ref{fig:BRKGAvsDILS}), \acs{LCVQE} (\ref{fig:LCVQEvsDILS}), \acs{RDPM} (\ref{fig:RDPMvsDILS}), \acs{TVClust} (\ref{fig:TVClustvsDILS}) and \acs{CECM} (\ref{fig:CECMvsDILS}) are very similar.
 They suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle.
-Bearing this in mind, and considering that \acs{ARI} is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that \acs{DILS}$_{CC}$ does represent a significant improvement over the named 5 methods, being the results offered by the latter non equivalent to those offered by \acs{DILS}$_{CC}$ and lower in quality.
-In the case of \acs{COPKM} (Figure \ref{fig:BRKGAvsDILS}) we find the cloud of triplets closer to the lower-middle of the diagram, indicating that results offered by \acs{COPKM} and \acs{DILS}$_{CC}$ remain being significantly different but in some cases the advantage would be for \acs{COPKM}.
-However, in most cases \acs{DILS}$_{CC}$ will offer better results.
+Bearing this in mind, and considering that \acs{ARI} is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that \acs{DILS}$_{CC}$ does represent a significant improvement over the named five methods, the results offered by the latter being non-equivalent to those offered by \acs{DILS}$_{CC}$ and lower in quality.
+In the case of \acs{COPKM} (Figure \ref{fig:BRKGAvsDILS}) we find the cloud of triplets closer to the lower-middle region of the diagram, indicating that the results offered by \acs{COPKM} and \acs{DILS}$_{CC}$ remain significantly different but in some cases \acs{COPKM} has an advantage.
+However, in most cases \acs{DILS}$_{CC}$ offers better results.
 
-\section{SHADE$_{CC}$ or DILS$_{CC}$?} \label{sec:SHADEorDILS}
+\section[\acsfont{SHADE}$_{CC}$ or \acsfont{DILS}$_{CC}$?]{SHADE$_{CC}$ or DILS$_{CC}$?} \label{sec:SHADEorDILS}
 
-Regarding the comparison between the two new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---, we have seen how both of them represent an improvement over previous constrained clustering approaches.
-In Section \ref{sec:NewPropvsSOTA} we saw how \acs{DILS}$_{CC}$ seemed to offer better results than \acs{SHADE}$_{CC}$, but it is worth empirically analyze this conclusions through the Bayesian sign test.
-We will refer to the results obtained \acs{SHADE}$_{CC}$ $A$, and to the results obtained with \acs{DILS}$_{CC}$ as sample $B$.
+Regarding the comparison between the two new proposals, we have seen how both of them represent an improvement over previous constrained clustering approaches.
+In Section \ref{sec:NewPropvsSOTA} we saw how \acs{DILS}$_{CC}$ seemed to offer better results than \acs{SHADE}$_{CC}$, but it is worth empirically analyzing these conclusions through the Bayesian sign test.
+We refer to the results obtained \acs{SHADE}$_{CC}$ as sample $A$, and to the results obtained with \acs{DILS}$_{CC}$ as sample $B$.
 Figure \ref{fig:SHADEvsDILS} shows the heatmap obtained with this setup.
@@
 	\includegraphics[scale=0.6]{gfx/AnalysisResults/SHADEvsDILS.pdf}
-	\caption[Heatmap for the comparison of SHADE$_{CC}$ with DILS$_{CC}$]{Heatmap for the comparison of \acs{SHADE}$_{CC}$ with \acs{DILS}$_{CC}$}\label{fig:SHADEvsDILS}
+	\caption{Heatmap comparing \acs{SHADE}$_{CC}$ with \acs{DILS}$_{CC}$.}\label{fig:SHADEvsDILS}
 \end{figure}
 
-As we could expect, the test assigns a high probability to $A - B < 0$, being the most of the triplets represented in the lower-left third of the triangle.
+As we could expect, the test assigns a high probability to $A - B < 0$, with most of the triplets being represented in the lower-left third of the triangle.
 This indicates that there are significant differences between both methods in most cases, and the test gives a clear lead to \acs{DILS}$_{CC}$.
-We conclude that the reason behind this is that a real-based problem representation does not provide any other advantage that the possibility to apply real-coding evolutionary methods.
-This way, given that the \acs{SHADE} method proved to be the best approach to real optimization problems we expected it to outperform \acs{BRKGA}, though \acs{SHADE} imposes an arithmetic relationship between individuals---through the mutation operator---which is hard to understand in the clustering domain.
-This is, applying an arithmetic operator to two labels does not provide easily understandable combination of them.
-Judging from the raw results and the empirical analysis, and with its straight integer-based problem representation and its strong exploitation capability \acs{DILS}$_{CC}$ stands out as the best of all methods we obtained results for.
+We conclude that the reason behind this is that a real-coded problem representation does not provide any advantages other  than the possibility to apply real-coded evolutionary methods.
+This way, given that \acs{SHADE} proved to be the best approach to real-coded optimization problems, we expected it to outperform \acs{BRKGA}, although \acs{SHADE} imposes an arithmetic relationship between individuals (through the mutation operator) which is difficult to understand in the clustering domain.
+This is, applying an arithmetic operator to two labels does not provide easily understandable combinations of them.
+Judging from the raw results and the empirical analysis, and with its straightforward integer-based problem representation and strong exploitation capabilities, \acs{DILS}$_{CC}$ stands out as the best of all methods we obtained results for.

### 2019-07-06 Memoria/Chapters/Clustering.tex
Acabada memoria excepto conclusiones y trabajo futuro
Cambios en memoria y en paper DILS

@@
 
-Firstly, it will be necessary to provide an introduction, albeit a brief one, to clustering and its applications, in order to ease the understanding of the following chapters.
-For this purpose we will follow the study carried out in \cite{Everitt:2009:CA:1538772}.
+Firstly, we need to provide an introduction, albeit a relatively brief one, to clustering and its applications in order to ease the understanding of the following chapters.
+For this purpose, the study carried out in \cite{Everitt:2009:CA:1538772} serves as a basis.
 In Section \ref{sec:ReasonsClassif} we present the reasons that lead humans to develop automatic classification methods.
-The numerical formulation for the latter is given in Section \ref{sec:NumMethClassif}.
-We give an intuitive definition for ``cluster'' in Section \ref{sec:WhatIsCluster}.
-Lastly a list of clustering application is presented in Section \ref{sec:ClusteringApplications}.
+The numerical formulation of the latter is shown in Section \ref{sec:NumMethClassif}.
+We give an intuitive definition of ``cluster'' in Section \ref{sec:WhatIsCluster}.
+Lastly, a list of clustering application is presented in Section \ref{sec:ClusteringApplications}.
 
@@
 
-Classification can be understood as a way of simplifying the information contained in large datasets in a way that is easily understood by humans.
-In this way, the processes of extracting useful information and applying it are simplified.
-Thus, if we are able to meaningfully divide a large dataset into subsets or groups, we can extract information common to elements of one of these subsets and provide a precise description that encompasses them.
+Classification can be understood as a way of simplifying the information contained in large datasets so that is easily understood by humans.
+Thus, the extraction and posterior use of the information are also simplified.
+If we are able to meaningfully divide a large dataset into subsets or groups, we can extract information common to elements of one of these subsets and provide a precise description that encompasses them.
 
-The need to analyze information in this way grows with the emergence and availability of large datasets in science.
+The need to analyze information grows with the emergence and availability of large datasets in science.
 The analysis of this type of information by classification, or clustering, is known today as \textit{Data Science}.
-In the 21st century a particular interest in data science arises from the appearance of the \textit{World Wide Web}, known as the Internet, where the aim has become to extract relevant information from the Web pages that make up this vast network.
+In the 21st century a particular interest in data science arose with the advent of the \textit{World Wide Web}, known as the Internet, where the aim has become to extract relevant information from the Web pages that make up this vast network.
 
@@
 
-Numerical methods for clustering arise in branches of natural sciences, such as biology or zoology, in an attempt to eliminate the subjectivity implicit in the classification process triggered by the discovery of a new species.
-The aim is to provide a non-subjective and stable method for classifying and grouping elements.
+Numerical methods for clustering appeared in branches of natural sciences, such as biology or zoology, in an attempt to eliminate the subjectivity implicit in the classification of new species.
+The aim is to provide an objective and stable method for classifying and grouping elements.
 
 These methods adopt different names that depend on the field in which they are applied: numerical taxonomy in biology, $Q$ analysis in psychology, or unsupervised pattern recognition in \acf{AI}.
-However, today, cluster analysis, or just clustering, are the most widely accepted and extended terms for tasks that involve the discovery of subgroups within a set of elements.
+Today, however, \textit{cluster analysis} or just \textit{clustering} are the most widely accepted and extended terms for tasks that involve the discovery of subgroups within a set of elements.
 
-In most clustering applications, the goal is to obtain a data partition, in which each instance or object belongs to a single cluster, and the union of all of them contains all the individual objects.
-That said, it should be noted that, in some circumstances, solutions where there is overlap between clusters are acceptable, as well as there may not be an acceptable data partition.
+In most clustering applications, the goal is to obtain a data partition, in which each instance or object belongs to a single cluster, and the union of all clusters contains all the individual objects.
+That said, it should be noted that, in some circumstances, solutions where there is an overlap between clusters are acceptable, or there may not be acceptable data partitions at all.
 
 In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
-In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
-A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
-In this work we will focus on partitional clustering.
+In hierarchical clustering the result is not a partition of the data with a certain number of clusters (as in partitional clustering) but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
+A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the \acsfont{CURE} method should be highlighted \cite{guha1998cure}.
+In this work we focus on partitional clustering.
 
-The most common way to represent the information on which clustering is applied is an array $X$ of dimension $n\times u$, in which each row corresponds to an instance or object to be processed, and each column corresponds to one of the variables that characterize those instances or objects.
-The commonly accepted term to refer to each row is \textit{features vector}, and the features vector set is referred as \textit{dataset}.
+The most common way to represent the data on which clustering is applied is an array $X$ of dimension $n\times u$.
+Each row corresponds to an instance or object to be processed, and each column corresponds to one of the variables that characterize those instances.
+The commonly accepted term to refer to each row is \textit{feature vector}, and the set of feature vectors is called \textit{dataset}.
 
@@
 
-The entry $x_{i,j}$ in $X$ corresponds to the value of the variable $j$th in the instance $i$.
+The entry $x_{i,j}$ in $X$ corresponds to the value of the $j$th variable in the $i$th instance.
 We will refer to the $i$th instance in $X$ as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$, which is a feature vector.
@@
 This mix of variable types and missing values can complicate the clustering task.
-However, there are methods to deal with these cases, such as inference of missing values or domain transformations.
+However, there are methods to deal with this, such as inference of missing values or domain transformations.
 
-In some applications, matrix rows may contain repeated measurements of the same variable, although under different conditions, or at different times, even at different spatial locations.
-An example of this can be the height measurements of a group of children in the same month over different years.
+In some applications, matrix rows may contain repeated measurements of the same variable under different conditions, at different times, or even at different spatial locations.
+An example of this can be the height measurements of a group of children in the same month over the years.
 This type of data has a structure that, again, can complicate the task of clustering.
 
-Some clustering methods involve making transformations on the $X$ matrix to convert it into a $n \times n$ matrix in which measurements extracted from the $X$ matrix are stored that relate one instance to all the others, such as similarity, distance or dissimilarity.
+Some clustering methods involve making transformations on the $X$ matrix.
+They convert it into a $n \times n$ matrix that stores measurements extracted from the $X$ matrix which relate one instance to all the others (such as similarity, distance or dissimilarity).
 
-Clustering is, put simply, the discovery of groups in data, and should never be confused with methods of discrimination or assignment, known in \acs{AI} as Supervised Learning, as we have already mentioned in Section \ref{sec:TypesOfML}.
-
-Once the general structure of clustering methods has been defined, it would be justified to ask, what is a cluster? Section \ref{sec:WhatIsCluster} will try to answer this question.
+Put simply, clustering is the discovery of groups in data, and should never be confused with methods of discrimination or assignment; these are known in \acs{AI} as Supervised Learning, as we have already mentioned in Section \ref{sec:TypesOfML}.
+Once the general structure of clustering methods has been defined, the next question would be: What is a cluster?
+Section \ref{sec:WhatIsCluster} tries to answer this question.
 
@@
 
-So far, the terms cluster, group and class have been used in a completely intuitive way, with no need for any formal definition, a further proof of the innate nature of these concepts in the human being.
+So far, the terms cluster, group and class have been used in a completely intuitive way, with no need for any formal definition, in a further proof of the innate nature of these concepts in the human being.
 In fact, giving a formal definition of cluster is a task, not only complicated, but in many occasions not very useful.
-For example, in \cite{lance1967general} a cluster definition completely dependent on the user's interpretation is proposed.
-As far as this definition is concerned, a cluster is what the user understands as a cluster without having proposed a formal definition of it.
+For example, in \cite{lance1967general} a cluster definition that is completely dependent on user interpretation is proposed.
+According to this definition, a cluster is just the preconceived notion of cluster that the user already had.
 
@@
 
-Although the definition above is accurate in a wide range of situations, a somewhat more analytical definition from the mathematical point of view, defining a cluster in terms of internal cohesion---homogeneity---, and external isolation---separation is proposed in \cite{cormack1971review} and \cite{gordon:1999}.
-Figure \ref{fig:ClustersProperties} informally illustrates the properties described above, so that any observer will find the clusters present in it apparent without the need for a formal cluster definition.
+Although the definition above is accurate in a wide range of situations, a somewhat more analytical approach from the mathematical point of view is needed.
+A definition of cluster in terms of internal cohesion (homogeneity) and external isolation (separation) is proposed in \cite{cormack1971review} and \cite{gordon:1999}.
+Figure \ref{fig:ClustersProperties} informally illustrates the properties described above, so that any observer will easily spot the clusters present in it without the need for a formal definition.
 This may explain why achieving a mathematically accurate definition of homogeneity and separation may become unnecessary.
@@
 	{\includegraphics[width=.3\linewidth]{gfx/Clustering/ThreeBasicClusters}}
-	\caption[Clusters with internal cohesion and/or external isolation]{Clusters with internal cohesion and/or external isolation}\label{fig:ClustersProperties}
+	\caption{Clusters with internal cohesion and/or external isolation.}\label{fig:ClustersProperties}
 \end{figure}
 
-It is not entirely clear how people recognize different clusters when they are represented on a plane, but one of the variables that certainly influences is the distribution of relative distances between objects or points.
+It is not entirely clear how people recognize different clusters when they are represented on a plane, but one of the variables that certainly has an influence is the distribution of relative distances between objects or points.
 
 On the other hand, as mentioned earlier in this section, it may be the case that there is no justified partition in a dataset.
-Figure \ref{fig:uniform_cloud} shows a dataset for which most observers would conclude that there are no distinct groups, just a cloud of uniformly distributed points.
+Figure \ref{fig:uniform_cloud} shows a dataset for which most observers would conclude that there are no distinct groups but only a cloud of uniformly distributed points.
 Ideally, one would expect a clustering method applied to this same dataset to reach the same conclusion.
@@
 However, most unsupervised learning methods will result in uniform partitioning as shown in Figure \ref{fig:uniform_partition}.
-The number of partitions found will depend on the method applied, although uniform partitioning will be achieved in any case.
+The number of partitions will depend on the method applied, although uniform partitioning will be achieved in any case.
 
@@
 The process of dividing a homogeneous distribution of data into different groups is known as dissection, and such a process may be useful in certain circumstances.
-However, since in most cases of actual application of cluster methods, the structure of the data is not known a priori, there is risk of interpreting all solutions in terms of the existence of subgroups, which would lead to the imposition of a fictitious structure on data in which there is no structure present.
+However, since in most cases the structure of the data is not known a priori, there is a risk of interpreting all solutions in terms of the existence of subgroups, which would lead to the imposition of a fictitious structure where there is none.
 
@@
 
-As it has been already mentioned, the general problem clustering attempts to solve is present in many disciplines: biology, botany, medicine, psychology, geography, marketing, image processing, psychiatry, archaeology, etc.
-This section presents some of the clustering applications related to these and others knowledge fields.
+As has been already mentioned, the general problem clustering attempts to solve is present in many disciplines: biology, botany, medicine, psychology, geography, marketing, image processing, psychiatry, archaeology, etc.
+This section presents some of the applications of clustering related to these and others knowledge fields.
 
@@
 
-Dividing customers into homogeneous groups is one of the most common marketing tasks.
-A marketing manager might wonder how to group potential customers according to the potential benefits of the product he or she is trying to introduce to the market.
-On the other hand, a marketing analyst might be interested in grouping companies according to their financial characteristics, in order to be able to analyze them and predict their market strategies.
+Separating customers into homogeneous groups is one of the most common marketing tasks.
+A marketing manager might wonder how to group potential customers according to the potential benefits of a new product being introduced into the market.
+On the other hand, a marketing analyst might be interested in grouping companies according to their financial characteristics, in order to analyze them and predict their market strategies.
 
 An example of the application of clustering in this field was published in \cite{green1967cluster}.
-With a large number of cities available for analysis, they had to restrict the places in which to carry out their studies for economic reasons.
-To this end they relied on cluster analysis, classifying cities into small groups based on 14 city features, including size and average income \textit{per capita}.
-Since cities in the same group were expected to be very similar, they chose one city from each of them to conduct their studies.
+With a large number of cities available for analysis, they had to restrict the places in which to carry out their study for economic reasons.
+To this end, they relied on cluster analysis to classify cities into small groups based on fourteen city features, including size and average income \textit{per capita}.
+Since cities in the same group were expected to be very similar, they chose one city from each of them to conduct their study.
 
 Another application of cluster analysis in marketing was described in \cite{chakrapani2004statistics}.
-In this case, a car manufacturer believes that buying a sports car is not a decision based solely on economic capabilities or age, but is a lifestyle decision made by those who decide to buy a sports car, as opposed to those who do not.
-Consequently, the manufacturer decides to carry out a study, using cluster analysis, which allows it to identify all features related to the people who would buy a sports car, in order to focus its marketing campaigns specifically on this sector.
+In this case, a car manufacturer believes that buying a sports car is not a decision based solely on economic capabilities or age, but instead a lifestyle decision made by those who decide to buy a sports car (as opposed to those who do not).
+Consequently, the manufacturer decides to carry out a study, using cluster analysis, to identify what types of person would buy a sports car in order to focus its marketing campaigns specifically on this sector.
 
@@
 
-Given a set of astronomical data, researchers want to know, for example, how many classes of stars are present in them, based on some statistical criteria.
-The most frequently asked questions in this area are: how many statistically different objects are present in the data and to which class should each object be assigned?
+Given a set of astronomical data, researchers want to know, for example, how many classes of stars are present in them according to some statistical criteria.
+The most frequently asked questions in this area are: How many statistically different objects are present in the data and which class should each object be assigned to?
 Do previously unknown classes of objects appear?
 Cluster analysis can be applied to answer these questions, helping to detect statistically anomalous objects, as well as to guide the process of classifying them.
-Examples include the discovery of high redshift quasars, type 2 quasars (highly luminous, active galactic nuclei often obscured by dust and gas), and brown dwarfs.
+Examples include the discovery of high redshift quasars, type 2 quasars---highly luminous, active galactic nuclei often obscured by dust and gas---and brown dwarfs.
 
 In \cite{faundez1996classification} a specific example of the above can be found.
-Clustering techniques are applied to data on the chemical composition of 192 planetary nebulae.
-Six different groups were identified, which were similar in many respects to a previous classification of such objects, but also showed interesting differences that researchers had so far missed.
+Clustering techniques were applied to data on the chemical composition of 192 planetary nebulae.
+Six different groups were identified, which were similar in many aspects to a previous classification of such objects, but also showed interesting differences that researchers had so far missed.
 
@@
 
-Mind disorders are often more difficult to diagnose than body pathologies, which is why there has been a growing interest in cluster analysis techniques to refine, or even redefine, diagnostic techniques for this type of disease in the field of psychiatry.
-Much of this work involves depressed patients, cases in which the interest lies in distinguishing between two types of depression, endogenous (congenital) and neurotic.
+Mind disorders are often more difficult to diagnose than body pathologies, which is why there has been a growing interest in cluster analysis techniques to refine, or even redefine, diagnostic techniques in the field of psychiatry.
+Much of this work involves depressed patients, cases where the interest lies in distinguishing between two types of depression: endogenous (congenital) and neurotic.
 
-In \cite{pledger2008using} clustering techniques are applied to 200 patients based on their responses to a questionnaire about depression, along with information about their sex, age, mental state, and illness.
+In \cite{pledger2008using} clustering techniques are applied to 200 patients based on their answers to a questionnaire about depression, along with information about their sex, age, mental state, and illnesses.
 This is a clear example of different types of variables included in the same dataset.
@@
 In \cite{paykel1978classification} a study on 236 cases of failed suicides recorded by the emergency service of a city in the United States of America is presented.
-From the set of available variables, 14 were selected as particularly relevant for classification and, therefore, were used in the analysis.
+From the set of available variables, fourteen were selected as particularly relevant for classification and, therefore, were used in the analysis.
 Among them were: age, number of suicide attempts, severity of depression and degree of hostility, in addition to a series of demographic features.
-Clustering methods were applied to the resulting dataset; the most significant result obtained corresponds to a division into three well-defined clusters.
+Clustering methods were applied to the resulting dataset; the most significant result corresponds to a division into three well-defined clusters.
 
@@
 
-In \cite{littmann2000empirical} clustering techniques are applied to data collected on daily changes in surface pressure in the Mediterranean basin, 20 groups explaining the variance of rainfall in the central Mediterranean regions were found.
-Another example can be found in \cite{liu2005mining}, in which the Fuzzy K-means algorithm is applied to spatial-temporal data of the USA southern-central regions climatology.
+In \cite{littmann2000empirical} clustering techniques are applied to data collected on daily changes in surface pressure in the Mediterranean basin, and twenty groups explaining the variance of rainfall in the central Mediterranean regions were found.
+Another example can be found in \cite{liu2005mining}, where the Fuzzy K-means algorithm is applied to spatial-temporal data on the climatology of southern-central regions of \acsfont{USA}.
 
@@
 Archaeology is another discipline in which clustering is useful.
-The classification of different objects found in deposits can help to discover their use, the periods to which they belong, as well as the population that used them.
-Similarly, studying fossilized materials can help reveal how prehistoric societies lived.
+The classification of different objects found in deposits can help to discover their use, the historical periods to which they belong, as well as the population that used them.
+Similarly, studying fossilized materials can help reveal how prehistorical societies lived.
 
-An early example of the application of clustering to archaeological objects is given in \cite{hodson1966some}, where clustering techniques are applied to a group of brooches dating from the Iron Age, finding a classification for them of proven archaeological relevance.
+An early example of the application of clustering to archaeological objects is given in \cite{hodson1966some}, where clustering techniques are applied to a group of brooches dating from the Iron Age, producing a classification for them of proven archaeological relevance.
 Another example found in \cite{hodson1971numerical} is the application of the K-means algorithm to construct a taxonomy of hand axes found in the British Isles.
@@
 
-Regarding fossilized materials, in \cite{sutton1995cluster} a study of 155 coprolites found at Antelope House, a prehistoric site at Chelly Canyon in Arizona is conducted.
+Regarding fossilized materials, in \cite{sutton1995cluster} a study of 155 coprolites found at Antelope House, a prehistoric site at Chelly Canyon in Arizona, is conducted.
 The study resulted in a diet-based interpretation of the differences between coprolites.
@@
 
-Recent times are witnessing a tremendous boom in interest in Bioinformatics, complemented by molecular biology, computer science, mathematics and statistics.
-Such an increase has been accelerated by the ever-growing database of genomic and protein data, which are themselves the result of a major advance in DNA sequencing techniques, gene expression measurements, and compression of macromolecular structures.
-Statistics have been relevant in the study of gene expression.
-The genes contained in the DNA of each cell provide the necessary templates for the generation of the proteins involved in most of the structural and biomechanical processes that take place in each of us.
+Recent times are witnessing a tremendous boom in interest in bioinformatics, complemented by molecular biology, computer science, mathematics and statistics.
+Such an increase has been accelerated by the ever-growing database of genomic and protein data, which are themselves the result of a major advance in \acsfont{DNA} sequencing techniques, gene expression measurements, and compression of macromolecular structures.
+Statistics has been relevant in the study of gene expression.
+The genes contained in the \acsfont{DNA} of each cell provide the necessary templates for the generation of the proteins involved in most of the structural and biomechanical processes that take place in each of us.
 However, although most cells in humans contain all the genetic components that make up the human genome, genes are expressed selectively in each cell depending on cell type, tissue and general conditions inside and outside the cell.
-Molecular biology has shown that most of the processes in a cell's life are regulated by factors that affect the expression of its genes.
+Molecular biology has shown that most of the processes that occur in a cell are regulated by factors that affect the expression of its genes.
 
 As we have seen, one of the most active fields of research today is the study of the processes that regulate gene expression.
-In order to store the information related to this area of study, microarrays arise \cite{cortese2000array}.
-From the data analysis point of view, one of the relevant features in this type of information is that the number of attributes of each instance ($u$) far exceeds the number of available instances ($n$); datasets featuring this property are called \textit{high dimensionality datasets}.
+Microarrays arose as a way to store the information related to this area of study \cite{cortese2000array}.
+From the point of view of data analysis, a relevant aspect of this type of information is that the number of attributes of each instance ($u$) far exceeds the number of available instances ($n$); datasets featuring this property are called \textit{high-dimensionality datasets}.
 
-Most classic statistical methods cannot be applied to this type of dataset without being substantially modified.
-However, cluster analysis accepts such datasets and can be used to identify groups of genes with similar expression patterns.
+Most classic statistical methods cannot be applied to this type of dataset without undergoing considerable modifications.
+However, cluster analysis is compatible with such datasets and can be used to identify groups of genes with similar expression patterns.
 This allows us to answer questions such as why a gene is affected by a certain disease, or which genes are responsible for hereditary genetic diseases.
 
-An example of application is found in \cite{selinski2008cluster}, who used clustering on single-nucleotide polymorphisms to detect differences between diseases at the genetic level.
+An example of application is found in \cite{selinski2008cluster}, where clustering was used on single-nucleotide polymorphisms to detect differences between diseases at a genetic level.
 
@@
 Although the examples of practical clustering applications presented so far are illustrative, it should be noted that clustering techniques are applied in a multitude of modern problems.
-Since it is not humanly possible to cover all the practical uses of clustering, a brief list of some of them is presented here: computer vision-guided traffic management \cite{kumaran2019computer}, near real-time detection of
-spatial clusters from large position data streams generated
-by moving objects \cite{junior2019dg2cep}, wireless sensor networks management \cite{wan2019similarity}, internet of things \cite{aranzazu2019anchor}, detecting taxi movements \cite{ibrahim2019detecting}, \textit{big data} distributed multi-target regression \cite{corizzo2019dencast},  cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
+Since it is not humanly possible to cover all the practical uses of clustering, a brief list of some of them is presented here: computer vision-guided traffic management \cite{kumaran2019computer}, near real-time detection of spatial clusters from large position data streams generated by moving objects \cite{junior2019dg2cep}, wireless sensor networks management \cite{wan2019similarity}, Internet of Things \cite{aranzazu2019anchor}, detecting taxi movements \cite{ibrahim2019detecting}, \textit{big data} distributed multi-target regression \cite{corizzo2019dencast},  cartographic labeling \cite{araujo2019improving}, and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
 
-It seems clear that clustering is an extremely powerful tool, no only for the development of any branch of science, but also because it can be applied to  many real-world problems and directly help people.
+It seems clear that clustering is an extremely powerful tool, not only for the development of any branch of science, but also because of its wide range of real-world and humanitarian applications.
 
@@
 
-Clustering consists of exploring datasets.
-Its goal is to determine whether or not they can be summarized in a meaningful way, in terms of a relatively small number of groups or clusters of objects or individuals which resemble each other and differ from those found in other clusters.
+Clustering consists in exploring datasets.
+Its goal is to determine whether they can be summarized in a meaningful way, in terms of a relatively small number of clusters of instances which resemble each other and differ from those found in other clusters.
 
-Many branches of science have successfully made use of clustering techniques to advance in their respective fields and to process large amounts of data, whose analysis would be unthinkable to tackle with traditional techniques.
-
-%Para referir un acronimo completo \acf
-%Para referir un acronimo solo con las siglas \acs
+Many branches of science have successfully made use of clustering techniques to facilitate the development of their respective fields and to process large amounts of data, whose analysis would be unthinkable to tackle with traditional techniques.

### 2019-07-06 Memoria/Chapters/ConclusionsAndFutureWork.tex
Cambios en memoria y en paper DILS

@@
 \chapter{Conclusions and Future Work}\label{ch:ConclFW}
+
+In this work we proposed the first application of \acf{DE} to the \acf{SSL} constrained clustering problem, taking as a basis its \acf{SHADE} variant and designing a new heuristic function and an unbiased \acf{LS} procedure.
+We called this applied method \acs{SHADE}$_{CC}$.
+We also presented the newly created \acf{DILS} heuristic method, along with its application to the constrained clustering problem, which we called \acs{DILS}$_{CC}$.
+Focusing on instance-level \acf{ML} and \acf{CL} constraints, \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$ have proven that nature-inspired and metaheuristic techniques constitute a competitive approach to constrained clustering, being able to scale the quality of the results in a directly proportional way to the number of constraints.
+
+Supported by Bayesian statistical tests, we were able to objectively prove that the \acs{DILS}$_{CC}$ approach is significantly better than the state-of-the-art.
+Its integer-based problem representation and its exploitation capabilities allow \acs{DILS}$_{CC}$ to outperform the current state-of-the-art, with \acs{SHADE}$_{CC}$ being able to achieve competitive results.
+Both \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$ have proven to be better in terms of quality and availability of the solutions when compared to the state-of-the-art, especially in cases where large constraint sets are analyzed.
+
+Regarding future research lines, we have been able to see how in some cases no constrained clustering algorithm is able to achieve quality results.
+Leaving aside parameter tuning, which will surely improve results in some cases, we find that there is no preprocessing method designed to consider side data-related information, such as constraints.
+A quality preprocessing phase is crucial to obtain quality results when applying data analysis techniques.
+This is why constraint-oriented preprocessing methods need to be developed in order to keep constrained clustering research on the right track.
+
+We can also mention two other unexplored approaches to constrained clustering: multi-objective optimization and data streaming.
+On one hand, as we have studied in this work, in constrained clustering the aim is to optimize two different measures: the intra-cluster distance and the number of unsatisfied constraints.
+Penalty-style fitness functions are able to reunite these two measures, but a multi-objective alternative might be able to do it better.
+On the other hand, we have shown how constraints can be automatically generated from a set of labeled data.
+This principle can be applied to data streaming clustering techniques so that constraints can also guide clustering methods within this framework.

### 2019-07-06 Memoria/Chapters/ConstrainedClustering.tex
Cambios en memoria y en paper DILS

@@
 
-Once the basics concepts about clustering have been introduced, we can focus im defining the concepts related to constrained clustering, as well as to give examples of its application and highlight the benefits and problems presented by its use.
+Once the basics concepts about clustering have been introduced, we can focus on defining the concepts specific to constrained clustering.
+We give example applications and highlight its advantages and disadvantages.
 To this end we take as our main reference the survey presented in \cite{davidson2007survey}.
-We will start by describing the motivation for constrained clustering in Section \ref{sec:CCMotiv}.
+We describe the motivation for constrained clustering in Section \ref{sec:CCMotiv}.
 We formalize the concept of constraints and discuss their use in Sections \ref{sec:ConstDef} and \ref{sec:ConstUse} respectively.
-Afterwards a list of constrained clustering application is presented in Section \ref{sec:CCApplications}.
+Afterwards a list of applications of constrained clustering is presented in Section \ref{sec:CCApplications}.
 Benefits and problems, as well as solutions for these problems, are discussed in Sections \ref{sec:ConstraintsBenefits} and \ref{sec:ConstraintsProblems} respectively.
@@
 is addressed.
-The problem consist of, given a large number of text documents, grouping them according to a taxonomy in which documents with similar topics are conceptually nearby.
-To solve this problem, unsupervised clustering methods are useful, as the information on the problem initially available is limited.
-However, in \cite{wagstaff2001constrained} is shown that applying unsupervised clustering to certain problems, such as grouping \acf{GPS} data in such a way that clusters define the lanes of a road, does not produce significant results, as the clusters obtained are far from the elongated shape that would be expected as a result.
-To tackle the problem, they introduced a new element in clustering, the constraints, which made it possible to include domain knowledge that would guide the clustering methods to obtain the expected results.
+The problem consists in, given a large number of text documents, grouping them according to a taxonomy in which documents with similar topics are conceptually nearby.
+To solve this problem, unsupervised clustering methods are useful, as there is limited initially available information on the problem.
+However, in \cite{wagstaff2001constrained} it is shown that applying unsupervised clustering to certain problems, such as grouping \acf{GPS} data in such a way that clusters define the lanes of a road, does not produce significant results, as the clusters obtained are far from the elongated shape that would be expected.
+To tackle the problem, they introduced a new element in clustering, the \textit{constraint}, which made it possible to include domain knowledge to guide clustering methods towards the expected results.
 It was enough to indicate that the lanes of the road on which the vehicles circulate are four meters wide, and therefore any vehicle that is at a distance of more than 4 meters from another, perpendicular to the motion direction, must be located in a different cluster.
 
-We are now in a new scenario: it is possible to incorporate additional information to the clustering process, in addition to that contained in the dataset itself, to guide it in the formation of the output partition and obtain more accurate results.
-This new approach is known as constrained clustering, a \acf{SSL} paradigm, unlike traditional clustering methods which fall within the area of unsupervised learning.
+We are now in a new scenario: it is possible to incorporate additional information into the clustering process, in addition to that contained in the dataset itself, to guide it in the formation of the output partition and obtain more accurate results.
+This new approach is known as constrained clustering, an \acf{SSL} paradigm, unlike traditional clustering methods which fall into the category of unsupervised learning.
 
@@
 
-The new type of information that we incorporate to clustering is given in the form of instance-level constraints, that is, to specify if two instances ($x_i$ and $x_j$) from the dataset ($X$) must be in the same cluster or, on the contrary, they must be in separate clusters.
+The new type of information that we incorporate into clustering is given in the form of instance-level constraints.
+Their purpose is to specify whether two instances ($x_i$ and $x_j$) from the dataset ($X$) must be in the same cluster or, on the contrary, they must be in separate clusters.
 
@@
 Although they may seem simple, constraints defined as above have some very interesting properties.
-\acs{ML} constraints are an example of an equivalence relationship, and therefore are symmetrical, reflexive and transitive.
+\acs{ML} constraints are an example of an equivalence relation, and therefore are symmetrical, reflexive and transitive.
 Observation \ref{ob:MLConstraintsRelation} formalizes this concept.
@@
 
-A clear example of constraint use can be found in clustering applications where there are distance measurement limitations, as in the case of \acs{GPS} data.
-Thus, if we want that instances forming two clusters to be separated by a distance greater or equal to $\delta$, it is enough to set \acs{ML} constraints between all instances separated by less than $\delta$ units.
+A clear example of the use of constraints can be found in clustering applications where there are distance measurement limitations, as in the case of \acs{GPS} data.
+Thus, if we want instances forming two clusters to be separated by a distance greater or equal to $\delta$, it is enough to set \acs{ML} constraints between all instances separated by less than $\delta$ units.
 Similarly, if we want the diameter of the clusters to be at most $\epsilon$, we must set \acs{CL} constraint between instances separated by more than $\epsilon$ units.
@@
 	\includegraphics[scale=0.45]{gfx/ConstClust/RestriccionesDeltaEpsilon.png} 
-	\caption[\textit{delta} and \textit{epsilon} constraints]{\textit{delta} and \textit{epsilon} constraints \cite{davidson2007survey}.}\label{fig:DistanceConstraints}
+	\caption[\textit{delta} $(\delta)$ and \textit{epsilon} $(\epsilon)$ constraints.]{\textit{delta} $(\delta)$ and \textit{epsilon} $(\epsilon)$ constraints \cite{davidson2007survey}.}\label{fig:DistanceConstraints}
 \end{figure}
@@
 On the other hand, in many domains the available information refers to relations between instances, and not to the specific class to which they belong.
-Moreover, in interactive clustering systems, a user who is not expert in the problem domain will probably be able to provide information in the form of constraints such as \acs{ML} and \acs{CL}, rather than providing information on what particular class certain instances belong to \cite{cohn2003semi,davidson2007hierarchical}.
+Moreover, in interactive clustering systems, a user who is not expert in the problem domain will probably be able to provide information in the form of constraints such as \acs{ML} and \acs{CL}, instead of providing information on what particular class certain instances belong to \cite{cohn2003semi,davidson2007hierarchical}.
 
 Typically, constraints are incorporated into clustering problems in two ways.
-They can be used to alter the instance clusters assignment rule of the method in hand, so that the solution satisfies as many constraints as possible.
+They can be used to alter the rules of assignment of instances to clusters of the method in question, so that the solution satisfies as many constraints as possible.
 Alternatively, it is possible to train the distance function used by the method based on the constraints, either before or during the application of the method.
-In any case, the initialization phase can take the constraints into account, so that instances associated with constraints \acs{ML} will be placed in the same cluster, and those between which there is a constraint \acs{CL} will be placed in different clusters.
-Based on this distinction, we identify two ways of approaching the problem, those based on explicit constraints (\textit{constraints-based}), and those based on distances (\textit{distance-based}).
+In any case, the initialization phase can take the constraints into account, so that instances appearing together in \acs{ML} constraints will be placed in the same cluster, and those between which there is a \acs{CL} constraint will be placed in different clusters.
+Based on this distinction, we identify two ways of approaching the problem: those based on explicit constraints (\textit{constraints-based}) and those based on distances (\textit{distance-based}).
 
@@
 
-	\item Initialize clusters based on constraints inferred from a set of labeled instances \cite{basu2002semi}.
+	\item Initializing clusters based on constraints inferred from a set of labeled instances \cite{basu2002semi}.
 
@@
 
-	\caption[Constraints and clustering over a dataset]{Constraints and clustering over a dataset \cite{davidson2007survey}.} \label{fig:ConstClustOverDataset}
+	\caption[Constraints and clustering over a dataset.]{Constraints and clustering over a dataset \cite{davidson2007survey}.} \label{fig:ConstClustOverDataset}
 \end{figure}
@@
 
-In distance-based approaches, classic clustering methods that use a distance measurement are employed, so that the distance measurement is modified to incorporate the constraints.
-In this context, satisfying constraints means that instances related to constraints \acs{ML} are placed together in the space, and those related by \acs{CL} are separated.
+In distance-based approaches, classic clustering methods that use a distance measurement are modified so that the distance measurement incorporates the constraints.
+In this context, constraint satisfaction means that instances co-occurring in \acs{ML} constraints are placed together in the space, and those co-occurring in \acs{CL} are separated.
 
 Figure \ref{fig:ConstrainsAndMetricLearned} shows a possible clustering based on a metric learned from the constraints.
-It should be noted that in Figure \ref{fig:MetricLearned} the data space has been compressed on the vertical axis and widened on the horizontal axis to match the learned distance metric.
+It should be noted that in Figure \ref{fig:MetricLearned} the data space has been compressed on the vertical axis and expanded on the horizontal axis to match the learned distance metric.
 
@@
 	\label{fig:MetricLearned}}
-	\caption[Clustering based on a metric learned from constraints.]{Clustering based on a metric learned based on constraints \cite{davidson2007survey}.} \label{fig:ConstrainsAndMetricLearned}
+	\caption[Clustering based on a metric learned from constraints.]{Clustering based on a metric learned from constraints \cite{davidson2007survey}.} \label{fig:ConstrainsAndMetricLearned}
 \end{figure}
@@
 
-This section shows some application cases in which constrained clustering has turned out to be a more useful tool than unsupervised clustering.
-For each case we will analyze how the constraints were obtained and how they improve the results in the resulting clustering.
+This section shows some application cases in which constrained clustering has turned out to be a more useful tool than fully unsupervised clustering.
+For each case we will analyze how the constraints were obtained and how they improve the resulting clustering.
 
@@
 
-Figure \ref{fig:CMUFacesDatabase} shows a sample from the CMU (Carnegie Mellon University) faces dataset, where the task is to group faces based on different criteria.
-In this case, the goal is to group faces according to their orientation.
+Figure \ref{fig:CMUFacesDatabase} shows a sample from the \acsfont{CMU} (Carnegie Mellon University) faces dataset, where the task is to group faces according to different criteria (in this case, face orientation).
 
@@
 	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Caras2}} \quad
-	\caption[CMU faces database.]{CMU faces database \cite{davidson2007survey}.}\label{fig:CMUFacesDatabase}
+	\caption[\acsfont{CMU} faces database.]{\acsfont{CMU} faces database \cite{davidson2007survey}.}\label{fig:CMUFacesDatabase}
 \end{figure}
 
-The method used to obtain the constraints is one of the most popular in the literature: set the number of clusters of the resulting partition equal to the number of classes in the database, and generate the constraints from a subset of labeled instances; that is, if two instances have different labels, set a \acs{CL} constraint between them, otherwise one of \acs{ML} type.
-Thus, between the images shown in Figure \ref{fig:FacesDatabaseCL} \acs{CL} constraints are set, since, although they belong to the same person, they do not have the same orientation.
+The method used to obtain the constraints is one of the most popular in the literature: let the number of clusters of the resulting partition be equal to the number of classes in the database, and generate the constraints from a subset of labeled instances; in other words, if two instances have different labels, set a \acs{CL} constraint between them, and otherwise set an \acs{ML} one.
+Thus, between the images shown in Figure \ref{fig:FacesDatabaseCL} \acs{CL} constraints are set, since, although they depict the same person, they do not have the same orientation.
 
@@
 
-Figure \ref{fig:AiboRobotClustSys} shows another set of image data on which constrained clustering techniques are applied.
+Figure \ref{fig:AiboRobotClustSys} shows another set of image data to which constrained clustering techniques are applied.
 In this case, the task is to perform object recognition to incorporate the method into the navigation system of the Aibo robot \cite{davidson2005clustering}.
 Distance constraints such as $\delta$ and $\epsilon$ are used as described in Figure \ref{fig:DistanceConstraints}.
-In this way well differentiated clusters are obtained and therefore they are useful for the path-finding techniques performed by the robot during navigation.
+In this way, well differentiated clusters are obtained and therefore they are useful for the path-finding tasks performed by the robot during navigation.
 
@@
 	\myfloatalign
-	\subfloat[Imagen original]
+	\subfloat[Original image.]
 	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo1}} \quad
-	\subfloat[Clustering sin restricciones]
+	\subfloat[Classic clustering.]
 	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo2}} \quad
-	\subfloat[Clustering con restricciones]
+	\subfloat[Constrained clustering.]
 	{\includegraphics[width=.3\linewidth]{gfx/ConstClust/AnalisisImagenes/Aibo3}} \quad
@@
 
-Video databases are one example where constraints can be generated directly from the data domain, especially if space-time data information is available \cite{yan2006discriminative}.
-In time sequenced data it is possible to set \acs{ML} constraints between groups of pixels of frames close in time.
+Video databases are one example where constraints can be generated directly from the data domain, especially if space-time metadata is available \cite{yan2006discriminative}.
+In time-sequenced data it is possible to set \acs{ML} constraints between groups of pixels of frames close in time.
 This is especially useful when the task is to perform object recognition based on clustering and segmentation.
-It is also possible to add \acs{CL} constraints to image segment in the same snapshot, as there is a low probability that they are associated with the same object after segmentation is done.
-In fact, in the video analysis domain problems there are a variety of constraint extraction methods \cite{yan2006discriminative}, Figure \ref{fig:ConstExtractionVideo} shows some examples.
+It is also possible to add \acs{CL} constraints to image segments in the same snapshot, as the probability of them being associated with the same object after segmentation is low.
+In fact, in video analysis problems there are a variety of constraint extraction methods \cite{yan2006discriminative}.
+Figure \ref{fig:ConstExtractionVideo} shows some examples.
 
@@
 
-In Figure \ref{fig:ConstExtractionVideo}, image \ref{fig:ConstExtractionVideoA} corresponds to constraints extracted from tracking a person over a period of time, \ref{fig:ConstExtractionVideoB} corresponds to spatial constraints associating two objects located in the same frame, image \ref{fig:ConstExtractionVideoC} corresponds to restrictions obtained through facial recognition and \ref{fig:ConstExtractionVideoD} to those provided by the user.
+In Figure \ref{fig:ConstExtractionVideo}, image \ref{fig:ConstExtractionVideoA} corresponds to constraints extracted from tracking a person over a period of time; \ref{fig:ConstExtractionVideoB} corresponds to spatial constraints associating two objects located in the same frame; image \ref{fig:ConstExtractionVideoC} corresponds to constraints obtained through facial recognition, and \ref{fig:ConstExtractionVideoD} to those provided by the user.
 
-With so many constraint extraction methods, one might ask: what happens if too many constraints are imposed?
+With so many constraint extraction methods, one might ask: What happens if too many constraints are imposed?
 Does this make the problem over-constrained?
@@
 
-In gene clustering based on micro-array data, genes are represented by their expression profile in different experiments and grouped using different constrained clustering methods.
-Figure \ref{fig:GeneticApp} shows an example: these are \acs{ML} constraints set between genes based on co-occurrence data stored in the protein interaction database, which contains information about which genes (and their associated proteins) are present in the same cell processes \cite{xenarios2001dip}.
+In gene clustering based on microarray data, genes are represented by their expression profile in different experiments and grouped using different constrained clustering methods.
+Figure \ref{fig:GeneticApp} shows an example: these are \acs{ML} constraints set between genes based on co-occurrence data stored in the protein interaction database, which contains information about which genes---and their associated proteins---are present in the same cell processes \cite{xenarios2001dip}.
 This information can be used to improve results provided by clustering methods.
@@
 	\includegraphics[scale=0.3]{gfx/ConstClust/Genetica/Genes} 
-	\caption[Gene clustering based on micro-array data.]{Gene clustering based on micro-array data \cite{davidson2007survey}.}\label{fig:GeneticApp}
+	\caption[Gene clustering based on microarray data.]{Gene clustering based on microarray data \cite{davidson2007survey}.}\label{fig:GeneticApp}
 \end{figure}
@@
 In this case it is possible to extract constraints from multiple auxiliary resources.
-For example, if two documents are in the same directory, a \acs{ML} constraints could be set between them.
-This way it is possible to adjust the resulting clustering to suit a particular criterion, such as creating a hierarchy of documents similar to the way they are organized in the input directories structure.
+For example, if two documents are in the same directory, an \acs{ML} constraint could be set between them.
+This way, it is possible to adjust the resulting clustering to suit a particular criterion, such as creating a hierarchy of documents akin to the way they are organized in the original directory structure.
 
@@
 Constrained clustering is very useful in web search data processing.
-Here, the aim is to automatically group the results of an ambiguous query in the search engine into clusters of \acs{URL}s, which refer to the concept introduced as a query in different contexts.
-In this area it is possible to extract the constraints from searches previously performed by users, so that a \acs{ML} constraints is set between \acs{URL}s visited in the same user session.
-Clustering using these constraints can help bias the result of searches towards user's preferences.
+Here, the aim is to automatically group the results of an ambiguous query to the search engine into clusters of \acsp{URL}, which represent the queried concept in different contexts.
+In this area it is possible to extract the constraints from previous searches performed by other users, so that an \acs{ML} constraint is set between \acsp{URL} visited in the same user session.
+Clustering using these constraints can help bias search results towards user preferences.
 
@@
 
-Constrained clustering is very useful in the processing of search data on web pages.
-Here, the aim is to automatically group the results of an ambiguous query in the search engine into clusters of \acs{URL}s, which refer to the concept introduced as a query in different contexts.
-In this area it is possible to extract constraints from searches previously performed by users, so that a \acs{ML} constraint is set between visited \acs{URL}s in the same user session.
-Clustering using these constraints can help biasing search results towards user preferences.
+In certain audio analysis tasks, the number of object classes present in the data may not be known, although constraints can be extracted directly from the data domain.
+This happens, for example, when applying clustering to speaker recognition in a conversation \cite{bar2003learning}.
+In this case the number of participants is not known a priori, but it is easy to detect if two speakers are different or similar and to set constraints accordingly.
 
-\subsection{GPS Data} \label{sec:GPSApp}
+\subsection[\acsfont{GPS} Data]{GPS Data} \label{sec:GPSApp}
 
-As mentioned at the beginning of Chapter \ref{ch:ConstrainedClustering}, constrained clustering can be applied to \acs{GPS} data to identify the lane in which each vehicle is running, as shown in Figure \ref{fig:GPSClustData}.
-Each instance is represented by the position it occupies on the road in two-dimensional Cartesian coordinates $(x,y)$, obtained from \acs{GPS} data.
-Figure \ref{fig:figure15} graphically shows this representation of the data (note that multiple instances can refer to the same vehicle at different times).
+As mentioned at the beginning of Chapter \ref{ch:ConstrainedClustering}, constrained clustering can be applied to \acs{GPS} data in order to identify the lane each vehicle is in, as shown in Figure \ref{fig:GPSClustData}.
+Each instance is represented by the position it occupies on the road in two-dimensional Cartesian coordinates $(x,y)$ obtained from \acs{GPS} data.
+Figure \ref{fig:figure15} graphically shows this representation of the data.
+Note that multiple instances can refer to the same vehicle at different times.
 
@@
 	\includegraphics[scale=0.3]{gfx/ConstClust/GPS/Coches} 
-	\caption[GPS data use.]{\acs{GPS} data use. \cite{davidson2007survey} \cite{wagstaff2001constrained}}\label{fig:GPSClustData}
+	\caption[\acs{GPS} data use.]{\acs{GPS} data use \cite{davidson2007survey, wagstaff2001constrained}.}\label{fig:GPSClustData}
 \end{figure}
@@
 In this domain, real clusters have an elongated shape on the horizontal axis and are aligned perpendicularly to the motion direction.
-To get the resulting clusters to have this shape, we can make use of the constraints.
-We will set a \acs{CL} constraint between those instances more than four meters apart perpendicularly to the motion direction (since the lanes have a maximum width of four meters), and \acs{ML} constraints between those instances that present continuity in the motion direction axis, since it is likely that the vehicles they represent are in the same lane.
-This clustering model has proven to be very useful in real-time navigation \cite{wagstaff2001constrained}, allowing the user to be notified when to change lanes, or when not to leave them.
+To produce clusters with this shape, we can make use of constraints.
+We will set \acs{CL} constraints between those instances more than four meters apart perpendicularly to the motion direction---since the lanes have a maximum width of four meters---and \acs{ML} constraints between those instances that present continuity in the motion direction axis, since it is likely that the vehicles they represent are in the same lane.
+This clustering model has proven to be very useful in real-time navigation \cite{wagstaff2001constrained}, allowing for the user to be notified when to change lanes, or when not to.
 
@@
 	\includegraphics[scale=0.32]{gfx/ConstClust/GPS/Instancias} 
-	\caption[Clustering of GPS data with no constraints.]{Clustering of \acs{GPS} data with no constraints \cite{davidson2007survey,wagstaff2001constrained}.}\label{fig:figure15}
+	\caption[Clustering of \acs{GPS} data with no constraints.]{Clustering of \acs{GPS} data with no constraints \cite{davidson2007survey,wagstaff2001constrained}.}\label{fig:figure15}
 \end{figure}
@@
 
-The already mentioned constrained clustering applications are the most studied in the literature, although the reader may have the impression that constrained clustering is no longer used in real-world problems.
-To prevent that we provide a list of some more recent applications of constrained clustering to modern problems: advanced robotics navigation systems \cite{semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, semantically similar relations identification \cite{wang2015constrained}, cooperative team building \cite{yang2014team}, video face clustering \cite{zhou2014video}, image clustering \cite{habashi2017semi} and storage location assignment in warehouses \cite{yang2016constrained} among others.
+The aforementioned applications are the most studied in the literature, although the reader may have the impression that constrained clustering is no longer used in real-world problems.
+To prevent that, we provide a list of more recent applications of constrained clustering to modern problems: advanced robotics navigation systems \cite{semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, semantically similar relations identification \cite{wang2015constrained}, cooperative team building \cite{yang2014team}, video face clustering \cite{zhou2014video}, image clustering \cite{habashi2017semi} and storage location assignment in warehouses \cite{yang2016constrained}, among others.
 
@@
 
-	\item Increased accuracy in label predictions by generating constraints based on a subset of labeled instances.
+	\item Increased accuracy in label prediction by generating constraints based on a subset of labeled instances.
 
-	\item Obtaining clusters with adaptable geometry to each problem.
+	\item Obtention of clusters with adaptable geometry to each problem.
 
@@
 
-Given $X = \{x_1 \cdots x_n\}$, a large set of unlabeled instances, and $L = \{(x_{n+1}, l_{n+1}) \cdots (x_{n+|L|}, l_{n+|L|})\}$, a small set of labeled instances, it is common to choose two elements from $L$ (with replacement) and set a \acs{ML} constraint between them if they belong to the same class, or a \acs{CL} constraints otherwise.
-An appropriate method to evaluate the results provided by a clustering method is to measure the level of accuracy it achieves when predicting dataset $X$ labels.
-This normally requires specifying the number of desired clusters equal to the number of known classes in $X$ ($k = k^*$).
+Given $X = \{x_1 \cdots x_n\}$, a large set of unlabeled instances, and $L = \{(x_{n+1}, l_{n+1}) \cdots (x_{n+|L|}, l_{n+|L|})\}$, a small set of labeled instances, it is common to choose two elements from $L$ (with replacement) and set an \acs{ML} constraint between them if they belong to the same class, or a \acs{CL} constraint otherwise.
+An appropriate way to evaluate the results provided by a clustering method is to measure the level of accuracy it achieves when predicting dataset $X$ labels.
+This normally requires the number of desired clusters to be equal to the number of known classes in $X$ ($k = k^*$).
 Methods such as the RandIndex \cite{rand1971objective} or its variants, which will be discussed later, are used to measure the degree of agreement between two given partitions (clusterings).
 
-In \cite{wagstaff2000clustering}, where constraints are generated as described above, it is shown that accuracy is increased up to a 20\% over classic methods when averaging results over different constraint sets.
+In \cite{wagstaff2000clustering}, where constraints are generated as described above, it is shown that accuracy is increased up to 20\% over classic methods when averaging results over different constraint sets.
 
@@
 
-This rule, however, is not true in all cases, as in datasets such as \textit{Tic-Tac-Toe Endgame}, no increase in predictions is achieved regardless of the number of constraints used.
-A possible explanation for these exceptions is based on the fact that setting $k = k^*$ may not be appropriate in these cases.
+This rule, however, is not always true, as in datasets such as \textit{Tic-Tac-Toe Endgame}, no increase in correct predictions is achieved regardless of the number of constraints used.
+A possible explanation for these exceptions is that setting $k = k^*$ may not be appropriate in these cases.
 
@@
 
-Although, as we have seen, incorporating constraints to clustering methods brings benefits in some applications, there are two main drawbacks that are discussed below, as well as possible solutions to them.
+Although, as we have seen, incorporating constraints into clustering methods brings benefits in some applications, there are two main drawbacks that are discussed below, as well as possible solutions to them.
 
@@
 
-Introducing constraints in clustering changes the problem it solves, which becomes: \textit{Find the best partition that satisfies all constraints}.
-This way, if constraints are not properly specified or if the extraction methods are inadequate, we can find that the constraints contradict each other, which means that there is no partition that satisfies them all.
+The introduction of constraints into clustering changes the problem it solves, which becomes: \textit{Find the best partition that satisfies all constraints}.
+This way, if constraints are not properly specified or if the extraction methods are inadequate, we can end up with contradictory constraints, which means that there is no partition that satisfies them all.
 For instance, there is no partition that satisfies the constraints $C_=(x_i,x_j)$ and $C_{\neq}(x_i,x_j)$, regardless of the value of $k$.
@@
 
-As shown in Table \ref{tab:CCComplexity}, the use of \acs{CL} constraints increases the constrained clustering complexity level to $\mathbf{NP}$-complete and therefore constrained clustering is intractable.
-Intuitively it can be easily understood that if finding a single partition that satisfies the constraints is a complex problem, finding the best of them is harder.
+As shown in Table \ref{tab:CCComplexity}, the use of \acs{CL} constraints increases the complexity level of constrained clustering to $\mathbf{NP}$-complete and therefore constrained clustering is intractable.
+Intuitively, it can be easily understood that if finding a single partition that satisfies the constraints is a complex problem, finding the best of them is even harder.
 
@@
 
-	\textbf{Knowing a feasible solution exists does not help us find it.} The consequences of this result on the constrained clustering complexity imply that, even if there is a feasible partition, it will not be easy to find, speaking in terms of algorithmic complexity \cite{davidson2007survey}.
+	\textbf{Knowing a feasible solution exists does not help us find it.} The consequences of this result for the complexity of constrained clustering imply that, even if there is a feasible partition, it will not be easy to find in terms of algorithmic complexity \cite{davidson2007survey}.
 	\label{ob:FeasibleSolution}
@@
 
-In \cite{wagstaff2002intelligent} and \cite{davidson2007hierarchical} is shown that, even in the number of clusters in the output partition is set equal to the number of classes in the dataset ($k = k*$), which guarantees that a feasible solution does exists, simple algorithms like COP-kmeans \cite{wagstaff2001constrained} may not converge due to the feasibility problem.
+In \cite{wagstaff2002intelligent} and \cite{davidson2007hierarchical} is shown that, even if the number of clusters in the output partition is equal to the number of classes in the dataset ($k = k^*$), which guarantees that a feasible solution does exist, simple algorithms like \acf{COPKM} \cite{wagstaff2001constrained} may not converge due to the feasibility problem.
 
@@
 
-In constrained clustering it is assumed that constraints are a guide for an algorithm to find de desired data partition.
-So, it is reasonable to think that the more additional information (constraints) we have available, the closer the result we get to the one we are looking for, as Observation \ref{ob:AccuracyIncrease} stated.
-However, and despite the stipulations of this observation, we find cases in which, even generating the constraints without noise and on the basis of the true labels, there are constraint sets that, far from improving the results, worsen them considerably \cite{davidson2006proceedings}.
-This seems to disagree with the Observation \ref{ob:AccuracyIncrease}, however, let us remember that it refers to the average case, and not to particular cases.
+In constrained clustering it is assumed that constraints are a guide for an algorithm to find the desired data partition.
+So, it would be reasonable to think that the more additional information (constraints) we have available, the closer the result will be to the one we are looking for, as Observation \ref{ob:AccuracyIncrease} stated.
+However, and despite the stipulations of this observation, we find cases where, even generating the constraints without noise and according to the true labels, there are constraint sets that, far from improving the results, worsen them considerably \cite{davidson2006proceedings}.
+This seems to disagree with Observation \ref{ob:AccuracyIncrease}; however, let us remember that it refers to the average case, and not to particular cases.
 
@@
 The feasibility problem can be tackled in several ways.
-The most immediate may be to keep the number of constraints low, in proportion to the number of total instances, to minimize the likelihood of inconsistencies.
+The most immediate may be to keep the number of constraints low (in proportion to the number of total instances) to minimize the likelihood of inconsistencies.
 However, not being able to increase the number of constraints if the problem requires it is not the ideal scenario.
-Therefore, interest should be given to analyzing when a problem becomes over-constrained, since, as we have studied in Section \ref{sec:ConstraintsProblems}, even generating constraints based on the ground truth, algorithms such as COP-kmeans are no longer effective as the number of constraints to be satisfied increases, even if the algorithm is randomly restarted several times.
+Attention should be paid to analyzing when a problem becomes over-constrained, since, as studied in Section \ref{sec:ConstraintsProblems}, even when constraint generation is based on the ground truth, algorithms such as \acs{COPKM} are no longer effective (not even with random restarts) as the number of constraints to be satisfied increases.
 
@@
 It has been proven that constrained clustering with \acs{CL} is equivalent to the graph coloring problem \cite{davidson2006identifying}.
-Thus, we find that solving a problem with \acs{CL} constraints using algorithms such as COP-Kmeans is, for practical purposes, solving the graph coloring problem.
+Thus, we find that solving a problem with \acs{CL} constraints using algorithms such as \acs{COPKM} is, in practice, solving the graph coloring problem.
 
@@
 
-This result allows you to transfer many of the properties of the graph coloring problem to the constrained clustering problem.
+This result allows us to transfer many of the properties of the graph coloring problem to the constrained clustering problem.
 For example, Brook's theorem states that graph coloring is simple when the number of available colors ($k$ in our case) is greater than the maximum degree of the graph.
@@
 
-With this, and although Observation \ref{ob:FeasibleSolution} indicates differently, when constrained clustering meets the condition set out in Observation \ref{ob:BrooksTheorem}, we can guarantee that a solution will always be found in polynomial time.
-To ensure Brook's condition, it is possible to build the constraint set so that no instance takes a side in more than $k$ \acs{CL} constraints \cite{davidson2006identifying}.
+With this, and although Observation \ref{ob:FeasibleSolution} indicates differently, when constrained clustering meets the condition seen in Observation \ref{ob:BrooksTheorem}, we can guarantee that a solution will always be found in polynomial time.
+To ensure Brook's condition, it is possible to build the constraint set so that no instance appears in more than $k$ \acs{CL} constraints \cite{davidson2006identifying}.
 
@@
 
-The solution to this problem is simple: identify those truly useful constraint sets.
-However, this involves applying some kind of metric that allows us to evaluate when a given set of constraints meets this condition.
+The solution to this problem is simple: truly useful constraint sets must be identified.
+However, this involves applying some kind of metric to evaluate the compliance of a given set of constraints with this condition.
 To this end two measures are proposed: informativity and coherence.
@@
 The \textbf{informativity} measure refers to the amount of information present in the constraint set that an algorithm cannot determine for itself.
-For example, in Figure \ref{fig:Informativity}, an algorithm such as COP-kmeans would tend to group nearby instances and place those that are far away in separate clusters; however, constraints bias the solution space preventing this from happening.
+For example, in Figure \ref{fig:Informativity}, an algorithm such as \acs{COPKM} would tend to group nearby instances and place those that are far away in separate clusters; however, constraints bias the solution space preventing this from happening.
 
@@
 Informativity is estimated using the constraint set as a test set, so the ability of the algorithm to predict the constraints present in it is measured.
-Formalizing, given a set of constraints $CS$ and a constrained clustering algorithm $A$, we get the partition $P_A$ by applying the algorithm to the input dataset specifying an empty constraint set.
-We can compute the number of unsatisfied constraints in $P_A$ a in Equation \ref{eq:Informativity} \cite{davidson2007survey}.
+Formally, given a set of constraints $CS$ and a constrained clustering algorithm $A$, we get the partition $C_A$ by applying the algorithm to the input dataset specifying an empty constraint set.
+We can compute the number of unsatisfied constraints in $C_A$ a in Equation \ref{eq:Informativity} \cite{davidson2007survey}.
 
 \begin{equation}
-I_A(R) = \frac{1}{|R|}\left[ \sum_{r \in R} \text{unsat}(r, P_A) \right].
+I_A(CS) = \frac{1}{|CS|}\left[ \sum_{r \in CS} \text{unsat}(r, C_A) \right].
 \label{eq:Informativity}
@@
 For instance, Figure \ref{fig:Coherence} shows two parallel and very close constraints of different type.
-It is in cases like this that the contradiction occurs, as the \acs{ML} constraints indicate that distance between instances involved in them is small, while \acs{CL}s indicate the opposite.
+It is in cases like this that the contradiction occurs, as the \acs{ML} constraints indicate that the distance between instances involved in them is small, while \acsp{CL} indicate the opposite.
 
@@
 	\includegraphics[scale=0.4]{gfx/ConstClust/Coherencia/Coher2}
-	\caption[Coherence measure representation.]{Coherence measure representation. \cite{davidson2007survey}}\label{fig:CoherenceOverlap}
+	\caption[Coherence measure representation.]{Coherence measure representation \cite{davidson2007survey}.}\label{fig:CoherenceOverlap}
 \end{figure}
@@
 Constrained clustering adds a new type of information to the original clustering problem, which is given in the form of specifications of belonging to the same or to different clusters on pairs of instances.
-Constraints, whether if they are the pairwise \acf{ML} and \acf{CL} constraints, or the distance-based $\delta$ and $\epsilon$ constraints, they are used to guide the clustering method we apply to a dataset in the search for a result partition that meets certain desired properties.
+Constraints, whether they are the pairwise \acf{ML} and \acf{CL} constraints, or the distance-based $\delta$ and $\epsilon$ constraints, are used to guide the clustering method being applied to a dataset in its search for a result partition that meets certain desired properties.
 
-Clustering methods derived from this concept have proved to be advantageous in many areas, as well as they present some problems that can be solved by deeply studying the constraints to be used to solve each individual problem.
+Clustering methods derived from this concept have proved to be advantageous in many areas; in addition to this, they suffer from some issues that can be tackled by carefully studying the constraints used to solve each individual problem.

### 2019-07-06 Memoria/Chapters/ExperimentalResults.tex
Cambios en memoria y en paper DILS

@@
 Being the mentioned the main purpose of this chapter, we firstly introduce a brief study on the influence of the parameter $\xi$ in the \acs{DILS} method in Section \ref{sec:XiInfl}, so the reader gets a better comprehension of theory presented in Section \ref{sec:DILS}.
-Afterwards we will carry out an in-depth comparison between \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$ in Section \ref{sec:BRKGAvsSHADE}, as both are population-based and heuristic-guided methods.
-Finally, in Section \ref{sec:NewPropvsSOTA}  we will present a more general comparison that will include all state-of-the-art methods and the new proposals.
+Afterwards, we carry out an in-depth comparison between \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$ in Section \ref{sec:BRKGAvsSHADE}, as both are population-based and heuristic-guided methods.
+Finally, in Section \ref{sec:NewPropvsSOTA} a more general comparison that will include all state-of-the-art methods and the new proposals is presented.
 
-It is worth noting that, since the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another.
-To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.
+It is worth noting that, since the methods we are comparing involve non-deterministic behavior, the results may vary between runs.
+To attenuate the effect this may have on the results, we will apply each method five times to every dataset and constraint set, so that the resulting measures correspond to a five-run average.
 
@@
 
-Parameter $\xi$ from \acs{DILS} plays a key role in the development of its optimization process.
-It was theoretically introduced in Section \ref{sec:DILS}.
-The aim of this section is to prove its influence in the \acs{DILS} constrained clustering application (\acs{DILS}$_{CC}$).
-With this in mind we set up the following experiment: apply \acs{DILS}$_{CC}$ to the Iris dataset and its associated $CS_{10}$ constraint set and save the number of worst individual ($m_w$) reinitializations in each generations.
-We have to keep in mind that \acs{DILS}$_{CC}$ is a non deterministic algorithm, due to its \acf{ILS} procedure.
-To mitigate the effect this could have of our study we will repeat the experiment 20 times, with the objective of analyzing average results and standard deviation.
+The parameter $\xi$ of \acs{DILS}, which was theoretically introduced in Section \ref{sec:DILS}, plays a key role in the optimization process.
+The aim of this section is to prove its influence in the application of \acs{DILS} to constrained clustering (\acs{DILS}$_{CC}$).
+With this in mind, we set up the following experiment: apply \acs{DILS}$_{CC}$ to the Iris dataset and its associated $CS_{10}$ constraint set and save the number of worst individual ($m_w$) reinitializations in each generation.
+We have to keep in mind that \acs{DILS}$_{CC}$ is a non-deterministic algorithm.
+To mitigate the effect this could have on our study, the experiment is repeated twenty times with the aim of analyzing average results and standard deviation.
 
@@
 	\myfloatalign
-	\subfloat[Restarts per generation with $\xi \le 0$]
+	\subfloat[Restarts per generation with $\xi \le 0$.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMenor.pdf}} 
 	\hspace{1cm}
-	\subfloat[Restarts per generation with $\xi \ge 0$]
+	\subfloat[Restarts per generation with $\xi \ge 0$.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/RestartsXiMayor.pdf}}
-	\caption[Visual representation of the influence of $\xi$ over the DILS$_{CC}$ optimization process.]{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
+	\caption{Visual representation of the influence of $\xi$ over the \acs{DILS}$_{CC}$ optimization process.}
 	\label{fig:RestartsLines}
@@
 Figure \ref{fig:RestartsLines} shows the number of total reinitializations per generation for $\xi \in [-0.5,0.5]$.
-Note that the $y$ axis is on a continuous scale due to the averaging process.
-As we expected this number grows in a logarithmic way with respect to the number of generations, which means that the algorithm tends to converge if a sufficient number of generations.
-Also it is clear that the higher $\xi$ the less reinitializations per generations are done.
+Note that the $y$ axis is real-valued due to the averaging process.
+As expected, this number grows in a logarithmic fashion with respect to the number of generations.
+This means that the algorithm tends to converge.
+It is also clear that the higher $\xi$ is, the fewer reinitializations per generation are done.
 
@@
 	\myfloatalign
-	\subfloat[Hisogram showing values]
+	\subfloat[Histogram of values.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/HistValues.pdf}} 
 	\hspace{1cm}
-	\subfloat[Hisogram showing variance]
+	\subfloat[Histogram of variance.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpResults/HistVar.pdf}}
-	\caption{Mean number of reinitializations histogram for $\xi \in [-0.5,0.5]$.}
+	\caption{Histogram of the average number of reinitializations for $\xi \in [-0.5,0.5]$.}
 	\label{fig:RestartsHist}
@@
 
-Figure \ref{fig:RestartsHist} aims to present a more clear visual representation of the average number of reinitializations and the standard deviation found in the total number of reinitializations for the 20 runs and for each $\xi$ value.
-We observe how the average number of reinitializations grows in an irregular way as $\xi$ decreases.
-This effect would tend to lessen as we increase the size of the sample runs increases.
-To put it clear: the more runs the more regular the grow of the average number reinitializations would be.
-Regarding the variance, we can clearly see how the more permissive the reinitilization rule is the higher it gets.
-Which is exactly what we would expect in a non deterministic method such as \acs{DILS}$_{CC}$.
+Figure \ref{fig:RestartsHist} aims to present a clearer visual representation of the twenty-run average number of reinitializations and its standard deviation for each value of $\xi$.
+We can observe how the average number of reinitializations grows in an irregular way as $\xi$ decreases.
+This effect would tend to lessen as the size of the sample runs increases.
+To put it clear: the higher the amount of runs, the more regular the growth of the average number of reinitializations.
+Regarding the variance, we can clearly see how the more permissive the reinitialization rule is, the higher it gets; this is exactly what would be expected in a non-deterministic method such as \acs{DILS}$_{CC}$.
 
-We can conclude that parameter $\xi$ has the expected influence over the \acs{DILS}$_{CC}$ optimization process.
-Note that in a problem in which the \acs{LS} procedure used is deterministic we would also find differences between consecutive runs over the same datasets, given that the \acs{DILS} general optimization scheme uses stochastic tools---such as the recombination and mutation operators---to make sure that diversity is introduced in the search of the global optima for the problem we apply it to.
+We can conclude that $\xi$ has the expected influence on the \acs{DILS}$_{CC}$ optimization process.
+Note that, in a problem in which the \acs{LS} procedure used was deterministic, we would also find differences between consecutive runs over the same datasets; the general optimization scheme in \acs{DILS} uses stochastic tools---such as the recombination and mutation operators---to make sure that diversity is introduced in the search for problem-specific global optima.
 
-\section{Comparing BRKGA$_{CC}$ with SHADE$_{CC}$} \label{sec:BRKGAvsSHADE}
+\section[Comparing \acsfont{BRKGA}$_{CC}$ with \acsfont{SHADE}$_{CC}$]{Comparing BRKGA$_{CC}$ with SHADE$_{CC}$} \label{sec:BRKGAvsSHADE}
 
 Tables from \ref{tab:resultsBRKGAvsSHADE10} to \ref{tab:resultsBRKGAvsSHADE20} present the comparison between \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.
-In them the \acs{ARI} columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
+In them the \acs{ARI} columns show the Adjusted Rand Index for each case, the Unsat column displays the percentage of violated constraints and the Time columns show the time taken for each algorithm to produce results, measured in seconds.
 
@@
 
-	\caption[Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.]{Experimental results obtained for $CS_{10}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
+	\caption{Experimental results obtained for $CS_{10}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
 	\label{tab:resultsBRKGAvsSHADE10}
@@
 
-	\caption[Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.]{Experimental results obtained for $CS_{15}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
+	\caption{Experimental results obtained for $CS_{15}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
 	\label{tab:resultsBRKGAvsSHADE15}
@@
 Table \ref{tab:resultsBRKGAvsSHADE15} presents the results obtained with the $CS_{15}$ constraint set.
-While conclusions for Unsat and Time remain unchanged, we see a clear difference as far as \acs{ARI} is concerned.
+While conclusions for Unsat and Time remain unchanged, we can see a clear difference as far as \acs{ARI} is concerned.
 We observe that both methods already report \acs{ARI} results above 0.7, which in an \acs{ARI} ranking is considered to be an acceptable result.
-We can even observe that the \acs{SHADE}$_{CC}$ method provides an \acs{ARI} higher than 0.9 for 7 datasets, compared to \acs{BRKGA}$_{CC}$ that only obtains similar results in 2 datasets.
+We can even observe that \acs{SHADE}$_{CC}$ achieves an \acs{ARI} higher than 0.9 for seven datasets, compared to \acs{BRKGA}$_{CC}$ that only obtains similar results in two datasets.
 
@@
 
-	\caption[Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.]{Experimental results obtained for $CS_{20}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
+	\caption{Experimental results obtained for $CS_{20}$ comparing \acs{SHADE}$_{CC}$ and \acs{BRKGA}$_{CC}$.}
 	\label{tab:resultsBRKGAvsSHADE20}
@@
 It is in Table \ref{tab:resultsBRKGAvsSHADE20}, which gathers results obtained for the constraint set $CS_{20}$, that we observe major differences between the two methods.
-We find that \acs{SHADE}$_{CC}$ is capable of producing clusters identical to the original ones for 4 datasets, regardless the single dataset for which \acs{BRKGA}$_{CC}$ achieves it.
-Let us remember that an \acs{ARI} of 1 represents a perfect match between the two partitions given to compute it.
-This improvement in \acs{ARI} is clearly reflected in the percentage of violated constraints---Unsat column---, value for which, once again, \acs{SHADE}$_{CC}$ outperforms \acs{BRKGA}$_{CC}$.
+We find that \acs{SHADE}$_{CC}$ is capable of producing clusters identical to the original ones for four datasets, against the single dataset for which \acs{BRKGA}$_{CC}$ achieves it.
+Let us remember that an \acs{ARI} of 1 represents a perfect match between the two partitions for which it is computed.
+This improvement in \acs{ARI} is clearly reflected in the percentage of violated constraints (Unsat column) for which, once again, \acs{SHADE}$_{CC}$ outperforms \acs{BRKGA}$_{CC}$.
 Conclusions on execution time remain the same.
@@
 
-\section{Comparing SHADE$_{CC}$ and DILS$_{CC}$ with The-State-Of-The-Art} \label{sec:NewPropvsSOTA}
+\section[Comparing \acsfont{SHADE}$_{CC}$ and \acsfont{DILS}$_{CC}$ with The-State-Of-The-Art]{Comparing SHADE$_{CC}$ and DILS$_{CC}$ with the State-Of-The-Art} \label{sec:NewPropvsSOTA}
 
-In this section we present Tables from \ref{tab:resultsSOTA10} to \ref{tab:resultsSOTA20}, which display the results obtained by the new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---and six state-of-the-art methods---including \acs{BRKGA}$_{CC}$---to be compared for each dataset and constraint set.
+In this section we present Tables from \ref{tab:resultsSOTA10} to \ref{tab:resultsSOTA20}, which display the results obtained by the new proposals---\acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$---and six state-of-the-art methods including \acs{BRKGA}$_{CC}$ to be compared for each dataset and constraint set.
 
-It should be noted that there are some missing results in these tables.
-In the case of the \acs{COPKM} algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
-It is possible that \acs{COPKM} cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels.
-In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
-We establish that in these cases the \acs{ARI} value considered for the mean calculation---and the forthcoming statistical analysis of results---is -1.000, which is the worst value the \acs{ARI} measure can deliver.
+It should be noted that there are some missing values in these tables.
+In the case of \acs{COPKM}, this is due to the fact that it is highly dependent on the order in which constraints are analyzed.
+It is possible that \acs{COPKM} cannot find a solution, even though it is always feasible since the constraints have been generated based on the true labels.
+In the case of \acs{CECM}, some of the results are not available because the in-memory data structures that hold the algorithm grow non-linearly with the number of classes and the number of features.
+We establish that in these cases the \acs{ARI} value considered for the mean calculation---and the forthcoming statistical analysis of results---is $-1.000$, which is the worst possible \acs{ARI} value.
 
@@
 Table \ref{tab:resultsSOTA10} presents the results obtained for the $CS_{10}$ constraint set.
-\acs{LCVQE}, \acs{RDPM} and \acs{TVClust} methods stand out of \acs{SHADE}$_{CC}$, achieving results as competitive as it with a low number of constraints.
-Nevertheless \acs{DILS}$_{CC}$ represents a consistent improvement over all other 6 state-of-the-art methods.
+\acs{LCVQE}, \acs{RDPM} and \acs{TVClust} stand out over \acs{SHADE}$_{CC}$, achieving comparable results with a low number of constraints.
+Nevertheless, \acs{DILS}$_{CC}$ represents a consistent improvement over all six state-of-the-art methods.
 This is due to the fact that the exploitation capacity of \acs{DILS}$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lowest level of constraint-based information we work with.
@@
 
-In one hand, Table \ref{tab:resultsSOTA15}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in precision in the results obtained by \acs{SHADE}$_{CC}$.
-On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that presented by \acs{SHADE}$_{CC}$, with the exception of \acs{COPKM}.
-Also we observe how \acs{DILS}$_{CC}$ and \acs{COPKM} are able to obtain \acs{ARI} values equal to 1, meaning that the resulting partitions of the datasets perfectly match the true partitions.
-The fact that \acs{DILS}$_{CC}$ is always able to output a partition of the datasets is a noteworthy advantage over \acs{COPKM}.
+On one hand, in Table \ref{tab:resultsSOTA15}, which shows the results for the $CS_{15}$ constraint set, we observe that there is a significant improvement in accuracy in the results obtained by \acs{SHADE}$_{CC}$.
+On the other hand, methods that obtained better results for smaller constraint sets do not present an improvement comparable to that of \acs{SHADE}$_{CC}$ (with the exception of \acs{COPKM}).
+We can also observe how \acs{DILS}$_{CC}$ and \acs{COPKM} are able to obtain \acs{ARI} values equal to 1, meaning that the resulting dataset partitions perfectly match the true partitions.
+The fact that \acs{DILS}$_{CC}$ is always able to output a partition is a noteworthy advantage over \acs{COPKM}.
 Even in the cases where \acs{COPKM} is unable to produce a partition, \acs{DILS} often finds the optimal or a near-optimal partition.
@@
 Table \ref{tab:resultsSOTA20} presents the results obtained for $CS_{20}$.
-For \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$ we continue to observe its progression in accuracy, as well as for methods such as \acs{BRKGA}$_{CC}$ or \acs{COPKM} (the latter being more limited in other aspects).
-As far as the state-of-the-art methods are concerned, we find a general improvement in \acs{ARI}, except for CECM.
-Even with this, it is \acs{SHADE}$_{CC}$ and \acs{COPKM} that seem to obtain the highest quality results, with the difference that \acs{SHADE}$_{CC}$ is able to produce results in all cases.
-It is also worth noting how \acs{DILS}$_{CC}$ provides the best average \acs{ARI} for each of the three constraint sets analyzed in this paper.
+For \acs{SHADE}$_{CC}$ and \acs{DILS}$_{CC}$ we continue to observe their progression in accuracy, as well as for methods such as \acs{BRKGA}$_{CC}$ or \acs{COPKM}---the latter being more limited in other aspects.
+As far as the state-of-the-art methods are concerned, we find a general improvement in \acs{ARI}, except for \acs{CECM}.
+The \acs{TVClust} method should be highlighted, since it is able to find the optimal solution in three cases, which implies a significant improvement over its performance on previous constraint sets.
+Even with this, it is \acs{SHADE}$_{CC}$ and \acs{COPKM} that seem to obtain the highest quality results, with the difference being that \acs{SHADE}$_{CC}$ is able to produce results in all cases.
+It is also worth noting how \acs{DILS}$_{CC}$ provides the best average \acs{ARI} for each of the three constraint sets analyzed here.
 
@@
 
-All comparisons carried out in this sections are merely supported by the raw results.
-To prove whether the new proposals represent an improvement over the state-of-the-art methods it is necessary to perform an statistic-based empirical analysis of the results.
+All comparisons carried out in these sections are merely supported by the raw results.
+To prove whether the new proposals represent an improvement over the state-of-the-art methods it is necessary to perform a statistics-based empirical analysis of the results.
 The last chapter (Chapter \ref{ch:AnalysisResults}) of Part \ref{pt:Experiments} will address this question.

### 2019-07-06 Memoria/Chapters/ExperimentalSetup.tex
Cambios en memoria y en paper DILS

@@
 
-In this chapter we present the experimental setup used to compare the two new proposals introduced in Chapter \ref{ch:NewProposals} with the current state of the art.
-We will start by briefly introducing the state-of-the-art methods in Section \ref{sec:SOTA}.
-Afterwards we specify the dataset used in experimentation (Section \ref{sec:Datasets}), as well as the constraint sets (Section \ref{sec:ConstGen}) and the results evaluation method (Section \ref{sec:EvalMet}).
-In Section \ref{sec:ValidtnMethod} we introduce the validation method we will apply to the results.
-We also include the required specifications to replicate the experiments in Section \ref{sec:Calibration}.
+In this chapter we present the experimental setup used to compare the two new proposals introduced in Chapter \ref{ch:NewProposals} with the current state-of-the-art.
+We start by briefly introducing the state-of-the-art methods in Section \ref{sec:SOTA}.
+Afterwards, we specify the dataset used in experimentation (Section \ref{sec:Datasets}), as well as the constraint sets (Section \ref{sec:ConstGen}) and the result evaluation method (Section \ref{sec:EvalMet}).
+The validation method that will be applied to the results in later chapters (Chapter \ref{ch:AnalysisResults} is discussed in Section \ref{sec:ValidtnMethod}).
+We also include the required configuration to replicate the experiments in Section \ref{sec:Calibration}.
 
@@
 The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}.
-It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL.
+It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known \acs{ML} and \acs{CL}.
 This method was named \acf{COPKM}, it introduces a modification to the assignation rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.
 
-In \cite{antoine2012cecm} \acf{CECM}, a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
+In \cite{antoine2012cecm} \acf{CECM}, a variant of the Evidential c-means (\acsfont{ECM} \cite{masson2008ecm}) algorithm is proposed, within the fuzzy clustering family of methods.
 The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function.
@@
 
-A modification of the Constrained Vector Quantization Error algorithm (CVQE \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}.
-The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity.
-\acf{LCVQE} introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex.
+A modification of the Constrained Vector Quantization Error algorithm (\acsfont{CVQE} \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}.
+The \acsfont{CVQE} algorithm proved to produce high quality results, at the cost of a very high computational complexity.
+\acf{LCVQE} introduces a modification of the cost function of \acsfont{CVQE} to make it more intuitive and less computationally complex.
 The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.
@@
 
-For our experiments we will compare the results obtained by \acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$ with those obtained by state-of-the-art methods and \acs{BRKGA}$_{CC}$ over 20 datasets and 3 constraint sets for each one of them.
+For our experiments we will compare the results obtained by \acs{DILS}$_{CC}$ and \acs{SHADE}$_{CC}$ with those obtained by state-of-the-art methods and \acs{BRKGA}$_{CC}$ over twenty datasets and three constraint sets for each one of them.
 Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
 \href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
+We also include three artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
 Table \ref{tab:datasets} displays a summary of every dataset.
@@
 Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets.
-This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.
+This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an \acs{ML} or \acs{CL} constraint depending on whether the labels are the same or different.
 
@@
 			\cline{2-3} \cline{5-6} \cline{8-9}
-			& ML & CL && ML & CL && ML & CL \\
+			& \acs{ML} & \acs{CL} && \acs{ML} & \acs{CL} && \acs{ML} & \acs{CL} \\
 			\hline
@@
 The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it.
-Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
+Note that the greater the number of classes present in the dataset, the fewer \acs{ML} constraints obtained with the method proposed in \cite{wagstaff2001constrained}.
 This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.
@@
 	\myfloatalign
-	\subfloat[$CS_{10}$ \acs{ML} constraints]
+	\subfloat[$CS_{10}$ \acs{ML} constraints.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML10.png}} 
-	\subfloat[$CS_{10}$ \acs{CL} constraints]
+	\subfloat[$CS_{10}$ \acs{CL} constraints.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL10.png}} \quad
-	\subfloat[$CS_{15}$ \acs{ML} constraints]
+	\subfloat[$CS_{15}$ \acs{ML} constraints.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML15.png}} 
-	\subfloat[$CS_{15}$ \acs{CL} constraints]
+	\subfloat[$CS_{15}$ \acs{CL} constraints.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL15.png}} \quad
-	\subfloat[$CS_{20}$ \acs{ML} constraints]
+	\subfloat[$CS_{20}$ \acs{ML} constraints.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/ML20.png}} 
-	\subfloat[$CS_{20}$ \acs{CL} constraints]
+	\subfloat[$CS_{20}$ \acs{CL} constraints.]
 	{\includegraphics[width=0.45\linewidth]{gfx/ExpSetup/CL20.png}} 
-	\caption{Visual representation of the three constraints sets the Iris dataset.}
+	\caption{Visual representation of the three constraint sets over the Iris dataset.}
 	\label{fig:IrisConst}
@@
 
-Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper.
+Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this work.
 To accomplish this task we just set one of the two partitions given to compute \acs{ARI} as the ground truth labels.
 
-\section{Validation of results} \label{sec:ValidtnMethod}
+\section{Validation of Results} \label{sec:ValidtnMethod}
 
@@
 \textit{"In a nutshell: \acs{NHST} do not answer the question we ask"}.
-To put it clear, the disadvantages of the \acs{NHST} that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?
+To put it clear, the disadvantages of the \acs{NHST} that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: To reject, or not to reject?
 
-To start with, \acs{NHST} do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different?
+To start with, \acs{NHST} do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: What is the probability that two methods are different?
 Another pitfall of \acs{NHST} is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis.
@@
 	%\resizebox{\textwidth}{!}{
-		\begin{tabular}{>{\centering\arraybackslash}c >{\arraybackslash}L{5cm} cc}
+		\begin{tabular}{c l cc}
 			\hline
@@
 			Evals & Fitness function evaluations & 300000 & 300000 \\
-			$P_e$ & Size of the elite set in population & $0.2 * |P|$ & $0.25 * |P|$ \\
-			$P_m$ & Number of mutants to be introduced in the population in each generation & $0.2 * |P|$ & - \\
-			$p_\text{inherit}$ & Probability that a feature is inherited from an elite parent & $50\%$ & - \\
-			$k$ & Output partition number of clusters & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
+			$P_e$ & \makecell[l]{Size of the elite set in \\ population} & $0.2 * |P|$ & $0.25 * |P|$ \\
+			$P_m$ & \makecell[l]{Number of mutants to be \\ introduced in the population in \\ each generation} & $0.2 * |P|$ & - \\
+			$p_\text{inherit}$ & \makecell[l]{Probability that a feature is \\ inherited  from an elite parent} & $50\%$ & - \\
+			$k$ & \makecell[l]{Output partition number of \\ clusters} & \multicolumn{2}{m{4cm}}{No. Classes (Table \ref{tab:datasets})} \\
 			\hline
@@
 
-	\caption[Parameters setup used for BRKGA$_{CC}$ and SHADE$_{CC}$]{Parameters setup used for \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.}
+	\caption{Parameter setup used for \acs{BRKGA}$_{CC}$ and \acs{SHADE}$_{CC}$.}
 	\label{tab:paramsSHADE}
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabular}{c L{6cm} c}
+		\begin{tabular}{c l c}
 			\hline
@@
 			Evals & Fitness function evaluations & 300000 \\
-			$p_s$ & Segment size for the mutation operator & $0.3 \times n$ \\
-			$p_m$ & Maximum number of neighbors generated in each call to the LS procedure & Evals $\times 0.01$ \\
-			$p_r$ & Probability that a feature is selected from $m_b$ for recombination & $0.3$ \\
+			$p_s$ & \makecell[l]{Segment size for the mutation \\ operator} & $0.3 \times n$ \\
+			$p_m$ & \makecell[l]{Maximum number of neighbors \\ generated in each call to the \acs{LS} \\ procedure} & Evals $\times 0.01$ \\
+			$p_r$ & \makecell[l]{Probability that a feature is selected \\ from $m_b$ for recombination} & $0.3$ \\
 			$\xi$ & Reinitialization method tolerance & $0.2$ \\
@@
 
-	\caption[Parameters setup used for DILS$_{CC}$]{Parameters setup used for \acs{DILS}$_{CC}$.}
+	\caption{Parameter setup used for \acs{DILS}$_{CC}$.}
 	\label{tab:paramsDILS}
@@
 
-	\caption{Parameters setup used for the state-of-the-art algorithms.}
+	\caption{Parameter setup used for the state-of-the-art algorithms.}
 	\label{tab:paramsSOTA}
@@
 Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.
-All parameters described above are meant to be used for the Python implementation that can be found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}, which will be used to obtain our results.
+Equivalent results can be obtained with the parameters described above when used in the Python implementation found at \href{https://github.com/GermangUgr}{GitHub}\footnote{https://github.com/GermangUgr}.

### 2019-07-06 Memoria/Chapters/Introduction.tex
Cambios en memoria y en paper DILS

--- 2019-06-29 Memoria/Chapters/Introduction.tex
@@
 \begin{quotation}{\slshape
-		An intelligent being cannot treat every object it sees as unique entity unlike anything else in the universe.
-It has to put objects in categories so that it may apply its hard-won knowledge about similar objects encountered in the past, to he object at hand.}
+		An intelligent being cannot treat every object it sees as a unique entity unlike anything else in the universe.
+It has to put objects in categories so that it may apply its hard-won knowledge about similar objects encountered in the past, to the object at hand.}
 		\begin{flushright}
@@
 One of the most basic and primitive abilities that human beings are endowed with is being able to group similar objects to produce a classification that is useful to them.
-For example, they should have been able to realize which objects were edible, which were poisonous, and which would try to kill them.
+For example, they were able to realize which objects were edible, which were poisonous, and which would try to kill them.
 
@@
 Aristotle built an elaborated animal classification system, dividing all creatures into two groups: red-blooded and not red-blooded.
-Later he proposed a subdivision that classified them according to the way in which new individuals came into the world, either alive, by means of eggs, chrysalis, etc.
+Later he proposed a subdivision that classified them according to the way in which new individuals came into the world (either alive, by means of eggs, chrysalis, etc).
 
-Following Aristotle, Theophrastus wrote the first document compiling guidelines for the classification of plants.
+Following Aristotle, Theophrastus wrote the first document that compiled guidelines for the classification of plants.
 The resulting works were so extensive and detailed that they laid the foundation for research in biology over the following centuries.
@@
 In particular, this classification laid the foundation for the development of Darwin's theory of evolution.
-But it has also been of high relevance in knowledge areas such as chemistry and physics, with the classification of the elements in the periodic table, proposed by Mendeleyev in the 1860s; or in astronomy, with the classification of dwarf or giant stars using the Hertzsprung-Russell guidelines, shown in Figure \ref{fig:intro_HRdiagram}.
+But it has also been highly relevant in knowledge areas such as chemistry and physics, with the classification of the elements in the periodic table, proposed by Mendeleyev in the 1860s; or in astronomy, with the classification of dwarf or giant stars using the Hertzsprung-Russell guidelines, shown in Figure \ref{fig:intro_HRdiagram}.
 
@@
 	\includegraphics[scale=0.35]{gfx/Intro/HR_diagram} 
-	\caption[Hertzsprung-Russell stars classification diagram.]{Hertzsprung-Russell stars classification diagram.}\label{fig:intro_HRdiagram}
+	\caption{Hertzsprung-Russell stars classification diagram.}\label{fig:intro_HRdiagram}
 \end{figure}
@@
 We humans, like many other beings, are capable of learning a multitude of concepts throughout our lives, although generally the learning process is extremely slow.
-It takes us about six years to be able to begin dedicated and specific learning in primary education, and another eighteen years---generally speaking---to become computer scientists or experts in any other branch of knowledge.
+It takes us about six years to be able to begin dedicated and specific learning in primary education, and another eighteen years (generally speaking) to become computer scientists or experts in any other branch of knowledge.
 
@@
 So, do we want to teach machines to learn the way humans do?
-Machines have proven to be outstanding at repetitive tasks, as well as they reach calculation speeds that humans can only dream of.
-It is worth asking if it is possible to develop a learning method for machines that is orders of magnitude faster than the human and equal or more effective.
-Although there is a major impediment to the development of such a method: we do not know if it exists!
-The search for a method to avoid the disadvantages of the human learning process is what drives research in the science we know today as \textit{machine learning}.
+Machines have proven to be outstanding at repetitive tasks, and they also reach calculation speeds that humans can only dream of.
+It is worth asking if it is possible to develop a learning method for machines that is orders of magnitude faster than the human method and equally or more effective.
+Nevertheless, there is a major impediment to the development of such a method: we do not know whether it exists!
+The search for a method to avoid the disadvantages of the human learning process is what drives research in the field of science we know today as \textit{machine learning}.
 
-One of the major differences between human learning and machine learning, and we must admit that to this respect machines again lead us, is the way in which knowledge is transferred.
-In machines it is enough to copy and paste, that is to say, once a program performs correctly the task for which it has been designed in a computer, it performs it in all of them.
+One of the major differences between human learning and machine learning, and we must admit that in this respect machines are again ahead of us, is the way in which knowledge is transferred.
+In machines it is enough to copy and paste: once a program correctly performs the task for which it has been designed in a computer, it can perform it in all of them.
 Just copy and paste.
-In humans this is not the case, we cannot extract knowledge from one brain and insert it into another, we don't even know how it is encoded!
+In humans this is not the case.
+We cannot extract knowledge from one brain and insert it into another.
+We don't even know how it is encoded!
 We humans must use our inefficient input-output devices to transmit knowledge.
@@
 
-So far, references have been made to learning without formally defining it.
+So far, references have been made to learning without formally providing a definition.
 The concept of learning is deeply anchored in human beings.
-However, when trying to apply learning to computers, which are still no more than glorified calculators, we must give some kind of definition: \textit{Learning denotes changes in the system that are adaptive in the sense that they enable the system to do the same task or tasks drawn from the same population more efficiently and more effectively the next time} \cite{Michalski:2013:MLA:2588013}.
+However, when trying to apply learning to computers, which are still no more than glorified calculators, the following definition should suffice: \textit{Learning denotes changes in the system that are adaptive in the sense that they enable the system to do the same task or tasks drawn from the same population more efficiently and more effectively the next time} \cite{Michalski:2013:MLA:2588013}.
 
 Machine learning is a field derived from computer science, and a branch of \acf{AI}, whose ultimate goal is to develop methods that make it possible for machines to learn.
-In other words, the techniques embedded in machine learning aim to develop programs that guide computers to learn from a set of examples.
-Machine learning is part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work and help them.
+In other words, the techniques embedded in machine learning aim to develop programs that guide computers in learning from a set of examples.
+Machine learning is a part of \acs{AI} in the sense that \acs{AI} is often directed at simulating human beings behavior through computers, so that we can better understand how humans work.
 
@@
 
-There are several approaches within the field of machine learning, these are classified according to the information that is provided to the machine for it to learn.
-Thus, following the classification proposed in \cite{ayodele2010types}, we find four different types of learning:
+There are several approaches within the field of machine learning, which are classified according to the information that is provided to the machine for its learning process.
+Thus, following the classification proposed in \cite{ayodele2010types}, we have four different types of learning:
 
@@
 
-	\item \textbf{Supervised Learning}: complete information is provided for learning.
-The machine is provided with a set of examples and the class (label) associated with each of them.
-The goal is to generate a function that approximates the behavior of the function that maps the examples in their respective classes.
+	\item \textbf{Supervised Learning}: complete information is available for learning.
+The machine is provided with a set of examples and the class label associated with each of them.
+The goal is approximate the behavior of the function that maps the examples to their respective classes.
 
-	\item \textbf{Unsupervised Learning}: in this case the class to which the examples belong to is not available. The goal is to extract the class from the information contained in the set of examples.
+	\item \textbf{Unsupervised Learning}: in this case the class which the examples belong to is not available. The goal is to deduce the class from the information contained in the set of examples.
 
 	\item \textbf{Semi-supervised Learning}: it arises from adding incomplete information to unsupervised learning.
-This information is usually given in the form of a subset of labeled data, but there is no restrictions to this concern.
+This information is usually given in the form of a subset of labeled data, but there are no restrictions in this regard.
 
-	\item \textbf{Reinforcement Learning}: the algorithm learns how to act under the influence of its environment, which provides feedback on the actions taken by the algorithm.
+	\item \textbf{Reinforcement Learning}: the algorithm learns how to act within its environment, which provides feedback on the actions taken by the algorithm.
 
@@
 The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
-The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.
+The second tries to obtain better-defined clusters incorporating background information to the clustering process; this is known as constrained clustering and is the main subject of the study presented in this work \cite{triguero2015self}.
+
+\section{Objectives}
+
+The main objective of this work is to provide an overview of the concepts and particularities related to constrained clustering.
+In order for the new approaches to constrained clustering presented in this work to be properly understood, some of these concepts must be addressed in sufficient depth.
+It is also in our best interest to provide an example of proper experimental methodology to obtain relevant and subjectivity-free results and conclusions.
 
 \section{Structure}
+
+Regarding the organization of this work, background knowledge concerning both classic and constrained clustering are presented in Chapters \ref{ch:IntroClustering} and \ref{ch:ConstrainedClustering} respectively.
+In Chapter \ref{ch:NewProposals} two new approaches to the constrained clustering problem are discussed.
+The experimental setup and the experimental results of the aforementioned techniques are presented in Chapters \ref{ch:ExperimentalSetup} and \ref{ch:ExperimentalResults} respectively.
+Afterward, an empirical analysis of these results is performed in Chapter \ref{ch:AnalysisResults}.
+Finally, conclusions and future work are discussed in Chapter \ref{ch:ConclFW}.

### 2019-07-06 Memoria/Chapters/NewProposals.tex
Cambios en memoria y en paper DILS

@@
 
-Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem.
-As we have already mentioned in Section \ref{sec:FeasibilityProblem}, the incorporation of constraints may modify the complexity of the clustering problem, making it $\mathbf{NP}$-complete and therefore intractable \cite{davidson2005clustering}.
+Finding the optimal partition in a dataset according to any kind of reasonable criteria is known to be a $\mathbf{NP}$-hard problem.
+As we have already mentioned in Section \ref{sec:FeasibilityProblem}, the use of constraints may modify the complexity of the clustering problem, making it $\mathbf{NP}$-complete and therefore intractable \cite{davidson2005clustering}.
 Bearing this in mind, in this chapter we propose two new approaches to find approximate solutions to the constrained clustering problem.
@@
 In this section we propose the first \acs{DE} application to find quality solutions for the constrained clustering problem.
-In particular, we will take as basis for our proposal one of the \acs{DE} variants which showed excellent behavior in worldwide competitions, known as SHADE \cite{molina2018insight}.
+In particular, we will take as basis for our proposal one of the \acs{DE} variants which showed excellent behavior in worldwide competitions, known as \acs{SHADE} \cite{molina2018insight}.
 We will make use of the Random-key concept from \acs{BRKGA}, along with proposing a new fitness function, to build the already mentioned \acs{DE} application.
 
-\subsection{A Glimpse to Differential Evolution}
+\subsection{A Glimpse into Differential Evolution}
 
@@
 
-To guide the optimization process, we use a fitness function.
+To guide the optimization process, we use a fitness function $f(\cdot)$.
 The task of \acs{DE} is to seek the parameter vector $p_i^*$ that minimizes a function such that $f(p_i)(f: \Omega \subseteq \mathfrak{R}^{v} \rightarrow \mathfrak{R})$, where $f(p_i^*) < f(p_i)$ for all $p_i \in \Omega$ with $\Omega$ a nonempty finite set that is the search domain.
@@
 	\includegraphics[scale=0.3]{gfx/NewProp/SHADE/DEop.png}
-	\caption[The optimization process of DE.]{The optimization process of \acs{DE} \cite{das2011differential}.}\label{fig:DE}
+	\caption[The optimization process of \acs{DE}.]{The optimization process of \acs{DE} \cite{das2011differential}.}\label{fig:DE}
 \end{figure}
@@
 Even though \acs{DE} was a major breakthrough in terms of real optimization, many variants have been developed since its invention that provide better results.
-We will focus on the SHADE variant, which adaptively optimizes some of the parameters of \acs{DE}.
-To carry out this task SHADE uses a historic record of values for these parameters that allows it to set them in favor of the optimization process, based on how good the algorithm behaved with certain parameter values in the past (see Section \ref{sec:SHADE}).
+We will focus on the \acs{SHADE} variant, which adaptively optimizes some of the parameters of \acs{DE}.
+To carry out this task \acs{SHADE} uses a historic record of values for these parameters that allows it to set them in favor of the optimization process, based on how good the algorithm behaved with certain parameter values in the past (see Section \ref{sec:SHADE}).
 
-\subsection{Brief Review of the BRKGA Algorithm} \label{sec:brkga}
+\subsection[Brief Review of the \acsfont{BRKGA} Algorithm]{Brief Review of the BRKGA Algorithm} \label{sec:brkga}
 
@@
 The mutation operator is used by \acsp{GA} to avoid local optima, and it consists in introducing random changes into the population to favor diversity.
-The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function that assigns a value to each chromosome to assess its quality as a solution for the problem.
+The selection operator is used to determine whether a chromosome is suitable to survive for the next generation, and to accomplish this task it employs a fitness function $f(\cdot)$ that assigns a value to each chromosome to assess its quality as a solution for the problem.
 \acsp{GA} apply these operators through a series of generations to hopefully obtain a solution for the problem to solve that is close to the optimal \cite{fausto2019ants}.
 
-\acf{RKGA} emerged in \cite{bean1994genetic} as a solid approach to high constrained problems.
-It uses vectors of random numbers to represent solutions to the problem (individuals), which are used as keys to obtain the actual solutions.
+\acf{RKGA} emerged in \cite{bean1994genetic} as a solid approach to highly constrained problems.
+It uses vectors of random numbers to represent solutions to the problem (individuals) which are used as keys to obtain the actual solutions.
 Random-key vectors are sampled from the space $[0,1]^{v}$, which is used as a surrogate space for the literal solution space.
@@
 Each random-key vector $p_i$ has $v$ random-keys in it.
-The population is sorted by the fitness value $f_i$ of each $p_i$ to select the first $P_e$ individuals, this is, the elite of the population.
+The population is sorted by the fitness value $f(p_i)$ of each individual to select the first $P_e$ individuals, this is, the elite of the population.
 The elite will be preserved in the next generation without modification.
@@
 
-\subsection{BRKGA optimization scheme for Constrained Clustering} \label{sec:AdaptationofBRKGA}
+\subsection[\acsfont{BRKGA} Optimization Scheme for Constrained Clustering]{BRKGA Optimization Scheme for Constrained Clustering} \label{sec:AdaptationofBRKGA}
 
@@
 The decoder divides the interval $[0,1]$ in $k$ intervals, so there exists a correspondence between each random-key (instance $x_i$) and the integer (label $l_i$) corresponding to the interval which it lies in.
-Table \ref{tab:decodingrk} shows an example of random-key decoding for a dataset with 10 instances ($n$ = 10) and $k = 3$.
+Table \ref{tab:decodingrk} shows an example of random-key decoding for a dataset withteninstances ($n$ = 10) and $k = 3$.
 Note that extreme values 0 and 1 can also appear in a random-key vector.
@@
 
-Once the decoded solution has been obtained, the fitness value of the solution is computed as in Equation \eqref{eq1}.
+Once the decoded solution has been obtained, the fitness value of the solution (individual $p_i$) is computed as in Equation \eqref{eq1}.
 
 \begin{equation}
-f_i = z_i + \overbrace{(\mu * n * \text{infeasibility}_i)}^\text{penalty},
+f_1(p_i) = z_i + \overbrace{(\mu * n * \text{infeasibility}_i)}^\text{penalty},
 \label{eq1}
@@
 Thus, the individuals produced by the \acs{LS} are only used to update the best solution found so far if needed.
-Figure \ref{fig:BRKGA_CC} shows a visual representation of the overall BRKGA optimization process for constrained clustering.
+Figure \ref{fig:BRKGA_CC} shows a visual representation of the overall \acs{BRKGA} optimization process for constrained clustering.
 For more details on \acs{BRKGA}$_{CC}$ see \cite{de2017comparison}.
+Arrows in red indicate population sorting by fitness.
 
@@
 	\includegraphics[scale=0.31]{gfx/NewProp/SHADE/BRKGA_Diagram.jpg} 
-	\caption[BRKGA$_{CC}$ optimization process for constrained clustering.]{\acs{BRKGA}$_{CC}$ optimization process for constrained clustering. \cite{de2017comparison}.}\label{fig:BRKGA_CC}
+	\caption[\acs{BRKGA}$_{CC}$ optimization process for constrained clustering.]{\acs{BRKGA}$_{CC}$ optimization process for constrained clustering \cite{de2017comparison}.}\label{fig:BRKGA_CC}
 \end{figure}
 
-\section{The SHADE Optimization Method} \label{sec:SHADE}
+\subsection[The \acsfont{SHADE} Optimization Method]{The SHADE Optimization Method} \label{sec:SHADE}
 
-\acf{SHADE} is an optimization method based on \acs{DE}. Proposed in \cite{tanabe2013success}, it was an improvement for the JADE model, presented in \cite{zhang2009jade}.
+\acf{SHADE} is an optimization method based on \acs{DE}. Proposed in \cite{tanabe2013success}, it was an improvement for the \acsfont{JADE} model, presented in \cite{zhang2009jade}.
 
-Unlike JADE, which only considers the parameters used in the previous generation, \acs{SHADE} uses a historical record of successful parameters as a mechanism for adapting parameters involved in the process of creating new generations.
+Unlike \acsfont{JADE}, which only considers the parameters used in the previous generation, \acs{SHADE} uses a historical record of successful parameters as a mechanism for adapting parameters involved in the process of creating new generations.
 
@@
 
-\subsection{SHADE Operators}
+\subsubsection[\acsfont{SHADE} Operators]{SHADE Operators}
 
@@
 
-\subsection{SHADE Parameter Adaptation Method}
+\subsubsection[\acsfont{SHADE} Parameter Adaptation Method]{SHADE Parameter Adaptation Method}
 
@@
 	\end{tabular}}
-	\caption[Historical memory $M_{CR}$, $M_{F}$ used by SHADE.]{Historical memory $M_{CR}$, $M_{F}$ used by \acs{SHADE} \cite{tanabe2013success}.}
+	\caption[Historical memory $M_{CR}$, $M_{F}$ used by \acs{SHADE}.]{Historical memory $M_{CR}$, $M_{F}$ used by \acs{SHADE} \cite{tanabe2013success}.}
 	\label{tab:SHADEmemory}
@@
 
-\subsection{SHADE optimization scheme for constrained clustering} \label{sec:SHADEadapt}
+\subsection[\acsfont{SHADE} Optimization Scheme for Constrained Clustering]{SHADE Optimization Scheme for Constrained Clustering} \label{sec:SHADEadapt}
 
@@
 \begin{equation}
-f_i = z_i * (\text{infeasibility}_i + 1).
+f_2(p_i) = z_i * (\text{infeasibility}_i + 1).
 \label{eq14}
@@
 To make these concepts clear we consider a toy dataset and three partitions over it.
-The dataset from Figure \ref{fig:toydatasets} contains 100 instances and a single ML constraint between two instances from the same class.
+The dataset from Figure \ref{fig:toydatasets} contains 100 instances and a single \acs{ML} constraint between two instances from the same class.
 Partitions $C_0$---displaying the true labels---and $C_1$ do not violate the constraint ($\text{infeasibility} = 0$), whereas $C_2$ does violate it ($\text{infeasibility} = 1$).
-However, $C_2$ shows a possibility, for later iterations, of adding both instances to their nearest cluster while also satisfying the ML constraint.
+However, $C_2$ shows a possibility, for later iterations, of adding both instances to their nearest cluster while also satisfying the \acs{ML} constraint.
 
@@
 	{\includegraphics[width=.41\linewidth]{gfx/NewProp/SHADE/C2}} \quad
-	\caption[Three partitions over a toy dataset.]{Three partitions over a toy dataset.}
+	\caption{Three partitions over a toy dataset.}
 	\label{fig:toydatasets}
@@
 
-	\caption[Expression and value of fitness functions over three partitions ($\mu = 10$).]{Expression and value of fitness functions over three partitions ($\mu = 10$).}
+	\caption{Expression and value of fitness functions over three partitions ($\mu = 10$).}
 	\label{tab:fitnessfunctions}
@@
 Figure \ref{fig:SHADE_CC} shows a representation of the \acs{SHADE}$_{CC}$ optimization process.
+Arrows in red indicate population sorting by fitness.
 
@@
 	\centering
-	\includegraphics[scale=0.31]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
-	\caption[SHADE$_{CC}$ optimization process for constrained clustering.]{\acs{SHADE}$_{CC}$ optimization process for constrained clustering.}\label{fig:SHADE_CC}
+	\includegraphics[scale=0.3]{gfx/NewProp/SHADE/SHADE_Diagram.jpg} 
+	\caption{\acs{SHADE}$_{CC}$ optimization process for constrained clustering.}\label{fig:SHADE_CC}
 \end{figure}
@@
 	}
-	\caption[SHADE$_{CC}$]{\acs{SHADE}$_{CC}$}\label{alg:SHADE}
+	\caption{\acs{SHADE}$_{CC}$}\label{alg:SHADE}
 \end{algorithm}
@@
 
-	\caption[SHADE$_{CC}$ Local Search]{\acs{SHADE}$_{CC}$ Local Search}\label{alg:SHADE_LS}
+	\caption{\acs{SHADE}$_{CC}$ Local Search}\label{alg:SHADE_LS}
 \end{algorithm}
@@
 
-Following the stream of thought introduced at the beginning of this chapter, it is justified to think that metaheuristics constitute a promising approach to fin approximate quality solutions to the tough challenge that constrained clustering constitutes.
+Following the stream of thought introduced at the beginning of this chapter, it is justified to think that metaheuristics constitute a promising approach to find approximate quality solutions to the tough challenge that constrained clustering represents.
 Metaheuristic algorithms are designed to explore the solution space of a problem with the guidance of a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}.
@@
 This is the reason why the classic \acf{LS} heuristic algorithm derived in the \acf{ILS} metaheuristic algorithm, which introduces periodic perturbations in the exploration of the solution space to escape local optima \cite{lourencco2010iterated}.
-\acs{ILS}---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
+\acs{ILS} (or variants of it) has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}.
 \acs{ILS} is also very commonly used in the design of hybrid methods with good exploration-exploitation tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.
@@
 
-\subsection{A Glimpse to Iterated Local Search}
+\subsection{A Glimpse into Iterated Local Search}
 
-Before defining the \acs{ILS}---from now on \acs{ILS}---,it is necessary to understand the \acs{LS} and the motivation that leads to designing methods with greater exploratory capacity.
+Before defining \acs{ILS}, it is necessary to understand \acs{LS} and the motivation behind the design of methods with greater exploratory capabilities.
 
-An \acs{LS} is a procedure that consists of performing an heuristics-guided iterative improvement.
-Taking as initial state a well defined solution to the problem, the \acs{LS} uses a neighbor generation operator to explore new solutions.
-When a solution better than the current one is found, it replaces the current one.
-In order to evaluate the solutions, the \acs{LS} uses a function that assigns a value to each one of them, this is known as a target function.
-***This value will be better the closer the solution that evaluates the optimal solution.***
+\acs{LS} is a procedure that consists in performing an heuristics-guided iterative improvement.
+With a valid solution to the problem as an initial state, \acs{LS} uses a neighbor generation operator to explore new solutions.
+The current solution is replaced when a better one is found.
+In order to evaluate solutions, a function to assess their suitability, known as \textit{fitness function}, is used.
+Ideally, the closer a solution is to the global optimum, the better its fitness value will be.
 
-Both the neighbor generation operator and the target function used by the \acs{LS} must be defined for each problem.
-The target function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem.
-The neighbor generation operator takes as base a solution to the problem $s$, and generates a new solution $s^\prime$ by modifying some of its components in a local way.
+Both the neighbor generation operator and the fitness function used by \acs{LS} must be defined for each particular problem.
+The fitness function $f(\cdot)$ determines how good the solution achieved is with respect to the purpose of the problem.
+The neighbor generation operator attempts to improve upon a solution $S$ to the problem by modifying a few of its components, and thus generating a new solution $S^\prime$.
 
 The Algorithm \ref{alg:BasicLS} presents a basic scheme for an \acs{LS} procedure.
-It is worth noting that the condition of line 4 must be adapted if the problem to be solved is one of maximization or minimization.
-In addition, the general scheme considers that the optimization process ends when the current solution is not improved, however the stop condition can be adapted to each problem.
+It is worth noting that the condition of line 4 must be adapted depending on whether the problem to be solved is one of maximization or minimization.
+In addition, the general scheme assumes that the optimization process ends when the current solution is not improved.
+However, the stop condition can be adapted to each problem.
 
@@
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Initial solution $s$}
+	\KwIn{Initial solution $S$}
 	\BlankLine
@@
 
-		$s^\prime \leftarrow $ \GenerateNeighbor{$s$}
+		$S^\prime \leftarrow $ \GenerateNeighbor{$S$}
 
-		\If{$f(s^\prime) < f(s)$}{
+		\If{$f(S^\prime) < f(S)$}{
 
-			$s \leftarrow s^\prime$\\
+			$S \leftarrow S^\prime$\\
 			$improvement \leftarrow$ \texttt{true} \\
@@
 	\BlankLine
-	\KwRet ($s$)
+	\KwRet ($S$)
 
@@
 
-With the \acs{LS} defined in this way, it is easy to see that the solution achieved will be the optimal local $s^*$ closest to the initial solution $s$.
-Figure \ref{img:LS} shows a pictorical representation of the above.
+With \acs{LS} defined in this way, it is easy to see that the solution achieved will be the local optimum $S^*$ that is closest to the initial solution $S$.
+Figure \ref{img:LS} shows a pictorial representation of the above.
 
@@
 	\includegraphics[scale=0.25]{gfx/NewProp/DILS/LS.png}
-	\caption[Local optimum achieved by an LS procedure]{Local optimum achieved by an \acs{LS} procedure}\label{img:LS}
+	\caption{Local optimum achieved by an \acs{LS} procedure.}\label{img:LS}
 \end{figure}
 
-It is therefore necessary to appeal to techniques that introduce diversity in the search to avoid the problem of local optimal. This is exactly what \acs{ILS} is trying to do.
-
-With \acs{ILS}, once we have reached a local optimum $s^*$, we apply a perturbation to it that, with a high probability, will result in a new solution $s^\prime$ that will not be a local optimum and will be far enough away from $s^*$.
-After that, we will apply an \acs{LS} procedure to find a new local optimal $s^{*\prime}$.
-Figure \ref{img:ILS} shows a representation of the above concepts.
+It is therefore necessary to turn to techniques that introduce diversity in the search procedure to avoid the problem of local optimum.
+This is exactly what \acs{ILS} is trying to do.
+With \acs{ILS}, once we have reached a local optimum $S^*$, we apply a perturbation to it that will result in a new solution $S^\prime$.
+With a high probability, $S^\prime$ will not be another local optimum and instead will be far enough away from $S^*$.
+After that, we will apply an \acs{LS} procedure to find a new local optimum $S^{*\prime}$.
+Figure \ref{img:ILS} shows a representation of the aforementioned concepts.
 
@@
 	\includegraphics[scale=0.25]{gfx/NewProp/DILS/ILS.png}
-	\caption[ILS perturbation over LS local optima]{\acs{ILS} perturbation over \acs{LS} local optima}\label{img:ILS}
+	\caption{\acs{ILS} perturbation over an \acs{LS} local optimum.}\label{img:ILS}
 \end{figure}
 
-It should be noted that, if the disturbance applied to $s*$ is too small, the most likely is that the local optimum closest to $s^\prime$ is $s^*$ and therefore the solution space will not be explored effectively.
-On the other hand, if the disturbance is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random reboot-type method.
+It should be noted that, if the perturbation applied to $S*$ is too small, most likely the local optimum closest to $S^\prime$ is again $S^*$ and therefore the exploration of the solution space is not be effective.
+On the other hand, if the perturbation is too big, there will be no bias in the generation of new solutions, and therefore the procedure becomes a random-reboot method.
 
-The method for generating the perturbation applied to the local optimum $s^*$ may vary depending on the problem, and may be based on previous solutions.
-In this way, it is possible to keep a history of the solutions reached so far, so they can be used when applying the perturbation.
-Algorithm \ref{alg:ILS} summarizes the \acs{ILS} overall optimization process.
+The generation of perturbations for the local optimum $S^*$ may vary depending on the problem, and may be based on previous solutions.
+In this way, it is possible to keep a history of solutions reached so far, so they can be used when applying the perturbation.
+Algorithm \ref{alg:ILS} summarizes the overall optimization process of \acs{ILS}.
 
@@
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Initial solution $s$}
+	\KwIn{Initial solution $S$}
 	\BlankLine
-	$s^* \leftarrow$ \LocalSearch{$s$}\\
+	$S^* \leftarrow$ \LocalSearch{$S$}\\
 	\While{Termination criteria are not met}{
 
-		$s^\prime \leftarrow $ \Perturbation{$s^*$, $history$}\\
-		$s^{*\prime} \leftarrow$ \LocalSearch{$s^\prime$}
+		$S^\prime \leftarrow $ \Perturbation{$S^*$, $history$}\\
+		$S^{*\prime} \leftarrow$ \LocalSearch{$S^\prime$}
 
-		\If{$f(s^{*\prime}) < f(s^*)$}{
+		\If{$f(S^{*\prime}) < f(S^*)$}{
 
-			$s^* \leftarrow s^{*\prime}$
+			$S^* \leftarrow S^{*\prime}$
 
@@
 	\BlankLine
-	\KwRet ($s*$)
+	\KwRet ($S*$)
 
@@
 
-In the \acs{ILS} scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only use the difference in the cost of the solutions to be compared to choose between them.
-This is the most simple scheme and does not depends on previous solutions recorded in the history.
+In the \acs{ILS} scheme described in Algorithm \ref{alg:ILS} the acceptance criterion only uses the difference in the cost of the compared solutions to choose between them.
+This is the simplest scheme and does not rely on the history of previous solutions.
 
@@
 
-The \acf{DILS} is a new variant of the classic \acs{ILS} method.
+The Dual Iterative Local Search (DILS) is a new variant of the classic ILS method.
 Its goal is to perform a search in the solution space to find the solution with the best fitness value (given by fitness function $f$) by introducing diversity in an adaptive and guided way.
-While \acs{ILS} works with a single individual, \acs{DILS} keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima.
+While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima.
 The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value at the end of each stage (generation) of the optimization process.
 
-In each generation $G$ of the optimization process \acs{DILS} builds a new individual by applying a recombination operator to $m_b$ and $m_w$.
+In each generation $G$ of the optimization process DILS builds a new individual by applying a recombination operator to $m_b$ and $m_w$.
 After that, it applies a strong mutation operator to the newly generated individual.
-This is the individual \acs{DILS} applies the \acs{LS} procedure to in order to improve its fitness value, resulting in the trial individual $m_t$.
-The trial individual $m_t$ replaces the worst of its predecessors, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for \acs{DILS}.
-Equation \ref{eq2.1} displays the expression for this criterion in the case of a minimization problem.
+This is the individual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trial individual $m_t$.
+The trial individual $m_t$ replaces the worst of its predecessors, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for DILS.
+Equation \ref{eq1.2} displays the expression for this criterion in the case of a minimization problem.
 
@@
 \right..
-\label{eq2.1}
+\label{eq1.2}
 \end{equation}
@@
 Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solution space.
-To avoid it \acs{DILS} implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory ($\{m_b, m_w\}$).
+To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory ($\{m_b, m_w\}$).
 Equation \ref{eq2.2} shows the reinitialization criterion for a minimization problem.
+RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
 
 \begin{equation}
+\resizebox{0.85 \textwidth}{!}{$
 	m_{w,G+1} = \left\{ \begin{array}{lc}
 	\text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
-m_{w,G} &  \text{otherwise}
+	m_{w,G} &  \text{if} \;\; f(m_b) - f(m_w) \le f(m_b) * \xi \; \wedge \; f(m_b) < f(m_w) \;\;\\
+	m_{b,G} &  \text{if} \;\; f(m_b) - f(m_w) \le f(m_b) * \xi \; \wedge \; f(m_b) > f(m_w) \;\;\\
 	\end{array}
-\right.,
+	\right.$}.
 \label{eq2.2}
@@
 
-\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
+To maintain consistency between generations we also need to reassign $m_b$ to the true best individual once the reinitialization criterion has been tested, Equation \ref{eq5.2} formalize this simple step.
 This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best predecessor.
-Figure \ref{img:DILS_diag} summarizes the \acs{DILS} optimization process.
+
+\begin{equation}
+m_{b,G+1} = \left\{ \begin{array}{lc}
+m_{b,G} &  \text{if} \;\; f(m_b) < f(m_w) \;\;\\
+m_{w,G} &  \text{if} \;\; f(m_b) > f(m_w) \;\;\\
+\end{array}
+\right..
+\label{eq5.2}
+\end{equation}
+
+Figure \ref{img:DILS_diag} summarizes the DILS optimization process.
+Arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively.
+This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
 
@@
 	\centering
-	\includegraphics[scale=0.27]{gfx/NewProp/DILS/DILSDiagram.jpg}
-	\caption{Diagram summarizing \acs{DILS} optimization process.}\label{img:DILS_diag}
+	\includegraphics[scale=0.25]{gfx/NewProp/DILS/DILSDiagram.jpg}
+	\caption{Diagram summarizing DILS optimization process.}\label{img:DILS_diag}
 \end{figure}
-
-It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively.
-This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
 
@@
 On one hand, if $\xi > 0$, then the true worst individual is only reinitialized when $m_t$ replaces $m_w$ and is also better than $m_b$ by a margin.
-On the other hand, if $\xi < 0$, we allow \acs{DILS} to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
+On the other hand, if $\xi < 0$, we allow DILS to reinitialize the true worst individual even when $m_t$ is not better than $m_b$.
 It becomes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$.
@@
 The reason for this is that the fitness value of the best individual $f(m_b)$ never worsens ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it becomes increasingly harder for $m_w$ to achieve similar or better scores.
-Algorithm \ref{alg:DILS} summarizes the overall \acs{DILS} optimization process.
+Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.
 
@@
 	\SetNlSkip{0.5em}
-	\setstretch{1.2}
+	\setstretch{1.1}
 	\SetKwFunction{RandInit}{RandInit}
@@
 	\KwRet{$m_b$}
-	\caption{DILS}\label{alg:DILS}
+	\caption{\acs{DILS}}\label{alg:DILS}
 \end{algorithm}
@@
 
-\subsection{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}
+\subsection[\acsfont{DILS} Optimization Scheme for Constrained Clustering]{DILS Optimization Scheme for Constrained Clustering} \label{sec:DILS_CC}
 
@@
 Additionally, in order not to bias the exploration of the solution space, we will use random initialization to set $m_b$ and $m_w$; this procedure will also be used in the reinitialization of $m_w$.
-
-Regarding the fitness function $f$, for each individual $m$ its fitness value $f_m$ can be calculated as in Equation \ref{eq2.3}.
-
-\begin{equation}
-f_m = z_m * (\text{infeasibility}_m + 1),
-\label{eq2.3}
-\end{equation}
-
-\noindent where $\text{infeasibility}_m$ is the number of non-satisfied constraints and $z_m$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2.4}.
-
-\begin{equation}
-z_m = \sum_{c_i \in C_m} \left[ \frac{\sum_{x_a, x_b \in c_i} d^2(x_a,x_b)}{|c_i|}\right].
-\label{eq2.4}
-\end{equation}
+Regarding the fitness function $f(\cdot)$, for each individual $m$ its fitness value $f(m)$ can be calculated as in Equation \ref{eq14}.
 
@@
 	\SetNlSkip{0.5em}
-	\setstretch{1.2}
+	\setstretch{1.1}
 	\SetKwFunction{RandomShuffle}{RandomShuffle}
@@
 	\SetKwRepeat{Do}{do}{while}
-	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $p_m$.}
+	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $S$, number of clusters $k$, maximum number of neighbors that can be generated $p_m$.}
 	\BlankLine
@@
 		$improvement \leftarrow$ \texttt{false} \\
-		\tcp{Select random index (object) from $m$}
+		\tcp{Select random index (object) from $S$}
 		$i \leftarrow $ \Rand{$1$,$n$} \\
@@
 		\For{$l \in RSL$}{
-			$m^\prime \leftarrow m$\\
-			\tcc{Move object $i$ from $m$ to the cluster associated with label $l$}
-			$m^\prime_i \leftarrow l$\\
-			\If{$f(m^\prime) < f(m)$}{
-				$m \leftarrow m^\prime$\\
-				$improvement \leftarrow$ \texttt{true} \\
+			$S^\prime \leftarrow S$\\
+			\tcc{Move object $i$ from $S$ to the cluster associated with label $l$}
+			$S^\prime_i \leftarrow l$\\
+			\If{$f(S^\prime) < f(S)$}{
+				$S \leftarrow S^\prime$\texttt{;} $improvement \leftarrow$ \texttt{true} \\
 			}
@@
 	\BlankLine
-	\KwRet{$m$}
+	\KwRet{$S$}
 
-	\caption[DILS$_{CC}$ Local Search]{\acs{DILS}$_{CC}$ Local Search}\label{alg:DISL_LS}
+	\caption{\acs{DILS}$_{CC}$ Local Search}\label{alg:DISL_LS}
 \end{algorithm}

### 2019-07-06 Memoria/FrontBackmatter/Abstract.tex
Cambios en memoria y en paper DILS

@@
 Constraints can be given in the form of instance-level must-link and cannot-link constraints, which this paper focuses on.
-We propose the first application of Differential Evolution to the constrained clustering problem, as well as a new Local Search variant and its application to the problem at hands.
+We propose the first application of Differential Evolution to the constrained clustering problem, as well as a new Local Search variant and its application to the problem at hand.
 We will compare the results obtained by these new proposals to those obtained by the state-of-the-art algorithms on twenty datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.\\
 
-\noindent\textbf{Keywords:} constrained clustering, instance-level, must-link, cannot-link, differential evolution, local search.
+\noindent\textbf{Keywords:} constrained clustering, instance-level, must-link, cannot-link, Differential Evolution, Local Search.
 

### 2019-07-06 Memoria/FrontBackmatter/Declaration.tex
Cambios en memoria y en paper DILS

@@
 
-\noindent Que el Trabajo Fin de Grado presentado en la Escuela Técnica Superior de Ingenierías Informática y de Telecomunicaciones de la Universidad de  Granada, con fecha 17 de Junio de 2018, titulado <<\textit{Constrained Clustering: New Proposals}>> es original, no es copia de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona.\\
+\noindent Que el Trabajo Fin de Máster presentado en la Escuela Internacional de Posgrado de la Universidad de Granada, con fecha 4 de julio de 2019, titulado <<\textit{\myTitle: \mySubtitle}>> es original, no es copia de ningún otro trabajo, inédito, y no ha sido difundido por ningún medio, ni presentado anteriormente por quien suscribe o por otra persona.\\
 
-\noindent Y para que así conste a los efectos oportunos, firmo la presente en Granada, a 17 de junio de 2018.
+\noindent Y para que así conste a los efectos oportunos, firmo la presente en Granada, a 4 de julio de 2019.
 

### 2019-07-06 Memoria/FrontBackmatter/Legality.tex
Cambios en memoria y en paper DILS

@@
 
-Yo, \textbf{Germán González Almagro}, alumno de la titulación en ingeniería informática de la \textbf{Escuela Técnica Superior
-de Ingenierías Informática y de Telecomunicación de la Universidad de Granada}, con DNI $\mathrm{76593910}$T, autorizo la
-ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca del Centro para que pueda ser
+Yo, \textbf{Germán González Almagro}, alumno del Máster Universitario Oficial en Ciencia de Datos e Ingeniería de Computadores de la \textbf{Escuela Internacional de Posgrado de la Universidad de Granada}, con DNI $\mathrm{76593910}$T: Autorizo la
+ubicación de la siguiente copia de mi Trabajo Fin de Máster en la biblioteca de la Universidad de Granada para que pueda ser
 consultada por las personas que lo deseen.
@@
 \begin{flushleft}
-	Granada, 17 de junio de 2018.
+	Granada, 4 de julio de 2019.
 \end{flushleft}
@@
 
-D. \textbf{\myProf}, Profesor del Departamento Ciencias de la Computación en Inteligencia Artificial de la Universidad de Granada.
+D. \textbf{\myProf}, Profesor Titular del Departamento Ciencias de la Computación en Inteligencia Artificial de la Universidad de Granada.
 
@@
 
-Que el presente trabajo, titulado \textit{\textbf{<<Constrained Clustering: New Proposals>>}},
+Que el presente trabajo, titulado \textit{\textbf{<<\myTitle: \mySubtitle>>}},
 ha sido realizado bajo su supervisión por \textbf{Germán González Almagro}, y autorizo la defensa de dicho trabajo ante el tribunal
@@
 
-Y para que conste, expido y firmo el presente informe en Granada a 17 de junio de 2018.
+Y para que conste, expido y firmo el presente informe en Granada a 4 de julio de 2019.
 
@@
 
-\vspace{5cm}
+\vspace{4cm}
 
 \begin{table}
-	\begin{minipage}{.4\textwidth}
-		\centering
+		\flushright
 		\begin{tabular}{c}
@@
 		\end{tabular}
-	\end{minipage}
 \end{table}

### 2019-07-06 Memoria/FrontBackmatter/Resumen.tex
Cambios en memoria y en paper DILS

--- 2019-06-18 Memoria/FrontBackmatter/Resumen.tex
@@
 
+El clustering siempre ha sido una potente herramienta en tareas de análisis de datos.
+Aunque tradicionalmente se ha ubicado dentro del marco del aprendizaje no supervisado, su potencial se vio ampliado tras la incorporación de nuevos tipos de información al mismo, dando lugar a un nuevo tipo de aprendizaje semisupervisado: el clustering con restricciones.
+Esta técnica es una generalización del clustering que considera información adicional, dada en forma de restricciones.
+Éstas pueden venir especificadas a nivel de instancia, centrándose este trabajo en las de tipo \textit{must-link} y \textit{cannot-link}.
+En este trabajo proponemos la primera aplicación de \textit{Differential Evolution} al clustering con restricciones, así como una nueva variante de la Búsqueda Local, incluyendo su esquema de optimización general y su aplicación al clustering con restricciones.
+Compararemos los resultados obtenidos por estas nuevas propuestas con los obtenidos por el estado del arte sobre veinte conjuntos de datos con niveles incrementales de información basada en restricciones.
+Apoyaremos nuestras conclusiones mediante el uso de tests Bayesianos.
+\\
+
+\noindent\textbf{Palabras Clave:} Clustering con restricciones, restricciones a nivel de instancia, \textit{must-link}, \textit{cannot-link}, \textit{Differential Evolution}, Búsqueda Local.
+
 \endgroup

### 2019-07-06 Papers/DILS/paperDILS.tex
Cambios en memoria y en paper DILS

@@
 
-\title{Improving instance-level constrained clustering through Dual Iterative Local Search}
+\title{DILS: a new approach to constrained clustering}
 
@@
 We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to produce quality results for the constrained clustering problem.
-We compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
+We compare the results obtained by this proposal to those obtained by the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
 \end{abstract}
@@
 It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled.
-Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics,\cite{Everitt:2009:CA:1538772} cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.
-
-In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering.
-In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}.
-A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}.
-In this paper we will focus on partitional clustering.
-
 We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them.
@@
 
-Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning.
-We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}.
-The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances.
-The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{triguero2015self}.
-
-Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
+Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning \cite{chapelle2009semi}.
+Following this paradigm we can incorporate background information to the clustering process, resulting in constrained clustering, which is the main subject of the study presented in this paper \cite{triguero2015self}.
+The constrained clustering goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set.
 It has been successfully applied in many knowledge fields, among which it is worth mentioning: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, and lane finding in GPS data \cite{wagstaff2001constrained} among others.
@@
 Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.
-
 In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints.
 In one hand, pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively.
-On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure \cite{davidson2007survey}.
-This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
+This paper focuses on these types of constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.
 
@@
 In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptive and guided way.
-We apply it to the constrained clustering problem, proving that it constitutes a competitive approach to obtain quality results.
+We build the DILS application to constrained clustering with the aid of an integer-based problem representation and a penalty-style fitness function.
+Also we prove that it constitutes a competitive approach to obtain quality results to the constrained clustering problem.
 
@@
 Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem.
-The feasibility problem for non-hierarchical instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
+The feasibility problem for instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.
 
@@
 While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima.
-The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value.
+The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value at the end of each stage (generation) of the optimization process.
 
@@
 This is the individual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trial individual $m_t$.
-The trial individual $m_t$ replaces the worst of its parents, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for DILS.
+The trial individual $m_t$ replaces the worst of its predecessors, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for DILS.
 Equation \ref{eq1} displays the expression for this criterion in the case of a minimization problem.
@@
 Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.
+RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
 
 \begin{equation}
+\resizebox{0.9 \textwidth}{!}{$
 m_{w,G+1} = \left\{ \begin{array}{lc}
 \text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
-m_{w,G} &  \text{otherwise}
+m_{w,G} &  \text{if} \;\; f(m_b) - f(m_w) \le f(m_b) * \xi \; \wedge \; f(m_b) < f(m_w) \;\;\\
+m_{b,G} &  \text{if} \;\; f(m_b) - f(m_w) \le f(m_b) * \xi \; \wedge \; f(m_b) > f(m_w) \;\;\\
 \end{array}
-\right.,
+\right.$}.
 \label{eq2}
@@
 
-\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method.
-This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best parent.
+To maintain consistency between generations we also need to reassign $m_b$ to the true best individual once the reinitialization criterion has been tested, Equation \ref{eq5} formalize this simple step.
+This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best predecessor.
+
+\begin{equation}
+m_{b,G+1} = \left\{ \begin{array}{lc}
+m_{b,G} &  \text{if} \;\; f(m_b) < f(m_w) \;\;\\
+m_{w,G} &  \text{if} \;\; f(m_b) > f(m_w) \;\;\\
+\end{array}
+\right..
+\label{eq5}
+\end{equation}
+
 Figure \ref{img:DILS_diag} summarizes the DILS optimization process.
+Arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively.
+This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
 
@@
 \end{figure}
-
-It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively.
-This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.
 
@@
 
-In this section, we present an optimization scheme for the DILS algorithm when applied to constrained clustering.
+In this section, we build the DILS application scheme for constrained clustering.
 We will call this new approach DILS$_{CC}$.
@@
 For our experiments we will compare the results obtained by DILS$_{CC}$ and state-of-the-art methods over 25 datasets and 3 constraint sets for each one of them.
-Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
-\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
+Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\footnote{https://sci2s.ugr.es/keel/category.php?cat=clas} \cite{triguero2017keel}, though some of them have been obtained via
+\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package}\footnote{https://scikit-learn.org/stable/datasets/index.html} \cite{scikit-learn}.
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}\footnote{https://github.com/GermangUgr/DILS\_CC}.
 Table \ref{tab:datasets} displays a summary of every dataset.
@@
 
-Since we have the true labels associated to each of the datasets, we can use them to evaluate the results provided by each method.
+Since we have the true labels associated with each of the datasets, we can use them to evaluate the results provided by each method.
 We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}.
@@
 The stop criterion is given by the number of evaluations of the fitness function $f$, which at most will be 300000.
-The implementation for DILS$_{CC}$ can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.
+The implementation for DILS$_{CC}$ can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}\footnote{https://github.com/GermangUgr/DILS\_CC}.
 
-To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}\footnote{https://github.com/GermangUgr/TFG/tree/master/Software}.
 The implementation for the BRKGA algorithm can be found in the same link as DILS$_{CC}$.
 In all cases the value for $k$ is equal to the number of classes displayed in Table \ref{tab:datasets}.
-
-\textcolor{red}{Meter todos los algoritmos en el mismo repo para no tener que poner tantos enlaces a github? Poner los enlaces como notas a pie de pagina tambien?}
 
@@
 In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.
+We establish that in these cases the ARI value considered for the mean calculation---and the forthcoming statistical analysis of results---is $-1.000$, which is the worst possible ARI value.
 
@@
 			\hline
-			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
+			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CECM \\
 			\hline
@@
 Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set.
-It is possible to notice that DILS$_{CC}$ represents a consistent improvement over all other 5 methods.
+It is possible to notice that DILS$_{CC}$ represents a consistent improvement over all other 6 methods.
 This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lowest level of constraint-based information we work with.
@@
 			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
-			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
+			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & \textbf{1.000} & 0.138 \\
 			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
@@
 			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
-			Heart & \textbf{1.000} & 0.942 & \textbf{1.000} & 0.025 & 0.033 & 1.000 & 0.000 \\
+			Heart & \textbf{1.000} & 0.942 & \textbf{1.000} & 0.025 & 0.033 & \textbf{1.000} & 0.000 \\
 			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
@@
 			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
-			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
+			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & \textbf{1.000} & 0.180 \\
 			Movement Libras & 0.018 & 0.006 & - & \textbf{0.293} & \textbf{0.293} & 0.000 & - \\
@@
 In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how the quality of the results of DILS$_{CC}$ scales with the number of constraints, as well as that of methods such as BRKGA or COPKM (the latter being more limited in other aspects).
+The TVClust method should also be highlighted, since it is able to find the optimal solution in 4 cases, which implies a significant improvement over its performance on previous constraint sets.
 It is also worth noting how DILS$_{CC}$ provides the best average ARI for each of the three constraint sets analyzed in this paper.
@@
 
-Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with DILS$_{CC}$ as sample $B$.
+Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with DILS$_{CC}$ as sample $B$.
 
@@
 		\includegraphics[width=1\linewidth]{Figures/DILSvsCEKM.pdf}
-		\caption{CEKM}
-		\label{fig:DILSvsCEKM}
+		\caption{CECM}
+		\label{fig:DILSvsCECM}
 	\end{subfigure}

### 2019-07-06 Papers/SHADE/paperSHADE.tex
Cambios en memoria y en paper DILS

@@
 For our experiments we will compare the results obtained by BRKGA$_{CC}$ and SHADE$_{CC}$ over 25 datasets and 3 constraint sets for each one of them.
-Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
-\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}.
-We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/SHADE_CC}{GitHub}.
+Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository}\footnote{https://sci2s.ugr.es/keel/category.php?cat=clas} \cite{triguero2017keel}, though some of them have been obtained via
+\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package}\footnote{https://scikit-learn.org/stable/datasets/index.html} \cite{scikit-learn}.
+We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/SHADE_CC}{GitHub}\footnote{https://github.com/GermangUgr/SHADE\_CC}.
 Table \ref{tab:datasets} displays a summary of every dataset.
@@
 In both cases---BRKGA$_{CC}$ and SHADE$_{CC}$---the population size will be 100 individuals, and the stop criterion is given by the number of evaluations of the fitness function, which at most will be 300000.
-A Python implementation for both algorithms can be found at \href{https://github.com/GermangUgr/SHADE\_CC}{GitHub}
-
-To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}.
+A Python implementation for both algorithms can be found at \href{https://github.com/GermangUgr/SHADE_CC}{GitHub}\footnote{https://github.com/GermangUgr/SHADE\_CC}
+
+To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}\footnote{https://github.com/GermangUgr/TFG/tree/master/Software}.
 
@@
 
-In Tables from \ref{tab:results10} to \ref{tab:results20} the ARI columns shows the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
+In Tables from \ref{tab:results10} to \ref{tab:results20} the ARI columns show the Adjusted Rand Index for each case, the Unsat columns displays the percentage of violated constraints and the Time columns shows the time taken for each algorithm to deliver results, measured in seconds.
 
@@
 As far as the state-of-the-art methods are concerned, we find a general improvement in ARI, except for CECM.
-Even with this, it is SHADE$_{CC}$ and COPKM that seem to obtain the highest quality results, with the difference that SHADE$_{CC}$ is able to produce results in all cases.
+Even with this, it is SHADE$_{CC}$ and COPKM that obtain the highest quality results, with the difference that SHADE$_{CC}$ is able to produce results in all cases.
+The TVClust method should also be highlighted, since it is able to find the optimal solution in 4 cases, which implies a significant improvement over its performance on previous constraint sets
 
@@
 	\end{subfigure}
-	\caption{Heat diagrams for the comparison with BRKGA$_{CC}$.}
+	\caption{Heatmaps for the comparison with BRKGA$_{CC}$.}
 \end{figure}
